---
title: "11-Make_XeniumSPE_forSamui"
output: html_document
date: "2024-10-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(sf)
library(ggplot2)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
## load RAW xenium SPE (i.e., before the coordinates were transformed for banksy); load main clustering results and their annots, and drop cells from sample-specific clusters
hypx <- readRDS("processed-data/03_make_SPE-SFE/01_sfe_raw.RDS")

## drop control probes
keeprn <- grep(rownames(hypx),pattern="Deprecated|Unassigned|NegControl",value=T,invert=T)

hypx <- hypx[keeprn,]
##

## add manuscript sample ids
mscriptids <- fread("../standardized_sampleids_for_plotting.txt")
tmpcd <- as.data.table(colData(hypx),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,mscriptids[,.(sample_id,manuscript_id)],by.x="Sample",by.y="sample_id",all.x=T)
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(hypx),]
tmpcd$rn <- NULL
colData(hypx) <- tmpcd
## 

## append cell types, drop the sample-specific clusters, and add the Banksy clusters + labels
bkcl <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"cl")

bkanno <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
bkanno <- unique(bkanno[,.(clus,bjm_annot)])

vmharccellanno <- fread("../manuscript_plot_code/xARCxVMH_cluster_detailedlabs_andplotformatnames.txt")
vmharccellanno[,dom.clus:=paste0("x",dom.clus)]
bkanno <- merge.data.table(bkanno,vmharccellanno,by="bjm_annot",all.x=T)
bkanno[!is.na(dom.clus),bjm_annot:=dom.clus]

## drop clusters labeled for, well, dropping
bkanno <- bkanno[!(bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T))]

## subset the cellwise labels to these clusters, also get cluster groups, and label those cells/cluster groups
bkcl <- merge.data.table(bkcl,bkanno,by.y="clus",by.x="cl")

hypx <- hypx[,colnames(hypx) %in% bkcl$rn]

bkcl[,supercl:="ifthisisplottedigoofed"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="VMH",value=T),supercl:="VMH (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="ARC",value=T),supercl:="ARC (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Tanyc|Alar",value=T),supercl:="Tanycytes, Portal Vasc. (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Oligo",value=T),supercl:="Oligo (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Astro",value=T),supercl:="Astro (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Microg",value=T),supercl:="Microglia (3)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Supraopt",value=T),supercl:="SON (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Unsure_AVP",value=T),supercl:="Non-SON *AVP*+*OXT*+ (1)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Macrop|Periph|Vascul|Endothel",value=T),supercl:="Vascular, Periph.<br>Immune (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="GABA",value=T),supercl:="Other GABAergic (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Periven",value=T),supercl:="PeriVN (Excitatory) (1)"]

bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]

hypx$Banksy_cluster <- bkcl$bjm_annot
hypx$banksy_cluster_group <- bkcl$supercl
### also add the domain labels from knn
bknn <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt")

bknn <- bknn[rn %in% bkcl$rn,.(rn,dualVMHARC4)]
bknn <- DataFrame(bknn,row.names=bknn$rn)[colnames(hypx),]
hypx$domain_of_cell <- bknn$dualVMHARC4

### adjust coordinates to zero-based for each sample
samps <- unique(hypx$sample_id)
xens <- lapply(samps,FUN=function(s){hypx[,hypx$sample_id==s]})
names(xens) <- samps

xens <- lapply(xens,FUN=function(x){
  spatialCoords(x)[,1] <- spatialCoords(x)[,1]-min(spatialCoords(x)[,1])
  spatialCoords(x)[,2] <- spatialCoords(x)[,2]-min(spatialCoords(x)[,2])
  ## account for samui using a reverse_y type paradigm for the y axis by reversing the axis beforehand. but stay in 0-positive integer range
  spatialCoords(x)[,2] <- abs(spatialCoords(x)[,2]-max(spatialCoords(x)[,2]))
  return(x)
})

i<-1
for (i in c(1:length(xens))){
   if (i == 1){
      hypx2 <- xens[[i]]
   } 
   else {
      hypx2 <- cbind(hypx2,xens[[i]]) 
   }
}

## plot test: using y_reverse should put vmh arc in bottom left corner
# test <- as.data.table(cbind(spatialCoords(hypx2[,hypx2$sample_id=="X36_5459A"]),hypx2[,hypx2$sample_id=="X36_5459A"]$domain_of_cell))
# test[,x_centroid:=as.numeric(x_centroid)]
# test[,y_centroid:=as.numeric(y_centroid)]
# ggplot(test,aes(x=x_centroid,y=y_centroid,col=V3))+geom_point(size=0.5,stroke=0)+theme(axis.text=element_blank())+scale_y_continuous(transform="reverse")

rm(xens,hypx)
gc()

### save the xenium SPE (new version with y axis reversed--don't overwrite the previous one without)

saveRDS(hypx2,"processed-data/11-XeniumSPE_forSamui_yreversed.RDS")
```

save just X97_reg3 SPE as a one-sample test SPE for ryan to test xenium-samui interplay.
```{r}
saveRDS(hypx2[,hypx2$sample_id=="X97_reg3"],"processed-data/11tmp-XeniumSPE_onesample_test/X97reg3_only_testSPE.RDS")


```

### make a straight-black png with coordinates equal to the full-res image. also make a variant of this with VMH, ARC outlines overlaid.

```{r}
samps <- unique(hypx2$sample_id)
xens <- lapply(samps,FUN=function(s){hypx2[,hypx2$sample_id==s]})
names(xens) <- samps
# # initialize tables into a samplewise list
# dombounds <- as.data.table(bknn)[dualVMHARC4 %in% c("VMH","ARC")]
# dombounds[,sampleid:=gsub(rn,pattern="^(.*)_[[:alnum:]]{8}-.$",replacement="\\1")]
# 
# 
# # make into a samplewise list
# dombounds <- split(dombounds,by="sampleid")
# dombounds <- dombounds[names(xens)]

dombounds <- lapply(xens,FUN=function(x){
    v <- st_as_sf(as.data.frame(spatialCoords(x[,x$domain_of_cell=="VMH"])),coords = c("x_centroid","y_centroid"))
    a <- st_as_sf(as.data.frame(spatialCoords(x[,x$domain_of_cell=="ARC"])),coords = c("x_centroid","y_centroid"))
    
    vhullpt <- st_convex_hull(st_union(v))
    ahullpt <- st_concave_hull(st_union(a),ratio=0.1)
    ## create the data table we will use for overlaying the boundary polygons. geom_path will need the first point to also be popped into the last slot to close the shape
    vdompoly <- rbind(vhullpt[[1]][[1]],
                      vhullpt[[1]][[1]][1,])
    vdompoly <- cbind(vdompoly,rep("VMH",nrow(vdompoly)))

    adompoly <- rbind(ahullpt[[1]][[1]],
                      ahullpt[[1]][[1]][1,])
    adompoly <- cbind(adompoly,rep("ARC",nrow(adompoly)))
    
    vdompoly <- as.data.table(vdompoly)
    adompoly <- as.data.table(adompoly)
    
    setnames(vdompoly,c("xpol","ypol","dompol"))
    setnames(adompoly,c("xpol","ypol","dompol"))
    
    dompoly <- rbind(vdompoly,adompoly)
    
    dompoly[,xpol:=as.numeric(xpol)]
    dompoly[,ypol:=as.numeric(ypol)]
    return(dompoly)
})

## make sure the ordering of dombounds is the same as samps
dombounds <- dombounds[samps]
```

#pngs of black bkgd and domain outlines for all samples
```{r}
## load domain palette
pal <- readRDS("../manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")[[1]]

mapply(s=samps,d=dombounds,FUN=function(s,d){
   tmp <- hypx2[,hypx2$sample_id==s]
   blnkbkgddim <- list(w=max(spatialCoords(tmp)[,1]),
                       h=max(spatialCoords(tmp)[,2]))

   p <- ggplot()+
      geom_rect(aes(xmin=0,xmax=blnkbkgddim$w,ymin=0,ymax=blnkbkgddim$h),fill="black")+
      
      theme_void()
   
   png(paste0("processed-data/11-XeniumSPE_imageunderlays/",s,"__",mscriptids[sample_id==s,manuscript_id],"_blackbkgd.png"),width=blnkbkgddim$w,height=blnkbkgddim$h,units="px",res=300,bg="black")
   
   print(p)
   
   dev.off()
   
   ## add domain polygons
   p <- p+
      geom_path(data=d,aes(x=xpol,y=ypol,col=dompol),linewidth=5)+
      scale_color_manual(values=pal)+
      guides(color="none")

   
   png(paste0("processed-data/11-XeniumSPE_imageunderlays/",s,"__",mscriptids[sample_id==s,manuscript_id],"_domainbounds.png"),width=blnkbkgddim$w,height=blnkbkgddim$h,units="px",res=300,bg="black")
   
   print(p)
   
   dev.off()

})

```


### reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.38.0             
 [3] colorout_1.3-0.2               ggplot2_3.5.1                 
 [5] sf_1.0-16                      SpatialFeatureExperiment_1.6.1
 [7] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [9] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[11] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
[13] IRanges_2.38.1                 S4Vectors_0.42.1              
[15] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[17] matrixStats_1.3.0              data.table_1.15.4             
[19] rlang_1.1.4                   

loaded via a namespace (and not attached):
 [1] DBI_1.2.3                 bitops_1.0-7              deldir_2.0-4             
 [4] s2_1.1.6                  magrittr_2.0.3            e1071_1.7-14             
 [7] compiler_4.4.1            DelayedMatrixStats_1.26.0 png_0.1-8                
[10] sfheaders_0.4.4           fftwtools_0.9-11          vctrs_0.6.5              
[13] pkgconfig_2.0.3           wk_0.9.2                  crayon_1.5.3             
[16] fastmap_1.2.0             magick_2.8.3              XVector_0.44.0           
[19] labeling_0.4.3            scuttle_1.14.0            utf8_1.2.4               
[22] rmarkdown_2.27            UCSC.utils_1.0.0          xfun_0.45                
[25] zlibbioc_1.50.0           beachmat_2.20.0           jsonlite_1.8.8           
[28] rhdf5filters_1.16.0       DelayedArray_0.30.1       Rhdf5lib_1.26.0          
[31] jpeg_0.1-10               tiff_0.1-12               terra_1.7-78             
[34] parallel_4.4.1            R6_2.5.1                  limma_3.60.3             
[37] boot_1.3-30               Rcpp_1.0.13               knitr_1.48               
[40] R.utils_2.12.3            tidyselect_1.2.1          Matrix_1.7-0             
[43] rstudioapi_0.16.0         abind_1.4-5               yaml_2.3.9               
[46] EBImage_4.46.0            codetools_0.2-20          lattice_0.22-6           
[49] tibble_3.2.1              withr_3.0.0               evaluate_0.24.0          
[52] units_0.8-5               spData_2.3.1              proxy_0.4-27             
[55] pillar_1.9.0              KernSmooth_2.23-24        generics_0.1.3           
[58] rprojroot_2.0.4           sp_2.1-4                  RCurl_1.98-1.14          
[61] sparseMatrixStats_1.16.0  munsell_0.5.1             scales_1.3.0             
[64] class_7.3-22              glue_1.7.0                tools_4.4.1              
[67] BiocNeighbors_1.22.0      locfit_1.5-9.10           rhdf5_2.48.0             
[70] grid_4.4.1                spdep_1.3-5               DropletUtils_1.24.0      
[73] edgeR_4.2.0               colorspace_2.1-0          GenomeInfoDbData_1.2.12  
[76] HDF5Array_1.32.0          cli_3.6.3                 fansi_1.0.6              
[79] S4Arrays_1.4.1            dplyr_1.1.4               gtable_0.3.5             
[82] R.methodsS3_1.8.2         zeallot_0.1.0             digest_0.6.36            
[85] classInt_0.4-10           SparseArray_1.4.8         dqrng_0.4.1              
[88] farver_2.1.2              rjson_0.2.21              htmlwidgets_1.6.4        
[91] htmltools_0.5.8.1         R.oo_1.26.0               lifecycle_1.0.4          
[94] httr_1.4.7                here_1.0.1                statmod_1.5.0            
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-11-18
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
