---
title: "02-Registration between Visium and Xenium"
output: html_document
date: "2024-08-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
#options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load spatialLIBD 1vsall marker results from visium and xenium
```{r}
## visium marker analyses
slibdvis <- readRDS("../spatial_HYP/processed-data/07-Marker_genes/01b-BSpace_allruns_spatialLIBD-regwrap-enrichment_full.RDS")
## subset to those using top 10%ile shared SVGs and harmony reduction with lambda=NA
slibdvis <- slibdvis[grep(names(slibdvis),pattern="nnsvg10",value=T)]
slibdvis <- slibdvis[grep(names(slibdvis),pattern="lmbna",value=T)]

## xenium markers
slibdxen <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_spLIBD_markers.RDS")
slibdxen <- as.data.table(slibdxen[["enrichment"]])

## xenium cell cluster annots
bkanno <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
bkanno <- unique(bkanno[,c("clus","bjm_annot")])
bkanno[,clus:=paste0("X",clus)]

```

tried the spatialLIBD registratoin function first, though it's given trouble before and indeed still does. no matter how much i do to wrangle the inputs into the format it seems to be looking for, it says eitherr
Error: It looks like 'stats' does not have ENSEMBL gene names on the rownames or none of the genes are matching.
or 
Error in `colnames<-`(`*tmp*`, value = gsub("[f|t]_stat_", "", colnames(tstats))) : attempt to set 'colnames' on an object with less than two dimensions

Even when it's passed matrices pre-subsetted to matching row names and data frames that are of course two dimensions. So fuck it, use the visium hyp scripts' re-write of this in ugly iterative code (see ../spatial_HYP/code/10-Spatial Registration / 02-Spatial registration vs Yao... , line 295):
	
```{r}
### I wrote this code to use inputs already subsetted to just t-statistics from the enrichment test tables. we can put these together in the following way. the xenium's simpler
keepnames <- c("ensembl",grep(names(slibdxen),pattern="t_stat",value=T))
slibdxen <- slibdxen[,..keepnames]
# switch the xenium column names from Xnumber to the annotation
swapnames <- as.data.table(cbind(colnames(slibdxen)[2:ncol(slibdxen)],
                   gsub(names(slibdxen)[2:ncol(slibdxen)],pattern="t_stat_",replacement="")))
swapnames <- merge(swapnames,bkanno,by.x="V2",by.y="clus")
# first, drop the discarded sample-specific clusters from the marker table
swapnames <- swapnames[bjm_annot!="DISCARD"]
keepnames <- c("ensembl",swapnames$V1)
slibdxen <- slibdxen[,..keepnames]
setnames(slibdxen,swapnames$V1,swapnames$bjm_annot)
####

### the visium will take some more work
slibdvis <- lapply(slibdvis,FUN=function(x){
    keepnames <- c("ensembl",grep(names(x),pattern="t_stat",value=T))
    y <- x[,..keepnames]
    setnames(y,c(2:ncol(y)),gsub(names(y)[c(2:ncol(y))],pattern="t_stat_",replacement=""))
})
## fix the labels for the collapsed clustering (everything except ARC and VMH starts XX)
setnames(slibdvis[[4]],gsub(names(slibdvis[[4]]),pattern="XX",replacement="X"))

## now annotate the respective visium clusters with ARC/VMH where applicable
setnames(slibdvis[[4]],gsub(names(slibdvis[[4]]),pattern="XVMH",replacement="VMH"))
setnames(slibdvis[[4]],gsub(names(slibdvis[[4]]),pattern="XARC",replacement="ARC"))
setnames(slibdvis[[4]],"X13","ARC.ventral")


### assign VMH(k).n labels to the corresponding uncollapsed k15/20k31 results
setnames(slibdvis[[1]],"X4","ARC.1")
setnames(slibdvis[[1]],"X6","ARC.2")
setnames(slibdvis[[1]],"X7","VMH.1")
setnames(slibdvis[[1]],"X12","VMH.2")
setnames(slibdvis[[1]],"X13","ARC.ventral")

### add the k20/k31 ARC and VMH cluster identifiers
setnames(slibdvis[[2]],"X2","VMH20.1")
setnames(slibdvis[[2]],"X3","VMH20.2")
setnames(slibdvis[[2]],"X1","ARC20.1")
setnames(slibdvis[[2]],"X4","ARC20.2")
setnames(slibdvis[[2]],"X11","ARC20.ventral1")
setnames(slibdvis[[2]],"X12","ARC20.ventral2")

# k=31
setnames(slibdvis[[3]],"X6","VMH31.1")
setnames(slibdvis[[3]],"X21","VMH31.2")
setnames(slibdvis[[3]],"X25","VMH31.3")
setnames(slibdvis[[3]],"X26","VMH31.4")
setnames(slibdvis[[3]],"X31","VMH31.5")

setnames(slibdvis[[3]],"X2","ARC31.1")
setnames(slibdvis[[3]],"X4","ARC31.2")
setnames(slibdvis[[3]],"X11","ARC31.3")
setnames(slibdvis[[3]],"X22","ARC31.4")
setnames(slibdvis[[3]],"X17","ARC31.ventral1")
setnames(slibdvis[[3]],"X24","ARC31ventral2")

regres <- bpmapply(m=slibdvis,z=names(slibdvis),BPPARAM=MulticoreParam(4),SIMPLIFY=FALSE,FUN=function(m,z){
    
    j<-seq(10,300,by=10)
    i<-1
    outlist <- list()
    cortab <- merge.data.table(m,slibdxen,by="ensembl")
    cortab$ensembl <- NULL
    for (i in c(1:(length(j)+1))){
        if(i<=length(j)){
            ngene <- j[i]
        }
        else{
            # no value for top_n --> all genes used
            ngene <- nrow(cortab)
        }
        # initialize cormat
        nclxen <- ncol(slibdxen)
        nclhyp <- ncol(m)
        cormat <- as.data.frame(matrix(nrow=nclhyp-1,ncol=nclxen-1))
        # skim the first column, the gene id
        k <- 2
        for (k in c(2:nclhyp)){
            rownames(cormat)[k-1] <- colnames(m)[k]
            l <- 2
            for (l in c(2:nclxen)){
                colnames(cormat)[l-1] <- colnames(slibdxen)[l]
                
                curhyp <- colnames(m)[k]
                curxen <- colnames(slibdxen)[l]
                setorderv(cortab,curhyp,-1)
                
                cormat[k,l] <- cor(cortab[c(1:ngene),..curhyp],cortab[c(1:ngene),..curxen],use="pairwise.complete.obs",method = "spearman")
            }
        }
        plttab <- as.data.table(cormat,keep.rownames=T)
        setnames(plttab,"rn","sphyp_domain")
        outlist[[i]] <- plttab
        names(outlist)[i] <- paste0("Xenium_to_Visium_",z,"_top_",ngene,"_sphypdomainmks")
    }
    rm(cortab)
    return(outlist)
})

saveRDS(regres,"processed-data/10_Spatial_Registration/02-Visium_to_Xenium_spatialLIBD_registration.RDS")

```

reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.38.0         parallelly_1.38.0          
 [3] colorout_1.3-0.2            spatialLIBD_1.16.2         
 [5] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
 [7] SummarizedExperiment_1.34.0 Biobase_2.64.0             
 [9] GenomicRanges_1.56.1        GenomeInfoDb_1.40.1        
[11] IRanges_2.38.1              S4Vectors_0.42.1           
[13] BiocGenerics_0.50.0         MatrixGenerics_1.16.0      
[15] matrixStats_1.3.0           data.table_1.15.4          
[17] rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3        rstudioapi_0.16.0         jsonlite_1.8.8           
  [4] magrittr_2.0.3            ggbeeswarm_0.7.2          magick_2.8.3             
  [7] rmarkdown_2.27            BiocIO_1.14.0             fields_16.2              
 [10] zlibbioc_1.50.0           vctrs_0.6.5               DelayedMatrixStats_1.26.0
 [13] memoise_2.0.1             config_0.3.2              Rsamtools_2.20.0         
 [16] paletteer_1.6.0           RCurl_1.98-1.14           benchmarkme_1.0.8        
 [19] htmltools_0.5.8.1         S4Arrays_1.4.1            AnnotationHub_3.12.0     
 [22] curl_5.2.1                BiocNeighbors_1.22.0      SparseArray_1.4.8        
 [25] sass_0.4.9                bslib_0.7.0               htmlwidgets_1.6.4        
 [28] plotly_4.10.4             cachem_1.1.0              GenomicAlignments_1.40.0 
 [31] mime_0.12                 lifecycle_1.0.4           iterators_1.0.14         
 [34] pkgconfig_2.0.3           rsvd_1.0.5                Matrix_1.7-0             
 [37] R6_2.5.1                  fastmap_1.2.0             GenomeInfoDbData_1.2.12  
 [40] shiny_1.8.1.1             digest_0.6.36             colorspace_2.1-0         
 [43] rematch2_2.1.2            AnnotationDbi_1.66.0      scater_1.32.0            
 [46] rprojroot_2.0.4           irlba_2.3.5.1             ExperimentHub_2.12.0     
 [49] RSQLite_2.3.7             beachmat_2.20.0           filelock_1.0.3           
 [52] fansi_1.0.6               httr_1.4.7                abind_1.4-5              
 [55] compiler_4.4.1            here_1.0.1                bit64_4.0.5              
 [58] doParallel_1.0.17         attempt_0.3.1             viridis_0.6.5            
 [61] DBI_1.2.3                 maps_3.4.2                sessioninfo_1.2.2        
 [64] rappdirs_0.3.3            DelayedArray_0.30.1       rjson_0.2.21             
 [67] tools_4.4.1               vipor_0.4.7               beeswarm_0.4.0           
 [70] httpuv_1.6.15             glue_1.7.0                restfulr_0.0.15          
 [73] promises_1.3.0            grid_4.4.1                generics_0.1.3           
 [76] gtable_0.3.5              tidyr_1.3.1               ScaledMatrix_1.12.0      
 [79] BiocSingular_1.20.0       utf8_1.2.4                XVector_0.44.0           
 [82] ggrepel_0.9.5             BiocVersion_3.19.1        foreach_1.5.2            
 [85] pillar_1.9.0              spam_2.10-0               limma_3.60.3             
 [88] later_1.3.2               benchmarkmeData_1.0.4     dplyr_1.1.4              
 [91] BiocFileCache_2.12.0      lattice_0.22-6            rtracklayer_1.64.0       
 [94] bit_4.0.5                 tidyselect_1.2.1          locfit_1.5-9.10          
 [97] scuttle_1.14.0            Biostrings_2.72.1         knitr_1.48               
[100] gridExtra_2.3             edgeR_4.2.0               xfun_0.45                
[103] statmod_1.5.0             DT_0.33                   UCSC.utils_1.0.0         
[106] lazyeval_0.2.2            yaml_2.3.9                shinyWidgets_0.8.6       
[109] evaluate_0.24.0           codetools_0.2-20          tibble_3.2.1             
[112] BiocManager_1.30.23       cli_3.6.3                 xtable_1.8-4             
[115] jquerylib_0.1.4           munsell_0.5.1             golem_0.4.1              
[118] Rcpp_1.0.13               dbplyr_2.5.0              png_0.1-8                
[121] XML_3.99-0.17             ggplot2_3.5.1             blob_1.2.4               
[124] dotCall64_1.1-1           sparseMatrixStats_1.16.0  bitops_1.0-7             
[127] viridisLite_0.4.2         scales_1.3.0              purrr_1.0.2              
[130] crayon_1.5.3              cowplot_1.1.3             KEGGREST_1.44.1          
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-05
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings             2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra              2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
