---
title: "01-One-vs-all marker gene identification"
author: "Bernie Mulvey"
date: "2024-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(spatialLIBD)
library(parallel)
library(scran)
library(BiocParallel)

```

load required files
```{r}
hyp2 <- readRDS("processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

## this retrieves clustering results from both the 10k-iteration-pass bayesspace runs AND the subset that were run at 60k iterations
bscl <- list.files("processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out",pattern=".txt",recursive = F,full.names = T)

bscls <- lapply(bscl,FUN=fread)
names(bscls) <- bscl
# replace file paths with informative short labels
names(bscls) <- gsub(names(bscls),pattern="^.*60kiter.*\\/BSpace_(.*)\\.txt$",replacement=paste0("\\1","_60kiter"))

## add one additional clustering result where we collapse the VMH and ARC domains into one cluster each. do this for the primary clustering of interest -- bayesspace60kiter k15 harmony-lambda=NA with nnsvg10. we don't really need to do this over and over since the collective VMH or ARC spots will be ~the same regardless of how many subdivisions of it we made via different k values
## VMH: 11,12; ARC 3,5
collapse <- copy(bscls[["k15_HARMONYlmbna_nnsvg10_60kiter"]])
setnames(collapse,2,"cl")
collapse[,cl:=paste0("X",cl)]

collapse[cl %in% c("X7","X12"),cl:="VMH"]
collapse[cl %in% c("X4","X6",),cl:="ARC"]

bscls[["k15_HARMONYlmbna_nnsvg10_60kiter_collapsed"]] <- collapse


```

```{r}
# extract gene annotation table to join back to results
tmprowdat <- as.data.table(rowData(hyp2))

## iterate through clustering results: assign current clustering result to the hyp2$label column, identify markers within those assignments, return marker test outputs, proceed to next assignment set 

# parallelize within individual calls to scran, otherwise we're cloning the whole spe over and over
bpparam <- MulticoreParam(workers=8)
register(bpparam)

i<-1
bs.markers <- list()
for (i in c(1:length(bscls))){
    y <- bscls[[i]]
    y <- as.data.frame(y)
    rownames(y) <- y$rn
    # safeguard against reordered rows
    y <- y[rownames(colData(hyp2)),]
    colnames(y)[2] <- "cl"
    y$cl <- paste0("X",y$cl)
    # row names as gene symbols for sanity
    colLabels(hyp2) <- as.factor(y$cl)
    rm(y)
    gc(full=T)
    
    bs.markers[[i]] <- scoreMarkers(hyp2,colLabels(hyp2),BPPARAM=bpparam)
    names(bs.markers)[i] <- names(bscls)[i]
}

```

### collapse to top 100 markers (for each metric) per measure and clusters per clustering run, then to one giganto table
```{r}
i<-1
for (i in c(1:length(bs.markers))){
    k<-1
    for (k in c(1:length(bs.markers[[i]]))){
        tmp <- as.data.frame(bs.markers[[i]]@listData[[k]])
        # gene ids are only in the rownames so hang onto those when converting to d.t.
        tmp <- as.data.table(tmp,keep.rownames=T)
        # top 100 by rank per metric
        tmp <- tmp[rank.logFC.cohen<=100|rank.AUC<=100|rank.logFC.detected<=100]
        # clustering info
        tmp[,cluster_set:=names(bs.markers)[i]]
        tmp[,cluster:=names(bs.markers[[i]])[k]]
        # add information from rowData so we have gene symbols
        tmp <- merge.data.table(tmp,tmprowdat[,.(gene_id,gene_name)],by.x="rn",by.y="gene_id")
        if(i==1&k==1){
            bs.markers.tab <- tmp
        }
        else{
            bs.markers.tab <- rbind(bs.markers.tab,tmp)
        }
        rm(tmp)
    }
    rm(k)
}
rm(i)

## put clustering info, gene info, rank columns first
setnames(bs.markers.tab,"rn","gene_id")
bmtnamesout <- c("cluster_set","cluster","gene_id","gene_name",colnames(bs.markers.tab)[!colnames(bs.markers.tab)%in%c("cluster_set","cluster","gene_id","gene_name")])
bs.markers.tab <- bs.markers.tab[,..bmtnamesout]

saveRDS(bs.markers.tab,"processed-data/07-Marker_genes/01a-ScoreMarkers_BSpace_allruns_top100_markers_1vsAll_perdomain_perrun.RDS")

# clean up to free memory for spatiallibd below
rm(bs.markers.tab,bs.markers,tmprowdat,i,bmtnamesout,bpparam)
gc(full=T)
```

### get additional marker stats using spatialLIBD::registration_wrapper (return "enrichment" analyses)
```{r}
# we can't use biocparallel in this function, so bplapply it across just a couple-three cores to avoid overloading the machine

slibdenrichments <- list()
i<-1
for (i in c(1:length(bscls))){
    y <- as.data.frame(bscls[[i]])
    rownames(y) <- y$rn
    # safeguard against reordered rows
    y <- y[colnames(hyp2),]
    colnames(y)[2] <- "cl"
    y$cl <- paste0("X",y$cl)

    colLabels(hyp2) <- y$cl
    
    tmp <- registration_wrapper(hyp2,"label","sample_id",gene_ensembl="gene_id",gene_name="gene_name",min_ncells = 10)
    z <- as.data.table(tmp[["enrichment"]])
    slibdenrichments[[i]] <- z
    names(slibdenrichments)[i] <- names(bscls)[i]
    
    rm(y,z,tmp)
}



# save full outputs for each clustering run
saveRDS(slibdenrichments,"processed-data/07-Marker_genes/01b-BSpace_allruns_spatialLIBD-regwrap-enrichment_full.RDS")

```

reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] BiocParallel_1.38.0         scran_1.32.0               
 [3] scuttle_1.14.0              gridExtra_2.3              
 [5] spatialLIBD_1.16.2          SpatialExperiment_1.14.0   
 [7] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
 [9] Biobase_2.64.0              GenomicRanges_1.56.1       
[11] GenomeInfoDb_1.40.1         IRanges_2.38.1             
[13] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[15] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[17] colorout_1.3-0.2            data.table_1.15.4          
[19] rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               prismatic_1.1.2           BiocIO_1.14.0            
  [4] bitops_1.0-7              filelock_1.0.3            fields_16.2              
  [7] tibble_3.2.1              XML_3.99-0.17             lifecycle_1.0.4          
 [10] edgeR_4.2.0               doParallel_1.0.17         rprojroot_2.0.4          
 [13] lattice_0.22-6            magrittr_2.0.3            limma_3.60.3             
 [16] plotly_4.10.4             sass_0.4.9                rmarkdown_2.27           
 [19] jquerylib_0.1.4           yaml_2.3.9                metapod_1.12.0           
 [22] httpuv_1.6.15             spam_2.10-0               sessioninfo_1.2.2        
 [25] cowplot_1.1.3             DBI_1.2.3                 RColorBrewer_1.1-3       
 [28] golem_0.4.1               maps_3.4.2                abind_1.4-5              
 [31] zlibbioc_1.50.0           purrr_1.0.2               RCurl_1.98-1.14          
 [34] rappdirs_0.3.3            GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [37] irlba_2.3.5.1             dqrng_0.4.1               DelayedMatrixStats_1.26.0
 [40] codetools_0.2-20          DelayedArray_0.30.1       DT_0.33                  
 [43] tidyselect_1.2.1          UCSC.utils_1.0.0          farver_2.1.2             
 [46] ScaledMatrix_1.12.0       viridis_0.6.5             shinyWidgets_0.8.6       
 [49] BiocFileCache_2.12.0      GenomicAlignments_1.40.0  jsonlite_1.8.8           
 [52] BiocNeighbors_1.22.0      scater_1.32.0             iterators_1.0.14         
 [55] foreach_1.5.2             tools_4.4.1               Rcpp_1.0.13              
 [58] glue_1.7.0                SparseArray_1.4.8         xfun_0.45                
 [61] here_1.0.1                dplyr_1.1.4               withr_3.0.0              
 [64] BiocManager_1.30.23       fastmap_1.2.0             bluster_1.14.0           
 [67] fansi_1.0.6               digest_0.6.36             rsvd_1.0.5               
 [70] R6_2.5.1                  mime_0.12                 colorspace_2.1-0         
 [73] RSQLite_2.3.7             config_0.3.2              utf8_1.2.4               
 [76] tidyr_1.3.1               generics_0.1.3            rtracklayer_1.64.0       
 [79] httr_1.4.7                htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [82] pkgconfig_2.0.3           gtable_0.3.5              blob_1.2.4               
 [85] XVector_0.44.0            htmltools_0.5.8.1         dotCall64_1.1-1          
 [88] scales_1.3.0              png_0.1-8                 attempt_0.3.1            
 [91] knitr_1.48                rstudioapi_0.16.0         rjson_0.2.21             
 [94] curl_5.2.1                cachem_1.1.0              BiocVersion_3.19.1       
 [97] vipor_0.4.7               AnnotationDbi_1.66.0      restfulr_0.0.15          
[100] pillar_1.9.0              grid_4.4.1                vctrs_0.6.5              
[103] promises_1.3.0            BiocSingular_1.20.0       dbplyr_2.5.0             
[106] beachmat_2.20.0           cluster_2.1.6             xtable_1.8-4             
[109] beeswarm_0.4.0            paletteer_1.6.0           evaluate_0.24.0          
[112] magick_2.8.3              cli_3.6.3                 locfit_1.5-9.10          
[115] compiler_4.4.1            Rsamtools_2.20.0          crayon_1.5.3             
[118] labeling_0.4.3            rematch2_2.1.2            ggbeeswarm_0.7.2         
[121] viridisLite_0.4.2         munsell_0.5.1             Biostrings_2.72.1        
[124] lazyeval_0.2.2            Matrix_1.7-0              ExperimentHub_2.12.0     
[127] benchmarkme_1.0.8         sparseMatrixStats_1.16.0  bit64_4.0.5              
[130] ggplot2_3.5.1             KEGGREST_1.44.1           statmod_1.5.0            
[133] shiny_1.8.1.1             AnnotationHub_3.12.0      igraph_2.0.3             
[136] memoise_2.0.1             bslib_0.7.0               benchmarkmeData_1.0.4    
[139] bit_4.0.5                
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-24
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings             2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bluster                1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   cluster                2.1.6     2023-12-01 [1] CRAN (R 4.4.1)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                  0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.0.3     2024-03-13 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   metapod                1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   prismatic              1.1.2     2024-04-10 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.1)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scran                * 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
