---
title: "03-Marker Gene Supp Tab Making"
author: "Bernard Mulvey"
date: "2024-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

```

Load up the spatialLIBD results and write out
```{r}
slibdmk <- readRDS("processed-data/07-Marker_genes/01b-BSpace_allruns_spatialLIBD-regwrap-enrichment_full.RDS")

slibdmk <- slibdmk[names(slibdmk) %in% grep(names(slibdmk),pattern="HARMONYlmbna_nnsvg10_60kiter",value=T)]

### fix VMH/ARC cluster labels for the collapsed case
setnames(slibdmk[[4]],gsub(names(slibdmk[[4]]),pattern="XVMH",replacement="VMH"))
setnames(slibdmk[[4]],gsub(names(slibdmk[[4]]),pattern="XARC",replacement="ARC"))
setnames(slibdmk[[4]],gsub(names(slibdmk[[4]]),pattern="XX",replacement="X"))
setnames(slibdmk[[4]],gsub(names(slibdmk[[4]]),pattern="X13",replacement="ARC.ventral"))


### assign VMH(k).n labels to the corresponding uncollapsed k15/20k31 results
setnames(slibdmk[[1]],gsub(names(slibdmk[[1]]),pattern="X4",replacement="ARC.1"))
setnames(slibdmk[[1]],gsub(names(slibdmk[[1]]),pattern="X6",replacement="ARC.2"))
setnames(slibdmk[[1]],gsub(names(slibdmk[[1]]),pattern="X7",replacement="VMH.1"))
setnames(slibdmk[[1]],gsub(names(slibdmk[[1]]),pattern="X12",replacement="VMH.2"))
setnames(slibdmk[[1]],gsub(names(slibdmk[[1]]),pattern="X13",replacement="ARC.ventral"))

### add the k20/k31 ARC and VMH cluster identifiers
setnames(slibdmk[[2]],gsub(names(slibdmk[[2]]),pattern="X2$",replacement="VMH20.1"))
setnames(slibdmk[[2]],gsub(names(slibdmk[[2]]),pattern="X3",replacement="VMH20.2"))
setnames(slibdmk[[2]],gsub(names(slibdmk[[2]]),pattern="X1$",replacement="ARC20.1"))
setnames(slibdmk[[2]],gsub(names(slibdmk[[2]]),pattern="X4",replacement="ARC20.2"))
setnames(slibdmk[[2]],gsub(names(slibdmk[[2]]),pattern="X11",replacement="ARC20.3"))
setnames(slibdmk[[2]],gsub(names(slibdmk[[2]]),pattern="X12",replacement="ARC20.4"))

# k=31
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X6",replacement="VMH31.1"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X21",replacement="VMH31.2"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X25",replacement="VMH31.3"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X26",replacement="VMH31.4"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X31",replacement="VMH31.5"))

setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X2$",replacement="ARC31.1"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X4",replacement="ARC31.2"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X11",replacement="ARC31.3"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X17",replacement="ARC31.4"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X22",replacement="ARC31.5"))
setnames(slibdmk[[3]],gsub(names(slibdmk[[3]]),pattern="X24",replacement="ARC31.6"))

### 
slmn <- names(slibdmk)
mapply(x=slibdmk,y=slmn,FUN=function(x,y){
    z <- melt.data.table(x,id.vars=c("ensembl","gene"),measure.vars=names(x)[c(1:(ncol(x)-2))])
    z[,domain:=gsub(variable,pattern="^.*_(X.*)$",replacement="\\1")]
    z[variable %in% grep(variable,pattern="VMH",value=T),domain:=gsub(variable,pattern="^.*_(VMH.*)$",replacement="\\1")]
    z[variable %in% grep(variable,pattern="ARC",value=T),domain:=gsub(variable,pattern="^.*_(ARC.*)$",replacement="\\1")]
    
    z[,variable:=gsub(variable,pattern="^(.*)_X.*$",replacement="\\1")]
    z[variable %in% grep(variable,pattern="VMH",value=T),variable:=gsub(variable,pattern="^(.*)_VMH.*$",replacement="\\1")]
    z[variable %in% grep(variable,pattern="ARC",value=T),variable:=gsub(variable,pattern="^(.*)_ARC.*$",replacement="\\1")]
    
    z <- dcast.data.table(z,ensembl+gene+domain~variable)
    fwrite(z,paste0("local_supptab_assembly/Marker Analysis",y,".txt"),sep="\t",quote=F)
})

```

## export one-to-one comparisons of ARC and VMH from each. since there's only one cluster compared in any iteration, we only need the mean and rank columns (which == min, max, median) 
```{r}
### get gene info to merge using the SPE
genedat <- fread("raw-data/USEFOR10X_ensg_lut_hg38_fromEns98_111123.txt")
genedat[,chromosome_name:=paste0("chr",chromosome_name)]
genedat <- unique(genedat[,.(ensembl_gene_id,external_gene_name,hgnc_symbol,chromosome_name)])
setnames(genedat,c("ensembl_gene_id","external_gene_name"),c("gene_id","gene_name"))
### load pairwise results
comparatives <- readRDS("processed-data/07-Marker_genes/02-nnsvg10_HRM-lmbna_bspace_VMHandARC-clustPair-diffmarkers.RDS")


### since these were run only with the two clusters of interest as input, we only need the first output (there's outputs for cluster A vs cluster B and B vs A from each, which are equivalent as depleted in A must then == enriched in B)

comparewrangle <- list()
comparewrangle[[1]] <- as.data.table(comparatives[[1]][[1]],keep.rownames=T)
comparewrangle[[2]] <- as.data.table(comparatives[[2]][[1]],keep.rownames=T)
names(comparewrangle) <- names(comparatives)[1:2]

### subset to only the mean and rank columns + relevant meta
keepnames <- c("rn",grep(names(comparewrangle[[1]]),pattern="self|other|mean|rank",value=T))
comparewrangle <- lapply(comparewrangle,FUN=function(x){
  x <- x[,..keepnames]
  return(x)
})

### add gene info
comparewrangle <- lapply(comparewrangle,FUN=function(x){
  x <- merge.data.table(genedat[,.(gene_id,gene_name)],x,by.x="gene_id",by.y="rn")
  return(x)
})

mapply(x=comparewrangle,y=names(comparewrangle),FUN=function(x,y){
    fwrite(x,paste0("local_supptab_assembly/Sup Tab-VMHorARC pairwise Marker Analyses",y,".txt"),sep="\t",quote=F)
})


```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] S4Vectors_0.42.1    BiocGenerics_0.50.0 colorout_1.3-0.2    data.table_1.15.4  
[5] rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] digest_0.6.36     fastmap_1.2.0     xfun_0.45         knitr_1.48       
 [5] htmltools_0.5.8.1 rmarkdown_2.27    cli_3.6.3         sessioninfo_1.2.2
 [9] compiler_4.4.1    rprojroot_2.0.4   here_1.0.1        rstudioapi_0.16.0
[13] tools_4.4.1       evaluate_0.24.0   yaml_2.3.9       
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-30
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package      * version date (UTC) lib source
   BiocGenerics * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   data.table   * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest         0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   evaluate       0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.48    2024-07-07 [1] CRAN (R 4.4.1)
 P rlang        * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors    * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   xfun           0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
