---
title: "04-Supptab_k15_Visium_Xn-to-inferred-type-lookup-tab"
output: html_document
date: "2024-08-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

read in BayesSpace k15 nnsvg10 harmony lambda NA sheet
```{r}
bscl <- fread("processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt")


setnames(bscl,2,"cl")
bscl[,cl:=paste0("X",cl)]
bscl[cl=="X7",cl2:="VMH.1"]
bscl[cl=="X12",cl2:="VMH.2"]
bscl[cl=="X4",cl2:="ARC.1"]
bscl[cl=="X6",cl2:="ARC.2"]
# WM (optic tract, OT): Clusters 3 and 10
bscl[cl=="X3",cl2:="OT.1"]
bscl[cl=="X10",cl2:="OT.2"]
bscl[cl=="X5",cl2:="OT.3"]
# others:
bscl[cl=="X1",cl2:="GABA.1"]
bscl[cl=="X2",cl2:="PeriVN"]
bscl[cl=="X9",cl2:="Vascular"]
bscl[cl=="X8",cl2:="SON"]
bscl[cl=="X13",cl2:="Portal Vasc."]
bscl[cl=="X14",cl2:="Astro"]
bscl[cl=="X15",cl2:="GABA.2"]
bscl[cl=="X11",cl2:="Glut.1"]

### long names
bscl[cl=="X7",cl3:="Ventromedial nucleus cluster 1"]
bscl[cl=="X12",cl3:="Ventromedial nucleus cluster 2"]
bscl[cl=="X4",cl3:="Arcuate nucleus cluster 1"]
bscl[cl=="X6",cl3:="Arcuate nucleus cluster 2"]
# WM (optic tract, OT): Clusters 3 and 10
bscl[cl=="X3",cl3:="Optic tract cluster 1"]
bscl[cl=="X10",cl3:="Optic tract cluster 2"]
bscl[cl=="X5",cl3:="Optic tract cluster 3"]
# others:
bscl[cl=="X1",cl3:="GABAergic cluster 1"]
bscl[cl=="X2",cl3:="Periventricular nucleus"]
bscl[cl=="X9",cl3:="Vascular"]
bscl[cl=="X8",cl3:="Supraoptic nucleus"]
bscl[cl=="X13",cl3:="Hypothalamic-Pituitary Portal Vascular System"]
bscl[cl=="X14",cl3:="Astrocytes"]
bscl[cl=="X15",cl3:="GABAergic cluster 2"]
bscl[cl=="X11",cl3:="Glutamatergic cluster 1"]

### other notes
bscl[cl=="X1",cl4:="Other hypothalamic GABAergic neurons, identity not established"]
bscl[cl=="X15",cl4:="Other hypothalamic GABAergic neurons, identity not established"]
bscl[cl=="X11",cl4:="Other hypothalamic or hypothalamus-adjacent excitatory neurons, identity not established"]


bscl <- unique(bscl[,.(cl,cl2,cl3,cl4)])

setnames(bscl,c("cl","cl2","cl3","cl4"),c("Cluster ID", "Cluster Short Name","Cluster Full Name","Other Notes"))

fwrite(bscl,"local_supptab_assembly/k15_nnSVG10_cluster-Xn-ID_to_inferred_type_lookuptable.txt",sep='\t',quote=F)

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-0.2  data.table_1.15.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] compiler_4.4.1    here_1.0.1        fastmap_1.2.0     rprojroot_2.0.4  
 [5] cli_3.6.3         htmltools_0.5.8.1 tools_4.4.1       rstudioapi_0.16.0
 [9] yaml_2.3.9        rmarkdown_2.27    knitr_1.48        xfun_0.45        
[13] digest_0.6.36     evaluate_0.24.0   sessioninfo_1.2.2
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-09
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package     * version date (UTC) lib source
 P cli           3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout    * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   data.table  * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest        0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   evaluate      0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   here          1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools     0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr         1.48    2024-07-07 [1] CRAN (R 4.4.1)
 P rlang       * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown     2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot     2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi    0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   sessioninfo   1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   xfun          0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml          2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────