---
title: "05-IntraVMH or IntraARC 1vothers nnsvg10_k15_20_31"
output: html_document
date: "2024-09-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(spatialLIBD)
library(scran)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load required files
```{r}
hyp2 <- readRDS("processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

## remove image data to parallelize
imgData(hyp2) <- NULL
gc(full=T)

bscls <- list.files("processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out",full.names=T,pattern="HARMONYlmbna_nnsvg10")

bscls.dt <- lapply(bscls,FUN=function(t){
    b <- fread(t)
    setnames(b,2,"cl")
    b[,cl:=paste0("X",cl)]
    return(b)
})
names(bscls.dt) <- c("k15","k20","k31")

bscls.dt[["k15"]][cl=="X4",cl:="ARC1"]
bscls.dt[["k15"]][cl=="X6",cl:="ARC2"]
bscls.dt[["k15"]][cl=="X7",cl:="VMH1"]
bscls.dt[["k15"]][cl=="X12",cl:="VMH2"]

bscls.dt[["k20"]][cl=="X2",cl:="VMH20_1"]
bscls.dt[["k20"]][cl=="X3",cl:="VMH20_2"]
bscls.dt[["k20"]][cl=="X1",cl:="ARC20_1"]
bscls.dt[["k20"]][cl=="X4",cl:="ARC20_2"]

# k31: x6 21 25 26 31 = VMH; 2,4,11,22 ARC
bscls.dt[["k31"]][cl=="X6",cl:="VMH31_1"]
bscls.dt[["k31"]][cl=="X21",cl:="VMH31_2"]
bscls.dt[["k31"]][cl=="X25",cl:="VMH31_3"]
bscls.dt[["k31"]][cl=="X26",cl:="VMH31_4"]
bscls.dt[["k31"]][cl=="X31",cl:="VMH31_5"]
bscls.dt[["k31"]][cl=="X2",cl:="ARC31_1"]
bscls.dt[["k31"]][cl=="X4",cl:="ARC31_2"]
bscls.dt[["k31"]][cl=="X11",cl:="ARC31_3"]
bscls.dt[["k31"]][cl=="X22",cl:="ARC31_4"]
```

## Use spatialLIBD registration tools for pairwise markers.
```{r}
## didn't work as an lapply for some reason, but can do this iteratively.
select.vmharc.markers <- list()
i<-1

for (i in c(1:length(bscls.dt))){
    b <- copy(bscls.dt[[i]])
    
    vmhs <- grep(unique(b$cl),pattern="VMH",value=T)
    vmhres <- lapply(vmhs,FUN=function(v){
        btmp <- copy(b)
        btmp[cl %in% vmhs & cl!=v,cl:="zotherVMHs"]
        btmp <- btmp[cl %in% c(v,"zotherVMHs")]
        hyptmp <- hyp2[,hyp2$key %in% btmp$rn]
        btmp <- DataFrame(btmp,row.names = btmp$rn)[colnames(hyptmp),]
        colLabels(hyptmp) <- btmp$cl
        
        ## stepwise set up to run registration_pairwise
        hyptmp.pb <- registration_pseudobulk(hyptmp,var_registration = "label",var_sample_id = "sample_id")
        mkmod <- registration_model(hyptmp.pb,var_registration = "label")
        bcor <- registration_block_cor(hyptmp.pb,registration_model = mkmod,var_sample_id = "sample_id")
        
        # run pairwise reg
        mk <- registration_stats_pairwise(sce_pseudo = hyptmp.pb, registration_model = mkmod, var_registration = "label",block_cor = bcor, var_sample_id = "sample_id", gene_ensembl = "gene_id",gene_name = "gene_name")
        mkres <- as.data.table(mk,keep.rownames=T)
        return(mkres)
    })
    names(vmhres) <- vmhs
    
    arcs <- grep(unique(b$cl),pattern="ARC",value=T)
    arcres <- lapply(arcs,FUN=function(v){
        btmp <- copy(b)
        btmp[cl %in% arcs & cl!=v,cl:="zotherARCs"]
        btmp <- btmp[cl %in% c(v,"zotherARCs")]
        hyptmp <- hyp2[,hyp2$key %in% btmp$rn]
        btmp <- DataFrame(btmp,row.names = btmp$rn)[colnames(hyptmp),]
        colLabels(hyptmp) <- btmp$cl
        
        ## stepwise set up to run registration_pairwise
        hyptmp.pb <- registration_pseudobulk(hyptmp,var_registration = "label",var_sample_id = "sample_id")
        mkmod <- registration_model(hyptmp.pb,var_registration = "label")
        bcor <- registration_block_cor(hyptmp.pb,registration_model = mkmod,var_sample_id = "sample_id")
        
        # run pairwise reg
        mk <- registration_stats_pairwise(sce_pseudo = hyptmp.pb, registration_model = mkmod, var_registration = "label",block_cor = bcor, var_sample_id = "sample_id", gene_ensembl = "gene_id",gene_name = "gene_name")
        mkres <- as.data.table(mk,keep.rownames=T)
        return(mkres)
    })
    names(arcres) <- arcs
    
    retlist <- list(vmhres,arcres)
    names(retlist) <- c("vmh","arc")
    select.vmharc.markers[[i]] <- retlist
    
    rm(vmhs,arcs,arcres,vmhres,b)
    gc(full=T)
}

names(select.vmharc.markers) <- names(bscls.dt)
```



```{r}
### save results as RDS
saveRDS(select.vmharc.markers,file="processed-data/07-Marker_genes/05-IntraVMH-intraARC_1clus_vallother_clus_slibdEnrichment.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  grid      stats4    stats     graphics  grDevices utils     methods  
[9] base     

other attached packages:
 [1] scran_1.32.0                scuttle_1.14.0             
 [3] BiocParallel_1.38.0         parallelly_1.38.0          
 [5] colorout_1.3-0.2            ComplexHeatmap_2.20.0      
 [7] spatialLIBD_1.16.2          SpatialExperiment_1.14.0   
 [9] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
[11] Biobase_2.64.0              GenomicRanges_1.56.1       
[13] GenomeInfoDb_1.40.1         IRanges_2.38.1             
[15] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[17] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[19] gridExtra_2.3               ggplot2_3.5.1              
[21] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0             bitops_1.0-7             
  [4] filelock_1.0.3            fields_16.2               tibble_3.2.1             
  [7] XML_3.99-0.17             lifecycle_1.0.4           edgeR_4.2.0              
 [10] doParallel_1.0.17         rprojroot_2.0.4           lattice_0.22-6           
 [13] magrittr_2.0.3            limma_3.60.3              plotly_4.10.4            
 [16] sass_0.4.9                rmarkdown_2.27            jquerylib_0.1.4          
 [19] yaml_2.3.9                metapod_1.12.0            httpuv_1.6.15            
 [22] spam_2.10-0               sessioninfo_1.2.2         cowplot_1.1.3            
 [25] DBI_1.2.3                 RColorBrewer_1.1-3        golem_0.4.1              
 [28] maps_3.4.2                abind_1.4-5               zlibbioc_1.50.0          
 [31] purrr_1.0.2               RCurl_1.98-1.14           rappdirs_0.3.3           
 [34] circlize_0.4.16           GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [37] irlba_2.3.5.1             dqrng_0.4.1               DelayedMatrixStats_1.26.0
 [40] codetools_0.2-20          DelayedArray_0.30.1       DT_0.33                  
 [43] shape_1.4.6.1             tidyselect_1.2.1          UCSC.utils_1.0.0         
 [46] ScaledMatrix_1.12.0       viridis_0.6.5             shinyWidgets_0.8.6       
 [49] BiocFileCache_2.12.0      GenomicAlignments_1.40.0  jsonlite_1.8.8           
 [52] GetoptLong_1.0.5          BiocNeighbors_1.22.0      scater_1.32.0            
 [55] iterators_1.0.14          foreach_1.5.2             tools_4.4.1              
 [58] Rcpp_1.0.13               glue_1.7.0                SparseArray_1.4.8        
 [61] xfun_0.45                 here_1.0.1                dplyr_1.1.4              
 [64] withr_3.0.0               BiocManager_1.30.23       fastmap_1.2.0            
 [67] bluster_1.14.0            fansi_1.0.6               digest_0.6.36            
 [70] rsvd_1.0.5                R6_2.5.1                  mime_0.12                
 [73] colorspace_2.1-0          RSQLite_2.3.7             config_0.3.2             
 [76] utf8_1.2.4                tidyr_1.3.1               generics_0.1.3           
 [79] rtracklayer_1.64.0        httr_1.4.7                htmlwidgets_1.6.4        
 [82] S4Arrays_1.4.1            pkgconfig_2.0.3           gtable_0.3.5             
 [85] blob_1.2.4                XVector_0.44.0            htmltools_0.5.8.1        
 [88] dotCall64_1.1-1           clue_0.3-65               scales_1.3.0             
 [91] png_0.1-8                 attempt_0.3.1             knitr_1.48               
 [94] rstudioapi_0.16.0         rjson_0.2.21              curl_5.2.1               
 [97] cachem_1.1.0              GlobalOptions_0.1.2       BiocVersion_3.19.1       
[100] vipor_0.4.7               AnnotationDbi_1.66.0      restfulr_0.0.15          
[103] pillar_1.9.0              vctrs_0.6.5               promises_1.3.0           
[106] BiocSingular_1.20.0       dbplyr_2.5.0              beachmat_2.20.0          
[109] cluster_2.1.6             xtable_1.8-4              beeswarm_0.4.0           
[112] paletteer_1.6.0           evaluate_0.24.0           magick_2.8.3             
[115] cli_3.6.3                 locfit_1.5-9.10           compiler_4.4.1           
[118] Rsamtools_2.20.0          crayon_1.5.3              rematch2_2.1.2           
[121] ggbeeswarm_0.7.2          viridisLite_0.4.2         munsell_0.5.1            
[124] Biostrings_2.72.1         lazyeval_0.2.2            Matrix_1.7-0             
[127] ExperimentHub_2.12.0      benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[130] bit64_4.0.5               KEGGREST_1.44.1           statmod_1.5.0            
[133] shiny_1.8.1.1             AnnotationHub_3.12.0      igraph_2.0.3             
[136] memoise_2.0.1             bslib_0.7.0               benchmarkmeData_1.0.4    
[139] bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-20
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings             2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bluster                1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   circlize               0.4.16    2024-02-20 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   clue                   0.3-65    2023-09-23 [1] CRAN (R 4.4.0)
   cluster                2.1.6     2023-12-01 [1] CRAN (R 4.4.1)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   ComplexHeatmap       * 2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                  0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   GetoptLong             1.0.5     2020-12-15 [1] CRAN (R 4.4.0)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   GlobalOptions          0.1.2     2020-06-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.0.3     2024-03-13 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   metapod                1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scran                * 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────