---
title: "Untitled"
author: "Bernard Mulvey"
date: "2023-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
require(colorout)
library(RcppML)
library(Matrix)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

## makevars for RcppML with openML support on ventura + apple silicon (M1/M2 processors):
## ref: https://github.com/Rdatatable/data.table/issues/5419

LOC = /opt/gfortran
CC=$(LOC)/bin/aarch64-apple-darwin20.0-gcc -fopenmp
CXX=/usr/bin/g++ #-fopenmp
CXX11 = /usr/bin/g++ #-fopenmp # for fst package

CFLAGS=-g -O3 -Wall -pedantic -std=gnu99 -mtune=native -pipe
CXXFLAGS=-g -O3 -Wall -pedantic -std=c++11 -mtune=native -pipe
LDFLAGS=-L$(LOC)/lib -Wl,-rpath,$(LOC)/lib
CPPFLAGS=-I$(LOC)/include -I/Library/Developer/CommandLineTools/SDKs/MacOSX12.3.sdk/usr/include/c++/v1
LDFLAGS += -L/opt/homebrew/opt/libomp/lib -lomp
CPPFLAGS += -I/opt/homebrew/opt/libomp/include -Xpreprocessor -fopenmp


### Command line paths mannually added to R "site-wide" (entire installation on this mac) so that it would pick up command line executables installed with brew:
### See /Library/R.Framework/etc/Renviron.site





```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_D07-075_B1 crop 070223.RDS")
colnames(hyp2) <- hyp2$key

counts <- assay(hyp2,"logcounts")
spscts <- as(counts,"dgCMatrix")

options(RcppML.threads=8)
x <- nmf(spscts,18,tol = 10E-8,maxit = 1000,verbose = T,L1 = c(0,0),seed = 42,mask_zeros = FALSE,diag = TRUE,nonneg = TRUE)





```
