---
title: "01-Plot clusters on dimensionality reductions"
author: "Bernard Mulvey"
date: "2024-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
require(colorout)
library(ggrastr)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load reduced dim slot with the PCA, UMAP, harmony dimreds, and mnn dimreds for the feature sets. load bayesspace cluster assignments.
```{r}
hyp2dimreds <- readRDS("processed-data/04-feature_selection/02-hypfiltered_4featureset-pca-umap-harmonydefault-harmonylambdanull-mnn30_reducedDim_slot.RDS")

bs_clusts <- list.files("processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out",pattern=".txt",full.names=T,recursive = F)
```

Plot on feature-set-matched dimreds

```{r}
sbp <- MulticoreParam(8)
bplapply(bs_clusts,BPPARAM=sbp,FUN=function(x){
    # get feature set (hvg20, svg10, svg20), k val
    bsk <- gsub(x,pattern="^.*_out/BSpa.*_k(.*)_.*_.*\\.txt",replacement="\\1")
    drandfeats <- gsub(x,pattern="^.*_out/BSpa.*_k.*_(.*_.*)\\.txt",replacement="\\1")
    feats <- gsub(drandfeats,pattern="^.*_(.*)$",replacement="\\1")
    
    # get batch correction method short name
    curbatcor <- gsub(x,pattern="^.*k15-20-31_out/BSpace_k.._(.*)_[[:alnum:]]{5,8}\\.txt$",replacement="\\1")
    
    # get corresponding dimred slots
    curdimrs <- hyp2dimreds[grep(drandfeats,names(hyp2dimreds))]
    
    # make sure we have the ones we want (PCA, plain UMAP, and the batch method)
    keepdimrs <- c(paste0(c("PCA_","UMAP_"),feats),drandfeats)
    keepdimrs <- unlist(lapply(keepdimrs,FUN=function(x){grep(names(curdimrs),pattern=x,value=T)}))
    # for MNN dimreds, also include the corresponding UMAP-of-MNN
    if(length(grep(drandfeats,pattern="mnn30",value=T))>0){keepdimrs <- c(keepdimrs,paste0(drandfeats,"_UMAP"))}
    
    curdimrs <- curdimrs[keepdimrs]
    
    ## read in cluster assignments
    curclusts <- fread(x)
    setnames(curclusts,2,"cl")
    curclusts[,cl:=paste0("X",cl)]

    # plot on the feature set's uncorrected dimreds, and that feature set's corresponding batch correction method's dimreds
    
    j <- 1
    for (j in c(1:length(keepdimrs))){
        if(length(grep(keepdimrs[j],pattern="UMAP",value=T))>0){
            zq <- rbind(c(1,2))
        }
        else{
            zq <- rbind(c(1,2),c(1,3),c(2,3))
        }
        
        
    # iterate thru dimension pairs
        m <- 1
        for (m in c(1:nrow(zq))){
            z <- zq[m,1]
            q <- zq[m,2]
            pltdat <- as.data.table(curdimrs[[keepdimrs[j]]][,c(z,q)],keep.rownames=T)
            pltdat <- merge.data.table(pltdat,curclusts,by="rn")
            pltdat[,rn:=NULL]
            
            setnames(pltdat,c(paste0(keepdimrs[j],paste0("_Dim",z)),paste0(keepdimrs[j],paste0("_Dim",q)),paste0(drandfeats,"_k",bsk)))
            xvar <- names(pltdat)[1]
            yvar <- names(pltdat)[2]
            colvar <- names(pltdat)[3]
        
            set(pltdat,j=xvar,value=as.numeric(pltdat[,get(xvar)]))
            set(pltdat,j=yvar,value=as.numeric(pltdat[,get(yvar)]))
            set(pltdat,j=colvar,value=factor(pltdat[,get(colvar)],levels=paste0("X",c(1:length(unique(pltdat[,get(colvar)]))))))
        
            plt <- ggplot(pltdat,aes(x=.data[[xvar]],y=.data[[yvar]],col=.data[[colvar]]))+
                rasterize(geom_point(size=0.1),dpi = 300)+
                ggtitle(paste0(keepdimrs[j]," dims ",z, " and ",
                               q,"\n","BayesSpace k = ",bsk))
        
            pdf(paste0("plots/06-BayesSpace/01-Clusters on dimreds/",drandfeats,"-Bspc-k",bsk,"_dims",q,"_and_",z,".pdf"),width = 10,height = 8)
            print(plt)
            dev.off()
            rm(z,q,pltdat,xvar,yvar)
        }
    }
    rm(drandfeats,feats,bsk)
})

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] ggrastr_1.0.2       S4Vectors_0.42.1    BiocGenerics_0.50.0 BiocParallel_1.38.0
 [5] parallelly_1.38.0   colorout_1.3-0.2    gridExtra_2.3       ggplot2_3.5.1      
 [9] data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.5      dplyr_1.1.4       compiler_4.4.1    tidyselect_1.2.1 
 [5] ggbeeswarm_0.7.2  parallel_4.4.1    scales_1.3.0      yaml_2.3.9       
 [9] fastmap_1.2.0     here_1.0.1        R6_2.5.1          generics_0.1.3   
[13] knitr_1.48        tibble_3.2.1      munsell_0.5.1     rprojroot_2.0.4  
[17] pillar_1.9.0      utf8_1.2.4        xfun_0.45         cli_3.6.3        
[21] withr_3.0.0       magrittr_2.0.3    digest_0.6.36     grid_4.4.1       
[25] rstudioapi_0.16.0 beeswarm_0.4.0    lifecycle_1.0.4   vipor_0.4.7      
[29] vctrs_0.6.5       evaluate_0.24.0   glue_1.7.0        sessioninfo_1.2.2
[33] codetools_0.2-20  fansi_1.0.6       colorspace_2.1-0  rmarkdown_2.27   
[37] tools_4.4.1       pkgconfig_2.0.3   htmltools_0.5.8.1
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-31
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package      * version date (UTC) lib source
   beeswarm       0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   BiocGenerics * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table   * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest         0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate       0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi          1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics       0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggbeeswarm     0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2      * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr      * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   glue           1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable         0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly   * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar         1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
 P rlang        * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors    * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8           1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor          0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   withr          3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun           0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
