---
title: "01-Spatial Registration of Xenium Types to Yao23 Mouse HYP subclasses"
output: html_document
date: "2024-08-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load registration results (table of cluster-cluster marker t stat correlations)
```{r}
yaoreg <- readRDS("processed-data/10_Spatial_Registration/01-NonSampleSpecific_cluster_reg_to_Yao23_mouseHyp_subclasses_and_VMHARCsupertypes.RDS")

# collapse lists of results into one big list
plotlist <- c(
        yaoreg[[1]],
        yaoreg[[2]],
        yaoreg[[3]]
)

## account for any duplicate cases where the top n genes (all)==250 (i.e., duplicate runs)
plotlist <- plotlist[unique(names(plotlist))]

## load the cluster annotations to annotate our clusters on the plots
bkanno <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

bkanno <- unique(bkanno[,.(clus,bjm_annot)])

## add more precise ARC/VMH cluster labels
vmharclab <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04e_ARC_and_VMH_cluster_identities_based_on_cellgroup_1vall.txt")
bkanno <- merge.data.table(bkanno,vmharclab,by="bjm_annot",all.x=T)
bkanno[!is.na(subclus_annot),bjm_annot:=subclus_annot]

## drop sample specific clusters
bkanno <- bkanno[!(bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T))]
```

# plots, as heatmaps
```{r}
sbp <- MulticoreParam(8)
register(sbp)

bpmapply(m=plotlist,n=as.list(names(plotlist)),SIMPLIFY = F,BPPARAM = sbp,FUN = function(m,n){
        plttab <- as.data.table(m,keep.rownames=T)
        plttab <- melt(plttab,id.vars="sphyp_domain")
        plttab <- merge.data.table(plttab,bkanno,by.x="sphyp_domain",by.y="clus")
        plt <- ggplot(plttab,aes(x=bjm_annot,y=variable,fill=value))+
                geom_tile()+
                ggtitle(gsub(n,pattern="_",replacement=" "))+
                theme(axis.text.x = element_text(angle=90,size=9,hjust=1),axis.text.y=element_text(size=9),title = element_text(size=12))+
                scale_fill_viridis_c()
        n <-  ifelse(as.numeric(gsub(
                n,pattern="^.*_top_(.*)_sp.*$",replacement="\\1"))>1000,
                yes=gsub(n,pattern="^(.*)_top_.*_(sp.*)$",replacement=paste0("\\1","_allgenes_","\\2")),
                no=n)
        pdf(paste0("plots/10_Spatial_Registration/01a-Yao23_markerRegistration_Xeniumclus_Heatmaps/Heatmap_",n,".pdf"),height = 0.2*ncol(m),width=10)
        print(plt)
        dev.off()
})
```

# subclass plot as line plots with x = number of top genes
```{r}
plotlist2 <- rbind(rbindlist(yaoreg[[1]],idcol="ntop"),rbindlist(yaoreg[[3]],idcol="ntop"),fill=T)

plotlist2[,ntop:=as.numeric(gsub(ntop,pattern="^.*_top_(.*)_sp.*$",replacement="\\1"))]
plotlist3 <- unique(rbind(
   melt(plotlist2[!is.na(VMHall)],id.vars=c("sphyp_domain","ntop")),
   melt(plotlist2[!is.na(ARHall)],id.vars=c("sphyp_domain","ntop")),
   melt(plotlist2[is.na(VMHall)&is.na(ARHall)],id.vars=c("sphyp_domain","ntop"))))

bplapply(X=unique(plotlist2$sphyp_domain),BPPARAM = sbp,FUN = function(m){
        plttab <- as.data.table(plotlist2[sphyp_domain==m],keep.rownames=T)
        plttab <- melt(plttab,id.vars=c("sphyp_domain","ntop"))
        plttab[,ntop:=as.numeric(ntop)]
        plttab <- plttab[!is.na(value)]
        plttab <- merge.data.table(plttab,bkanno,by.x="sphyp_domain",by.y="clus")
        plt <- ggplot(plttab,aes(x=ntop,y=value))+
                geom_line(color="blue")+
                geom_hline(yintercept = 0,color="black",linetype = "dashed")+
                facet_wrap(.~variable)+
                ggtitle(paste0("Xenium Human ",unique(plttab$bjm_annot), " to\nYao 2023 Mouse Subclasses"))+
                theme(axis.text.x = element_text(angle=90,size=9,hjust=1),axis.text.y=element_text(size=9),title = element_text(size=12),strip.text=element_text(size=7))
        pdf(paste0("plots/10_Spatial_Registration/01b-Yao23_markerRegistration_Xeniumclus_Lineplots/Lineplot_Xen-",bkanno[clus==m,bjm_annot],"_to_Yao23Subclasses.pdf"),height = 25,width=25)
        print(plt)
        dev.off()
})

```



### Lineplots showing all VMH subclasses in one plot for each xenium VMH cluster
```{r}
pltsuper <- rbindlist(yaoreg[[2]],idcol="ntop")
pltsuper[,ntop:=as.numeric(gsub(ntop,pattern="^.*_top_(.*)_sp.*$",replacement="\\1"))]

pltsuper <- merge.data.table(pltsuper,bkanno,by.x="sphyp_domain",by.y="clus")

## get VMH supertypes and VMH xenium clusters, melt, plot
vnames <- c("ntop","bjm_annot",grep(names(pltsuper),pattern="VMH",value=T))
vmhsuper <- pltsuper[bjm_annot %in% grep(bjm_annot,pattern="VMH",value=T),..vnames]
vmhsuper <- melt(vmhsuper,id.vars=c("bjm_annot","ntop"))

## same for arc
anames <- c("ntop","bjm_annot",grep(names(pltsuper),pattern="ARH",value=T))
arcsuper <- pltsuper[bjm_annot %in% grep(bjm_annot,pattern="ARC",value=T),..anames]
arcsuper <- melt(arcsuper,id.vars=c("bjm_annot","ntop"))

## VMH plot set:
pdf("plots/10_Spatial_Registration/01c-VMH_to_YaoVMHsupertypes.pdf",height=10,width=6)
ggplot(vmhsuper,aes(x=ntop,y=value,col=variable))+
        geom_line()+
        geom_hline(yintercept = 0,color="black",linetype = "dashed")+
        facet_wrap(.~bjm_annot)+
        labs(color="Yao 23 Mouse\nVMH Supertype")+
        ylab("Spearman correlation of human\n to mouse 1-vs-all marker t stats")+
           xlab("Top n human markers by t stat")+
        ggtitle("Xenium Human VMH to\nYao 2023 Mouse VMH Supertypes")+
        theme(axis.text.x = element_text(angle=90,size=9,hjust=1),axis.text.y=element_text(size=9),title = element_text(size=12),strip.text=element_text(size=7))
dev.off()
## ARC plot set:
pdf("plots/10_Spatial_Registration/01c-ARC_to_YaoARHsupertypes.pdf",height=14,width=9)
ggplot(arcsuper,aes(x=ntop,y=value,col=variable))+
        geom_line()+
        geom_hline(yintercept = 0,color="black",linetype = "dashed")+
        facet_wrap(.~bjm_annot)+
        labs(color="Yao 23 Mouse\nARC Supertype")+
        ylab("Spearman correlation of human\n to mouse 1-vs-all marker t stats")+
        xlab("Top n human markers by t stat")+
        guides(color=guide_legend(ncol=2))+
        ggtitle("Xenium Human ARC to\nYao 2023 Mouse ARC Supertypes")+
        theme(axis.text.x = element_text(angle=90,size=9,hjust=1),axis.text.y=element_text(size=9),title = element_text(size=12),strip.text=element_text(size=7),legend.position = "bottom")
dev.off()
```

rep inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] gridExtra_2.3       ggplot2_3.5.1       parallelly_1.38.0   orthogene_1.10.0   
[5] BiocParallel_1.38.0 magrittr_2.0.3      colorout_1.3-0.2    data.table_1.15.4  
[9] rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.5              babelgene_22.9            xfun_0.45                
 [4] htmlwidgets_1.6.4         rstatix_0.7.2             lattice_0.22-6           
 [7] bitops_1.0-7              vctrs_0.6.5               tools_4.4.1              
[10] generics_0.1.3            yulab.utils_0.1.4         tibble_3.2.1             
[13] fansi_1.0.6               pkgconfig_2.0.3           Matrix_1.7-0             
[16] homologene_1.4.68.19.3.27 ggplotify_0.1.2           lifecycle_1.0.4          
[19] farver_2.1.2              compiler_4.4.1            treeio_1.28.0            
[22] munsell_0.5.1             codetools_0.2-20          carData_3.0-5            
[25] ggtree_3.12.0             gprofiler2_0.2.3          ggfun_0.1.5              
[28] htmltools_0.5.8.1         RCurl_1.98-1.14           yaml_2.3.9               
[31] lazyeval_0.2.2            plotly_4.10.4             pillar_1.9.0             
[34] car_3.1-2                 ggpubr_0.6.0              tidyr_1.3.1              
[37] cachem_1.1.0              grr_0.9.5                 sessioninfo_1.2.2        
[40] abind_1.4-5               nlme_3.1-165              tidyselect_1.2.1         
[43] aplot_0.2.3               digest_0.6.36             dplyr_1.1.4              
[46] purrr_1.0.2               labeling_0.4.3            rprojroot_2.0.4          
[49] fastmap_1.2.0             grid_4.4.1                here_1.0.1               
[52] colorspace_2.1-0          cli_3.6.3                 patchwork_1.2.0          
[55] utf8_1.2.4                broom_1.0.6               ape_5.8                  
[58] withr_3.0.0               scales_1.3.0              backports_1.5.0          
[61] httr_1.4.7                rmarkdown_2.27            ggsignif_0.6.4           
[64] memoise_2.0.1             evaluate_0.24.0           knitr_1.48               
[67] viridisLite_0.4.2         gridGraphics_0.5-1        Rcpp_1.0.13              
[70] glue_1.7.0                tidytree_0.4.6            rstudioapi_0.16.0        
[73] jsonlite_1.8.8            R6_2.5.1                  fs_1.6.4                 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-20
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package      * version        date (UTC) lib source
   abind          1.4-5          2016-07-21 [1] CRAN (R 4.4.0)
   ape            5.8            2024-04-11 [1] CRAN (R 4.4.0)
   aplot          0.2.3          2024-06-17 [1] CRAN (R 4.4.0)
   babelgene      22.9           2022-09-29 [1] CRAN (R 4.4.0)
   backports      1.5.0          2024-05-23 [1] CRAN (R 4.4.0)
   BiocParallel * 1.38.0         2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bitops         1.0-7          2021-04-24 [1] CRAN (R 4.4.0)
   broom          1.0.6          2024-05-17 [1] CRAN (R 4.4.0)
   cachem         1.1.0          2024-05-16 [1] CRAN (R 4.4.0)
   car            3.1-2          2023-03-30 [1] CRAN (R 4.4.0)
   carData        3.0-5          2022-01-06 [1] CRAN (R 4.4.0)
 P cli            3.6.3          2024-06-21 [2] CRAN (R 4.4.0)
   codetools      0.2-20         2024-03-31 [1] CRAN (R 4.4.1)
   colorout     * 1.3-0.2        2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-0          2023-01-23 [1] CRAN (R 4.4.0)
   data.table   * 1.15.4         2024-03-30 [2] CRAN (R 4.4.0)
   digest         0.6.36         2024-06-23 [1] CRAN (R 4.4.0)
   dplyr          1.1.4          2023-11-17 [1] CRAN (R 4.4.0)
   evaluate       0.24.0         2024-06-10 [1] CRAN (R 4.4.0)
   fansi          1.0.6          2023-12-08 [1] CRAN (R 4.4.0)
   farver         2.1.2          2024-05-13 [1] CRAN (R 4.4.0)
   fastmap        1.2.0          2024-05-15 [1] CRAN (R 4.4.0)
   fs             1.6.4          2024-04-25 [1] CRAN (R 4.4.0)
   generics       0.1.3          2022-07-05 [1] CRAN (R 4.4.0)
   ggfun          0.1.5          2024-05-28 [1] CRAN (R 4.4.0)
   ggplot2      * 3.5.1          2024-04-23 [1] CRAN (R 4.4.0)
   ggplotify      0.1.2          2023-08-09 [1] CRAN (R 4.4.0)
   ggpubr         0.6.0          2023-02-10 [1] CRAN (R 4.4.0)
   ggsignif       0.6.4          2022-10-13 [1] CRAN (R 4.4.0)
   ggtree         3.12.0         2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   glue           1.7.0          2024-01-09 [1] CRAN (R 4.4.0)
   gprofiler2     0.2.3          2024-02-23 [1] CRAN (R 4.4.0)
   gridExtra    * 2.3            2017-09-09 [1] CRAN (R 4.4.0)
   gridGraphics   0.5-1          2020-12-13 [1] CRAN (R 4.4.0)
   grr            0.9.5          2016-08-26 [1] CRAN (R 4.4.0)
   gtable         0.3.5          2024-04-22 [1] CRAN (R 4.4.0)
   here           1.0.1          2020-12-13 [1] CRAN (R 4.4.0)
   homologene     1.4.68.19.3.27 2019-03-28 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1        2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets    1.6.4          2023-12-06 [1] CRAN (R 4.4.0)
   httr           1.4.7          2023-08-15 [1] CRAN (R 4.4.0)
   jsonlite       1.8.8          2023-12-04 [1] CRAN (R 4.4.0)
   knitr          1.48           2024-07-07 [1] CRAN (R 4.4.1)
   labeling       0.4.3          2023-08-29 [1] CRAN (R 4.4.0)
   lattice        0.22-6         2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval       0.2.2          2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle      1.0.4          2023-11-07 [1] CRAN (R 4.4.0)
   magrittr     * 2.0.3          2022-03-30 [1] CRAN (R 4.4.0)
   Matrix         1.7-0          2024-04-26 [1] CRAN (R 4.4.1)
   memoise        2.0.1          2021-11-26 [1] CRAN (R 4.4.0)
   munsell        0.5.1          2024-04-01 [1] CRAN (R 4.4.0)
   nlme           3.1-165        2024-06-06 [1] CRAN (R 4.4.0)
   orthogene    * 1.10.0         2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   parallelly   * 1.38.0         2024-07-27 [1] CRAN (R 4.4.0)
   patchwork      1.2.0          2024-01-08 [1] CRAN (R 4.4.0)
   pillar         1.9.0          2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig      2.0.3          2019-09-22 [1] CRAN (R 4.4.0)
   plotly         4.10.4         2024-01-13 [1] CRAN (R 4.4.0)
   purrr          1.0.2          2023-08-10 [1] CRAN (R 4.4.0)
   R6             2.5.1          2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp           1.0.13         2024-07-17 [1] CRAN (R 4.4.0)
   RCurl          1.98-1.14      2024-01-09 [1] CRAN (R 4.4.0)
 P rlang        * 1.1.4          2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27           2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4          2023-11-05 [1] CRAN (R 4.4.0)
   rstatix        0.7.2          2023-02-01 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0         2024-03-24 [1] CRAN (R 4.4.0)
   scales         1.3.0          2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2          2021-12-06 [1] CRAN (R 4.4.0)
   tibble         3.2.1          2023-03-20 [1] CRAN (R 4.4.0)
   tidyr          1.3.1          2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect     1.2.1          2024-03-11 [1] CRAN (R 4.4.0)
   tidytree       0.4.6          2023-12-12 [1] CRAN (R 4.4.0)
   treeio         1.28.0         2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   utf8           1.2.4          2023-10-22 [1] CRAN (R 4.4.0)
   vctrs          0.6.5          2023-12-01 [1] CRAN (R 4.4.0)
   viridisLite    0.4.2          2023-05-02 [1] CRAN (R 4.4.0)
   withr          3.0.0          2024-01-16 [1] CRAN (R 4.4.0)
   xfun           0.45           2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9          2024-07-05 [1] CRAN (R 4.4.0)
   yulab.utils    0.1.4          2024-01-28 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
