---
title: "02-Visium-Xenium Registration Plots"
output: html_document
date: "2024-08-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

read in results
```{r}
reglist <- readRDS("processed-data/10_Spatial_Registration/02-Visium_to_Xenium_spatialLIBD_registration.RDS")

## each list is one visium bayesspace k value; each sublist is the result matrix from a cross-correlation set using the T top genes (among those on Xenium) marking that Visium domain. T was 10,20...300, all overlapping genes, so 31 sublists per list.
## add a column to each of the result matrices with the T value used, then collapse these into one table, and then melt that for plotting of everything from that visium set.


## also, these results somehow come out a little weird -- an extra row for a nonexistent cluster, in place of the 1st cluster, and twice the number of columns there ought to be, with the results for the first xenium column instead of being the results for every extra column 
## we'll deal with the former issue in the lapply below; for the latter, get the colnames (the visium domain and the 34 xenium clusters) we actually want to keep from each table. all the ones we don't want conveniently begin with V34:
keepnames <- grep(names(reglist[[1]][[1]]),pattern="V34",value=T,invert=T)

reglist <- lapply(reglist,FUN=function(x){
    # get the j value for each list, append it as a column before binding all the results together
    js <- gsub(names(x),pattern="^.*_top_(.*)_sph.*$",replacement="\\1")
    x <- mapply(X=x,Y=js,SIMPLIFY=FALSE,FUN=function(X,Y){
        # first, grab the weirdly repeating results for the first xenium cluster, Other_GABAergic_1, which are in V35.V1 (through V35.V35)
        z <- X
        z[,Other_GABAergic_1:=V34.V1]
        # then dump all the BS columns
        z <- z[,..keepnames]
        # find the row where a real visium cluster got no scores at all (this is the row that instead ends up as the weird extra)
        z[,nacols:=rowSums(is.na(z))]
        realdomain <- z[nacols==33,sphyp_domain]
        dropdomain <- grep(z$sphyp_domain,pattern="^X|ARC|VMH",value=T,invert=T)
        z[sphyp_domain==dropdomain,sphyp_domain:=realdomain]
        z <- z[nacols!=33]
        z[,nacols:=NULL]
        z[,ntopgenes:=as.numeric(Y)]
        z <- melt.data.table(z,id.vars=c("sphyp_domain","ntopgenes"))
        return(z)
    })
    m <- rbindlist(x)
    return(m)
})
```


### phew. now, first churn out line plots for each visium domain, of x=ntopgenes, y=correlation, faceted by xenium cluster
```{r}
mapply(l=reglist,n=names(reglist),SIMPLIFY=FALSE,FUN=function(l,n){
    bplapply(unique(l$sphyp_domain),BPPARAM=MulticoreParam(8),FUN=function(d){
        p <- ggplot(l[sphyp_domain==d],aes(x=ntopgenes,y=value,group=variable,color=variable))+
            geom_line()+
            facet_wrap(~variable)+
            labs(title=d)+
            theme(legend.position="none",strip.text.x=element_text(size=9))
        pdf(paste0("plots/10_Spatial_Registration/02a-Xenium_to_",n,"/Visium_",d,".pdf"),width=30,height=30)
        print(p)
        dev.off()
    })
})
```

It looks like where we're getting decent mapping between known areas (VMH, ARC, WM) without roping in accessory cell types (endothelium etc) is approx the top 20 genes for k=15, 30 for k=20, 40 for k=31. So use those respectively for making heatmaps.

```{r}
mapply(l=reglist,n=names(reglist),t=c(20,30,40,20),SIMPLIFY=FALSE,FUN=function(l,n,t){
    # reorder so that ARCs and VMHs are first along X, top along Y
    vislvl <- unique(c(grep(l$sphyp_domain,pattern="ARC",value=T),
                       grep(l$sphyp_domain,pattern="VMH",value=T),
                       grep(l$sphyp_domain,pattern="ARC|VMH",value=T,invert=T)))
    
    xenlvl <- unique(c(grep(l$variable,pattern="ARC",value=T),
                       grep(l$variable,pattern="VMH",value=T),
                       grep(l$variable,pattern="ARC|VMH",value=T,invert=T)))
    
    # make a hard copy of the data table so it doesn't come back changed, otherwise the apply only works once without reloading and running all the above code again. 
    pltdat <- copy(l)
    pltdat[,sphyp_domain:=factor(sphyp_domain,levels=vislvl)]
    pltdat[,variable:=factor(variable,levels=xenlvl)]
    
    # plot
    p <- ggplot(pltdat[ntopgenes==t],aes(x=sphyp_domain,y=variable,fill=value))+
        geom_tile()+
        labs(title=n)+
        scale_fill_continuous(type="viridis")+
        xlab("Visium Domain")+
        ylab("Xenium Cluster")+
        theme(axis.text=element_text(size=8),axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
    # render
    pdf(paste0("plots/10_Spatial_Registration/02b-XenVisRegist_",n,"_top-",t,"-genes.pdf"),width=10,height=10)
    print(p)
    dev.off()
})

```

repro inf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.38.0 parallelly_1.38.0   colorout_1.3-0.2    gridExtra_2.3      
[5] ggplot2_3.5.1       data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.5      dplyr_1.1.4       compiler_4.4.1    Rcpp_1.0.13      
 [5] tidyselect_1.2.1  stringr_1.5.1     parallel_4.4.1    scales_1.3.0     
 [9] yaml_2.3.9        fastmap_1.2.0     here_1.0.1        plyr_1.8.9       
[13] R6_2.5.1          labeling_0.4.3    generics_0.1.3    knitr_1.48       
[17] viridis_0.6.5     tibble_3.2.1      munsell_0.5.1     rprojroot_2.0.4  
[21] pillar_1.9.0      utf8_1.2.4        stringi_1.8.4     xfun_0.45        
[25] viridisLite_0.4.2 cli_3.6.3         withr_3.0.0       magrittr_2.0.3   
[29] digest_0.6.36     grid_4.4.1        rstudioapi_0.16.0 lifecycle_1.0.4  
[33] vctrs_0.6.5       evaluate_0.24.0   glue_1.7.0        farver_2.1.2     
[37] sessioninfo_1.2.2 codetools_0.2-20  fansi_1.0.6       colorspace_2.1-0 
[41] reshape2_1.4.4    rmarkdown_2.27    tools_4.4.1       pkgconfig_2.0.3  
[45] htmltools_0.5.8.1
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-06
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package      * version date (UTC) lib source
   BiocParallel * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table   * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest         0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate       0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi          1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver         2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics       0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggplot2      * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue           1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable         0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling       0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly   * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar         1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   plyr           1.8.9   2023-10-02 [1] CRAN (R 4.4.0)
   R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp           1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   reshape2       1.4.4   2020-04-09 [1] CRAN (R 4.4.0)
 P rlang        * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   stringi        1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
   stringr        1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
   tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8           1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   viridis        0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite    0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
   withr          3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun           0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
