---
title: "01b-Broad ARC domain by knn (first smoothing)"
output: html_document
date: "2024-06-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(FastKNN)
library(caTools)
library(FNN)
library(SpatialExperiment)
library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

## adapting from script (despite filename, is specifically ARC step 1 smoothing): 
../../0_untidied_Xenium_code_preJune2024/xenium_2runs_together/withQC/cellLevel/lambda0_updatedDomain_May14_2024/step3_post_smooth_ARC_arc/step1_knn/postSmooth_ARC_ARC_knn_k50.R

## load SFE, re-add banksy annots
```{r}
hypx <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

bcl <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bcl,2,"banksy_m0l0kg6_res2")
bcl <- DataFrame(bcl,row.names=bcl$rn)
bcl <- bcl[colnames(hypx),]

colData(hypx)$banksy_m0l0kg6_res2 <- bcl$banksy_m0l0kg6_res2

rm(bcl)
```

### adapting yi's code, make a table of x-y coordinates and ARC (yes/no) assignment. only use the five clusters that are exclusively or extend into ARC and are found across the donors (X15, X19, X23, X31, X33). in order to perform smoothing on each sample individually, split the resulting table into a list of tables, one table per sample.

```{r}
# sanity check: which column is which coordinate in spatialcoords, since we can't call them directly as spatialCoords(hypx)$columnname
# colnames(spatialCoords(hypx))
# [1] "sdimx" "sdimy"
hypx$xcoord <- spatialCoords(hypx)[,1]
hypx$ycoord <- spatialCoords(hypx)[,2]

## make a data table of only the essential columns for smoothing and later return of these results to the SPE: namely, the x y coords, ARC status (one of the ARC clusters or not), and unique cell id (==colnames(hypx)) 
xyARC <- as.data.table(colData(hypx),keep.rownames=T)
xyARC[,isARC:=0]
xyARC[banksy_m0l0kg6_res2 %in% c("X15","X19","X23","X31","X33"),isARC:=1]
sum(xyARC$isARC==1)
# 74037 cells total

## subset to a frame with rownames (cell id), sample id (for making samplewise subtables and later for samplewise plotting), x-y coords, and ARC y/n
xyARC <- xyARC[,.(rn,sample_id,xcoord,ycoord,isARC)]

## make samplewise table list
xyARC_list <- split(xyARC,xyARC$sample_id)

## clean up to maximize RAM avail. for bplapply below
rm(xyARC, hypx)
gc(full=T)
```

## step 1 knn smoothing of ARC with k=50
#### note that the data.table-internalized sapply call is not the absolute fastest way to run that step (but still only takes a matter of seconds at scale here).
#### however, this approach is more transparent in that the user can take an entry of xyARC_list into the environment as v and step through the bplapply to see what's being done here, and can see that cells are always matched in the process (as opposed to working with vectors where we don't have cell ids matched up anymore. which always makes me unccomfortable, however rationally/irrationally.)
```{r}
knnreslist <- bplapply(xyARC_list,BPPARAM=MulticoreParam(10),FUN=function(v){
    # get nearest neighbors for each cell in the sample
    knnout <- FNN::get.knn(v[,c("xcoord","ycoord")],k = 50)
    
    # get the row indices of those k cells
    neighborslist <- apply(knnout$nn.index,MARGIN=1,FUN=list)
    # goofily, this ^ returns each entry as a list within a list (e.g., [[1]][[1]], [[2]][[1]], etc.). so fix that:
    neighborslist <- lapply(neighborslist,FUN=unlist)
     # make the list names the cell identifiers as a failsafe against reordering; we will use the list entry names to retrieve the neighbor indices when calculating means below
    names(neighborslist) <- v$rn
    
    # now, get the mean of ARC y/n for the k nearest neighbors indexed for each cell
    # (in knnout, the first list item, nn.index, has v's row indices corresponding to those neighbors) 
    # this value is thus a probability 0-1 that, a given cell is a ARC type given the ARC/nonARC identity of its 50 spatial neighbors
    v[,ARCpred.smooth1.k50:=
          apply(.SD, MARGIN = 1, FUN=function(x){
              mean(v[neighborslist[[x]],isARC])}
              )
      ,.SDcols = "rn"]
    return(v)
})
rm(xyARC_list)

## concatenate back into a single full-dataset table of KNN smoothing predictions based on k=10 nearest cells
xyARCout <- rbindlist(knnreslist)

## save the resulting smoothing predictions as a table of cell ID, coords, and ARCknn smoothing step 1 prediction values
fwrite(xyARCout,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01b-step1smoothing_ARC_knn-k50.txt",sep='\t',quote=F)
```

### reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.37.1             
 [3] colorout_1.3-0.2               SpatialFeatureExperiment_1.6.1
 [5] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [7] SummarizedExperiment_1.34.0    Biobase_2.64.0                
 [9] GenomicRanges_1.56.1           MatrixGenerics_1.16.0         
[11] matrixStats_1.3.0              FNN_1.1.4                     
[13] caTools_1.18.2                 FastKNN_0.0.1                 
[15] Biostrings_2.72.1              GenomeInfoDb_1.40.1           
[17] XVector_0.44.0                 IRanges_2.38.0                
[19] S4Vectors_0.42.0               BiocGenerics_0.50.0           
[21] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
 [1] EBImage_4.46.0            R.utils_2.12.3           
 [3] bitops_1.0-7              fastmap_1.2.0            
 [5] RCurl_1.98-1.14           digest_0.6.36            
 [7] lifecycle_1.0.4           sf_1.0-16                
 [9] statmod_1.5.0             terra_1.7-78             
[11] magrittr_2.0.3            compiler_4.4.1           
[13] tools_4.4.1               yaml_2.3.8               
[15] knitr_1.47                S4Arrays_1.4.1           
[17] dqrng_0.4.1               htmlwidgets_1.6.4        
[19] sp_2.1-4                  classInt_0.4-10          
[21] here_1.0.1                DelayedArray_0.30.1      
[23] abind_1.4-5               KernSmooth_2.23-24       
[25] HDF5Array_1.32.0          R.oo_1.26.0              
[27] grid_4.4.1                beachmat_2.20.0          
[29] e1071_1.7-14              Rhdf5lib_1.26.0          
[31] edgeR_4.2.0               zeallot_0.1.0            
[33] cli_3.6.3                 rmarkdown_2.27           
[35] crayon_1.5.3              rstudioapi_0.16.0        
[37] httr_1.4.7                rjson_0.2.21             
[39] sessioninfo_1.2.2         DelayedMatrixStats_1.26.0
[41] spdep_1.3-5               scuttle_1.14.0           
[43] DBI_1.2.3                 proxy_0.4-27             
[45] rhdf5_2.48.0              zlibbioc_1.50.0          
[47] assertthat_0.2.1          parallel_4.4.1           
[49] s2_1.1.6                  tiff_0.1-12              
[51] boot_1.3-30               pdist_1.2.1              
[53] Matrix_1.7-0              spData_2.3.1             
[55] jsonlite_1.8.8            fftwtools_0.9-11         
[57] BiocNeighbors_1.22.0      jpeg_0.1-10              
[59] magick_2.8.3              locfit_1.5-9.10          
[61] limma_3.60.3              units_0.8-5              
[63] DropletUtils_1.24.0       codetools_0.2-20         
[65] deldir_2.0-4              UCSC.utils_1.0.0         
[67] htmltools_0.5.8.1         rhdf5filters_1.16.0      
[69] GenomeInfoDbData_1.2.12   R6_2.5.1                 
[71] wk_0.9.1                  sparseMatrixStats_1.16.0 
[73] sfheaders_0.4.4           rprojroot_2.0.4          
[75] evaluate_0.24.0           lattice_0.22-6           
[77] R.methodsS3_1.8.2         png_0.1-8                
[79] class_7.3-22              Rcpp_1.0.12              
[81] SparseArray_1.4.8         xfun_0.45                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-27
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   assertthat                 0.2.1     2019-03-21 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings               * 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   caTools                  * 1.18.2    2021-03-28 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   FastKNN                  * 0.0.1     2015-02-12 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   FNN                      * 1.1.4     2024-01-12 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47      2024-05-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pdist                      1.2.1     2022-05-02 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XVector                  * 0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8     2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
