o---
title: "02a-Step2smoothing_knn_VMHk200-ARCk500_yisthresholds_VMHARC_pbulkDE"
output: html_document
date: "2024-06-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(FastKNN)
library(caTools)
library(FNN)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

the following code stems from the following two R scripts: 
• ../../code/0_untidied_Xenium_code_preJune2024/xenium_2runs_together/withQC/cellLevel/lambda0_updatedDomain_May14_2024/step3_post_smooth_vmh_arc/step2_knn/postSmooth_ARC_knn_k500.R
• ../../code/0_untidied_Xenium_code_preJune2024/xenium_2runs_together/withQC/cellLevel/lambda0_updatedDomain_May14_2024/step3_post_smooth_vmh_arc/step2_knn/postSmooth_VMH_knn_k200.R


##### start by reading in the first round KNN results (coordinates, sample and cell ids, and knn predictions). 
```{r}
vmhknn1 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01a-step1smoothing_VMH_knn-k10.txt")

arcknn1 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01b-step1smoothing_arc_knn-k50.txt")

## remove the cluster-based isvmh/isarc labels, since we will be redefining those here based on knn prediction values
vmhknn1[,isVMH:=NULL]
arcknn1[,isARC:=NULL]
```


```{r}
threshes <- seq(0,1,0.1)

# divy up samplewise for sample-level smoothing, remove concatenated tables to free a bit of RAM
vmhsamps <- split(vmhknn1,vmhknn1$sample_id)
arcsamps <- split(arcknn1,arcknn1$sample_id)
rm(arcknn1,vmhknn1)
gc(full=T)

## iterate over sampelwise smoothing with k=200 (VMH) or 500 (ARC) for each smoothing1 threshold for defining preliminary VMH/ARC 
vmharc_dualthreshtestres <- lapply(threshes,FUN=function(t){
    # for each sample at this threshold, define the VMH and ARC cells based on their round 1 KNN smoothing score
    vmhsamps <- lapply(vmhsamps,function(samp){
        samp[,isVMH:=ifelse(vmhpred.smooth1.k10>t,yes=1,no=0)]
    })
    arcsamps <- lapply(arcsamps,function(samp){
        samp[,isARC:=ifelse(arcpred.smooth1.k50>t,yes=1,no=0)]
    })
    
    # within each sample, get the k (200 vmh, 500 ARC) nearest neighbors of each cell. we only care about the first item in get.knn's list output, the indices of neighboring cells
    vmh_knnres <- bplapply(vmhsamps,BPPARAM=MulticoreParam(10), FUN=function(v){
        vknnres <- FNN::get.knn(v[,.(xcoord,ycoord)],k = 200)
        
         # get the row indices of each cells k neighbors in list format
        vneighbs <- apply(vknnres$nn.index,MARGIN=1,FUN=list)
         # goofily, this is returns each entry as a list within a list (e.g., [[1]][[1]], [[2]][[1]], etc.). so fix that:
        vneighbs <- lapply(vneighbs,FUN=unlist)
        
        # make the list names the cell identifiers as a failsafe against reordering; we will use the list entry names to retrieve the neighbor indices when calculating means below
        names(vneighbs) <- v$rn
        # remoove intermediate objs
        rm(vknnres)
        
        # make a column name to go into the vmh table corresponding to this threshold (we'll collapse them altogether after)
        vthreshname <- paste0("vmhsmooth2_k200_thresh",t)
        # make the column and fill it with means of isVMH for a cell's k neighbors 
        v[,newcol:=
          apply(.SD, MARGIN = 1, FUN=function(x){
              mean(v[vneighbs[[x]],isVMH])
          })
          ,.SDcols = "rn"]
        
        setnames(v,"newcol",vthreshname)
        return(v)
    })
    gc(full=T)
    # same procedure for ARC
    
    arc_knnres <- bplapply(arcsamps,BPPARAM=MulticoreParam(10), FUN=function(a){
        aknnres <- FNN::get.knn(a[,.(xcoord,ycoord)],k = 500)
        
         # get the row indices of each cells k neighbors in list format
        aneighbs <- apply(aknnres$nn.index,MARGIN=1,FUN=list)
        # goofily, this is returns each entry as a list within a list (e.g., [[1]][[1]], [[2]][[1]], etc.). so fix that:
        aneighbs <- lapply(aneighbs,FUN=unlist)
        
        # make the list names the cell identifiers as a failsafe against reordering; we will use the list entry names to retrieve the neighbor indices when calculating means below
        names(aneighbs) <- a$rn
        # remoove intermediate obj
        rm(aknnres)
        
        # make a column name to go into the arc table corresponding to this threshold (we'll collapse them altogether after)
        athreshname <- paste0("arcsmooth2_k500_thresh",t)
        # make the column and fill it with means of isarc for a cell's k neighbors 
        a[,newcol:=
          apply(.SD, MARGIN = 1, FUN=function(x){
              mean(a[aneighbs[[x]],isARC])
          })
      ,.SDcols = "rn"]
        
        setnames(a,"newcol",athreshname)
        return(a)
    })
    gc(full=T)
    
    outlist <- list(vmh_knnres,arc_knnres)
    names(outlist) <- paste0(c("VMHk200_smooth2_tSmooth1_","ARCk500_smooth2_tSmooth1_"),t)
    return(outlist)
})

names(vmharc_dualthreshtestres) <- paste0("tSmooth1_",threshes)
```

### Now, collapse each second-level list entry 
into one table of vmh and arc results from smoothing 2 at a threshold. These 2 results are contained within the first-level lists from the above output, and we will now have a list of 13 tables.
Then, merge each of thoseeeeee by cell_id and threshold score to get one unified table of all cell ids, their coordinates (identical in every table output), and scores for all thresholds tested.
```{r}
vmharc_dualthreshtestres <- bplapply(vmharc_dualthreshtestres, BPPARAM=MulticoreParam(10), FUN=function(x){
    vmh <- rbindlist(x[[1]])
    arc <- rbindlist(x[[2]])
    # subset to key names (rn in both for merging, x and y coords from one, vmhsmooth2 result column or arcsmooth2 column)
    keepvnames <- c("rn","xcoord","ycoord",grep(names(vmh),pattern="smooth2",value=T))
    keepanames <- c("rn",grep(names(arc),pattern="smooth2",value=T))
    vmh <- vmh[,..keepvnames]
    arc <- arc[,..keepanames]
    # merge by cell id
    out <- merge(vmh,arc,by="rn")
    return(out)
})
## save this output
### save output from this step
saveRDS(vmharc_dualthreshtestres,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02a-step2knnscoring_of_dual-step1-thresh-series_vmharc_knn_k200_k500.RDS")


### now merge each of these tables together for a unified table of one cell id, one x-y pair, and the VMH and ARC scores for each of the 10 smoothing round 1 thresholds tested
i<-1
for (i in c(1:length(vmharc_dualthreshtestres2))){
    if(i==1){
        allthreshres <- vmharc_dualthreshtestres2[[i]]
    }
    else{
        tmp <- vmharc_dualthreshtestres2[[i]]
        keepnames <- c("rn",grep(names(tmp),pattern="smooth2",value=T))
        tmp <- tmp[,..keepnames]
        allthreshres <- merge.data.table(allthreshres,tmp,by="rn")
        rm(tmp,keepnames)
    }
}
rm(i)

fwrite(allthreshres,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02a-step2knnscoring_of_dual-step1-thresh-series_vmharc_knn_k200_k500_collapsed.txt",sep='\t',quote=F)

### clean up 
```

### clear the workspace except the list-per-round1-thresh object and load spatialLIBD and the SFE for marker analyses
```{r}
rm(vmhsamps,arcsamps,allthreshres,threshes)
gc(full=T)

library(spatialLIBD)
library(SpatialFeatureExperiment)
hypx <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")
```

### test marker enrichment using the VMH/ARC thresholds from Yi's code, and check for how frequently cells pass the respective thresholds for both layers. these cutoffs were
ARC (stage 1, k50): >0.1
ARC (stage 2, k500): >0.3
VMH (stage 1, k10): >0.2
VMH (stage 2, k200): >0.2

```{r}
arc.yi <- vmharc_dualthreshtestres[["tSmooth1_0.1"]]
arc.yi[,isARCsmooth2:=ifelse(arcsmooth2_k500_thresh0.1>0.3, yes="ARC_yi_2stage",no="other")]

vmh.yi <- vmharc_dualthreshtestres[["tSmooth1_0.2"]]
vmh.yi[,isVMHsmooth2:=ifelse(vmhsmooth2_k200_thresh0.2>0.2, yes="VMH_yi_2stage",no="other")]

### how many cells passed each thresholds doing this?
### more importantly, how many cells pass BOTH thresholds doing this, which as far as i can tell was not accounted for in those analyses?
vmharc.yi <- merge.data.table(arc.yi[,.(rn,arcsmooth2_k500_thresh0.1,xcoord,ycoord,isARCsmooth2)],vmh.yi[,.(rn,vmhsmooth2_k200_thresh0.2,isVMHsmooth2)],by="rn")

vmharc.yi[,.N,by=c("isARCsmooth2","isVMHsmooth2")]
## ~148k are ARC, ~110k are VMH; 599851 are other
## and another ~10k overlap (~7% of ARC, ~9% of VMH cells are not actually a specific domain)
```

## for dually-defined cells, assign the domain to that with the greater knn score (which may well not be the best way to do this but is the simplest for testing enrichments on this parameter set).
```{r}
vmharc.yi[isARCsmooth2=="ARC_yi_2stage"&isVMHsmooth2=="VMH_yi_2stage",dualassignedSmooth2:=ifelse(arcsmooth2_k500_thresh0.1>vmhsmooth2_k200_thresh0.2, yes="ARC_yi_2stage",no="VMH_yi_2stage")]

## transfer the other domain assignments to the dualassigned column
vmharc.yi[isARCsmooth2=="ARC_yi_2stage"&isVMHsmooth2=="other",dualassignedSmooth2:=isARCsmooth2]
vmharc.yi[isVMHsmooth2=="VMH_yi_2stage"&isARCsmooth2=="other",dualassignedSmooth2:=isVMHsmooth2]
vmharc.yi[isVMHsmooth2=="other"&isARCsmooth2=="other",dualassignedSmooth2:="other"]

# ### sanity check: make sure the "other" domain is the same size as before, which shouldn't have changed here
# vmharc.yi[,.N,by="dualassignedSmooth2"]
# other 599851 -- same as before. ok.
## also note here that of 11990 passing both final domain definition thresholds, 4562 were assigned to VMH and 7158 were assigned to ARC using the greater knn score.

## clean up
rm(arc.yi,vmh.yi,vmharc_dualthreshtestres)
```
#### other 599851 -- same as before. ok.
#### also note here that of 11990 passing both final domain definition thresholds, 4562 were assigned to VMH and 7158 were assigned to ARC using the greater knn score.

### now, append the assignments using our de-conflicted results of Yi's parameters to the SFE and test head-to-head, FULL domain DE of VMH vs. ARC
```{r} 
vmharc.yi <- DataFrame(vmharc.yi,row.names = vmharc.yi$rn)
vmharc.yi <- vmharc.yi[colnames(hypx),]
colLabels(hypx) <- vmharc.yi$dualassignedSmooth2

## test enrichment
yivmharc.de <- spatialLIBD::registration_wrapper(hypx,var_registration = "label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol")

## extract the VMH-ARC pairwise enrichments from the output list
yivmharc.de <- as.data.table(yivmharc.de[["pairwise"]])
keepnames <- grep(names(yivmharc.de),pattern="VMH",value=T)
keepnames <- c("gene",grep(keepnames,pattern="ARC",value=T))

yivmharc.de <- yivmharc.de[,..keepnames]
## label previously ascertained VMH markers and ARC markers and subset to these
yivmharc.de[gene %in% c("POMC","GHRH","ESR1","TAC3","AGRP","NPY","GAD1","GAD2"),knownmarkerof:="ARC"]
yivmharc.de[gene %in% c("NRGN","ADCYAP1","SLC17A6","NR5A1","LAMP5","FEZF1","ANKRD34B"),knownmarkerof:="VMH"]

yivmharc.de <- yivmharc.de[knownmarkerof %in% c("ARC","VMH")]
## notably, POMC (which is definitely pan-ARC specific, confirmed by RNAscope as well as visium) is rel. weakly enriched (logFC = 2.9) compared to some other ARC subtype markers e.g. GHRH.

## save these DE stats for comparison to other results depending on the plots.
fwrite(yivmharc.de,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02a-VMHARC_domains_yi-thresholds_pseudobulkDE.txt",sep='\t',quote=F)

## clean up
rm(yivmharc.de)
gc()
```

### reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] SpatialFeatureExperiment_1.6.1 spatialLIBD_1.16.2            
 [3] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [5] SummarizedExperiment_1.34.0    Biobase_2.64.0                
 [7] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
 [9] IRanges_2.38.0                 S4Vectors_0.42.0              
[11] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[13] matrixStats_1.3.0              BiocParallel_1.38.0           
[15] parallelly_1.37.1              colorout_1.3-0.2              
[17] FNN_1.1.4                      caTools_1.18.2                
[19] FastKNN_0.0.1                  data.table_1.15.4             
[21] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0            
  [3] bitops_1.0-7              filelock_1.0.3           
  [5] fields_15.2               R.oo_1.26.0              
  [7] tibble_3.2.1              XML_3.99-0.17            
  [9] lifecycle_1.0.4           sf_1.0-16                
 [11] edgeR_4.2.0               doParallel_1.0.17        
 [13] rprojroot_2.0.4           lattice_0.22-6           
 [15] magrittr_2.0.3            sass_0.4.9               
 [17] limma_3.60.3              plotly_4.10.4            
 [19] rmarkdown_2.27            jquerylib_0.1.4          
 [21] yaml_2.3.8                httpuv_1.6.15            
 [23] spam_2.10-0               sp_2.1-4                 
 [25] sessioninfo_1.2.2         cowplot_1.1.3            
 [27] DBI_1.2.3                 RColorBrewer_1.1-3       
 [29] golem_0.4.1               maps_3.4.2               
 [31] abind_1.4-5               zlibbioc_1.50.0          
 [33] R.utils_2.12.3            purrr_1.0.2              
 [35] RCurl_1.98-1.14           rappdirs_0.3.3           
 [37] GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [39] irlba_2.3.5.1             terra_1.7-78             
 [41] units_0.8-5               dqrng_0.4.1              
 [43] DelayedMatrixStats_1.26.0 DropletUtils_1.24.0      
 [45] codetools_0.2-20          DelayedArray_0.30.1      
 [47] DT_0.33                   scuttle_1.14.0           
 [49] tidyselect_1.2.1          UCSC.utils_1.0.0         
 [51] ScaledMatrix_1.12.0       viridis_0.6.5            
 [53] shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [55] GenomicAlignments_1.40.0  jsonlite_1.8.8           
 [57] BiocNeighbors_1.22.0      e1071_1.7-14             
 [59] scater_1.32.0             iterators_1.0.14         
 [61] foreach_1.5.2             tools_4.4.1              
 [63] Rcpp_1.0.12               glue_1.7.0               
 [65] gridExtra_2.3             SparseArray_1.4.8        
 [67] xfun_0.45                 here_1.0.1               
 [69] EBImage_4.46.0            HDF5Array_1.32.0         
 [71] dplyr_1.1.4               BiocManager_1.30.23      
 [73] fastmap_1.2.0             boot_1.3-30              
 [75] rhdf5filters_1.16.0       fansi_1.0.6              
 [77] pdist_1.2.1               spData_2.3.1             
 [79] digest_0.6.36             rsvd_1.0.5               
 [81] R6_2.5.1                  mime_0.12                
 [83] wk_0.9.1                  colorspace_2.1-0         
 [85] jpeg_0.1-10               RSQLite_2.3.7            
 [87] R.methodsS3_1.8.2         config_0.3.2             
 [89] utf8_1.2.4                tidyr_1.3.1              
 [91] generics_0.1.3            class_7.3-22             
 [93] rtracklayer_1.64.0        httr_1.4.7               
 [95] htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [97] spdep_1.3-5               pkgconfig_2.0.3          
 [99] gtable_0.3.5              blob_1.2.4               
[101] XVector_0.44.0            htmltools_0.5.8.1        
[103] dotCall64_1.1-1           fftwtools_0.9-11         
[105] scales_1.3.0              png_0.1-8                
[107] attempt_0.3.1             knitr_1.47               
[109] rstudioapi_0.16.0         rjson_0.2.21             
[111] curl_5.2.1                proxy_0.4-27             
[113] rhdf5_2.48.0              cachem_1.1.0             
[115] KernSmooth_2.23-24        BiocVersion_3.19.1       
[117] parallel_4.4.1            vipor_0.4.7              
[119] AnnotationDbi_1.66.0      s2_1.1.6                 
[121] restfulr_0.0.15           pillar_1.9.0             
[123] grid_4.4.1                vctrs_0.6.5              
[125] promises_1.3.0            BiocSingular_1.20.0      
[127] dbplyr_2.5.0              beachmat_2.20.0          
[129] sfheaders_0.4.4           xtable_1.8-4             
[131] beeswarm_0.4.0            paletteer_1.6.0          
[133] evaluate_0.24.0           zeallot_0.1.0            
[135] magick_2.8.3              cli_3.6.3                
[137] locfit_1.5-9.10           compiler_4.4.1           
[139] Rsamtools_2.20.0          crayon_1.5.3             
[141] classInt_0.4-10           rematch2_2.1.2           
[143] ggbeeswarm_0.7.2          deldir_2.0-4             
[145] viridisLite_0.4.2         assertthat_0.2.1         
[147] munsell_0.5.1             Biostrings_2.72.1        
[149] tiff_0.1-12               lazyeval_0.2.2           
[151] Matrix_1.7-0              ExperimentHub_2.12.0     
[153] benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[155] bit64_4.0.5               Rhdf5lib_1.26.0          
[157] ggplot2_3.5.1             KEGGREST_1.44.1          
[159] statmod_1.5.0             shiny_1.8.1.1            
[161] AnnotationHub_3.12.0      memoise_2.0.1            
[163] bslib_0.7.0               benchmarkmeData_1.0.4    
[165] bit_4.0.5                
> sessioninfo()
Error in sessioninfo() : could not find function "sessioninfo"
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-27
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   assertthat                 0.2.1     2019-03-21 [1] CRAN (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   caTools                  * 1.18.2    2021-03-28 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   FastKNN                  * 0.0.1     2015-02-12 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     15.2      2023-08-17 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   FNN                      * 1.1.4     2024-01-12 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                    3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                  2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47      2024-05-29 [1] CRAN (R 4.4.0)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pdist                      1.2.1     2022-05-02 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.4.0)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8     2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
