---
title: "02b-Benchmark_previousKNNsmooveParams_VMHARC_boundary_DE"
output: html_document
date: "2024-06-29"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(FastKNN)
library(caTools)
library(FNN)
library(sf)
library(SpatialFeatureExperiment)
library(SpatialExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

### test marker enrichment using the VMH/ARC thresholds from Yi's code, specificially in the boundary area of VMH and ARC to test for expected marker DE in the domains where the overlap may be blurry. 

### first, apply yi's cutoffs as before.
ARC (stage 1, k50): >0.1
ARC (stage 2, k500): >0.3
VMH (stage 1, k10): >0.2
VMH (stage 2, k200): >0.2

```{r}
## load the results of the second step KNN smoothings from a variety of step 1 and step 2 cutoffs, extract those results corresp to yi's previous params
vmharc_dualthreshtestres <- readRDS("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02a-step2knnscoring_of_dual-step1-thresh-series_vmharc_knn_k200_k500.RDS")

arc.yi <- vmharc_dualthreshtestres[["tSmooth1_0.1"]]
arc.yi[,isARCsmooth2:=ifelse(arcsmooth2_k500_thresh0.1>0.3, yes="ARC_yi_2stage",no="other")]

vmh.yi <- vmharc_dualthreshtestres[["tSmooth1_0.2"]]
vmh.yi[,isVMHsmooth2:=ifelse(vmhsmooth2_k200_thresh0.2>0.2, yes="VMH_yi_2stage",no="other")]

### how many cells passed each thresholds doing this?
vmharc.yi <- merge.data.table(arc.yi[,.(rn,arcsmooth2_k500_thresh0.1,xcoord,ycoord,isARCsmooth2)],vmh.yi[,.(rn,vmhsmooth2_k200_thresh0.2,isVMHsmooth2)],by="rn")

vmharc.yi[,.N,by=c("isARCsmooth2","isVMHsmooth2")]
## ~148k are ARC, ~110k are VMH; 599851 are other
## and another ~10k overlap (~7% of ARC, ~9% of VMH cells are not actually a specific domain)
```

## as in script 2a: for dually-defined cells, assign the domain to that with the greater knn score (which may well not be the best way to do this but is the simplest for testing enrichments on this parameter set).
```{r}
vmharc.yi[isARCsmooth2=="ARC_yi_2stage"&isVMHsmooth2=="VMH_yi_2stage",dualassignedSmooth2:=ifelse(arcsmooth2_k500_thresh0.1>vmhsmooth2_k200_thresh0.2, yes="ARC_yi_2stage",no="VMH_yi_2stage")]

## transfer the other domain assignments to the dualassigned column
vmharc.yi[isARCsmooth2=="ARC_yi_2stage"&isVMHsmooth2=="other",dualassignedSmooth2:=isARCsmooth2]
vmharc.yi[isVMHsmooth2=="VMH_yi_2stage"&isARCsmooth2=="other",dualassignedSmooth2:=isVMHsmooth2]
vmharc.yi[isVMHsmooth2=="other"&isARCsmooth2=="other",dualassignedSmooth2:="other"]

# ### sanity check: make sure the "other" domain is the same size as before, which shouldn't have changed here
# vmharc.yi[,.N,by="dualassignedSmooth2"]
# other 599851 -- same as before. ok.
## also note here that of 11990 passing both final domain definition thresholds, 4562 were assigned to VMH and 7158 were assigned to ARC using the greater knn score.

## clean up
rm(arc.yi,vmh.yi,vmharc_dualthreshtestres)
```
#### other 599851 -- same as before. ok.


## extracting the ARC - VMH boundary regions 

#### github copilot explains that functions in the package `sf` can be used to extend the boundary of two polygonal areas by first creating a buffer zone around each, then finding the overlap OF those buffer zones. let's try that here.
```{r}
hypx <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

## make a samplewise list of VMH, ARC domains and define their convex hulls at yi's cutoffs
## we'll need to extract the sample ids from the cell identifiers first
vmharc.yi[,sample_id:=gsub(rn,pattern="^(.*)_[[:alpha:]]{5,20}-.$",replacement="\\1")]
# sanity check
stopifnot(length(unique(vmharc.yi$sample_id))==length(unique(hypx$sample_id)))
# split
vmharc.yi.samplist <- split(vmharc.yi,vmharc.yi$sample_id)

## divy each sample into VMH and ARC subtables for defining their respective polygons; extract convex hulls, convert to sf objects and create additional buffer zones, then find their overlaps (which define the approx ARC-VMH boundary region) and extract the VMH/ARC domain cells in that boundary region for downstream enrichment testing 
vmharc.yi.boundarycells <- bplapply(vmharc.yi.samplist,BPPARAM=MulticoreParam(10),FUN=function(s){
    vmh <- s[dualassignedSmooth2=="VMH_yi_2stage"]
    arc <- s[dualassignedSmooth2=="ARC_yi_2stage"]

    ## get the points defining the convex hull using sf by running st_union to make a single geometry of all component points, then st_convex_hull to get the hull
    vmh.sf <- st_as_sf(vmh,coords=c("xcoord","ycoord"))
    vmh.hull <- st_convex_hull(st_union(vmh.sf))
    
    ## arc is not necessarily a CONVEX hull, however. e.g, it's concave where it meets the VMH (in some cases). again, using X86_reg1, calling st_concave_hull with ratio 0.1 (concavity vs convexity parameter, lower = more concavity; 0 just gives a polygon tracing all points sequentially and is not useful) works beautifully, and this is probably the scoop-iest shaped ARC in the dataset. so let's try that for all samples
    arc.sf <- st_as_sf(arc,coords=c("xcoord","ycoord"))
    arc.hull <- st_concave_hull(st_union(arc.sf),ratio=0.1)

    ## add a buffer zone the domain hull polygons. our coordinates span about 10000 unitless values in the x and y planes for each tissue section, so a buffer adding about 4% to the hull should be good. we can approximate this by calculating the median of the pairwise distances between points defining the domain hull
    
    vmh.buffdist <- median(dist(st_coordinates(vmh.hull)))
    vmh.buffhull <- st_buffer(vmh.hull,dist=0.04*vmh.buffdist)

    arc.buffdist <- median(dist(st_coordinates(arc.hull)))
    arc.buffhull <- st_buffer(arc.hull,dist=0.04*arc.buffdist)

    ## then fucking finally, get the polygon defining where these buffered hulls that overlap, which should thus encompass the vicinity of where a "true" VMH and ARC boundary lies. extract the portions of the overlapping buffers that were assigned ARC or VMH domains by the original smoothing. these are what we want to run DE on next. 
    vmharc.buffintsec <- st_intersection(vmh.buffhull,arc.buffhull)

    buffintsec.invmh <- st_intersection(vmharc.buffintsec,vmh.hull)
    buffintsec.inarc <- st_intersection(vmharc.buffintsec,arc.hull)

    ## really, truly finally, get the cells that fall within this VMH-arc boundary polygon in each respective domain.
    vmh.inboundary <- st_within(vmh.sf,buffintsec.invmh,sparse = F)
    vmh.boundarycells <- as.data.frame(vmh)[vmh.inboundary,"rn"]
    
    arc.inboundary <- st_within(arc.sf,buffintsec.inarc,sparse=F)
    arc.boundarycells <- as.data.frame(arc)[arc.inboundary,"rn"]
    return(list(vmhboundcells=vmh.boundarycells,arcboundcells=arc.boundarycells))
})

```

load spatialLIBD
```{r}
library(spatialLIBD)
```

### extract the cells from the above result, subset the SFE to those cells, label them according to which domain they're assigned within the boundary region, and test DE between VMH-boundary and ARC-boundary.
```{r}
vmh.boundaryids <- unlist(lapply(vmharc.yi.boundarycells,FUN=function(x){unlist(x[[1]])}))
arc.boundaryids <- unlist(lapply(vmharc.yi.boundarycells,FUN=function(x){unlist(x[[2]])}))

## to test DE, we need a third cell type in order for registration_wrapper to run, which is way easier than any other option for pseudobulk DE, so just grab a few hundred exatra cells from the SFE and label them as "other" for this test--we're not considering results involving those cells anyhow.
## subset the SFE with some random cells in addition
othercells <- sample(colnames(hypx)[!(colnames(hypx) %in% c(vmh.boundaryids,arc.boundaryids))],size=15000,replace = F)

hypx.boundary <- hypx[,unique(c(vmh.boundaryids,arc.boundaryids,othercells))]

## make domain labels of VMH(boundary), ARC(boundary), or other (throwaway cells)
boundarylabs <- as.data.table(colnames(hypx.boundary))
setnames(boundarylabs,1,"rn")

boundarylabs[rn %in% othercells,domain:="other"]
boundarylabs[rn %in% vmh.boundaryids,domain:="VMHboundary"]
boundarylabs[rn %in% arc.boundaryids,domain:="ARCboundary"]

boundarylabs <- DataFrame(boundarylabs,row.names=boundarylabs$rn)
boundarylabs <- boundarylabs[colnames(hypx.boundary),]

colLabels(hypx.boundary) <- boundarylabs$domain

## run de with registration_wrapper --> grab pairwise results between portions of the boundary region
boundaryde <- spatialLIBD::registration_wrapper(hypx.boundary,var_registration = "label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol")
## clean up 
rm(hypx.boundary,boundarylabs)

boundaryde <- as.data.table(boundaryde[["pairwise"]])
keepnames <- c("gene",grep(names(boundaryde),pattern="boundary",value=T))
keepnames <- grep(keepnames,pattern="other",value=T,invert=T)
boundaryde <- boundaryde[,..keepnames]

## save result
fwrite(boundaryde,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b-VMHARC_domains_yi-threshes_boundaryarea-VMH-ARC-DE.txt",sep='\t',quote=F)
```

### enrichments are as expected in boundary cells for the respective domains, but 1-3 log2FCs lower in magnitude, and in the case of pan-ARC marker POMC, not even significant (FDR 0.10something).
### let's try grabbing the next "slice" of boundary over, not including the above boundary region, and see what that looks like in terms of arc-vmh DE.
#### since we used 0.04*median(hull vertex-hull vertex distances) as the buffer size the first time, we'll use 0.08 this time, then subset to those cells NOT in the first boundary region.

```{r}
vmharc.yi.boundarycells_slice2 <- bplapply(vmharc.yi.samplist,BPPARAM=MulticoreParam(10),FUN=function(s){
    vmh <- s[dualassignedSmooth2=="VMH_yi_2stage"]
    arc <- s[dualassignedSmooth2=="ARC_yi_2stage"]

    ## get the points defining the convex hull using sf by running st_union to make a single geometry of all component points, then st_convex_hull or st_concave_hull (for arc) to get the hull
    vmh.sf <- st_as_sf(vmh,coords=c("xcoord","ycoord"))
    vmh.hull <- st_convex_hull(st_union(vmh.sf))
    arc.sf <- st_as_sf(arc,coords=c("xcoord","ycoord"))
    arc.hull <- st_concave_hull(st_union(arc.sf),ratio=0.1)

    vmh.buffdist <- median(dist(st_coordinates(vmh.hull)))
    vmh.buffhull <- st_buffer(vmh.hull,dist=0.08*vmh.buffdist)

    arc.buffdist <- median(dist(st_coordinates(arc.hull)))
    arc.buffhull <- st_buffer(arc.hull,dist=0.08*arc.buffdist)
    
    ## get the overlapping parts of these buffer regions and their VMH/ARC components
    vmharc.buffintsec <- st_intersection(vmh.buffhull,arc.buffhull)

    buffintsec.invmh <- st_intersection(vmharc.buffintsec,vmh.hull)
    buffintsec.inarc <- st_intersection(vmharc.buffintsec,arc.hull)

    ## cell ids that fall within this VMH-arc boundary polygon in each respective domain
    vmh.inboundary <- st_within(vmh.sf,buffintsec.invmh,sparse = F)
    vmh.boundarycells <- as.data.frame(vmh)[vmh.inboundary,"rn"]
    
    arc.inboundary <- st_within(arc.sf,buffintsec.inarc,sparse=F)
    arc.boundarycells <- as.data.frame(arc)[arc.inboundary,"rn"]
    return(list(vmhboundcells=vmh.boundarycells,arcboundcells=arc.boundarycells))
})
    
### extract the cells into single vectors per domain, remove those in the 4%*median buffer zone to get the second buffer slice

vmh.boundaryids2 <- unlist(lapply(vmharc.yi.boundarycells_slice2,FUN=function(x){unlist(x[[1]])}))
arc.boundaryids2 <- unlist(lapply(vmharc.yi.boundarycells_slice2,FUN=function(x){unlist(x[[2]])}))

vmh.boundaryids2 <- vmh.boundaryids2[!(vmh.boundaryids2 %in% vmh.boundaryids)]
arc.boundaryids2 <- arc.boundaryids2[!(arc.boundaryids2 %in% arc.boundaryids)]

### spot check that this worked by plotting full VMh and ARC + slice 2 areas for a sample

plotcheck <- vmharc.yi.samplist[["X86_reg2"]]
plotcheck <- plotcheck[dualassignedSmooth2 %in% c("ARC_yi_2stage","VMH_yi_2stage")]

plotcheck[rn %in% vmh.boundaryids2,dualassignedSmooth2:="VMHbound2"]
plotcheck[rn %in% arc.boundaryids2,dualassignedSmooth2:="ARCbound2"]

library(ggplot2)
ggplot2::ggplot(plotcheck,aes(x=xcoord,y=ycoord,color=dualassignedSmooth2)) + geom_point(size=0.2) + theme_minimal()
dev.off()
## k cool

## use the othercells from before (to satisfy spatialLIBD requirements only), make table to annotate a subsetted SFE, test DE in the second border slice
## make domain labels of VMH(boundary), ARC(boundary), or other (throwaway cells)
hypx.boundary <- hypx[,colnames(hypx) %in% c(vmh.boundaryids2,arc.boundaryids2,othercells)]

boundarylabs <- as.data.table(colnames(hypx.boundary))
setnames(boundarylabs,1,"rn")

boundarylabs[rn %in% othercells,domain:="other"]
boundarylabs[rn %in% vmh.boundaryids2,domain:="VMHboundary_slice2"]
boundarylabs[rn %in% arc.boundaryids2,domain:="ARCboundary_slice2"]

boundarylabs <- DataFrame(boundarylabs,row.names=boundarylabs$rn)
boundarylabs <- boundarylabs[colnames(hypx.boundary),]

colLabels(hypx.boundary) <- boundarylabs$domain

### run DE
boundary2de <- spatialLIBD::registration_wrapper(hypx.boundary,var_registration = "label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol")

boundary2de <- as.data.table(boundary2de[["pairwise"]])
keepnames <- c("gene",grep(names(boundary2de),pattern="other|gene|ensembl",value=T,invert=T))
boundary2de <- boundary2de[,..keepnames]

## save result
fwrite(boundary2de,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b-VMHARC_domains_yi-threshes_boundaryarea_slice2-VMH-ARC-DE.txt",sep='\t',quote=F)

```

## reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] SpatialFeatureExperiment_1.6.1 spatialLIBD_1.16.2            
 [3] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [5] SummarizedExperiment_1.34.0    Biobase_2.64.0                
 [7] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
 [9] IRanges_2.38.0                 S4Vectors_0.42.0              
[11] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[13] matrixStats_1.3.0              BiocParallel_1.38.0           
[15] parallelly_1.37.1              colorout_1.3-0.2              
[17] FNN_1.1.4                      caTools_1.18.2                
[19] FastKNN_0.0.1                  data.table_1.15.4             
[21] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0            
  [3] bitops_1.0-7              filelock_1.0.3           
  [5] fields_15.2               R.oo_1.26.0              
  [7] tibble_3.2.1              XML_3.99-0.17            
  [9] lifecycle_1.0.4           sf_1.0-16                
 [11] edgeR_4.2.0               doParallel_1.0.17        
 [13] rprojroot_2.0.4           lattice_0.22-6           
 [15] magrittr_2.0.3            sass_0.4.9               
 [17] limma_3.60.3              plotly_4.10.4            
 [19] rmarkdown_2.27            jquerylib_0.1.4          
 [21] yaml_2.3.8                httpuv_1.6.15            
 [23] spam_2.10-0               sp_2.1-4                 
 [25] sessioninfo_1.2.2         cowplot_1.1.3            
 [27] DBI_1.2.3                 RColorBrewer_1.1-3       
 [29] golem_0.4.1               maps_3.4.2               
 [31] abind_1.4-5               zlibbioc_1.50.0          
 [33] R.utils_2.12.3            purrr_1.0.2              
 [35] RCurl_1.98-1.14           rappdirs_0.3.3           
 [37] GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [39] irlba_2.3.5.1             terra_1.7-78             
 [41] units_0.8-5               dqrng_0.4.1              
 [43] DelayedMatrixStats_1.26.0 DropletUtils_1.24.0      
 [45] codetools_0.2-20          DelayedArray_0.30.1      
 [47] DT_0.33                   scuttle_1.14.0           
 [49] tidyselect_1.2.1          UCSC.utils_1.0.0         
 [51] ScaledMatrix_1.12.0       viridis_0.6.5            
 [53] shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [55] GenomicAlignments_1.40.0  jsonlite_1.8.8           
 [57] BiocNeighbors_1.22.0      e1071_1.7-14             
 [59] scater_1.32.0             iterators_1.0.14         
 [61] foreach_1.5.2             tools_4.4.1              
 [63] Rcpp_1.0.12               glue_1.7.0               
 [65] gridExtra_2.3             SparseArray_1.4.8        
 [67] xfun_0.45                 here_1.0.1               
 [69] EBImage_4.46.0            HDF5Array_1.32.0         
 [71] dplyr_1.1.4               BiocManager_1.30.23      
 [73] fastmap_1.2.0             boot_1.3-30              
 [75] rhdf5filters_1.16.0       fansi_1.0.6              
 [77] pdist_1.2.1               spData_2.3.1             
 [79] digest_0.6.36             rsvd_1.0.5               
 [81] R6_2.5.1                  mime_0.12                
 [83] wk_0.9.1                  colorspace_2.1-0         
 [85] jpeg_0.1-10               RSQLite_2.3.7            
 [87] R.methodsS3_1.8.2         config_0.3.2             
 [89] utf8_1.2.4                tidyr_1.3.1              
 [91] generics_0.1.3            class_7.3-22             
 [93] rtracklayer_1.64.0        httr_1.4.7               
 [95] htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [97] spdep_1.3-5               pkgconfig_2.0.3          
 [99] gtable_0.3.5              blob_1.2.4               
[101] XVector_0.44.0            htmltools_0.5.8.1        
[103] dotCall64_1.1-1           fftwtools_0.9-11         
[105] scales_1.3.0              png_0.1-8                
[107] attempt_0.3.1             knitr_1.47               
[109] rstudioapi_0.16.0         rjson_0.2.21             
[111] curl_5.2.1                proxy_0.4-27             
[113] rhdf5_2.48.0              cachem_1.1.0             
[115] KernSmooth_2.23-24        BiocVersion_3.19.1       
[117] parallel_4.4.1            vipor_0.4.7              
[119] AnnotationDbi_1.66.0      s2_1.1.6                 
[121] restfulr_0.0.15           pillar_1.9.0             
[123] grid_4.4.1                vctrs_0.6.5              
[125] promises_1.3.0            BiocSingular_1.20.0      
[127] dbplyr_2.5.0              beachmat_2.20.0          
[129] sfheaders_0.4.4           xtable_1.8-4             
[131] beeswarm_0.4.0            paletteer_1.6.0          
[133] evaluate_0.24.0           zeallot_0.1.0            
[135] magick_2.8.3              cli_3.6.3                
[137] locfit_1.5-9.10           compiler_4.4.1           
[139] Rsamtools_2.20.0          crayon_1.5.3             
[141] classInt_0.4-10           rematch2_2.1.2           
[143] ggbeeswarm_0.7.2          deldir_2.0-4             
[145] viridisLite_0.4.2         assertthat_0.2.1         
[147] munsell_0.5.1             Biostrings_2.72.1        
[149] tiff_0.1-12               lazyeval_0.2.2           
[151] Matrix_1.7-0              ExperimentHub_2.12.0     
[153] benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[155] bit64_4.0.5               Rhdf5lib_1.26.0          
[157] ggplot2_3.5.1             KEGGREST_1.44.1          
[159] statmod_1.5.0             shiny_1.8.1.1            
[161] AnnotationHub_3.12.0      memoise_2.0.1            
[163] bslib_0.7.0               benchmarkmeData_1.0.4    
[165] bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   assertthat                 0.2.1     2019-03-21 [1] CRAN (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   caTools                  * 1.18.2    2021-03-28 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   FastKNN                  * 0.0.1     2015-02-12 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     15.2      2023-08-17 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   FNN                      * 1.1.4     2024-01-12 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                    3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                  2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47      2024-05-29 [1] CRAN (R 4.4.0)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pdist                      1.2.1     2022-05-02 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.4.0)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8     2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
