---
title: "03b-Domain definition from 2step smoothing of ARC w.o periventricular types"
output: html_document
date: "2024-08-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(sf)
library(ggplot2)
library(gridExtra)
library(ggrastr)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##
```

### based on visual inspection of plots 07/03a, smoothing using only the four ventrally-restricted/preferentially located ARC cell types for domain definition, the ideal parameters VISUALLY are for MOST samples (after round 1 smoothing at 0.2 (VMH) or 0.1 (ARC)):
VMH threshold 0.2, ARC threshold 0.5

### there are two donors for which this does not apply in the case of VMH: 
##### Br8741 has holes with VMH thresholds of 0.2, use 0.15 to get a fuller one and then extract convex hull of that --> make the VMH domain for Br8741 samps

##### Br8667 has additional little spots of VMH at 0.2, so remove those.

to make these domain assignments, load in the table of KNN calculations and apply labels based on the above thresholds/exceptions
```{r}
arcvmh4 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03a-Alt_2stepKNN-ARCVMH_ARC-no-X23.txt")

arcvmh4.list <- split(arcvmh4,arcvmh4$sample_id)

## make VMH and ARC assignments to samples suited to the selected thresholds 
arcvmh4ready <- arcvmh4.list[c("X36_5459A","X36_5459B","X86_reg1","X86_reg2","X86_reg3","X97_reg1","X97_reg2","X97_reg3","X99_1225A","X99_1225B")]
arcvmh4ready <- rbindlist(arcvmh4ready)

arcvmh4ready[VMHpred_thresh1_0.2_smooth2.k200>0.2&ARCpred_thresh1_0.1_smooth2.k500<=0.5,dualVMHARC4:="VMH"]
arcvmh4ready[VMHpred_thresh1_0.2_smooth2.k200<=0.2&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:="ARC"]
arcvmh4ready[VMHpred_thresh1_0.2_smooth2.k200>0.2&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:=ifelse(VMHpred_thresh1_0.2_smooth2.k200>ARCpred_thresh1_0.1_smooth2.k500,yes="VMH",no="ARC")]
arcvmh4ready[is.na(dualVMHARC4),dualVMHARC4:="other"]
```

### handling for 8667, the easier one to deal with
```{r}
br8667 <- arcvmh4.list[["X36_8667C"]]
br8667[VMHpred_thresh1_0.2_smooth2.k200>0.2&ARCpred_thresh1_0.1_smooth2.k500<=0.5,dualVMHARC4:="VMH"]
br8667[VMHpred_thresh1_0.2_smooth2.k200<=0.2&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:="ARC"]
br8667[VMHpred_thresh1_0.2_smooth2.k200>0.2&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:=ifelse(VMHpred_thresh1_0.2_smooth2.k200>ARCpred_thresh1_0.1_smooth2.k500,yes="VMH",no="ARC")]
br8667[is.na(dualVMHARC4),dualVMHARC4:="other"]

### find the little extra guys, should be easy enough with a ggplot by coordinates
ggplot(br8667[dualVMHARC4=="VMH"],aes(x=xcoord,y=ycoord))+geom_point(size=0.1)
### 5 bits to remove: x > 51500 & y > 2500 (2 small areas), x > 54000 (one small area), x < 51000 (literally one cell), x<50000 max(y) (literally one cell)

br8667[dualVMHARC4=="VMH"&xcoord>51500&ycoord>2500,dualVMHARC4:="other"]
br8667[dualVMHARC4=="VMH"&xcoord>54000,dualVMHARC4:="other"]
### replot:
ggplot(br8667[dualVMHARC4=="VMH"],aes(x=xcoord,y=ycoord))+geom_point(size=0.1)
### get the singletons
br8667[dualVMHARC4=="VMH"&xcoord<49000,dualVMHARC4:="other"]
br8667[dualVMHARC4=="VMH"&xcoord<50000&ycoord==max(br8667[xcoord<50000&dualVMHARC4=="VMH",ycoord]),dualVMHARC4:="other"]
### replot:
ggplot(br8667[dualVMHARC4=="VMH"],aes(x=xcoord,y=ycoord))+geom_point(size=0.1)

### k excellent
dev.off()


### we can add this to the table of sampels with finished labels now

arcvmh4ready <- rbind(arcvmh4ready,br8667)

rm(br8667)
```

### now for the gnarlier ones, br8741
```{r}
arcvmh4.list <- arcvmh4.list[c("X99_8741C","X99_8741D")]

## first, apply the general labels
arcvmh4.list <- lapply(arcvmh4.list,FUN=function(x){
   x[VMHpred_thresh1_0.2_smooth2.k200>0.15&ARCpred_thresh1_0.1_smooth2.k500<=0.5,dualVMHARC4:="VMH"]
   x[VMHpred_thresh1_0.2_smooth2.k200<=0.15&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:="ARC"]
   x[VMHpred_thresh1_0.2_smooth2.k200>0.15&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:=ifelse(VMHpred_thresh1_0.2_smooth2.k200>ARCpred_thresh1_0.1_smooth2.k500,yes="VMH",no="ARC")]
   x[is.na(dualVMHARC4),dualVMHARC4:="other"]
   return(x)
})

### now get the convex hulls for VMH in each sample, and assign cells falling within that hull to VMH
### we can test this and see how the convex hull closes the shape
### test <- st_as_sf(arcvmh4.list[[1]][dualVMHARC4=="VMH],coords=c("xcoord","ycoord"))
### test2 <- st_convex_hull(st_union(test2))
### plot(test) -- compare that to:
### plot(test2)

arcvmh4.list <- lapply(arcvmh4.list,FUN=function(x){
   ### get the convex hull
   vmh <- st_as_sf(x[dualVMHARC4=="VMH"],coords=c("xcoord","ycoord"))
   vmh <- st_convex_hull(st_union(vmh))
   
   ### get cells within the convex hull, label as vmh 
   fetchnames <- st_within(st_as_sf(x,coords=c("xcoord","ycoord")),vmh)
   names(fetchnames) <- x$rn
   
   x[rn %in% names(unlist(fetchnames)==1),dualVMHARC4:="VMH"]
   
   ### now deconflict labels with ARC without explicitly requiring a VMH threshold (which is what the convex hull is bypassing), i.e. replace anything labeled VMH by the hull but passing the ARC threshold back to ARC
   x[dualVMHARC4=="VMH"&ARCpred_thresh1_0.1_smooth2.k500>0.5,dualVMHARC4:="ARC"]
   ### 
   
   return(x)
})

arcvmh4ready <- rbind(arcvmh4ready,rbindlist(arcvmh4.list))

## final cell totals:
arcvmh4ready[,.N,by="dualVMHARC4"]
# other 624k
# VMH 132k
# ARC 115k

## get rid of some excess columns
arcvmh4ready <- arcvmh4ready[,.(rn,sample_id,xcoord,ycoord,dualVMHARC4)]
```

### check for remaining holes or little pockets outside of VMH/ARC proper
```{r}
pal <- c("VMH"="purple","ARC"="limegreen","other"="white")

plt <- lapply(unique(arcvmh4ready$sample_id),FUN=function(x){
   tmp <- arcvmh4ready[sample_id==x&dualVMHARC4!="other"]
   ggplot(tmp,aes(x=xcoord,y=ycoord,color=dualVMHARC4))+
      ggtitle(x)+
      scale_color_manual(values=pal)+
      geom_hline(yintercept=seq(min(tmp$ycoord),max(tmp$ycoord),by=500),color="black",linetype="dotted",linewidth=0.2)+
      geom_vline(xintercept = seq(min(tmp$xcoord),max(tmp$xcoord),by=500),color="black",linetype="dotted",linewidth=0.2)+
      rasterize(
         geom_point(size=0.1,stroke=0),layers="Points",dpi=900,dev="cairo_png")+
      theme_minimal()+
      theme(plot.background = element_blank(),line = element_blank(),axis.text.x=element_text(size=4,angle=90,hjust=0.5),axis.text.y=element_text(size=4),axis.title=element_text(size=5))
})

pdf("~/checkXenDomainsForStrayCells.pdf",width=10,height=10)
do.call(gridExtra::grid.arrange,c(plt,list(ncol=3)))
dev.off()
```

### visual check...

### X36_5459A: two ARC points at y<0 & x<16500
```{r}
ggplot(arcvmh4ready[sample_id=="X36_5459A"&dualVMHARC4=="ARC"&ycoord<0&xcoord<16500],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() # the only two points at x<16200

arcvmh4ready <- arcvmh4ready[sample_id=="X36_5459A"&dualVMHARC4=="ARC"&xcoord<16200&ycoord<0,dualVMHARC4:="other"]
```


### X36_5459B: two bunches of ARC points at x<32500 & y<0
```{r}
ggplot(arcvmh4ready[sample_id=="X36_5459B"&dualVMHARC4=="ARC"&ycoord<0&xcoord<32500],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() # xcoord<32200 &  ycoord<0

arcvmh4ready <- arcvmh4ready[sample_id=="X36_5459B"&dualVMHARC4=="ARC"&xcoord<32200&ycoord<0,dualVMHARC4:="other"]
```

### X86_reg1: one VMH way beyond y 9000; several VMH points at x>70000 & y < 2750; one ARC point x<66000 & y < 2000
```{r}
ggplot(arcvmh4ready[sample_id=="X86_reg1"&dualVMHARC4=="VMH"&xcoord>70100&ycoord<3000],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() ## these are at x > 70200 & y < 2600

ggplot(arcvmh4ready[sample_id=="X86_reg1"&dualVMHARC4=="ARC"&xcoord<66000&ycoord<2000],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() # x<65400 & y< 1600

arcvmh4ready <- arcvmh4ready[sample_id=="X86_reg1"&dualVMHARC4=="VMH"&xcoord>70200&ycoord<2600,dualVMHARC4:="other"]
arcvmh4ready <- arcvmh4ready[sample_id=="X86_reg1"&dualVMHARC4=="ARC"&xcoord<65400&ycoord<1600,dualVMHARC4:="other"]
arcvmh4ready <- arcvmh4ready[sample_id=="X86_reg1"&dualVMHARC4=="VMH"&ycoord>9000,dualVMHARC4:="other"]
```


### X86_reg2: one ARC point somewhere at x<81000
```{r}
ggplot(arcvmh4ready[sample_id=="X86_reg2"&dualVMHARC4=="ARC"&xcoord<82000],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() # actually three: one at x<80000, another at x<80500 & y < 3000, and a third that needs closer inspection to get coords
ggplot(arcvmh4ready[sample_id=="X86_reg2"&dualVMHARC4=="ARC"&xcoord<82000&ycoord>5000],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() ## third one is x<80150 & y< 5500&y>5000

## make doubly sure here
stopifnot(nrow(arcvmh4ready[sample_id=="X86_reg2"&dualVMHARC4=="ARC"&(xcoord<80000|(xcoord<80500&ycoord<3000)|(xcoord<80150&ycoord>5000&ycoord<5500))])==3)

## do it
arcvmh4ready <- arcvmh4ready[sample_id=="X86_reg2"&dualVMHARC4=="ARC"&(xcoord<80000|(xcoord<80500&ycoord<3000)|(xcoord<80150&ycoord>5000&ycoord<5500)),dualVMHARC4:="other"]
```


### X86_reg3: 4 ARC points somewhere x<96500 & y<3500
```{r}
ggplot(arcvmh4ready[sample_id=="X86_reg3"&dualVMHARC4=="ARC"&xcoord<96500&ycoord<3500],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() ## x<96100 & y < 2600

arcvmh4ready <- arcvmh4ready[sample_id=="X86_reg3"&dualVMHARC4=="ARC"&xcoord<96100&ycoord<2600,dualVMHARC4:="other"]
```

### X97_Reg3: a few ARC stragglers around xcoord<145000&ycoord>1500
```{r}
ggplot(arcvmh4ready[sample_id=="X97_reg3"&dualVMHARC4=="ARC"&xcoord<144300&ycoord>1500],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() ### xcoord < 143800 & ycoord> 1750

arcvmh4ready <- arcvmh4ready[sample_id=="X97_reg3"&dualVMHARC4=="ARC"&xcoord<143800&ycoord>1750,dualVMHARC4:="other"]
```

### X99_8741_C: 1 lousy arc point way out somewhere past xcoord<192000
```{r}
ggplot(arcvmh4ready[sample_id=="X99_8741C"&dualVMHARC4=="ARC"&xcoord<192000],aes(x=xcoord,y=ycoord))+geom_point()
dev.off() ## specifically just above x 191200

arcvmh4ready <- arcvmh4ready[sample_id=="X99_8741C"&dualVMHARC4=="ARC"&xcoord<191300,dualVMHARC4:="other"]
```

### X99_8741_D: 4 ARC points at xcoord < 207250
```{r}
arcvmh4ready <- arcvmh4ready[sample_id=="X99_8741D"&dualVMHARC4=="ARC"&xcoord<207250,dualVMHARC4:="other"]
```

### 1225A: two groups of VMH stragglers past y > 7000
```{r}
arcvmh4ready <- arcvmh4ready[sample_id=="X99_1225A"&dualVMHARC4=="VMH"&ycoord>7000,dualVMHARC4:="other"]
```

### 1225B has a couple super-far out VMH cells at ycoord > 8500, plus a little run of VMH stragglers around x=179500 y 4000. will need a refined plot to find those
```{r}
ggplot(arcvmh4ready[sample_id=="X99_1225B"&dualVMHARC4=="VMH"&xcoord>179500],aes(x=xcoord,y=ycoord))+geom_point()
dev.off()
## specifically x>179850 & y > 3750

arcvmh4ready <- arcvmh4ready[sample_id=="X99_1225B"&dualVMHARC4=="VMH"&(ycoord>8500|(xcoord>179850&ycoord>3750)),dualVMHARC4:="other"]
```

## recheck:

### check for remaining holes or little pockets outside of VMH/ARC proper
```{r}
pal <- c("VMH"="purple","ARC"="limegreen","other"="white")

plt <- lapply(unique(arcvmh4ready$sample_id),FUN=function(x){
   tmp <- arcvmh4ready[sample_id==x&dualVMHARC4!="other"]
   ggplot(tmp,aes(x=xcoord,y=ycoord,color=dualVMHARC4))+
      ggtitle(x)+
      scale_color_manual(values=pal)+
      geom_hline(yintercept=seq(min(tmp$ycoord),max(tmp$ycoord),by=500),color="black",linetype="dotted",linewidth=0.2)+
      geom_vline(xintercept = seq(min(tmp$xcoord),max(tmp$xcoord),by=500),color="black",linetype="dotted",linewidth=0.2)+
      rasterize(
         geom_point(size=0.1,stroke=0),layers="Points",dpi=900,dev="cairo_png")+
      theme_minimal()+
      theme(plot.background = element_blank(),line = element_blank(),axis.text.x=element_text(size=4,angle=90,hjust=0.5),axis.text.y=element_text(size=4),axis.title=element_text(size=5))
})

pdf("~/checkXenDomainsForStrayCells.pdf",width=10,height=10)
do.call(gridExtra::grid.arrange,c(plt,list(ncol=3)))
dev.off()
```

## okay, now we're good.


### save
```{r}
fwrite(arcvmh4ready,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt",sep='\t',quote=F)

```

### reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] gridExtra_2.3     colorout_1.3-0.2  ggrastr_1.0.2     ggplot2_3.5.1    
[5] sf_1.0-16         data.table_1.15.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] gtable_0.3.5       dplyr_1.1.4        compiler_4.4.1     tidyselect_1.2.1  
 [5] Rcpp_1.0.13        ggbeeswarm_0.7.2   png_0.1-8          scales_1.3.0      
 [9] yaml_2.3.9         fastmap_1.2.0      here_1.0.1         R6_2.5.1          
[13] labeling_0.4.3     generics_0.1.3     classInt_0.4-10    knitr_1.48        
[17] tibble_3.2.1       units_0.8-5        munsell_0.5.1      rprojroot_2.0.4   
[21] DBI_1.2.3          pillar_1.9.0       utf8_1.2.4         xfun_0.45         
[25] cli_3.6.3          withr_3.0.0        magrittr_2.0.3     class_7.3-22      
[29] digest_0.6.36      grid_4.4.1         rstudioapi_0.16.0  beeswarm_0.4.0    
[33] lifecycle_1.0.4    vipor_0.4.7        vctrs_0.6.5        KernSmooth_2.23-24
[37] proxy_0.4-27       evaluate_0.24.0    glue_1.7.0         farver_2.1.2      
[41] fansi_1.0.6        e1071_1.7-14       colorspace_2.1-0   rmarkdown_2.27    
[45] tools_4.4.1        pkgconfig_2.0.3    htmltools_0.5.8.1 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package     * version date (UTC) lib source
   beeswarm      0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   class         7.3-22  2023-05-03 [1] CRAN (R 4.4.1)
   classInt      0.4-10  2023-09-05 [1] CRAN (R 4.4.0)
 P cli           3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout    * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace    2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table  * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   DBI           1.2.3   2024-06-02 [1] CRAN (R 4.4.0)
   digest        0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr         1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   e1071         1.7-14  2023-12-06 [1] CRAN (R 4.4.0)
   evaluate      0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi         1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver        2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics      0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggbeeswarm    0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2     * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr     * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   glue          1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra   * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable        0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here          1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools     0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   KernSmooth    2.23-24 2024-05-17 [1] CRAN (R 4.4.1)
   knitr         1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling      0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle     1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr      2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell       0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   pillar        1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig     2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   png           0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
   proxy         0.4-27  2022-06-09 [1] CRAN (R 4.4.0)
   R6            2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp          1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
 P rlang       * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown     2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot     2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi    0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales        1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo   1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   sf          * 1.0-16  2024-03-24 [1] CRAN (R 4.4.0)
   tibble        3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect    1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   units         0.8-5   2023-11-28 [1] CRAN (R 4.4.0)
   utf8          1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs         0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor         0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   withr         3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun          0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml          2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
