---
title: "03c-DE benchmarks for VMH-ARC smoothing with 4 non-periV arc cell types"
output: html_document
date: "2024-07-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialFeatureExperiment)
library(SpatialExperiment)
library(sf)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

### load SFE, apply labels from scripts 03a/03b
```{r}
arcvmh4 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt")

hypb <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

arcvmh4 <- DataFrame(arcvmh4,row.names=arcvmh4$rn)
arcvmh4 <- arcvmh4[colnames(hypb),]
colLabels(hypb) <- arcvmh4$dualVMHARC4
```

### run pseudobulk DE for full domain diffs
```{r}
fulldomde <- registration_wrapper(hypb,var_registration="label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol")

### get ARC-VMH pairwise DE  results
fulldomde <- as.data.table(fulldomde[["pairwise"]])
keepname <- grep(names(fulldomde),pattern="other",value=T,invert=T)
fulldomde <- fulldomde[,..keepname]

## save
fwrite(fulldomde,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03c-ARCVMHdomains_2stepsmooth-4ARCtypes_fullARC-VMH_DE.txt",sep='\t',quote=F)

rm(fulldomde)
```

### run pseudobulk DE on boundary domains as we tried with yi's parameter sets to compare

### the package `sf` can be used to extend the boundary of two polygonal areas by first creating a buffer zone around each, then finding the overlap OF those buffer zones. where that buffer overlaps the defined VMH domain or the defined ARC domain can then be compared with DE analysis to determine how well the expected markers are enriched in the boundary zones of the respective domains.
```{r}
## return to data.table format; make a samplewise list
arcvmh4 <- as.data.table(arcvmh4)
arcvmh4.list <- split(arcvmh4,arcvmh4$sample_id)

## divy each sample into VMH and ARC subtables for defining their respective polygons; extract convex hulls, convert to sf objects and create additional buffer zones, then find their overlaps (which define the approx ARC-VMH boundary region) and extract the VMH/ARC domain cells in that boundary region for downstream enrichment testing 
arcvmh4.boundarycells <- bplapply(arcvmh4.list,BPPARAM=MulticoreParam(10),FUN=function(s){
    vmh <- s[dualVMHARC4=="VMH"]
    arc <- s[dualVMHARC4=="ARC"]

    ## get the points defining the convex hull using sf by running st_union to make a single geometry of all component points, then st_convex_hull to get the hull
    vmh.sf <- st_as_sf(vmh,coords=c("xcoord","ycoord"))
    vmh.hull <- st_convex_hull(st_union(vmh.sf))
    
    ## arc is not necessarily a CONVEX hull, however. e.g, it's concave where it meets the VMH (in some cases). again, using X86_reg1, calling st_concave_hull with ratio 0.1 (concavity vs convexity parameter, lower = more concavity; 0 just gives a polygon tracing all points sequentially and is not useful) works beautifully, and this is probably the scoop-iest shaped ARC in the dataset. so let's try that for all samples
    arc.sf <- st_as_sf(arc,coords=c("xcoord","ycoord"))
    arc.hull <- st_concave_hull(st_union(arc.sf),ratio=0.1)

    ## add a buffer zone the domain hull polygons. our coordinates span about 10000 unitless values in the x and y planes for each tissue section, so a buffer adding about 4% to the hull should be good. we can approximate this by calculating the median of the pairwise distances between points defining the domain hull
    
    vmh.buffdist <- median(dist(st_coordinates(vmh.hull)))
    vmh.buffhull <- st_buffer(vmh.hull,dist=0.04*vmh.buffdist)

    arc.buffdist <- median(dist(st_coordinates(arc.hull)))
    arc.buffhull <- st_buffer(arc.hull,dist=0.04*arc.buffdist)

    ## then fucking finally, get the polygon defining where these buffered hulls that overlap, which should thus encompass the vicinity of where a "true" VMH and ARC boundary lies. extract the portions of the overlapping buffers that were assigned ARC or VMH domains by the original smoothing. these are what we want to run DE on next. 
    vmharc.buffintsec <- st_intersection(vmh.buffhull,arc.buffhull)

    buffintsec.invmh <- st_intersection(vmharc.buffintsec,vmh.hull)
    buffintsec.inarc <- st_intersection(vmharc.buffintsec,arc.hull)

    ## really, truly finally, get the cells that fall within this VMH-arc boundary polygon in each respective domain.
    vmh.inboundary <- st_within(vmh.sf,buffintsec.invmh,sparse = F)
    vmh.boundarycells <- as.data.frame(vmh)[vmh.inboundary,"rn"]
    
    arc.inboundary <- st_within(arc.sf,buffintsec.inarc,sparse=F)
    arc.boundarycells <- as.data.frame(arc)[arc.inboundary,"rn"]
    return(list(vmhboundcells=vmh.boundarycells,arcboundcells=arc.boundarycells))
})

```

extract VMH-boundary and ARC-boundary cells, subset SFE to those, then run DE again (which will thus only analyze the pseudobulked boundary zones of VMH and ARC)
```{r}
### extract the cells from the above result, subset the SFE to those cells, label them according to which domain they're assigned within the boundary region, and test DE between VMH-boundary and ARC-boundary.
vmh.boundaryids <- unlist(lapply(arcvmh4.boundarycells,FUN=function(x){unlist(x[[1]])}))
arc.boundaryids <- unlist(lapply(arcvmh4.boundarycells,FUN=function(x){unlist(x[[2]])}))

bounddoms <- as.data.table(colnames(hypb))
setnames(bounddoms,1,"rn")
bounddoms[,boundarydomain:="other"]
bounddoms[rn %in% vmh.boundaryids,boundarydomain:="VMHboundary"]
bounddoms[rn %in% arc.boundaryids,boundarydomain:="ARCboundary"]

### assign to SFE
bounddoms <- DataFrame(bounddoms,row.names=bounddoms$rn)
colLabels(hypb) <- bounddoms$boundarydomain

### pseudobulk DE
bounddomde <- registration_wrapper(hypb,var_registration="label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol")

### get ARC-VMH pairwise DE  results
bounddomde <- as.data.table(bounddomde[["pairwise"]])
keepname <- grep(names(bounddomde),pattern="other",value=T,invert=T)
bounddomde <- bounddomde[,..keepname]

### save
fwrite(bounddomde,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03c-ARCVMHdomains_2stepsmooth-4ARCtypes_boundaryARC-VMH_DE.txt",sep='\t',quote=F)
```

### reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.37.1             
 [3] colorout_1.3-0.2               spatialLIBD_1.16.2            
 [5] sf_1.0-16                      SpatialExperiment_1.14.0      
 [7] SingleCellExperiment_1.26.0    SummarizedExperiment_1.34.0   
 [9] Biobase_2.64.0                 GenomicRanges_1.56.1          
[11] GenomeInfoDb_1.40.1            IRanges_2.38.0                
[13] S4Vectors_0.42.0               BiocGenerics_0.50.0           
[15] MatrixGenerics_1.16.0          matrixStats_1.3.0             
[17] SpatialFeatureExperiment_1.6.1 data.table_1.15.4             
[19] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0            
  [3] bitops_1.0-7              filelock_1.0.3           
  [5] fields_15.2               tibble_3.2.1             
  [7] R.oo_1.26.0               XML_3.99-0.17            
  [9] lifecycle_1.0.4           edgeR_4.2.0              
 [11] doParallel_1.0.17         rprojroot_2.0.4          
 [13] lattice_0.22-6            magrittr_2.0.3           
 [15] sass_0.4.9                limma_3.60.3             
 [17] plotly_4.10.4             rmarkdown_2.27           
 [19] jquerylib_0.1.4           yaml_2.3.8               
 [21] httpuv_1.6.15             spam_2.10-0              
 [23] sessioninfo_1.2.2         sp_2.1-4                 
 [25] cowplot_1.1.3             DBI_1.2.3                
 [27] RColorBrewer_1.1-3        golem_0.4.1              
 [29] maps_3.4.2                abind_1.4-5              
 [31] zlibbioc_1.50.0           purrr_1.0.2              
 [33] R.utils_2.12.3            RCurl_1.98-1.14          
 [35] rappdirs_0.3.3            GenomeInfoDbData_1.2.12  
 [37] ggrepel_0.9.5             irlba_2.3.5.1            
 [39] terra_1.7-78              units_0.8-5              
 [41] dqrng_0.4.1               DelayedMatrixStats_1.26.0
 [43] codetools_0.2-20          DropletUtils_1.24.0      
 [45] DelayedArray_0.30.1       DT_0.33                  
 [47] scuttle_1.14.0            tidyselect_1.2.1         
 [49] UCSC.utils_1.0.0          viridis_0.6.5            
 [51] ScaledMatrix_1.12.0       shinyWidgets_0.8.6       
 [53] BiocFileCache_2.12.0      GenomicAlignments_1.40.0 
 [55] jsonlite_1.8.8            BiocNeighbors_1.22.0     
 [57] e1071_1.7-14              scater_1.32.0            
 [59] iterators_1.0.14          foreach_1.5.2            
 [61] tools_4.4.1               Rcpp_1.0.12              
 [63] glue_1.7.0                gridExtra_2.3            
 [65] SparseArray_1.4.8         xfun_0.45                
 [67] here_1.0.1                EBImage_4.46.0           
 [69] dplyr_1.1.4               HDF5Array_1.32.0         
 [71] BiocManager_1.30.23       fastmap_1.2.0            
 [73] boot_1.3-30               rhdf5filters_1.16.0      
 [75] fansi_1.0.6               spData_2.3.1             
 [77] rsvd_1.0.5                digest_0.6.36            
 [79] R6_2.5.1                  mime_0.12                
 [81] colorspace_2.1-0          wk_0.9.1                 
 [83] jpeg_0.1-10               RSQLite_2.3.7            
 [85] R.methodsS3_1.8.2         config_0.3.2             
 [87] utf8_1.2.4                tidyr_1.3.1              
 [89] generics_0.1.3            rtracklayer_1.64.0       
 [91] class_7.3-22              httr_1.4.7               
 [93] htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [95] spdep_1.3-5               pkgconfig_2.0.3          
 [97] gtable_0.3.5              blob_1.2.4               
 [99] XVector_0.44.0            htmltools_0.5.8.1        
[101] dotCall64_1.1-1           fftwtools_0.9-11         
[103] scales_1.3.0              png_0.1-8                
[105] attempt_0.3.1             knitr_1.47               
[107] rstudioapi_0.16.0         rjson_0.2.21             
[109] curl_5.2.1                proxy_0.4-27             
[111] cachem_1.1.0              rhdf5_2.48.0             
[113] BiocVersion_3.19.1        KernSmooth_2.23-24       
[115] vipor_0.4.7               parallel_4.4.1           
[117] AnnotationDbi_1.66.0      restfulr_0.0.15          
[119] s2_1.1.6                  pillar_1.9.0             
[121] grid_4.4.1                vctrs_0.6.5              
[123] promises_1.3.0            BiocSingular_1.20.0      
[125] dbplyr_2.5.0              beachmat_2.20.0          
[127] xtable_1.8-4              sfheaders_0.4.4          
[129] beeswarm_0.4.0            paletteer_1.6.0          
[131] evaluate_0.24.0           zeallot_0.1.0            
[133] magick_2.8.3              Rsamtools_2.20.0         
[135] cli_3.6.3                 locfit_1.5-9.10          
[137] compiler_4.4.1            crayon_1.5.3             
[139] classInt_0.4-10           rematch2_2.1.2           
[141] ggbeeswarm_0.7.2          viridisLite_0.4.2        
[143] deldir_2.0-4              munsell_0.5.1            
[145] Biostrings_2.72.1         lazyeval_0.2.2           
[147] tiff_0.1-12               Matrix_1.7-0             
[149] ExperimentHub_2.12.0      benchmarkme_1.0.8        
[151] sparseMatrixStats_1.16.0  bit64_4.0.5              
[153] ggplot2_3.5.1             Rhdf5lib_1.26.0          
[155] KEGGREST_1.44.1           statmod_1.5.0            
[157] shiny_1.8.1.1             AnnotationHub_3.12.0     
[159] memoise_2.0.1             bslib_0.7.0              
[161] benchmarkmeData_1.0.4     bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-05
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     15.2      2023-08-17 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                    3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                  2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47      2024-05-29 [1] CRAN (R 4.4.0)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.4.0)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8     2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
