---
title: "04-Markers_Distinguishing_VMHARC_celltypes"
output: html_document
date: "2024-08-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

```{r}
hypx <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

bkcl <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
bkanno <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

## merge annotations to raw banksy cluster-cell-assignment output; drop sample specific clusters (DISCARD)
bkanno <- unique(bkanno[,.(clus,bjm_annot)])
bkanno <- bkanno[!(bjm_annot %in% grep(bjm_annot, pattern="DISCARD",value=T))]

setnames(bkcl,2,"clus")
bkcl <- merge(bkcl,bkanno,by="clus")

## subset SPE to cells not labeled as donor spec clust
hypx <- hypx[,bkcl$rn]

## carefully, probably excessively, assign annotated cellc lusters to coldata
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
colLabels(hypx) <- bkcl$bjm_annot

## now append the domain definitions
smoove <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt")
smoove <- smoove[rn %in% colnames(hypx)]
smoove <- DataFrame(smoove,row.names=smoove$rn)[colnames(hypx),]
hypx$smoovedom <- smoove$dualVMHARC4

## for computational effish subset to ARC and VMH cells within their respective domains
hyparc <- hypx[,hypx$smoovedom=="ARC"&hypx$label %in% paste0("ARC_",c(1:5))]
hypvmh <- hypx[,hypx$smoovedom=="VMH"&hypx$label %in% paste0("VMH_",c(1:3,"Lateral_border"))]
```


### run the marker analyses two ways: one vs "all" (i.e., other ARC or VMH types in the domain), aand pairwise. conveniently both are wrapped into spatialLIBD registration_wrapper calls
```{r}
arcdomclusmk <- registration_wrapper(sce = hyparc,var_registration = "label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)

vmhdomclusmk <- registration_wrapper(sce = hypvmh,var_registration = "label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol",min_ncells = 20)
```

# ARC: make these into melted tables
## 1vall:
```{r}
arcdomclus1vall <- as.data.table(arcdomclusmk$enrichment)

arcdomclus1vall <- melt(arcdomclus1vall,id.vars=c("ensembl","gene"))
arcdomclus1vall[,stat:=gsub(variable,pattern="^(.*)_ARC.*$",replacement="\\1")]
arcdomclus1vall[,clust:=gsub(variable,pattern="^.*_(ARC.*)$",replacement="\\1")]
arcdomclus1vall.l <- arcdomclus1vall[stat=="logFC"&value>0]
arcdomclus1vall.t <- arcdomclus1vall[stat=="t_stat"]
arcdomclus1vall.p <- arcdomclus1vall[stat=="p_value"]
arcdomclus1vall.f <- arcdomclus1vall[stat=="fdr"]

setnames(arcdomclus1vall.l,"value","logFC")
setnames(arcdomclus1vall.t,"value","t_stat")
setnames(arcdomclus1vall.p,"value","p_value")
setnames(arcdomclus1vall.f,"value","fdr")

arcdomclus1va.out <- merge.data.table(arcdomclus1vall.l[,.(ensembl,gene,clust,logFC)],arcdomclus1vall.t[,.(ensembl,gene,clust,t_stat)],by=c("clust","ensembl","gene"))
arcdomclus1va.out <- merge.data.table(arcdomclus1va.out,arcdomclus1vall.p[,.(ensembl,gene,clust,p_value)],by=c("clust","ensembl","gene"))
arcdomclus1va.out <- merge.data.table(arcdomclus1va.out,arcdomclus1vall.f[,.(ensembl,gene,clust,fdr)],by=c("clust","ensembl","gene"))

fwrite(arcdomclus1va.out,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04a-ARCtypes-inARCdom_1vall_markers.txt",sep='\t',quote=F)

## also save a copy as a supplemental table sheet
fwrite(arcdomclus1va.out,"local_supptab_assembly/ARCtypesOnly-inARCdom_1vall_markers.txt",sep='\t',quote=F)

rm(arcdomclus1vall,arcdomclus1vall.l,arcdomclus1vall.t,arcdomclus1vall.p,arcdomclus1vall.f,arcdomclus1va.out)
```

## pairwise, ARC. note here that some comparisons (e.g. 2 vs 1 enriched) are the negative LFC values of 1 vs 2, so we need to extract things both ways to get comprehensive pairwise results in long format.
```{r}
arcdomcluspair <- as.data.table(arcdomclusmk$pairwise)

arcdomcluspair <- melt(arcdomcluspair,id.vars=c("ensembl","gene"))

arcdomcluspair[,stat:=gsub(variable,pattern="^(.*)_ARC.*$",replacement="\\1")]
arcdomcluspair[,clust1:=gsub(variable,pattern="^.*_(ARC.*)-ARC.*$",replacement="\\1")]
arcdomcluspair[,clust2:=gsub(variable,pattern="^.*_ARC.*-(ARC.*)$",replacement="\\1")]
    
arcdomcluspair.l <- arcdomcluspair[stat=="logFC"]
arcdomcluspair.t <- arcdomcluspair[stat=="t_stat"]
arcdomcluspair.p <- arcdomcluspair[stat=="p_value"]
arcdomcluspair.f <- arcdomcluspair[stat=="fdr"]

setnames(arcdomcluspair.l,"value","logFC")
setnames(arcdomcluspair.t,"value","t_stat")
setnames(arcdomcluspair.p,"value","p_value")
setnames(arcdomcluspair.f,"value","fdr")

arcdomcluspair.out <- merge.data.table(arcdomcluspair.l[,.(ensembl,gene,clust1,clust2,logFC)],arcdomcluspair.t[,.(ensembl,gene,clust1,clust2,t_stat)],by=c("clust1","clust2","ensembl","gene"))
arcdomcluspair.out <- merge.data.table(arcdomcluspair.out,arcdomcluspair.p[,.(ensembl,gene,clust1,clust2,p_value)],by=c("clust1","clust2","ensembl","gene"))
arcdomcluspair.out <- merge.data.table(arcdomcluspair.out,arcdomcluspair.f[,.(ensembl,gene,clust1,clust2,fdr)],by=c("clust1","clust2","ensembl","gene"))

# for FC < 0, flip the cluster assignments and sign of the logFC so that we have the enriched cell types listed consistently
arcdomcluspair.out[logFC<0,clust1.2:=clust2]
arcdomcluspair.out[logFC<0,clust2.2:=clust1]
arcdomcluspair.out[logFC<0,clust1:=clust1.2]
arcdomcluspair.out[logFC<0,clust2:=clust2.2]
arcdomcluspair.out[logFC<0,t_stat:=-t_stat]
arcdomcluspair.out[logFC<0,logFC:=-logFC]

set(arcdomcluspair.out,j=c("clust1.2","clust2.2"),value=NULL)

## for the pairwise, further subset to fdr<0.2 to make this less messy
arcdomcluspair.out <- arcdomcluspair.out[fdr<0.2]
setnames(arcdomcluspair.out,"clust1","enriched_cluster")
setnames(arcdomcluspair.out,"clust2","comparator_cluster")

fwrite(arcdomcluspair.out,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04d-ARCtypes-inARCdom_pairwise_markers.txt",sep='\t',quote=F)

## also save a copy as a supplemental table sheet
fwrite(arcdomcluspair.out,"local_supptab_assembly/ARCtypesOnly-inARCdom_pairwise_markers.txt",sep='\t',quote=F)

rm(arcdomcluspair,arcdomcluspair.l,arcdomcluspair.t,arcdomcluspair.p,arcdomcluspair.f,arcdomcluspair.out)
```

## now for VMH make the melted 1vall and pairwise tables
```{r}
vmhdomclus1vall <- as.data.table(vmhdomclusmk$enrichment)

vmhdomclus1vall <- melt(vmhdomclus1vall,id.vars=c("ensembl","gene"))
vmhdomclus1vall[,stat:=gsub(variable,pattern="^(.*)_VMH.*$",replacement="\\1")]
vmhdomclus1vall[,clust:=gsub(variable,pattern="^.*_(VMH.*)$",replacement="\\1")]

vmhdomclus1vall.l <- vmhdomclus1vall[stat=="logFC"&value>0]
vmhdomclus1vall.t <- vmhdomclus1vall[stat=="t_stat"]
vmhdomclus1vall.p <- vmhdomclus1vall[stat=="p_value"]
vmhdomclus1vall.f <- vmhdomclus1vall[stat=="fdr"]

setnames(vmhdomclus1vall.l,"value","logFC")
setnames(vmhdomclus1vall.t,"value","t_stat")
setnames(vmhdomclus1vall.p,"value","p_value")
setnames(vmhdomclus1vall.f,"value","fdr")

vmhdomclus1va.out <- merge.data.table(vmhdomclus1vall.l[,.(ensembl,gene,clust,logFC)],vmhdomclus1vall.t[,.(ensembl,gene,clust,t_stat)],by=c("clust","ensembl","gene"))
vmhdomclus1va.out <- merge.data.table(vmhdomclus1va.out,vmhdomclus1vall.p[,.(ensembl,gene,clust,p_value)],by=c("clust","ensembl","gene"))
vmhdomclus1va.out <- merge.data.table(vmhdomclus1va.out,vmhdomclus1vall.f[,.(ensembl,gene,clust,fdr)],by=c("clust","ensembl","gene"))

fwrite(vmhdomclus1va.out,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04c-VMHtypes-inVMHdom_1vall_markers.txt",sep='\t',quote=F)

## also save a copy as a supplemental table sheet
fwrite(vmhdomclus1va.out,"local_supptab_assembly/VMHtypesOnly-inVMHdom_1vall_markers.txt",sep='\t',quote=F)

rm(vmhdomclus1vall,vmhdomclus1vall.l,vmhdomclus1vall.t,vmhdomclus1vall.p,vmhdomclus1vall.f,vmhdomclus1va.out)
```

aaand for pairwise
```{r}
vmhdomcluspair <- as.data.table(vmhdomclusmk$pairwise)

vmhdomcluspair <- melt(vmhdomcluspair,id.vars=c("ensembl","gene"))

vmhdomcluspair[,stat:=gsub(variable,pattern="^(.*)_VMH.*$",replacement="\\1")]
vmhdomcluspair[,clust1:=gsub(variable,pattern="^.*_(VMH.*)-VMH.*$",replacement="\\1")]
vmhdomcluspair[,clust2:=gsub(variable,pattern="^.*_VMH.*-(VMH.*)$",replacement="\\1")]
    
vmhdomcluspair.l <- vmhdomcluspair[stat=="logFC"]
vmhdomcluspair.t <- vmhdomcluspair[stat=="t_stat"]
vmhdomcluspair.p <- vmhdomcluspair[stat=="p_value"]
vmhdomcluspair.f <- vmhdomcluspair[stat=="fdr"]

setnames(vmhdomcluspair.l,"value","logFC")
setnames(vmhdomcluspair.t,"value","t_stat")
setnames(vmhdomcluspair.p,"value","p_value")
setnames(vmhdomcluspair.f,"value","fdr")

vmhdomcluspair.out <- merge.data.table(vmhdomcluspair.l[,.(ensembl,gene,clust1,clust2,logFC)],vmhdomcluspair.t[,.(ensembl,gene,clust1,clust2,t_stat)],by=c("clust1","clust2","ensembl","gene"))
vmhdomcluspair.out <- merge.data.table(vmhdomcluspair.out,vmhdomcluspair.p[,.(ensembl,gene,clust1,clust2,p_value)],by=c("clust1","clust2","ensembl","gene"))
vmhdomcluspair.out <- merge.data.table(vmhdomcluspair.out,vmhdomcluspair.f[,.(ensembl,gene,clust1,clust2,fdr)],by=c("clust1","clust2","ensembl","gene"))

# flip the cluster assignments and sign of the logFC so that we have the enriched cell types listed consistently
vmhdomcluspair.out[logFC<0,clust1.2:=clust2]
vmhdomcluspair.out[logFC<0,clust2.2:=clust1]
vmhdomcluspair.out[logFC<0,clust1:=clust1.2]
vmhdomcluspair.out[logFC<0,clust2:=clust2.2]
vmhdomcluspair.out[logFC<0,t_stat:=-t_stat]
vmhdomcluspair.out[logFC<0,logFC:=-logFC]

set(vmhdomcluspair.out,j=c("clust1.2","clust2.2"),value=NULL)

## for the pairwise, further subset to fdr<0.2 to make this less messy
vmhdomcluspair.out <- vmhdomcluspair.out[fdr<0.2]
setnames(vmhdomcluspair.out,"clust1","enriched_cluster")
setnames(vmhdomcluspair.out,"clust2","comparator_cluster")

fwrite(vmhdomcluspair.out,"processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04d-VMHtypes-inVMHdom_pairwise_markers.txt",sep='\t',quote=F)

## also save a copy as a supplemental table sheet
fwrite(vmhdomcluspair.out,"local_supptab_assembly/VMHtypesOnly-inVMHdom_pairwise_markers.txt",sep='\t',quote=F)

rm(vmhdomcluspair,vmhdomcluspair.l,vmhdomcluspair.t,vmhdomcluspair.p,vmhdomcluspair.f,vmhdomcluspair.out)
```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.38.0             
 [3] spatialLIBD_1.16.2             SpatialFeatureExperiment_1.6.1
 [5] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [7] SummarizedExperiment_1.34.0    Biobase_2.64.0                
 [9] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
[11] IRanges_2.38.1                 S4Vectors_0.42.1              
[13] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[15] matrixStats_1.3.0              gridExtra_2.3                 
[17] colorout_1.3-0.2               ggrastr_1.0.2                 
[19] ggplot2_3.5.1                  sf_1.0-16                     
[21] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0             fields_16.2              
  [4] bitops_1.0-7              filelock_1.0.3            tibble_3.2.1             
  [7] R.oo_1.26.0               XML_3.99-0.17             lifecycle_1.0.4          
 [10] doParallel_1.0.17         edgeR_4.2.0               rprojroot_2.0.4          
 [13] lattice_0.22-6            magrittr_2.0.3            sass_0.4.9               
 [16] limma_3.60.3              plotly_4.10.4             rmarkdown_2.27           
 [19] jquerylib_0.1.4           yaml_2.3.9                httpuv_1.6.15            
 [22] spam_2.10-0               sp_2.1-4                  sessioninfo_1.2.2        
 [25] cowplot_1.1.3             DBI_1.2.3                 RColorBrewer_1.1-3       
 [28] golem_0.4.1               maps_3.4.2                abind_1.4-5              
 [31] zlibbioc_1.50.0           purrr_1.0.2               R.utils_2.12.3           
 [34] RCurl_1.98-1.14           rappdirs_0.3.3            GenomeInfoDbData_1.2.12  
 [37] ggrepel_0.9.5             irlba_2.3.5.1             terra_1.7-78             
 [40] units_0.8-5               dqrng_0.4.1               DelayedMatrixStats_1.26.0
 [43] codetools_0.2-20          DropletUtils_1.24.0       DelayedArray_0.30.1      
 [46] DT_0.33                   scuttle_1.14.0            tidyselect_1.2.1         
 [49] UCSC.utils_1.0.0          farver_2.1.2              viridis_0.6.5            
 [52] ScaledMatrix_1.12.0       shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [55] GenomicAlignments_1.40.0  jsonlite_1.8.8            BiocNeighbors_1.22.0     
 [58] e1071_1.7-14              scater_1.32.0             iterators_1.0.14         
 [61] foreach_1.5.2             tools_4.4.1               Rcpp_1.0.13              
 [64] glue_1.7.0                SparseArray_1.4.8         xfun_0.45                
 [67] here_1.0.1                EBImage_4.46.0            dplyr_1.1.4              
 [70] HDF5Array_1.32.0          withr_3.0.0               BiocManager_1.30.23      
 [73] fastmap_1.2.0             boot_1.3-30               rhdf5filters_1.16.0      
 [76] fansi_1.0.6               spData_2.3.1              rsvd_1.0.5               
 [79] digest_0.6.36             R6_2.5.1                  mime_0.12                
 [82] colorspace_2.1-0          wk_0.9.2                  jpeg_0.1-10              
 [85] RSQLite_2.3.7             R.methodsS3_1.8.2         config_0.3.2             
 [88] utf8_1.2.4                tidyr_1.3.1               generics_0.1.3           
 [91] rtracklayer_1.64.0        class_7.3-22              httr_1.4.7               
 [94] htmlwidgets_1.6.4         S4Arrays_1.4.1            spdep_1.3-5              
 [97] pkgconfig_2.0.3           gtable_0.3.5              blob_1.2.4               
[100] XVector_0.44.0            htmltools_0.5.8.1         dotCall64_1.1-1          
[103] fftwtools_0.9-11          scales_1.3.0              attempt_0.3.1            
[106] png_0.1-8                 knitr_1.48                rstudioapi_0.16.0        
[109] rjson_0.2.21              curl_5.2.1                proxy_0.4-27             
[112] cachem_1.1.0              rhdf5_2.48.0              BiocVersion_3.19.1       
[115] KernSmooth_2.23-24        parallel_4.4.1            vipor_0.4.7              
[118] AnnotationDbi_1.66.0      restfulr_0.0.15           s2_1.1.6                 
[121] pillar_1.9.0              grid_4.4.1                vctrs_0.6.5              
[124] promises_1.3.0            BiocSingular_1.20.0       dbplyr_2.5.0             
[127] beachmat_2.20.0           xtable_1.8-4              sfheaders_0.4.4          
[130] beeswarm_0.4.0            paletteer_1.6.0           evaluate_0.24.0          
[133] zeallot_0.1.0             magick_2.8.3              Rsamtools_2.20.0         
[136] cli_3.6.3                 locfit_1.5-9.10           compiler_4.4.1           
[139] crayon_1.5.3              labeling_0.4.3            classInt_0.4-10          
[142] rematch2_2.1.2            ggbeeswarm_0.7.2          viridisLite_0.4.2        
[145] deldir_2.0-4              munsell_0.5.1             Biostrings_2.72.1        
[148] lazyeval_0.2.2            tiff_0.1-12               Matrix_1.7-0             
[151] ExperimentHub_2.12.0      benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[154] bit64_4.0.5               Rhdf5lib_1.26.0           KEGGREST_1.44.1          
[157] statmod_1.5.0             shiny_1.8.1.1             AnnotationHub_3.12.0     
[160] memoise_2.0.1             bslib_0.7.0               benchmarkmeData_1.0.4    
[163] bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
