---
title: "01-Graph-based hvg, svg harmony clusters k9-15-20-31 markers"
author: "Bernie Mulvey"
date: "2023-05-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater) # addPerCellQC
library(BiocParallel)
library(scran)
library(parallel)
# library(scCustomize)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

```{r}
allwalks <- readRDS("analysis/data/H02-clustering/02-walktraps_snn5.10_HVG1020-OR-SVG-1020_cut9.15.20.31.RDS")

hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")
# row names as gene symbols for sanity
rownames(hyp2) <- rowData(hyp2)$gene_name

# pcas <- readRDS("analysis/data/H01-feature_selection/SPEs_sampleBlock-hvgs-10or20pctile_nnsvg10or20pctile_Harmony-PCA-UMAP.RDS")

```

the function findMarkers in scran is used with binomial tests (proportion of cells expressing per cluster; "This is a more stringent test than the default t-tests, and tends to select genes that are easier to interpret and validate experimentally." per osta). we also specify upregulated genes in order to identify positive markers per cluster.

Using the scoreMarkers function, we can run all scoring algorithms and get composite and full stats

```{r}
bpparam <- MulticoreParam(workers=8)
walktrap.markers <- bplapply(allwalks,BPPARAM=bpparam,FUN=function(x){
    colLabels(hyp2) <- factor(x)
    mks <- scoreMarkers(x = hyp2)
})
names(walktrap.markers) <- names(allwalks)

#### cut these to the union of the top 2.5k per ranking metric
walktrap.markers2 <- bplapply(walktrap.markers,BPPARAM = bpparam, FUN=function(x){
    y <- lapply(x,FUN=function(x){
        x <- as.data.table(x,keep.rownames=T)
        x[rank.logFC.cohen<=2500|rank.AUC<=2500|rank.logFC.detected<=2500]
    })
})

# saveRDS(walktrap.markers2,"analysis/data/H03-marker_genes/walktrap-cut-at-9-15-20-31_HVG1020-SVG1020_SNNk5-10_scoremarkers_out.RDS")
rm(walktrap.markers)
gc(full=T)
```

### Extract top 10 rank gene(s) (genes with FDR = minimum FDR for cluster) per cluster per cluster set into plaintext tabular format
```{r}
i<-1
for (i in c(1:length(walktrap.markers2))){
    k<-1
    for (k in c(1:length(walktrap.markers2[[i]]))){
        tmp <- walktrap.markers2[[i]][[k]]
        tmp <- as.data.table(tmp,keep.rownames=T)
        tmp <- tmp[rank.logFC.cohen<=10|rank.AUC<=10|rank.logFC.detected<=10]
        tmp[,cluster_set:=names(walktrap.markers2)[i]]
        tmp[,cluster:=names(walktrap.markers2[[i]])[k]]
        setnames(tmp,"rn","gene")
        if(i==1&k==1){
            walktrap.markers.plt <- tmp
        }
        else{
            walktrap.markers.plt <- rbind(walktrap.markers.plt,tmp)
        }
        rm(tmp)
    }
    rm(k)
}
rm(i)

write.table(walktrap.markers.plt,"local_results_ref/topRanks_markers_hvg1020-svg1020-snnk5-10_cut9-15-20-31.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

### OLD CODE: Extract top FDR gene(s) (genes with FDR = minimum FDR for cluster) per cluster per cluster set into plaintext tabular format

```{r}
# i<-1
# for (i in c(1:length(walktrap.markers))){
#     k<-1
#     for (k in c(1:length(walktrap.markers[[i]]@listData))){
#         tmp <- walktrap.markers[[i]]@listData[[k]]
#         tmp <- rownames(tmp[which(tmp$FDR==min(tmp$FDR)),])
#         tmp2 <- rep(names(walktrap.markers)[i],times=length(tmp))
#         tmp3 <- rep(names(walktrap.markers[[i]]@listData)[k],times=length(tmp))
#         tmpres <- cbind(tmp,tmp2,tmp3)
#         colnames(tmpres) <- c("gene","cluster_set","cluster")
#         if(i==1&k==1){
#             walktrap.markers.plt <- tmpres
#         }
#         else{
#             walktrap.markers.plt <- rbind(walktrap.markers.plt,tmpres)
#         }
#         rm(tmp,tmp2,tmp3,tmpres)
#     }
#     rm(k)
# }
# rm(i)
# 
# walktrap.markers.plt <- as.data.table(as.data.frame(walktrap.markers.plt))
# 
# write.table(walktrap.markers.plt,"local_results_ref/topFDR_markers_hvg1020-svg1020-snnk5-10_cut9-15-20-31.txt",sep='\t',quote=F,row.names=F,col.names=T)
```
