---
title: "02-Plot top 20 markers per cluster for select BSpace-hvg-svg clusterings"
author: "Bernie Mulvey"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(scater)
library(SpatialExperiment)
library(Polychrome)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```


Read in markers and hyp2 spe
```{r}
bs.markers <- readRDS("analysis/data/H03-marker_genes/BSpace_HVG1020-SVG1020_q9-15-20-31_findmarkers_out.RDS")

hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")

# set hyp2 rownames to gene symbols as markers were identified after making this change
rownames(hyp2) <- rowData(hyp2)$gene_name

# and set hyp2 coldata rownames to "key" so there aren't non-unique rowname problems
rownames(colData(hyp2)) <- colData(hyp2)$key

# load clustering assignments, change names to be same format as markers list
bscl <- list.files("analysis/data/H02-clustering/bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out",pattern=".txt",full.names = T)

bscls <- lapply(bscl,fread)
names(bscls) <- gsub(bscl,pattern="^.*9-15-20-31_out\\/BSpace_Harmony_(.*)\\.txt$",replacement="\\1")

bscls <- lapply(bscls,FUN=function(x){
    as.data.frame(x)
    rownames(x) <- x$rn
    x
})
```

# The "select" conditions, based on visual inspection of Harmony PCA/UMAP and/or spotplots of clustering assignments, are the following:

•HVGs 10%ile, q=15 (umap and spotplot)***
•HVGs 10%ile, q=20 (umap)
•HVGs 10%ile q=31 (umap and spotplot)***
•HVGs 20%ile, q=15 (umap)
•HVGs 20%ile, q=20 (umap)
•HVGs 20%ile, q=31 (umap)
•nnsvgs 10%ile q=31 (UMAP and spotplot)***
•nnsvgs 10%ile q=20 (umap)
•svgs 20%ile q=15 (spotplot)
•svgs 20%ile q=31 (spotplot)--distinct subarchitecture of VMH oval not seen with SVGs @ 10%ile

```{r}
mklistnames <- list("hvg10pctile_15","hvg10pctile_20","hvg10pctile_31","hvg20pctile_15","hvg20pctile_20","hvg20pctile_31","nnsvg10pctile_31","nnsvg10pctile_20","nnsvg20pctile_15","nnsvg20pctile_31")

i<-1
for (i in c(1:2)){
    curname <- mklistnames[[i]]
    curmkset <- bs.markers[[curname]]
    y <- bscls[[curname]]
    y <- as.data.frame(y)
    rownames(y) <- y$rn
    # safeguard against reordered rows
    if((sum(rownames(colData(hyp2))==rownames(y)))!=ncol(hyp2)){
        y <- y[rownames(colData(hyp2)),]
    }
    # assign clusters
    colLabels(hyp2) <- factor(y[,2])
    rm(y)
    k<-1
    for (k in c(1:length(curmkset))){
        curmks <- as.data.table(curmkset[[k]],keep.rownames=T)
        setnames(curmks,"rn","gene_name")
        setorder(curmks,FDR)
        pltmks <- curmks[1:20,gene_name]
        plt <- plotExpression(hyp2,features = pltmks,x="label",colour_by = "label")
        png(paste0("plots/H03-marker_genes/BSpace_",curname,"_top20markers/cluster_",names(curmkset)[k],".png"),width=2000,height=800)
        print(plt)
        dev.off()
        rm(plt,pltmks,curmks)
    }
    rm(k,curmkset,curname,y)
}

```
