---
title: "02-Plot top 20 markers per cluster for select BSpace-hvg-svg clusterings"
author: "Bernie Mulvey"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(scater)
library(SpatialExperiment)
library(Polychrome)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```


Read in markers and hyp2 spe
```{r}
bs.markers <- readRDS("analysis/data/spe_053123/H03-marker_genes/02-BSpace_HVG1020-SVG1020_q9-15-20-31_findmarkers_out.RDS")

hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")

# set hyp2 rownames to gene symbols as markers were identified after making this change
rownames(hyp2) <- rowData(hyp2)$gene_name

# and set hyp2 coldata rownames to "key" so there aren't non-unique rowname problems
rownames(colData(hyp2)) <- colData(hyp2)$key

# load clustering assignments, change names to be same format as markers list
bscl <- list.files("analysis/data/spe_053123/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out",pattern=".txt",full.names = T)

bscls <- lapply(bscl,fread)
names(bscls) <- gsub(bscl,pattern="^.*9-15-20-31_out\\/BSpace_Harmony_(.*)\\.txt$",replacement="\\1")

bscls <- lapply(bscls,FUN=function(x){
    as.data.frame(x)
    rownames(x) <- x$rn
    x
})
```

# The "select" conditions, based on visual inspection of Harmony PCA/UMAP and/or spotplots of clustering assignments, are the following:

•HVGs 10%ile, q=15 (umap and spotplot)***
•HVGs 10%ile, q=20 (umap)
•HVGs 10%ile q=31 (umap and spotplot)***
•HVGs 20%ile, q=15 (umap)
•HVGs 20%ile, q=20 (umap)
•HVGs 20%ile, q=31 (umap)
•nnsvgs 10%ile q=31 (UMAP and spotplot)***
•nnsvgs 10%ile q=20 (umap)
•svgs 20%ile q=15 (spotplot)
•svgs 20%ile q=31 (spotplot)--distinct subarchitecture of VMH oval not seen with SVGs @ 10%ile
•svgs 10%ile q=15 (spotplot/allen atlas expectations)

# note 09-09-23: these plots were previously rendered (and now have been removed from the repo). the cluster labels were mixed up relative to the spotplots of the cluster assignments because in this code, was setting collabel(spe) <- factor(clusterresults[,"cluster_assignment_col"]), where the cluster assignment column was just a numeral (and so the levels were not equivalent to the actual numeral assigned a cluster).
```{r}
# just fix nnsvg10pctile_15 for now, since this is the preliminary dataset
# still not working -_-
mklistnames <- list("nnsvg10pctile_15")

i<-1
for (i in c(1:1)){
    curname <- mklistnames[[i]]
    curmkset <- bs.markers[[curname]]
    y <- bscls[[curname]]
    y <- as.data.frame(y)
    rownames(y) <- y$rn
    # safeguard against reordered rows
    if((sum(rownames(colData(hyp2))==rownames(y)))!=ncol(hyp2)){
        y <- y[rownames(colData(hyp2)),]
    }
    colnames(y)[2] <- "cl"
    y$cl <- paste0("X",y$cl)
    # assign clusters
    colLabels(hyp2) <- as.factor(y$cl)
    rm(y)
    k<-1
    for (k in c(1:length(curmkset))){
        curmks <- as.data.table(curmkset[[k]],keep.rownames=T)
        setnames(curmks,"rn","gene_name")
        setorder(curmks,mean.logFC.cohen)
        pltmks <- curmks[1:20,gene_name]
        plt <- plotExpression(hyp2,features = pltmks,x="label",colour_by = "label")
        png(paste0("plots/spe_053123/H03-marker_genes/BSpace_",curname,"/cluster_",names(curmkset)[k],".png"),width=2000,height=800)
        print(plt)
        dev.off()
        rm(plt,pltmks,curmks)
    }
    rm(k,curmkset,curname,y)
}

```
