---
title: "02-nnSVGs of ARC spots only"
author: "Bernard Mulvey"
date: "2023-07-18"
output: html_document
---

```{r setup, include=FALSE}
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(ggspavis)
library(scater) # addPerCellQC
require(colorout)
ColorOut()
library(BiocParallel)
library(scran)
library(parallel)
library(nnSVG)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### Spatially variable genes--nnSVG, per sample
### the workhorse function nnSVG can use parallel processing through BiocParallel. Set up the preferred (i.e., compatible) param. the parallelization is within the nnsvg call, so no need to match cpu # to sample 

#### crop the preprocessed, nnsvg10-k=15 clustered sample V12-D075_B1 to the high-confidence VMH area, then proceed with SVG detection. Handled in plots code and plots output 04 k2 consensus / 04 nnsvgs.

```{r}
hyp2.vmh <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_061323_HarmonyBSpc-q2VMH_rowname-key.RDS")

### crop V12D07-075_B1 first -- see code/00 blah blah .rmd
coordstmp <- as.data.frame(spatialCoords(hyp2.vmh))
rownames(coordstmp) <- colnames(hyp2.vmh)
coordstmp <- as.data.table(coordstmp,keep.rownames=T)
coordstmp <- coordstmp[,diag:=pxl_col_in_fullres+pxl_row_in_fullres]
coordstmp <- coordstmp[!(rn %in% grep(rn,pattern="V12D07-075_B1",value=T)&diag>30000)]

hyp2 <- hyp2.vmh[,colnames(hyp2.vmh)%in%coordstmp$rn]
stopifnot(dim(hyp2)[2]==nrow(coordstmp))

### now cut to ARC spots (clusters 8 and 13) only
bscl <- fread("analysis/data/spe_053123/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/BSpace_Harmony_nnsvg10pctile_15.txt")
bscl <- as.data.frame(bscl)
rownames(bscl) <- bscl$rn
bscl <- DataFrame(bscl)
bscl <- bscl[colnames(hyp2),]
colLabels(hyp2) <- as.character(bscl$BShar_nnsvg10pctile_15)
rm(bscl)

hyp2 <- hyp2[,which(hyp2$label %in% c("8","13"))]

#cleanup before nnsvg
rm(hyp2.vmh)

sbp <- MulticoreParam(10)
register(sbp)

# With vignette as a guide...:
caps <- as.list(unique(colData(hyp2)$sample_id))
res_list2 <- bplapply(caps,FUN=function(x){
    hyp2.samp <- hyp2[,colData(hyp2)$sample_id==x]
    hyp2.samp <- filter_genes(hyp2.samp,filter_genes_ncounts = 3,filter_genes_pcspots = 1.5,filter_mito = F)
    hyp2.samp <- computeLibraryFactors(hyp2.samp)
    hyp2.samp <- logNormCounts(hyp2.samp)
    out <- nnSVG(hyp2.samp,assay_name = "logcounts")
    out
})

for (i in c(1:length(caps))){
    cap <- caps[i]
    message(paste0("Starting sample ",i," of ",length(caps)))
    hyp2.samp <- hyp2[,colData(hyp2)$sample_id==cap]
    # dim(hyp2.samp)
    hyp2.samp <- filter_genes(hyp2.samp,filter_genes_ncounts = 3,filter_genes_pcspots = 1.5,filter_mito = F)
    #recalculate logcounts
    hyp2.samp <- computeLibraryFactors(hyp2.samp)
    hyp2.samp <- logNormCounts(hyp2.samp)
    # main call -- defaults to logcounts but shown for posterity; X if using covariates; n_neighbors defaults to 10, slower but potentially more granular with higher values eg 15, order = don't touch unless you have < 70 spots total; n_threads or BPPARAM for parallelization
  res_list2[[i]] <- nnSVG(hyp2.samp,assay_name = "logcounts",BPPARAM = sbp)
  rm(hyp2.samp,cap)
  gc(full=T)
}

# extract SVG results
res_list3 <- lapply(res_list2,FUN=function(x){
    # this first line renders rownames invalid, so don't carry them along
    rowData(x) <- rowData(x)[order(rowData(x)$rank),]
    colnames(rowData(x))[8:21] <- paste0("nnSVG_",colnames(rowData(x))[8:21])
    as.data.table(as.data.frame(rowData(x)),keep.rownames=F)
})
#store,save, clean up
nnsvg_N3Spt5 <- res_list3
saveRDS(res_list3,"analysis/data/spe_053123/H05-bspace15-nnsvg10_ARC_clusters/02-nnSVG-3-1pt5_ARC8and13-spots-only_postcropD07075B1.RDS")
rm(i,caps,sbp,res_list2,res_list3)

### get mean ranks
i<-1
for (i in c(1:sum(unique(colData(hyp2)$sample_id)%in%names(nnsvg_N3Spt5)))){
  if(i==1){
  nnsvg_ranks <- nnsvg_N3Spt5[[i]][nnSVG_padj<0.05]
  nnsvg_ranks <- nnsvg_ranks[,.(gene_id,gene_name,nnSVG_rank)]
  setnames(nnsvg_ranks,3,names(nnsvg_N3Spt5)[i])
  }
  else{
  tmp <- nnsvg_N3Spt5[[i]][nnSVG_padj<0.05]
  tmp <- tmp[,.(gene_id,gene_name,nnSVG_rank)]
  setnames(tmp,3,names(nnsvg_N3Spt5)[i])
  nnsvg_ranks <- merge.data.table(nnsvg_ranks,tmp,by=c("gene_id","gene_name"),all.x=T,all.y=T,incomparables = NA)
  rm(tmp)
  }
}
rm(i)

nnsvg_ranks[,meanrank:=rowMeans(.SD,na.rm=T),.SDcols=c(3:10)]

nnsvg_ranks[,n_naSamps:=rowSums(is.na(.SD),na.rm=F),.SDcols=c(3:10)]


write.table(nnsvg_ranks,"analysis/data/spe_053123/H05-bspace15-nnsvg10_ARC_clusters/02-meangeneranks_nnSVG_samplewise_n3s1.5_ARCspotsonly_postcropD07075B1.txt",sep='\t',quote=F,row.names=F,col.names=T)





```
