---
title: "Fig5-TFs_sexDE_disease"
author: "Bernie Mulvey"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(escheR)
library(ggplot2)
library(gridExtra)
library(ggtext)
library(Cairo)
library(ggrastr)
library(ggbeeswarm)
library(ggstance) # vertical dodge

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


A: TF-targ enrichments in visium
```{r}
tft <- readRDS("spatial_HYP/processed-data/11-GSEA/02a-Marker TF-target GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")

tft <- tft[c("k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_XARC","k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_XVMH")]
tft <- rbindlist(tft)
tft[spdom=="k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_XARC",spdom:="vARC"]
tft[spdom=="k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_XVMH",spdom:="vVMH"]

pltdat <- tft

## extract TFs of note
pltdat <- pltdat[pathway %in% c(
    "PMC10242764-CCHTS-26-1560_SD1.pdf-5-RXRA_RXRA_Rummagene_transcription_factors",
    "PMC6815046-13073_2019_678_MOESM2_ESM.xlsx-MYT1L-Unnamed_0_MYT1L_Rummagene_transcription_factors",
    "PMC8683635-MGG3-9-e1749-s005.docx-0-NR3C1_NR3C1_Rummagene_transcription_factors",
    "RARA_KD_HUMAN_GSE26298_CREEDSID_GENE_69_DOWN_TF_Perturbations_Followed_by_Expression",
    "NR2F2_KD_HUMAN_GSE47438_CREEDSID_GENE_2217_DOWN_TF_Perturbations_Followed_by_Expression",
    "EGR1_KO_MOUSE_GSE16974_CREEDSID_GENE_553_DOWN_TF_Perturbations_Followed_by_Expression",
    "EGR1_KO_MOUSE_GSE16974_CREEDSID_GENE_553_UP_TF_Perturbations_Followed_by_Expression",
    "PMC7005888-41598_2020_58845_MOESM2_ESM.xlsx-Insulinoma-INSM1_INSM1_Rummagene_transcription_factors",
    "PMC9160255-41467_2022_30710_MOESM5_ESM.xlsx-AR-targeted_Enhancer_Peaks_anno-GeneExpression_Symbol_AR_Rummagene_transcription_factors",
    "PMC4537001-oncotarget-06-13088-s001.pdf-11-Activators_Of_AR_Transcriptional_Activity_AR_Rummagene_transcription_factors")]

## add column for database source
pltdat[pathway=="PMC10242764-CCHTS-26-1560_SD1.pdf-5-RXRA_RXRA_Rummagene_transcription_factors",source:="Rummagene"]
pltdat[pathway=="PMC6815046-13073_2019_678_MOESM2_ESM.xlsx-MYT1L-Unnamed_0_MYT1L_Rummagene_transcription_factors",source:="Rummagene"]
pltdat[pathway=="PMC8683635-MGG3-9-e1749-s005.docx-0-NR3C1_NR3C1_Rummagene_transcription_factors",source:="Rummagene"]
pltdat[pathway=="RARA_KD_HUMAN_GSE26298_CREEDSID_GENE_69_DOWN_TF_Perturbations_Followed_by_Expression",source:="GEO TF Perturbations"]
pltdat[pathway=="NR2F2_KD_HUMAN_GSE47438_CREEDSID_GENE_2217_DOWN_TF_Perturbations_Followed_by_Expression",source:="GEO TF Perturbations"]
pltdat[pathway=="EGR1_KO_MOUSE_GSE16974_CREEDSID_GENE_553_DOWN_TF_Perturbations_Followed_by_Expression",source:="GEO TF Perturbations"]
pltdat[pathway=="EGR1_KO_MOUSE_GSE16974_CREEDSID_GENE_553_UP_TF_Perturbations_Followed_by_Expression",source:="GEO TF Perturbations"]
pltdat[pathway=="PMC7005888-41598_2020_58845_MOESM2_ESM.xlsx-Insulinoma-INSM1_INSM1_Rummagene_transcription_factors",source:="Rummagene"]
pltdat[pathway=="PMC9160255-41467_2022_30710_MOESM5_ESM.xlsx-AR-targeted_Enhancer_Peaks_anno-GeneExpression_Symbol_AR_Rummagene_transcription_factors",source:="Rummagene"]
pltdat[pathway=="PMC4537001-oncotarget-06-13088-s001.pdf-11-Activators_Of_AR_Transcriptional_Activity_AR_Rummagene_transcription_factors",source:="Rummagene"]

## fix source names to just TFs
setnames(pltdat,"pathway","tf")

pltdat[tf=="PMC10242764-CCHTS-26-1560_SD1.pdf-5-RXRA_RXRA_Rummagene_transcription_factors",tf:="RXRA"]
pltdat[tf=="PMC6815046-13073_2019_678_MOESM2_ESM.xlsx-MYT1L-Unnamed_0_MYT1L_Rummagene_transcription_factors",tf:="MYT1L"]
pltdat[tf=="PMC8683635-MGG3-9-e1749-s005.docx-0-NR3C1_NR3C1_Rummagene_transcription_factors",tf:="NR3C1"]
pltdat[tf=="RARA_KD_HUMAN_GSE26298_CREEDSID_GENE_69_DOWN_TF_Perturbations_Followed_by_Expression",tf:="RARA (KD Down)"]
pltdat[tf=="NR2F2_KD_HUMAN_GSE47438_CREEDSID_GENE_2217_DOWN_TF_Perturbations_Followed_by_Expression",tf:="NR2F2 (KD Down)"]
pltdat[tf=="EGR1_KO_MOUSE_GSE16974_CREEDSID_GENE_553_DOWN_TF_Perturbations_Followed_by_Expression",tf:="EGR1 (KO Down)"]
pltdat[tf=="EGR1_KO_MOUSE_GSE16974_CREEDSID_GENE_553_UP_TF_Perturbations_Followed_by_Expression",tf:="EGR1 (KO Up)"]
pltdat[tf=="PMC7005888-41598_2020_58845_MOESM2_ESM.xlsx-Insulinoma-INSM1_INSM1_Rummagene_transcription_factors",tf:="INSM1"]
pltdat[tf=="PMC9160255-41467_2022_30710_MOESM5_ESM.xlsx-AR-targeted_Enhancer_Peaks_anno-GeneExpression_Symbol_AR_Rummagene_transcription_factors",tf:="AR"]
pltdat[tf=="PMC4537001-oncotarget-06-13088-s001.pdf-11-Activators_Of_AR_Transcriptional_Activity_AR_Rummagene_transcription_factors",tf:="AR (Activators)"]

pltdat[source=="GEO TF Perturbations",source:="GEO TF\n Perturbations"]
## make the plot
breaks <- c(seq(-1, 1, 1),seq(1.5,2.5, 0.5))
n_breaks <- length(breaks)
labels <- c(breaks, rep("", n_breaks))
shapes <- c(rep(16,n_breaks),rep(17,n_breaks))
breaks2 <- rep(breaks, 2)
## 
pdf("manuscript_plots/Fig5/5A-marker_TFtargs_GSEA.pdf",height=3,width=3.5)
ggplot(pltdat,aes(y=tf,x=-log10(pval)))+
    geom_point(aes(col=spdom,size=NES,shape=source))+
    ylab("TF")+
    scale_color_manual(values=c("#5dd959","#8c63cf"),na.value = NA)+
    xlab("-log10 (*p*-value)")+
    labs(col="Domain",shape="So`urce Database")+
    geom_vline(xintercept=-log10(tft[padj<0.05,max(pval)]),linetype="dashed",linewidth=0.5)+
    guides(col=guide_legend(byrow=T))+
    # ^ necess. to get spacing to work in theme
    guides(col=guide_legend(override.aes=list(size=2.5)),
           shape=guide_legend(override.aes=list(size = 2.5)))+
    # ^ make legend points bigger
    scale_size_continuous(transform="exp",range=c(0.75,3),limits = c(-2,2.5),
                          breaks = breaks2, 
                          labels = labels,
                          guide = guide_legend(ncol=2, nrow=n_breaks,bycol = TRUE,override.aes = list(shape = shapes), label.vjust = 0.5,direction="vertical",label.hjust=0.05,vjust=0))+
    # ^ show both shapes plotted on the scale
    theme(axis.text.y = element_text(size = 8), axis.title.x = element_markdown(size = 9), axis.text.x = element_text(size = 8), legend.text = element_text(size=8,hjust = 0.5), legend.title = element_text(size=9,hjust=0.5),legend.spacing.y = ggplot2::unit(0.005, "in"),axis.title.y=element_markdown(size=9,margin=margin(0,0,0,0.02,"in")),legend.box.background = element_blank(),legend.margin = margin(0,0,0,-0.135,unit="in"),plot.margin = margin(0,-0.14,0,0.03))
dev.off()

```

5B/5D: sex de GSEA for msigdb and TF targets -- with shared legend and combined output plot.
## define a palette for region-sex
```{r}
sexregpal <- c("vVMH Fem. Up"="#8c63cf","vVMH Male Up"="#6A0D53","vARC Fem. Up"="#5dd959","vARC Male Up"="#053b05")
```

5D- Sex DE msigdb GSEA
```{r}
sexreslist <- readRDS("spatial_HYP/processed-data/11-GSEA/03a-Sex DE msigdb_1,2,3,5,6,8 GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")

vmhs <- sexreslist[["BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_VMH"]][pathway %in% c("HP_ABNORMAL_AGGRESSIVE_IMPULSIVE_OR_VIOLENT_BEHAVIOR", "GOMF_CHROMATIN_BINDING","HP_ABNORMALITY_OF_ACID_BASE_HOMEOSTASIS", "YAO_TEMPORAL_RESPONSE_TO_PROGESTERONE_CLUSTER_17","GABRIELY_MIR21_TARGETS", "MIR30A_3P","MCLACHLAN_DENTAL_CARIES_UP", "FAN_OVARY_CL6_PUTATIVE_EARLY_ATRETIC_FOLLICLE_THECAL_CELL_2")]

arcs <- sexreslist[["BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_ARC"]][pathway %in% c("HP_ABNORMAL_AGGRESSIVE_IMPULSIVE_OR_VIOLENT_BEHAVIOR", "GOMF_CHROMATIN_BINDING","HP_ABNORMALITY_OF_ACID_BASE_HOMEOSTASIS", "YAO_TEMPORAL_RESPONSE_TO_PROGESTERONE_CLUSTER_17","GABRIELY_MIR21_TARGETS", "MIR30A_3P","MCLACHLAN_DENTAL_CARIES_UP", "FAN_OVARY_CL6_PUTATIVE_EARLY_ATRETIC_FOLLICLE_THECAL_CELL_2")]

msigpltdat <- rbind(vmhs,arcs)
msigpltdat[,spdom:=gsub(spdom,pattern="BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_(.*)",replacement="\\1")]
# msigpltdat[spdom=="VMH",sexreg:=ifelse(NES>0,yes="vVMH Male Up",no="vVMH Fem. Up")]
# msigpltdat[spdom=="ARC",sexreg:=ifelse(NES>0,yes="vARC Male Up",no="vARC Fem. Up")]
msigpltdat[spdom=="VMH",sexreg:=ifelse(dir=="male_up",yes="vVMH Male Up",no="vVMH Fem. Up")]
msigpltdat[spdom=="ARC",sexreg:=ifelse(dir=="male_up",yes="vARC Male Up",no="vARC Fem. Up")]
# make all NES for female test w/ enrichment, (i.e. negative NES) positive
msigpltdat[dir=="female_up"&NES<0,NES:=abs(NES)]

rm(vmhs,arcs)
### extract pathway source db

i<-1
for (i in c(1:nrow(msigpltdat))){
    if(i==1){gsrc <- unlist(strsplit(msigpltdat[i,pathway],"_"))[1]}
    else{gsrc <- c(gsrc,
                    unlist(strsplit(msigpltdat[i,pathway],"_"))[1])}
}

## tidy gene sources
msigpltdat$genesource <- gsrc
msigpltdat[genesource %in% c("FAN","GABRIELY","MCLACHLAN","YAO"),genesource:="Literature-Curated"]
msigpltdat[genesource=="GOMF",genesource:="GO Mol. Fxn"]
msigpltdat[genesource=="HP",genesource:="Human Phenotypes"]
msigpltdat[genesource=="MIR30A",genesource:="miRNA Targets"]

rm(i,gsrc)

### simplify pathways
msigpltdat[pathway=="HP_ABNORMAL_AGGRESSIVE_IMPULSIVE_OR_VIOLENT_BEHAVIOR",pathway:="Aggressive/im-\npulsive behavior"]
msigpltdat[pathway=="GOMF_CHROMATIN_BINDING",pathway:="Chromatin binding"]
msigpltdat[pathway=="HP_ABNORMALITY_OF_ACID_BASE_HOMEOSTASIS",pathway:="Acid-base ab-\nnormality"]
msigpltdat[pathway=="YAO_TEMPORAL_RESPONSE_TO_PROGESTERONE_CLUSTER_17",pathway:="Temporal progest-\nerone response"]
msigpltdat[pathway=="GABRIELY_MIR21_TARGETS",pathway:="miR-21"]
msigpltdat[pathway=="MIR30A_3P",pathway:="miR-30a_3P"]
msigpltdat[pathway=="MCLACHLAN_DENTAL_CARIES_UP",pathway:="Dental caries"]
msigpltdat[pathway=="FAN_OVARY_CL6_PUTATIVE_EARLY_ATRETIC_FOLLICLE_THECAL_CELL_2",pathway:="Thecal cell of ovary\n(early atretic follicle)"]

### flip sign on negative NES scores
# msigpltdat[NES<0,NES:=-1*NES]

## change source info slightly so we can have msigdb and non-msigdb stuff in one legend
msigpltdat[genesource=="MSigDB Literature-Curated",source:="Literature-Curated"]
msigpltdat[genesource=="MSigDB Human Phenotypes",source:="Human Phenotypes"]
msigpltdat[genesource=="MSigDB miRNA Targets",source:="miRNA Targets"]


# pause here to get the TF-targ table made so we can make the shared NES scale+legend for it.
```

## 6B - TFs ##
Male both:
MYT1L: PMC6815046-13073_2019_678_MOESM3_ESM.xlsx-MYT1L-Unnamed_0_MYT1L_Rummagene_transcription_factors


VMH ~UNIQUE Male (FDR 10E-8 vs 0.04):
TCF20 (rare LOF NDD gene Vetrini '19): PMC6815046-13073_2019_678_MOESM3_ESM.xlsx-TCF20-Unnamed_0_TCF20_Rummagene_transcription_factors


VMH UNIQUE Fem:
PPARA: PPARA_KO_MOUSE_GSE6864_CREEDSID_GENE_1372_DOWN_TF_Perturbations_Followed_by_Expression, PPARA_KO_MOUSE_GSE6864_CREEDSID_GENE_1373_UP_TF_Perturbations_Followed_by_Expression

PPARG: PPARG_OE_MOUSE_GSE2192_CREEDSID_GENE_861_UP_TF_Perturbations_Followed_by_Expression, PPARG_OE_MOUSE_GSE2192_CREEDSID_GENE_865_DOWN_TF_Perturbations_Followed_by_Expression


VMH shared-ish
ESR1 (KD-down, female; ESR1 (targets, interactors male): ESR1_KD_HUMAN_GSE26298_CREEDSID_GENE_479_DOWN_TF_Perturbations_Followed_by_Expression, 	
PMC7162552-pcbi.1007800.s007.txt--ESR1_Interactors_ESR1_Rummagene_transcription_factors

AR (KD-bidir DE, female; targets male): AR_35413990_ChIP-Seq_Human_ARposCRProstateCancer_ChEA_2022, AR_KD_HUMAN_GSE22483_CREEDSID_GENE_2194_DOWN_TF_Perturbations_Followed_by_Expression, AR_KD_HUMAN_GSE22483_CREEDSID_GENE_2194_UP_TF_Perturbations_Followed_by_Expression


ARC unique male:
LIN28B (growth phenos): PMC7335061-41467_2020_17150_MOESM4_ESM.xlsx-Table_S3-LIN28B_Correlates_LIN28B_Rummagene_transcription_factors

ASH1L (ID/ASD in mice https://www.nature.com/articles/s42003-021-02282-z ): PMC6815046-13073_2019_678_MOESM2_ESM.xlsx-ASH1L-Unnamed_0_ASH1L_Rummagene_transcription_factors


ARC unique female:
VDR (KO-bidir DE): VDR_KO_MOUSE_GSE2421_CREEDSID_GENE_1153_UP_TF_Perturbations_Followed_by_Expression, VDR_KO_MOUSE_GSE2421_CREEDSID_GENE_1153_DOWN_TF_Perturbations_Followed_by_Expression

ESR1 (KO-bidir DE): ESR1_KO_MOUSE_GSE53812_CREEDSID_GENE_2898_UP_TF_Perturbations_Followed_by_Expression, ESR1_KO_MOUSE_GSE53812_CREEDSID_GENE_2898_DOWN_TF_Perturbations_Followed_by_Expression

ARC general: 
STAT3 (required for hypothalamic response to leptin Buettner 06): PMC7801680-41598_2020_79740_MOESM9_ESM.xlsx-STAT3_Target_Genes_In_AD1-STAT3_Target_Genes_Predicted_By_CHIP-seq_Overlap_With_Genes_In_AD_M1_STAT3_Rummagene_transcription_factors, STAT3_human_TRRUST_Transcription_Factors_2019
```{r}
tfres <- readRDS("spatial_HYP/processed-data/11-GSEA/04a-Sex DE TF-target GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")
tfres <- tfres[names(tfres) %in% c("BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_VMH","BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_ARC")]
tfres <- rbindlist(tfres)

# remove the leadingedge column, which is a list and doesn't play nice with table format
tfres[,leadingEdge:=NULL]

## clean up domain name
tfres[,spdom:=gsub(spdom,pattern="BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_(.*)$",replacement="\\1")]



## subset to TF sets of interest, clean names up
tfres <- tfres[pathway %in% c(
    "PMC6815046-13073_2019_678_MOESM4_ESM.xlsx-MYT1L-Unnamed_0_MYT1L_Rummagene_transcription_factors",
    "PMC6815046-13073_2019_678_MOESM3_ESM.xlsx-TCF20-Unnamed_0_TCF20_Rummagene_transcription_factors",
    "PPARA_KO_MOUSE_GSE6864_CREEDSID_GENE_1372_DOWN_TF_Perturbations_Followed_by_Expression",
    "PPARA_KO_MOUSE_GSE6864_CREEDSID_GENE_1373_UP_TF_Perturbations_Followed_by_Expression",
    "PPARG_OE_MOUSE_GSE2192_CREEDSID_GENE_861_UP_TF_Perturbations_Followed_by_Expression",
    "PPARG_OE_MOUSE_GSE2192_CREEDSID_GENE_865_DOWN_TF_Perturbations_Followed_by_Expression",
    "ESR1_KD_HUMAN_GSE26298_CREEDSID_GENE_479_DOWN_TF_Perturbations_Followed_by_Expression",
    "PMC7162552-pcbi.1007800.s007.txt--ESR1_Interactors_ESR1_Rummagene_transcription_factors",
    "AR_35413990_ChIP-Seq_Human_ARposCRProstateCancer_ChEA_2022",
    "AR_KD_HUMAN_GSE22483_CREEDSID_GENE_2194_DOWN_TF_Perturbations_Followed_by_Expression",
    "AR_KD_HUMAN_GSE22483_CREEDSID_GENE_2194_UP_TF_Perturbations_Followed_by_Expression",
    "PMC7335061-41467_2020_17150_MOESM4_ESM.xlsx-Table_S3-LIN28B_Correlates_LIN28B_Rummagene_transcription_factors",
    "PMC6815046-13073_2019_678_MOESM2_ESM.xlsx-ASH1L-Unnamed_0_ASH1L_Rummagene_transcription_factors",
    "VDR_KO_MOUSE_GSE2421_CREEDSID_GENE_1153_UP_TF_Perturbations_Followed_by_Expression",
    "VDR_KO_MOUSE_GSE2421_CREEDSID_GENE_1153_DOWN_TF_Perturbations_Followed_by_Expression",
    "ESR1_KO_MOUSE_GSE53812_CREEDSID_GENE_2898_UP_TF_Perturbations_Followed_by_Expression",
    "ESR1_KO_MOUSE_GSE53812_CREEDSID_GENE_2898_DOWN_TF_Perturbations_Followed_by_Expression",
    "PMC7801680-41598_2020_79740_MOESM9_ESM.xlsx-STAT3_Target_Genes_In_AD1-STAT3_Target_Genes_Predicted_By_CHIP-seq_Overlap_With_Genes_In_AD_M1_STAT3_Rummagene_transcription_factors",
    "ZNF142_SIRNA_MCF7_HUMAN_GSE79586_RNASEQ_UP_TF_Perturbations_Followed_by_Expression",
    "ZNF142_SIRNA_MCF7_HUMAN_GSE79586_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression")]



## data sourcesr
tfres[pathway %in% grep(pathway,pattern="Rummagene",value=T),source:="Rummagene TFs"]
# tfres[pathway %in% grep(pathway,pattern="TRRUST",value=T),source:="TRRUST"]
tfres[pathway %in% grep(pathway,pattern="CREEDSID|_GSE",value=T),source:="GEO TF\nPerturbations"]
tfres[pathway %in% grep(pathway,pattern="ChEA",value=T),source:="ChEA 2022"]

## TF specifics where applicable
tfres[pathway %in% c("PPARA_KO_MOUSE_GSE6864_CREEDSID_GENE_1372_DOWN_TF_Perturbations_Followed_by_Expression",
    "PPARA_KO_MOUSE_GSE6864_CREEDSID_GENE_1373_UP_TF_Perturbations_Followed_by_Expression"),tf:="PPARA (KO DE)"]
tfres[pathway %in% c("PPARG_OE_MOUSE_GSE2192_CREEDSID_GENE_861_UP_TF_Perturbations_Followed_by_Expression",
    "PPARG_OE_MOUSE_GSE2192_CREEDSID_GENE_865_DOWN_TF_Perturbations_Followed_by_Expression"),tf:="PPARG (OE DE)"]
tfres[pathway=="ESR1_KD_HUMAN_GSE26298_CREEDSID_GENE_479_DOWN_TF_Perturbations_Followed_by_Expression",tf:="ESR1 (KD Down)"]
tfres[pathway=="PMC7162552-pcbi.1007800.s007.txt--ESR1_Interactors_ESR1_Rummagene_transcription_factors",tf:="ESR1 (Interactors)"]
tfres[pathway %in% c("AR_KD_HUMAN_GSE22483_CREEDSID_GENE_2194_DOWN_TF_Perturbations_Followed_by_Expression",
    "AR_KD_HUMAN_GSE22483_CREEDSID_GENE_2194_UP_TF_Perturbations_Followed_by_Expression"),tf:="AR (KD DE)"]
tfres[pathway %in% c("VDR_KO_MOUSE_GSE2421_CREEDSID_GENE_1153_UP_TF_Perturbations_Followed_by_Expression",
    "VDR_KO_MOUSE_GSE2421_CREEDSID_GENE_1153_DOWN_TF_Perturbations_Followed_by_Expression"),tf:="VDR (KO DE)"]
tfres[pathway %in% c(
    "ESR1_KO_MOUSE_GSE53812_CREEDSID_GENE_2898_UP_TF_Perturbations_Followed_by_Expression",
     "ESR1_KO_MOUSE_GSE53812_CREEDSID_GENE_2898_DOWN_TF_Perturbations_Followed_by_Expression"),tf:="ESR1 (KO DE)"]
tfres[pathway=="ZNF142_SIRNA_MCF7_HUMAN_GSE79586_RNASEQ_UP_TF_Perturbations_Followed_by_Expression",tf:="ZNF142 (KD Up)"]
tfres[pathway=="ZNF142_SIRNA_MCF7_HUMAN_GSE79586_RNASEQ_DOWN_TF_Perturbations_Followed_by_Expression",tf:="ZNF142 (KD Down)"]
tfres[pathway %in% c("PMC7801680-41598_2020_79740_MOESM9_ESM.xlsx-STAT3_Target_Genes_In_AD1-STAT3_Target_Genes_Predicted_By_CHIP-seq_Overlap_With_Genes_In_AD_M1_STAT3_Rummagene_transcription_factors","STAT3_human_TRRUST_Transcription_Factors_2019"),tf:="STAT3"]
tfres[pathway=="PMC7335061-41467_2020_17150_MOESM4_ESM.xlsx-Table_S3-LIN28B_Correlates_LIN28B_Rummagene_transcription_factors",tf:="LIN28B"]
tfres[pathway=="PMC6815046-13073_2019_678_MOESM2_ESM.xlsx-ASH1L-Unnamed_0_ASH1L_Rummagene_transcription_factors",tf:="ASH1L"]
tfres[pathway=="PMC6815046-13073_2019_678_MOESM3_ESM.xlsx-TCF20-Unnamed_0_TCF20_Rummagene_transcription_factors",tf:="TCF20"]
tfres[pathway=="AR_35413990_ChIP-Seq_Human_ARposCRProstateCancer_ChEA_2022",tf:="AR"]
tfres[pathway=="PMC6815046-13073_2019_678_MOESM4_ESM.xlsx-MYT1L-Unnamed_0_MYT1L_Rummagene_transcription_factors",tf:="MYT1L"]



### group by sex•region, flip all female enrichments (negative NES and female test) to positive
tfres[dir=="female_up"&NES<0,NES:=abs(NES)]

tfres[dir=="female_up"&spdom=="ARC",sexreg:="vARC Fem. Up"]
tfres[dir=="male_up"&spdom=="ARC",sexreg:="vARC Male Up"]
tfres[dir=="female_up"&spdom=="VMH",sexreg:="vVMH Fem. Up"]
tfres[dir=="male_up"&spdom=="VMH",sexreg:="vVMH Male Up"]


### make data source name consistent across the two result tables
setnames(tfres,"source","genesource")
```

### for this to work, we have to manually make a double legend kinda clunkily. modified slightly from https://stackoverflow.com/questions/62541720/ggplot2-show-multiple-keys-shapes-in-size-legend -- speciifically, defining these variables below and then the scale_size_manual call.

we also need to extract just one legend from a not-actually-used plot where the two analyses are strung together for purposes of axis scaling, using this to grab said legend to pass back later to gridExtra:
```{r}
### helper function to extract a legend and pipe back to a diff plot: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
extract_legend <- function(my_ggp) {
    step1 <- ggplot_gtable(ggplot_build(my_ggp))
    step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
    step3 <- step1$grobs[[step2]]
    return(step3)
}

## 
tmpt <- copy(tfres)
tmpt[,pathway:=tf]

legdat <- rbind(tmpt[,.(pathway,pval,padj,NES,spdom,sexreg,genesource)],msigpltdat[,.(pathway,pval,padj,NES,spdom,sexreg,genesource)])

## get the breaks based on generating the plot once quick, then get a vector of the same values to feed into the legend override

breaks <- seq(min(legdat$NES),(ceiling(max(legdat$NES))-0.5),0.5)
n_breaks <- length(breaks)
labels <- c(c(rep("",n_breaks),breaks,rep("",2*n_breaks)))
## need 8 shapes: use 16 17 18 15 23 25 7 9 or 16-18 15 7 9 10 14
shapes <- c(rep(16,n_breaks),rep(17,n_breaks),rep(18,n_breaks),rep(15,n_breaks),rep(7,n_breaks),rep(9,n_breaks),rep(23,n_breaks),rep(25,n_breaks))
breaks2 <- rep(breaks, length(unique(shapes)))

## order source as a factor so the first plots' sets are all the first entries in the legend
legdat[,genesource:=factor(genesource,levels=c("Literature-Curated","GO Mol. Fxn","Human Phenotypes","miRNA Targets","Rummagene TFs","GEO TF\nPerturbations","ChEA 2022"))]

shapelabs <- c("Literature-Curated"=7,"GO Mol. Fxn"=9,"Human Phenotypes"=10,"miRNA Targets"=14,"Rummagene TFs"=16,"GEO TF\nPerturbations"=18,"ChEA 2022"=17)

shapelabs <- shapelabs[levels(legdat$genesource)]

legplt <- ggplot(legdat,aes(x=-log10(padj),y=pathway,size=NES))+
   geom_point(aes(col=sexreg,fill=sexreg,size=NES,shape=genesource,y=pathway),alpha=0.7,position = position_dodge2v(preserve = "single",height=0.75))+
   scale_color_manual(values=sexregpal)+
   scale_shape_manual(values=shapelabs)+
   guides(col=guide_legend(override.aes=list(stroke = 0,size=3.5,alpha=0.7)))+
   labs(size="NES",shape="Geneset Source",col="DE Pattern",fill="DE Pattern")+
    # ^ make size legend with all point types
    theme(legend.text = element_text(size=6), 
          legend.title = element_text(size=7,hjust=0.5,vjust=0.5,margin=margin(0,0,-0.01,0,"in")),
          legend.key.height = ggplot2::unit(0.15,"in"),
          legend.spacing.x = ggplot2::unit(-0.015,"in"),
          legend.spacing.y = ggplot2::unit(0.05,"in"),
          legend.background=element_blank(),
          legend.key=element_blank(),
          legend.box.background = element_blank(),
          legend.frame = element_blank(),
          legend.margin = margin(0,0,0,-0.05,"in"))
   

### extract and hold onto legend
legcd <- extract_legend(legplt)
dev.off()

rm(tmpt,legplt,legdat)
```

### now make the two plots as entries in a list using the same breaks for NES, same shape palette, etc. arrangegrob with the legend into a single row of panels
```{r}
fig5bd <- list()

## make sure the extracted legend and the actual plot shapes are consistent as follows:
tfres[,genesource:=factor(genesource,levels=c("Rummagene TFs","GEO TF\nPerturbations","ChEA 2022"))]
msigpltdat[,genesource:=factor(genesource,levels=c("Literature-Curated","GO Mol. Fxn","Human Phenotypes","miRNA Targets"))]

tfshapes <- shapelabs[levels(tfres$genesource)]
msigshapes <- shapelabs[levels(msigpltdat$genesource)]

#### 5B: TF results #### 

## reorder tfs as factor so TFs are in vaguely functionally-related groups
tfres[,tf:=factor(tf,levels=c("AR", "AR (KD DE)", "ESR1 (KD Down)", "ESR1 (KO DE)", "ESR1 (Interactors)", "PPARA (KO DE)", "PPARG (OE DE)", "LIN28B", "STAT3", "VDR (KO DE)", "TCF20", "ASH1L", "ZNF142 (KD Down)", "ZNF142 (KD Up)", "MYT1L"))]


fig5bd[[1]] <- ggplot(tfres,aes(y=tf,x=-log10(padj),size=NES))+
    ylab("TF")+
    scale_color_manual(values=sexregpal)+
    xlab("-log10 (*p*-value)")+
    geom_vline(xintercept=-log10(0.05),linetype="dashed",linewidth=0.25,color="darkred")+
    scale_shape_manual(values=tfshapes)+
    #add minor gridlines between TFs to contain jittered points
    geom_hline(yintercept = seq(0.5,length(unique(tfres$tf))-0.5),color="grey",linewidth=.375)+
    geom_point(aes(col=sexreg,fill=sexreg,size=NES,shape=genesource,y=tf),alpha=0.7,position = position_dodge2v(preserve = "single",height=0.75))+
    scale_size_continuous(range = c(breaks[1],breaks[length(breaks)]),breaks = rep(breaks,length(tfshapes)))+
    guides(col="none",size="none",shape="none",fill="none")+
    theme(axis.text.y = element_text(size = 8),
          axis.title.y=element_text(size=9,margin=margin(0,-0.125,0,0,"in")),
          axis.title.x = element_markdown(size = 9),
          axis.text.x = element_text(size = 8), 
          panel.grid.major.y = element_line(color = "#595959",linewidth = 0.25),
          panel.grid.minor.y = element_line(color = "#595959",linewidth = 0.25),
          panel.grid.major.x=element_line(color="#bfbfbf",linewidth=0.25),
          panel.grid.minor.x=element_line(color="#bfbfbf",linewidth=0.25),
          axis.ticks.x = element_line(linewidth=0.25,color="#bfbfbf"),
          axis.ticks.y=element_line(color = "#595959",linewidth = 0.25))

#### 5D: MSigDB results
msigpltdat[,pathway:=factor(pathway,levels=c("Aggressive/im-\npulsive behavior","Dental caries","Acid-base ab-\nnormality","miR-21","miR-30a_3P","Chromatin binding","Thecal cell of ovary\n(early atretic follicle)","Temporal progest-\nerone response"))]

fig5bd[[2]] <- ggplot(msigpltdat,aes(y=pathway,x=-log10(padj),size=NES))+
    ylab("MSigDB Set")+
    scale_color_manual(values=sexregpal)+
    xlab("-log10 (*p*-value)")+
    labs(shape="none",col="none",fill="none",size="none")+
    geom_vline(xintercept=-log10(0.05),linetype="dashed",linewidth=0.25,color="darkred")+
    scale_shape_manual(values=msigshapes)+
    #add minor gridlines between TFs to contain jittered points
    geom_hline(yintercept = seq(0.5,length(unique(msigpltdat$pathway))-0.5),color="grey",linewidth=.375)+
    scale_size_continuous(range = c(breaks[1],breaks[length(breaks)]),breaks = rep(breaks,length(msigshapes)))+
    geom_point(aes(col=sexreg,fill=sexreg,size=NES,shape=genesource,y=pathway),alpha=0.7,position = position_dodge2v(preserve = "single",height=0.75))+
    guides(col="none",size="none",shape="none",fill="none")+
    theme(axis.text.y = element_text(size = 8),
          axis.title.y=element_text(size=9,margin=margin(0,-0.175,0,0,"in")),
          axis.title.x = element_markdown(size = 9),
          axis.text.x = element_text(size = 8), 
          panel.grid.major.y = element_line(color = "#595959",linewidth = 0.25),
          panel.grid.minor.y = element_line(color = "#595959",linewidth = 0.25),
          panel.grid.major.x=element_line(color="#bfbfbf",linewidth=0.25),
          panel.grid.minor.x=element_line(color="#bfbfbf",linewidth=0.25),
          axis.ticks.x = element_line(linewidth=0.25,color="#bfbfbf"),
          axis.ticks.y=element_line(color = "#595959",linewidth = 0.25),
          legend.spacing.x = ggplot2::unit(-0.015,"in"))

```

## save 6C-D combined
```{r}
pdf("manuscript_plots/Fig5/5BandD-sexDE_msigdb_and_TFtargGSEAs.pdf",width=4.25,height=6,onefile = FALSE)
do.call("grid.arrange",list(
    arrangeGrob(grobs=fig5bd,ncol=1,widths=3.15),
    legcd,
    ncol=2,
    widths=c(3.15,1.1)))
dev.off()

```

C-Sex DE/Disease GSEA 
```{r}
# read GSEA results of  of sex DEGs in GWAS-derived gene sets
sxgs <- readRDS("spatial_HYP/processed-data/11-GSEA/06-Common variant disease gene enrichment in sex DEGs.RDS")
sxgs <- sxgs[genesets_from=="union"]

sxgsunsign <- readRDS("spatial_HYP/processed-data/11-GSEA/06b-Unsigned_Common variant disease gene enrichment in sex DEGs.RDS")
sxgsunsign <- sxgsunsign[genesets_from=="union"]
sxgsunsign[,dir:="both"]

sxgs <- rbind(sxgs,sxgsunsign)
rm(sxgsunsign)

sxgs[dir=="female_up",dir:="Female-\nUpreg."]
sxgs[dir=="male_up",dir:="Male-\nUpreg."]
sxgs[dir=="both",dir:="Bidir.\nSex DE"]

# plot
dsx <- sxgs[spdom %in% c("BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_ARC","BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_VMH")]

dsx[spdom=="BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_ARC",spdom:="ARC"]
dsx[spdom=="BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH_VMH",spdom:="VMH"]

## reorder traits of interest into broad groups (not that they will be explicitly labeled as such, but still)
dsx[,pathway:=factor(pathway,levels=c("Lefthanded","BMI","Anorexia","Panic","PTSD","GAD","OCD","TS","ADHD","NDD","ASD","OUD","AUD","Cannabis","SCZ","BIP","MDD","AD"))]


# absolute value NES since we have one-sided tests now
dsx[dir!="Bidir. Sex DE",NES:=abs(NES)]
#dsx[NES<0,NES:=0]

dsx[,dir:=factor(dir,levels = c("Male-\nUpreg.","Female-\nUpreg.","Bidir.\nSex DE"))]

## make plot; use a rearranged version of the wesanderson:: palette "Darjeeling2" (colors 3,1,4,2)
scalefact <- max(dsx$NES)-min(dsx$NES)

cairo_pdf("manuscript_plots/Fig5/5C-DomainSexDE_Dis_GSEA.pdf",height=3,width=3.5)
ggplot(dsx,aes(x=spdom,y=pathway,fill=NES))+
    geom_tile(linewidth = 0.25,colour = "black")+
    geom_text(label=ifelse(dsx$padj<0.05,"*",""),size = 5,nudge_y = -0.2)+
    geom_text(label=ifelse(dsx$padj>0.05&dsx$pval<0.05,"•",""),size = 4)+
    facet_grid(.~dir)+
    # scale_fill_viridis_c(begin=0.15,end=0.37,option="H",direction = 1)+
    scale_fill_gradientn(colors=c("#ffffff","#F9EFDE","#E8C696","#D69C4E"),values = c(0,min(dsx[dir=="Bidir.\nSex DE",NES])/scalefact,max(dsx[dir=="Bidir.\nSex DE",NES])/scalefact,1))+
    scale_x_discrete(expand = c(0,0)) +
    scale_y_discrete(expand = c(0,0)) +
    labs(fill="Normalized\nEnrichment\nScore (<0:\nDepleted)")+
    xlab("Visium Domain")+
    ylab("Disease / Trait\nGene Set") +
    theme(plot.background = element_blank(),axis.text.x = element_text(size=9,angle=90,vjust=0.5),axis.text.y=element_text(size=9),axis.title.x=element_text(size=10,hjust=0.5),axis.title.y=element_text(size=10,vjust=0.5,margin=margin(0,0.05,0,0,"in")),legend.title = element_text(size=9,hjust=0.5),legend.text=element_text(size=8),legend.margin=margin(0,0,0,-0.1,"in"),strip.text=element_text(size=10),strip.background = element_rect(fill=NA))
dev.off()

rm(dsx,scalefact)
```


## 5E: looking at the overlap between leadingEdge genes driving male ASD enrichment and genes on Xenium, MYT1L, an obesity+ASD syndrome, is sex DE in xVMH CRHR2 but not LAMP5. let's take a look at this warlock

```{r}
hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
fineanno <- fread("manuscript_plot_code/xARCxVMH_cluster_detailedlabs_andplotformatnames.txt")

## asterisk-markdown for italicization isn't working with the legend extraction + arrangement thing we use for making this plot, so just italicize the gene names in illustrator
fineanno[,plotclusname:=gsub(plotclusname,pattern="\\*",replacement="")]

smoove <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt")

setnames(bkcl,2,"cl")
bkanno <- unique(bkanno[,.(clus,bjm_annot)])

# drop DISCARDS
bkanno <- bkanno[!(bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T))]
bkcl <- merge.data.table(bkcl,bkanno,by.x="cl",by.y="clus")
bkcl <- merge.data.table(bkcl,fineanno[,.(bjm_annot,dom.clus,plotclusname)],by="bjm_annot",all.x=T)
bkcl[!is.na(plotclusname),bjm_annot:=plotclusname]

# add clusters and smoothed domains
hypx$dom <- DataFrame(smoove[,.(rn,dualVMHARC4)],row.names=smoove$rn)[colnames(hypx),"dualVMHARC4"]

hypx$bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),"bjm_annot"]

```

load cell type palettes
```{r}
pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")

xenvmhpal <- pals$XenVMH
# as above, markdown for facet titles isn't working so just kick the asterisks
names(xenvmhpal) <- gsub(names(xenvmhpal),pattern="\\*",replacement="")


stopifnot(names(xenvmhpal) %in% unique(bkcl[dom.clus %in% grep(dom.clus,pattern="VMH",value=T),plotclusname]))
```

### plot MYT1L in two male, two female samples showing VMH CRHR2 and LAMP5. append manuscript sample ids. markdown legends with the extract-and-arrange approach here doesn't work so italicize genes in legends manually in illustrator.
```{r}
mytx <- hypx["MYT1L",]
mytx <- mytx[,mytx$dom=="VMH"]
mytx <- mytx[,mytx$bkcl %in% c("CRHR2","LAMP5")]

# add manuscript identifiers
mscriptid <- fread("standardized_sampleids_for_plotting.txt")
tmpcd <- as.data.table(colData(mytx),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,mscriptid[,.(sample_id,manuscript_id)],all.x=T,by="sample_id")
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$rn)[colnames(mytx),]
tmpcd$rn <- NULL
colData(mytx) <- tmpcd
mytx$sample_id <- mytx$manuscript_id

mytxa <- mytx[,mytx$sample_id %in% c("x1735C_M","x6197B_M","x5993B_F","x6588C_F")]
mytxa$`log counts` <- logcounts(mytxa)["MYT1L",]

mytx.m1 <- mytxa[,mytxa$sample_id=="x1735C_M"]

mytx.m2 <- mytxa[,mytxa$sample_id=="x6197B_M"]

mytx.f1 <- mytxa[,mytxa$sample_id=="x5993B_F"]

mytx.f2 <-  mytxa[,mytxa$sample_id=="x6588C_F"]

mytxs <- list(mytx.m1,mytx.f1,mytx.m2,mytx.f2)

# make all plots, extract legend from the first one (which scales  the fill by expression across the whole set of samples, i.e. mytxa)
mytxplt <- lapply(mytxs,FUN=function(m){
    # m <- m[,m$bkcl=="*TAC3*-*ESR1*"]
    colData(m)$`log counts` <- logcounts(m)["MYT1L",]
    p <- as.data.table(cbind(m$`log counts`,m$bkcl))
    setnames(p,c("log counts","bkcl"))
    p[,`log counts`:=as.numeric(`log counts`)]
    p2 <- ggplot(p,aes(x=bkcl,y=`log counts`,col=bkcl,fill=`log counts`))+
        #gbeeswarm::geom_beeswarm(size=1.25,stroke =0.45,shape=21,corral.width = 0.1,)+
        rasterize(geom_quasirandom(dodge.width=1.5,width = 0.475, size=0.35,stroke =0.175,shape=21),layers="Points",dpi = 600,dev="cairo_png")+
        scale_color_manual(values=xenvmhpal)+
        scale_fill_gradient(low="white",high="black",limits=c(min(mytxa$`log counts`),max(mytxa$`log counts`)))+
        scale_y_continuous(limits=c(0,max(mytxa$`log counts`)))+
        ggtitle(paste0("*MYT1L* ",unique(m$sample_id)))+
        labs(fill="log counts",color="xARC Cluster")+
        theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_markdown(size=7),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=7.5,hjust=0.5),axis.text.x=element_blank(),axis.text.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),axis.ticks.x = element_blank(),axis.ticks.y=element_blank(),panel.grid = element_blank())
    return(p2)
})

leg.tmp <- extract_legend(mytxplt[[1]])
dev.off()

# now toss the legends
mytxplt<- lapply(mytxplt,FUN=function(x){x+guides(fill="none")+guides(color="none")})
```

```{r}
pdf("manuscript_plots/Fig5/5E-MYT1L beeswarms.pdf",height=3,width=3.75,onefile = FALSE)
do.call("grid.arrange",list(arrangeGrob(grobs=mytxplt,nrow=2),leg.tmp,ncol=2,widths=c(3,0.75)))
dev.off()

```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```


R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.4

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0            parallelly_1.43.0             
 [3] colorout_1.3-0.2               ggstance_0.3.7                
 [5] ggbeeswarm_0.7.2               ggrastr_1.0.2                 
 [7] Cairo_1.6-2                    ggtext_0.1.2                  
 [9] gridExtra_2.3                  escheR_1.6.0                  
[11] ggplot2_3.5.1                  SpatialFeatureExperiment_1.8.6
[13] SpatialExperiment_1.16.0       SingleCellExperiment_1.28.1   
[15] SummarizedExperiment_1.36.0    Biobase_2.66.0                
[17] GenomicRanges_1.58.0           GenomeInfoDb_1.42.3           
[19] IRanges_2.40.1                 S4Vectors_0.44.0              
[21] BiocGenerics_0.52.0            MatrixGenerics_1.18.1         
[23] matrixStats_1.5.0              data.table_1.17.0             
[25] rlang_1.1.5                   

loaded via a namespace (and not attached):
  [1] rstudioapi_0.17.1         jsonlite_1.9.1            wk_0.9.4                 
  [4] magrittr_2.0.3            TH.data_1.1-3             magick_2.8.6             
  [7] farver_2.1.2              rmarkdown_2.29            zlibbioc_1.52.0          
 [10] vctrs_0.6.5               spdep_1.3-10              DelayedMatrixStats_1.28.1
 [13] RCurl_1.98-1.17           terra_1.8-29              htmltools_0.5.8.1        
 [16] S4Arrays_1.6.0            BiocNeighbors_2.0.1       Rhdf5lib_1.28.0          
 [19] s2_1.1.7                  SparseArray_1.6.2         rhdf5_2.50.2             
 [22] LearnBayes_2.15.1         spData_2.3.4              KernSmooth_2.23-26       
 [25] htmlwidgets_1.6.4         sandwich_3.1-1            zoo_1.8-13               
 [28] commonmark_1.9.5          lifecycle_1.0.4           pkgconfig_2.0.3          
 [31] Matrix_1.7-3              R6_2.6.1                  fastmap_1.2.0            
 [34] GenomeInfoDbData_1.2.13   digest_0.6.37             colorspace_2.1-1         
 [37] rprojroot_2.0.4           dqrng_0.4.1               beachmat_2.22.0          
 [40] labeling_0.4.3            spatialreg_1.3-6          httr_1.4.7               
 [43] abind_1.4-8               compiler_4.4.3            here_1.0.1               
 [46] proxy_0.4-27              withr_3.0.2               tiff_0.1-12              
 [49] DBI_1.2.3                 HDF5Array_1.34.0          R.utils_2.13.0           
 [52] MASS_7.3-65               DelayedArray_0.32.0       rjson_0.2.23             
 [55] classInt_0.4-11           tools_4.4.3               units_0.8-7              
 [58] vipor_0.4.7               beeswarm_0.4.0            R.oo_1.27.0              
 [61] glue_1.8.0                nlme_3.1-167              EBImage_4.48.0           
 [64] rhdf5filters_1.18.1       gridtext_0.1.5            grid_4.4.3               
 [67] sf_1.0-19                 generics_0.1.3            gtable_0.3.6             
 [70] R.methodsS3_1.8.2         class_7.3-23              sp_2.2-0                 
 [73] xml2_1.3.8                XVector_0.46.0            stringr_1.5.1            
 [76] markdown_2.0              pillar_1.10.1             limma_3.62.2             
 [79] splines_4.4.3             dplyr_1.1.4               lattice_0.22-6           
 [82] survival_3.8-3            deldir_2.0-4              tidyselect_1.2.1         
 [85] locfit_1.5-9.12           scuttle_1.16.0            sfheaders_0.4.4          
 [88] knitr_1.50                litedown_0.6              edgeR_4.4.2              
 [91] xfun_0.51                 statmod_1.5.0             DropletUtils_1.26.0      
 [94] stringi_1.8.4             UCSC.utils_1.2.0          fftwtools_0.9-11         
 [97] yaml_2.3.10               boot_1.3-31               evaluate_1.0.3           
[100] codetools_0.2-20          tibble_3.2.1              cli_3.6.4                
[103] munsell_0.5.1             Rcpp_1.0.14               zeallot_0.1.0            
[106] coda_0.19-4.1             png_0.1-8                 parallel_4.4.3           
[109] jpeg_0.1-11               sparseMatrixStats_1.18.0  bitops_1.0-9             
[112] mvtnorm_1.3-3             scales_1.3.0              e1071_1.7-16             
[115] crayon_1.5.3              multcomp_1.4-28          
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sonoma 14.7.4
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2025-03-26
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-8     2024-09-12 [1] CRAN (R 4.4.1)
   beachmat                   2.22.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   Biobase                  * 2.66.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics             * 0.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocNeighbors              2.0.1     2024-11-28 [1] Bioconductor 3.20 (R 4.4.2)
   BiocParallel             * 1.40.0    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   bitops                     1.0-9     2024-10-03 [1] CRAN (R 4.4.1)
   boot                       1.3-31    2024-08-28 [1] CRAN (R 4.4.3)
   Cairo                    * 1.6-2     2023-11-28 [1] CRAN (R 4.4.0)
   class                      7.3-23    2025-01-01 [1] CRAN (R 4.4.3)
   classInt                   0.4-11    2025-01-08 [1] CRAN (R 4.4.1)
 P cli                        3.6.4     2025-02-13 [2] CRAN (R 4.4.1)
   coda                       0.19-4.1  2024-01-31 [1] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.3)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-1     2024-07-26 [1] CRAN (R 4.4.0)
   commonmark                 1.9.5     2025-03-17 [1] CRAN (R 4.4.1)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.17.0    2025-02-22 [1] CRAN (R 4.4.1)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.32.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   DelayedMatrixStats         1.28.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.37    2024-08-19 [1] CRAN (R 4.4.1)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.26.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   e1071                      1.7-16    2024-09-16 [1] CRAN (R 4.4.1)
   EBImage                    4.48.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   edgeR                      4.4.2     2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
   escheR                   * 1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   evaluate                   1.0.3     2025-01-10 [1] CRAN (R 4.4.1)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.42.3    2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData           1.2.13    2024-12-12 [1] Bioconductor
   GenomicRanges            * 1.58.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm               * 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggstance                 * 0.3.7     2024-04-05 [1] CRAN (R 4.4.0)
   ggtext                   * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   glue                       1.8.0     2024-09-30 [1] CRAN (R 4.4.1)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext                   0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                     0.3.6     2024-10-25 [1] CRAN (R 4.4.1)
   HDF5Array                  1.34.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.40.1    2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   jpeg                       0.1-11    2025-03-21 [1] CRAN (R 4.4.1)
   jsonlite                   1.9.1     2025-03-03 [1] CRAN (R 4.4.1)
   KernSmooth                 2.23-26   2025-01-01 [1] CRAN (R 4.4.3)
   knitr                      1.50      2025-03-16 [1] CRAN (R 4.4.2)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.3)
   LearnBayes                 2.15.1    2018-03-18 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.62.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   litedown                   0.6       2025-02-27 [1] CRAN (R 4.4.3)
   locfit                     1.5-9.12  2025-03-05 [1] CRAN (R 4.4.1)
   magick                     2.8.6     2025-03-23 [1] CRAN (R 4.4.1)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   markdown                   2.0       2025-03-23 [1] CRAN (R 4.4.3)
   MASS                       7.3-65    2025-02-28 [1] CRAN (R 4.4.1)
   Matrix                     1.7-3     2025-03-11 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.18.1    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   matrixStats              * 1.5.0     2025-01-07 [1] CRAN (R 4.4.1)
   multcomp                   1.4-28    2025-01-29 [1] CRAN (R 4.4.1)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   mvtnorm                    1.3-3     2025-01-10 [1] CRAN (R 4.4.1)
   nlme                       3.1-167   2025-01-27 [1] CRAN (R 4.4.3)
   parallelly               * 1.43.0    2025-03-24 [1] CRAN (R 4.4.1)
   pillar                     1.10.1    2025-01-07 [1] CRAN (R 4.4.1)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.27.0    2024-11-01 [1] CRAN (R 4.4.1)
   R.utils                    2.13.0    2025-02-24 [1] CRAN (R 4.4.1)
   R6                         2.6.1     2025-02-15 [1] CRAN (R 4.4.1)
   Rcpp                       1.0.14    2025-01-12 [1] CRAN (R 4.4.1)
   RCurl                      1.98-1.17 2025-03-22 [1] CRAN (R 4.4.1)
   rhdf5                      2.50.2    2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   rhdf5filters               1.18.1    2025-03-06 [1] Bioconductor 3.20 (R 4.4.3)
   Rhdf5lib                   1.28.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   rjson                      0.2.23    2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                    * 1.1.5     2025-01-17 [2] CRAN (R 4.4.1)
   rmarkdown                  2.29      2024-11-04 [1] CRAN (R 4.4.1)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.17.1    2024-10-22 [1] CRAN (R 4.4.1)
   s2                         1.1.7     2024-07-17 [1] CRAN (R 4.4.0)
   S4Arrays                   1.6.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors                * 0.44.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sandwich                   3.1-1     2024-09-15 [1] CRAN (R 4.4.1)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scuttle                    1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   sessioninfo                1.2.3     2025-02-05 [1] CRAN (R 4.4.1)
   sf                         1.0-19    2024-11-05 [1] CRAN (R 4.4.1)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.28.1    2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   sp                         2.2-0     2025-02-01 [1] CRAN (R 4.4.1)
   SparseArray                1.6.2     2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
   sparseMatrixStats          1.18.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialExperiment        * 1.16.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   SpatialFeatureExperiment * 1.8.6     2025-01-30 [1] Bioconductor 3.20 (R 4.4.3)
   spatialreg                 1.3-6     2024-12-02 [1] CRAN (R 4.4.1)
   spData                     2.3.4     2025-01-08 [1] CRAN (R 4.4.1)
   spdep                      1.3-10    2025-01-20 [1] CRAN (R 4.4.1)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   stringi                    1.8.4     2024-05-06 [1] CRAN (R 4.4.0)
   stringr                    1.5.1     2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.36.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   survival                   3.8-3     2024-12-17 [1] CRAN (R 4.4.3)
   terra                      1.8-29    2025-02-26 [1] CRAN (R 4.4.1)
   TH.data                    1.1-3     2025-01-17 [1] CRAN (R 4.4.1)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.2.0     2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   units                      0.8-7     2025-03-11 [2] CRAN (R 4.4.1)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   withr                      3.0.2     2024-10-28 [1] CRAN (R 4.4.1)
   wk                         0.9.4     2024-10-11 [1] CRAN (R 4.4.1)
   xfun                       0.51      2025-02-19 [1] CRAN (R 4.4.1)
   xml2                       1.3.8     2025-03-14 [1] CRAN (R 4.4.1)
   XVector                    0.46.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                       2.3.10    2024-07-26 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.52.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   zoo                        1.8-13    2025-02-22 [1] CRAN (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────────