---
title: "SF-Negprobe, gene metrics per Xenium sample, Xenium cell, and per cell in
  Xenium VMH or ARC clusters"
output: html_document
date: "2024-09-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggrastr)
library(gridExtra)
library(ggtext)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

# here we only need the Xenium object after QC, but with negative probe counts. since we only used banksy to cluster QC-passing cells, we can get this from the SPE and cluster label tables and the SPE colData.
```{r}
hypx <- readRDS("xenium_HYP/processed-data/04_general_QC_and_normalization/01b-sfe_controlOutliers_and_giantNuclei_filtered.RDS")

bksmooth <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt") ## xenium domain assignments after smoothing, by xenium cell

## we'll want two sets of data: one after filtering, before clustering, and another only with retained clusters. drop discarded cell clusters, too
bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"cl")

bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

bkanno <- unique(bkanno[,.(clus,bjm_annot)])

## convert ARC_1..etc to more descriptive cluster IDs defined based on ARC/VMH specific cluster marker analyses
arcvmhanno <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04e_ARC_and_VMH_cluster_identities_based_on_cellgroup_1vall.txt")
## simplify these by removing VMH/ARC and underscores; asterisk-ize so the gene names  can be italiczed in ggplot elements; remove vmh and arc from the names but add a temporary column here so we know they're VMH and ARC for mapping onto the de result sets
arcvmhanno[,subclus_domain:=gsub(subclus_annot,pattern="^(VMH)_.*$",replacement="\\1")]
arcvmhanno[,subclus_annot:=gsub(subclus_annot,pattern="VMH_",replacement="")]

arcvmhanno[subclus_domain!="VMH",subclus_domain:=gsub(subclus_annot,pattern="^(ARC)_.*$",replacement="\\1")]
arcvmhanno[,subclus_annot:=gsub(subclus_annot,pattern="ARC_",replacement="")]

arcvmhanno[,subclus_annot:=gsub(subclus_annot,pattern="_",replacement=" ")]
arcvmhanno[subclus_annot=="SLC17A6-CRHR2",subclus_annot:="*SLC17A6*-*CRHR2*"]
arcvmhanno[subclus_annot=="FEZF1-LAMP5-NRGN",subclus_annot:="*FEZF1*-*LAMP5*-*NRGN*"]
arcvmhanno[subclus_annot=="HCRT-SLC6A3-TAC1",subclus_annot:="*HCRT*-*SLC6A3*-*TAC1*"]
arcvmhanno[subclus_annot=="AGRP",subclus_annot:="*AGRP*"]
arcvmhanno[subclus_annot=="TAC3-AR-ESR1",subclus_annot:="*TAC3*-*AR*-*ESR1*"]
arcvmhanno[subclus_annot=="POMC-Oligo-Mixed",subclus_annot:="*POMC*-Oligo"]
arcvmhanno[subclus_annot=="Excit-SLC17A7-Glial-Mixed",subclus_annot:="*SLC17A7*-Glia"]
arcvmhanno[subclus_annot=="GAL-TRH-GHRH",subclus_annot:="*GAL*-*TRH*-*GHRH*"]

## append these to the other annots and replace bjm_annot with the ARC/VMH cell type labels where applicable
bkanno <- merge.data.table(bkanno,arcvmhanno,by="bjm_annot",all.x=T)
bkanno[!is.na(subclus_annot),bjm_annot:=subclus_annot]

## now merge the cluster assignments with annotations
bkcl <- merge.data.table(bkcl,bkanno,by.x="cl",by.y="clus")
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
hypx$banksyclus <- bkcl$bjm_annot

## label domains in the SPE
bksmooth[dualVMHARC4=="other",dualVMHARC4:="Other"]
bksmooth <- DataFrame(bksmooth,row.names=bksmooth$rn)[colnames(hypx),]
hypx$dom <- bksmooth$dualVMHARC4
```

### extract key metrics and metadata to a cellwise data table:
[4] "transcript_counts"                          
 [5] "control_probe_counts"                       
 [6] "control_codeword_counts"                    
 [7] "unassigned_codeword_counts"                 
 [8] "deprecated_codeword_counts"                 
[10] "cell_area"                                  
[11] "nucleus_area"                               
[12] "sample_id"                                  
[15] "nGenes"
"banksyclus" and "dom"
```{r}
hypx.allclus <- as.data.table(colData(hypx))[,.(transcript_counts,control_probe_counts,control_codeword_counts,unassigned_codeword_counts,deprecated_codeword_counts,cell_area,nucleus_area,sample_id,nGenes,banksyclus,dom)]
```

we can then filter this down later by removing rows (cells) that were assigned a DISCARD cluster or "VMH_4_DISCARD".

the first plot we want is a simple dotplot of each sample median for each metric of interest. the control probe fields all have medians of 0 or 1 so dont' bother with those for this first set of plots.
```{r}
meds <- melt(hypx.allclus,id.vars=c("sample_id","banksyclus","dom"))
meds <- meds[,median(value),by=c("sample_id","variable")]

meds <- meds[variable %in% c("transcript_counts","cell_area","nucleus_area","nGenes")]

meds[variable=="transcript_counts",variable:="Transcript\nCounts"]
meds[variable=="cell_area",variable:="Cell\nArea"]
meds[variable=="nucleus_area",variable:="Nucleus\nArea"]
meds[variable=="nGenes",variable:="Unique\nGenes"]

xenqcplots <- list()

xenqcplots[[1]] <- ggplot(meds,aes(x=1,y=V1))+
    geom_point(position=position_jitter(width=0.3))+
    scale_x_continuous(limits=c(0,2),expand=c(0,0))+
    facet_wrap(~variable,scales="free_y",nrow=1) +
    ylab("Sample Median (all cells)")+
    theme(axis.text.x = element_blank(),axis.ticks.x=element_blank(),axis.title.x=element_blank())

rm(meds)
```

then, make violin plots of all individual cell values for each metric, faceted by sample:
```{r}
cellvals <- melt(hypx.allclus,id.vars=c("sample_id","banksyclus","dom"))

# make metric names nice
cellvals[variable=="transcript_counts",variable:="Transcript\nCounts"]
cellvals[variable=="cell_area",variable:="Cell\nArea"]
cellvals[variable=="nucleus_area",variable:="Nucleus\nArea"]
cellvals[variable=="nGenes",variable:="Unique\nGenes"]
cellvals[variable=="control_probe_counts",variable:="Control\nProbe\nCounts"]
cellvals[variable=="control_codeword_counts",variable:="Control\nCodeword\nCounts"]
cellvals[variable=="unassigned_codeword_counts",variable:="Unassigned\nCodeword\nCounts"]
cellvals[variable=="deprecated_codeword_counts",variable:="Deprecated\nCodeword\nCounts"]

# violin with dashed lines for the 10/90%ile, solid for the 50%ile
# simplest approach is two violins: https://stackoverflow.com/questions/48936704/manual-specification-of-draw-quantile-argument-in-violin-plot-call-in-r

xenqcplots[[2]] <- ggplot(cellvals[variable %in% c("Transcript\nCounts","Cell\nArea","Nucleus\nArea","Unique\nGenes")],aes(x=sample_id,y=value))+
    geom_violin(aes(fill=sample_id),show.legend = FALSE,draw_quantiles = c(0.1,0.9),linetype="dashed",linewidth=0.5)+
    geom_violin(aes(fill=sample_id),show.legend=FALSE,draw_quantiles=0.5,linewidth=0.8,fill="transparent")+
    facet_wrap(~variable,scales="free_y",nrow = 1,)+
    theme(strip.text = element_text(size=6),axis.text.x=element_text(angle=90),title = element_text(size=14),plot.title.position = "plot")+
    ylab("Cell Values (All 41 Banksy clusters)")+
    xlab("Sample ID")
```

## for the negative probes, save a supp table instead (silly to make 870870 points that range integers 0-10).
```{r}
tabulaterange <- unique(c(hypx.allclus$control_probe_counts,hypx$control_codeword_counts,hypx$unassigned_codeword_counts,hypx$deprecated_codeword_counts))

negprobecttab <- as.data.frame(matrix(nrow=length(tabulaterange)*length(unique(hypx.allclus$sample_id)),ncol=6))
colnames(negprobecttab) <- c("sample_id","count","control_probe_counts","control_codeword_counts","unassigned_codeword_counts","deprecated_codeword_counts")

i<-1
for (i in c(1:length(unique(hypx.allclus$sample_id)))){
    cursamp <- unique(hypx.allclus$sample_id)[i]
    j<-1
    for (j in c(1:length(tabulaterange))){
        curind <- (i-1)*length(tabulaterange)+j
        curval <- tabulaterange[j]
        negprobecttab[curind,1] <- cursamp
        negprobecttab[curind,2] <- curval
        negprobecttab[curind,3] <- nrow(hypx.allclus[sample_id==cursamp&control_probe_counts==curval])
        negprobecttab[curind,4] <- nrow(hypx.allclus[sample_id==cursamp&control_codeword_counts==curval])
        negprobecttab[curind,5] <- nrow(hypx.allclus[sample_id==cursamp&unassigned_codeword_counts==curval])
        negprobecttab[curind,6] <- nrow(hypx.allclus[sample_id==cursamp&deprecated_codeword_counts==curval])
    }
}

negprobecttab <- as.data.table(negprobecttab)
setnames(negprobecttab,c(3:6),c("cells_w_X_control_probe_counts","cells_w_X_control_codeword_counts","cells_w_X_unassigned_codeword_counts","cells_w_X_deprecated_codeword_counts"))

fwrite(negprobecttab,"xenium_HYP/local_supptab_assembly/Ncells_with_negative_element_counts_0-max_41clusters.txt",sep='\t',quote=F,row.names=F)
```

### sigh, now do all this again after removing the DISCARDable clusters
```{r}
hypx.retain <- hypx.allclus[!(banksyclus %in% c("VMH_4_DISCARD","DISCARD"))]

meds <- melt(hypx.retain,id.vars=c("sample_id","banksyclus","dom"))
meds <- meds[,median(value),by=c("sample_id","variable")]

meds <- meds[variable %in% c("transcript_counts","cell_area","nucleus_area","nGenes")]

meds[variable=="transcript_counts",variable:="Transcript\nCounts"]
meds[variable=="cell_area",variable:="Cell\nArea"]
meds[variable=="nucleus_area",variable:="Nucleus\nArea"]
meds[variable=="nGenes",variable:="Unique\nGenes"]

xenqcplots[[3]] <- ggplot(meds,aes(x=1,y=V1))+
    geom_point(position=position_jitter(width=0.3))+
    scale_x_continuous(limits=c(0,2),expand=c(0,0))+
    facet_wrap(~variable,scales="free_y",nrow=1) +
    ylab("Sample Median (Cells in 33 Retained Clusters)")+
    theme(axis.text.x = element_blank(),axis.ticks.x=element_blank(),axis.title.x=element_blank())

rm(meds)


# violin with dashed lines for the 10/90%ile, solid for the 50%ile
# simplest approach is two violins: https://stackoverflow.com/questions/48936704/manual-specification-of-draw-quantile-argument-in-violin-plot-call-in-r

retaincells <- melt(hypx.retain,id.vars=c("sample_id","banksyclus","dom"))
rm(hypx.retain)
# make metric names nice
retaincells[variable=="transcript_counts",variable:="Transcript\nCounts"]
retaincells[variable=="cell_area",variable:="Cell\nArea"]
retaincells[variable=="nucleus_area",variable:="Nucleus\nArea"]
retaincells[variable=="nGenes",variable:="Unique\nGenes"]
                    
xenqcplots[[4]] <- ggplot(retaincells[variable %in% c("Transcript\nCounts","Cell\nArea","Nucleus\nArea","Unique\nGenes")],aes(x=sample_id,y=value))+
    geom_violin(aes(fill=sample_id),show.legend = FALSE,draw_quantiles = c(0.1,0.9),linetype="dashed",linewidth=0.5)+
    geom_violin(aes(fill=sample_id),show.legend=FALSE,draw_quantiles=0.5,linewidth=0.8,fill="transparent")+
    facet_wrap(~variable,scales="free_y",nrow = 1,)+
    theme(strip.text = element_text(size=6),axis.text.x=element_text(angle=90),title = element_text(size=14),plot.title.position = "plot")+
    ylab("Cell Values (33 retained Banksy clusters)")+
    xlab("Sample ID")
```


### lastly, make violins per QC metric for cells in each VMH cluster and each ARC cluster, by sample. (we don't need to do this twice because the discarding is being done by cluster label, which doesnt involve any of these clusters).
```{r}
vmhcellvals <- cellvals[variable %in% c("Transcript\nCounts","Cell\nArea","Nucleus\nArea","Unique\nGenes")&banksyclus %in% c("*FEZF1*-*LAMP5*-*NRGN*","Microglia","*HCRT*-*SLC6A3*-*TAC1*","*SLC17A6*-*CRHR2*")]
vmhcellvals[,banksyclus:=paste0("VMH ",banksyclus)]

xenqcplots[[5]] <- ggplot(vmhcellvals,aes(x=sample_id,y=value))+
    geom_violin(aes(fill=sample_id),show.legend = FALSE,draw_quantiles = c(0.1,0.9),linetype="dashed",linewidth=0.5)+
    geom_violin(aes(fill="sample_id"),show.legend=FALSE,draw_quantiles=0.5,linewidth=0.8,fill="transparent")+
    facet_grid(variable~banksyclus,scales="free_y")+
    theme(strip.text = element_markdown(size=6),axis.text.x=element_text(angle=90))+
    ylab("Cell Values")+
    xlab("Sample ID")

arccellvals <- cellvals[variable %in% c("Transcript\nCounts","Cell\nArea","Nucleus\nArea","Unique\nGenes")&banksyclus %in% arcvmhanno[subclus_domain=="ARC",subclus_annot]]
arccellvals[,banksyclus:=paste0("ARC ",banksyclus)]

xenqcplots[[6]] <- ggplot(arccellvals,aes(x=sample_id,y=value))+
    geom_violin(aes(fill=sample_id),show.legend = FALSE,draw_quantiles = c(0.1,0.9),linetype="dashed",linewidth=0.5)+
    geom_violin(aes(fill="sample_id"),show.legend=FALSE,draw_quantiles=0.5,linewidth=0.8,fill="transparent")+
    facet_grid(variable~banksyclus,scales="free_y")+
    theme(strip.text = element_markdown(size=6),axis.text.x=element_text(angle=90))+
    ylab("Cell Values")+
    xlab("Sample ID")
```

save a 3 page PDF with these in them (page 1: plot sets 1-4, indiv. page per cell type plots)
```{r}
pdf("manuscript_plots/Supplemental/SF-Negprobe_gene_metrics_per_Xenium_sample_Xenium_cell_and_per_cell_in_Xenium_VMH_or_ARC_clusters.pdf",width=8.5,height=11)

do.call("grid.arrange",list(arrangeGrob(grobs=xenqcplots[c(1:4)],ncol=1,nrow=4,heights = c(2,3.75,2,3.75))))

xenqcplots[[5]]

xenqcplots[[6]]

dev.off()
```

## for the negative probes, save a second supp table covering the 33 retained clusters instead of trying to plot individual cell valuess
```{r}
tabulaterange <- unique(retaincells[variable %in% c("control_probe_counts","control_codeword_counts","unassigned_codeword_counts","deprecated_codeword_counts"),value])

ret.negprobecttab <- as.data.frame(matrix(nrow=length(tabulaterange)*length(unique(retaincells$sample_id)),ncol=6))
colnames(ret.negprobecttab) <- c("sample_id","count","control_probe_counts","control_codeword_counts","unassigned_codeword_counts","deprecated_codeword_counts")

i<-1
for (i in c(1:length(unique(hypx.allclus$sample_id)))){
    cursamp <- unique(hypx.allclus$sample_id)[i]
    j<-1
    for (j in c(1:length(tabulaterange))){
        curind <- (i-1)*length(tabulaterange)+j
        curval <- tabulaterange[j]
        ret.negprobecttab[curind,1] <- cursamp
        ret.negprobecttab[curind,2] <- curval
        ret.negprobecttab[curind,3] <- nrow(hypx.allclus[sample_id==cursamp&control_probe_counts==curval])
        ret.negprobecttab[curind,4] <- nrow(hypx.allclus[sample_id==cursamp&control_codeword_counts==curval])
        ret.negprobecttab[curind,5] <- nrow(hypx.allclus[sample_id==cursamp&unassigned_codeword_counts==curval])
        ret.negprobecttab[curind,6] <- nrow(hypx.allclus[sample_id==cursamp&deprecated_codeword_counts==curval])
    }
}

ret.negprobecttab <- as.data.table(ret.negprobecttab)
setnames(ret.negprobecttab,c(3:6),c("cells_w_X_control_probe_counts","cells_w_X_control_codeword_counts","cells_w_X_unassigned_codeword_counts","cells_w_X_deprecated_codeword_counts"))

fwrite(ret.negprobecttab,"xenium_HYP/local_supptab_assembly/Ncells_with_negative_element_counts_0-max_RetainedClustersOnly.txt",sep='\t',quote=F,row.names=F)
```