---
title: "SF-PCA, Harmony, and UMAPs colored by dataset variables"
output: html_document
date: "2024-09-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(ggrastr)
library(scater)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## all of the myriad dimensionality reductions tried are saved in a separate RDS
## but we need the spe itself for spot metadata
```{r}
hyp2 <- readRDS("spatial_HYP/processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

hyp2drs <- readRDS("spatial_HYP/processed-data/04-feature_selection/02-hypfiltered_4featureset-pca-umap-harmonydefault-harmonylambdanull-mnn30_reducedDim_slot.RDS")

keepdrs <- grep(names(hyp2drs),pattern="nnsvg10|HARMONYlmbNA",value=T)
keepdrs <- grep(keepdrs,pattern="HARMONYdflt|mnn30",invert=T,value=T)

hyp2drs <- hyp2drs[keepdrs]
names(hyp2drs)
names(hyp2drs) <- c("Unadjusted PCA","Unadjusted UMAP","Harmony PCA","Harmony UMAP")



```


## replace sample_ids with manuscript sample_ids, make into data tables, change a couple of column names for plot readability

```{r}
mids <- fread("standardized_sampleids_for_plotting.txt")

hyp2$sample_id <- mids[match(hyp2$sample_id,sample_id),manuscript_id]

## get variance explained for PCs 1-3 for the uncorrected PCA.
varexps <- attr(hyp2drs[[1]],"percentVar")[1:3]

hyp2drs <- lapply(hyp2drs,FUN=function(d){
    pd <- as.data.table(d,keep.rownames=T)
    pd <- merge.data.table(as.data.table(colData(hyp2),keep.rownames=T),pd,by="rn")
    setnames(pd,c("sum_umi","sum_gene","brnum","PMI","bmi"),c("Spot_UMIs","Spot_Unique_Genes","Donor","Postmortem_Interval","Body_Mass_Index"))
    return(pd)
})

## fix harmony names
names(hyp2drs[[3]]) <- gsub(names(hyp2drs[[3]]),pattern="^HARMONYlmbna_nnsvg10_(.*)$",replacement=paste0("Harmony","\\1"))
# 

# get variables to plot
# names(hyp2drs[[1]])

# variables to plot by: 
colorvars <- c("sample_id", "Spot_UMIs", "Spot_Unique_Genes", "expr_chrM_ratio", "Donor", "AgeDeath", "Sex", "Postmortem_Interval", "best_RIN_PFC", "Body_Mass_Index", "slide")
```


## uncorrected PCAs
```{r}
pcpairs <- list(c("PC1","PC2"),c("PC2","PC3"),c("PC1","PC3"))
x <- hyp2drs[[1]]
pcp <- lapply(colorvars,FUN=function(v){

      pa <- ggplot(x,aes_string(x=pcpairs[[1]][1],y=pcpairs[[1]][2],color=v))+
         geom_point(size=0.15,stroke=0.15) +
         ggtitle(paste0(names(hyp2drs)[1]," colored by\n",v))+
         xlab(paste0(pcpairs[[1]][[1]]," (",round(varexps[1]),"% variance explained)"))+
         ylab(paste0(pcpairs[[1]][[2]]," (",round(varexps[2]),"% variance explained)"))
      
        pb <- ggplot(x,aes_string(x=pcpairs[[2]][1],y=pcpairs[[2]][2],color=v))+
           geom_point(size=0.15,stroke=0.15) +
           ggtitle(paste0(names(hyp2drs)[1]," colored by\n",v))+
           xlab(paste0(pcpairs[[2]][[1]]," (",round(varexps[2]),"% variance explained)"))+
           ylab(paste0(pcpairs[[2]][[2]]," (",round(varexps[3]),"% variance explained)"))
        
        pc <- ggplot(x,aes_string(x=pcpairs[[3]][1],y=pcpairs[[3]][2],color=v))+
           geom_point(size=0.15,stroke=0.15) +
           ggtitle(paste0(names(hyp2drs)[1]," colored by\n",v))+
           xlab(paste0(pcpairs[[3]][[1]]," (",round(varexps[1]),"% variance explained)"))+
           ylab(paste0(pcpairs[[3]][[2]]," (",round(varexps[3]),"% variance explained)"))
        
    retlist <- list(pa,pb,pc)
    return(retlist)
})

i<-1
pcplt <- list()
for (i in c(1:length(pcp))){
    pcplt <- c(pcplt,pcp[[i]])
}

rm(pcp,pcpairs,x)
```

## uncorrected UMAPs
```{r}
x <- hyp2drs[[2]]
ump <- lapply(colorvars,FUN=function(v){
        pa <- ggplot(x,aes_string(x="UMAP1",y="UMAP2",color=v)) + geom_point(size=0.15,stroke=0.15) + ggtitle(paste0(names(hyp2drs)[2]," colored by\n",v))
    return(pa)
})
rm(x)
```

uncorrected haromnies
```{r}
hpairs <- list(c("Harmony1","Harmony2"),c("Harmony2","Harmony3"),c("Harmony1","Harmony3"))
x <- hyp2drs[[3]]
hp <- lapply(colorvars,FUN=function(v){
        pa <- ggplot(x,aes_string(x=hpairs[[1]][1],y=hpairs[[1]][2],color=v)) + geom_point(size=0.15,stroke=0.15) + ggtitle(paste0(names(hyp2drs)[3]," colored by\n",v))
        pb <- ggplot(x,aes_string(x=hpairs[[2]][1],y=hpairs[[2]][2],color=v)) + geom_point(size=0.15,stroke=0.15) + ggtitle(paste0(names(hyp2drs)[3]," colored by\n",v))
        pc <- ggplot(x,aes_string(x=hpairs[[3]][1],y=hpairs[[3]][2],color=v)) + geom_point(size=0.15,stroke=0.15) + ggtitle(paste0(names(hyp2drs)[3]," colored by\n",v))

    retlist <- list(pa,pb,pc)
    return(retlist)
})

i<-1
hplt <- list()
for (i in c(1:length(hp))){
    hplt <- c(hplt,hp[[i]])
}

rm(hp,hpairs,x)
```

## Harmony UMAPS
```{r}
x <- hyp2drs[[4]]
hump <- lapply(colorvars,FUN=function(v){
        pa <- ggplot(x,aes_string(x="UMAP1",y="UMAP2",color=v)) + geom_point(size=0.15,stroke=0.15) + ggtitle(paste0(names(hyp2drs)[4]," colored by\n",v))
    return(pa)
})
rm(x)
```

plot em
```{r}
pcplt <- lapply(pcplt,FUN=function(x){rasterize(x,dpi=300,dev="cairo_png")})
ump <- lapply(ump,FUN=function(x){rasterize(x,dpi=300,dev="cairo_png")})
hplt <- lapply(hplt,FUN=function(x){rasterize(x,dpi=300,dev="cairo_png")})
hump <- lapply(hump,FUN=function(x){rasterize(x,dpi=300,dev="cairo_png")})

pdf("manuscript_plots/Supplemental/SF-PCA_UMAP_Harmony_UMAP_coloredByMetadataVars.pdf",height=23,width=36)
do.call("grid.arrange",c(pcplt,ncol=5))
do.call("grid.arrange",c(ump,ncol=3))
do.call("grid.arrange",c(hplt,ncol=5))
do.call("grid.arrange",c(hump,ncol=3))
dev.off()
```


repro inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.38.0         parallelly_1.38.0          
 [3] colorout_1.3-0.2            scater_1.32.0              
 [5] scuttle_1.14.0              ggrastr_1.0.2              
 [7] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
 [9] SummarizedExperiment_1.34.0 Biobase_2.64.0             
[11] GenomicRanges_1.56.1        GenomeInfoDb_1.40.1        
[13] IRanges_2.38.1              S4Vectors_0.42.1           
[15] BiocGenerics_0.50.0         MatrixGenerics_1.16.0      
[17] matrixStats_1.3.0           gridExtra_2.3              
[19] ggplot2_3.5.1               data.table_1.15.4          
[21] rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1          viridisLite_0.4.2         dplyr_1.1.4              
 [4] vipor_0.4.7               farver_2.1.2              viridis_0.6.5            
 [7] fastmap_1.2.0             digest_0.6.36             rsvd_1.0.5               
[10] lifecycle_1.0.4           magrittr_2.0.3            compiler_4.4.1           
[13] tools_4.4.1               utf8_1.2.4                yaml_2.3.9               
[16] knitr_1.48                S4Arrays_1.4.1            labeling_0.4.3           
[19] here_1.0.1                DelayedArray_0.30.1       abind_1.4-5              
[22] withr_3.0.0               grid_4.4.1                fansi_1.0.6              
[25] beachmat_2.20.0           colorspace_2.1-0          scales_1.3.0             
[28] cli_3.6.3                 rmarkdown_2.27            crayon_1.5.3             
[31] generics_0.1.3            rstudioapi_0.16.0         httr_1.4.7               
[34] rjson_0.2.21              DelayedMatrixStats_1.26.0 ggbeeswarm_0.7.2         
[37] zlibbioc_1.50.0           XVector_0.44.0            vctrs_0.6.5              
[40] Matrix_1.7-0              jsonlite_1.8.8            BiocSingular_1.20.0      
[43] BiocNeighbors_1.22.0      ggrepel_0.9.5             irlba_2.3.5.1            
[46] beeswarm_0.4.0            magick_2.8.3              glue_1.7.0               
[49] codetools_0.2-20          gtable_0.3.5              UCSC.utils_1.0.0         
[52] ScaledMatrix_1.12.0       munsell_0.5.1             tibble_3.2.1             
[55] pillar_1.9.0              htmltools_0.5.8.1         GenomeInfoDbData_1.2.12  
[58] R6_2.5.1                  sparseMatrixStats_1.16.0  rprojroot_2.0.4          
[61] evaluate_0.24.0           lattice_0.22-6            png_0.1-8                
[64] Rcpp_1.0.13               SparseArray_1.4.8         xfun_0.45                
[67] pkgconfig_2.0.3          
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-10-07
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                  1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   beachmat               2.20.0  2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors          1.22.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   DelayedArray           0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi                  1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr              * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5   2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   png                    0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
   S4Arrays               1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   ScaledMatrix           1.12.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.32.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray            1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector                0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────