---
title: "SF-VisXen_marker-and-sexDE-registration_to_Yao23mouseHYP-regionmatched-subclasses"
author: "Bernie Mulvey"
date: "2024-11-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(ggtext)
library(Polychrome)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


# Visium-spatial registration to Yao '23. just do domain-pseudobulked ms domain subclasses for now.
```{r}
spatregs <- readRDS("spatial_HYP/processed-data/10-Spatial Registration/02g-sphyp-k15-20-31-15clps_ABAms-47subclass-47withVMHARCclps-VMHARCsupertypes.RDS")

sexregs <- readRDS("spatial_HYP/processed-data/12-Mouse comparisons/03-Yao23-VMHARCcollapsedSubclasses-MvFsexDE-tstat_correls_to_k15-20-31-15collapsed_bytopNhumanDEGs.RDS")

## get collapsed VMHARC at k15 results
spatregs <- spatregs[["subclass_collapsed_to_sphypk15_collapsed"]]
sexregs <- sexregs[["BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH"]]
sexregs[,pcc:=NULL]
## get the spatial reg table into the same collapsed format as the sex reg table

j <- 1
for (j in c(1:length(spatregs))){
    newdat <- copy(spatregs[[j]])
    newdat[,ngenes:=as.numeric(gsub(names(spatregs)[j],pattern="^ABA.*_top_(.*)_sphyp.*$",replacement="\\1"))]
    if(j==1){
        curres <- copy(newdat)
    }
    else{curres <- rbind(curres,newdat)}
    rm(newdat)
}
spatregs <- curres
rm(curres,j)

spatregs <- spatregs[sphyp_domain %in% c("ARC","VMH")]
spatregs[,sphyp_domain:=paste0("v",sphyp_domain," Markers")]
keepcols <- c("sphyp_domain","ngenes",grep(names(spatregs),pattern="VMHall|ARHall",value=T))
spatregs <- spatregs[,..keepcols]
spatregs <- melt.data.table(spatregs,id.vars=c("sphyp_domain","ngenes"))
setnames(spatregs,c("variable","ngenes","value"),c("mouse_pseudobulk_domain","top_n_human","spearmancor"))
spatregs[mouse_pseudobulk_domain=="ARHall",mouse_pseudobulk_domain:="Ms ARC Aggregated Subclasses"]
spatregs[mouse_pseudobulk_domain=="VMHall",mouse_pseudobulk_domain:="Ms VMH Aggregated Subclasses"]

sexregs <- sexregs[human_domain %in% c("ARC","VMH")]
sexregs[,human_domain:=paste0("v",human_domain," Sex DE")]
setnames(sexregs,"human_domain","sphyp_domain")
sexregs[mouse_pseudobulk_domain=="VMH",mouse_pseudobulk_domain:="Ms VMH Aggregated Subclasses"]
sexregs[mouse_pseudobulk_domain=="ARC",mouse_pseudobulk_domain:="Ms ARC Aggregated Subclasses"]

jtres <- rbind(spatregs,sexregs)

jtres[sphyp_domain %in% grep(sphyp_domain,pattern="VMH",value=T),domain:="VMH"]
jtres[sphyp_domain %in% grep(sphyp_domain,pattern="ARC",value=T),domain:="ARC"]

## subset to vmh-vmh and arc-arc
jtres <- rbind(jtres[domain=="VMH"&mouse_pseudobulk_domain %in% unique(grep(mouse_pseudobulk_domain,pattern="VMH",value=T))],
               jtres[domain=="ARC"&mouse_pseudobulk_domain %in% unique(grep(mouse_pseudobulk_domain,pattern="ARC",value=T))])



# pdf("manuscript_plots/Supplemental/SF-VisDomain-Marker-and-SexDE_registration_to_aggregated-Yao23-MouseHYP-domainsubclasses.pdf",height=11,width=8.5)
plts <- list()
plts[[1]] <- ggplot(jtres,aes(x=log10(top_n_human),y=spearmancor,col=sphyp_domain))+
    geom_line()+
    facet_wrap(~mouse_pseudobulk_domain,scales="free",nrow = 1)+
    labs(y="Spearman rho (Human-<br>Ms Marker/Sex DE *t*)",x="log10 (Top n Human DEGs or Markers)",col="Human Domain Genes")+
    theme(strip.text = element_text(size=10),strip.background = element_blank())

rm(spatregs,sexregs,jtres,keepcols)
```


####### ####### #######

## xenium clusters to subclasses
```{r}
clustids <- fread("manuscript_plot_code/xARCxVMH_cluster_detailedlabs_andplotformatnames.txt")

###
xenspat <- readRDS("xenium_HYP/processed-data/10_Spatial_Registration/01-NonSampleSpecific_cluster_reg_to_Yao23_mouseHyp_subclasses.RDS")
xenspat <- xenspat[[1]]
xenspat <- rbindlist(xenspat,idcol="ngenes")
xenspat[,ngenes:=as.numeric(gsub(ngenes,pattern="^.*_top_(.*)_sphyp.*$",replacement="\\1"))]

# this step conveniently also filters down to xVMH and xARC neuron types
xenspat <- merge.data.table(xenspat,clustids[,.(cluster_id,plotclusname,subclus_domain)],by.x="sphyp_domain",by.y="cluster_id")
# subset to ARH/VMH mouse subclasses
keepn <- c("plotclusname","ngenes","subclus_domain",grep(names(xenspat),pattern="ARH|VMH",value=T))
xenspat <- xenspat[,..keepn]
# melt
## append xenium domain to the front of the xenium cluster name 
xenspat[,analysis:=paste0(plotclusname," Markers")]
xenspat <- melt.data.table(xenspat,id.vars=c("plotclusname","ngenes","subclus_domain","analysis"))
```

### setup continued 
### xenium sex DE - already melted into same layout as above
```{r}
xensex <- readRDS("xenium_HYP/processed-data/10_Spatial_Registration/03-Yao23-subclass-MvFsexDE-tstat_correls_to_celltypes-in-xenARCVMHdomains.RDS")
###

xensex[,analysis_domain:=gsub(human_domain_celltype,pattern="^(...)_.*$",replacement="\\1")]

## match plotting IDs
xensex[,human_domain_celltype:=gsub(human_domain_celltype,pattern="^..._(.*)$",replacement="\\1")]
xensex <- merge.data.table(xensex,clustids[,.(bjm_annot,plotclusname,subclus_domain)],by.x="human_domain_celltype",by.y="bjm_annot")

## subset to ARC type - ARC type and VMH type - VMH type correlations
xensex <- xensex[
   (mouse_subclass %in% grep(mouse_subclass,pattern="VMH",value=T)&analysis_domain=="VMH"&subclus_domain=="VMH")|(mouse_subclass %in% grep(mouse_subclass,pattern="ARH",value=T) & analysis_domain=="ARC"&subclus_domain=="ARC")]


xensex[,analysis:=paste0(plotclusname," Sex DE")]

setnames(xensex,c("mouse_subclass","spearmancor","top_n_human"),c("variable","value","ngenes"))

## remove temporary which-domain-was-the-cell-type-being-analyzed-in column
xensex[,analysis_domain:=NULL]

xenjt <- rbind(xenspat,xensex[,names(xenspat),with=F])

# tidy up mouse subclass names
xenjt[,variable:=gsub(variable,pattern="_",replacement=" ")]
xenjt[,variable:=gsub(variable,pattern="^x... (.*)$",replacement="\\1")]
xenjt[,variable:=gsub(variable,pattern="ARH",replacement="ARC")]
## remove additional areas attributed to some ARC cell groups
xenjt[,variable:=gsub(variable,pattern="PVa |PVp |TU ",replacement="")]
## make cell type markers of mouse cells italic in output
xenjt[,variable:=gsub(variable,pattern="(Tbx3|Six3|Six6|Otp|Fezf1|Tbx3|Nr5a1)",replacement=paste0("*","\\1","*"))]

# make primary transmitter type more presentable
xenjt[,variable:=gsub(variable,pattern="Gaba",replacement="GABAergic")]
xenjt[,variable:=gsub(variable,pattern="Glut",replacement="Glutamatergic")]
xenjt[,variable:=gsub(variable,pattern="Dopa",replacement="DAergic")]

## lastly, append the name of the xenium domain to the xenium clusters (to be used as  plot titles)
xenjt[,analysis:=paste0("x",subclus_domain," ",analysis)]

## but get rid of any redundant xVMH xVMH or xARC xARC
xenjt[,analysis:=gsub(analysis,pattern="xVMH xVMH",replacement="xVMH")]
xenjt[,analysis:=gsub(analysis,pattern="xARC xARC ",replacement="xARC ")]

## format the analysis column to be used as plot titles
xenjt[analysis %in% grep(analysis,pattern="POMC|SLC17A7|GHRH",value=T),analysis:=gsub(analysis,pattern="(Mixed|\\*GHRH*)",replacement=paste0("<br>","\\1"))]
xenjt[,analysis:=gsub(analysis,pattern="^(.*)<br>Mixed (.*)$",replacement=paste0("\\1","<br>Mixed (","\\2",")"))]


xenjt[analysis %in% grep(analysis,pattern="POMC|SLC17A7|GHRH",value=T,invert=T),analysis:=gsub(analysis,pattern=" (Marker)",
                       replacement=paste0("<br>","\\1"))]
xenjt[analysis %in% grep(analysis,pattern="POMC|SLC17A7|GHRH",value=T,invert=T),analysis:=gsub(analysis,pattern=" (Sex)",
                       replacement=paste0("<br>","\\1"))]


## make ARC table
arcreg <- xenjt[variable %in% grep(variable,pattern="ARC",value=T)&subclus_domain=="ARC"]

## make vmh table
vmhreg <- xenjt[variable %in% grep(variable,pattern="VMH",value=T)&subclus_domain=="VMH"]

rm(xensex,xenspat,clustids,keepn)
```

## make unified and shared legend for arc-arc and vmh-vmh plots with one 
## use xenium celltype palette for colors here
```{r}
pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")
arcpal <- pals[["XenARC"]][c(1:3,5)]
vmhpal <- pals[["XenVMH"]][c(1,4)]

names(arcpal) <- unique(arcreg$variable)
names(vmhpal) <- unique(vmhreg$variable)
jtpal <- c(arcpal,vmhpal)

# helper function to extract a legend and use as grob: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}

legdat <- rbind(arcreg,vmhreg)

## arrange so that ARCs are first, VMHs second. put palette in same factor order to be sure it comes out right
levs <- c(grep(unique(legdat$variable),pattern="ARC",value=T),grep(unique(legdat$variable),pattern="VMH",value=T))
legdat[,variable:=factor(variable,levels=levs)]
jtpal <- jtpal[levs]

legplt <- ggplot(legdat,aes(x=variable,y=ngenes,col=variable))+
    geom_line()+
    scale_color_manual(values=jtpal)+
    facet_grid(~plotclusname)+
    labs(col="Mouse Subclass")+
    guides(color=guide_legend(ncol=2))+
    theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 9), legend.title = element_text(size = 9), legend.text = element_markdown(size = 8), title = element_text(size = 8))

legout <- extract_legend(legplt)
dev.off()

rm(legdat,legplt,pals,levs)
```

make plots in pairs of marker-sex by human cluster
```{r}
# plot shared numbers of genes tested: 50, 100, 150, 200, 250, 280
arcreg <- arcreg[ngenes %in% c(50,100,150,200,250,280)]
vmhreg <- vmhreg[ngenes %in% c(50,100,150,200,250,280)]

## continue plotting in the list that we dropped the Visium plot pair into several blocks ago

# go thru making arc plots first
i<-1
for (i in c(1:length(unique(arcreg$plotclusname)))){
   curclus <- unique(arcreg$plotclusname)[i]
   
   ## marker plot 
   mkdat <- arcreg[plotclusname==curclus&analysis %in% grep(analysis,pattern="Markers",value=T)]
   plts[[length(plts)+1]] <- ggplot(mkdat,aes(x=ngenes,y=value,col=variable))+
       geom_line()+
       scale_color_manual(values=jtpal)+
       guides(color="none")+
       ggtitle(unique(mkdat$analysis))+
       labs(y="Spearman rho (Human-<br>Ms Marker *t*))",x="Top n Xenium Markers")
   
   ## sex DE plot
   sexdat <- arcreg[plotclusname==curclus & analysis %in% grep(analysis,pattern="Sex DE",value=T)]
   plts[[length(plts)+1]] <-
       ggplot(sexdat,aes(x=ngenes,y=value,col=variable))+
       geom_line()+
       scale_color_manual(values=jtpal)+
       guides(color="none")+
       ggtitle(unique(sexdat$analysis))+
       labs(y="Spearman rho(Human-<br>Ms Sex DE *t*)",x="Top n Xenium DEGs")
}

## then vmh plots
i<-1
for (i in c(1:length(unique(vmhreg$plotclusname)))){
   curclus <- unique(vmhreg$plotclusname)[i]
   
   ## marker plot 
   mkdat <- vmhreg[plotclusname==curclus&analysis %in% grep(analysis,pattern="Markers",value=T)]
   plts[[length(plts)+1]] <- ggplot(mkdat,aes(x=ngenes,y=value,col=variable))+
       geom_line()+
       scale_color_manual(values=jtpal)+
       guides(color="none")+
       ggtitle(unique(mkdat$analysis))+
       labs(y="Spearman rho (Human-<br>Ms Marker *t*))",x="Top n Xenium Markers")
   
   ## sex DE plot
   sexdat <- vmhreg[plotclusname==curclus & analysis %in% grep(analysis,pattern="Sex DE",value=T)]
   plts[[length(plts)+1]] <-
       ggplot(sexdat,aes(x=ngenes,y=value,col=variable))+
       geom_line()+
       scale_color_manual(values=jtpal)+
       guides(color="none")+
       ggtitle(unique(sexdat$analysis))+
       labs(y="Spearman rho(Human-<br>Ms Sex DE *t*)",x="Top n Xenium DEGs")
}

plts <- lapply(plts,FUN=function(p){
   q <- p+theme(axis.text.y=element_text(size=8),axis.text.x=element_text(size=8),axis.title.y=element_markdown(size=9),axis.title.x=element_markdown(size=9),plot.title=element_markdown(size=10,hjust=0.5),strip.text=element_text(size=10))
return(q)
})

## append legout as the last plot of this so it appears in the incomplete row
plts[[length(plts)+1]] <- legout
```

## save. legend still needs to be scooched rightward in illustrator but no big deal there
```{r}
pdf("manuscript_plots/Supplemental/SF-VisiumXenium_Yao23HYPscseq_registration.pdf",height=11,width=8,onefile=FALSE)
do.call("grid.arrange",list(arrangeGrob(grobs=plts[1],nrow=1),
                            arrangeGrob(grobs=plts[2:length(plts)],ncol=4),
                        nrow=2,
                        heights=c(2.5,8.5)))
dev.off()
```


```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-0.2  Polychrome_1.5.1  ggtext_0.1.2      gridExtra_2.3    
[5] ggplot2_3.5.1     data.table_1.15.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] gtable_0.3.5         dplyr_1.1.4          compiler_4.4.1      
 [4] tidyselect_1.2.1     Rcpp_1.0.13          stringr_1.5.1       
 [7] xml2_1.3.6           scales_1.3.0         yaml_2.3.9          
[10] fastmap_1.2.0        here_1.0.1           R6_2.5.1            
[13] commonmark_1.9.1     labeling_0.4.3       generics_0.1.3      
[16] knitr_1.48           datasets_4.4.1       tibble_3.2.1        
[19] munsell_0.5.1        rprojroot_2.0.4      pillar_1.9.0        
[22] R.utils_2.12.3       utf8_1.2.4           stringi_1.8.4       
[25] xfun_0.45            cli_3.6.3            withr_3.0.0         
[28] magrittr_2.0.3       digest_0.6.36        grid_4.4.1          
[31] gridtext_0.1.5       rstudioapi_0.16.0    markdown_1.13       
[34] lifecycle_1.0.4      R.oo_1.26.0          R.methodsS3_1.8.2   
[37] vctrs_0.6.5          scatterplot3d_0.3-44 evaluate_0.24.0     
[40] glue_1.7.0           farver_2.1.2         sessioninfo_1.2.2   
[43] fansi_1.0.6          colorspace_2.1-0     rmarkdown_2.27      
[46] tools_4.4.1          pkgconfig_2.0.3      htmltools_0.5.8.1   
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-11-11
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package       * version date (UTC) lib source
 P cli             3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout      * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace      2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   commonmark      1.9.1   2024-01-30 [1] CRAN (R 4.4.0)
   data.table    * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest          0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr           1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate        0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi           1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver          2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap         1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics        0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggplot2       * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggtext        * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
   glue            1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra     * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gridtext        0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
   gtable          0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here            1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools       0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr           1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling        0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle       1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr        2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   markdown        1.13    2024-06-04 [1] CRAN (R 4.4.0)
   munsell         0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   pillar          1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig       2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   Polychrome    * 1.5.1   2022-05-03 [1] CRAN (R 4.4.0)
   R.methodsS3     1.8.2   2022-06-13 [1] CRAN (R 4.4.0)
   R.oo            1.26.0  2024-01-24 [1] CRAN (R 4.4.0)
   R.utils         2.12.3  2023-11-18 [1] CRAN (R 4.4.0)
   R6              2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp            1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
 P rlang         * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown       2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot       2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi      0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales          1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   scatterplot3d   0.3-44  2023-05-05 [1] CRAN (R 4.4.0)
   sessioninfo     1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   stringi         1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
   stringr         1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
   tibble          3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect      1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8            1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs           0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr           3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun            0.45    2024-06-16 [1] CRAN (R 4.4.0)
   xml2            1.3.6   2023-12-04 [1] CRAN (R 4.4.0)
   yaml            2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────










#### UNUSED: make plots of visium domains x individual mouse subclasses/supertypes
```{r}
# spatregs <- readRDS("spatial_HYP/processed-data/10-Spatial Registration/02g-sphyp-k15-20-31-15clps_ABAms-47subclass-47withVMHARCclps-VMHARCsupertypes.RDS")
# 
# ## get collapsed VMHARC at k15 results
# 
# jointrestabs <- list()
# i<-1
# for (i in c(1:length(spatregs))){
#     j <- 1
#     for (j in c(1:length(spatregs[[i]]))){
#         newdat <- copy(spatregs[[i]][[j]])
#         newdat[,ngenes:=as.numeric(gsub(names(spatregs[[i]])[j],pattern="^ABA.*_top_(.*)_sphyp.*$",replacement="\\1"))]
#         newdat[,comparison:=gsub(names(spatregs[[i]])[j],pattern="^(.*)_top.*$",replacement="\\1")]
#         if(j==1){
#             curres <- copy(newdat)
#         }
#         else{curres <- rbind(curres,newdat)}
#         rm(newdat)
#     }
#     jointrestabs[[i]] <- copy(curres)
#     names(jointrestabs)[i] <- unique(curres$comparison)
#     rm(j,curres)
# }
# rm(i)
### save these as Supp Tabs since we made them - done 111024
# i<-1
# for (i in c(1:12)){
#     fwrite(jointrestabs[[i]],paste0("spatial_HYP/local_supptab_assembly/Yao23 Spatreg sheets/",names(jointrestabs)[i],".txt"),sep='\t',quote=F)
# }
```


## similarly for sex DE regulatoin

### setup: VMH domain to VMH clusters / ARC domain to ARC clusters spatreg values as a function of n genes
```{r}
# plotspatregtabs <- mapply(reg=jointrestabs[c(10:12)],regpar=names(jointrestabs)[c(10:12)],SIMPLIFY=FALSE,FUN=function(reg,regpar){
#     reg <- reg[sphyp_domain %in% c("VMH","ARC")]
#     keepcols <- c("ngenes","sphyp_domain",grep(colnames(reg),pattern="VMH|ARH",value=T))
#     plotspatreg <- melt(reg[,..keepcols],id.vars=c("ngenes","sphyp_domain"),value.name="sprrho")
#     
#     plotspatreg[,sphyp_domain:=paste0("v",sphyp_domain)]
#     plotspatreg[,variable:=as.character(variable)]
#     
#     arcs <- grep(unique(plotspatreg$variable),pattern="ARH",value=T)
#     vmhs <- grep(unique(plotspatreg$variable),pattern="VMH",value=T)
#     
#     arcplots <- plotspatreg[sphyp_domain=="vARC"&variable%in%arcs]
#     vmhplots <- plotspatreg[sphyp_domain=="vVMH"&variable%in%vmhs]
# 
#     retl <- list(vARC=arcplots,vVMH=vmhplots)
#     names(retl) <- paste0(names(retl),"_",regpar)
#     
#     return(retl)
# })
# 
# plotspatregtabs2 <- list()
# i<-1
# comps <- c("Mouse HYPwide Subclass","Ms HYPwide, Domain Subclasses Combined","Ms Domain Supertypes")
# for (i in c(1:length(plotspatregtabs))){
#     curcomp <- comps[i]
#     j<-1
#     for (j in c(1:length(plotspatregtabs[[i]]))){
#       plotspatregtabs2[[length(plotspatregtabs2)+1]] <- plotspatregtabs[[i]][[j]]
#       curdom <- gsub(names(plotspatregtabs[[i]])[j],pattern="^(v...)_.*$",replacement="\\1")
#       names(plotspatregtabs2)[length(plotspatregtabs2)] <- paste0(curdom," to ",curcomp)
#     }
#     rm(j)
# }
# rm(i,curdom,curcomp,comps,plotspatregtabs,jointrestabs,spatregs)
```
# 
# ## use existing palettes to pull colors for these plots
```{r}
# pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")
# 
# vsubclpal <- pals[[2]][c("*CRHR2*","*LAMP5*")] # two purplishes
# asubclpal <- pals[[3]][names(pals[[3]])!="*SLC17A7*-Glia Mixed"] # three greenishes and a pink; make the pink a 6-digit hex value for polychrome compatibility
# asubclpal[4] <- "#DD00BB"
# 
# names(vsubclpal) <- NA
# names(asubclpal) <- NA
# 
# ## construct manual palettes for vmh supertypes and arc supertypes (7 and 10 respectively)
# set.seed(4042)
# vsuperpal <- Polychrome::createPalette(7,seedcolors=vsubclpal,range = c(25,75))
# names(vsuperpal) <- NA
# set.seed(4042)
# asuperpal <- Polychrome::createPalette(10,seedcolors=asubclpal,range = c(25,75))
# names(asuperpal) <- NA
# rm(pals)
```
# 
# helper function to extract a legend and use as grob: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
# extract_legend <- function(my_ggp) {
#   step1 <- ggplot_gtable(ggplot_build(my_ggp))
#   step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
#   step3 <- step1$grobs[[step2]]
#   return(step3)
# }
```
# 
# ## marker registration plots
```{r}
# arcplts <- plotspatregtabs2[grep(names(plotspatregtabs2),pattern="vARC",value=T)]
# # change variable ARHall in the collapsed analysis to Pseudobulked<br>Ms ARC Subclasses
# vmhplts <- plotspatregtabs2[grep(names(plotspatregtabs2),pattern="vVMH",value=T)]
# legtitles <- c("Mouse Subclass","Mouse Domain","Mouse Supertype")
# 
# # varc
# apals <- list(asub=asubclpal,pb=asubclpal[1],sup=asuperpal)
# names(apals[[2]]) <- "ARHall"
# 
# varc.mk.plt <- mapply(d=arcplts,pt=names(arcplts),lt=legtitles,p=apals,SIMPLIFY=FALSE,FUN=function(d,pt,lt,p){
#   if(length(unique(d$variable))!=1){
#     names(p) <- unique(d$variable)
#   }
#   
#   ggplot(d,aes(x=log10(ngenes),y=sprrho,col=variable))+
#     geom_line()+
#     scale_color_manual(values=p)+
#     ggtitle(pt)+
#     labs(col=lt)+
#     scale_y_continuous(limits=c(min(d$sprrho),max(d$sprrho)))+
#     theme(axis.title.x = element_blank(), axis.text.y = element_text(size = 8),axis.text.x=element_text(size=8,angle=90,vjust=1,hjust=0.5), axis.title.y = element_text(size=9),legend.title = element_text(size=8,hjust=0.5),legend.text=element_text(size=8),title = element_text(size=10))
# })
```