---
title: "SF-vVMH_vARC_Volcanos_wAllosomes"
author: "Bernie Mulvey"
date: "2025-11-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())
library(data.table) # Preferred data manipulation package
library(ggplot2) # Dependency for several plotting functions
library(ggtext) # more character types in ggplots
library(ggrepel) # text labels in volcano plots
library(ggrastr) # avoid making raster images one can't even open
library(gridExtra)
library(viridis) # palettes
library(Polychrome) # better palettes

## rstudio GUI tweaks
# require(colorout)
# ColorOut()
# options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw() + theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 9), plot.title = element_blank(), strip.text = element_text(size = 10), legend.text = element_text(size = 8), legend.title = element_text(size = 8, hjust = 0.5)))


## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### sex DE results
```{r}
visde <- fread("spatial_HYP/processed-data/09-Sex DE/01c-voomLmFit_nnsvg10-HmnylmbNA-BS-15-VMHARCclpsd.txt")
### exclude mito genes (need to append this info)
genelut <- fread("spatial_HYP/raw-data/USEFOR10X_ensg_lut_hg38_fromEns98_111123.txt")
genelut <- unique(genelut[,.(ensembl_gene_id,chromosome_name)])
setnames(genelut,c("ens","chr"))
visde <- merge.data.table(visde,genelut,by.x="gene_id",by.y="ens")
visde <- visde[!(chr %in% c("MT"))]

## clean up
rm(genelut)
```


## palette set up (shared)
```{r}
pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")

pal <- pals[["Domains"]]

 
### also load manuscript plotting ids
mscriptids <- fread("standardized_sampleids_for_plotting.txt")
```


Panel A, option 2: Visium VMH volcano
```{r}
visdeplt <- visde[assay=="VMH"]
## points to label
visdeplt <- visdeplt[assay=="VMH"&gene_name %in% c("XIST","DDX3Y"),label:=gene_name]

## no fdr sig genes here; min FDR 0.07 and corresponding p value is 6.904826e-05, so conveniently we can approximate FDR ~= p*10^-3 at this range, and use p 5e-05 to draw the line

p <- ggplot(visdeplt[assay=="VMH"], aes(y = -log10(P.Value), x = logFC,col=assay))+
    ggtitle("VMH (Visium)")+
    scale_color_manual(values=pal[1]) +
    geom_hline(yintercept = -log10(0.05),
        linetype = "dashed",
        color = "black",
        linewidth=0.25) +
    geom_hline(yintercept = -log10(5e-5),
        linetype = "dashed",
        color = "red",
        linewidth=0.25) +
    geom_text_repel(seed = 23942,aes(y = -log10(P.Value),
            x = logFC,
            label = ifelse(is.na(label), "", label)),
        min.segment.length = 3,
        size = 3,
        show.legend=FALSE,
        max.iter=100000,
        max.overlaps=10000,
        max.time=3,
        segment.size=0.125,fontface = "italic")+
    scale_x_continuous(limits=c(min(visdeplt$logFC),max(visdeplt$logFC)),expand=c(0.025,0.025))+
    xlab("logFC (M:F)")+
    ylab("-log10 (P-value)")+
    guides(color="none")+
    theme(axis.text=element_text(size=7),axis.title=element_text(size=7.5),plot.title=element_text(size=9,hjust=0.5,margin = margin(0,0,0.025,0,"in")))+
    rasterize(geom_point(size=0.4), layers="Points",dpi=450,dev="cairo_png")
    

pv <- rasterize(p,layers="Points",dpi=450,dev="cairo_png")
rm(p)
```


## panel B (or D, depending how we arrange panels): Visium ARC volcano plot
```{r}
visdeplt <- visde[assay=="ARC"][gene_name %in% c("XIST","DDX3Y"),label:=gene_name]


p <- ggplot(visdeplt, aes(y = -log10(P.Value), x = logFC,col=assay)) +
    ggtitle("ARC (Visium)")+
    scale_color_manual(values=pal) +
    geom_hline(yintercept = -log10(0.05),
        linetype = "dashed",
        color = "black",
        linewidth=0.25) +
    geom_hline(yintercept = -log10(5e-5),
        linetype = "dashed",
        color = "red",
        linewidth=0.25) +
    geom_text_repel(aes(y = -log10(P.Value),
            x = logFC,
            label = ifelse(is.na(label), "", label)),
        min.segment.length = 0,
        size = 3,
        show.legend=FALSE,
        #nudge_x=0.85*sign(visdeplt$logFC),
        seed=10403, 
        max.iter=100000,
        max.overlaps=10000,
        max.time=3,
        segment.size=0.125,
        fontface="italic")+
    # scale_x_continuous(limits=c(-4,2.5),expand=c(0,0))+
    xlab("logFC (M:F)")+
    ylab("-log10 (P-value)")+
    guides(color="none")+
    theme(axis.text=element_text(size=7),axis.title=element_text(size=7.5),plot.title=element_text(size=9,hjust=0.5,margin = margin(0,0,0.025,0,"in")))+
    rasterize(geom_point(size=0.4),"Points",dpi=450,dev="cairo_png")
    

# rasterize
pa <- rasterize(p,layers="Points",dpi=450,dev="cairo_png")

# save
pdf("manuscript_plots/Supplemental/SF-vVMH_vARC_volcanos_wAllosomes.pdf",height=4,width=8)
do.call("grid.arrange",c(list(pv,pa),ncol=2))
dev.off()
```

repinf
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sequoia 15.3.2

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.40.0 parallelly_1.43.0   Polychrome_1.5.1    viridis_0.6.5      
 [5] viridisLite_0.4.2   gridExtra_2.3       ggrastr_1.0.2       ggrepel_0.9.6      
 [9] ggtext_0.1.2        ggplot2_3.5.2       data.table_1.17.4   rlang_1.1.5        

loaded via a namespace (and not attached):
 [1] generics_0.1.4       xml2_1.3.8           digest_0.6.37       
 [4] magrittr_2.0.3       evaluate_1.0.3       grid_4.4.3          
 [7] RColorBrewer_1.1-3   fastmap_1.2.0        rprojroot_2.0.4     
[10] sessioninfo_1.2.3    scales_1.4.0         codetools_0.2-20    
[13] cli_3.6.4            scatterplot3d_0.3-44 withr_3.0.2         
[16] yaml_2.3.10          ggbeeswarm_0.7.2     tools_4.4.3         
[19] parallel_4.4.3       dplyr_1.1.4          colorspace_2.1-1    
[22] here_1.0.1           png_0.1-8            vctrs_0.6.5         
[25] R6_2.6.1             lifecycle_1.0.4      vipor_0.4.7         
[28] pkgconfig_2.0.3      beeswarm_0.4.0       pillar_1.10.2       
[31] gtable_0.3.6         glue_1.8.0           Rcpp_1.0.14         
[34] xfun_0.52            tibble_3.2.1         tidyselect_1.2.1    
[37] rstudioapi_0.17.1    knitr_1.50           dichromat_2.0-0.1   
[40] farver_2.1.2         htmltools_0.5.8.1    rmarkdown_2.29      
[43] labeling_0.4.3       compiler_4.4.3       gridtext_0.1.5      
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sequoia 15.3.2
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2025-11-10
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ─────────────────────────────────────────────────────────────────────────
 !  package       * version date (UTC) lib source
    beeswarm        0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
    BiocParallel  * 1.40.0  2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 VP cli             3.6.4   2025-04-23 [2] CRAN (R 4.4.1) (on disk 3.6.5)
    codetools       0.2-20  2024-03-31 [1] CRAN (R 4.4.3)
    colorspace      2.1-1   2024-07-26 [1] CRAN (R 4.4.0)
    data.table    * 1.17.4  2025-05-26 [1] CRAN (R 4.4.3)
    dichromat       2.0-0.1 2022-05-02 [1] CRAN (R 4.4.0)
    digest          0.6.37  2024-08-19 [1] CRAN (R 4.4.1)
    dplyr           1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
    evaluate        1.0.3   2025-01-10 [1] CRAN (R 4.4.1)
    farver          2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
    fastmap         1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
    generics        0.1.4   2025-05-09 [1] CRAN (R 4.4.1)
    ggbeeswarm      0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
    ggplot2       * 3.5.2   2025-04-09 [1] CRAN (R 4.4.1)
    ggrastr       * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
    ggrepel       * 0.9.6   2024-09-07 [1] CRAN (R 4.4.1)
    ggtext        * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
    glue            1.8.0   2024-09-30 [1] CRAN (R 4.4.1)
    gridExtra     * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
    gridtext        0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
    gtable          0.3.6   2024-10-25 [1] CRAN (R 4.4.1)
    here            1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
    htmltools       0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
    knitr           1.50    2025-03-16 [1] CRAN (R 4.4.2)
    labeling        0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
    lifecycle       1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
    magrittr        2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
    parallelly    * 1.43.0  2025-03-24 [1] CRAN (R 4.4.1)
    pillar          1.10.2  2025-04-05 [1] CRAN (R 4.4.1)
    pkgconfig       2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
    png             0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
    Polychrome    * 1.5.1   2022-05-03 [1] CRAN (R 4.4.0)
    R6              2.6.1   2025-02-15 [1] CRAN (R 4.4.1)
    RColorBrewer    1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
    Rcpp            1.0.14  2025-01-12 [1] CRAN (R 4.4.1)
 VP rlang         * 1.1.5   2025-04-11 [2] CRAN (R 4.4.1) (on disk 1.1.6)
    rmarkdown       2.29    2024-11-04 [1] CRAN (R 4.4.1)
    rprojroot       2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
    rstudioapi      0.17.1  2024-10-22 [1] CRAN (R 4.4.1)
    scales          1.4.0   2025-04-24 [1] CRAN (R 4.4.1)
    scatterplot3d   0.3-44  2023-05-05 [1] CRAN (R 4.4.0)
    sessioninfo     1.2.3   2025-02-05 [1] CRAN (R 4.4.1)
    tibble          3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
    tidyselect      1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
    vctrs           0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
    vipor           0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
    viridis       * 0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
    viridisLite   * 0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
    withr           3.0.2   2024-10-28 [1] CRAN (R 4.4.1)
    xfun            0.52    2025-04-02 [1] CRAN (R 4.4.1)
    xml2            1.3.8   2025-03-14 [1] CRAN (R 4.4.1)
    yaml            2.3.10  2024-07-26 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 V ── Loaded and on-disk version mismatch.
 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────────