---
title: "SF-vARC-vVMH and v-x sexDE comparisons"
output: html_document
date: "2024-11-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(ggtext)


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

sex DE table
```{r}
visde <- fread("spatial_HYP/processed-data/09-Sex DE/01c-voomLmFit_nnsvg10-HmnylmbNA-BS-15-VMHARCclpsd.txt")

## append chromosome info to filter out sex chr, mito genes
enslut <- fread("spatial_HYP/raw-data/USEFOR10X_ensg_lut_hg38_fromEns98_111123.txt")
enslut <- unique(enslut[,.(ensembl_gene_id,chromosome_name)])
enslut[,chromosome_name:=paste0("chr",chromosome_name)]

visde <- merge.data.table(visde,enslut,by.x="gene_id",by.y="ensembl_gene_id")
visde <- visde[chromosome_name %in% paste0("chr",c(1:22))]

vmh <- visde[assay=="VMH",.(gene_id,gene_name,logFC,P.Value,adj.P.Val,t,B)]
arc <- visde[assay=="ARC",.(gene_id,gene_name,logFC,P.Value,adj.P.Val,t,B)]
setnames(vmh,c(3:ncol(vmh)),paste0("vmh_",names(vmh)[3:ncol(vmh)]))
setnames(arc,c(3:ncol(arc)),paste0("arc_",names(arc)[3:ncol(arc)]))

jt <- merge.data.table(vmh,arc,by="gene_id")


# initialize plot list
plts <- list()

# first plot: t stat correlations
sprh <- round(cor(jt$vmh_t,jt$arc_t,use = "pairwise.complete.obs",method = "spearman"),digits = 3)

# make abline dataframe
xeqy <- as.data.table(data.frame(intercept=0,slope=1,`Line Color`="x=y"))
setnames(xeqy,3,"Line Color")

plts[[1]] <- ggplot(jt,aes(x=vmh_t,y=arc_t))+
    geom_point(size=0.375)+
    ggtitle(paste0("Autosomal sex DE *t*-stats<br>(Spearman rho: ",sprh,")"))+
    xlab("vVMH *t*-statistic")+
    ylab("vARC *t*-statistic")+
    scale_x_continuous(limits=c(min(c(jt$vmh_t,jt$arc_t)),max(c(jt$vmh_t,jt$arc_t))),expand=c(0.05,0.05))+
    scale_y_continuous(limits=c(min(c(jt$vmh_t,jt$arc_t)),max(c(jt$vmh_t,jt$arc_t))),expand=c(0.05,0.05))

rm(sprh)
## logFC correlations
sprh <- round(cor(jt$vmh_logFC,jt$arc_logFC,use = "pairwise.complete.obs",method = "spearman"),digits = 3)
pcc <- round(cor(jt$vmh_logFC,jt$arc_logFC,use = "pairwise.complete.obs",method = "pearson"),digits = 3)

plts[[2]] <- ggplot(jt,aes(x=vmh_logFC,y=arc_logFC))+
    geom_point(size=0.375)+
    ggtitle(paste0("Autosomal sex DE logFC<br>(Spearman ",sprh," / Pearson ",pcc,")"))+
    xlab("vVMH logFC")+
    ylab("vARC logFC")+
    scale_x_continuous(limits=c(min(c(jt$vmh_logFC,jt$arc_logFC)),max(c(jt$vmh_logFC,jt$arc_logFC))),expand=c(0.05,0.05))+
    scale_y_continuous(limits=c(min(c(jt$vmh_logFC,jt$arc_logFC)),max(c(jt$vmh_logFC,jt$arc_logFC))),expand=c(0.05,0.05))

rm(sprh,pcc)

## panel 3: unconditional *p*-value densities: genes nominal in either
jtun <- jt[(vmh_P.Value<0.05|arc_P.Value<0.05)&!is.na(vmh_P.Value)&!is.na(arc_P.Value)]
jtun <- melt.data.table(jtun,id.vars=c("gene_id"),measure.vars = c("arc_P.Value","vmh_P.Value"))
jtun[,variable:=gsub(variable,pattern="^(.*)_.*$",replacement="\\1")]
jtun[,variable:=paste0("v",toupper(variable))]


plts[[3]] <- ggplot(jtun,aes(x=value,color=variable))+
    geom_density()+
    ggtitle("Gene *p*-value density: *p*<0.05, either domain")+
    xlab("*p*-value")+
    ylab("Density")+
    labs(color="Domain")+
    scale_y_continuous(expand=c(0,0))+
    theme(legend.position = "inside",legend.position.inside=c(0.85,0.825),legend.background = element_blank(),legend.key=element_blank(),legend.margin = margin(0,0,-0.05,0,unit="in"),              axis.text.x = element_text(size = 8),
              axis.title.x = element_markdown(size = 9),
              axis.text.y = element_text(size = 8),
              axis.title.y = element_markdown(size =9))


rm(jtun)

# panel 4 using p<0.01
jtun <- jt[(vmh_P.Value<0.01|arc_P.Value<0.01)&!is.na(vmh_P.Value)&!is.na(arc_P.Value)]
jtun <- melt.data.table(jtun,id.vars=c("gene_id"),measure.vars = c("arc_P.Value","vmh_P.Value"))
jtun[,variable:=gsub(variable,pattern="^(.*)_.*$",replacement="\\1")]
jtun[,variable:=paste0("v",toupper(variable))]

plts[[4]] <- ggplot(jtun,aes(x=value,color=variable))+
    geom_density()+
    ggtitle("Gene *p*-value density: *p*<0.01, either domain")+
    xlab("*p*-value")+
    ylab("Density")+
    labs(color="Domain")+
    scale_y_continuous(expand=c(0,0))+
    theme(legend.position = "inside",legend.position.inside=c(0.8,0.825),legend.background = element_blank(),legend.key=element_blank(),legend.margin = margin(0,0,-0.05,0,unit="in"),axis.text.x = element_text(size = 8),
              axis.title.x = element_markdown(size = 9),
              axis.text.y = element_text(size = 8),
              axis.title.y = element_markdown(size =9))

rm(jtun)

#### conditional: ARC DEGs p<0.2, p<0.05, p<0.01 in VMH
vcon <- jt[arc_P.Value<0.2,.(gene_id,vmh_P.Value,arc_P.Value)]
vcon[,grp:=ifelse(arc_P.Value>0.1,"vARC *p* 0.1 to 0.2","tmp")]
vcon[arc_P.Value>0.05&arc_P.Value<0.1,grp:="vARC *p* 0.05 to 0.1"]
vcon[arc_P.Value<0.05,grp:="vARC *p*<0.05"]
vcon2 <- jt[arc_P.Value>0.2,.(gene_id,vmh_P.Value,arc_P.Value)]
vcon2[,grp:="vARC *p*>0.2"]
vcon <- rbind(vcon,vcon2)

plts[[5]] <- ggplot(vcon,aes(x=vmh_P.Value,color=grp))+
    geom_density()+
    ggtitle("vVMH gene *p*-value density by *p* in vARC")+
    xlab("vVMH *p*-value")+
    ylab("Density")+
    scale_y_continuous(expand=c(0,0))+
    theme(legend.position = "inside",legend.position.inside=c(0.8,0.75),legend.background = element_blank(),legend.key=element_blank(),legend.margin = margin(0,0,-0.05,0,unit="in"),legend.title=element_blank(),              axis.text.x = element_text(size = 8),
              axis.title.x = element_markdown(size = 9),
              axis.text.y = element_text(size = 8),
              axis.title.y = element_markdown(size =9))

### VMH DEGs in ARC
acon <- jt[vmh_P.Value<0.2,.(gene_id,vmh_P.Value,arc_P.Value)]
acon[,grp:=ifelse(vmh_P.Value>0.1,"vVMH *p*0.1 to 0.2","tmp")]
acon[vmh_P.Value>0.05&vmh_P.Value<0.1,grp:="vVMH *p* 0.05 to 0.1"]
acon[vmh_P.Value<0.05,grp:="vVMH *p*<0.05"]
acon2 <- jt[vmh_P.Value>0.2,.(gene_id,vmh_P.Value,arc_P.Value)]
acon2[,grp:="vVMH *p*>0.2"]
acon <- rbind(acon,acon2)

plts[[6]] <- ggplot(acon,aes(x=arc_P.Value,color=grp))+
    geom_density()+
    ggtitle("vARC gene *p*-value density by *p* in vVMH")+
    xlab("vARC *p*-value")+
    ylab("Density")+
    scale_y_continuous(expand=c(0,0))+
    theme(legend.position = "inside",legend.position.inside=c(0.8,0.75),legend.background = element_blank(),legend.key=element_blank(),legend.margin = margin(0,0,-0.05,0,unit="in"),legend.title=element_blank(),              axis.text.x = element_text(size = 8),
              axis.title.x = element_markdown(size = 9),
              axis.text.y = element_text(size = 8),
              axis.title.y = element_markdown(size =9))
```

## xenium-visium comparative plots (domain level)
```{r}
xendom <- fread("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/04-4typeARC-and-dualassignVMH_domainwise_sexDE.txt")

### grab ensembl ids (retrieved when assembling SPE) so we can match to ids in visium
enslut <- fread("xenium_HYP/processed-data/03_make_SPE-SFE/01_rowData_raw.txt")
xendom <- merge.data.table(xendom,enslut[,.(ID,Symbol)],by.x="gene_name",by.y="Symbol")

## check that symbol and ID give same visium results
stopifnot(identical(vmh[gene_name %in% unique(xendom$gene_name)&vmh_P.Value<0.1],vmh[gene_id %in% unique(xendom$ID)&vmh_P.Value<0.1]))

stopifnot(identical(arc[gene_name %in% unique(xendom$gene_name)&arc_P.Value<0.1],arc[gene_id %in% unique(xendom$ID)&arc_P.Value<0.1]))

###
xvmh <- xendom[domain=="VMH",.(ID,logFC_MvF,P.Value,adj.P.Val,t,B)]
xarc <- xendom[domain=="ARC",.(ID,logFC_MvF,P.Value,adj.P.Val,t,B)]
setnames(xvmh,1,"gene_id")
setnames(xarc,1,"gene_id")
setnames(xvmh,c(2:ncol(xvmh)),paste0("xvmh_",names(xvmh)[2:ncol(xvmh)]))
setnames(xarc,c(3:ncol(xarc)),paste0("xarc_",names(xarc)[3:ncol(xarc)]))

arcxv <- merge.data.table(arc,xarc,by="gene_id")
vmhxv <- merge.data.table(vmh,xvmh,by="gene_id")
```


xen-vis comparison plot series: ARC (all genes), VMH (all genes), ARC (36 prioritized genes), VMH (36 prioritized genes)
```{r}
xvplts <- list()

# arc
cor.test(arcxv$xarc_t,arcxv$arc_t,use = "pairwise.complete.obs",method = "spearman") # p < 2.2e-16
sprh <- round(cor(arcxv$xarc_t,arcxv$arc_t,use = "pairwise.complete.obs",method = "spearman"),3)

xvplts[[1]] <- ggplot(arcxv,aes(y=xarc_t,x=arc_t))+
    geom_point(size=0.375)+
    ggtitle(paste0("vARC-xARC Sex DE *t*<br>(277 shared genes;<br>rho= ",sprh,")"))+
    xlab("Visium ARC *t*-statistic")+
    ylab("Xenium ARC *t*-statistic")+
    scale_x_continuous(limits=c(min(c(arcxv$arc_t,arcxv$xarc_t)),max(c(arcxv$arc_t,arcxv$xarc_t))),expand=c(0.05,0.05))+
    scale_y_continuous(limits=c(min(c(arcxv$arc_t,arcxv$xarc_t)),max(c(arcxv$arc_t,arcxv$xarc_t))),expand=c(0.05,0.05))
    
rm(sprh)

# vmh
cor.test(vmhxv$xvmh_t,vmhxv$vmh_t,use = "pairwise.complete.obs",method = "spearman") # p < 5.46e-11

sprh <- round(cor(vmhxv$xvmh_t,vmhxv$vmh_t,use = "pairwise.complete.obs",method = "spearman"),3)

xvplts[[2]] <- ggplot(vmhxv,aes(y=xvmh_t,x=vmh_t))+
    geom_point(size=0.375)+
    ggtitle(paste0("vVMH-xVMH Sex DE *t*<br>(245 shared genes;<br>rho= ",sprh,")"))+
    xlab("Visium VMH *t*-statistic")+
    ylab("Xenium VMH *t*-statistic")+
    scale_x_continuous(limits=c(min(c(vmhxv$vmh_t,vmhxv$xvmh_t)),max(c(vmhxv$vmh_t,vmhxv$xvmh_t))),expand=c(0.05,0.05))+
    scale_y_continuous(limits=c(min(c(vmhxv$vmh_t,vmhxv$xvmh_t)),max(c(vmhxv$vmh_t,vmhxv$xvmh_t))),expand=c(0.05,0.05))

rm(sprh)
## subset to genes earmarked for sex de followup
fulist <- fread("xenium_HYP/raw-data/genes_prioritized_for_xeniumSexDEfollowup.txt",header=F)

arcxv <- arcxv[gene_name %in% fulist$V1]
vmhxv <- vmhxv[gene_name %in% fulist$V1]

cor.test(arcxv$xarc_t,arcxv$arc_t,use = "pairwise.complete.obs",method = "spearman") # p < 5.8e-7

sprh <- round(cor(arcxv$xarc_t,arcxv$arc_t,use = "pairwise.complete.obs",method = "spearman"),3)

xvplts[[3]] <- ggplot(arcxv,aes(y=xarc_t,x=arc_t))+
    geom_point(size=0.375)+
    ggtitle(paste0("vARC-xARC Sex DE *t*<br>(33 prioritized genes;<br>rho= ",sprh,")"))+
    xlab("vARC *t*-statistic")+
    ylab("xARC *t*-statistic")+
    scale_x_continuous(limits=c(min(c(arcxv$arc_t,arcxv$xarc_t)),max(c(arcxv$arc_t,arcxv$xarc_t))),expand=c(0.05,0.05))+
    scale_y_continuous(limits=c(min(c(arcxv$arc_t,arcxv$xarc_t)),max(c(arcxv$arc_t,arcxv$xarc_t))),expand=c(0.05,0.05))

rm(sprh)

# vmh, earmarked genes

cor.test(vmhxv$xvmh_t,vmhxv$vmh_t,use = "pairwise.complete.obs",method = "spearman") # p < 0.008

sprh <- round(cor(vmhxv$xvmh_t,vmhxv$vmh_t,use = "pairwise.complete.obs",method = "spearman"),3)

xvplts[[4]] <- ggplot(vmhxv,aes(y=xvmh_t,x=vmh_t))+
    geom_point(size=0.375)+
    ggtitle(paste0("vVMH-xVMH Sex DE *t*<br>(27 prioritized genes;<br>rho= ",sprh,")"))+
    xlab("vVMH *t*-statistic")+
    ylab("xVMH *t*-statistic")+
    scale_x_continuous(limits=c(min(c(vmhxv$vmh_t,vmhxv$xvmh_t)),max(c(vmhxv$vmh_t,vmhxv$xvmh_t))),expand=c(0.05,0.05))+
    scale_y_continuous(limits=c(min(c(vmhxv$vmh_t,vmhxv$xvmh_t)),max(c(vmhxv$vmh_t,vmhxv$xvmh_t))),expand=c(0.05,0.05))

```

## standardize theme and draw x=y line onto correlation plots. defaults defined in setup code block don't pass into lapply so reproduce here
```{r}
plts[c(1:2)] <- lapply(plts[c(1:2)],function(x){
    x + 
        theme_bw()+
        theme(legend.position = "inside", 
              legend.position.inside=c(0.1, 0.875), 
              legend.background = element_blank(),
              legend.key=element_blank(),
              legend.text=element_text(size=8, margin=margin(0, 0, 0, -0.06, unit = "in")), 
              axis.text.x = element_text(size = 8),
              axis.title.x = element_markdown(size = 9),
              axis.text.y = element_text(size = 8),
              axis.title.y = element_markdown(size =9),
              plot.title = element_markdown(size = 11, hjust=0.5),
              strip.text = element_text(size=11),
              legend.title = element_text(size=9, hjust=0.5))+
        geom_abline(data = xeqy, linetype="dashed", aes(slope = slope, intercept=intercept, col=`Line Color`))
})

xvplts[c(1:4)] <- lapply(xvplts[c(1:4)],function(x){
    x + theme_bw()+
        theme(legend.position = "inside",
              legend.position.inside=c(0.195,0.85),
              legend.background = element_blank(),
              legend.key=element_blank(),
              legend.text=element_text(size=8,margin=margin(0,0,0,-0.04,unit = "in")),
              axis.text.x = element_text(size = 8),
              axis.title.x = element_markdown(size = 9),
              axis.text.y = element_text(size = 8),
              axis.title.y = element_markdown(size =9),
              plot.title = element_markdown(size = 11, hjust=0.5),
              strip.text = element_text(size=11),
              legend.title = element_text(size=9, hjust=0.5))+
        geom_abline(data = xeqy,linetype="dashed",aes(slope = slope,intercept=intercept,col=`Line Color`))
})
```
### save
```{r}
pdf("manuscript_plots/Supplemental/SF-vVMH-vARC_sex_DE_comparative_stats.pdf",height=11,width=8.5,onefile = FALSE)
do.call("grid.arrange",list(arrangeGrob(grobs=plts,ncol=2),arrangeGrob(grobs=xvplts,nrow=1),nrow=2,heights=c(8.5,2.5)))
dev.off()
```

```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] colorout_1.3-0.2  ggtext_0.1.2      gridExtra_2.3     ggplot2_3.5.1    
[5] data.table_1.15.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] gtable_0.3.5      dplyr_1.1.4       compiler_4.4.1    tidyselect_1.2.1 
 [5] Rcpp_1.0.13       stringr_1.5.1     xml2_1.3.6        scales_1.3.0     
 [9] yaml_2.3.9        fastmap_1.2.0     here_1.0.1        R6_2.5.1         
[13] commonmark_1.9.1  labeling_0.4.3    generics_0.1.3    knitr_1.48       
[17] tibble_3.2.1      munsell_0.5.1     rprojroot_2.0.4   pillar_1.9.0     
[21] utf8_1.2.4        stringi_1.8.4     xfun_0.45         cli_3.6.3        
[25] withr_3.0.0       magrittr_2.0.3    digest_0.6.36     grid_4.4.1       
[29] gridtext_0.1.5    rstudioapi_0.16.0 markdown_1.13     lifecycle_1.0.4  
[33] vctrs_0.6.5       evaluate_0.24.0   glue_1.7.0        farver_2.1.2     
[37] sessioninfo_1.2.2 fansi_1.0.6       colorspace_2.1-0  rmarkdown_2.27   
[41] tools_4.4.1       pkgconfig_2.0.3   htmltools_0.5.8.1
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-11-21
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package     * version date (UTC) lib source
 P cli           3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout    * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace    2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   commonmark    1.9.1   2024-01-30 [1] CRAN (R 4.4.0)
   data.table  * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest        0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr         1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate      0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi         1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver        2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics      0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggplot2     * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggtext      * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
   glue          1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra   * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gridtext      0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
   gtable        0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here          1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools     0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr         1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling      0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle     1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr      2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   markdown      1.13    2024-06-04 [1] CRAN (R 4.4.0)
   munsell       0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   pillar        1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig     2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6            2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp          1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
 P rlang       * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown     2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot     2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi    0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales        1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo   1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   stringi       1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
   stringr       1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
   tibble        3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect    1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8          1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs         0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr         3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun          0.45    2024-06-16 [1] CRAN (R 4.4.0)
   xml2          1.3.6   2023-12-04 [1] CRAN (R 4.4.0)
   yaml          2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────