---
title: "SF-VMH and ARC clusters and markers at k15, 20, 31"
output: html_document
date: "2024-10-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(ComplexHeatmap)
library(ggrastr) # render plots with bajillions of points as raster so they can be manipulated in eg Illustrator--Fig 2B
library(Cairo) # pdf device for ^ to properly save; also used for complexheatmap raster functions
library(circlize) #colorRamp2
library(viridis) # magma
library(ggtext) # mixed font formats in ggplot elements

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

## load SPE, clusterings, one-v-all markers, within-VMH-markers, within-ARC markers
```{r}
hyp2 <- readRDS("spatial_HYP/processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

bscls <- list.files("spatial_HYP/processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out",full.names=T,pattern="HARMONYlmbna_nnsvg10")

globalmk <- readRDS("spatial_HYP/processed-data/07-Marker_genes/01b-BSpace_allruns_spatialLIBD-regwrap-enrichment_full.RDS")
## subset to the three main clustering approaches
globalmk <- globalmk[names(globalmk) %in% grep(names(globalmk),pattern="HARMONYlmbna_nnsvg10_60kiter$",value=T)]

# this was only done for those three to begin with:
intramk <- readRDS("spatial_HYP/processed-data/07-Marker_genes/05-IntraVMH-intraARC_1clus_vallother_clus_slibdEnrichment_full.RDS")

sampleplotnames <- fread("standardized_sampleids_for_plotting.txt")
```

## replace sample names for plotting. doing it once the manuscript_ids are in the SPE's coldata updates imgdata sample ids simultaneously.
```{r}
tmpcd <- as.data.table(colData(hyp2),keep.rownames=T)
tmpcd <- merge(tmpcd,sampleplotnames[,.(sample_id,manuscript_id)],by="sample_id",all.x=T)
tmpcd <- DataFrame(tmpcd,row.names = tmpcd$rn)[colnames(hyp2),]
colData(hyp2) <- tmpcd

## update colData and imgData
hyp2$sample_id <- hyp2$manuscript_id
```

## prepare clustering tables for assignments to hyp2 as labels and (global) marker tables to use the same nomenclature
```{r}
## clustering assignments:
bscls.dt <- lapply(bscls,FUN = function(b){
    bdt <- fread(b)
    setnames(bdt,2,"cl")
    bdt[,cl:=paste0("X",cl)]
    return(bdt)
})

## we also want to label white matter since it goes in the global heatmaps
# 15: WM is predom X3, X10
# 20: WM is predom  x15, X17
# 31: WM is predom X7, X15, X23

names(bscls.dt) <- c("k15","k20","k31")

bscls.dt[["k15"]][cl=="X4",cl:="ARC.1"]
bscls.dt[["k15"]][cl=="X6",cl:="ARC.2"]
bscls.dt[["k15"]][cl=="X7",cl:="VMH.1"]
bscls.dt[["k15"]][cl=="X12",cl:="VMH.2"]
bscls.dt[["k15"]][cl=="X3",cl:="WM.1"]
bscls.dt[["k15"]][cl=="X10",cl:="WM.2"]

bscls.dt[["k20"]][cl=="X2",cl:="VMH20.1"]
bscls.dt[["k20"]][cl=="X3",cl:="VMH20.2"]
bscls.dt[["k20"]][cl=="X1",cl:="ARC20.1"]
bscls.dt[["k20"]][cl=="X4",cl:="ARC20.2"]
bscls.dt[["k20"]][cl=="X15",cl:="WM20.1"]
bscls.dt[["k20"]][cl=="X17",cl:="WM20.2"]

# k31: x6 21 25 26 31 = VMH; 2,4,11,22 ARC
bscls.dt[["k31"]][cl=="X6",cl:="VMH31.1"]
bscls.dt[["k31"]][cl=="X21",cl:="VMH31.2"]
bscls.dt[["k31"]][cl=="X25",cl:="VMH31.3"]
bscls.dt[["k31"]][cl=="X26",cl:="VMH31.4"]
bscls.dt[["k31"]][cl=="X31",cl:="VMH31.5"]
bscls.dt[["k31"]][cl=="X2",cl:="ARC31.1"]
bscls.dt[["k31"]][cl=="X4",cl:="ARC31.2"]
bscls.dt[["k31"]][cl=="X11",cl:="ARC31.3"]
bscls.dt[["k31"]][cl=="X22",cl:="ARC31.4"]
bscls.dt[["k31"]][cl=="X7",cl:="WM31.1"]
bscls.dt[["k31"]][cl=="X15",cl:="WM31.2"]
bscls.dt[["k31"]][cl=="X23",cl:="WM31.3"]

## global markers pt 1: melts and list names to match bscls.dt
globalmk2 <- lapply(globalmk,FUN = function(g){
    gm <- melt(g,id.vars="gene")
    gm[,clus:=gsub(variable,pattern="^.*_([[:alnum:]]{2,5})$",replacement="\\1")]
    gm[,stat:=gsub(variable,pattern="^(.*)_([[:alnum:]]{2,5})$",replacement="\\1")]
    
    gml <- gm[stat=="logFC"&value>0]
    setnames(gml,"value","logFC_1vall")
    gml[,variable:=NULL]
    
    gmf <- gm[stat=="fdr"&value<0.2]
    setnames(gmf,"value","fdr_1vall")
    gmf[,variable:=NULL]
    
    gmo <- merge.data.table(gmf[,.(gene,clus,fdr_1vall)],gml[,.(gene,clus,logFC_1vall)],by=c("gene","clus"))
    gmo[,fdr_1vall:=as.numeric(fdr_1vall)]
    gmo[,logFC_1vall:=as.numeric(logFC_1vall)]
})
names(globalmk2) <- c("k15","k20","k31")

## global markers pt 2: rename clusters to match other data sources. we also want WMs in here for the globals
# 15: WM is predom X3, X10
# 20: WM is predom  x15, X17
# 31: WM is predom X7, X15, X23

globalmk2[["k15"]][clus=="X4",clus:="ARC.1"]
globalmk2[["k15"]][clus=="X6",clus:="ARC.2"]
globalmk2[["k15"]][clus=="X7",clus:="VMH.1"]
globalmk2[["k15"]][clus=="X12",clus:="VMH.2"]
globalmk2[["k15"]][clus=="X3",clus:="WM.1"]
globalmk2[["k15"]][clus=="X10",clus:="WM.2"]

globalmk2[["k20"]][clus=="X2",clus:="VMH20.1"]
globalmk2[["k20"]][clus=="X3",clus:="VMH20.2"]
globalmk2[["k20"]][clus=="X1",clus:="ARC20.1"]
globalmk2[["k20"]][clus=="X4",clus:="ARC20.2"]
globalmk2[["k20"]][clus=="X15",clus:="WM20.1"]
globalmk2[["k20"]][clus=="X17",clus:="WM20.2"]

# k31: x6 21 25 26 31 = VMH; 2,4,11,22 ARC
globalmk2[["k31"]][clus=="X6",clus:="VMH31.1"]
globalmk2[["k31"]][clus=="X21",clus:="VMH31.2"]
globalmk2[["k31"]][clus=="X25",clus:="VMH31.3"]
globalmk2[["k31"]][clus=="X26",clus:="VMH31.4"]
globalmk2[["k31"]][clus=="X31",clus:="VMH31.5"]
globalmk2[["k31"]][clus=="X2",clus:="ARC31.1"]
globalmk2[["k31"]][clus=="X4",clus:="ARC31.2"]
globalmk2[["k31"]][clus=="X11",clus:="ARC31.3"]
globalmk2[["k31"]][clus=="X22",clus:="ARC31.4"]
globalmk2[["k31"]][clus=="X7",clus:="WM31.1"]
globalmk2[["k31"]][clus=="X15",clus:="WM31.2"]
globalmk2[["k31"]][clus=="X23",clus:="WM31.3"]

# set other cluster IDs to "Other"
bscls.dt <- lapply(bscls.dt,FUN=function(x){x[cl %in% grep(cl,pattern="VMH|ARC|WM",value=T,invert=T),cl:="Other"]})
globalmk2 <- lapply(globalmk2,FUN=function(x){x[clus %in% grep(clus,pattern="VMH|ARC|WM",value=T,invert=T),clus:="Other"]})

```

## use the xenium palettes for VMH/ARC visium clusters for consistency
```{r}
pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")
vmhpal <- pals[["XenVMH"]]
arcpal <- pals[["XenARC"]]
# we have 5 VMH clusters max and 4 ARC max here, plus 3 max WM, so move one of the colors (#09979B) to VMH here; replace VMH gray with "#FA260D"; replace ARC grayish with #9F09E0; use blues for WM
vmhpal <- c(vmhpal,"#09979B")
arcpal <- arcpal[c(1,2,4,5)]
arcpal[which(arcpal=="#A997AB")] <- "#9F09E0"

# reorder VMH pal so puke orange is only in the highest clustering config
vmhpal[which(vmhpal=="darkgray")] <- "#a3690D"
vmhpal <- vmhpal[c(1,3,4,5,2)]

wmpal <- c("#ADD8E6","#0000FF","#00008B")
```

## prepare expression matrices for each set of global, domain markers
## # Z-scale logcounts, make a subsetted spe with the columns already ordered by cluster since complexheatmap was giving issues with this, make plot.
### this is not immediately cooperating for an axis flip so just do it in illustrator. we can dump the domain color legend (which is the same as A and B) in AI too.
```{r}
## get about 40 genes total
globalmkgn <- lapply(globalmk2,FUN=function(g){
    gs <- g[clus %in% grep(clus,pattern="VMH|ARC|WM",value=T)]
    mkperclus <- ceiling(40/(length(unique(gs$clus))))
    
    gns <- gs[fdr_1vall<0.05,.SD[order(-logFC_1vall)][1:mkperclus],by="clus"][,gene]
    gns <- unique(gns)
    gns <- gns[!is.na(gns)]
    return(gns)
})

## define function to z-scale expression
# z scale rows, from tony https://github.com/LieberInstitute/spatial_DG_lifespan/blob/9419eb91453cda1df494d93dc91156d819042c66/code/Pseudobulk/top_infant_DE_heatmap_enrichment.R#L106

scale_rows <- function(x) {
    m <- apply(x, 1, mean, na.rm = T)
    s <- apply(x, 1, sd, na.rm = T)
    return((x - m) / s)
}


# set hyp2 row names to gene names to make this simpler
rownames(hyp2) <- rowData(hyp2)$gene_name

# make sure our cluster and gene lists are in the same order
globalmkgn <- globalmkgn[names(bscls.dt)]

# get expression matrices
globalmkxpr <- mapply(g=globalmkgn,b=bscls.dt,SIMPLIFY=FALSE,FUN=function(g,b){
    b <- b[cl %in% grep(cl,pattern="VMH|ARC|WM",value=T)]
    
    
    # subset to relevant spots
    hyptmp <- hyp2[,hyp2$key %in% b$rn]
    
    # label spots
    b <- DataFrame(b,row.names=b$rn)[colnames(hyptmp),]
    colLabels(hyptmp) <- b$cl
    
    # order as factor so we can fetch the spots per cluster in a presentable order (ARCS, VMHs, WMs)
    arcs <- grep(unique(b$cl),pattern="ARC",value=T)
    vmhs <- grep(unique(b$cl),pattern="VMH",value=T)
    wms <- grep(unique(b$cl),pattern="WM",value=T)

    hyptmp$label <- factor(hyptmp$label,levels=c(arcs,vmhs,wms,"Other"))

        
   # make a subsetted spe with reordering of the columns to be in the order of the domains we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
    i <- 1
    for (i in c(1:length(levels(hyptmp$label)))) {
        curlev <- levels(hyptmp$label)[i]
        if (i == 1) {
            cnames <- colnames(hyptmp[, hyptmp$label == curlev])
        } else {
            cnames <- c(cnames, colnames(hyptmp[, hyptmp$label == curlev]))
        }
    }
    rm(i, curlev)
    
    ### extract log counts per marker, z-normalize, return both
    pltdat <- hyptmp[g, cnames]
    pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))
    
    retlist <- list(pltdat,pltdat2)
    
    return(retlist)
})
```

## make global heatmaps (we'll arrange these with SPEs + intra-domain mk heatmaps on the illustrator side).
```{r}
mapply(n=names(globalmkxpr),x=globalmkxpr,SIMPLIFY=FALSE,FUN=function(x,n){
    
    arcs <- unique(grep(x[[1]]$label,pattern="ARC",value=T))
    narc <- length(arcs)
    vmhs <- unique(grep(x[[1]]$label,pattern="VMH",value=T))
    nvmh <- length(vmhs)
    wms <- unique(grep(x[[1]]$label,pattern="WM",value=T))
    nwm <- length(wms)
    
    # nclusts <- length(unique(b[, 2]))
    
    # pal <- Polychrome::palette36.colors(n = (nclusts+2))[c(3:(nclusts+2))]
    pal <- c(arcpal[1:narc],
             vmhpal[1:nvmh],
             wmpal[1:nwm])
    
    names(pal) <- c(arcs,vmhs,wms)
    
    htmap <- ComplexHeatmap::Heatmap(
        x[[2]],
        top_annotation = HeatmapAnnotation(
            `BayesSpace cluster` = x[[1]]$label,
            gap = 0,
            col = list(`BayesSpace cluster`=pal),
            gp = gpar(alpha = 1)
        ),
        show_column_names = FALSE,
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        show_row_names = T,
        col = colorRamp2(
            breaks = c(-2, 0, 2),
            colors = c("blue", "white", "orange")
        ),
        cluster_columns = F, # handled by ordering the collumns of the spe above
        heatmap_legend_param = list(
            title = "Z scored log counts",
            direction = "horizontal",
            title_position = "topcenter"
        ),
        row_names_gp = gpar(fontsize = 6),
        raster_device = "CairoTIFF",
        raster_quality = 10
    )

### save plot
    CairoPDF(paste0("manuscript_plots/Supplemental/SFigparts-k152031spotplots_globalmks_intraVMHmks_intraARCmks/",n,"_globalmks.pdf"), height = 6, width = 5)
    draw(htmap,
        heatmap_legend_side = "bottom",
        annotation_legend_side = "right",
        legend_grouping = "original"
    )
    dev.off()
    
    return("kdone")
})
```


now do all this same shit with the intra-domain comparisons
```{r}
## intra-domain marker setups: extract the top 60/length(intramk[[i]][["vmh]])+length(intramk[[i]][["arc]]) genes
ng <- lapply(intramk,FUN=function(i){
    ng <- ceiling(60/(length(i[["vmh"]])+length(i[["arc"]])))
    return(ng)
})

intramk2 <- mapply(m=intramk,n=ng,SIMPLIFY=FALSE,FUN = function(m,n){
  g <- lapply(m,FUN=function(d){
    clusnames <- names(d)  
    gx <- mapply(x=d,cl=names(d),SIMPLIFY=FALSE,FUN=function(x,cl){
      if(length(clusnames)>2){
        keepnames <- c("rn",paste0(c("t_stat_","p_value_","fdr_","logFC_"),cl),"ensembl","gene")
        x <- x[,..keepnames]
      }
      setnames(x,c("rn","t","p","fdr","logFC","ens","symb"))
      
      y <- x[fdr<0.05&logFC>0]
      if (nrow(y)==0){
          y <- x[p<0.05&logFC>0]
      }
      y <- setorderv(y,"logFC",-1)
      gout <- y[1:n,symb]
      return(gout)
    })
    return(gx)
  })
  g2 <- c(g[[1]],g[[2]])
  names(g2) <- c(names(m[["vmh"]]),names(m[["arc"]]))
  return(g2)
})

intramkg <- lapply(intramk2,FUN=unlist)
intramkg <- lapply(intramkg,FUN=unique)
intramkg <- lapply(intramkg,FUN=function(x){x[!is.na(x)]})

### get expression matrices of these accordingly
intramkxpr <- mapply(g=intramkg,b=bscls.dt,SIMPLIFY=FALSE,FUN=function(g,b){
    b <- b[cl %in% grep(cl,pattern="VMH|ARC",value=T)]
    
    # subset to relevant spots
    hyptmp <- hyp2[,hyp2$key %in% b$rn]
    
    # label spots
    b <- DataFrame(b,row.names=b$rn)[colnames(hyptmp),]
    colLabels(hyptmp) <- b$cl
    
    # order as factor so we can fetch the spots per cluster in a presentable order (ARCS, VMHs, WMs)
    arcs <- grep(unique(b$cl),pattern="ARC",value=T)
    vmhs <- grep(unique(b$cl),pattern="VMH",value=T)

    hyptmp$label <- factor(hyptmp$label,levels=c(arcs,vmhs))

        
   # make a subsetted spe with reordering of the columns to be in the order of the domains we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
    i <- 1
    for (i in c(1:length(levels(hyptmp$label)))) {
        curlev <- levels(hyptmp$label)[i]
        if (i == 1) {
            cnames <- colnames(hyptmp[, hyptmp$label == curlev])
        } else {
            cnames <- c(cnames, colnames(hyptmp[, hyptmp$label == curlev]))
        }
    }
    rm(i, curlev)
    
    ### extract log counts per marker, z-normalize, return both
    pltdat <- hyptmp[g, cnames]
    pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))
    
    retlist <- list(pltdat,pltdat2)
    
    return(retlist)
})

mapply(n=names(intramkxpr),x=intramkxpr,SIMPLIFY=FALSE,FUN=function(x,n){
    arcs <- unique(grep(x[[1]]$label,pattern="ARC",value=T))
    narc <- length(arcs)
    vmhs <- unique(grep(x[[1]]$label,pattern="VMH",value=T))
    nvmh <- length(vmhs)
   
    pal <- c(arcpal[1:narc],
             vmhpal[1:nvmh])
    
    names(pal) <- c(arcs,vmhs)
    
    htmap <- ComplexHeatmap::Heatmap(
        x[[2]],
        top_annotation = HeatmapAnnotation(
            Domain = x[[1]]$label,
            gap = 0,
            col = list(Domain=pal),
            gp = gpar(alpha = 1)
        ),
        show_column_names = FALSE,
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        show_row_names = T,
        col = colorRamp2(
            breaks = c(-2, 0, 2),
            colors = c("blue", "white", "orange")
        ),
        cluster_columns = F, # handled by ordering the collumns of the spe above
        heatmap_legend_param = list(
            title = "Z scored log counts",
            direction = "horizontal",
            title_position = "topcenter"
        ),
        row_names_gp = gpar(fontsize = 6),
        raster_device = "CairoTIFF",
        raster_quality = 10
    )

### save plot
    CairoPDF(paste0("manuscript_plots/Supplemental/SFigparts-k152031spotplots_globalmks_intraVMHmks_intraARCmks/",n,"_intraDomainmks.pdf"), height = 6, width = 5)
    draw(htmap,
        heatmap_legend_side = "bottom",
        annotation_legend_side = "right",
        legend_grouping = "original"
    )
    dev.off()
    
    return("kdone")
})
```

## the last part is easy because we've already coded it all: the spotplots

helper function to extract a legend and pipe back to a diff plot: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

spotplots
```{r}
# first we need to reload the SPE with imgData and instead substitute sample names into the imgdata slot for spatialLIBD plotting functions to work, even though we're not using the imgdata
hyp2 <- readRDS("spatial_HYP/processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")



# replace sample names -- doing it from within the spe seems to update imgdata simultaneously somehow
tmpcd <- as.data.table(colData(hyp2))
tmpcd <- merge.data.table(tmpcd,sampleplotnames[,.(sample_id,manuscript_id)],by="sample_id",all.x=T)
# tmpcd[,sample_id:=manuscript_id]
tmpcd <- DataFrame(tmpcd,row.names=tmpcd$key)[colnames(hyp2),]
colData(hyp2) <- tmpcd

hyp2$sample_id <- hyp2$manuscript_id


### make the plots
mapply(x=bscls.dt,y=names(bscls.dt),SIMPLIFY=FALSE,FUN=function(x,y){
    arcs <- unique(x[cl %in% grep(cl,pattern="ARC",value=T),cl])
    narc <- length(arcs)
    vmhs <- unique(x[cl %in% grep(cl,pattern="VMH",value=T),cl])
    nvmh <- length(vmhs)
    wms <- unique(x[cl %in% grep(cl,pattern="WM",value=T),cl])
    nwm <- length(wms)
    
    b <- DataFrame(x,row.names=x$rn)[colnames(hyp2),]
    
    # nclusts <- length(unique(b[, 2]))
    
    # pal <- Polychrome::palette36.colors(n = (nclusts+2))[c(3:(nclusts+2))]
    pal <- c(arcpal[1:narc],
             vmhpal[1:nvmh],
             wmpal[1:nwm],
             "gray")
    
    names(pal) <- c(arcs,vmhs,wms,"Other")
    
    colLabels(hyp2) <- factor(b$cl,levels=c(arcs,vmhs,wms,"Other"))
    # next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
    names(pal) <- levels(hyp2$label)
    
    ## create a unified legend spanning all possible cluster values (regardless of presence in a given sample)
    leg.tmp <- ggplot(as.data.table(colData(hyp2))[,.(label,array_row,array_col)],aes(x=array_row,y=array_col,col=label))+geom_point()+
        scale_color_manual(values=as.character(pal))+theme(
                    legend.text = element_text(size = 10),
                    legend.title = element_text(size = 10, hjust = 0.5),
                    legend.spacing.y = unit(0, "cm"),
                    legend.spacing.x = unit(0.1, "cm"))
    leg.tmp2 <- extract_legend(leg.tmp)
    dev.off()
    ### make spotplots
    plt <-
        spatialLIBD::vis_grid_clus(
            hyp2,
            clustervar = "label",
            pdf_file = NULL,
            colors = pal,
            spatial = F,
            auto_crop = T,
            return_plots = T,
            point_size = 1.65,
            sort_clust = F
        )
    
    ### adjust font sizes, drop individual plot legends
    plt <- lapply(
        plt,
        FUN = function(x) {
            x +
                theme(
                    plot.title = element_text(size = 11, hjust = 0.5),
                    strip.text = element_text(size = 10),
                )+
                guides(fill="none")
        }
    )
    
    # save
    pdf(
        paste0(
            "manuscript_plots/Supplemental/SFigparts-k152031spotplots_globalmks_intraVMHmks_intraARCmks/",y,"_spotplot.pdf"),
        width = 11,
        height = 20
    )
    ### arrangeGrob() on the spotplot list --> 3 x 3 of plots from that list only, constituting the "first column" of the parent; leg.tmp2 is an additional thing being plotted in the second column of the parent. width of the plot grid is set to 10, width of the second column (legend only) set to 2.
    do.call("grid.arrange",list(arrangeGrob(grobs=plt,ncol=3,nrow=4),leg.tmp2,ncol=2,widths=c(9.5,1.5)))
    dev.off()
})
```

## reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] BiocParallel_1.38.0         parallelly_1.38.0          
 [3] colorout_1.3-0.2            ggtext_0.1.2               
 [5] viridis_0.6.5               viridisLite_0.4.2          
 [7] circlize_0.4.16             Cairo_1.6-2                
 [9] ggrastr_1.0.2               ComplexHeatmap_2.20.0      
[11] spatialLIBD_1.16.2          SpatialExperiment_1.14.0   
[13] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
[15] Biobase_2.64.0              GenomicRanges_1.56.1       
[17] GenomeInfoDb_1.40.1         IRanges_2.38.1             
[19] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[21] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[23] gridExtra_2.3               ggplot2_3.5.1              
[25] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0             bitops_1.0-7             
  [4] filelock_1.0.3            fields_16.2               tibble_3.2.1             
  [7] XML_3.99-0.17             lifecycle_1.0.4           edgeR_4.2.0              
 [10] doParallel_1.0.17         rprojroot_2.0.4           lattice_0.22-6           
 [13] magrittr_2.0.3            limma_3.60.3              plotly_4.10.4            
 [16] sass_0.4.9                rmarkdown_2.27            jquerylib_0.1.4          
 [19] yaml_2.3.9                httpuv_1.6.15             spam_2.10-0              
 [22] sessioninfo_1.2.2         cowplot_1.1.3             DBI_1.2.3                
 [25] RColorBrewer_1.1-3        golem_0.4.1               maps_3.4.2               
 [28] abind_1.4-5               zlibbioc_1.50.0           purrr_1.0.2              
 [31] RCurl_1.98-1.14           rappdirs_0.3.3            GenomeInfoDbData_1.2.12  
 [34] ggrepel_0.9.5             irlba_2.3.5.1             DelayedMatrixStats_1.26.0
 [37] codetools_0.2-20          DelayedArray_0.30.1       xml2_1.3.6               
 [40] DT_0.33                   scuttle_1.14.0            shape_1.4.6.1            
 [43] tidyselect_1.2.1          farver_2.1.2              UCSC.utils_1.0.0         
 [46] ScaledMatrix_1.12.0       shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [49] GenomicAlignments_1.40.0  jsonlite_1.8.8            GetoptLong_1.0.5         
 [52] BiocNeighbors_1.22.0      scater_1.32.0             iterators_1.0.14         
 [55] foreach_1.5.2             tools_4.4.1               Rcpp_1.0.13              
 [58] glue_1.7.0                SparseArray_1.4.8         xfun_0.45                
 [61] here_1.0.1                dplyr_1.1.4               withr_3.0.0              
 [64] BiocManager_1.30.23       fastmap_1.2.0             fansi_1.0.6              
 [67] digest_0.6.36             rsvd_1.0.5                R6_2.5.1                 
 [70] mime_0.12                 colorspace_2.1-0          RSQLite_2.3.7            
 [73] config_0.3.2              utf8_1.2.4                tidyr_1.3.1              
 [76] generics_0.1.3            rtracklayer_1.64.0        httr_1.4.7               
 [79] htmlwidgets_1.6.4         S4Arrays_1.4.1            scatterplot3d_0.3-44     
 [82] pkgconfig_2.0.3           gtable_0.3.5              blob_1.2.4               
 [85] XVector_0.44.0            htmltools_0.5.8.1         dotCall64_1.1-1          
 [88] clue_0.3-65               scales_1.3.0              png_0.1-8                
 [91] attempt_0.3.1             knitr_1.48                rstudioapi_0.16.0        
 [94] rjson_0.2.21              curl_5.2.1                cachem_1.1.0             
 [97] GlobalOptions_0.1.2       Polychrome_1.5.1          BiocVersion_3.19.1       
[100] parallel_4.4.1            vipor_0.4.7               AnnotationDbi_1.66.0     
[103] restfulr_0.0.15           pillar_1.9.0              vctrs_0.6.5              
[106] promises_1.3.0            BiocSingular_1.20.0       dbplyr_2.5.0             
[109] beachmat_2.20.0           cluster_2.1.6             xtable_1.8-4             
[112] beeswarm_0.4.0            paletteer_1.6.0           evaluate_0.24.0          
[115] magick_2.8.3              cli_3.6.3                 locfit_1.5-9.10          
[118] compiler_4.4.1            Rsamtools_2.20.0          crayon_1.5.3             
[121] labeling_0.4.3            rematch2_2.1.2            ggbeeswarm_0.7.2         
[124] munsell_0.5.1             Biostrings_2.72.1         tiff_0.1-12              
[127] lazyeval_0.2.2            Matrix_1.7-0              ExperimentHub_2.12.0     
[130] benchmarkme_1.0.8         sparseMatrixStats_1.16.0  bit64_4.0.5              
[133] KEGGREST_1.44.1           statmod_1.5.0             shiny_1.8.1.1            
[136] AnnotationHub_3.12.0      gridtext_0.1.5            memoise_2.0.1            
[139] bslib_0.7.0               benchmarkmeData_1.0.4     bit_4.0.5                
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-10-07
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings             2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   Cairo                * 1.6-2     2023-11-28 [1] CRAN (R 4.4.0)
   circlize             * 0.4.16    2024-02-20 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   clue                   0.3-65    2023-09-23 [1] CRAN (R 4.4.0)
   cluster                2.1.6     2023-12-01 [1] CRAN (R 4.4.1)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   ComplexHeatmap       * 2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   GetoptLong             1.0.5     2020-12-15 [1] CRAN (R 4.4.0)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr              * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   ggtext               * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   GlobalOptions          0.1.2     2020-06-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome             1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scatterplot3d          0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
   scuttle                1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shape                  1.4.6.1   2024-02-23 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                   0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis              * 0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite          * 0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xml2                   1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────