---
title: "Xenium UMAPs and nonARCVMH clusters heatmap"
author: "Bernie Mulvey"
date: "2024-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(escheR)
library(ComplexHeatmap)
library(ggrastr) # render plots with bajillions of points as raster so they can be manipulated in eg Illustrator--Fig 2B
library(Cairo) # pdf device for ^ to properly save; also used for complexheatmap raster functions
library(circlize) #colorRamp2
library(viridis) # magma
library(ggtext) # mixed font formats in ggplot elements

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


### Import the saved UMAP coords to churn out four versions of the plot: by cell supercluster, by sex, by donor, and by batch.
```{r}
rastumap <- fread("xenium_HYP/processed-data/04_general_QC_and_normalization/02a-genetarg-lctUMAP_rawctCAandUMAP_w_metadata.txt.gz")

bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"cl")

bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
bkanno <- unique(bkanno[,.(clus,bjm_annot)])
bkanno <- bkanno[bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T,invert=T)]

bkcl <- merge(bkcl,bkanno,by.x="cl",by.y="clus")
## create supercluster labels to group related cell types together
bkcl[,supercl:="ifthisisplottedigoofed"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="VMH",value=T),supercl:="VMH (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="ARC",value=T),supercl:="ARC (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Tanyc|Alar",value=T),supercl:="Tanycytes, Portal Vasc. (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Oligo",value=T),supercl:="Oligo (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Astro",value=T),supercl:="Astro (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Microg",value=T),supercl:="Microglia (3)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Supraopt",value=T),supercl:="SON (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Unsure_AVP",value=T),supercl:="Non-SON AVP+OXT+ (1)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Macrop|Periph|Vascul|Endothel",value=T),supercl:="Vascular and Peripheral Immune (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="GABA",value=T),supercl:="Other GABAergic Neurons (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Periven",value=T),supercl:="PeriVN (Excitatory) (1)"]
### the parenthetical numbers add up to 33, so we're good.

# key is a field from the colData slot of the SPE that was exported, and is identical to the rownames in case they are lost during switches between d.t. and d.f elsewhere
rastumap <- rastumap[key %in% bkcl$rn]
rastumap <- merge.data.table(rastumap,bkcl,by.x="key",by.y="rn")

### load palette for this
pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")

xenpal <- pals[["xen_cellgroups"]]

rastumap[supercl=="Vascular and Peripheral Immune (5)",supercl:="Vascular, Peripheral\nImmune (5)"]
names(xenpal)[names(xenpal)=="Vascular and Peripheral Immune (5)"] <- "Vascular, Peripheral\nImmune (5)"

stopifnot(all(unique(rastumap$supercl) %in% names(xenpal)))


## superclsuters
p <- ggplot(rastumap,aes(x=UMAP1_CA,y=UMAP2_CA,color=supercl))+ 
   xlab("UMAP dim 1")+
   ylab("UMAP dim 2")+
   labs(color="Cell group (N clusters)")+
   scale_color_manual(values=xenpal)+
   guides(color=guide_legend(ncol=3,override.aes=list(size=1.5)))+
   scale_x_continuous(expand=c(0,0))+
   scale_y_continuous(expand=c(0,0))+
   theme(legend.position = "bottom",
         legend.direction="vertical",
         title = element_blank(),
         axis.title.y = element_text(size=9,margin=margin(0.02,0,0,0,unit="in")),
         axis.title.x = element_text(size=9,margin=margin(0,0.02,0,0,unit="in")),
         axis.text.x=element_blank(),
         axis.text.y=element_blank(),
         legend.title = element_text(size=8,margin=margin(-0.04,0,0.01,0,"in")),
         legend.text = element_text(size=7),
         legend.background = element_blank(),
         legend.key.height = ggplot2::unit(0.1,"in"),
         legend.key.width=ggplot2::unit(0.05,"in"),
         plot.margin=margin(0.025,0.025,0,0.025,unit = "in"),
         legend.margin=margin(-0.06,0,0.025,0,unit="in"),
         legend.key.spacing.x = ggplot2::unit(0.05,"in"),
         legend.key.spacing.y=ggplot2::unit(0.025,"in"),
         panel.grid = element_blank(),
         axis.ticks = element_blank())+
   rasterize(geom_point(size=0.05,stroke=0),dpi=600,layer="Points",dev = "cairo_png")


pdf("manuscript_plots/Supplemental/Xenium UMAPs and nonARCVMH clusters heatmap/Xenium_supercluster_UMAP.pdf",width=4,height=3)
p
dev.off()

### by donor
rastumap[,BrNum:=paste0(BrNum,"_",gsub(Sex,pattern="^(.).*$",replacement="\\1"))]

p <- ggplot(rastumap,aes(x=UMAP1_CA,y=UMAP2_CA,color=BrNum))+ 
   xlab("UMAP dim 1")+
   ylab("UMAP dim 2")+
   labs(color="Donor")+
   guides(color=guide_legend(ncol=3,override.aes=list(size=1.5)))+
   scale_x_continuous(expand=c(0,0))+
   scale_y_continuous(expand=c(0,0))+
   theme(legend.position = "bottom",
         legend.direction="vertical",
         title = element_blank(),
         axis.title.y = element_text(size=9,margin=margin(0.02,0,0,0,unit="in")),
         axis.title.x = element_text(size=9,margin=margin(0,0.02,0,0,unit="in")),
         axis.text.x=element_blank(),
         axis.text.y=element_blank(),
         legend.title = element_text(size=8,margin=margin(-0.04,0,0.01,0,"in")),
         legend.text = element_text(size=7),
         legend.background = element_blank(),
         legend.key.height = ggplot2::unit(0.1,"in"),
         legend.key.width=ggplot2::unit(0.05,"in"),
         plot.margin=margin(0.025,0.025,0,0.025,unit = "in"),
         legend.margin=margin(0,0,0.0862,0,unit="in"),
         legend.key.spacing.x = ggplot2::unit(0.05,"in"),
         legend.key.spacing.y=ggplot2::unit(0.075,"in"),
         panel.grid = element_blank(),
         axis.ticks = element_blank())+
   rasterize(geom_point(size=0.05,stroke=0),dpi=600,layer="Points",dev = "cairo_png")


pdf("manuscript_plots/Supplemental/Xenium UMAPs and nonARCVMH clusters heatmap/Xenium_donor_UMAP.pdf",width=4,height=3)
p
dev.off()

## by sex
p <- ggplot(rastumap,aes(x=UMAP1_CA,y=UMAP2_CA,color=Sex))+ 
   xlab("UMAP dim 1")+
   ylab("UMAP dim 2")+
   labs(color="Sex")+
   guides(color=guide_legend(ncol=3,override.aes=list(size=1.5,alpha=1)))+
   scale_x_continuous(expand=c(0,0))+
   scale_y_continuous(expand=c(0,0))+
   theme(legend.position = "bottom",
         legend.direction="vertical",
         title = element_blank(),
         axis.title.y = element_text(size=9,margin=margin(0.02,0,0,0,unit="in")),
         axis.title.x = element_text(size=9,margin=margin(0,0.02,0,0,unit="in")),
         axis.text.x=element_blank(),
         axis.text.y=element_blank(),
         legend.title = element_text(size=8,margin=margin(-0.04,0,0.01,0,"in")),
         legend.text = element_text(size=7),
         legend.background = element_blank(),
         legend.key.height = ggplot2::unit(0.1,"in"),
         legend.key.width=ggplot2::unit(0.05,"in"),
         plot.margin=margin(0.025,0.025,0,0.025,unit = "in"),
         legend.margin=margin(-0.06,0,0.025,0,unit="in"),
         legend.key.spacing.x = ggplot2::unit(0.05,"in"),
         legend.key.spacing.y=ggplot2::unit(0.075,"in"),
         panel.grid = element_blank(),
         axis.ticks = element_blank())+
   rasterize(geom_point(size=0.05,stroke=0,alpha=0.2),dpi=600,layer="Points",dev = "cairo_png")


pdf("manuscript_plots/Supplemental/Xenium UMAPs and nonARCVMH clusters heatmap/Xenium_sex_UMAP.pdf",width=4,height=2.5)
p
dev.off()

# by batch
unique(rastumap$Date_Experiment_Began)
rastumap[Date_Experiment_Began=="2/21/24",Batch:="Batch 2"]
rastumap[Date_Experiment_Began=="10/17/23",Batch:="Batch 1"]


p <- ggplot(rastumap,aes(x=UMAP1_CA,y=UMAP2_CA,color=Batch))+ 
   xlab("UMAP dim 1")+
   ylab("UMAP dim 2")+
   labs(color="Xenium Slide Batch")+
   guides(color=guide_legend(ncol=3,override.aes=list(size=1.5,alpha=1)))+
   scale_x_continuous(expand=c(0,0))+
   scale_y_continuous(expand=c(0,0))+
   theme(legend.position = "bottom",
         legend.direction="vertical",
         title = element_blank(),
         axis.title.y = element_text(size=9,margin=margin(0.02,0,0,0,unit="in")),
         axis.title.x = element_text(size=9,margin=margin(0,0.02,0,0,unit="in")),
         axis.text.x=element_blank(),
         axis.text.y=element_blank(),
         legend.title = element_text(size=8,margin=margin(-0.04,0,0.01,0,"in")),
         legend.text = element_text(size=7),
         legend.background = element_blank(),
         legend.key.height = ggplot2::unit(0.1,"in"),
         legend.key.width=ggplot2::unit(0.05,"in"),
         plot.margin=margin(0.025,0.025,0,0.025,unit = "in"),
         legend.margin=margin(-0.06,0,0.025,0,unit="in"),
         legend.key.spacing.x = ggplot2::unit(0.05,"in"),
         legend.key.spacing.y=ggplot2::unit(0.075,"in"),
         panel.grid = element_blank(),
         axis.ticks = element_blank())+
   rasterize(geom_point(size=0.05,stroke=0,alpha=0.2),dpi=600,layer="Points",dev = "cairo_png")


pdf("manuscript_plots/Supplemental/Xenium UMAPs and nonARCVMH clusters heatmap/Xenium_batch_UMAP.pdf",width=4,height=2.5)
p
dev.off()

```


### Heatmap of top couplefew clustermarkers for ALL clusters
## prepare expression matrices for cluster markers
## # Z-scale logcounts, make a subsetted spe with the columns already ordered by cluster since complexheatmap was giving issues with this, make plot.

```{r}
# setup: SPE, cluster labels
hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

arcvmhanno <- fread("manuscript_plot_code/xARCxVMH_cluster_detailedlabs_andplotformatnames.txt")
bkcl <- merge.data.table(bkcl,arcvmhanno[,.(cluster_id,subclus_domain,plotclusname)],by.x="cl",by.y="cluster_id",all.x=T)
bkcl[!is.na(plotclusname),bjm_annot:=paste0("x",subclus_domain," ",gsub(plotclusname,pattern="\\*",replacement=""))]
bkcl[,bjm_annot:=gsub(bjm_annot,pattern="xVMH xVMH",replacement="xVMH")]


# label spots
hypx <- hypx[,colnames(hypx) %in% bkcl$rn]
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
colLabels(hypx) <- bkcl$bjm_annot

# get order that we'll eventually apply as factor with all xarcs and xvmhs together
labs <- unique(hypx$label)

clusord <- c(grep(labs,pattern="xARC",value=T),
             grep(labs,pattern="xVMH",value=T),
             grep(labs,pattern="Periven",value=T),
             grep(labs,pattern="Unsure_AVP",value=T),
             grep(labs,pattern="Supraoptic",value=T),
             grep(labs,pattern="GABA",value=T),
             grep(labs,pattern="Alar",value=T),
             grep(labs,pattern="Tanycy",value=T),
             grep(labs,pattern="^Oligo",value=T),
             grep(labs,pattern="Astro",value=T),
             grep(labs,pattern="Microgl",value=T),
             grep(labs,pattern="Macroph|immune",value=T),
             grep(labs,pattern="Endothel",value=T),
             grep(labs,pattern="Vascul",value=T))

stopifnot(length(unique(clusord)) == length(unique(labs)))

# get the column name order to make an spe with reordering of the cells to be in the order of the clusters we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
i <- 1
for (i in c(1:length(clusord))) {
    curlev <- clusord[i]
    if (i == 1) {
        cnames <- colnames(hypx[, hypx$label == curlev])
    } else {
        cnames <- c(cnames, colnames(hypx[, hypx$label == curlev]))
    }
}
rm(i, curlev)
    

### now get clusterwise markers, in the same order
xenclusmks <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/03a-M0l0kg6_reproducibleClustersOnly_spLIBD_markers.RDS")

xenclusmks <- melt.data.table(as.data.table(xenclusmks[[2]]),id.cols=c("gene","ensembl"))

xenclusmks[,clus:=gsub(variable,pattern="^.*_(X.*)$",replacement="\\1")]
xenclusmks[,stat:=gsub(variable,pattern="^(.*)_X.*$",replacement="\\1")]
xenclusmks <- xenclusmks[stat=="t_stat"]
xenclusmks <- xenclusmks[value>0]

xenclusmks <- merge.data.table(xenclusmks,unique(as.data.table(bkcl)[,.(cl,bjm_annot)]),by.x="clus",by.y="cl")
stopifnot(all(unique(xenclusmks$bjm_annot) %in% unique(as.character(hypx$label))))

### get top 4 genes per cluster, grabbing genes in the same cluster order that the clusters are factored into

i<-1
for (i in c(1:length(clusord))){
    tmp <- xenclusmks[bjm_annot==clusord[i]]
    setorderv(tmp,"value",-1)
    if(i==1){
        hmgs <- tmp[1:4,gene]
    } else{
        hmgs <- unique(c(hmgs,tmp[1:4,gene]))
    }
    rm(tmp)
}

### extract mean log counts per marker per cluster with the columns in per-cluster order. 
pltdat <- hypx[hmgs, cnames]
pltdat2 <- as.data.table(as.data.frame(as.matrix(logcounts(pltdat))),keep.rownames = T)

setnames(pltdat2,"rn","gene")
pltdat2 <- melt(pltdat2,id.vars ="gene")
pltdat2 <- merge.data.table(pltdat2,as.data.table(colData(pltdat),keep.rownames=T)[,.(rn,label)],by.x="variable",by.y="rn")

pltdat3 <- pltdat2[!is.na(value),mean(value),by=c("label","gene")]
setnames(pltdat3,"V1","meanlct")

pltdat4 <- dcast(pltdat3,gene~label,value.var="meanlct")
pltdat4 <- as.data.frame(pltdat4[,c(2:ncol(pltdat4)),with=F],row.names = pltdat4$gene)

pltdat4<-as.matrix(pltdat4)
pltdat4 <- t(pltdat4)

# ensure genes are in the order we want them
pltdat4 <- pltdat4[,hmgs]
# ensure clusters are in the order we want them
stopifnot(all(clusord %in% rownames(pltdat4)))
pltdat4 <- pltdat4[clusord,]

# remove underscores from cluster labels for plotting
rownames(pltdat4) <- gsub(rownames(pltdat4),pattern="_",replacement=" ")

## conserve memory for this big MF plot coming up
rm(hypx,rastumap,xenclusmks,pltdat,pltdat2)
gc(full=T)
```

# make + save plot
```{r}
## determine what the starting blue at 0 should be if the palette were symmetric about 1, use that for the real pal color
calcpal <- colorRamp2(
            breaks = c(1+(-max(pltdat3$meanlct)), 1, max(pltdat3$meanlct)),
            colors = c("blue", "white", "orange"),
        )
pltdat3[,palcol:=calcpal(meanlct)]

pal <- colorRamp2(breaks=c(0,1,max(pltdat3$meanlct)),
                  colors=c(pltdat3[meanlct==min(meanlct),palcol],"white","orange"))

htmap <- ComplexHeatmap::Heatmap(
        pltdat4,
        cluster_columns = FALSE, # handled by ordering the collumns of the spe above
        cluster_rows=FALSE,
        #top_annotation = HeatmapAnnotation(
        #     gap = 0,
        #     show_annotation_name = T,
        #     gp = gpar(alpha = 1)
        # ),
        show_column_names = TRUE,# TRUE prints EVERY column's name.
        show_column_dend = FALSE,
        show_row_dend = FALSE,
        show_row_names = T,
        col=pal,
        #col = colorRamp2(
        #    breaks = c(0, quantile(pltdat3$meanlct,0.25), max(pltdat3$meanlct)),
         #   colors = c("#8080ff", "white", "orange"),
            
        #),
        heatmap_legend_param = list(
            title = "Mean cluster log counts",
            direction = "horizontal",
            title_position = "topcenter",
            title_gp=gpar(fontsize=7,fontface="bold")
        ),
        row_names_gp = gpar(fontsize=6),
        column_names_gp = gpar(fontsize = 6,fontface="italic"),
        raster_device = "CairoTIFF",
        raster_quality = 10
    )

pdf("manuscript_plots/Supplemental/Xenium UMAPs and nonARCVMH clusters heatmap/allclusts_top4each_heatmap.pdf",height=4.5,width=8)
draw(htmap,heatmap_legend_side="bottom")
dev.off()

```

reprod inf
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] SpatialFeatureExperiment_1.6.1 BiocParallel_1.38.0           
 [3] parallelly_1.38.0              colorout_1.3-0.2              
 [5] ggtext_0.1.2                   viridis_0.6.5                 
 [7] viridisLite_0.4.2              circlize_0.4.16               
 [9] Cairo_1.6-2                    ggrastr_1.0.2                 
[11] ComplexHeatmap_2.20.0          escheR_1.4.0                  
[13] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[15] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[17] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
[19] IRanges_2.38.1                 S4Vectors_0.42.1              
[21] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[23] matrixStats_1.3.0              gridExtra_2.3                 
[25] ggplot2_3.5.1                  data.table_1.15.4             
[27] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3        wk_0.9.2                  rstudioapi_0.16.0        
  [4] jsonlite_1.8.8            shape_1.4.6.1             magrittr_2.0.3           
  [7] ggbeeswarm_0.7.2          magick_2.8.3              farver_2.1.2             
 [10] rmarkdown_2.27            GlobalOptions_0.1.2       zlibbioc_1.50.0          
 [13] vctrs_0.6.5               spdep_1.3-5               DelayedMatrixStats_1.26.0
 [16] RCurl_1.98-1.14           terra_1.7-78              htmltools_0.5.8.1        
 [19] S4Arrays_1.4.1            BiocNeighbors_1.22.0      s2_1.1.6                 
 [22] Rhdf5lib_1.26.0           SparseArray_1.4.8         rhdf5_2.48.0             
 [25] spData_2.3.1              KernSmooth_2.23-24        htmlwidgets_1.6.4        
 [28] lifecycle_1.0.4           iterators_1.0.14          pkgconfig_2.0.3          
 [31] Matrix_1.7-0              R6_2.5.1                  fastmap_1.2.0            
 [34] GenomeInfoDbData_1.2.12   clue_0.3-65               digest_0.6.36            
 [37] colorspace_2.1-0          rprojroot_2.0.4           dqrng_0.4.1              
 [40] beachmat_2.20.0           labeling_0.4.3            fansi_1.0.6              
 [43] httr_1.4.7                abind_1.4-5               compiler_4.4.1           
 [46] here_1.0.1                proxy_0.4-27              withr_3.0.0              
 [49] doParallel_1.0.17         tiff_0.1-12               DBI_1.2.3                
 [52] HDF5Array_1.32.0          R.utils_2.12.3            sessioninfo_1.2.2        
 [55] DelayedArray_0.30.1       scatterplot3d_0.3-44      rjson_0.2.21             
 [58] classInt_0.4-10           units_0.8-5               tools_4.4.1              
 [61] vipor_0.4.7               beeswarm_0.4.0            R.oo_1.26.0              
 [64] glue_1.7.0                EBImage_4.46.0            rhdf5filters_1.16.0      
 [67] gridtext_0.1.5            sf_1.0-16                 cluster_2.1.6            
 [70] generics_0.1.3            gtable_0.3.5              R.methodsS3_1.8.2        
 [73] class_7.3-22              sp_2.1-4                  xml2_1.3.6               
 [76] utf8_1.2.4                XVector_0.44.0            foreach_1.5.2            
 [79] pillar_1.9.0              limma_3.60.3              dplyr_1.1.4              
 [82] lattice_0.22-6            deldir_2.0-4              tidyselect_1.2.1         
 [85] locfit_1.5-9.10           sfheaders_0.4.4           scuttle_1.14.0           
 [88] knitr_1.48                edgeR_4.2.0               xfun_0.45                
 [91] statmod_1.5.0             DropletUtils_1.24.0       UCSC.utils_1.0.0         
 [94] fftwtools_0.9-11          boot_1.3-30               yaml_2.3.9               
 [97] evaluate_0.24.0           codetools_0.2-20          tibble_3.2.1             
[100] Polychrome_1.5.1          cli_3.6.3                 munsell_0.5.1            
[103] Rcpp_1.0.13               zeallot_0.1.0             png_0.1-8                
[106] parallel_4.4.1            jpeg_0.1-10               sparseMatrixStats_1.16.0 
[109] bitops_1.0-7              scales_1.3.0              e1071_1.7-14             
[112] crayon_1.5.3              GetoptLong_1.0.5         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-11-19
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   Cairo                    * 1.6-2     2023-11-28 [1] CRAN (R 4.4.0)
   circlize                 * 0.4.16    2024-02-20 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   clue                       0.3-65    2023-09-23 [1] CRAN (R 4.4.0)
   cluster                    2.1.6     2023-12-01 [1] CRAN (R 4.4.1)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   ComplexHeatmap           * 2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   escheR                   * 1.4.0     2024-05-16 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   GetoptLong                 1.0.5     2020-12-15 [1] CRAN (R 4.4.0)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggtext                   * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   GlobalOptions              0.1.2     2020-06-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext                   0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome                 1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scatterplot3d              0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shape                      1.4.6.1   2024-02-23 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                  * 0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite              * 0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   xml2                       1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────