---
title: "SF-Xenium VMH and ARC clusters by cluster by sample"
output: html_document
date: "2024-09-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table) # Preferred data manipulation package
library(ggplot2) # Dependency for several plotting functions
library(ggtext) # more character types in ggplots
library(ggrastr) # avoid making raster images one can't even open
library(SpatialExperiment) # Visium data framework
library(SpatialFeatureExperiment) # Xenium data framework
library(viridis) # palettes
library(Polychrome) # better palettes
library(ggh4x) # center rows of facets when uneven lengths

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### XENIUM setup ###
```{r}
## manuscript sample ids
mscriptsampleids <- fread("standardized_sampleids_for_plotting.txt")

hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

bksmooth <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt") ## xenium domain assignments after smoothing, by xenium cell

## but drop discarded cell clusters, too
bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"cl")

bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

## drop sample-specific clusters, append cluster annotations 
bkanno <- unique(bkanno[,.(clus,bjm_annot)])
bkanno <- bkanno[bjm_annot!="DISCARD"&bjm_annot!="VMH_4_DISCARD"]
## convert ARC_1..etc to more descriptive cluster IDs defined based on ARC/VMH specific cluster marker analyses

arcvmhanno <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04e_ARC_and_VMH_cluster_identities_based_on_cellgroup_1vall.txt")
## simplify these by removing VMH/ARC and underscores; asterisk-ize so the gene names  can be italiczed in ggplot elements; remove vmh and arc from the names but add a temporary column here so we know they're VMH and ARC for mapping onto the de result sets
arcvmhanno[,subclus_domain:=gsub(subclus_annot,pattern="^(VMH)_.*$",replacement="\\1")]
arcvmhanno[,subclus_annot:=gsub(subclus_annot,pattern="VMH_",replacement="")]

arcvmhanno[subclus_domain!="VMH",subclus_domain:=gsub(subclus_annot,pattern="^(ARC)_.*$",replacement="\\1")]
arcvmhanno[,subclus_annot:=gsub(subclus_annot,pattern="ARC_",replacement="")]

arcvmhanno[,subclus_annot:=gsub(subclus_annot,pattern="_",replacement=" ")]

arcvmhanno[subclus_annot=="VMH-Glia-Mixed",subclus_annot:="VMH-Glia Mixed"]
arcvmhanno[subclus_annot=="SLC17A6-CRHR2",subclus_annot:="*SLC17A6*-*CRHR2*"]
arcvmhanno[subclus_annot=="FEZF1-LAMP5-NRGN",subclus_annot:="*FEZF1*-*LAMP5*-*NRGN*"]
arcvmhanno[subclus_annot=="HCRT-SLC6A3-TAC1",subclus_annot:="*HCRT*-*SLC6A3*-*TAC1*"]
arcvmhanno[subclus_annot=="AGRP",subclus_annot:="*AGRP*"]
arcvmhanno[subclus_annot=="TAC3-AR-ESR1",subclus_annot:="*TAC3*-*AR*-*ESR1*"]
arcvmhanno[subclus_annot=="POMC-Oligo-Mixed",subclus_annot:="*POMC*-Oligo Mixed"]
arcvmhanno[subclus_annot=="Excit-SLC17A7-Glial-Mixed",subclus_annot:="*SLC17A7*-Glia Mixed"]
arcvmhanno[subclus_annot=="GAL-TRH-GHRH",subclus_annot:="*GAL*-*TRH*-*GHRH*"]

## append these to the other annots and replace bjm_annot with the ARC/VMH cell type labels where applicable
bkanno <- merge.data.table(bkanno,arcvmhanno,by="bjm_annot",all.x=T)
bkanno[!is.na(subclus_annot),bjm_annot:=subclus_annot]

## this drops rows that were assigned one of the discarded types (ie no longer in the cell type annotations)
bkcl <- merge.data.table(bkcl,bkanno,by.x="cl",by.y="clus")
## and now we can drop excluded clusters from xenium spe
hypx <- hypx[,colnames(hypx) %in% bkcl$rn]

## append cluster annots to xenium spe
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
bkcl$rn <- NULL
hypx$banksyclus <- bkcl$bjm_annot

## assign the domain labels to the xenium cells
bksmooth[dualVMHARC4=="other",dualVMHARC4:="Other"]
bksmooth <- DataFrame(bksmooth,row.names=bksmooth$rn)[colnames(hypx),]
hypx$dom <- bksmooth$dualVMHARC4


## we only need the VMH domain and VMH clusters or ARC domain and ARC clusters for these plots
hypv <- hypx[,hypx$dom=="VMH"]
hypv <- hypv[,hypv$banksyclus %in% arcvmhanno[subclus_domain=="VMH",subclus_annot]]
stopifnot(length(unique(hypv$banksyclus))==nrow(arcvmhanno[subclus_domain=="VMH"]))

hypa <- hypx[,hypx$dom=="ARC"]
hypa <- hypa[,hypa$banksyclus %in% arcvmhanno[subclus_domain=="ARC",subclus_annot]]
stopifnot(length(unique(hypa$banksyclus))==nrow(arcvmhanno[subclus_domain=="ARC"]))
```


```{r}
pals <- readRDS("manuscript_plot_code/domain_and_xencluster_palettes_CHECKPALNAMESBEFOREUS.RDS")

vmhpal <- pals[["XenVMH"]]
stopifnot(sum(names(vmhpal)%in%unique(hypx$banksyclus))==length(names(vmhpal)))

arcpal <- pals[["XenARC"]]
stopifnot(sum(names(arcpal)%in%unique(hypx$banksyclus))==length(names(arcpal)))
```


## plot clusters in each sample, faceted by cluster for a sample. then arrange those into a grid. don't replot the legend every time; extract it once, pin it to the side after.

helper function to extract a legend and pipe back to a diff plot: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}
```

### VMH
```{r}
## order the clusters so that the HCRT one is last
hypv$banksyclus <- factor(hypv$banksyclus,levels=c("*SLC17A6*-*CRHR2*","*FEZF1*-*LAMP5*-*NRGN*","VMH-Glia Mixed","*HCRT*-*SLC6A3*-*TAC1*"))

## make a plot w legend to extract from

pltspe <- hypv[,hypv$sample_id=="X86_reg1"]
pltspe <- as.data.table(cbind(spatialCoords(pltspe),as.data.table(colData(pltspe))))[,.(sdimx,sdimy,banksyclus)]
    
tmplegplt <- ggplot(pltspe,aes(x=sdimx,y=sdimy,color=banksyclus))+
    rasterize(geom_point(size=0.075,stroke=0.375),res=450,dev="cairo_png")+
    scale_color_manual(values=vmhpal)+
    guides(color=guide_legend(override.aes = list(size=1)))+
    labs(color="Xenium VMH\nCluster")+
    facet_wrap(~banksyclus)+
    theme(axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.title.y=element_blank(),
          axis.title.x=element_blank(),
          axis.ticks = element_blank(),
          legend.background = element_blank(),
          legend.box=element_blank(),
          #legend.margin=margin(0,0,0,-0.15,unit="in"),
          #plot.margin = margin(0.01,0.02,0,0,unit="in"),
          legend.text = element_markdown(margin=margin(0,0,0,-0.075,"in")),
          strip.background = element_blank(),
          strip.text = element_text(size=10))

leg.tmp <- extract_legend(tmplegplt)
dev.off()
rm(tmplegplt,pltspe)
###

## actual plots
allplts <- lapply(unique(hypv$sample_id),FUN=function(s){
    pltspe <- hypv[,hypv$sample_id==s]
    pltspe <- as.data.table(cbind(spatialCoords(pltspe),as.data.table(colData(pltspe))))[,.(sdimx,sdimy,banksyclus)]
    
    samplts <- ggplot(pltspe,aes(x=sdimx,y=sdimy,color=banksyclus))+
        rasterize(geom_point(size=0.075,stroke=0.375),res=450,dev="cairo_png")+
        facet_wrap(~banksyclus)+
        scale_color_manual(values = vmhpal)+
        guides(color="none")+
        ggtitle(mscriptsampleids[sample_id==s,manuscript_id])+
        theme(axis.text.x=element_blank(),
              axis.text.y=element_blank(),
              axis.title.y=element_blank(),
              axis.title.x=element_blank(),
              axis.ticks = element_blank(),
              legend.background = element_blank(),
              legend.box=element_blank(),
              panel.grid = element_blank(),
              #legend.margin=margin(0,0,0,-0.15,unit="in"),
              #plot.margin = margin(0.01,0.02,0,0,unit="in"),
              legend.text = element_markdown(margin=margin(0,0,0,-0.075,"in")),
              strip.background = element_blank(),
              strip.text = element_markdown(size=10),
              )
    
    return(samplts)
})

### arrangeGrob() on the spotplot list --> constituting the "first column" of the parent; leg.tmp is an additional thing being plotted in the second column of the parent. width of the plot grid is set to 30, width of the second column (legend only) set to 2.
pdf("manuscript_plots/Supplemental/SF-VMH_cluster_facets_allsamples.pdf",height=32,width=25)
    do.call("grid.arrange",list(arrangeGrob(grobs=allplts,ncol=4),leg.tmp,ncol=2,widths=c(30,2)))
dev.off()
```

### ARC:
```{r}
## make a plot w legend to extract from

pltspe <- hypa[,hypa$sample_id=="X86_reg1"]
pltspe <- as.data.table(cbind(spatialCoords(pltspe),as.data.table(colData(pltspe))))[,.(sdimx,sdimy,banksyclus)]
    
tmplegplt <- ggplot(pltspe,aes(x=sdimx,y=sdimy,color=banksyclus))+
    rasterize(geom_point(size=0.075,stroke=0.375),res=450,dev="cairo_png")+
    scale_color_manual(values=arcpal)+
    guides(color=guide_legend(override.aes = list(size=1)))+
    labs(color="Xenium VMH\nCluster")+
    facet_wrap(~banksyclus)+
    theme(axis.text.x=element_blank(),
          axis.text.y=element_blank(),
          axis.title.y=element_blank(),
          axis.title.x=element_blank(),
          axis.ticks = element_blank(),
          legend.background = element_blank(),
          legend.box=element_blank(),
          #legend.margin=margin(0,0,0,-0.15,unit="in"),
          #plot.margin = margin(0.01,0.02,0,0,unit="in"),
          legend.text = element_markdown(margin=margin(0,0,0,-0.075,"in")),
          strip.background = element_blank(),
          strip.text = element_text(size=10))

leg.tmp <- extract_legend(tmplegplt)
dev.off()
rm(tmplegplt,pltspe)
###

## actual plots
allplts <- lapply(unique(hypa$sample_id),FUN=function(s){
    pltspe <- hypa[,hypa$sample_id==s]
    pltspe <- as.data.table(cbind(spatialCoords(pltspe),as.data.table(colData(pltspe))))[,.(sdimx,sdimy,banksyclus)]
    
    samplts <- ggplot(pltspe,aes(x=sdimx,y=sdimy,color=banksyclus))+
        rasterize(geom_point(size=0.075,stroke=0.375),res=450,dev="cairo_png")+
        facet_wrap(~banksyclus,nrow=2)+
        scale_color_manual(values = arcpal)+
        guides(color="none")+
        ggtitle(mscriptsampleids[sample_id==s,manuscript_id])+
        theme(axis.text.x=element_blank(),
              axis.text.y=element_blank(),
              axis.title.y=element_blank(),
              axis.title.x=element_blank(),
              axis.ticks = element_blank(),
              legend.background = element_blank(),
              legend.box=element_blank(),
              panel.grid = element_blank(),
              #legend.margin=margin(0,0,0,-0.15,unit="in"),
              #plot.margin = margin(0.01,0.02,0,0,unit="in"),
              legend.text = element_markdown(margin=margin(0,0,0,-0.075,"in")),
              strip.background = element_blank(),
              strip.text = element_markdown(size=10),
              )
    
    return(samplts)
})

### arrangeGrob() on the spotplot list --> constituting the "first column" of the parent; leg.tmp is an additional thing being plotted in the second column of the parent. width of the plot grid is set to 32, width of the second column (legend only) set to 2.
pdf("manuscript_plots/Supplemental/SF-ARC_cluster_facets_allsamples.pdf",height=40,width=34)
    do.call("grid.arrange",list(arrangeGrob(grobs=allplts,ncol=3),leg.tmp,ncol=2,widths=c(32,2)))
dev.off()
```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] ggh4x_0.2.8                    Polychrome_1.5.1              
 [3] viridis_0.6.5                  viridisLite_0.4.2             
 [5] ggrastr_1.0.2                  SpatialFeatureExperiment_1.6.1
 [7] spatialLIBD_1.16.2             SpatialExperiment_1.14.0      
 [9] SingleCellExperiment_1.26.0    SummarizedExperiment_1.34.0   
[11] Biobase_2.64.0                 GenomicRanges_1.56.1          
[13] GenomeInfoDb_1.40.1            IRanges_2.38.1                
[15] S4Vectors_0.42.1               BiocGenerics_0.50.0           
[17] MatrixGenerics_1.16.0          matrixStats_1.3.0             
[19] Cairo_1.6-2                    ggtext_0.1.2                  
[21] BiocParallel_1.38.0            parallelly_1.38.0             
[23] colorout_1.3-0.2               gridExtra_2.3                 
[25] ggplot2_3.5.1                  data.table_1.15.4             
[27] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] bitops_1.0-7              sf_1.0-16                 EBImage_4.46.0           
  [4] httr_1.4.7                RColorBrewer_1.1-3        doParallel_1.0.17        
  [7] tools_4.4.1               utf8_1.2.4                R6_2.5.1                 
 [10] DT_0.33                   HDF5Array_1.32.0          lazyeval_0.2.2           
 [13] rhdf5filters_1.16.0       withr_3.0.0               sp_2.1-4                 
 [16] cli_3.6.3                 labeling_0.4.3            sass_0.4.9               
 [19] proxy_0.4-27              commonmark_1.9.1          Rsamtools_2.20.0         
 [22] R.utils_2.12.3            scater_1.32.0             sessioninfo_1.2.2        
 [25] attempt_0.3.1             maps_3.4.2                limma_3.60.3             
 [28] rstudioapi_0.16.0         RSQLite_2.3.7             generics_0.1.3           
 [31] BiocIO_1.14.0             spdep_1.3-5               dplyr_1.1.4              
 [34] wesanderson_0.3.7         Matrix_1.7-0              ggbeeswarm_0.7.2         
 [37] fansi_1.0.6               abind_1.4-5               terra_1.7-78             
 [40] R.methodsS3_1.8.2         lifecycle_1.0.4           scatterplot3d_0.3-44     
 [43] yaml_2.3.9                edgeR_4.2.0               rhdf5_2.48.0             
 [46] SparseArray_1.4.8         BiocFileCache_2.12.0      paletteer_1.6.0          
 [49] grid_4.4.1                blob_1.2.4                promises_1.3.0           
 [52] dqrng_0.4.1               ExperimentHub_2.12.0      crayon_1.5.3             
 [55] lattice_0.22-6            beachmat_2.20.0           cowplot_1.1.3            
 [58] KEGGREST_1.44.1           magick_2.8.3              zeallot_0.1.0            
 [61] pillar_1.9.0              knitr_1.48                rjson_0.2.21             
 [64] boot_1.3-30               codetools_0.2-20          wk_0.9.2                 
 [67] glue_1.7.0                vctrs_0.6.5               png_0.1-8                
 [70] spam_2.10-0               gtable_0.3.5              rematch2_2.1.2           
 [73] cachem_1.1.0              xfun_0.45                 S4Arrays_1.4.1           
 [76] mime_0.12                 DropletUtils_1.24.0       sfheaders_0.4.4          
 [79] iterators_1.0.14          units_0.8-5               fields_16.2              
 [82] statmod_1.5.0             bit64_4.0.5               filelock_1.0.3           
 [85] rprojroot_2.0.4           bslib_0.7.0               irlba_2.3.5.1            
 [88] vipor_0.4.7               KernSmooth_2.23-24        colorspace_2.1-0         
 [91] spData_2.3.1              DBI_1.2.3                 tidyselect_1.2.1         
 [94] bit_4.0.5                 compiler_4.4.1            curl_5.2.1               
 [97] BiocNeighbors_1.22.0      xml2_1.3.6                DelayedArray_0.30.1      
[100] plotly_4.10.4             rtracklayer_1.64.0        scales_1.3.0             
[103] classInt_0.4-10           rappdirs_0.3.3            stringr_1.5.1            
[106] tiff_0.1-12               digest_0.6.36             fftwtools_0.9-11         
[109] rmarkdown_2.27            benchmarkmeData_1.0.4     XVector_0.44.0           
[112] htmltools_0.5.8.1         pkgconfig_2.0.3           jpeg_0.1-10              
[115] sparseMatrixStats_1.16.0  dbplyr_2.5.0              fastmap_1.2.0            
[118] htmlwidgets_1.6.4         UCSC.utils_1.0.0          shiny_1.8.1.1            
[121] DelayedMatrixStats_1.26.0 farver_2.1.2              jquerylib_0.1.4          
[124] jsonlite_1.8.8            config_0.3.2              R.oo_1.26.0              
[127] BiocSingular_1.20.0       RCurl_1.98-1.14           magrittr_2.0.3           
[130] scuttle_1.14.0            GenomeInfoDbData_1.2.12   s2_1.1.6                 
[133] dotCall64_1.1-1           Rhdf5lib_1.26.0           munsell_0.5.1            
[136] Rcpp_1.0.13               stringi_1.8.4             zlibbioc_1.50.0          
[139] AnnotationHub_3.12.0      parallel_4.4.1            ggrepel_0.9.5            
[142] deldir_2.0-4              Biostrings_2.72.1         gridtext_0.1.5           
[145] locfit_1.5-9.10           markdown_1.13             ScaledMatrix_1.12.0      
[148] BiocVersion_3.19.1        XML_3.99-0.17             evaluate_0.24.0          
[151] golem_0.4.1               BiocManager_1.30.23       foreach_1.5.2            
[154] httpuv_1.6.15             tidyr_1.3.1               purrr_1.0.2              
[157] benchmarkme_1.0.8         rsvd_1.0.5                xtable_1.8-4             
[160] restfulr_0.0.15           e1071_1.7-14              later_1.3.2              
[163] class_7.3-22              tibble_3.2.1              memoise_2.0.1            
[166] beeswarm_0.4.0            AnnotationDbi_1.66.0      GenomicAlignments_1.40.0 
[169] shinyWidgets_0.8.6        here_1.0.1               
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   Cairo                    * 1.6-2     2023-11-28 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   commonmark                 1.9.1     2024-01-30 [1] CRAN (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggh4x                    * 0.2.8     2024-01-23 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   ggtext                   * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext                   0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   markdown                   1.13      2024-06-04 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome               * 1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scatterplot3d              0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   stringi                    1.8.4     2024-05-06 [1] CRAN (R 4.4.0)
   stringr                    1.5.1     2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                  * 0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite              * 0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   wesanderson                0.3.7     2023-10-31 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xml2                       1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────