---
title: "SF-Xenium VMH ARC cluster domain and intra-domain-cluster markers"
output: html_document
date: "2024-09-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(ComplexHeatmap)
library(ggrastr) # render plots with bajillions of points as raster so they can be manipulated in eg Illustrator--Fig 2B
library(Cairo) # pdf device for ^ to properly save; also used for complexheatmap raster functions
library(circlize) #colorRamp2
library(viridis) # magma
library(Polychrome)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_text(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load xenium data, banksy clusters, one-v-all markers, intra-VMH marker analysis, intra-ARC marker analysis 
```{r}
hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

### 
globalmk <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/03a-M0l0kg6_reproducibleClustersOnly_spLIBD_markers.RDS")

# intraVMH and intra ARC markers (performed using 1-vs-all analysis but only among VMH or ARC clusters constrained to domain boundaries)
vintramk <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04c-VMHtypes-inVMHdom_1vall_markers.txt")

aintramk <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04a-ARCtypes-inARCdom_1vall_markers.txt")
```

apply cluster labells to xenium data and subset to oligos, vmh, arc for these plots; get domain labels and also apply to xenium data
```{r}
bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"clus")

# fuller annots for VMH, ARC clusters
fineanno <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04e_ARC_and_VMH_cluster_identities_based_on_cellgroup_1vall.txt")

# vague annots - remove DISCARD clusters but do NOT use the marker analysis results from this table -- this was analysis of all 41 clusters
bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

bkanno <- unique(bkanno[,.(clus,bjm_annot)])
bkanno <- merge.data.table(bkanno,fineanno,by="bjm_annot",all.x=T)
bkanno[!(is.na(subclus_annot)),bjm_annot:=subclus_annot]

bkcl <- merge.data.table(bkcl,bkanno,by="clus",all.x=T)
bkcl <- bkcl[!(bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T))]

## we're only plotting oligodendrocyte, VMH, and ARC clusters here so subset to those and remove other cell types from the SPE
bkcl <- bkcl[bjm_annot %in% grep(bjm_annot,pattern="Oligo|VMH|ARC",value=T)]

hypx <- hypx[,bkcl$rn]

bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]

hypx$bkcl <- bkcl$bjm_annot


### domain labels for these cells
hypdom <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt") 
hypdom <- hypdom[rn %in% bkcl$rn]

hypdom <- DataFrame(hypdom,row.names=hypdom$rn)[colnames(hypx),]
hypx$domain <- hypdom$dualVMHARC4
```

global marker heatmap: top 7 genes by LFC per cluster x 13 clusters = 91 genes (max)
```{r}
clusts <- unique(bkcl$bjm_annot)
globalmk <- as.data.table(globalmk[["enrichment"]])
globalmk[,ensembl:=NULL]
globalmk <- melt.data.table(globalmk,id.vars="gene")
globalmk[,stat:=gsub(variable,pattern="^(.*)_X.*$",replacement="\\1")]
globalmk[,clus:=gsub(variable,pattern="^.*_(X.*)$",replacement="\\1")]
globalmk[,variable:=NULL]
globalmk <- merge.data.table(globalmk,unique(as.data.table(bkcl)[,.(clus,bjm_annot)]),by="clus")
globalmk[,clus:=NULL]

globalmkfdr <- globalmk[stat=="fdr"]
globalmkfdr[,stat:=NULL]
setnames(globalmkfdr,"value","fdr")

globalmklfc <- globalmk[stat=="logFC"]
globalmklfc[,stat:=NULL]
setnames(globalmklfc,"value","logFC")

globalmk2 <- merge.data.table(globalmkfdr,globalmklfc,by=c("gene","bjm_annot"))

gns <- globalmk2[fdr<0.05,.SD[order(-logFC)][1:7],by="bjm_annot"][,gene]
gns <- unique(gns)
```

get scaled expression data for global heatmap
```{r}

## define function to z-scale expression
# z scale rows, from tony https://github.com/LieberInstitute/spatial_DG_lifespan/blob/9419eb91453cda1df494d93dc91156d819042c66/code/Pseudobulk/top_infant_DE_heatmap_enrichment.R#L106

scale_rows <- function(x) {
    m <- apply(x, 1, mean, na.rm = T)
    s <- apply(x, 1, sd, na.rm = T)
    return((x - m) / s)
}


# order as factor so we can fetch the spots per cluster in a presentable order (ARCS, VMHs, WMs)
arcs <- grep(unique(hypx$bkcl),pattern="ARC",value=T)
vmhs <- grep(unique(hypx$bkcl),pattern="VMH",value=T)
wms <- grep(unique(hypx$bkcl),pattern="Oligo",value=T)
wms <- wms[!(wms %in% c(arcs,vmhs))]

hypx$bkcl <- factor(hypx$bkcl,levels=c(arcs,vmhs,wms))

    
# make a subsetted spe with reordering of the columns to be in the order of the domains we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
i <- 1
for (i in c(1:length(levels(hypx$bkcl)))) {
    curlev <- levels(hypx$bkcl)[i]
    if (i == 1) {
        cnames <- colnames(hypx[, hypx$bkcl == curlev])
    } else {
        cnames <- c(cnames, colnames(hypx[, hypx$bkcl == curlev]))
    }
}
rm(i, curlev)

### extract log counts per marker, z-normalize, hang onto both
pltdat <- hypx[gns, cnames]
pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))
```


make the plot. use the same ARC and VMH cluster palettes from other xenium plots, then get 4 separate colors using polychrome for the oligos
SFigparts-Xenium_VMHARC_global_domainwide_and_intra_markers
```{r}
nclus <- length(unique(hypx$bkcl))

vmhpal <- c("#8c63cf", "#d3d3d3", "#6A0D53", "#AD260D")
names(vmhpal) <- c("VMH_SLC17A6-CRHR2","VMH_VMH-Glia-Mixed","VMH_FEZF1-LAMP5-NRGN","VMH_HCRT-SLC6A3-TAC1")

arcpal <- c("#5dd959","#296656","#09979B","#A997AB","#D000B0")
names(arcpal) <-  c("ARC_AGRP","ARC_POMC-Oligo-Mixed","ARC_GAL-TRH-GHRH","ARC_Excit-SLC17A7-Glial-Mixed","ARC_TAC3-AR-ESR1")

fullpal <- Polychrome::createPalette(N=nclus,seedcolors = c(vmhpal,arcpal),range=c(25,75))

names(fullpal) <- c(names(vmhpal),names(arcpal),unique(bkcl$bjm_annot)[!(unique(bkcl$bjm_annot) %in% c(names(vmhpal),names(arcpal)))])


htmap <- ComplexHeatmap::Heatmap(
    pltdat2,
    top_annotation = HeatmapAnnotation(
        XeniumCluster = pltdat$bkcl,
        gap = 0,
        col = list(XeniumCluster=fullpal),
        gp = gpar(alpha = 1,cex=0.35)
    ),
    show_column_names = FALSE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = T,
    col = colorRamp2(
        breaks = c(-2, 0, 2),
        colors = c("blue", "white", "orange")
    ),
    cluster_columns = F, # handled by ordering the collumns of the spe above
    heatmap_legend_param = list(
        title = "Z scored cell log counts",
        direction = "horizontal",
        title_position = "topcenter"
    ),
    row_names_gp = gpar(fontsize = 6),
    raster_device = "CairoTIFF",
    raster_quality = 10
)

### save plot
CairoPDF("manuscript_plots/Supplemental/SFigparts-Xenium_VMHARC_global_domainwide_and_intra_markers/XenVMHARCWM_globalmks.pdf", height = 7, width = 9)
draw(htmap,
    heatmap_legend_side = "bottom",
    annotation_legend_side = "right",
    legend_grouping = "original"
)
dev.off()

rm(pltdat,pltdat2,gns,cnames,clusts,nclus,wms)
```


intra-domain markers: 8*9 clus = 72 max
we need the intermediate-depth annotations  again to relabel these ones
```{r}
bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

bkanno <- merge.data.table(bkanno,fineanno,by="bjm_annot",all.x=T)

vintramk <- merge.data.table(vintramk,unique(bkanno[,.(bjm_annot,subclus_annot)]),by.x="clust",by.y="bjm_annot")

vgns <- vintramk[fdr<0.05,.SD[order(-logFC)][1:8],by="subclus_annot"][,gene]

aintramk <- merge.data.table(aintramk,unique(bkanno[,.(bjm_annot,subclus_annot)]),by.x="clust",by.y="bjm_annot")

agns <- aintramk[fdr<0.05,.SD[order(-logFC)][1:8],by="subclus_annot"][,gene]

gns <- unique(c(vgns,agns))
```


get scaled expression data of intra-VMH or intra-ARC domain cells for intra markers heatmap
```{r}

## define function to z-scale expression
# z scale rows, from tony https://github.com/LieberInstitute/spatial_DG_lifespan/blob/9419eb91453cda1df494d93dc91156d819042c66/code/Pseudobulk/top_infant_DE_heatmap_enrichment.R#L106

scale_rows <- function(x) {
    m <- apply(x, 1, mean, na.rm = T)
    s <- apply(x, 1, sd, na.rm = T)
    return((x - m) / s)
}


# order as factor so we can fetch the spots per cluster in a presentable order (ARCS, VMHs, WMs)
arcs <- grep(unique(hypx$bkcl),pattern="ARC",value=T)
vmhs <- grep(unique(hypx$bkcl),pattern="VMH",value=T)

hypxintra <- hypx[,hypx$bkcl %in% c(arcs,vmhs)&hypx$domain %in% c("VMH","ARC")]
hypxintra$bkcl <- factor(hypxintra$bkcl,levels=c(arcs,vmhs))

    
# make a subsetted spe with reordering of the columns to be in the order of the domains we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
i <- 1
for (i in c(1:length(c(arcs,vmhs)))) {
    curlev <- c(arcs,vmhs)[i]
    if (i == 1) {
        cnames <- colnames(hypxintra[, hypxintra$bkcl == curlev])
    } else {
        cnames <- c(cnames, colnames(hypxintra[, hypxintra$bkcl == curlev]))
    }
}
rm(i, curlev)

### extract log counts per marker, z-normalize, hang onto both
pltdat <- hypxintra[gns, cnames]
pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))
```


make the plot. use the same ARC and VMH cluster palettes from other xenium plots, then get 4 separate colors using polychrome for the oligos
SFigparts-Xenium_VMHARC_global_domainwide_and_intra_markers
```{r}

vmhpal <- c("#8c63cf", "#d3d3d3", "#6A0D53", "#AD260D")
names(vmhpal) <- c("VMH_SLC17A6-CRHR2","VMH_VMH-Glia-Mixed","VMH_FEZF1-LAMP5-NRGN","VMH_HCRT-SLC6A3-TAC1")

arcpal <- c("#5dd959","#296656","#09979B","#A997AB","#D000B0")
names(arcpal) <-  c("ARC_AGRP","ARC_POMC-Oligo-Mixed","ARC_GAL-TRH-GHRH","ARC_Excit-SLC17A7-Glial-Mixed","ARC_TAC3-AR-ESR1")

fullpal <- c(arcpal,vmhpal)

htmap <- ComplexHeatmap::Heatmap(
    pltdat2,
    top_annotation = HeatmapAnnotation(
        XeniumCluster = pltdat$bkcl,
        gap = 0,
        col = list(XeniumCluster=fullpal),
        gp = gpar(alpha = 1,cex=0.35)
    ),
    show_column_names = FALSE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = T,
    col = colorRamp2(
        breaks = c(-2, 0, 2),
        colors = c("blue", "white", "orange")
    ),
    cluster_columns = F, # handled by ordering the collumns of the spe above
    heatmap_legend_param = list(
        title = "Z scored cell log counts\nacross ARC and VMH clusters",
        direction = "horizontal",
        title_position = "topcenter"
    ),
    row_names_gp = gpar(fontsize = 6),
    raster_device = "CairoTIFF",
    raster_quality = 10
)

### save plot
CairoPDF("manuscript_plots/Supplemental/SFigparts-Xenium_VMHARC_global_domainwide_and_intra_markers/XenVMHARC_intraClusmks.pdf", height = 8, width = 9)
draw(htmap,
    heatmap_legend_side = "bottom",
    annotation_legend_side = "right",
    legend_grouping = "original"
)
dev.off()

rm(pltdat,pltdat2,gns,cnames,hypxintra,agns,vgns,htmap)
gc(full=T)
```

## load domain level markers and get top 25 ea
```{r}
dommk <- readRDS("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03d-Allcellsof-ARCVMH-in-singeclusters-otherARCVMHcellsdropped_1vallReprodClus-markers.RDS")

dommk <- as.data.table(dommk[["enrichment"]])
keepn <- c("gene",grep(names(dommk),pattern="VMH|ARC",value=T))
dommk <- dommk[,..keepn]

dommk <- melt.data.table(dommk,id.vars="gene")
dommk[,stat:=gsub(variable,pattern="^(.*)_...*$",replacement="\\1")]
dommk[,clus:=gsub(variable,pattern="^.*_(...)$",replacement="\\1")]

vdf <- dommk[clus=="VMH"&stat=="fdr"&value<0.05]
adf <- dommk[clus=="ARC"&stat=="fdr"&value<0.05]
setnames(vdf,"value","fdr")
vdf[,stat:=NULL]
vdf[,variable:=NULL]
setnames(adf,"value","fdr")
adf[,stat:=NULL]
adf[,variable:=NULL]

vdl <- dommk[clus=="VMH"&stat=="logFC"&value>0]
adl <- dommk[clus=="ARC"&stat=="logFC"&value>0]
setnames(vdl,"value","logFC")
vdl[,stat:=NULL]
vdl[,variable:=NULL]
setnames(adl,"value","logFC")
adl[,stat:=NULL]
adl[,variable:=NULL]

vg <- merge.data.table(vdl,vdf,by="gene")
ag <- merge.data.table(adl,adf,by="gene")

gns <- unique(c(vg[order(-logFC)][1:25]$gene,ag[order(-logFC)][1:25]$gene))

rm(vg,ag,vdl,adl,vdf,adf,dommk,keepn)
```

for the expression values here, we need to reload the SPE, subset to all cells of all types within the VMH/ARC domains, and then z-normalize those for consistency with how the markers were obtained. (we also need to get rid of discardable clusters from the intra-domain sets first)
```{r}
hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

doms <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt")
doms <- doms[dualVMHARC4 %in% c("VMH","ARC")]

bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"clus")

bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
bkanno <- unique(bkanno[,.(clus,bjm_annot)])

bkanno <- bkanno[bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T,invert=T)]

bkcl <- merge.data.table(bkcl,bkanno,by="clus")

bkcl <- merge.data.table(bkcl,doms,by="rn")
stopifnot(length(unique(bkcl$dualVMHARC4))==2)

hypx <- hypx[,bkcl$rn]

bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
hypx$dom <- bkcl$dualVMHARC4
```

now get the domain markers' expression values
```{r}
# order as factor so we can fetch the spots per cluster in a presentable order (ARCS, VMHs, WMs)

# make a subsetted spe with reordering of the columns to be in the order of the domains we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
i <- 1
for (i in c(1:length(unique(hypx$dom)))) {
    curlev <- unique(hypx$dom)[i]
    if (i == 1) {
        cnames <- colnames(hypx[, hypx$dom == curlev])
    } else {
        cnames <- c(cnames, colnames(hypx[, hypx$dom == curlev]))
    }
}
rm(i, curlev)

### extract log counts per marker, z-normalize, hang onto both
pltdat <- hypx[gns, cnames]
pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))

```

# make the heatmap
```{r}
fullpal <- c("#5dd959","#8c63cf")
names(fullpal) <- c("ARC","VMH")

htmap <- ComplexHeatmap::Heatmap(
    pltdat2,
    top_annotation = HeatmapAnnotation(
        Domain = pltdat$dom,
        gap = 0,
        col = list(Domain=fullpal),
        gp = gpar(alpha = 1)
    ),
    show_column_names = FALSE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = T,
    col = colorRamp2(
        breaks = c(-2, 0, 2),
        colors = c("blue", "white", "orange")
    ),
    cluster_columns = F, # handled by ordering the collumns of the spe above
    heatmap_legend_param = list(
        title = "Z scored cell log counts\nacross all cells w/in ARC, VMH domains",
        direction = "horizontal",
        title_position = "topcenter"
    ),
    row_names_gp = gpar(fontsize = 6),
    raster_device = "CairoTIFF",
    raster_quality = 10
)

### save plot
CairoPDF("manuscript_plots/Supplemental/SFigparts-Xenium_VMHARC_global_domainwide_and_intra_markers/XenVMHARCWM_domainmks_vs_allcluscellsoutsidedoms.pdf", height = 6, width = 5)
draw(htmap,
    heatmap_legend_side = "bottom",
    annotation_legend_side = "right",
    legend_grouping = "original"
)
dev.off()

rm(pltdat,pltdat2,gns,cnames,hypxintra,agns,vgns,htmap)
gc(full=T)
```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] Polychrome_1.5.1               SpatialFeatureExperiment_1.6.1
 [3] BiocParallel_1.38.0            parallelly_1.38.0             
 [5] colorout_1.3-0.2               ggtext_0.1.2                  
 [7] viridis_0.6.5                  viridisLite_0.4.2             
 [9] circlize_0.4.16                Cairo_1.6-2                   
[11] ggrastr_1.0.2                  ComplexHeatmap_2.20.0         
[13] spatialLIBD_1.16.2             SpatialExperiment_1.14.0      
[15] SingleCellExperiment_1.26.0    SummarizedExperiment_1.34.0   
[17] Biobase_2.64.0                 GenomicRanges_1.56.1          
[19] GenomeInfoDb_1.40.1            IRanges_2.38.1                
[21] S4Vectors_0.42.1               BiocGenerics_0.50.0           
[23] MatrixGenerics_1.16.0          matrixStats_1.3.0             
[25] gridExtra_2.3                  ggplot2_3.5.1                 
[27] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] bitops_1.0-7              sf_1.0-16                 EBImage_4.46.0           
  [4] httr_1.4.7                RColorBrewer_1.1-3        doParallel_1.0.17        
  [7] tools_4.4.1               utf8_1.2.4                R6_2.5.1                 
 [10] DT_0.33                   HDF5Array_1.32.0          lazyeval_0.2.2           
 [13] rhdf5filters_1.16.0       GetoptLong_1.0.5          withr_3.0.0              
 [16] sp_2.1-4                  cli_3.6.3                 sass_0.4.9               
 [19] proxy_0.4-27              Rsamtools_2.20.0          R.utils_2.12.3           
 [22] scater_1.32.0             sessioninfo_1.2.2         attempt_0.3.1            
 [25] maps_3.4.2                limma_3.60.3              rstudioapi_0.16.0        
 [28] RSQLite_2.3.7             generics_0.1.3            shape_1.4.6.1            
 [31] BiocIO_1.14.0             spdep_1.3-5               dplyr_1.1.4              
 [34] Matrix_1.7-0              ggbeeswarm_0.7.2          fansi_1.0.6              
 [37] abind_1.4-5               terra_1.7-78              R.methodsS3_1.8.2        
 [40] lifecycle_1.0.4           scatterplot3d_0.3-44      yaml_2.3.9               
 [43] edgeR_4.2.0               rhdf5_2.48.0              SparseArray_1.4.8        
 [46] BiocFileCache_2.12.0      paletteer_1.6.0           blob_1.2.4               
 [49] promises_1.3.0            dqrng_0.4.1               ExperimentHub_2.12.0     
 [52] crayon_1.5.3              lattice_0.22-6            beachmat_2.20.0          
 [55] cowplot_1.1.3             KEGGREST_1.44.1           magick_2.8.3             
 [58] zeallot_0.1.0             pillar_1.9.0              knitr_1.48               
 [61] boot_1.3-30               rjson_0.2.21              codetools_0.2-20         
 [64] wk_0.9.2                  glue_1.7.0                vctrs_0.6.5              
 [67] png_0.1-8                 spam_2.10-0               gtable_0.3.5             
 [70] rematch2_2.1.2            cachem_1.1.0              xfun_0.45                
 [73] S4Arrays_1.4.1            mime_0.12                 DropletUtils_1.24.0      
 [76] sfheaders_0.4.4           iterators_1.0.14          units_0.8-5              
 [79] fields_16.2               statmod_1.5.0             bit64_4.0.5              
 [82] filelock_1.0.3            rprojroot_2.0.4           bslib_0.7.0              
 [85] irlba_2.3.5.1             vipor_0.4.7               KernSmooth_2.23-24       
 [88] colorspace_2.1-0          spData_2.3.1              DBI_1.2.3                
 [91] tidyselect_1.2.1          bit_4.0.5                 compiler_4.4.1           
 [94] curl_5.2.1                BiocNeighbors_1.22.0      xml2_1.3.6               
 [97] DelayedArray_0.30.1       plotly_4.10.4             rtracklayer_1.64.0       
[100] scales_1.3.0              classInt_0.4-10           rappdirs_0.3.3           
[103] tiff_0.1-12               digest_0.6.36             fftwtools_0.9-11         
[106] rmarkdown_2.27            benchmarkmeData_1.0.4     XVector_0.44.0           
[109] htmltools_0.5.8.1         pkgconfig_2.0.3           jpeg_0.1-10              
[112] sparseMatrixStats_1.16.0  dbplyr_2.5.0              fastmap_1.2.0            
[115] GlobalOptions_0.1.2       htmlwidgets_1.6.4         UCSC.utils_1.0.0         
[118] shiny_1.8.1.1             DelayedMatrixStats_1.26.0 jquerylib_0.1.4          
[121] jsonlite_1.8.8            config_0.3.2              R.oo_1.26.0              
[124] BiocSingular_1.20.0       RCurl_1.98-1.14           magrittr_2.0.3           
[127] scuttle_1.14.0            GenomeInfoDbData_1.2.12   s2_1.1.6                 
[130] dotCall64_1.1-1           Rhdf5lib_1.26.0           munsell_0.5.1            
[133] Rcpp_1.0.13               zlibbioc_1.50.0           AnnotationHub_3.12.0     
[136] parallel_4.4.1            ggrepel_0.9.5             deldir_2.0-4             
[139] Biostrings_2.72.1         gridtext_0.1.5            locfit_1.5-9.10          
[142] ScaledMatrix_1.12.0       BiocVersion_3.19.1        XML_3.99-0.17            
[145] evaluate_0.24.0           golem_0.4.1               BiocManager_1.30.23      
[148] foreach_1.5.2             httpuv_1.6.15             tidyr_1.3.1              
[151] purrr_1.0.2               benchmarkme_1.0.8         clue_0.3-65              
[154] rsvd_1.0.5                xtable_1.8-4              restfulr_0.0.15          
[157] e1071_1.7-14              later_1.3.2               class_7.3-22             
[160] tibble_3.2.1              memoise_2.0.1             beeswarm_0.4.0           
[163] AnnotationDbi_1.66.0      GenomicAlignments_1.40.0  cluster_2.1.6            
[166] shinyWidgets_0.8.6        here_1.0.1               
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-22
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   Cairo                    * 1.6-2     2023-11-28 [1] CRAN (R 4.4.0)
   circlize                 * 0.4.16    2024-02-20 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   clue                       0.3-65    2023-09-23 [1] CRAN (R 4.4.0)
   cluster                    2.1.6     2023-12-01 [1] CRAN (R 4.4.1)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   ComplexHeatmap           * 2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   GetoptLong                 1.0.5     2020-12-15 [1] CRAN (R 4.4.0)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   ggtext                   * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   GlobalOptions              0.1.2     2020-06-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext                   0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome               * 1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scatterplot3d              0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shape                      1.4.6.1   2024-02-23 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                  * 0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite              * 0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xml2                       1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────