---
title: "diseases_tfs_and_sexDE_tabulations"
output: html_document
date: "2024-10-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

library(parallel)

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```


```{r}
visde <- fread("spatial_HYP/processed-data/09-Sex DE/01c-voomLmFit_nnsvg10-HmnylmbNA-BS-15-VMHARCclpsd.txt")

xendea <- fread("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/02a-4ARCtypesmoothed_withinARC_celltypeSexDE.txt")

xendev <- fread("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/02b-4ARCtypesmoothed_withinVMH_celltypeSexDE.txt")

betterclustnames <- fread("manuscript_plot_code/xARCxVMH_cluster_detailedlabs_andplotformatnames.txt")

xendea <- merge.data.table(xendea,betterclustnames[,.(bjm_annot,dom.clus)],by.x="cluster_annotated",by.y="bjm_annot",all.x=T)
xendea[!is.na(dom.clus),cluster_annotated:=dom.clus]

xendev <- merge.data.table(xendev,betterclustnames[,.(bjm_annot,dom.clus)],by.x="cluster_annotated",by.y="bjm_annot",all.x=T)
xendev[!is.na(dom.clus),cluster_annotated:=dom.clus]

## disease gene sets for reference:
dissets <- fread("spatial_HYP/raw-data/Genesets from GWAS pubs, TWAS, and disgenet/nonredundant_union_set_genesymbols.txt")

## gsea results from marker, sex de enrichments in disease, tf targets
## WHERE IS THE EXPRESSION FILTERING? the info from the filtering of these on expression tho?
mkdisres <- readRDS("spatial_HYP/processed-data/11-GSEA/05-Common variant disease gene enrichment in markers.RDS")

sexdisres <- fread("local_supptab_assembly/sex-disgene-gsea.txt")

mktfres <- readRDS("spatial_HYP/processed-data/11-GSEA/08-Marker TF-target GSEA results from 02a filtered to TFs expressed in avg 5pct of cluster spots and at least 1pct of cluster spots in 5samps.RDS")
mktfres <- mktfres[c("k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_XARC","k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_XVMH")]
names(mktfres) <- c("vARC","vVMH")

sextfres <- readRDS("spatial_HYP/processed-data/11-GSEA/09-Sex DE TF-target GSEA results from 04a filtered to TFs expressed in at least one sex at avg 2.5pct of cluster spots and at least 1pct of cluster spots in 3samps.RDS")
sextfres <- sextfres[c("k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_ARC","k15_HARMONYlmbna_nnsvg10_60kiter_collapsedXVMH")]
names(sextfres) <- c("vARC","vVMH")
```


### apply the "expressed TF" cutoffs for all genes to get genes qualifying as expressed in vVMH, vARC, or that domain for a single sex
```{r}
hyp2 <- readRDS("spatial_HYP/processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

# set rownames of spe to gene symbol, since those are the identifiers we're stuck with for TF-targ sets
# make all upper case to be sure
dropg <- as.data.table(rowData(hyp2),keep.rownames=T)[,.(.N,gene_id),by=gene_name][N>1,gene_id]
hyp2 <- hyp2[!(rownames(hyp2)%in%dropg),]
rownames(hyp2) <- rowData(hyp2)$gene_name

bscl <- fread("spatial_HYP/processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt")
setnames(bscl,2,"cl")
bscl[,cl:=paste0("X",cl)]

bscl[cl %in% c("X4","X6"),cl:="vARC"]
bscl[cl %in% c("X12","X7"),cl:="vVMH"]

bscl <- DataFrame(bscl,row.names=bscl$rn)[colnames(hyp2),]
colLabels(hyp2) <- bscl$cl

doms <- c(rep("vARC",3),rep("vVMH",3))
sexes <- list(c("Male","Female"),"Male","Female",c("Male","Female"),"Male","Female")

setDTthreads(2,restore_after_fork = FALSE)
domxprgs <- mcmapply(S=sexes,D=doms,mc.cores=6,FUN=function(S,D){

    cursamps <- unique(colData(hyp2[,hyp2$Sex %in% S])[,"sample_id"])
    lct <- as.data.table(as.data.frame(as.matrix(logcounts(hyp2[,hyp2$label==D&hyp2$sample_id %in% cursamps]))),keep.rownames=T)
    setnames(lct,"rn","symb")
    lct <- melt(lct,id.vars="symb",value.name="logcounts")
    ### get sample ID from the unique spot identifiers
    lct[,samp:=gsub(variable,pattern="^.*_(.*_.*)$",replacement="\\1")]
    lct[,Nspot:=.N/length(unique(symb)),by="samp"]
            
    # this is a clunky approach, but the TOTAL %age of spots expressing the TF in question (ie logcounts>0) for that sample will fill in for any rows where logcounts are >0. we can then get a table of unique combinations of sample X TF x %age of spots
    lct[logcounts>0,pctxpr:=.N/Nspot,by=c("symb","samp")]
    # set NAs in pctxpr to 0
    lct[is.na(pctxpr),pctxpr:=0]
    # get each unique pair of tf-sample-pctxpr, where we fetch the max row value of pctxpr for a tf-sample pair (which will be >0 and = the % of spots expressing in any row where the gene was detected for that sample)
    lctfilt <- lct[,max(pctxpr),by=c("symb","samp")]
    # calculate mean % spots expressing dataset wide and number of samples with expression over 1%. each tf occurs once per sample, so take the mean across rows with a given TF value for samplewise means, or tabulate the number of TF occurrences with the pctxpr filter to get the number of passing rows
    setnames(lctfilt,"V1","pctxpr")
    lctfilt[,meanxpr:=mean(pctxpr),by="symb"]
    lctfilt[pctxpr>0.01,Nxpr:=.N,by="symb"]
    # na rows returned in Nxpr will be meaningless in this result, since we just will subset to max Nxpr analogous to what we did above, resulting in a table of X tfs with their mean + spot %age and number of samples over threshold. set na values to 0 to avoid errors relating to NAs here.
    lctfilt[is.na(Nxpr),Nxpr:=0]
    lctfilt2 <- lctfilt[,max(Nxpr),by="symb"]
    setnames(lctfilt2,"V1","Nxpr")
    ## add baack the mean expression to the number of threshold-passing samples by merging parts of the old data ttable to the above result
    lctfilt2 <- merge.data.table(lctfilt2,unique(lctfilt[,.(symb,meanxpr)],by="symb"))

    # FINALLY  get the genes meeting criteria (expressed in at least 2.5% of spots of cluster for at least 3 samples from the given sex, or 5% / 5 samples for the combined sex case) back and filter the result table to them
    
    if(length(S)==1){
        xprg <- lctfilt2[Nxpr>3&meanxpr>0.025,symb]
    }
    else{
        xprg <- lctfilt2[Nxpr>5&meanxpr>0.05,symb]
    }
    
    # return qualifying genes
    return(xprg)
})
sexes[c(1,4)] <- "both"
names(domxprgs) <- mapply(X=doms,Y=sexes,FUN=function(X,Y){paste0(X,"_",Y)})
```


supp tab of sex DE genes from Visium paired to GSEA disease sets(s); ibid for xenium: one long table of gene IDs (one for TFs, one for disease•gene pairs, including columns for whether vVMH, vARC markers andor vVMH/vARC sex DE was enriched for that disease) with expressed y/n, vVMH top 1k 1 vs. all, sex DE (encoding p < 0.05, fdr < 0.05, or neither) , vARC sexDE, ...xVMH CRHR2 top 20 marker y/n, xVMH CRHR2 sex DE y/n...
```{r}
gtab <- as.data.table(rownames(hyp2))
setnames(gtab,1,"gene_symbol")

gtab[,disorders:=apply(.SD,MARGIN = 1,FUN=function(x){
    dxes <- unique(dissets[symbol %in% x[1],dx])
    if(length(dxes)==0){
        return(NA)
    }
    else{
        return(paste0(dxes,collapse=", "))
}})]

gtab[gene_symbol %in% unique(c(mktfres$vVMH,mktfres$vARC$tf,sextfres$vVMH$tf,sextfres$vARC$tf)),tested_tf:="YES"]
gtab[is.na(tested_tf),tested_tf:="NO"]

gtab[gene_symbol %in% domxprgs[["vVMH_both"]],xpr_vVMH:="YES"]
gtab[gene_symbol %in% domxprgs[["vVMH_Male"]],xpr_vVMH_M:="YES"]
gtab[gene_symbol %in% domxprgs[["vVMH_Female"]],xpr_vVMH_F:="YES"]
gtab[gene_symbol %in% domxprgs[["vARC_both"]],xpr_vARC:="YES"]
gtab[gene_symbol %in% domxprgs[["vARC_Male"]],xpr_vARC_M:="YES"]
gtab[gene_symbol %in% domxprgs[["vARC_Female"]],xpr_vARC_F:="YES"]

gtab[is.na(xpr_vVMH),xpr_vVMH:="NO"]
gtab[is.na(xpr_vVMH_M),xpr_vVMH_M:="NO"]
gtab[is.na(xpr_vVMH_F),xpr_vVMH_F:="NO"]
gtab[is.na(xpr_vARC),xpr_vARC:="NO"]
gtab[is.na(xpr_vARC_M),xpr_vARC_M:="NO"]
gtab[is.na(xpr_vARC_F),xpr_vARC_F:="NO"]



gtab[,vARC_sexDE_fdrsig:=apply(.SD,MARGIN = 1,function(x){ifelse(visde[assay=="ARC"&adj.P.Val<0.05&gene_name %in% x,.N]>0,"YES","NO")}),.SDcols = "gene_symbol"]
gtab[,vVMH_sexDE_fdrsig:=apply(.SD,MARGIN = 1,function(x){ifelse(visde[assay=="VMH"&adj.P.Val<0.05&gene_name %in% x,.N]>0,"YES","NO")}),.SDcols = "gene_symbol"]

gtab[,vARC_sexDE_nomsig:=apply(.SD,MARGIN = 1,function(x){ifelse(visde[assay=="ARC"&P.Value<0.05&gene_name==x,.N]>0,"YES","NO")}),.SDcols = "gene_symbol"]
gtab[,vVMH_sexDE_nomsig:=apply(.SD,MARGIN = 1,function(x){ifelse(visde[assay=="VMH"&P.Value<0.05&gene_name==x,.N]>0,"YES","NO")}),.SDcols = "gene_symbol"]

gtab[,xARC_sexDE_FDRpt1clusts:=apply(.SD,MARGIN = 1,function(x){ifelse(xendea[FDR<0.1&gene_name==x,.N]>0,paste0(xendea[FDR<0.1&gene_name==x,cluster_annotated],collapse=", "),"NO")}),.SDcols = "gene_symbol"]

gtab[,xVMH_sexDE_FDRpt1clusts:=apply(.SD,MARGIN = 1,function(x){ifelse(xendev[FDR<0.1&gene_name==x,.N]>0,paste0(xendev[FDR<0.1&gene_name==x,cluster_annotated],collapse=", "),"NO")}),.SDcols = "gene_symbol"]
```

## lastly, add a column for whether or not the gene was assayed on Xenium
```{r}
# check if the DE tables cover all 366 genes otherwise load up the full panel list
stopifnot(length(unique(c(xendea$gene_name,xendev$gene_name)))==366)
xeng <- unique(c(xendea$gene_name,xendev$gene_name))

gtab[,xenium_gene:="No"]
gtab[gene_symbol %in% xeng,xenium_gene:="Yes"]

names(gtab)
gtab <- gtab[,c(1,16,2:15),with=F]
```


save
```{r}
fwrite(gtab,"manuscript_plots/Supplemental/STable-Genes_diseases_TFs_and_sexDE.txt",sep='\t',quote=F,row.names=F,col.names=T)


```