---
title: "06-Sex DE heatmaps for cell types within ARC or VMH domains, nuclear counts only"
output: html_document
date: "2024-07-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(ggrastr)
library(ComplexHeatmap)
library(colorRamp2)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### read in sex de results
```{r}
indomde <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/06-4celltypeARC-and-VMH-dualassignedDomains_celltypeSexDE_nuclearcounts.txt")

## get custom panel genes
panelgenes <- fread("raw-data/humanbrain_basepanel_genes.txt")
hypgenes <- unique(indomde[!(ID %in% panelgenes$Gene),ID])
rm(panelgenes)

## get all genes with signif DE or in custom panel and P < 0.01
pltgene <- indomde[FDR_inARCnuc<0.05|FDR_inVMHnuc<0.05|(ID %in% hypgenes & (P.Value_inARCnuc<0.01|P.Value_inVMHnuc<0.01)),ID]

## get interesting cell types
keeptypes <- grep(unique(indomde$bjm_annot),pattern="ARC|VMH|Tanycyte|Periventricular|Supraoptic|TRH",value=T)

## get all data for these genes and cell types
indomde.plt <- indomde[ID %in% pltgene & bjm_annot %in% keeptypes]
```

### ARC sex DE heatmap -- get a trio of logFC (plot values) and pvalue (asterisked cells) tables for this 
```{r}
indomde.plt.arc <- dcast(indomde.plt[,.(ID,bjm_annot,logFC_MvF_inARCnuc)],ID~bjm_annot,value.var="logFC_MvF_inARCnuc")

indomde.plt.arc.pvals <- dcast(indomde.plt[,.(ID,bjm_annot,P.Value_inARCnuc)],ID~bjm_annot,value.var="P.Value_inARCnuc")

indomde.plt.arc.fdrs <- dcast(indomde.plt[,.(ID,bjm_annot,FDR_inARCnuc)],ID~bjm_annot,value.var="FDR_inARCnuc")

names(indomde.plt.arc) <- gsub(names(indomde.plt.arc),pattern="_", replacement=" ")
names(indomde.plt.arc.pvals) <- gsub(names(indomde.plt.arc.pvals),pattern="_", replacement=" ")
names(indomde.plt.arc.fdrs) <- gsub(names(indomde.plt.arc.fdrs),pattern="_", replacement=" ")


## remove NAs from p value and fdr tables
indomde.plt.arc.pvals <- 
    indomde.plt.arc.pvals[,.(ID,
                             apply(.SD,MARGIN=c(1,2),FUN=function(x){
                                 if(is.na(x)){return(1)}
                                 else{return(x)}
                             })
                            ),
                          .SDcols=c(2:ncol(indomde.plt.arc.pvals))]

indomde.plt.arc.fdrs <- 
    indomde.plt.arc.fdrs[,.(ID,
                             apply(.SD,MARGIN=c(1,2),FUN=function(x){
                                 if(is.na(x)){return(1)}
                                 else{return(x)}
                             })
                            ),
                          .SDcols=c(2:ncol(indomde.plt.arc.fdrs))]


indomde.plt.arc <- as.data.frame(indomde.plt.arc)
indomde.plt.arc.pvals <- as.data.frame(indomde.plt.arc.pvals)
indomde.plt.arc.fdrs <- as.data.frame(indomde.plt.arc.fdrs)

rownames(indomde.plt.arc) <- indomde.plt.arc$ID
rownames(indomde.plt.arc.pvals) <- indomde.plt.arc.pvals$ID
rownames(indomde.plt.arc.fdrs) <- indomde.plt.arc.fdrs$ID

indomde.plt.arc$ID <- NULL
indomde.plt.arc.pvals$ID <- NULL
indomde.plt.arc.fdrs$ID <- NULL

indomde.plt.arc.pvals <- as.matrix(indomde.plt.arc.pvals)
indomde.plt.arc.fdrs <- as.matrix(indomde.plt.arc.fdrs)

htmap <- ComplexHeatmap::Heatmap(
    as.matrix(indomde.plt.arc),
    show_column_names = TRUE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = T,
    col = colorRamp2(
        breaks = c(-2, 0, 2),
        colors = c("blue", "white", "orange")
    ),
    cluster_columns = F,
    heatmap_legend_param = list(
        title = "log2FC (M:F)",
        direction = "horizontal",
        title_position = "topcenter"
    ),
    row_names_gp = gpar(fontsize = 8),
    column_names_gp=grid::gpar(fontsize=6),
    raster_device = "CairoTIFF",
    raster_quality = 10,
    cell_fun = function(j, i, x, y, w, h, fill){
        if(indomde.plt.arc.fdrs[i, j] < 0.1) {
            grid.text('**', x, y,vjust=.75)}
        else if (indomde.plt.arc.pvals[i,j]<0.05){
            grid.text('*', x, y,vjust=.75)
        }
    }
)

png("plots/08_VMH-ARC cell type sex DE within domains/06-ARC domain cell type sex DE heatmap FDR.1 p05 nuclear counts.png",width=6,height=7,units = "in",res=144)
draw(htmap,heatmap_legend_side="bottom")
dev.off()

```

### VMH sex DE heatmap
### there's so much missing data for VMH with nuclear counts that we need to subset this frame down to get rid of genes not analyzed in any of these cell types / cell types not analyzed for any of these genes
```{r}
indomde.plt.vmh <- dcast(indomde.plt[,.(ID,bjm_annot,logFC_MvF_inVMHnuc)],ID~bjm_annot,value.var="logFC_MvF_inVMHnuc")

## drop cell types with no genes analyzed
dropcols <- colSums(is.na(indomde.plt.vmh))==nrow(indomde.plt.vmh)
indomde.plt.vmh <- indomde.plt.vmh[,!dropcols,with=FALSE]
## drop genes not analyzed in any cell type (ncol()-1 beacuse a column is the gene id here)
droprows <- rowSums(is.na(indomde.plt.vmh))==ncol(indomde.plt.vmh)-1
indomde.plt.vmh <- indomde.plt.vmh[!droprows]

## pvalue table, subsetted to same genes and cell types
indomde.plt.vmh.pvals <- dcast(indomde.plt[,.(ID,bjm_annot,P.Value_inVMHnuc)],ID~bjm_annot,value.var="P.Value_inVMHnuc")
indomde.plt.vmh.pvals <- indomde.plt.vmh.pvals[ID %in% indomde.plt.vmh$ID,names(indomde.plt.vmh),with=FALSE]

## fdr table, subsetted to same genes and cell types
indomde.plt.vmh.fdrs <- dcast(indomde.plt[,.(ID,bjm_annot,FDR_inVMHnuc)],ID~bjm_annot,value.var="FDR_inVMHnuc")
indomde.plt.vmh.fdrs<- indomde.plt.vmh.fdrs[ID %in% indomde.plt.vmh$ID,names(indomde.plt.vmh),with=FALSE]

names(indomde.plt.vmh) <- gsub(names(indomde.plt.vmh),pattern="_", replacement=" ")
names(indomde.plt.vmh.pvals) <- gsub(names(indomde.plt.vmh.pvals),pattern="_", replacement=" ")
names(indomde.plt.vmh.fdrs) <- gsub(names(indomde.plt.vmh.fdrs),pattern="_", replacement=" ")


## remove NAs from p value and fdr tables
indomde.plt.vmh.pvals <- 
    indomde.plt.vmh.pvals[,.(ID,
                             apply(.SD,MARGIN=c(1,2),FUN=function(x){
                                 if(is.na(x)){return(1)}
                                 else{return(x)}
                             })
                            ),
                          .SDcols=c(2:ncol(indomde.plt.vmh.pvals))]

indomde.plt.vmh.fdrs <- 
    indomde.plt.vmh.fdrs[,.(ID,
                             apply(.SD,MARGIN=c(1,2),FUN=function(x){
                                 if(is.na(x)){return(1)}
                                 else{return(x)}
                             })
                            ),
                          .SDcols=c(2:ncol(indomde.plt.vmh.fdrs))]


indomde.plt.vmh <- as.data.frame(indomde.plt.vmh)
indomde.plt.vmh.pvals <- as.data.frame(indomde.plt.vmh.pvals)
indomde.plt.vmh.fdrs <- as.data.frame(indomde.plt.vmh.fdrs)

rownames(indomde.plt.vmh) <- indomde.plt.vmh$ID
rownames(indomde.plt.vmh.pvals) <- indomde.plt.vmh.pvals$ID
rownames(indomde.plt.vmh.fdrs) <- indomde.plt.vmh.fdrs$ID

indomde.plt.vmh$ID <- NULL
indomde.plt.vmh.pvals$ID <- NULL
indomde.plt.vmh.fdrs$ID <- NULL

indomde.plt.vmh.pvals <- as.matrix(indomde.plt.vmh.pvals)
indomde.plt.vmh.fdrs <- as.matrix(indomde.plt.vmh.fdrs)

htmap <- ComplexHeatmap::Heatmap(
    as.matrix(indomde.plt.vmh),
    show_column_names = TRUE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = T,
    col = colorRamp2(
        breaks = c(-2, 0, 2),
        colors = c("blue", "white", "orange")
    ),
    cluster_columns = F,
    cluster_rows=T,
    heatmap_legend_param = list(
        title = "log2FC (M:F)",
        direction = "horizontal",
        title_position = "topcenter"
    ),
    row_names_gp = gpar(fontsize = 8),
    column_names_gp=grid::gpar(fontsize=6),
    raster_device = "CairoTIFF",
    raster_quality = 10,
    cell_fun = function(j, i, x, y, w, h, fill){
        if(indomde.plt.vmh.fdrs[i, j] < 0.1) {
            grid.text('**', x, y,vjust=.75)}
        else if (indomde.plt.vmh.pvals[i,j]<0.05){
            grid.text('*', x, y,vjust=.75)
        }
    }
)

png("plots/08_VMH-ARC cell type sex DE within domains/06-VMH domain cell type sex DE heatmap FDR.1 p05 nuclear counts.png",width=6,height=7,units = "in",res=144)
draw(htmap,heatmap_legend_side="bottom")
dev.off()
```


### reprod info
```{r}
sessionInfo()

sessioninfo::session_info()

```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     datasets 
[8] methods   base     

other attached packages:
 [1] colorRamp2_0.1.0      ComplexHeatmap_2.20.0 ggrastr_1.0.2        
 [4] BiocParallel_1.38.0   parallelly_1.37.1     colorout_1.3-0.2     
 [7] gridExtra_2.3         ggplot2_3.5.1         Biostrings_2.72.1    
[10] GenomeInfoDb_1.40.1   XVector_0.44.0        IRanges_2.38.0       
[13] S4Vectors_0.42.0      BiocGenerics_0.50.0   data.table_1.15.4    
[16] rlang_1.1.4          

loaded via a namespace (and not attached):
 [1] shape_1.4.6.1           gtable_0.3.5            beeswarm_0.4.0         
 [4] circlize_0.4.16         rjson_0.2.21            xfun_0.45              
 [7] GlobalOptions_0.1.2     paletteer_1.6.0         vctrs_0.6.5            
[10] tools_4.4.1             generics_0.1.3          parallel_4.4.1         
[13] tibble_3.2.1            fansi_1.0.6             cluster_2.1.6          
[16] pkgconfig_2.0.3         RColorBrewer_1.1-3      lifecycle_1.0.4        
[19] GenomeInfoDbData_1.2.12 compiler_4.4.1          munsell_0.5.1          
[22] codetools_0.2-20        clue_0.3-65             vipor_0.4.7            
[25] htmltools_0.5.8.1       yaml_2.3.8              pillar_1.9.0           
[28] crayon_1.5.3            sessioninfo_1.2.2       magick_2.8.3           
[31] viridis_0.6.5           iterators_1.0.14        foreach_1.5.2          
[34] tidyselect_1.2.1        digest_0.6.36           rematch2_2.1.2         
[37] dplyr_1.1.4             rprojroot_2.0.4         fastmap_1.2.0          
[40] here_1.0.1              colorspace_2.1-0        cli_3.6.3              
[43] magrittr_2.0.3          utf8_1.2.4              withr_3.0.0            
[46] scales_1.3.0            UCSC.utils_1.0.0        ggbeeswarm_0.7.2       
[49] rmarkdown_2.27          httr_1.4.7              matrixStats_1.3.0      
[52] png_0.1-8               GetoptLong_1.0.5        evaluate_0.24.0        
[55] knitr_1.47              doParallel_1.0.17       viridisLite_0.4.2      
[58] Rcpp_1.0.12             glue_1.7.0              rstudioapi_0.16.0      
[61] jsonlite_1.8.8          R6_2.5.1                zlibbioc_1.50.0        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-06
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package          * version date (UTC) lib source
   beeswarm           0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   BiocGenerics     * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel     * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings       * 2.72.1  2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   circlize           0.4.16  2024-02-20 [1] CRAN (R 4.4.0)
 P cli                3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   clue               0.3-65  2023-09-23 [1] CRAN (R 4.4.0)
   cluster            2.1.6   2023-12-01 [1] CRAN (R 4.4.1)
   codetools          0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout         * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorRamp2       * 0.1.0   2022-12-21 [1] CRAN (R 4.4.0)
   colorspace         2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   ComplexHeatmap   * 2.20.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   crayon             1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table       * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest             0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   doParallel         1.0.17  2022-02-07 [1] CRAN (R 4.4.0)
   dplyr              1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate           0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi              1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   fastmap            1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   foreach            1.5.2   2022-02-02 [1] CRAN (R 4.4.0)
   generics           0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb     * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData   1.2.12  2024-05-01 [1] Bioconductor
   GetoptLong         1.0.5   2020-12-15 [1] CRAN (R 4.4.0)
   ggbeeswarm         0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2          * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr          * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   GlobalOptions      0.1.2   2020-06-10 [1] CRAN (R 4.4.0)
   glue               1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra        * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable             0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here               1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools          0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr               1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges          * 2.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   iterators          1.0.14  2022-02-05 [1] CRAN (R 4.4.0)
   jsonlite           1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr              1.47    2024-05-29 [1] CRAN (R 4.4.0)
   lifecycle          1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick             2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr           2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   matrixStats        1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell            0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   paletteer          1.6.0   2024-01-21 [1] CRAN (R 4.4.0)
   parallelly       * 1.37.1  2024-02-29 [1] CRAN (R 4.4.0)
   pillar             1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig          2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   png                0.1-8   2022-11-29 [1] CRAN (R 4.4.0)
   R6                 2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   RColorBrewer       1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp               1.0.12  2024-01-09 [1] CRAN (R 4.4.0)
   rematch2           2.1.2   2020-05-01 [1] CRAN (R 4.4.0)
   rjson              0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang            * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown          2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot          2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi         0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors        * 0.42.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales             1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo        1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   shape              1.4.6.1 2024-02-23 [1] CRAN (R 4.4.0)
   tibble             3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect         1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils         1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8               1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs              0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor              0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   viridis            0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite        0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
   withr              3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun               0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector          * 0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml               2.3.8   2023-12-11 [1] CRAN (R 4.4.0)
   zlibbioc           1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
