---
title: "03-Sex DE msigdb-GSEA"
author: "Bernard Mulvey"
date: "2024-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(BiocParallel)
library(parallel)
library(parallelly)
library(fgsea)
library(msigdbr)

require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
options(bphost="localhost")

```

load sex DE stats, subsetted to the four clusterings of interest (nnsvg10, harmony lambda null, q=15,20,31 and q=15 collapsed VMH/ARC)

```{r}
sexde <- readRDS("processed-data/09-Sex DE/01-voomLmFit_svg10-svg20-hvg20_Hmnydflt-mnn30-HmnylmbNA_BS-15-20-31-15VMHARCclpsd.RDS")

## subset to the nnsvg10 ones
sexde <- sexde[grep(names(sexde),pattern="HARMONYlmbna_nnsvg10")]


# check:
# names(sexde)
# [1] "k15_HARMONYlmbna_nnsvg10_60kiter"          
# [2] "k20_HARMONYlmbna_nnsvg10_60kiter"          
# [3] "k31_HARMONYlmbna_nnsvg10_60kiter"          
# [4] "k15_HARMONYlmbna_nnsvg10_60kiter_clpsdARC_clpsdVMH"

```


### fgsea sets retrieval
```{r}
allsigs <- msigdbr("Homo sapiens",category="C1")

# tack on the others, make d.t.
msigs <- paste0("C",c(2,3,5,6,8))
for (msig in msigs){
    allsigs <- rbind(allsigs,msigdbr("Homo sapiens",category=msig))
}
rm(msig,msigs)
allsigs <- as.data.table(allsigs)

# remove legacy sets
allsigs <- allsigs[gs_subcat %in% grep(gs_subcat,pattern="legacy|LEGACY|Legacy",value=T,invert=T)]

# make a list object of identifiers per  gene set
# we can map these back to the msigdb table later for reference purposes by their list name (gs_name)...
sets <- split(allsigs$ensembl_gene,allsigs$gs_name)
```


# SEX DE gsea
```{r}
### drop mito and sex chr genes
sexde <- lapply(sexde,FUN=function(x){
    x[!(chromosome_name %in% grep(chromosome_name,pattern="MT|X|Y",value=T))]})

### get named vectors of marker t-stats for each cluster in each result set (15+20+31+13)=79 (if no clusters dropped out of sex de analysis)
sexdelist <- list()
i<-1
for (i in c(1:length(sexde))){
    sdoms <- unique(sexde[[i]]$assay)
    setnames(sexde[[i]],"t","tstat")
    j<-1
    for (j in c(1:length(sdoms))){
        tmp <- copy(sexde[[i]][assay==sdoms[j]])[,.(gene_id,tstat)]
        setorderv(tmp,"tstat",-1)
        ## get the stat column as a vector and give it the gene ids as names for fgsea
        ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
        tmp <- as.data.frame(tmp)
        v <- tmp$tstat
        names(v) <- tmp$gene_id
        sexdelist[[length(sexdelist)+1]] <- v
        names(sexdelist)[length(sexdelist)] <- paste0(names(sexde)[i],"_",sdoms[j])
        rm(v,tmp)
    }
    rm(j,sdoms)
}
rm(i)
```

### loop through each domain (each run is very fast with 10 cpus in biocparallel, so just serially run 10cpu GSEA for each cluster)
```{r}
sbp <- MulticoreParam(10)
register(sbp)
sexderes <- mapply(X=sexdelist,Y=names(sexdelist),SIMPLIFY=FALSE,FUN=function(X,Y){
    z<-as.data.table(fgseaMultilevel(pathways=sets,stats = X,minSize = 15,maxSize = 500,eps=0.0,BPPARAM = sbp,nPermSimple = 50000))
    z[,spdom:=Y]
    return(z)
})
names(sexderes) <- names(sexdelist)

saveRDS(sexderes,"processed-data/11-GSEA/03a-Sex DE msigdb_1,2,3,5,6,8 GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")

# clear some memory for the next step which is more intensive
rm(allsigs)
gc(full=T)
### collapse pathways-this part is WAY WAY slower than the GSEA itself

sexderes2 <- mcmapply(X=sexderes,Y=sexdelist,SIMPLIFY = FALSE, mc.cores=7,FUN=function(X,Y){
    tmpres <- fgsea::collapsePathways(fgseaRes = X,stats=Y,pathways = sets)
    return(tmpres)
})
names(sexderes2) <- names(sexderes)
saveRDS(sexderes2,"processed-data/11-GSEA/03b-Sex DE mSigDB_1,2,3,5,6,8 fGSEA-collapsePathways output.RDS")


rm(sexdelist,sexde)
gc(full=T)

### subset to main enrichments
sexderes3 <- mapply(X=sexderes,Y=sexderes2,SIMPLIFY = FALSE,FUN=function(X,Y){
    X[pathway %in% Y[[1]]]
})
saveRDS(sexderes3,"processed-data/11-GSEA/03c-Sex DE mSigDB_1,2,3,5,6,8 GSEA clpspath mainPathways-only.RDS")
# clean up
rm(sexderes,sexderes2)
gc(full=T)
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] default_1.0.0       colorout_1.3-0.2    msigdbr_7.5.1      
 [4] fgsea_1.30.0        parallelly_1.37.1   BiocParallel_1.38.0
 [7] Biostrings_2.72.1   GenomeInfoDb_1.40.1 XVector_0.44.0     
[10] IRanges_2.38.1      S4Vectors_0.42.1    BiocGenerics_0.50.0
[13] data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] utf8_1.2.4              generics_0.1.3          lattice_0.22-6         
 [4] digest_0.6.36           magrittr_2.0.3          evaluate_0.24.0        
 [7] grid_4.4.1              fastmap_1.2.0           rprojroot_2.0.4        
[10] jsonlite_1.8.8          Matrix_1.7-0            httr_1.4.7             
[13] fansi_1.0.6             UCSC.utils_1.0.0        scales_1.3.0           
[16] codetools_0.2-20        cli_3.6.3               crayon_1.5.3           
[19] cowplot_1.1.3           munsell_0.5.1           withr_3.0.0            
[22] yaml_2.3.9              tools_4.4.1             dplyr_1.1.4            
[25] colorspace_2.1-0        ggplot2_3.5.1           fastmatch_1.1-4        
[28] GenomeInfoDbData_1.2.12 here_1.0.1              babelgene_22.9         
[31] vctrs_0.6.5             R6_2.5.1                lifecycle_1.0.4        
[34] zlibbioc_1.50.0         pkgconfig_2.0.3         pillar_1.9.0           
[37] gtable_0.3.5            glue_1.7.0              Rcpp_1.0.13            
[40] xfun_0.45               tibble_3.2.1            tidyselect_1.2.1       
[43] rstudioapi_0.16.0       knitr_1.48              htmltools_0.5.8.1      
[46] rmarkdown_2.27          compiler_4.4.1         
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-26
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package          * version date (UTC) lib source
   babelgene          22.9    2022-09-29 [1] CRAN (R 4.4.0)
   BiocGenerics     * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel     * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings       * 2.72.1  2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools          0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout         * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace         2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   cowplot            1.1.3   2024-01-22 [1] CRAN (R 4.4.0)
   crayon             1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table       * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   default          * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
   digest             0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr              1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate           0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi              1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   fastmap            1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   fastmatch          1.1-4   2023-08-18 [1] CRAN (R 4.4.0)
   fgsea            * 1.30.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   generics           0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb     * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData   1.2.12  2024-05-01 [1] Bioconductor
   ggplot2            3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue               1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gtable             0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here               1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools          0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr               1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges          * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jsonlite           1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr              1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lattice            0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle          1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr           2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix             1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   msigdbr          * 7.5.1   2022-03-30 [1] CRAN (R 4.4.0)
   munsell            0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly       * 1.37.1  2024-02-29 [1] CRAN (R 4.4.1)
   pillar             1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig          2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6                 2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp               1.0.13  2024-07-17 [1] CRAN (R 4.4.1)
 P rlang            * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown          2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot          2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi         0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors        * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales             1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo        1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble             3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect         1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils         1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8               1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs              0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr              3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun               0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector          * 0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml               2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc           1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
