---
title: "04-Sex DE TF-Target GSEA"
author: "Bernard Mulvey"
date: "2024-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(BiocParallel)
library(parallel)
library(parallelly)
library(fgsea)

require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
options(bphost="localhost")

```

load dreamlet random effect sex DE stats, which were only run for the four clusterings of interest to begin with

```{r}
sexde <- readRDS("processed-data/09-Sex DE/01-voomLmFit_svg10-svg20-hvg20_Hmnydflt-mnn30-HmnylmbNA_BS-15-20-31-15VMHARCclpsd.RDS")

sexde <- sexde[grep(names(sexde),pattern="HARMONYlmbna_nnsvg10")]

# check:
# names(sexde)
# [1] "k15_HARMONYlmbna_nnsvg10"          
# [2] "k20_HARMONYlmbna_nnsvg10"          
# [3] "k31_HARMONYlmbna_nnsvg10"          
# [4] "k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH"

```

### Enrichr TF-target databases gene sets retrieval
```{r}
tftargsets <- readRDS("data/11-GSEA/00a-enrichrTFsets_forfgseainput.RDS")
```

# sex de gsea for TF-target sets
# note that we need SYMBOLS ("gene" column), not ens IDs in this case. the TF-target lists are only available as symbols
```{r}
### drop mito genes and chr X, Y genes
sexde <- lapply(sexde,FUN=function(x){x[!(chromosome_name %in% grep(chromosome_name,pattern="MT|X|Y",value=T))]})

### get named vectors of marker t-stats for each cluster in each result set (max 15+20+31+13)=79; probably less for clusters that dropped due to poor representation
sexdelist <- list()
i<-1
for (i in c(1:length(sexde))){
    sdoms <- unique(sexde[[i]]$assay)
    setnames(sexde[[i]],"t","tstat")
    j<-1
    for (j in c(1:length(sdoms))){
        tmp <- copy(sexde[[i]][assay==sdoms[j]])[,.(gene_name,tstat)]
        setorderv(tmp,"tstat",-1)
        ## get the stat column as a vector and give it the gene ids as names for fgsea
        ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
        tmp <- as.data.frame(tmp)
        v <- tmp$tstat
        names(v) <- tmp$gene_name
        sexdelist[[length(sexdelist)+1]] <- v
        names(sexdelist)[length(sexdelist)] <- paste0(names(sexde)[i],"_",sdoms[j])
        rm(v,tmp)
    }
    rm(j,sdoms)
}
rm(i)
```

### run fgsea
### looping through each domain (each run is very fast with 10 cpus in biocparallel, so just serially run 10cpu GSEA for each cluster)

```{r}
sbp <- MulticoreParam(10)
register(sbp)
sexderes <- mapply(X=sexdelist,Y=names(sexdelist),SIMPLIFY=FALSE,FUN=function(X,Y){
    z<-as.data.table(fgseaMultilevel(pathways=tftargsets,stats = X,minSize = 15,maxSize = 500,eps=0.0,BPPARAM = sbp,nPermSimple = 50000))
    z[,spdom:=Y]
    return(z)
})
names(sexderes) <- names(sexdelist)

saveRDS(sexderes,"processed-data/11-GSEA/04a-Sex DE TF-target GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")

### collapse pathways-this part is WAY WAY slower than the GSEA itself

sexderes2 <- mcmapply(X=sexderes,Y=sexdelist,SIMPLIFY = FALSE, mc.cores=7,FUN=function(X,Y){
    tmpres <- fgsea::collapsePathways(fgseaRes = X,stats=Y,pathways = tftargsets)
    return(tmpres)
})
names(sexderes2) <- names(sexderes)

saveRDS(sexderes2,"processed-data/11-GSEA/04b-Sex DE TF-target fGSEA-collapsePathways output.RDS")

rm(sexdelist,sexde)
gc(full=T)

### subset to main enrichments
sexderes3 <- mapply(X=sexderes,Y=sexderes2,SIMPLIFY = FALSE,FUN=function(X,Y){
    X[pathway %in% Y[[1]]]
})
saveRDS(sexderes3,"processed-data/11-GSEA/04c-Sex DE TF-target GSEA clpspath mainPathways-only.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
