---
title: "08-Filter marker TF-targ results to expressed TFs"
author: "Bernard Mulvey"
date: "2024-01-29"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(SpatialExperiment)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

Reload marker - TFtarg GSEA results before having run collapsePathways (which could have e.g., removed a result for a TF that's relevant in the data but retained one for a TF with similar targets despite the TF NOT being expressed in the data) and the SPE (which we need for filtering on expression)
```{r}
## redundancied results
mkreslist <- readRDS("processed-data/11-GSEA/02a-Marker TF-target GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")

hyp2 <- readRDS("processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

# set rownames of spe to gene symbol, since those are the identifiers we're stuck with for TF-targ sets
# make all upper case to be sure; drop genes with multiple entries
dropg <- as.data.table(rowData(hyp2),keep.rownames=T)[,.(.N,gene_id),by=gene_name][N>1,gene_id]
hyp2 <- hyp2[!(rownames(hyp2)%in%dropg),]
rownames(hyp2) <- rowData(hyp2)$gene_name
rownames(hyp2) <- toupper(rownames(hyp2))

#### separate out the TF symbol from the full set name so we can filter the SPE accordingly
mkreslist <- lapply(mkreslist,FUN=function(x){
    
    y <- x[!(pathway %in% grep(pathway,pattern="Rummagene",value=T))]
    y[,tf:=apply(.SD,MARGIN = 1,FUN=function(h){strsplit(h,"_",1)[[1]][1]}),.SDcols="pathway"]
    ### rummagene TFs are named differently:
    z <- x[pathway %in% grep(pathway,pattern="Rummagene",value=T)]
    # reverse the output of a strsplit of a strsplit o.0 to get the string immediately preceding "_Rummagene_.."
    z[,tf:=apply(.SD,MARGIN=1,FUN=function(p){
        rev.default(strsplit(
            strsplit(p,"_Rummagene_transcription_factors",-1)[[1]][1],
            "_")[[1]]
        )[1]
    }),.SDcols="pathway"]
    #z[,tf:=gsub(pathway,pattern=".*-.*_(.*)_Rummagene_transcription_factors)",replacement="\\1")]
    fix <- rbind(z,y)
    # make all upper case
    fix[,tf:=toupper(tf)]
    return(fix)
})

keepgenes <- unique(unlist(lapply(mkreslist,FUN=function(x){x$tf})))
keepgenes <- keepgenes[keepgenes %in% rownames(hyp2)]
hyp2 <- hyp2[keepgenes,]

rm(keepgenes)
gc(full=T)

### finally, rearrange the results into a list of lists -- each parent list a clustering run, each child list one of the clusters
names(mkreslist)

mkreslist <- list(
    mkreslist[grep(names(mkreslist),pattern="k15_HARMONYlmbna_nnsvg10_60kiter_X",value=T)],
    mkreslist[grep(names(mkreslist),pattern="k20_HARMONYlmbna_nnsvg10_60kiter_X",value=T)],
    mkreslist[grep(names(mkreslist),pattern="k31_HARMONYlmbna_nnsvg10_60kiter_X",value=T)],
    mkreslist[grep(names(mkreslist),pattern="k15_HARMONYlmbna_nnsvg10_60kiter_collapsed_X",value=T)]
)

names(mkreslist) <- c("k15_HARMONYlmbna_nnsvg10_60kiter","k20_HARMONYlmbna_nnsvg10_60kiter","k31_HARMONYlmbna_nnsvg10_60kiter","k15_HARMONYlmbna_nnsvg10_60kiter_collapsed")
```

### get the clustering sets for the four Bayesspace runs we're reporting so we can extract and filter to TFs expressed in the cluster at >1% of spots in ≥ 5 samples and with a mean expression in at least 5% of spots cohort wide (for each clustering paradigm separately)
```{r}
bscl <- list.files("processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out",recursive=F,pattern=".txt",full.names = T)
bscls <- lapply(bscl,fread)


names(bscls) <- gsub(bscl,pattern="^.*31_out/(.*)\\.txt$",replacement="\\1")

bscls <- bscls[names(bscls) %in% grep(names(bscls),pattern="HARMONYlmbna_nnsvg10",value=T)]

## (re)create the collapsed set with the same list name as its listed in the per-cluster GSEA results list 
bscls[[4]] <- copy(bscls[["BSpace_k15_HARMONYlmbna_nnsvg10"]])
bscls[[4]] <- as.data.table(bscls[[4]])
setnames(bscls[[4]],2,"label")
bscls[[4]][,label:=paste0("X",label)]

# VMH and ARC were labeled XVMH / XARC in this analysis; others were labeled XX# (clunky thing about manually appending this cluster assignment set in previous datasets)
bscls[[4]][label %in% c("X7","X12"),label:="XVMH"]
bscls[[4]][label %in% c("X4","X6"),label:="XARC"]
bscls[[4]][!(label %in% c("XARC","XVMH")),label:=paste0("X",label)]

### format for spe compatability (as data frame with matched order of rownames)
bscls <- c(
    lapply(bscls[1:3],FUN=function(x){
        setnames(x,2,"label")
        x[,label:=paste0("X",label)]
        x <- DataFrame(x)
        rownames(x) <- x$rn
        x <- x[colnames(hyp2),]
        return(x)
    }),
    bscls[4]
)
## formatting for the collapsed entry
bscls[[4]] <- DataFrame(bscls[[4]])
rownames(bscls[[4]]) <- bscls[[4]]$rn
bscls[[4]] <- bscls[[4]][colnames(hyp2),]
names(bscls)[4] <- "k15_HARMONYlmbna_nnsvg10_60kiter_collapsed"

# change bsls names to the values used to index results in mkres
names(bscls) <- c("k15_HARMONYlmbna_nnsvg10_60kiter","k20_HARMONYlmbna_nnsvg10_60kiter","k31_HARMONYlmbna_nnsvg10_60kiter","k15_HARMONYlmbna_nnsvg10_60kiter_collapsed")
```

### get TFs qualifying as expressed in each cluster

### since the data table parlance is a little hard to read here, the nonbeliever may want to set R <- mkreslist[[1]], C <- bscls[[1]], N <- as.list(names(bscls)[1]) and step through the calls inside the apply loop below to view the table and spot check the results as they please. it does work though--and quickly! 
```{r}
sbp <- MulticoreParam(4)
register(sbp)
mkreslist.filt <- bpmapply(R=mkreslist,C=bscls,N=as.list(names(bscls)),BPPARAM=sbp,FUN=function(R,C,N){
    
    colLabels(hyp2) <- C$label
    
    filtres <- list()
    i<-1
    for (i in c(1:length(R))){
        curclus <- gsub(names(R)[i],pattern=paste0(N,"_(.*)$"),replacement="\\1")
        tflogcts <- as.data.table(as.data.frame(as.matrix(logcounts(hyp2[,hyp2$label==curclus]))),keep.rownames=T)
        setnames(tflogcts,"rn","tf")
        tflogcts <- melt(tflogcts,id.vars="tf",value.name="logcounts")
        ### get sample ID from the unique spot identifiers
        tflogcts[,samp:=gsub(variable,pattern="^.*_(.*_.*)$",replacement="\\1")]
        
        ## get the number of spots: since every spot will have SOME value for each TF, the number of spots is the number of sample-specific expression values (.N, number of row entries) DIVIDED BY the number of tfs themselves, which will be a constant across samples.
        tflogcts[,Nspot:=.N/length(unique(tf)),by="samp"]
        
        # this is a clunky approach, but the TOTAL %age of spots expressing the TF in question (ie logcounts>0) for that sample will fill in for any rows where logcounts are >0. we can then get a table of unique combinations of sample X TF x %age of spots
        tflogcts[logcounts>0,pctxpr:=.N/Nspot,by=c("tf","samp")]
        # set NAs in pctxpr to 0
        tflogcts[is.na(pctxpr),pctxpr:=0]
        # get each unique pair of tf-sample-pctxpr, where we fetch the max row value of pctxpr for a tf-sample pair (which will be all non-zero (previously NA) cells with the % of spots expressing in any row where the gene was detected for that sample). surprisingly this returns a table without repeats, which would be expected for a given tf/sample/pctxpr triplet where the values are the same for all spots with expression >0. but it makes things that much cleaner
        tffilt <- tflogcts[,max(pctxpr),by=c("tf","samp")]
        # calculate mean % spots expressing dataset wide and number of samples with expression over 1%. each tf occurs once per sample, so take the mean across rows with a given TF value for samplewise means, or tabulate the number of TF occurrences with the pctxpr filter to get the number of passing rows. this again produces repetivie rows (the total number of samples expressing is entered in that TF row for every sample, whether that particular sample was an expresser or not.)
        setnames(tffilt,"V1","pctxpr")
        tffilt[,meanxpr:=mean(pctxpr),by="tf"]
        tffilt[pctxpr>.01,Nxpr:=.N,by="tf"]
        # na rows returned in Nxpr will be meaningless in this result, since we just will subset to max Nxpr analogous to what we did above, resulting in a table of X tfs with their mean + spot %age and number of samples over threshold. set na values to 0 to avoid errors relating to NAs here.
        tffilt[is.na(Nxpr),Nxpr:=0]
        tffilt2 <- tffilt[,max(Nxpr),by="tf"]
        setnames(tffilt2,"V1","Nxpr")
        ## add baack the mean expression to the number of threshold-passing samples by merging parts of the old data ttable to the above result
        tffilt2 <- merge.data.table(tffilt2,unique(tffilt[,.(tf,meanxpr)],by="tf"))
        
        # FINALLY  get the TFs meeting criteria (majority of samples expressing in >1% of spots and >5% of spots in the clustering expressing on average across the dataset) and filter the result table to them
        
        keeptfs <- tffilt2[Nxpr>4&meanxpr>0.05,tf]
        # filter the current cluster's results to these TFs
        filtres[[i]] <- R[[i]][tf %in% keeptfs]
        names(filtres)[i] <- names(R)[i]
    }
    return(filtres)
})

```

### return to one giant list and save
```{r}
megalist.filt <- list()
i<-1
for (i in c(1:4)){
    j<-1
    for(j in c(1:length(mkreslist.filt[[i]]))){
        megalist.filt[[length(megalist.filt)+1]] <- mkreslist.filt[[i]][[j]]
        names(megalist.filt)[length(megalist.filt)] <- names(mkreslist.filt[[i]])[j]
    }
    rm(j)
}
rm(i)

saveRDS(megalist.filt,"processed-data/11-GSEA/08-Marker TF-target GSEA results from 02a filtered to TFs expressed in avg 5pct of cluster spots and at least 1pct of cluster spots in 5samps.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] default_1.0.0               SpatialExperiment_1.14.0   
 [3] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
 [5] Biobase_2.64.0              GenomicRanges_1.56.1       
 [7] MatrixGenerics_1.16.0       matrixStats_1.3.0          
 [9] colorout_1.3-0.2            parallelly_1.38.0          
[11] BiocParallel_1.38.0         hypeR_2.2.0                
[13] Biostrings_2.72.1           GenomeInfoDb_1.40.1        
[15] XVector_0.44.0              IRanges_2.38.1             
[17] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[19] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1        viridisLite_0.4.2       dplyr_1.1.4            
 [4] farver_2.1.2            fastmap_1.2.0           tweenr_2.0.3           
 [7] promises_1.3.0          digest_0.6.36           mime_0.12              
[10] lifecycle_1.0.4         magrittr_2.0.3          compiler_4.4.1         
[13] tools_4.4.1             igraph_2.0.3            utf8_1.2.4             
[16] yaml_2.3.9              knitr_1.48              S4Arrays_1.4.1         
[19] htmlwidgets_1.6.4       DelayedArray_0.30.1     here_1.0.1             
[22] plyr_1.8.9              xml2_1.3.6              abind_1.4-5            
[25] babelgene_22.9          withr_3.0.0             purrr_1.0.2            
[28] grid_4.4.1              polyclip_1.10-6         fansi_1.0.6            
[31] xtable_1.8-4            colorspace_2.1-0        ggplot2_3.5.1          
[34] scales_1.3.0            MASS_7.3-61             cli_3.6.3              
[37] rmarkdown_2.27          crayon_1.5.3            generics_0.1.3         
[40] rstudioapi_0.16.0       rjson_0.2.21            httr_1.4.7             
[43] reshape2_1.4.4          visNetwork_2.1.2        ggforce_0.4.2          
[46] stringr_1.5.1           zlibbioc_1.50.0         parallel_4.4.1         
[49] msigdbr_7.5.1           vctrs_0.6.5             Matrix_1.7-0           
[52] jsonlite_1.8.8          magick_2.8.3            systemfonts_1.1.0      
[55] glue_1.7.0              codetools_0.2-20        stringi_1.8.4          
[58] gtable_0.3.5            later_1.3.2             UCSC.utils_1.0.0       
[61] munsell_0.5.1           tibble_3.2.1            pillar_1.9.0           
[64] htmltools_0.5.8.1       reactable_0.4.4         GenomeInfoDbData_1.2.12
[67] R6_2.5.1                rprojroot_2.0.4         lattice_0.22-6         
[70] evaluate_0.24.0         kableExtra_1.4.0        shiny_1.8.1.1          
[73] openxlsx_4.2.5.2        httpuv_1.6.15           Rcpp_1.0.13            
[76] zip_2.3.1               SparseArray_1.4.8       svglite_2.1.3          
[79] xfun_0.45               pkgconfig_2.0.3        
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                  1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   babelgene              22.9    2022-09-29 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings           * 2.72.1  2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   default              * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi                  1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggforce                0.4.2   2024-02-19 [1] CRAN (R 4.4.0)
   ggplot2                3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue                   1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4   2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15  2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   hypeR                * 2.2.0   2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   igraph                 2.0.3   2024-03-13 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   kableExtra             1.4.0   2024-01-24 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   later                  1.3.2   2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   MASS                   7.3-61  2024-06-13 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   mime                   0.12    2021-09-28 [1] CRAN (R 4.4.0)
   msigdbr                7.5.1   2022-03-30 [1] CRAN (R 4.4.0)
   munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   openxlsx               4.2.5.2 2023-02-06 [1] CRAN (R 4.4.0)
   parallelly           * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   plyr                   1.8.9   2023-10-02 [1] CRAN (R 4.4.0)
   polyclip               1.10-6  2023-09-27 [1] CRAN (R 4.4.0)
   promises               1.3.0   2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2   2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   reactable              0.4.4   2023-03-12 [1] CRAN (R 4.4.0)
   reshape2               1.4.4   2020-04-09 [1] CRAN (R 4.4.0)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Arrays               1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1 2024-04-02 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray            1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   stringi                1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
   stringr                1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   svglite                2.1.3   2023-12-08 [1] CRAN (R 4.4.0)
   systemfonts            1.1.0   2024-05-15 [1] CRAN (R 4.4.0)
   tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   tweenr                 2.0.3   2024-02-26 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
   visNetwork             2.1.2   2022-09-29 [1] CRAN (R 4.4.0)
   withr                  3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   xml2                   1.3.6   2023-12-04 [1] CRAN (R 4.4.0)
   xtable                 1.8-4   2019-04-21 [1] CRAN (R 4.4.0)
   XVector              * 0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zip                    2.3.1   2024-01-27 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
