---
title: "02-Harmony and MNN"
author: "Bernard Mulvey"
date: "2023-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
require(colorout)
library(batchelor)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### spe prep
```{r}
hyp2 <- readRDS("data/03-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

```

### load feature sets
```{r}
hyp2.hvg10 <- readRDS("data/04-feature_selection/01-hyp2n9_HVG10pctile_sample-blocked.RDS")
hyp2.hvg20 <- readRDS("data/04-feature_selection/01-hyp2n9_HVG20pctile_sample-blocked.RDS")
nnsvg10 <- readRDS("data/04-feature_selection/01-hyp2n9_nnsvg10pctile.RDS")
nnsvg20 <- readRDS("data/04-feature_selection/01-hyp2n9_nnsvg20pctile.RDS")

feats <- list(hyp2.hvg10,hyp2.hvg20,nnsvg10,nnsvg20)
names(feats) <- c("hvg10","hvg20","nnsvg10","nnsvg20")
```

### Dimensionality reductions with each feature set: standard PCA, standard UMAP; PCA/UMAP with harmony (defaults); PCA/UMAP with harmony (lambda=NULL); PCA/UMAP with MNN

### for harmony, install and run github development version (1.2), which is in beta to address overfitting by setting the initial lambda to NULL such that harmony estimates the input value. 
```{r}
# do not build vignettes; dependency conflicts with Seurat5 as of 111523
# devtools::install_github("immunogenomics/harmony", build_vignettes=FALSE)
library(harmony)

# set up BiocParallel for mnn
sbp <- MulticoreParam(8)

i<-1
for (i in c(1:length(feats))){
    # standard PCAs
    set.seed(42)
    hyp2 <- scater::runPCA(hyp2,subset_row=feats[[i]],name=paste0("PCA_",names(feats)[i]))

    # standard UMAPs (which require PCA first)
    set.seed(42)
    hyp2 <- scater::runUMAP(hyp2,dimred=paste0("PCA_",names(feats)[i]),name=paste0("UMAP_",names(feats)[i]))
    
    # harmony runs -- we need to temporarily rename the PCA for the feature set to "PCA" for runharmony to accept it
    reducedDimNames(hyp2)[which(reducedDimNames(hyp2)==paste0("PCA_",names(feats)[i]))] <- "PCA"
    
    # vanilla harmony (lambda a pre-specified default value of 1)
    set.seed(42)
    hyp2 <- RunHarmony(hyp2,group.by.vars = "sample_id",reduction.save=paste0("HARMONYdflt_",names(feats)[i]))
    
    # data-based initialization of lambda (development ver of Harmony; requires explicit lambda=NULL)
    hyp2 <- RunHarmony(hyp2,group.by.vars="sample_id",reduction.save=paste0("HARMONYlmbna_",names(feats)[i]))
    
    # restore the feature-set-specific name for the uncorrected PCA
    reducedDimNames(hyp2)[which(reducedDimNames(hyp2)=="PCA")] <- paste0("PCA_",names(feats)[i])
    
    # MNN by donor with k=30
    set.seed(42)
    mnnout <- reducedMNN(reducedDim(hyp2,paste0("PCA_",names(feats)[i])),batch = hyp2$brnum,k=30,BPPARAM=sbp)
    # extract reduced dimension table and append to SPE
    reducedDim(hyp2,paste0("mnn30_",names(feats)[i])) <- mnnout$corrected
    # remove intermediate object
    rm(mnnout)
    
    # UMAP of uncorrected PCA
    set.seed(42)
    hyp2 <- scater::runUMAP(hyp2,dimred=paste0("PCA_",names(feats)[i]),name=paste0("UMAP_",names(feats)[i]))
    
    # UMAP of vanilla harmony
    set.seed(42)
    hyp2 <- scater::runUMAP(hyp2,dimred=paste0("HARMONYdflt_",names(feats)[i]),name=paste0("HARMONYdflt_",names(feats)[i],"_UMAP"))
    
    # UMAP of data-based initialization of lambda
    set.seed(42)
    hyp2 <- scater::runUMAP(hyp2,dimred=paste0("HARMONYlmbna_",names(feats)[i]),name=paste0("HARMONYlmbna_",names(feats)[i],"_UMAP"))
    
    # UMAP of MNN
    set.seed(42)
    hyp2 <- scater::runUMAP(hyp2,dimred=paste0("mnn30_",names(feats)[i]),name=paste0("mnn30_",names(feats)[i],"_UMAP"))
    
}

```

### save RDS with all the dimreds ####
```{r}
saveRDS(hyp2,"data/04-feature_selection/02-hypfiltered_4featureset-pca-umap-harmonydefault-harmonylambdanull-mnn30.RDS")
```

### reproducibility information
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.1 (2023-06-16)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] batchelor_1.18.0            harmony_1.2.0              
 [3] Rcpp_1.0.11                 BiocParallel_1.36.0        
 [5] parallelly_1.36.0           default_1.0.0              
 [7] colorout_1.3-0              spatialLIBD_1.13.4         
 [9] SpatialExperiment_1.12.0    SingleCellExperiment_1.24.0
[11] SummarizedExperiment_1.32.0 Biobase_2.62.0             
[13] GenomicRanges_1.54.1        MatrixGenerics_1.14.0      
[15] matrixStats_1.1.0           gridExtra_2.3              
[17] ggplot2_3.4.4               Biostrings_2.70.1          
[19] GenomeInfoDb_1.38.1         XVector_0.42.0             
[21] IRanges_2.36.0              S4Vectors_0.40.1           
[23] BiocGenerics_0.48.1         data.table_1.14.8          
[25] rlang_1.1.2                

loaded via a namespace (and not attached):
  [1] RcppAnnoy_0.0.21              later_1.3.1                  
  [3] BiocIO_1.12.0                 bitops_1.0-7                 
  [5] filelock_1.0.2                fields_15.2                  
  [7] tibble_3.2.1                  XML_3.99-0.15                
  [9] lifecycle_1.0.4               edgeR_4.0.1                  
 [11] doParallel_1.0.17             rprojroot_2.0.4              
 [13] processx_3.8.2                lattice_0.22-5               
 [15] magrittr_2.0.3                limma_3.58.1                 
 [17] plotly_4.10.3                 sass_0.4.7                   
 [19] rmarkdown_2.25                remotes_2.4.2.1              
 [21] jquerylib_0.1.4               yaml_2.3.7                   
 [23] httpuv_1.6.12                 spam_2.10-0                  
 [25] ggside_0.2.2                  pkgbuild_1.4.2               
 [27] sessioninfo_1.2.2             cowplot_1.1.1                
 [29] DBI_1.1.3                     RColorBrewer_1.1-3           
 [31] ResidualMatrix_1.12.0         pkgload_1.3.3                
 [33] golem_0.4.1                   maps_3.4.1.1                 
 [35] abind_1.4-5                   zlibbioc_1.48.0              
 [37] purrr_1.0.2                   RCurl_1.98-1.13              
 [39] rappdirs_0.3.3                GenomeInfoDbData_1.2.11      
 [41] ggrepel_0.9.4                 irlba_2.3.5.1                
 [43] DelayedMatrixStats_1.24.0     codetools_0.2-19             
 [45] DelayedArray_0.28.0           DT_0.30                      
 [47] scuttle_1.12.0                tidyselect_1.2.0             
 [49] farver_2.1.1                  ScaledMatrix_1.10.0          
 [51] viridis_0.6.4                 shinyWidgets_0.8.0           
 [53] BiocFileCache_2.10.1          GenomicAlignments_1.38.0     
 [55] jsonlite_1.8.7                BiocNeighbors_1.20.0         
 [57] ellipsis_0.3.2                scater_1.30.1                
 [59] iterators_1.0.14              foreach_1.5.2                
 [61] tools_4.3.1                   glue_1.6.2                   
 [63] SparseArray_1.2.2             xfun_0.41                    
 [65] here_1.0.1                    usethis_2.2.2                
 [67] dplyr_1.1.4                   withr_2.5.2                  
 [69] BiocManager_1.30.22           fastmap_1.1.1                
 [71] fansi_1.0.5                   callr_3.7.3                  
 [73] digest_0.6.33                 rsvd_1.0.5                   
 [75] R6_2.5.1                      mime_0.12                    
 [77] colorspace_2.1-0              RSQLite_2.3.3                
 [79] RhpcBLASctl_0.23-42           config_0.3.2                 
 [81] utf8_1.2.4                    tidyr_1.3.0                  
 [83] generics_0.1.3                rtracklayer_1.62.0           
 [85] prettyunits_1.2.0             httr_1.4.7                   
 [87] htmlwidgets_1.6.2             S4Arrays_1.2.0               
 [89] scatterplot3d_0.3-44          uwot_0.1.16                  
 [91] pkgconfig_2.0.3               gtable_0.3.4                 
 [93] blob_1.2.4                    htmltools_0.5.7              
 [95] profvis_0.3.8                 dotCall64_1.1-0              
 [97] scales_1.2.1                  png_0.1-8                    
 [99] attempt_0.3.1                 knitr_1.45                   
[101] rstudioapi_0.15.0             rjson_0.2.21                 
[103] curl_5.1.0                    cachem_1.0.8                 
[105] Polychrome_1.5.1              stringr_1.5.1                
[107] BiocVersion_3.18.1            parallel_4.3.1               
[109] miniUI_0.1.1.1                vipor_0.4.5                  
[111] AnnotationDbi_1.64.1          restfulr_0.0.15              
[113] pillar_1.9.0                  grid_4.3.1                   
[115] vctrs_0.6.4                   urlchecker_1.0.1             
[117] promises_1.2.1                BiocSingular_1.18.0          
[119] dbplyr_2.4.0                  beachmat_2.18.0              
[121] xtable_1.8-4                  beeswarm_0.4.0               
[123] paletteer_1.5.0               evaluate_0.23                
[125] magick_2.8.1                  cli_3.6.1                    
[127] locfit_1.5-9.8                compiler_4.3.1               
[129] Rsamtools_2.18.0              crayon_1.5.2                 
[131] labeling_0.4.3                rematch2_2.1.2               
[133] ps_1.7.5                      fs_1.6.3                     
[135] ggbeeswarm_0.7.2              stringi_1.8.1                
[137] viridisLite_0.4.2             munsell_0.5.0                
[139] lazyeval_0.2.2                devtools_2.4.5               
[141] Matrix_1.6-3                  ExperimentHub_2.10.0         
[143] benchmarkme_1.0.8             sparseMatrixStats_1.14.0     
[145] bit64_4.0.5                   KEGGREST_1.42.0              
[147] statmod_1.5.0                 shiny_1.8.0                  
[149] interactiveDisplayBase_1.40.0 AnnotationHub_3.10.0         
[151] igraph_1.5.1                  ggspavis_1.8.0               
[153] memoise_2.0.1                 bslib_0.5.1                  
[155] benchmarkmeData_1.0.4         bit_4.0.5                    

###

─ Session info ────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.1 (2023-06-16)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2023-11-22
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────
 package                * version   date (UTC) lib source
 abind                    1.4-5     2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1    2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0    2023-10-24 [1] Bioconductor
 attempt                  0.3.1     2020-05-03 [1] CRAN (R 4.3.0)
 batchelor              * 1.18.0    2023-10-24 [1] Bioconductor
 beachmat                 2.18.0    2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0     2021-06-01 [1] CRAN (R 4.3.0)
 benchmarkme              1.0.8     2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4     2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0    2023-10-26 [1] Bioconductor
 BiocFileCache            2.10.1    2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1    2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0    2023-10-24 [1] Bioconductor
 BiocManager              1.30.22   2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.0    2023-10-24 [1] Bioconductor
 BiocParallel           * 1.36.0    2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0    2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1    2023-11-15 [1] Bioconductor
 Biostrings             * 2.70.1    2023-10-26 [1] Bioconductor
 bit                      4.0.5     2022-11-15 [1] CRAN (R 4.3.0)
 bit64                    4.0.5     2020-08-30 [1] CRAN (R 4.3.0)
 bitops                   1.0-7     2021-04-24 [1] CRAN (R 4.3.0)
 blob                     1.2.4     2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.5.1     2023-08-11 [1] CRAN (R 4.3.0)
 cachem                   1.0.8     2023-05-01 [1] CRAN (R 4.3.0)
 callr                    3.7.3     2022-11-02 [1] CRAN (R 4.3.0)
 cli                      3.6.1     2023-03-23 [1] CRAN (R 4.3.0)
 codetools                0.2-19    2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0     2023-11-01 [1] local (/Users/bmulvey/Desktop/colorout)
 colorspace               2.1-0     2023-01-23 [1] CRAN (R 4.3.0)
 config                   0.3.2     2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.1     2020-12-30 [1] CRAN (R 4.3.0)
 crayon                   1.5.2     2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.1.0     2023-10-02 [1] CRAN (R 4.3.1)
 data.table             * 1.14.8    2023-02-17 [1] CRAN (R 4.3.0)
 DBI                      1.1.3     2022-06-18 [1] CRAN (R 4.3.0)
 dbplyr                   2.4.0     2023-10-26 [1] CRAN (R 4.3.1)
 default                * 1.0.0     2017-08-07 [1] CRAN (R 4.3.0)
 DelayedArray             0.28.0    2023-11-06 [1] Bioconductor
 DelayedMatrixStats       1.24.0    2023-10-24 [1] Bioconductor
 devtools                 2.4.5     2022-10-11 [1] CRAN (R 4.3.0)
 digest                   0.6.33    2023-07-07 [1] CRAN (R 4.3.0)
 doParallel               1.0.17    2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-0     2023-10-17 [1] CRAN (R 4.3.1)
 dplyr                    1.1.4     2023-11-17 [1] CRAN (R 4.3.1)
 DT                       0.30      2023-10-05 [1] CRAN (R 4.3.1)
 edgeR                    4.0.1     2023-10-29 [1] Bioconductor
 ellipsis                 0.3.2     2021-04-29 [1] CRAN (R 4.3.0)
 evaluate                 0.23      2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0    2023-10-24 [1] Bioconductor
 fansi                    1.0.5     2023-10-08 [1] CRAN (R 4.3.1)
 farver                   2.1.1     2022-07-06 [1] CRAN (R 4.3.0)
 fastmap                  1.1.1     2023-02-24 [1] CRAN (R 4.3.0)
 fields                   15.2      2023-08-17 [1] CRAN (R 4.3.0)
 filelock                 1.0.2     2018-10-05 [1] CRAN (R 4.3.0)
 foreach                  1.5.2     2022-02-02 [1] CRAN (R 4.3.0)
 fs                       1.6.3     2023-07-20 [1] CRAN (R 4.3.0)
 generics                 0.1.3     2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.1    2023-11-08 [1] Bioconductor
 GenomeInfoDbData         1.2.11    2023-11-10 [1] Bioconductor
 GenomicAlignments        1.38.0    2023-10-24 [1] Bioconductor
 GenomicRanges          * 1.54.1    2023-10-30 [1] Bioconductor
 ggbeeswarm               0.7.2     2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4     2023-10-12 [1] CRAN (R 4.3.1)
 ggrepel                  0.9.4     2023-10-13 [1] CRAN (R 4.3.1)
 ggside                   0.2.2     2022-12-04 [1] CRAN (R 4.3.0)
 ggspavis                 1.8.0     2023-10-24 [1] Bioconductor
 glue                     1.6.2     2022-02-24 [1] CRAN (R 4.3.0)
 golem                    0.4.1     2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3       2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4     2023-08-21 [1] CRAN (R 4.3.0)
 harmony                * 1.2.0     2023-11-16 [1] Github (immunogenomics/harmony@f054b03)
 here                     1.0.1     2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7     2023-11-03 [1] CRAN (R 4.3.1)
 htmlwidgets              1.6.2     2023-03-17 [1] CRAN (R 4.3.0)
 httpuv                   1.6.12    2023-10-23 [1] CRAN (R 4.3.1)
 httr                     1.4.7     2023-08-15 [1] CRAN (R 4.3.0)
 igraph                   1.5.1     2023-08-10 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0    2023-10-24 [1] Bioconductor
 IRanges                * 2.36.0    2023-10-26 [1] Bioconductor
 irlba                    2.3.5.1   2022-10-03 [1] CRAN (R 4.3.1)
 iterators                1.0.14    2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4     2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.7     2023-06-29 [1] CRAN (R 4.3.0)
 KEGGREST                 1.42.0    2023-10-26 [1] Bioconductor
 knitr                    1.45      2023-10-30 [1] CRAN (R 4.3.1)
 labeling                 0.4.3     2023-08-29 [1] CRAN (R 4.3.0)
 later                    1.3.1     2023-05-02 [1] CRAN (R 4.3.0)
 lattice                  0.22-5    2023-10-24 [1] CRAN (R 4.3.1)
 lazyeval                 0.2.2     2019-03-15 [1] CRAN (R 4.3.0)
 lifecycle                1.0.4     2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1    2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8   2023-06-11 [1] CRAN (R 4.3.0)
 magick                   2.8.1     2023-10-22 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3     2022-03-30 [1] CRAN (R 4.3.0)
 maps                     3.4.1.1   2023-11-03 [1] CRAN (R 4.3.1)
 Matrix                   1.6-3     2023-11-14 [1] CRAN (R 4.3.1)
 MatrixGenerics         * 1.14.0    2023-10-26 [1] Bioconductor
 matrixStats            * 1.1.0     2023-11-07 [1] CRAN (R 4.3.1)
 memoise                  2.0.1     2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12      2021-09-28 [1] CRAN (R 4.3.0)
 miniUI                   0.1.1.1   2018-05-18 [1] CRAN (R 4.3.0)
 munsell                  0.5.0     2018-06-12 [1] CRAN (R 4.3.0)
 paletteer                1.5.0     2022-10-19 [1] CRAN (R 4.3.0)
 parallelly             * 1.36.0    2023-05-26 [1] CRAN (R 4.3.0)
 pillar                   1.9.0     2023-03-22 [1] CRAN (R 4.3.0)
 pkgbuild                 1.4.2     2023-06-26 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3     2019-09-22 [1] CRAN (R 4.3.0)
 pkgload                  1.3.3     2023-09-22 [1] CRAN (R 4.3.1)
 plotly                   4.10.3    2023-10-21 [1] CRAN (R 4.3.1)
 png                      0.1-8     2022-11-29 [1] CRAN (R 4.3.0)
 Polychrome               1.5.1     2022-05-03 [1] CRAN (R 4.3.0)
 prettyunits              1.2.0     2023-09-24 [1] CRAN (R 4.3.1)
 processx                 3.8.2     2023-06-30 [1] CRAN (R 4.3.0)
 profvis                  0.3.8     2023-05-02 [1] CRAN (R 4.3.0)
 promises                 1.2.1     2023-08-10 [1] CRAN (R 4.3.0)
 ps                       1.7.5     2023-04-18 [1] CRAN (R 4.3.0)
 purrr                    1.0.2     2023-08-10 [1] CRAN (R 4.3.0)
 R6                       2.5.1     2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3     2021-01-31 [1] CRAN (R 4.3.0)
 RColorBrewer             1.1-3     2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                   * 1.0.11    2023-07-06 [1] CRAN (R 4.3.1)
 RcppAnnoy                0.0.21    2023-07-02 [1] CRAN (R 4.3.0)
 RCurl                    1.98-1.13 2023-11-02 [1] CRAN (R 4.3.1)
 rematch2                 2.1.2     2020-05-01 [1] CRAN (R 4.3.0)
 remotes                  2.4.2.1   2023-07-18 [1] CRAN (R 4.3.0)
 ResidualMatrix           1.12.0    2023-10-24 [1] Bioconductor
 restfulr                 0.0.15    2022-06-16 [1] CRAN (R 4.3.0)
 RhpcBLASctl              0.23-42   2023-02-11 [1] CRAN (R 4.3.0)
 rjson                    0.2.21    2022-01-09 [1] CRAN (R 4.3.0)
 rlang                  * 1.1.2     2023-11-04 [1] CRAN (R 4.3.1)
 rmarkdown                2.25      2023-09-18 [1] CRAN (R 4.3.1)
 rprojroot                2.0.4     2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0    2023-10-24 [1] Bioconductor
 RSQLite                  2.3.3     2023-11-04 [1] CRAN (R 4.3.1)
 rstudioapi               0.15.0    2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5     2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0    2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0     2023-10-26 [1] Bioconductor
 S4Vectors              * 0.40.1    2023-10-26 [1] Bioconductor
 sass                     0.4.7     2023-07-15 [1] CRAN (R 4.3.0)
 ScaledMatrix             1.10.0    2023-10-24 [1] Bioconductor
 scales                   1.2.1     2022-08-20 [1] CRAN (R 4.3.0)
 scater                   1.30.1    2023-11-16 [1] Bioconductor
 scatterplot3d            0.3-44    2023-05-05 [1] CRAN (R 4.3.0)
 scuttle                  1.12.0    2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2     2021-12-06 [1] CRAN (R 4.3.0)
 shiny                    1.8.0     2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.0     2023-08-30 [1] CRAN (R 4.3.0)
 SingleCellExperiment   * 1.24.0    2023-11-06 [1] Bioconductor
 spam                     2.10-0    2023-10-23 [1] CRAN (R 4.3.1)
 SparseArray              1.2.2     2023-11-08 [1] Bioconductor
 sparseMatrixStats        1.14.0    2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0    2023-10-24 [1] Bioconductor
 spatialLIBD            * 1.13.4    2023-05-25 [1] Bioconductor
 statmod                  1.5.0     2023-01-06 [1] CRAN (R 4.3.0)
 stringi                  1.8.1     2023-11-13 [1] CRAN (R 4.3.1)
 stringr                  1.5.1     2023-11-14 [1] CRAN (R 4.3.1)
 SummarizedExperiment   * 1.32.0    2023-11-06 [1] Bioconductor
 tibble                   3.2.1     2023-03-20 [1] CRAN (R 4.3.0)
 tidyr                    1.3.0     2023-01-24 [1] CRAN (R 4.3.0)
 tidyselect               1.2.0     2022-10-10 [1] CRAN (R 4.3.0)
 urlchecker               1.0.1     2021-11-30 [1] CRAN (R 4.3.0)
 usethis                  2.2.2     2023-07-06 [1] CRAN (R 4.3.0)
 utf8                     1.2.4     2023-10-22 [1] CRAN (R 4.3.1)
 uwot                     0.1.16    2023-06-29 [1] CRAN (R 4.3.0)
 vctrs                    0.6.4     2023-10-12 [1] CRAN (R 4.3.1)
 vipor                    0.4.5     2017-03-22 [1] CRAN (R 4.3.0)
 viridis                  0.6.4     2023-07-22 [1] CRAN (R 4.3.0)
 viridisLite              0.4.2     2023-05-02 [1] CRAN (R 4.3.0)
 withr                    2.5.2     2023-10-30 [1] CRAN (R 4.3.1)
 xfun                     0.41      2023-11-01 [1] CRAN (R 4.3.1)
 XML                      3.99-0.15 2023-11-02 [1] CRAN (R 4.3.1)
 xtable                   1.8-4     2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0    2023-10-26 [1] Bioconductor
 yaml                     2.3.7     2023-01-23 [1] CRAN (R 4.3.0)
 zlibbioc                 1.48.0    2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

───────────────────────────────────────────────────────────────────────────
