---
title: "Fig 2 Panels"
author: "Bernard Mulvey"
date: "2023-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10, fig.width = 7, include = FALSE)
knitr::opts_chunk$set(fig.width = 7, fig.height = 10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table) # Preferred data manipulation package
library(Biostrings) # General
library(ggplot2) # Dependency for several plotting functions
library(SpatialExperiment) # Dependency
library(spatialLIBD) # Fig 2A
require(colorout) # Utility for RStudio
ColorOut()

# code reformatting in Rstudio
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

# set plotting defaults for ggplot
theme_set(theme_bw() + theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 11), axis.text.y = element_text(size = 10), axis.title.y = element_text(size = 11), plot.title = element_text(size = 12, hjust = 0.5), strip.text = element_text(size = 18), legend.text = element_text(size = 9), legend.title = element_text(size = 10, hjust = 0.5)))
```

### Setup: load data, cluster assignments, and rename from generic to appropriate region (VMH/ARC/WM)
```{r}
hyp2 <- readRDS("data/03-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

bscl <- fread("data/06-BayesSpace/02-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt")

setnames(bscl, c("rn", "cl"))
bscl[, cl := paste0("X", cl)]
bscl <- DataFrame(bscl)
rownames(bscl) <- bscl$rn
bscl <- bscl[colnames(hyp2), ]

colLabels(hyp2) <- factor(bscl$cl, levels = paste0("X", c(1:15)))

### relabel as VMH/ARC/WM for spotplottings
tmpcd <- as.data.table(colData(hyp2), keep.rownames = T)
tmpcd[label == "X11", label := "VMH.1"]
tmpcd[label == "X12", label := "VMH.2"]
tmpcd[label == "X3", label := "ARC.1"]
tmpcd[label == "X5", label := "ARC.2"]
tmpcd[label == "X1", label := "WM.1"]
tmpcd[label == "X9", label := "WM.2"]
tmpcd <- DataFrame(tmpcd)
rownames(tmpcd) <- tmpcd$rn
tmpcd <- tmpcd[colnames(hyp2), ]
colData(hyp2) <- tmpcd
### set ordering so named clusters lead and the rest are in numerical order
hyp2$label <- factor(hyp2$label, levels = c("VMH.1", "VMH.2", "ARC.1", "ARC.2", "WM.1", "WM.2", paste0("X", c(2, 4, 6:8, 10, 13:15))))

### make rownames == gene names
rownames(hyp2) <- rowData(hyp2)$gene_name

```

### Panel C (part): NR5A1 spotplot in V12Y31-080_A1 ###
```{r}
tmp <- hyp2[, hyp2$sample_id == "V12Y31-080_A1"]

plt <- vis_gene(tmp,sampleid="V12Y31-080_A1",geneid = "NR5A1",spatial = T,na_color = NA,viridis = FALSE,minCount = 0,auto_crop = T,point_size = 0.6,alpha=0.65)+
    ggtitle("NR5A1")+
    theme(plot.title = element_text(size = 11, hjust = 0.5,face="italic",margin=margin(0,0,0,0)),
          legend.title = element_text(size=8,hjust=0.5),
          legend.text=element_text(size=8),
          panel.border = element_blank(),
          plot.margin = margin(-0.075,0.05,-0.45,-0.35,unit="in"),
          legend.margin = margin(0.5,0,0.5,0,"in"),
          legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"))

pdf("plots/Manuscript/Fig1C-NR5A1.pdf",height=1.57,width=1.87)
plt
dev.off()

```

### Panel D (part): NPY in V12Y31-080_A1
```{r}
tmp <- hyp2[, hyp2$sample_id == "V12D05-350_D1"]

plt <- vis_gene(tmp,sampleid="V12D05-350_D1",geneid = "NPY",spatial = T,na_color = NA,viridis = FALSE,minCount = 0,auto_crop = T,point_size = 0.6,alpha=0.65)+
    ggtitle("NPY")+
    theme(plot.title = element_text(size = 11, hjust = 0.5,face="italic",margin=margin(0,0,0,0)),
          legend.title = element_text(size=8,hjust=0.5),
          legend.text=element_text(size=8),
          panel.border = element_blank(),
          plot.margin = margin(-0.075,0.075,-0.475,-0.425,unit="in"),
          legend.margin = margin(0.5,0,0.5,0.025,"in"),
          legend.key.height = unit(0.083,"in"),
          legend.key.width = unit(0.156,"in"),
          )

pdf("plots/Manuscript/Fig1D-NPY.pdf",height=1.65,width=1.87)
plt
dev.off()
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] gridExtra_2.3               colorout_1.3-0.1           
 [3] spatialLIBD_1.14.1          SpatialExperiment_1.12.0   
 [5] SingleCellExperiment_1.24.0 SummarizedExperiment_1.32.0
 [7] Biobase_2.62.0              GenomicRanges_1.54.1       
 [9] MatrixGenerics_1.14.0       matrixStats_1.2.0          
[11] ggplot2_3.4.4               Biostrings_2.70.2          
[13] GenomeInfoDb_1.38.5         XVector_0.42.0             
[15] IRanges_2.36.0              S4Vectors_0.40.2           
[17] BiocGenerics_0.48.1         data.table_1.14.10         
[19] rlang_1.1.3                

loaded via a namespace (and not attached):
  [1] later_1.3.2                   BiocIO_1.12.0                
  [3] bitops_1.0-7                  filelock_1.0.3               
  [5] fields_15.2                   tibble_3.2.1                 
  [7] XML_3.99-0.16.1               lifecycle_1.0.4              
  [9] edgeR_4.0.14                  doParallel_1.0.17            
 [11] rprojroot_2.0.4               lattice_0.22-5               
 [13] magrittr_2.0.3                limma_3.58.1                 
 [15] plotly_4.10.4                 sass_0.4.8                   
 [17] rmarkdown_2.25                jquerylib_0.1.4              
 [19] yaml_2.3.8                    httpuv_1.6.13                
 [21] spam_2.10-0                   sessioninfo_1.2.2            
 [23] cowplot_1.1.3                 DBI_1.2.1                    
 [25] RColorBrewer_1.1-3            golem_0.4.1                  
 [27] maps_3.4.2                    abind_1.4-5                  
 [29] zlibbioc_1.48.0               purrr_1.0.2                  
 [31] RCurl_1.98-1.14               rappdirs_0.3.3               
 [33] circlize_0.4.15               GenomeInfoDbData_1.2.11      
 [35] ggrepel_0.9.5                 irlba_2.3.5.1                
 [37] pheatmap_1.0.12               DelayedMatrixStats_1.24.0    
 [39] codetools_0.2-19              DelayedArray_0.28.0          
 [41] DT_0.31                       scuttle_1.12.0               
 [43] shape_1.4.6                   tidyselect_1.2.0             
 [45] farver_2.1.1                  ScaledMatrix_1.10.0          
 [47] viridis_0.6.5                 shinyWidgets_0.8.1           
 [49] BiocFileCache_2.10.1          GenomicAlignments_1.38.2     
 [51] jsonlite_1.8.8                GetoptLong_1.0.5             
 [53] BiocNeighbors_1.20.2          ellipsis_0.3.2               
 [55] scater_1.30.1                 iterators_1.0.14             
 [57] foreach_1.5.2                 tools_4.3.2                  
 [59] Rcpp_1.0.12                   glue_1.7.0                   
 [61] SparseArray_1.2.3             xfun_0.41                    
 [63] here_1.0.1                    dplyr_1.1.4                  
 [65] withr_3.0.0                   BiocManager_1.30.22          
 [67] fastmap_1.1.1                 fansi_1.0.6                  
 [69] digest_0.6.34                 rsvd_1.0.5                   
 [71] R6_2.5.1                      mime_0.12                    
 [73] colorspace_2.1-0              RSQLite_2.3.5                
 [75] config_0.3.2                  utf8_1.2.4                   
 [77] tidyr_1.3.1                   generics_0.1.3               
 [79] rtracklayer_1.62.0            httr_1.4.7                   
 [81] htmlwidgets_1.6.4             S4Arrays_1.2.0               
 [83] scatterplot3d_0.3-44          pkgconfig_2.0.3              
 [85] gtable_0.3.4                  blob_1.2.4                   
 [87] ComplexHeatmap_2.18.0         htmltools_0.5.7              
 [89] dotCall64_1.1-1               clue_0.3-65                  
 [91] scales_1.3.0                  png_0.1-8                    
 [93] attempt_0.3.1                 knitr_1.45                   
 [95] rstudioapi_0.15.0             rjson_0.2.21                 
 [97] curl_5.2.0                    GlobalOptions_0.1.2          
 [99] cachem_1.0.8                  Polychrome_1.5.1             
[101] BiocVersion_3.18.1            parallel_4.3.2               
[103] vipor_0.4.7                   AnnotationDbi_1.64.1         
[105] restfulr_0.0.15               pillar_1.9.0                 
[107] grid_4.3.2                    vctrs_0.6.5                  
[109] promises_1.2.1                BiocSingular_1.18.0          
[111] dbplyr_2.4.0                  beachmat_2.18.0              
[113] cluster_2.1.6                 xtable_1.8-4                 
[115] beeswarm_0.4.0                paletteer_1.6.0              
[117] evaluate_0.23                 magick_2.8.2                 
[119] cli_3.6.2                     locfit_1.5-9.8               
[121] compiler_4.3.2                Rsamtools_2.18.0             
[123] crayon_1.5.2                  labeling_0.4.3               
[125] rematch2_2.1.2                ggbeeswarm_0.7.2             
[127] viridisLite_0.4.2             BiocParallel_1.36.0          
[129] munsell_0.5.0                 lazyeval_0.2.2               
[131] Matrix_1.6-5                  ExperimentHub_2.10.0         
[133] benchmarkme_1.0.8             sparseMatrixStats_1.14.0     
[135] bit64_4.0.5                   KEGGREST_1.42.0              
[137] statmod_1.5.0                 shiny_1.8.0                  
[139] interactiveDisplayBase_1.40.0 AnnotationHub_3.10.0         
[141] memoise_2.0.1                 bslib_0.6.1                  
[143] benchmarkmeData_1.0.4         bit_4.0.5                    
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-02-08
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package                * version     date (UTC) lib source
 abind                    1.4-5       2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1      2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0      2023-10-24 [1] Bioconductor
 attempt                  0.3.1       2020-05-03 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0      2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0       2021-06-01 [1] CRAN (R 4.3.2)
 benchmarkme              1.0.8       2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4       2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0      2023-10-24 [1] Bioconductor
 BiocFileCache            2.10.1      2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1      2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0      2023-10-26 [1] Bioconductor
 BiocManager              1.30.22     2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.2      2024-01-07 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel             1.36.0      2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0      2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1      2023-11-18 [1] Bioconductor 3.18 (R 4.3.2)
 Biostrings             * 2.70.2      2024-01-28 [1] Bioconductor 3.18 (R 4.3.2)
 bit                      4.0.5       2022-11-15 [1] CRAN (R 4.3.2)
 bit64                    4.0.5       2020-08-30 [1] CRAN (R 4.3.2)
 bitops                   1.0-7       2021-04-24 [1] CRAN (R 4.3.2)
 blob                     1.2.4       2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.6.1       2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8       2023-05-01 [1] CRAN (R 4.3.2)
 circlize                 0.4.15      2022-05-10 [1] CRAN (R 4.3.0)
 cli                      3.6.2       2023-12-11 [1] CRAN (R 4.3.2)
 clue                     0.3-65      2023-09-23 [1] CRAN (R 4.3.2)
 cluster                  2.1.6       2023-12-01 [1] CRAN (R 4.3.2)
 codetools                0.2-19      2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0.1     2024-01-11 [1] local
 colorspace               2.1-0       2023-01-23 [1] CRAN (R 4.3.2)
 ComplexHeatmap           2.18.0      2023-10-26 [1] Bioconductor
 config                   0.3.2       2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.3       2024-01-22 [1] CRAN (R 4.3.1)
 crayon                   1.5.2       2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0       2023-12-08 [1] CRAN (R 4.3.2)
 data.table             * 1.14.10     2023-12-08 [1] CRAN (R 4.3.2)
 DBI                      1.2.1       2024-01-12 [1] CRAN (R 4.3.2)
 dbplyr                   2.4.0       2023-10-26 [1] CRAN (R 4.3.1)
 DelayedArray             0.28.0      2023-10-24 [1] Bioconductor
 DelayedMatrixStats       1.24.0      2023-11-06 [1] Bioconductor
 digest                   0.6.34      2024-01-11 [1] CRAN (R 4.3.2)
 doParallel               1.0.17      2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1       2023-11-28 [1] CRAN (R 4.3.2)
 dplyr                    1.1.4       2023-11-17 [1] CRAN (R 4.3.2)
 DT                       0.31        2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.14      2024-01-29 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2       2021-04-29 [1] CRAN (R 4.3.2)
 evaluate                 0.23        2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0      2023-10-26 [1] Bioconductor
 fansi                    1.0.6       2023-12-08 [1] CRAN (R 4.3.2)
 farver                   2.1.1       2022-07-06 [1] CRAN (R 4.3.2)
 fastmap                  1.1.1       2023-02-24 [1] CRAN (R 4.3.2)
 fields                   15.2        2023-08-17 [1] CRAN (R 4.3.2)
 filelock                 1.0.3       2023-12-11 [1] CRAN (R 4.3.2)
 foreach                  1.5.2       2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3       2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.5      2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData         1.2.11      2024-01-11 [1] Bioconductor
 GenomicAlignments        1.38.2      2024-01-16 [1] Bioconductor 3.18 (R 4.3.2)
 GenomicRanges          * 1.54.1      2023-10-29 [1] Bioconductor
 GetoptLong               1.0.5       2020-12-15 [1] CRAN (R 4.3.0)
 ggbeeswarm               0.7.2       2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4       2023-10-12 [1] CRAN (R 4.3.1)
 ggrepel                  0.9.5       2024-01-10 [1] CRAN (R 4.3.2)
 GlobalOptions            0.1.2       2020-06-10 [1] CRAN (R 4.3.0)
 glue                     1.7.0       2024-01-09 [1] CRAN (R 4.3.2)
 golem                    0.4.1       2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3         2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4       2023-08-21 [1] CRAN (R 4.3.0)
 here                     1.0.1       2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7       2023-11-03 [1] CRAN (R 4.3.2)
 htmlwidgets              1.6.4       2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13      2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7       2023-08-15 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0      2023-10-26 [1] Bioconductor
 IRanges                * 2.36.0      2023-10-24 [1] Bioconductor
 irlba                    2.3.5.1     2022-10-03 [1] CRAN (R 4.3.2)
 iterators                1.0.14      2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4       2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8       2023-12-04 [1] CRAN (R 4.3.2)
 KEGGREST                 1.42.0      2023-10-26 [1] Bioconductor
 knitr                    1.45        2023-10-30 [1] CRAN (R 4.3.2)
 labeling                 0.4.3       2023-08-29 [1] CRAN (R 4.3.0)
 later                    1.3.2       2023-12-06 [1] CRAN (R 4.3.2)
 lattice                  0.22-5      2023-10-24 [1] CRAN (R 4.3.2)
 lazyeval                 0.2.2       2019-03-15 [1] CRAN (R 4.3.2)
 lifecycle                1.0.4       2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1      2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8     2023-06-11 [1] CRAN (R 4.3.2)
 magick                   2.8.2       2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3       2022-03-30 [1] CRAN (R 4.3.2)
 maps                     3.4.2       2023-12-15 [1] CRAN (R 4.3.1)
 Matrix                   1.6-5       2024-01-11 [1] CRAN (R 4.3.2)
 MatrixGenerics         * 1.14.0      2023-10-26 [1] Bioconductor
 matrixStats            * 1.2.0       2023-12-11 [1] CRAN (R 4.3.2)
 memoise                  2.0.1       2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12        2021-09-28 [1] CRAN (R 4.3.2)
 munsell                  0.5.0       2018-06-12 [1] CRAN (R 4.3.0)
 paletteer                1.6.0       2024-01-21 [1] CRAN (R 4.3.1)
 pheatmap                 1.0.12      2019-01-04 [1] CRAN (R 4.3.0)
 pillar                   1.9.0       2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3       2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.4      2024-01-13 [1] CRAN (R 4.3.1)
 png                      0.1-8       2022-11-29 [1] CRAN (R 4.3.2)
 Polychrome               1.5.1       2022-05-03 [1] CRAN (R 4.3.0)
 promises                 1.2.1       2023-08-10 [1] CRAN (R 4.3.2)
 purrr                    1.0.2       2023-08-10 [1] CRAN (R 4.3.2)
 R6                       2.5.1       2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3       2021-01-31 [1] CRAN (R 4.3.2)
 RColorBrewer             1.1-3       2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.12      2024-01-09 [1] CRAN (R 4.3.2)
 RCurl                    1.98-1.14   2024-01-09 [1] CRAN (R 4.3.1)
 rematch2                 2.1.2       2020-05-01 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15      2022-06-16 [1] CRAN (R 4.3.2)
 rjson                    0.2.21      2022-01-09 [1] CRAN (R 4.3.2)
 rlang                  * 1.1.3       2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown                2.25        2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot                2.0.4       2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0      2023-10-24 [1] Bioconductor
 RSQLite                  2.3.5       2024-01-21 [1] CRAN (R 4.3.2)
 rstudioapi               0.15.0      2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5       2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0      2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0       2023-10-24 [1] Bioconductor
 S4Vectors              * 0.40.2      2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8       2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0      2023-11-06 [1] Bioconductor
 scales                   1.3.0       2023-11-28 [1] CRAN (R 4.3.2)
 scater                   1.30.1      2023-11-16 [1] Bioconductor
 scatterplot3d            0.3-44      2023-05-05 [1] CRAN (R 4.3.0)
 scuttle                  1.12.0      2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2       2021-12-06 [1] CRAN (R 4.3.0)
 shape                    1.4.6       2021-05-19 [1] CRAN (R 4.3.0)
 shiny                    1.8.0       2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.1       2024-01-10 [1] CRAN (R 4.3.1)
 SingleCellExperiment   * 1.24.0      2023-11-06 [1] Bioconductor
 spam                     2.10-0      2023-10-23 [1] CRAN (R 4.3.2)
 SparseArray              1.2.3       2023-12-25 [1] Bioconductor 3.18 (R 4.3.2)
 sparseMatrixStats        1.14.0      2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0      2023-10-26 [1] Bioconductor
 spatialLIBD            * 1.14.1      2023-11-30 [1] Bioconductor 3.18 (R 4.3.2)
 statmod                  1.5.0       2023-01-06 [1] CRAN (R 4.3.2)
 SummarizedExperiment   * 1.32.0      2023-11-06 [1] Bioconductor
 tibble                   3.2.1       2023-03-20 [1] CRAN (R 4.3.2)
 tidyr                    1.3.1       2024-01-24 [1] CRAN (R 4.3.1)
 tidyselect               1.2.0       2022-10-10 [1] CRAN (R 4.3.0)
 utf8                     1.2.4       2023-10-22 [1] CRAN (R 4.3.2)
 vctrs                    0.6.5       2023-12-01 [1] CRAN (R 4.3.2)
 vipor                    0.4.7       2023-12-18 [1] CRAN (R 4.3.1)
 viridis                  0.6.5       2024-01-29 [1] CRAN (R 4.3.2)
 viridisLite              0.4.2       2023-05-02 [1] CRAN (R 4.3.0)
 withr                    3.0.0       2024-01-16 [1] CRAN (R 4.3.1)
 xfun                     0.41        2023-11-01 [1] CRAN (R 4.3.2)
 XML                      3.99-0.16.1 2024-01-22 [1] CRAN (R 4.3.1)
 xtable                   1.8-4       2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0      2023-10-24 [1] Bioconductor
 yaml                     2.3.8       2023-12-11 [1] CRAN (R 4.3.2)
 zlibbioc                 1.48.0      2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
