---
title: "Fig 2 Panels"
author: "Bernard Mulvey"
date: "2023-12-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10, fig.width = 7, include = FALSE)
knitr::opts_chunk$set(fig.width = 7, fig.height = 10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table) # Preferred data manipulation package
library(Biostrings) # General
library(ggrastr) # render plots with bajillions of points as raster so they can be manipulated in eg Illustrator--Fig 2B
library(Cairo) # pdf device for ^ to properly save; also used for complexheatmap raster functions
library(ggplot2) # Dependency for several plotting functions
library(gridExtra) # Fig 2A
library(SpatialExperiment) # Dependency
library(spatialLIBD) # Fig 2A
library(scater) # Fig 2B
require(colorout) # Utility for RStudio
library(Polychrome) # Generate a visually-discernible palette for 15 colors
library(ComplexHeatmap)
ColorOut()

# code reformatting in Rstudio
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

# set plotting defaults for ggplot
theme_set(theme_bw() + theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 11), axis.text.y = element_text(size = 10), axis.title.y = element_text(size = 11), plot.title = element_text(size = 12, hjust = 0.5), strip.text = element_text(size = 18), legend.text = element_text(size = 9), legend.title = element_text(size = 10, hjust = 0.5)))
```

### Setup: load data, cluster assignments, and rename from generic to appropriate region (VMH/ARC/WM)
```{r}
hyp2 <- readRDS("data/03-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

bscl <- fread("data/06-BayesSpace/02-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt")

setnames(bscl, c("rn", "cl"))
bscl[, cl := paste0("X", cl)]
bscl <- DataFrame(bscl)
rownames(bscl) <- bscl$rn
bscl <- bscl[colnames(hyp2), ]

colLabels(hyp2) <- factor(bscl$cl, levels = paste0("X", c(1:15)))

### relabel as VMH/ARC/WM for spotplottings
tmpcd <- as.data.table(colData(hyp2), keep.rownames = T)
tmpcd[label == "X11", label := "VMH.1"]
tmpcd[label == "X12", label := "VMH.2"]
tmpcd[label == "X3", label := "ARC.1"]
tmpcd[label == "X5", label := "ARC.2"]
tmpcd[label == "X1", label := "WM.1"]
tmpcd[label == "X9", label := "WM.2"]
tmpcd <- DataFrame(tmpcd)
rownames(tmpcd) <- tmpcd$rn
tmpcd <- tmpcd[colnames(hyp2), ]
colData(hyp2) <- tmpcd
### set ordering so named clusters lead and the rest are in numerical order
hyp2$label <- factor(hyp2$label, levels = c("VMH.1", "VMH.2", "ARC.1", "ARC.2", "WM.1", "WM.2", paste0("X", c(2, 4, 6:8, 10, 13:15))))
```

## ONE-TIME RUN: build the palette for ALL figures. feed in the colors we intend to use for VMH (first two), ARC (next two), and WM (last two) as seedColors to prevent similar colors from being included in the palette.
```{r}
# set.seed(424242)
# pal <- Polychrome::createPalette(N = 15,M=50000,seedcolors = c("#4a2881","#660053","#058a00","#009682","#4d4d4d","#bababa"),range=c(20,70))
# preview
# swatch(pal)
# keep 1:9,12
# newseed <- pal[c(1:9,12)]
# set.seed(84848484)
# pal2 <- Polychrome::createPalette(N = 15,M=90000,seedcolors = newseed,range=c(20,80))
# swatch(pal2)

## ugh, alright so now also keep 10, 11, 13, 14. 11 and 12 look too much alike though and 15 is too ambiguously like several others.
# newseed2 <- pal2[c(1:11,13:14)]
# set.seed(84848484)
# pal3 <- Polychrome::createPalette(N = 15,M=90000,seedcolors = newseed2,range=c(20,80))
# swatch(pal3)

## okay, make a 16 color palette with the first 15, and then drop 15 (still that salmon color that's too indistinct)
# set.seed(84848484)
# pal4 <- Polychrome::createPalette(N = 16,M=90000,seedcolors = pal3,range=c(20,80))
# swatch(pal4)

# THERE we go.
# pal <- pal4[c(1:14,16)]
# pal values: "#3B3B5C","#6A0D53","#269422","#169482","#3D4047","#A9A7A9","#F70022","#551CFD","#F59B26","#00B3FD","#FF0DF7","#D193FE","#D1CA1C","#FB479F"

# swatch(pal)
# beautiful

# tweaked the first color to be more vibrant after feedback from JHU spatial group.
# pal <- c("#4a2881",pal[2:14])
# saveRDS(pal,"plots/Manuscript/k15_palette.RDS")


# rm(pal,pal2,pal3,pal4,newseed,newseed2)
```

# load palette and assign names for consistency in color-cluster assignments between libdspatial and the generated legend below

```{r}
pal <- readRDS("plots/Manuscript/k15_palette.RDS")
names(pal) <- levels(hyp2$label)
```

### Panel A: k15 clustering ###


### custom function to extrct legend from one plot to extend to all ###
```{r}
extract_legend <- function(my_ggp) {
    step1 <- ggplot_gtable(ggplot_build(my_ggp))
    step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
    step3 <- step1$grobs[[step2]]
    return(step3)
}
```

### run this block twice -- the first time its run, axes will come out with numbers on the first sample ###
```{r}
### extract two samples to plot; rename to change appearance in the plotted clustering results ###
tmp <- hyp2[, hyp2$sample_id == "V12Y31-080_A1"]
tmp$sample_id <- "V12Y31-080_A1 (Male)"
hyp2samp <- hyp2[, hyp2$sample_id == "V12D05-350_D1"]
hyp2samp$sample_id <- "V12D05-350_D1 (Female)"

hyp2samp <- cbind(hyp2samp, tmp)
rm(tmp)
gc(full = T)


## create a unified legend spanning all possible cluster values (regardless of presence in a given sample)
# ## make sure the legend is generated from data with all 15 clusters present
# length(unique(hyp2samp$label))
# 15

leg.tmp <- ggplot(as.data.table(colData(hyp2samp))[, .(label, array_row, array_col)], aes(x = array_row, y = array_col, col = label)) +
    geom_point() +
    scale_color_manual(values = as.character(pal)) +
    theme(
        legend.text = element_text(size = 9),
        legend.title = element_blank(),
        legend.spacing.y = unit(0, "cm"),
        legend.spacing.x = unit(0, "cm")
    )
leg.tmp2 <- extract_legend(leg.tmp)
dev.off()

### make spotplots
plt <-
    spatialLIBD::vis_grid_clus(
        hyp2samp,
        clustervar = "label",
        pdf_file = NULL,
        colors = pal,
        spatial = T,
        alpha = 0.7,
        auto_crop = T,
        return_plots = T,
        point_size = 0.975,
        sort_clust = F
    )

### adjust font sizes, drop individual plot legends, drop axis text since this function sometimes glitches out; expand plotting area margins to reduce whitespace waste
plt <- lapply(
    plt,
    FUN = function(x) {
        x +
            theme(
                plot.title = element_text(size = 10, hjust = 0.5, margin = margin(0, 0, 0, 0)),
                legend.title = element_blank(),
                axis.text = element_blank()
            ) +
            # plot.margin = margin(0,-0.125,-0.25,-0.125,"in"))+
            guides(fill = "none")
    }
)

# give the second sample more space to be plotted so the plots are same size (since one is rotated and other isnt)
plt[[1]] <- plt[[1]] + theme(plot.margin = margin(0, -0.1, -0.2, -0.1, "in"))
plt[[2]] <- plt[[2]] + theme(plot.margin = margin(0, -0.175, -0.35, -0.175, "in"))


pdf("plots/Manuscript/2A-Bayesspace k15 cluster spotplots.pdf", height = 5, width = 3.75)
### arrangeGrob() on the spotplot list --> 2x1 of plots from that list only, constituting the "first column" of the parent; leg.tmp2 is an additional thing being plotted in the second column of the parent.
do.call("grid.arrange", list(arrangeGrob(grobs = plt, ncol = 1, nrow = 2), leg.tmp2, ncol = 2, widths = c(3, 1)))
dev.off()

rm(plt, leg.tmp, leg.tmp2, hyp2samp)
gc(full = T)
```


### 2B: Marker violins ###
```{r}
rownames(hyp2) <- rowData(hyp2)$gene_name
hypselect <- hyp2[, hyp2$label %in% c("VMH.1", "VMH.2", "ARC.1", "ARC.2", "WM.1", "WM.2")]

## easiest way to get strip titles to show as desired is to just rename the variable (rowname) being used to title them:
hypselect <- hypselect[c("NR5A1", "AGRP", "FEZF1", "NPY", "LAMP5", "POMC", "ADCYAP1", "GHRH", "MBP", "ERMN"), ]
stopifnot(nrow(hypselect) == 10)
# check that the rownames are in the order we listed above before modifying
rownames(hypselect)
# okay, so we can do this:
mknm <- cbind(rownames(hypselect), c(rep(c(" (VMH)", " (ARC)"), 4), " (WM)", " (WM.2)"))
rownames(hypselect) <- unlist(lapply(c(1:10), function(x) {
    paste0(mknm[x, 1], mknm[x, 2])
}))

#

# save plot
cairo_pdf("plots/Manuscript/2B-Marker violins.pdf", height = 2.5, width = 3.75)

plotExpression(hypselect, features = rownames(hypselect), colour_by = "label", x = "label", scales = "free_y", ncol = 2, point_fun = function(...) geom_quasirandom_rast(..., width = 0.4, groupOnX = TRUE, bandwidth = 1, size = 0.2, alpha = 0.8, dev = "cairo", raster.dpi = 600)) +
    scale_color_manual(values = pal) +
    guides(color = "none") +
    scale_y_continuous(n.breaks = 3) +
    theme(axis.title = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.x = element_blank(), axis.text.x = element_blank(), strip.background = element_blank(), strip.text = element_text(size = 9))

dev.off()

rm(hypselect)
```

# width of 7 --> four 1.75x1.5 spotplots with inch of spacing total between

### Fig 2C: GHRH in V12D05-350_D1 ###
```{r}
hypselect <- hyp2[, hyp2$sample_id == "V12D05-350_D1"]
# change rownames to gene symbols
rownames(hypselect) <- rowData(hypselect)$gene_name

pdf("plots/Manuscript/2C-GHRH in V12D05-350_D1.pdf", height = 1.75, width = 1.5)
vis_gene(hypselect, geneid = "GHRH", spatial = T, auto_crop = T, point_size = 0.525, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA, viridis = FALSE) +
    ggtitle("GHRH V12D05-350_D1") +
    # counter vis_gene's plot margins to use the allocated area
    theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 8, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_text(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.4, 0, 0, 0, unit = "in"),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, -0.025, 0, -.45, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
dev.off()

rm(hypselect)
```

Fig 2D: POMC in V12D05-348_C1
```{r}
hypselect <- hyp2[, hyp2$sample_id == "V12D05-348_C1"]
# change rownames to gene symbols
rownames(hypselect) <- rowData(hypselect)$gene_name

pdf("plots/Manuscript/2D-POMC in V12D05-348_C1.pdf", height = 1.75, width = 1.5)
vis_gene(hypselect, geneid = "POMC", spatial = T, auto_crop = T, point_size = 0.525, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA, viridis = FALSE) +
    ggtitle("POMC V12D05-348_C1") +
    # counter vis_gene's plot margins to use the allocated area
    theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 8, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_text(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.4, 0, 0, 0, unit = "in"),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, -0.025, 0, -.45, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
dev.off()

rm(hypselect)
```

Fig 2E: FEZF1 in V12D07-075_D1
```{r}
hypselect <- hyp2[, hyp2$sample_id == "V12D07-075_D1"]
# change rownames to gene symbols
rownames(hypselect) <- rowData(hypselect)$gene_name

pdf("plots/Manuscript/2E-FEZF1 in V12D05-075_D1.pdf", height = 1.75, width = 1.5)
vis_gene(hypselect, geneid = "FEZF1", spatial = T, auto_crop = T, point_size = 0.525, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA, viridis = FALSE) +
    ggtitle("FEZF1 V12D07-075_D1") +
    # counter vis_gene's plot margins to use the allocated area
    theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 8, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_text(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.4, 0, 0, 0, unit = "in"),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, -0.025, 0, -.45, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
dev.off()

rm(hypselect)
```


Fig 2F: KISS1R in V12Y31-080_A1
```{r}
hypselect <- hyp2[, hyp2$sample_id == "V12Y31-080_A1"]
# change rownames to gene symbols
rownames(hypselect) <- rowData(hypselect)$gene_name

pdf("plots/Manuscript/2F-KISS1R in V12Y31-080_A1.pdf", height = 1.75, width = 1.5)
vis_gene(hypselect, geneid = "KISS1R", spatial = T, auto_crop = T, point_size = 0.525, alpha=0.8, pdf_file = NULL, return_plots = T, na_color = NA, viridis = FALSE) +
    ggtitle("KISS1R V12Y31-080_A1") +
    # counter vis_gene's plot margins to use the allocated area
    theme(
        axis.text = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        legend.text = element_text(size = 8, margin = margin(-0.1, 0, 0, 0, "in")),
        legend.title = element_blank(),
        plot.title = element_text(
            size = 9,
            hjust = 0.5,
            margin = margin(0, 0, 0, 0, "in")
        ),
        legend.key.size = unit(0.125, "in"),
        legend.margin = margin(-0.4, 0, 0, 0, unit = "in"),
        axis.ticks = element_blank(),
        plot.margin = margin(0.05, -0.025, 0, -.45, unit = "in"),
        legend.position = "bottom",
        legend.justification = 0.5,
        legend.spacing.x = unit(0, "in"),
        legend.box.margin = margin(0, 0, 0, 0, "in"),
        legend.background = element_blank()
    )
dev.off()
# rm(hypselect)
```

Fig 2G: markers as heatmap
# load spatialLIBD marker results, take top 6 per cluster of interest + top 2 per other = 6*6 + 2*9 = 54
### Get marker genes to be plotted -- top AUCs per cluster
```{r}
slibdmk <- readRDS("data/07-Marker_genes/01a-BSpace_allruns_top100_1vsAll-markers_perdomain_perrun.RDS")
# get NONcollapsed clusters
slibdmk <- slibdmk[cluster_set == "k15_HARMONYlmbna_nnsvg10_60kiter"]

# apply matching label names
slibdmk[cluster == "X11", cluster := "VMH.1"]
slibdmk[cluster == "X12", cluster := "VMH.2"]
slibdmk[cluster == "X3", cluster := "ARC.1"]
slibdmk[cluster == "X5", cluster := "ARC.2"]
slibdmk[cluster == "X1", cluster := "WM.1"]
slibdmk[cluster == "X9", cluster := "WM.2"]

pltg <- lapply(unique(slibdmk$cluster), function(x) {
    tmp <- copy(slibdmk[slibdmk$cluster == x])
    setorderv(tmp, "mean.AUC", -1)
    return(tmp[1:6, gene_name])
})
names(pltg) <- unique(slibdmk$cluster)

# reduce other clust lists to top 2
pltg[!(names(pltg) %in% c("VMH.1", "VMH.2", "ARC.1", "ARC.2", "WM.1", "WM.2"))] <- lapply(pltg[!(names(pltg) %in% c("VMH.1", "VMH.2", "ARC.1", "ARC.2", "WM.1", "WM.2"))], function(x) {
    return(x[1:2])
})

# combine into one list
pltg <- unique(unlist(pltg, recursive = F))

rm(slibdmk)
gc(full = T)
```

### 2G, continued: Z-scale logcounts, make a subsetted spe with the columns already ordered by cluster since complexheatmap was giving issues with this, make plot.
```{r}
# z scale rows, from tony https://github.com/LieberInstitute/spatial_DG_lifespan/blob/9419eb91453cda1df494d93dc91156d819042c66/code/Pseudobulk/top_infant_DE_heatmap_enrichment.R#L106

scale_rows <- function(x) {
    m <- apply(x, 1, mean, na.rm = T)
    s <- apply(x, 1, sd, na.rm = T)
    return((x - m) / s)
}

### extract log counts per marker, z-normalize


### getting a problem here trying to use column_order=levels(hyp2$label),so instead, make a subsetted spe with reordering of the columns to be in the order of the domains we want plotted, then disable column clustering in the heatmap. (thatll make the plotting quicker too).
i <- 1
for (i in c(1:length(unique(hyp2$label)))) {
    curlev <- levels(hyp2$label)[i]
    if (i == 1) {
        cnames <- colnames(hyp2[, hyp2$label == curlev])
    } else {
        cnames <- c(cnames, colnames(hyp2[, hyp2$label == curlev]))
    }
}
rm(i, curlev)


rownames(hyp2) <- rowData(hyp2)$gene_name
pltdat <- hyp2[pltg, cnames]
pltdat2 <- scale_rows(as.matrix(logcounts(pltdat)))

rm(cnames)
gc(full = T)

### pre-make
htmap <- ComplexHeatmap::Heatmap(
    pltdat2,
    top_annotation = HeatmapAnnotation(
        Domain = pltdat$label,
        col = list(Domain = pal),
        gap = 0,
        gp = gpar(alpha = 1)
    ),
    show_column_names = FALSE,
    show_column_dend = FALSE,
    show_row_dend = FALSE,
    show_row_names = T,
    col = colorRamp2(
        breaks = c(-2, 0, 2),
        colors = c("blue", "white", "orange")
    ),
    cluster_columns = F,
    heatmap_legend_param = list(
        title = "Z scored log counts",
        direction = "horizontal",
        title_position = "topcenter"
    ),
    row_names_gp = gpar(fontsize = 6),
    raster_device = "CairoTIFF",
    raster_quality = 10
)

### save plot
CairoPDF("plots/Manuscript/2G-Marker_heatmap.pdf", height = 4, width = 3.5)
draw(htmap,
    heatmap_legend_side = "bottom",
    annotation_legend_side = "right",
    legend_grouping = "original"
)

dev.off()
```

## old heatmap code
```{r}
### add axis labels: https://stackoverflow.com/questions/17538830/x-axis-and-y-axis-labels-in-pheatmap-in-r
library(grid)
# pdf("plots/Manuscript/2G-Marker_heatmap.pdf",height=6,width=3)
setHook("grid.newpage", function() pushViewport(viewport(x = 1, y = 1, width = 0.95, height = 0.97, name = "vp", just = c("right", "top"))), action = "prepend")
pheatmap::pheatmap(pltmk, cluster_rows = T, cluster_cols = F, treeheight_row = 0, fontsize_row = 8, fontsize_col = 10, fontsize = 8)
setHook("grid.newpage", NULL, "replace")
grid.text("Visium Domain (k=15)", y = -0.005, gp = gpar(fontsize = 11))
grid.text("Gene Symbol", x = -0.02, rot = 90, gp = gpar(fontsize = 11))


dev.off()
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] grid      stats4    stats     graphics  grDevices utils     datasets 
[8] methods   base     

other attached packages:
 [1] colorRamp2_0.1.0            ComplexHeatmap_2.18.0      
 [3] Polychrome_1.5.1            colorout_1.3-0.1           
 [5] scater_1.30.1               scuttle_1.12.0             
 [7] spatialLIBD_1.14.1          SpatialExperiment_1.12.0   
 [9] SingleCellExperiment_1.24.0 SummarizedExperiment_1.32.0
[11] Biobase_2.62.0              GenomicRanges_1.54.1       
[13] MatrixGenerics_1.14.0       matrixStats_1.2.0          
[15] gridExtra_2.3               ggplot2_3.4.4              
[17] Cairo_1.6-2                 ggrastr_1.0.2              
[19] Biostrings_2.70.2           GenomeInfoDb_1.38.5        
[21] XVector_0.42.0              IRanges_2.36.0             
[23] S4Vectors_0.40.2            BiocGenerics_0.48.1        
[25] data.table_1.14.10          rlang_1.1.3                

loaded via a namespace (and not attached):
  [1] later_1.3.2                   BiocIO_1.12.0                
  [3] bitops_1.0-7                  filelock_1.0.3               
  [5] fields_15.2                   R.oo_1.26.0                  
  [7] tibble_3.2.1                  XML_3.99-0.16.1              
  [9] lifecycle_1.0.4               edgeR_4.0.14                 
 [11] doParallel_1.0.17             rprojroot_2.0.4              
 [13] lattice_0.22-5                magrittr_2.0.3               
 [15] limma_3.58.1                  plotly_4.10.4                
 [17] sass_0.4.8                    rmarkdown_2.25               
 [19] jquerylib_0.1.4               yaml_2.3.8                   
 [21] httpuv_1.6.13                 spam_2.10-0                  
 [23] sessioninfo_1.2.2             cowplot_1.1.3                
 [25] DBI_1.2.1                     RColorBrewer_1.1-3           
 [27] golem_0.4.1                   maps_3.4.2                   
 [29] abind_1.4-5                   zlibbioc_1.48.0              
 [31] R.cache_0.16.0                R.utils_2.12.3               
 [33] purrr_1.0.2                   RCurl_1.98-1.14              
 [35] styler_1.10.2                 rappdirs_0.3.3               
 [37] circlize_0.4.15               GenomeInfoDbData_1.2.11      
 [39] ggrepel_0.9.5                 irlba_2.3.5.1                
 [41] pheatmap_1.0.12               DelayedMatrixStats_1.24.0    
 [43] codetools_0.2-19              DelayedArray_0.28.0          
 [45] DT_0.31                       shape_1.4.6                  
 [47] tidyselect_1.2.0              farver_2.1.1                 
 [49] ScaledMatrix_1.10.0           viridis_0.6.5                
 [51] shinyWidgets_0.8.1            BiocFileCache_2.10.1         
 [53] GenomicAlignments_1.38.2      jsonlite_1.8.8               
 [55] GetoptLong_1.0.5              BiocNeighbors_1.20.2         
 [57] ellipsis_0.3.2                iterators_1.0.14             
 [59] foreach_1.5.2                 tools_4.3.2                  
 [61] Rcpp_1.0.12                   glue_1.7.0                   
 [63] SparseArray_1.2.3             xfun_0.41                    
 [65] here_1.0.1                    usethis_2.2.2                
 [67] dplyr_1.1.4                   withr_3.0.0                  
 [69] BiocManager_1.30.22           fastmap_1.1.1                
 [71] fansi_1.0.6                   digest_0.6.34                
 [73] rsvd_1.0.5                    R6_2.5.1                     
 [75] mime_0.12                     colorspace_2.1-0             
 [77] RSQLite_2.3.5                 R.methodsS3_1.8.2            
 [79] config_0.3.2                  utf8_1.2.4                   
 [81] tidyr_1.3.1                   generics_0.1.3               
 [83] rtracklayer_1.62.0            httr_1.4.7                   
 [85] htmlwidgets_1.6.4             S4Arrays_1.2.0               
 [87] scatterplot3d_0.3-44          pkgconfig_2.0.3              
 [89] gtable_0.3.4                  rsconnect_1.2.0              
 [91] blob_1.2.4                    htmltools_0.5.7              
 [93] dotCall64_1.1-1               clue_0.3-65                  
 [95] scales_1.3.0                  png_0.1-8                    
 [97] attempt_0.3.1                 biocthis_1.12.0              
 [99] knitr_1.45                    rstudioapi_0.15.0            
[101] rjson_0.2.21                  curl_5.2.0                   
[103] GlobalOptions_0.1.2           cachem_1.0.8                 
[105] BiocVersion_3.18.1            parallel_4.3.2               
[107] vipor_0.4.7                   AnnotationDbi_1.64.1         
[109] restfulr_0.0.15               pillar_1.9.0                 
[111] vctrs_0.6.5                   promises_1.2.1               
[113] BiocSingular_1.18.0           dbplyr_2.4.0                 
[115] beachmat_2.18.0               cluster_2.1.6                
[117] xtable_1.8-4                  beeswarm_0.4.0               
[119] paletteer_1.6.0               evaluate_0.23                
[121] magick_2.8.2                  cli_3.6.2                    
[123] locfit_1.5-9.8                compiler_4.3.2               
[125] Rsamtools_2.18.0              crayon_1.5.2                 
[127] labeling_0.4.3                rematch2_2.1.2               
[129] fs_1.6.3                      ggbeeswarm_0.7.2             
[131] viridisLite_0.4.2             BiocParallel_1.36.0          
[133] munsell_0.5.0                 tiff_0.1-12                  
[135] lazyeval_0.2.2                Matrix_1.6-5                 
[137] ExperimentHub_2.10.0          benchmarkme_1.0.8            
[139] sparseMatrixStats_1.14.0      bit64_4.0.5                  
[141] KEGGREST_1.42.0               statmod_1.5.0                
[143] shiny_1.8.0                   interactiveDisplayBase_1.40.0
[145] AnnotationHub_3.10.0          memoise_2.0.1                
[147] bslib_0.6.1                   benchmarkmeData_1.0.4        
[149] bit_4.0.5                    
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-02-05
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────
 package                * version     date (UTC) lib source
 abind                    1.4-5       2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1      2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0      2023-10-24 [1] Bioconductor
 attempt                  0.3.1       2020-05-03 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0      2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0       2021-06-01 [1] CRAN (R 4.3.2)
 benchmarkme              1.0.8       2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4       2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0      2023-10-24 [1] Bioconductor
 BiocFileCache            2.10.1      2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1      2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0      2023-10-26 [1] Bioconductor
 BiocManager              1.30.22     2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.2      2024-01-07 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel             1.36.0      2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0      2023-10-24 [1] Bioconductor
 biocthis                 1.12.0      2023-10-26 [1] Bioconductor
 BiocVersion              3.18.1      2023-11-18 [1] Bioconductor 3.18 (R 4.3.2)
 Biostrings             * 2.70.2      2024-01-28 [1] Bioconductor 3.18 (R 4.3.2)
 bit                      4.0.5       2022-11-15 [1] CRAN (R 4.3.2)
 bit64                    4.0.5       2020-08-30 [1] CRAN (R 4.3.2)
 bitops                   1.0-7       2021-04-24 [1] CRAN (R 4.3.2)
 blob                     1.2.4       2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.6.1       2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8       2023-05-01 [1] CRAN (R 4.3.2)
 Cairo                  * 1.6-2       2023-11-28 [1] CRAN (R 4.3.1)
 circlize                 0.4.15      2022-05-10 [1] CRAN (R 4.3.0)
 cli                      3.6.2       2023-12-11 [1] CRAN (R 4.3.2)
 clue                     0.3-65      2023-09-23 [1] CRAN (R 4.3.2)
 cluster                  2.1.6       2023-12-01 [1] CRAN (R 4.3.2)
 codetools                0.2-19      2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0.1     2024-01-11 [1] local
 colorRamp2             * 0.1.0       2022-12-21 [1] CRAN (R 4.3.0)
 colorspace               2.1-0       2023-01-23 [1] CRAN (R 4.3.2)
 ComplexHeatmap         * 2.18.0      2023-10-26 [1] Bioconductor
 config                   0.3.2       2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.3       2024-01-22 [1] CRAN (R 4.3.1)
 crayon                   1.5.2       2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0       2023-12-08 [1] CRAN (R 4.3.2)
 data.table             * 1.14.10     2023-12-08 [1] CRAN (R 4.3.2)
 DBI                      1.2.1       2024-01-12 [1] CRAN (R 4.3.2)
 dbplyr                   2.4.0       2023-10-26 [1] CRAN (R 4.3.1)
 DelayedArray             0.28.0      2023-10-24 [1] Bioconductor
 DelayedMatrixStats       1.24.0      2023-11-06 [1] Bioconductor
 digest                   0.6.34      2024-01-11 [1] CRAN (R 4.3.2)
 doParallel               1.0.17      2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1       2023-11-28 [1] CRAN (R 4.3.2)
 dplyr                    1.1.4       2023-11-17 [1] CRAN (R 4.3.2)
 DT                       0.31        2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.14      2024-01-29 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2       2021-04-29 [1] CRAN (R 4.3.2)
 evaluate                 0.23        2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0      2023-10-26 [1] Bioconductor
 fansi                    1.0.6       2023-12-08 [1] CRAN (R 4.3.2)
 farver                   2.1.1       2022-07-06 [1] CRAN (R 4.3.2)
 fastmap                  1.1.1       2023-02-24 [1] CRAN (R 4.3.2)
 fields                   15.2        2023-08-17 [1] CRAN (R 4.3.2)
 filelock                 1.0.3       2023-12-11 [1] CRAN (R 4.3.2)
 foreach                  1.5.2       2022-02-02 [1] CRAN (R 4.3.0)
 fs                       1.6.3       2023-07-20 [1] CRAN (R 4.3.0)
 generics                 0.1.3       2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.5      2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData         1.2.11      2024-01-11 [1] Bioconductor
 GenomicAlignments        1.38.2      2024-01-16 [1] Bioconductor 3.18 (R 4.3.2)
 GenomicRanges          * 1.54.1      2023-10-29 [1] Bioconductor
 GetoptLong               1.0.5       2020-12-15 [1] CRAN (R 4.3.0)
 ggbeeswarm               0.7.2       2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4       2023-10-12 [1] CRAN (R 4.3.1)
 ggrastr                * 1.0.2       2023-06-01 [1] CRAN (R 4.3.0)
 ggrepel                  0.9.5       2024-01-10 [1] CRAN (R 4.3.2)
 GlobalOptions            0.1.2       2020-06-10 [1] CRAN (R 4.3.0)
 glue                     1.7.0       2024-01-09 [1] CRAN (R 4.3.2)
 golem                    0.4.1       2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3         2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4       2023-08-21 [1] CRAN (R 4.3.0)
 here                     1.0.1       2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7       2023-11-03 [1] CRAN (R 4.3.2)
 htmlwidgets              1.6.4       2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13      2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7       2023-08-15 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0      2023-10-26 [1] Bioconductor
 IRanges                * 2.36.0      2023-10-24 [1] Bioconductor
 irlba                    2.3.5.1     2022-10-03 [1] CRAN (R 4.3.2)
 iterators                1.0.14      2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4       2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8       2023-12-04 [1] CRAN (R 4.3.2)
 KEGGREST                 1.42.0      2023-10-26 [1] Bioconductor
 knitr                    1.45        2023-10-30 [1] CRAN (R 4.3.2)
 labeling                 0.4.3       2023-08-29 [1] CRAN (R 4.3.0)
 later                    1.3.2       2023-12-06 [1] CRAN (R 4.3.2)
 lattice                  0.22-5      2023-10-24 [1] CRAN (R 4.3.2)
 lazyeval                 0.2.2       2019-03-15 [1] CRAN (R 4.3.2)
 lifecycle                1.0.4       2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1      2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8     2023-06-11 [1] CRAN (R 4.3.2)
 magick                   2.8.2       2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3       2022-03-30 [1] CRAN (R 4.3.2)
 maps                     3.4.2       2023-12-15 [1] CRAN (R 4.3.1)
 Matrix                   1.6-5       2024-01-11 [1] CRAN (R 4.3.2)
 MatrixGenerics         * 1.14.0      2023-10-26 [1] Bioconductor
 matrixStats            * 1.2.0       2023-12-11 [1] CRAN (R 4.3.2)
 memoise                  2.0.1       2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12        2021-09-28 [1] CRAN (R 4.3.2)
 munsell                  0.5.0       2018-06-12 [1] CRAN (R 4.3.0)
 paletteer                1.6.0       2024-01-21 [1] CRAN (R 4.3.1)
 pheatmap                 1.0.12      2019-01-04 [1] CRAN (R 4.3.0)
 pillar                   1.9.0       2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3       2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.4      2024-01-13 [1] CRAN (R 4.3.1)
 png                      0.1-8       2022-11-29 [1] CRAN (R 4.3.2)
 Polychrome             * 1.5.1       2022-05-03 [1] CRAN (R 4.3.0)
 promises                 1.2.1       2023-08-10 [1] CRAN (R 4.3.2)
 purrr                    1.0.2       2023-08-10 [1] CRAN (R 4.3.2)
 R.cache                  0.16.0      2022-07-21 [1] CRAN (R 4.3.0)
 R.methodsS3              1.8.2       2022-06-13 [1] CRAN (R 4.3.0)
 R.oo                     1.26.0      2024-01-24 [1] CRAN (R 4.3.1)
 R.utils                  2.12.3      2023-11-18 [1] CRAN (R 4.3.1)
 R6                       2.5.1       2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3       2021-01-31 [1] CRAN (R 4.3.2)
 RColorBrewer             1.1-3       2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.12      2024-01-09 [1] CRAN (R 4.3.2)
 RCurl                    1.98-1.14   2024-01-09 [1] CRAN (R 4.3.1)
 rematch2                 2.1.2       2020-05-01 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15      2022-06-16 [1] CRAN (R 4.3.2)
 rjson                    0.2.21      2022-01-09 [1] CRAN (R 4.3.2)
 rlang                  * 1.1.3       2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown                2.25        2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot                2.0.4       2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0      2023-10-24 [1] Bioconductor
 rsconnect                1.2.0       2023-12-15 [1] CRAN (R 4.3.1)
 RSQLite                  2.3.5       2024-01-21 [1] CRAN (R 4.3.2)
 rstudioapi               0.15.0      2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5       2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0      2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0       2023-10-24 [1] Bioconductor
 S4Vectors              * 0.40.2      2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8       2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0      2023-11-06 [1] Bioconductor
 scales                   1.3.0       2023-11-28 [1] CRAN (R 4.3.2)
 scater                 * 1.30.1      2023-11-16 [1] Bioconductor
 scatterplot3d            0.3-44      2023-05-05 [1] CRAN (R 4.3.0)
 scuttle                * 1.12.0      2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2       2021-12-06 [1] CRAN (R 4.3.0)
 shape                    1.4.6       2021-05-19 [1] CRAN (R 4.3.0)
 shiny                    1.8.0       2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.1       2024-01-10 [1] CRAN (R 4.3.1)
 SingleCellExperiment   * 1.24.0      2023-11-06 [1] Bioconductor
 spam                     2.10-0      2023-10-23 [1] CRAN (R 4.3.2)
 SparseArray              1.2.3       2023-12-25 [1] Bioconductor 3.18 (R 4.3.2)
 sparseMatrixStats        1.14.0      2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0      2023-10-26 [1] Bioconductor
 spatialLIBD            * 1.14.1      2023-11-30 [1] Bioconductor 3.18 (R 4.3.2)
 statmod                  1.5.0       2023-01-06 [1] CRAN (R 4.3.2)
 styler                   1.10.2      2023-08-29 [1] CRAN (R 4.3.0)
 SummarizedExperiment   * 1.32.0      2023-11-06 [1] Bioconductor
 tibble                   3.2.1       2023-03-20 [1] CRAN (R 4.3.2)
 tidyr                    1.3.1       2024-01-24 [1] CRAN (R 4.3.1)
 tidyselect               1.2.0       2022-10-10 [1] CRAN (R 4.3.0)
 tiff                     0.1-12      2023-11-28 [1] CRAN (R 4.3.1)
 usethis                  2.2.2       2023-07-06 [1] CRAN (R 4.3.0)
 utf8                     1.2.4       2023-10-22 [1] CRAN (R 4.3.2)
 vctrs                    0.6.5       2023-12-01 [1] CRAN (R 4.3.2)
 vipor                    0.4.7       2023-12-18 [1] CRAN (R 4.3.1)
 viridis                  0.6.5       2024-01-29 [1] CRAN (R 4.3.2)
 viridisLite              0.4.2       2023-05-02 [1] CRAN (R 4.3.0)
 withr                    3.0.0       2024-01-16 [1] CRAN (R 4.3.1)
 xfun                     0.41        2023-11-01 [1] CRAN (R 4.3.2)
 XML                      3.99-0.16.1 2024-01-22 [1] CRAN (R 4.3.1)
 xtable                   1.8-4       2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0      2023-10-24 [1] Bioconductor
 yaml                     2.3.8       2023-12-11 [1] CRAN (R 4.3.2)
 zlibbioc                 1.48.0      2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

───────────────────────────────────────────────────────────────────────────
