---
title: "Fig 4-Sex DE"
author: "Bernard Mulvey"
date: "2023-12-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table) # Preferred data manipulation package
library(Biostrings) # General
library(ggrastr) # render plots with bajillions of points as raster so they can be manipulated in eg Illustrator
library(Cairo) # pdf device for ^ to properly save
library(ggplot2) # Dependency for several plotting functions
library(gridExtra)
library(SpatialExperiment) # Dependency
library(escheR) # spotplots with domain outlines
library(ggrepel) # for plotting labels on a volcano cleanly
require(colorout) # Utility for RStudio
ColorOut() 

# code reformatting in Rstudio
options("styler.addins_style_transformer" = "biocthis::bioc_style()") 

# set plotting defaults for ggplot
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 10), axis.title.x = element_text(size = 11), axis.text.y = element_text(size = 10), axis.title.y = element_text(size =11), plot.title = element_text(size = 12,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=9), legend.title = element_text(size=10,hjust=0.5)))
```

### Setup: load data, cluster assignments, and rename from generic to appropriate region (VMH/ARC/WM)
```{r}
hyp2 <- readRDS("data/03-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

bscl <- fread("data/06-BayesSpace/02-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt")

setnames(bscl,c("rn","cl"))
bscl[,cl:=paste0("X",cl)]
bscl <- DataFrame(bscl)
rownames(bscl) <- bscl$rn
bscl <- bscl[colnames(hyp2),]

colLabels(hyp2) <- factor(bscl$cl,levels=paste0("X",c(1:15)))

### relabel as VMH/ARC/WM for spotplottings
tmpcd <- as.data.table(colData(hyp2),keep.rownames=T)
tmpcd[label %in% c("X11","X12"),label:="VMH"]
tmpcd[label %in% c("X3","X5"),label:="ARC"]
tmpcd[label %in% c("X1","X9"),label:="WM"]

tmpcd <- DataFrame(tmpcd)
rownames(tmpcd) <- tmpcd$rn
tmpcd <- tmpcd[colnames(hyp2),]
colData(hyp2) <- tmpcd
hyp2$label <- factor(hyp2$label,levels=c("VMH","ARC","WM",paste0("X",c(2,4,6:8,10,13:15))))

### also define cluster tables that only have VMH or ARC and otehrs as NA for escher plotting outlines specific to VMH/ARC
bscl.vmhonly <- as.data.table(bscl,keep.rownames=T)
bscl.arconly <- as.data.table(bscl,keep.rownames=T)
bscl.vmharc <- as.data.table(bscl,keep.rownames=T)

bscl.vmhonly <- bscl.vmhonly[cl %in% c("X11","X12"),cl2:="VMH"]
bscl.arconly <- bscl.arconly[cl %in% c("X3","X5"),cl2:="ARC"]
bscl.vmharc <- bscl.vmharc[cl %in% c("X11","X12"),cl2:="VMH"]
bscl.vmharc <- bscl.vmharc[cl %in% c("X3","X5"),cl2:="ARC"]

rm(tmpcd)
```


4A-4C: Sex DE volcanos for ARC VMH WM
```{r}
sexderes <- fread("data/09-Sex DE/01c-dreamlet-ranfx_nnsvg10-HmnylmbNA-BS-15-VMHARCclpsd.txt")

sexderes <- sexderes[!(chromosome_name %in% c("chrX","chrY","chrMT"))]

vmhde <- sexderes[assay=="VMH"]
vmhde[,genelabel:=as.character(NA)]
vmhde[gene_name %in% c("CACNA1D", "ANKRD17", "PTGR1", "HACE1"),genelabel:=gene_name]

vmhvol <- ggplot(vmhde,aes(y=-log10(P.Value),x=logFC))+
    geom_point(size=0.6,col="#001859")+
    geom_hline(yintercept=-log10(0.05),linetype="dashed",color="black")+
    geom_hline(yintercept=-log10(max(sexderes[adj.P.Val<0.05&assay=="VMH",P.Value])),linetype="dashed",color="red")+
    geom_text_repel(aes(y=-log10(P.Value),x=logFC,label = ifelse(is.na(genelabel), "",gene_name)),nudge_y =-0.2,nudge_x=0.5,min.segment.length = 0)+
    xlab("VMH logFC\n(Male vs. Female)")+
    ylab("-log10 (p-value)")+
    xlim(-2.25,2.25)+
    theme(axis.title.x=element_text(size=11),axis.text.x = element_text(size=10),axis.title.y = element_text(size=11),axis.text.y=element_text(size=10))

#dl.jt.res[assay=="VMH"&-log10(P.Value)>12] # label PTGR1, ANKRD17, NAB2, BCR

arcde <- sexderes[assay=="ARC"]
arcde[,genelabel:=as.character(NA)]
arcde[gene_name %in% c("PENK", "IGFLR1", "IGFBP6","CAPS2"),genelabel:=gene_name]

arcvol <- ggplot(arcde,aes(y=-log10(P.Value),x=logFC))+
    geom_point(size=0.6,col="#058a00")+
    geom_hline(yintercept=-log10(0.05),linetype="dashed",color="black")+
    geom_hline(yintercept=-log10(max(sexderes[adj.P.Val<0.05&assay=="ARC",P.Value])),linetype="dashed",color="red")+
    geom_text_repel(aes(y=-log10(P.Value),x=logFC,label = ifelse(is.na(genelabel), "",gene_name)),nudge_y = 6,nudge_x=sign(arcde$logFC)*0.65,min.segment.length = 0)+
    xlab("ARC logFC\n(Male vs. Female)")+
    ylab("-log10 (p-value)")+
    xlim(-3,3)+
    theme(axis.title.x=element_text(size=11),axis.text.x = element_text(size=10),axis.title.y = element_text(size=11),axis.text.y=element_text(size=10))

arcvmh <- merge.data.table(sexderes[assay=="VMH",.(ID,logFC)],sexderes[assay=="ARC",.(ID,logFC)],by="ID",suffixes=c("_VMH","_ARC"))

cor(arcvmh$logFC_ARC,arcvmh$logFC_VMH,method="pearson")

arcvmhcor <- ggplot(arcvmh,aes(x=logFC_VMH,y=logFC_ARC))+
    geom_point(size=0.9,col="#157e82")+
    geom_abline(slope=1,intercept=0,linetype="dashed",col="black")+
    geom_text(aes(x=-1.2,y=2.6),label="Pearson Coef=0.512",size=2.5)+
    xlim(-2.6,2.6)+
    ylim(-2.6,2.6)+
    xlab("logFC VMH")+
    ylab("logFC ARC")+
    theme(plot.title = element_text(hjust = 0.5))+
    theme(axis.title.x=element_text(size=11),axis.text.x = element_text(size=10),axis.title.y = element_text(size=11),axis.text.y=element_text(size=10),title = element_text(size=11))

arcvmhcor
dev.off()
plts <- list(vmhvol,arcvol,arcvmhcor)

pdf("plots/Manuscript/4A-C-ARC, VMH volcanos and lfc correls.pdf",height=2.5,width=7)
do.call("grid.arrange",c(plts,ncol=3))
dev.off()


rm(arcvmhcor,vmhvol,arcvol,plts,arcvmh)
```

## Fig 4D-E/F-G: sex DEG spotplots
D-E: PTGR1 in VMH (2E-9)
F-G: RXRG in ARC (FDR 0.09)
```{r}
ptgrM <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(ptgrM) <- rowData(ptgrM)$gene_name
colData(ptgrM)$`log counts` <- logcounts(ptgrM)["PTGR1",]
ptgrlab <- copy(bscl.vmharc)[rn %in% colnames(ptgrM)]
ptgrlab <- DataFrame(ptgrlab)
rownames(ptgrlab) <- ptgrlab$rn


ptgrM$Domain <- factor(ptgrlab$cl2,levels=c("VMH","ARC"))
ptgrM <- ptgrM[,ptgrM$Domain %in% c("VMH","ARC")]
### make spotplots
p <- make_escheR(ptgrM)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Domain",stroke=0.15,point_size = 0.6)

pdf("plots/Manuscript/4D-PTGR1_VMHsexDE_F_malesamp.pdf",height=1.75,width=2)
p+
    scale_color_manual(values=c("purple","darkgreen"),na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("PTGR1 (V12Y31-\n080_A1 - Male)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_text(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = unit(0.125,"in"),legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
    #theme(plot.title.position = "plot",plot.title = element_text(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = unit(0.125,"in"),legend.title=element_text(size=9),plot.margin = margin(-0.3,0,-0.1,0,unit = "in"))
dev.off()

rm(ptgrM,p,ptgrlab)

### 
ptgrF <- hyp2[,hyp2$sample_id=="V13M13-362_D1"]
rownames(ptgrF) <- rowData(ptgrF)$gene_name
colData(ptgrF)$`log counts` <- logcounts(ptgrF)["PTGR1",]
ptgrlab <- copy(bscl.vmharc)[rn %in% colnames(ptgrF)]
ptgrlab <- DataFrame(ptgrlab)
rownames(ptgrlab) <- ptgrlab$rn

ptgrF$Domain <- factor(ptgrlab$cl2,levels=c("VMH","ARC"))
ptgrF <- ptgrF[,ptgrF$Domain %in% c("VMH","ARC")]
### make spotplots
p <- make_escheR(ptgrF)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Domain",stroke=0.15,point_size = 0.6)

pdf("plots/Manuscript/4E-PTGR1_VMHsexDE_F_femsamp.pdf",height=1.75,width=2)
p+
    scale_color_manual(values=c("purple","darkgreen"),na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("PTGR1 (V13M13-\n362_D1 - Female)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_text(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = unit(0.125,"in"),legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(ptgrF.p,ptgrF,p,ptgrlab)
```

```{r}
rxrM <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(rxrM) <- rowData(rxrM)$gene_name
colData(rxrM)$`log counts` <- logcounts(rxrM)["RXRG",]
rxrlab <- copy(bscl.vmharc)[rn %in% colnames(rxrM)]
rxrlab <- DataFrame(rxrlab)
rownames(rxrlab) <- rxrlab$rn

rxrM$Domain <- factor(rxrlab$cl2,levels=c("VMH","ARC"))
rxrM <- rxrM[,rxrM$Domain %in% c("VMH","ARC")]
### make spotplots
p <- make_escheR(rxrM)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Domain",stroke=0.15,point_size = 0.6)

pdf("plots/Manuscript/4F-RXRG_ARCsexDE_M_malesamp.pdf",height=1.75,width=2)
p+
    scale_color_manual(values=c("purple","darkgreen"),na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("RXRG (V12Y31-\n080_A1 - Male)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_text(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = unit(0.125,"in"),legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(rxrM,p,rxrlab)

### 
rxrF <- hyp2[,hyp2$sample_id=="V12D05-350_C1"]
rownames(rxrF) <- rowData(rxrF)$gene_name
colData(rxrF)$`log counts` <- logcounts(rxrF)["RXRG",]
rxrlab <- copy(bscl.vmharc)[rn %in% colnames(rxrF)]
rxrlab <- DataFrame(rxrlab)
rownames(rxrlab) <- rxrlab$rn

rxrF$Domain <- factor(rxrlab$cl2,levels=c("VMH","ARC"))
rxrF <- rxrF[,rxrF$Domain %in% c("VMH","ARC")]
### make spotplots
p <- make_escheR(rxrF)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Domain",stroke=0.15,point_size = 0.6)

pdf("plots/Manuscript/4G-RXRG_ARCsexDE_M_femsamp.pdf",height=1.75,width=2)
p+
    scale_color_manual(values=c("purple","darkgreen"),na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("RXRG (V12D05-\n350_C1 - Female)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_text(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = unit(0.125,"in"),legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(rxrF,p,rxrlab)
```

Shared legend: 4D thru 4G
```{r}
### custom function to extrct legend from one plot to extend to all ###
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}

### run the first part of the code used to generate the marker panels but with the fill legend changed to be the domains factored in the same order
gabre <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(gabre) <- rowData(gabre)$gene_name
colData(gabre)$`log counts` <- logcounts(gabre)["GABRE",]
gabrelab <- copy(bscl.vmharc)[rn %in% colnames(gabre)]
gabrelab <- DataFrame(gabrelab)
rownames(gabrelab) <- gabrelab$rn

gabre$Domain <- factor(gabrelab$cl2,levels=c("VMH","ARC"))
### make a ggplot of nothing in particular, we just want the legend
p <- ggplot(as.data.table(colData(gabre))[!is.na(Domain)],aes(x=sum_umi,y=expr_chrM_ratio,color=Domain))+
    geom_point()+
    scale_color_manual(values=c("purple","darkgreen"),na.value = NA)+
    theme(plot.title.position = "plot",plot.title = element_text(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = unit(0.125,"in"),legend.title=element_text(size=9),plot.margin = margin(-0.3,0,-0.1,0,unit = "in"))

pl <- extract_legend(p)
dev.off()


# save in dimension ~equiv to the dimensions of a legend included in one panel
pdf("plots/Manuscript/4D-G shared legend.pdf",height=0.6,width=0.45)
do.call("grid.arrange",list(pl))
dev.off()
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.1 (2023-06-16)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] spatialLIBD_1.14.1          escheR_1.2.0               
 [3] SpatialExperiment_1.12.0    SingleCellExperiment_1.24.0
 [5] SummarizedExperiment_1.32.0 Biobase_2.62.0             
 [7] GenomicRanges_1.54.1        MatrixGenerics_1.14.0      
 [9] matrixStats_1.2.0           Cairo_1.6-2                
[11] ggrastr_1.0.2               BiocParallel_1.36.0        
[13] parallelly_1.36.0           default_1.0.0              
[15] colorout_1.3-0              gridExtra_2.3              
[17] ggplot2_3.4.4               Biostrings_2.70.1          
[19] GenomeInfoDb_1.38.2         XVector_0.42.0             
[21] IRanges_2.36.0              S4Vectors_0.40.2           
[23] BiocGenerics_0.48.1         data.table_1.14.10         
[25] rlang_1.1.2                

loaded via a namespace (and not attached):
  [1] later_1.3.2                   BiocIO_1.12.0                
  [3] fields_15.2                   bitops_1.0-7                 
  [5] filelock_1.0.3                tibble_3.2.1                 
  [7] polyclip_1.10-6               XML_3.99-0.16                
  [9] lifecycle_1.0.4               edgeR_4.0.3                  
 [11] doParallel_1.0.17             rprojroot_2.0.4              
 [13] lattice_0.22-5                MASS_7.3-60                  
 [15] magrittr_2.0.3                sass_0.4.8                   
 [17] limma_3.58.1                  openxlsx_4.2.5.2             
 [19] plotly_4.10.3                 rmarkdown_2.25               
 [21] jquerylib_0.1.4               yaml_2.3.8                   
 [23] httpuv_1.6.13                 spam_2.10-0                  
 [25] zip_2.3.0                     sessioninfo_1.2.2            
 [27] cowplot_1.1.2                 DBI_1.2.0                    
 [29] RColorBrewer_1.1-3            golem_0.4.1                  
 [31] maps_3.4.2                    abind_1.4-5                  
 [33] zlibbioc_1.48.0               rvest_1.0.3                  
 [35] purrr_1.0.2                   msigdbr_7.5.1                
 [37] RCurl_1.98-1.13               tweenr_2.0.2                 
 [39] rappdirs_0.3.3                reactable_0.4.4              
 [41] GenomeInfoDbData_1.2.11       ggrepel_0.9.4                
 [43] irlba_2.3.5.1                 DelayedMatrixStats_1.24.0    
 [45] svglite_2.1.3                 codetools_0.2-19             
 [47] DelayedArray_0.28.0           scuttle_1.12.0               
 [49] DT_0.31                       xml2_1.3.6                   
 [51] ggforce_0.4.1                 tidyselect_1.2.0             
 [53] farver_2.1.1                  viridis_0.6.4                
 [55] ScaledMatrix_1.10.0           shinyWidgets_0.8.0           
 [57] BiocFileCache_2.10.1          webshot_0.5.5                
 [59] GenomicAlignments_1.38.0      jsonlite_1.8.8               
 [61] BiocNeighbors_1.20.1          ellipsis_0.3.2               
 [63] scater_1.30.1                 iterators_1.0.14             
 [65] systemfonts_1.0.5             foreach_1.5.2                
 [67] tools_4.3.1                   Rcpp_1.0.11                  
 [69] glue_1.6.2                    SparseArray_1.2.2            
 [71] xfun_0.41                     here_1.0.1                   
 [73] dplyr_1.1.4                   withr_2.5.2                  
 [75] BiocManager_1.30.22           fastmap_1.1.1                
 [77] fansi_1.0.6                   rsvd_1.0.5                   
 [79] digest_0.6.33                 R6_2.5.1                     
 [81] mime_0.12                     colorspace_2.1-0             
 [83] RSQLite_2.3.4                 config_0.3.2                 
 [85] utf8_1.2.4                    tidyr_1.3.0                  
 [87] generics_0.1.3                rtracklayer_1.62.0           
 [89] httr_1.4.7                    htmlwidgets_1.6.4            
 [91] S4Arrays_1.2.0                pkgconfig_2.0.3              
 [93] gtable_0.3.4                  blob_1.2.4                   
 [95] htmltools_0.5.7               dotCall64_1.1-1              
 [97] scales_1.3.0                  kableExtra_1.3.4             
 [99] attempt_0.3.1                 png_0.1-8                    
[101] knitr_1.45                    rstudioapi_0.15.0            
[103] reshape2_1.4.4                rjson_0.2.21                 
[105] visNetwork_2.1.2              curl_5.2.0                   
[107] cachem_1.0.8                  stringr_1.5.1                
[109] BiocVersion_3.18.1            parallel_4.3.1               
[111] vipor_0.4.7                   AnnotationDbi_1.64.1         
[113] restfulr_0.0.15               pillar_1.9.0                 
[115] grid_4.3.1                    vctrs_0.6.5                  
[117] promises_1.2.1                BiocSingular_1.18.0          
[119] beachmat_2.18.0               dbplyr_2.4.0                 
[121] xtable_1.8-4                  beeswarm_0.4.0               
[123] paletteer_1.5.0               evaluate_0.23                
[125] magick_2.8.2                  Rsamtools_2.18.0             
[127] locfit_1.5-9.8                cli_3.6.2                    
[129] compiler_4.3.1                crayon_1.5.2                 
[131] hypeR_2.0.0                   labeling_0.4.3               
[133] rematch2_2.1.2                plyr_1.8.9                   
[135] ggbeeswarm_0.7.2              stringi_1.8.3                
[137] viridisLite_0.4.2             babelgene_22.9               
[139] munsell_0.5.0                 lazyeval_0.2.2               
[141] Matrix_1.6-4                  ExperimentHub_2.10.0         
[143] benchmarkme_1.0.8             sparseMatrixStats_1.14.0     
[145] bit64_4.0.5                   statmod_1.5.0                
[147] KEGGREST_1.42.0               shiny_1.8.0                  
[149] interactiveDisplayBase_1.40.0 AnnotationHub_3.10.0         
[151] igraph_1.6.0                  memoise_2.0.1                
[153] bslib_0.6.1                   benchmarkmeData_1.0.4        
[155] bit_4.0.5                    
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.1 (2023-06-16)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2023-12-28
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package                * version   date (UTC) lib source
 abind                    1.4-5     2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1    2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0    2023-10-24 [1] Bioconductor
 attempt                  0.3.1     2020-05-03 [1] CRAN (R 4.3.0)
 babelgene                22.9      2022-09-29 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0    2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0     2021-06-01 [1] CRAN (R 4.3.0)
 benchmarkme              1.0.8     2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4     2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0    2023-10-26 [1] Bioconductor
 BiocFileCache            2.10.1    2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1    2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0    2023-10-24 [1] Bioconductor
 BiocManager              1.30.22   2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.1    2023-12-19 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel           * 1.36.0    2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0    2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1    2023-11-15 [1] Bioconductor
 Biostrings             * 2.70.1    2023-10-26 [1] Bioconductor
 bit                      4.0.5     2022-11-15 [1] CRAN (R 4.3.0)
 bit64                    4.0.5     2020-08-30 [1] CRAN (R 4.3.0)
 bitops                   1.0-7     2021-04-24 [1] CRAN (R 4.3.0)
 blob                     1.2.4     2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.6.1     2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8     2023-05-01 [1] CRAN (R 4.3.0)
 Cairo                  * 1.6-2     2023-11-28 [1] CRAN (R 4.3.1)
 cli                      3.6.2     2023-12-11 [1] CRAN (R 4.3.1)
 codetools                0.2-19    2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0     2023-11-01 [1] local (/Users/bmulvey/Desktop/colorout)
 colorspace               2.1-0     2023-01-23 [1] CRAN (R 4.3.0)
 config                   0.3.2     2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.2     2023-12-15 [1] CRAN (R 4.3.1)
 crayon                   1.5.2     2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0     2023-12-08 [1] CRAN (R 4.3.1)
 data.table             * 1.14.10   2023-12-08 [1] CRAN (R 4.3.1)
 DBI                      1.2.0     2023-12-21 [1] CRAN (R 4.3.1)
 dbplyr                   2.4.0     2023-10-26 [1] CRAN (R 4.3.1)
 default                * 1.0.0     2017-08-07 [1] CRAN (R 4.3.0)
 DelayedArray             0.28.0    2023-11-06 [1] Bioconductor
 DelayedMatrixStats       1.24.0    2023-10-24 [1] Bioconductor
 digest                   0.6.33    2023-07-07 [1] CRAN (R 4.3.0)
 doParallel               1.0.17    2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1     2023-11-28 [1] CRAN (R 4.3.1)
 dplyr                    1.1.4     2023-11-17 [1] CRAN (R 4.3.1)
 DT                       0.31      2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.3     2023-12-09 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2     2021-04-29 [1] CRAN (R 4.3.0)
 escheR                 * 1.2.0     2023-10-26 [1] Bioconductor
 evaluate                 0.23      2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0    2023-10-24 [1] Bioconductor
 fansi                    1.0.6     2023-12-08 [1] CRAN (R 4.3.1)
 farver                   2.1.1     2022-07-06 [1] CRAN (R 4.3.0)
 fastmap                  1.1.1     2023-02-24 [1] CRAN (R 4.3.0)
 fields                   15.2      2023-08-17 [1] CRAN (R 4.3.0)
 filelock                 1.0.3     2023-12-11 [1] CRAN (R 4.3.1)
 foreach                  1.5.2     2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3     2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.2    2023-12-13 [1] Bioconductor 3.18 (R 4.3.1)
 GenomeInfoDbData         1.2.11    2023-11-10 [1] Bioconductor
 GenomicAlignments        1.38.0    2023-10-24 [1] Bioconductor
 GenomicRanges          * 1.54.1    2023-10-30 [1] Bioconductor
 ggbeeswarm               0.7.2     2023-04-29 [1] CRAN (R 4.3.0)
 ggforce                  0.4.1     2022-10-04 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4     2023-10-12 [1] CRAN (R 4.3.1)
 ggrastr                * 1.0.2     2023-06-01 [1] CRAN (R 4.3.0)
 ggrepel                  0.9.4     2023-10-13 [1] CRAN (R 4.3.1)
 glue                     1.6.2     2022-02-24 [1] CRAN (R 4.3.0)
 golem                    0.4.1     2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3       2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4     2023-08-21 [1] CRAN (R 4.3.0)
 here                     1.0.1     2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7     2023-11-03 [1] CRAN (R 4.3.1)
 htmlwidgets              1.6.4     2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13    2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7     2023-08-15 [1] CRAN (R 4.3.0)
 hypeR                    2.0.0     2023-10-24 [1] Bioconductor
 igraph                   1.6.0     2023-12-11 [1] CRAN (R 4.3.1)
 interactiveDisplayBase   1.40.0    2023-10-24 [1] Bioconductor
 IRanges                * 2.36.0    2023-10-26 [1] Bioconductor
 irlba                    2.3.5.1   2022-10-03 [1] CRAN (R 4.3.1)
 iterators                1.0.14    2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4     2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8     2023-12-04 [1] CRAN (R 4.3.1)
 kableExtra               1.3.4     2021-02-20 [1] CRAN (R 4.3.0)
 KEGGREST                 1.42.0    2023-10-26 [1] Bioconductor
 knitr                    1.45      2023-10-30 [1] CRAN (R 4.3.1)
 labeling                 0.4.3     2023-08-29 [1] CRAN (R 4.3.0)
 later                    1.3.2     2023-12-06 [1] CRAN (R 4.3.1)
 lattice                  0.22-5    2023-10-24 [1] CRAN (R 4.3.1)
 lazyeval                 0.2.2     2019-03-15 [1] CRAN (R 4.3.0)
 lifecycle                1.0.4     2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1    2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8   2023-06-11 [1] CRAN (R 4.3.0)
 magick                   2.8.2     2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3     2022-03-30 [1] CRAN (R 4.3.0)
 maps                     3.4.2     2023-12-15 [1] CRAN (R 4.3.1)
 MASS                     7.3-60    2023-05-04 [1] CRAN (R 4.3.0)
 Matrix                   1.6-4     2023-11-30 [1] CRAN (R 4.3.1)
 MatrixGenerics         * 1.14.0    2023-10-26 [1] Bioconductor
 matrixStats            * 1.2.0     2023-12-11 [1] CRAN (R 4.3.1)
 memoise                  2.0.1     2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12      2021-09-28 [1] CRAN (R 4.3.0)
 msigdbr                  7.5.1     2022-03-30 [1] CRAN (R 4.3.0)
 munsell                  0.5.0     2018-06-12 [1] CRAN (R 4.3.0)
 openxlsx                 4.2.5.2   2023-02-06 [1] CRAN (R 4.3.0)
 paletteer                1.5.0     2022-10-19 [1] CRAN (R 4.3.0)
 parallelly             * 1.36.0    2023-05-26 [1] CRAN (R 4.3.0)
 pillar                   1.9.0     2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3     2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.3    2023-10-21 [1] CRAN (R 4.3.1)
 plyr                     1.8.9     2023-10-02 [1] CRAN (R 4.3.1)
 png                      0.1-8     2022-11-29 [1] CRAN (R 4.3.0)
 polyclip                 1.10-6    2023-09-27 [1] CRAN (R 4.3.1)
 promises                 1.2.1     2023-08-10 [1] CRAN (R 4.3.0)
 purrr                    1.0.2     2023-08-10 [1] CRAN (R 4.3.0)
 R6                       2.5.1     2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3     2021-01-31 [1] CRAN (R 4.3.0)
 RColorBrewer             1.1-3     2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.11    2023-07-06 [1] CRAN (R 4.3.1)
 RCurl                    1.98-1.13 2023-11-02 [1] CRAN (R 4.3.1)
 reactable                0.4.4     2023-03-12 [1] CRAN (R 4.3.0)
 rematch2                 2.1.2     2020-05-01 [1] CRAN (R 4.3.0)
 reshape2                 1.4.4     2020-04-09 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15    2022-06-16 [1] CRAN (R 4.3.0)
 rjson                    0.2.21    2022-01-09 [1] CRAN (R 4.3.0)
 rlang                  * 1.1.2     2023-11-04 [1] CRAN (R 4.3.1)
 rmarkdown                2.25      2023-09-18 [1] CRAN (R 4.3.1)
 rprojroot                2.0.4     2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0    2023-10-24 [1] Bioconductor
 RSQLite                  2.3.4     2023-12-08 [1] CRAN (R 4.3.1)
 rstudioapi               0.15.0    2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5     2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0    2023-10-26 [1] Bioconductor
 rvest                    1.0.3     2022-08-19 [1] CRAN (R 4.3.0)
 S4Arrays                 1.2.0     2023-10-26 [1] Bioconductor
 S4Vectors              * 0.40.2    2023-11-25 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8     2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0    2023-10-24 [1] Bioconductor
 scales                   1.3.0     2023-11-28 [1] CRAN (R 4.3.1)
 scater                   1.30.1    2023-11-16 [1] Bioconductor
 scuttle                  1.12.0    2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2     2021-12-06 [1] CRAN (R 4.3.0)
 shiny                    1.8.0     2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.0     2023-08-30 [1] CRAN (R 4.3.0)
 SingleCellExperiment   * 1.24.0    2023-11-06 [1] Bioconductor
 spam                     2.10-0    2023-10-23 [1] CRAN (R 4.3.1)
 SparseArray              1.2.2     2023-11-08 [1] Bioconductor
 sparseMatrixStats        1.14.0    2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0    2023-10-24 [1] Bioconductor
 spatialLIBD            * 1.14.1    2023-11-30 [1] Bioconductor 3.18 (R 4.3.1)
 statmod                  1.5.0     2023-01-06 [1] CRAN (R 4.3.0)
 stringi                  1.8.3     2023-12-11 [1] CRAN (R 4.3.1)
 stringr                  1.5.1     2023-11-14 [1] CRAN (R 4.3.1)
 SummarizedExperiment   * 1.32.0    2023-11-06 [1] Bioconductor
 svglite                  2.1.3     2023-12-08 [1] CRAN (R 4.3.1)
 systemfonts              1.0.5     2023-10-09 [1] CRAN (R 4.3.1)
 tibble                   3.2.1     2023-03-20 [1] CRAN (R 4.3.0)
 tidyr                    1.3.0     2023-01-24 [1] CRAN (R 4.3.0)
 tidyselect               1.2.0     2022-10-10 [1] CRAN (R 4.3.0)
 tweenr                   2.0.2     2022-09-06 [1] CRAN (R 4.3.0)
 utf8                     1.2.4     2023-10-22 [1] CRAN (R 4.3.1)
 vctrs                    0.6.5     2023-12-01 [1] CRAN (R 4.3.1)
 vipor                    0.4.7     2023-12-18 [1] CRAN (R 4.3.1)
 viridis                  0.6.4     2023-07-22 [1] CRAN (R 4.3.0)
 viridisLite              0.4.2     2023-05-02 [1] CRAN (R 4.3.0)
 visNetwork               2.1.2     2022-09-29 [1] CRAN (R 4.3.0)
 webshot                  0.5.5     2023-06-26 [1] CRAN (R 4.3.0)
 withr                    2.5.2     2023-10-30 [1] CRAN (R 4.3.1)
 xfun                     0.41      2023-11-01 [1] CRAN (R 4.3.1)
 XML                      3.99-0.16 2023-11-29 [1] CRAN (R 4.3.1)
 xml2                     1.3.6     2023-12-04 [1] CRAN (R 4.3.1)
 xtable                   1.8-4     2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0    2023-10-26 [1] Bioconductor
 yaml                     2.3.8     2023-12-11 [1] CRAN (R 4.3.1)
 zip                      2.3.0     2023-04-17 [1] CRAN (R 4.3.0)
 zlibbioc                 1.48.0    2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
