---
title: "Supplemental Fig-Spotplots of NR5A1, NPY, MBP in all samples without rotations
  or mirrorings"
author: "Bernard Mulvey"
date: "2024-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
library(viridis)
library(ggtext)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load raw spe (i.e., no rotations, no mirrors); subset to spots and genes in the final filtered object for exactly comparable non-visual data;

```{r}
load("processed-data/02_build_spe/spe_raw_w_V13M13-362.Rdata")

hyp2 <- readRDS("data/03-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

spe_raw <- spe_raw[rownames(spe_raw) %in% rownames(hyp2),][rownames(hyp2),]
spe_raw <- spe_raw[,colnames(spe_raw) %in% colnames(hyp2)][,colnames(hyp2)]

# move over logcounts info
assays(spe_raw)$logcounts <- assays(hyp2)$logcounts


rm(hyp2)
gc(full=T)

rownames(spe_raw) <- rowData(spe_raw)$gene_name
```

### override the legend title spatialLIBD leaves us stuck with by overwriting their function definition
```{r}
vis_gene_bm <- function (spe, 
                         sampleid = unique(spe$sample_id)[1], 
                         geneid = rowData(spe)$gene_search[1], 
                         spatial = TRUE, 
                         assayname = "logcounts", 
                         minCount = 0, 
                         viridis = FALSE, 
                         image_id = "lowres", 
                         alpha = 0.7, 
                         cont_colors = magma(256,begin=0,end=0.7),
                         auto_crop=TRUE,
                         na_color=NA,
                         point_size=0.6){
    
    spe_sub <- spe[, spe$sample_id == sampleid]
    d <- as.data.frame(cbind(colData(spe_sub), SpatialExperiment::spatialCoords(spe_sub)), 
        optional = TRUE)
    if (geneid %in% colnames(colData(spe_sub))) {
        d$COUNT <- colData(spe_sub)[[geneid]]
    }
    else if (geneid %in% rowData(spe_sub)$gene_search) {
        d$COUNT <- assays(spe_sub)[[assayname]][which(rowData(spe_sub)$gene_search == 
            geneid), ]
    }
    else if (geneid %in% rownames(spe_sub)) {
        d$COUNT <- assays(spe_sub)[[assayname]][which(rownames(spe_sub) == 
            geneid), ]
    }
    else {
        stop("Could not find the 'geneid' ", geneid, call. = FALSE)
    }
    d$COUNT[d$COUNT <= minCount] <- NA
    p <- vis_gene_p(
        spe = spe_sub,
        d = d,
        sampleid = sampleid,
        spatial = spatial,
        title = paste0(sampleid," ", paste0("*",geneid,"*")),
        viridis = viridis,
        image_id = image_id,
        alpha = alpha,
        cont_colors = cont_colors,
        point_size = point_size,
        auto_crop = auto_crop,
        na_color = na_color,
        legend_title = "log\ncounts"
    )
    return(p)
}
```


spotplot MBP, NPY, NR5A1 for comparisons w/ VMH/ARC localization RNAscopes in supp fig. plot three ways for heena: full pages by gene, strips w/ each gene for one sample, and individual plots.
 
### format 1: full pages by gene
```{r}
plotg <- c("MBP","NPY","NR5A1")
samps <- unique(spe_raw$sample_id)


i <- 1
for (i in c(1:length(plotg))){
    j<-1
    plt <- list()
    for (j in c(1:length(samps))){
        cursamp <- samps[j]
        spe_tmp <- spe_raw[,spe_raw$sample_id==cursamp]
        plt[[j]] <- vis_gene_bm(spe = spe_tmp,
                           geneid=plotg[i],
                           spatial=T,
                           point_size=1,
                           alpha=0.65,
                           viridis=FALSE,
                           na_color = NA,
                           auto_crop = T)
    }
    # adjust formatting for pdf
    plt <- lapply(plt,FUN=function(x){x+
            theme(plot.title = element_markdown(size = 11,hjust=0.5),
                  strip.text = element_text(size=10),
                  legend.text = element_text(size=10),
                  legend.title = element_text(size=10,hjust=0.5),
                  panel.border=element_blank(),
                  axis.text=element_blank(),
                  axis.title=element_blank())
            })
    pdf(paste0("plots/Manuscript/Supp Panels-NR5A1-NPY-MBP spotplots/Full page by gene/",plotg[i],".pdf"),height=8,width=10)
    do.call("grid.arrange",plt)
    dev.off()
}

```


### format 2: panel strips of all genes, single samples. (height = 1.2 inch, strip width = 3.75 in)
### geez, that took a lot of margin tinkering
```{r}
sbp <- MulticoreParam(4)
register(sbp)

samps <- unique(spe_raw$sample_id)
bplapply(samps,BPPARAM = sbp,function(s){
    plt<-list()
    i <- 1
    for (i in c(1:length(plotg))){
        plt[[i]] <- vis_gene_bm(spe = spe_raw,
                             sampleid=s,
                             geneid = plotg[i],
                             spatial=T,
                             point_size=0.4,
                             alpha=0.65,
                             viridis=FALSE,
                             na_color = NA,
                             auto_crop = T)
    # adjust formatting for pdf
    plt[[i]] <- plt[[i]]+
        theme(plot.title = element_markdown(size=8,
                                        hjust=0.5,
                                        margin=
                                            margin(0,0,0,0,unit="in")),
              legend.box = element_blank(),
              legend.justification = "right",
              strip.text = element_blank(),
              legend.text = element_text(size=4),
              legend.title = element_blank(),
              plot.margin = margin(0.025,0,-0.075,-0.05,unit = "in"),
              legend.spacing.x = unit(0.005,"in"),
              legend.key.height = unit(0.1,"in"),
              legend.key.width = unit(0.05,"in"),
              legend.margin = margin(-0.15,0,0,0,"in"),
              panel.border = element_blank(),
              axis.text=element_blank(),
              axis.title=element_blank())+
        ggtitle(paste0("*",plotg[i],"*"))
    }
    pdf(paste0("plots/Manuscript/Supp Panels-NR5A1-NPY-MBP spotplots/Strips per sample/",s,".pdf"),height=1.2,width=3.75)
    do.call("grid.arrange",c(plt,ncol=3))
    dev.off()
})

```

## format 3: individual plots
```{r}
sbp <- MulticoreParam(8)
register(sbp)

samps <- unique(spe_raw$sample_id)
bplapply(samps,BPPARAM = sbp,function(s){
    i <- 1
    for (i in c(1:length(plotg))){
        plt <- vis_gene_bm(spe = spe_raw,
                             sampleid=s,
                             geneid = plotg[i],
                             spatial=T,
                             point_size=0.4,
                             alpha=0.65,
                             viridis=FALSE,
                             na_color = NA,
                             auto_crop = T)
    # adjust formatting for pdf
        plt <- plt+
            theme(plot.title = element_markdown(size=8,
                                            hjust=0.5,
                                            margin=
                                                margin(0,0,0,0,unit="in")),
                  legend.box = element_blank(),
                  legend.justification = "right",
                  strip.text = element_blank(),
                  legend.text = element_text(size=4),
                  legend.title = element_blank(),
                  plot.margin = margin(0.1,0,-0.25,-0.3,unit = "in"),
                  legend.spacing.x = unit(0.005,"in"),
                  legend.key.height = unit(0.1,"in"),
                  legend.key.width = unit(0.05,"in"),
                  legend.margin = margin(-0.15,0,0,0,"in"),
                  panel.border=element_blank())+
            ggtitle(paste0("*",plotg[i],"*"))
    
        pdf(paste0("plots/Manuscript/Supp Panels-NR5A1-NPY-MBP spotplots/Single panels/",s,"_",plotg[i],".pdf"),height=1.2,width=1.25)
        print(plt)
        dev.off()
    }
})
```

```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6.4

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] ggtext_0.1.2                viridis_0.6.5              
 [3] viridisLite_0.4.2           BiocParallel_1.36.0        
 [5] parallelly_1.36.0           default_1.0.0              
 [7] colorout_1.3-0.1            spatialLIBD_1.14.1         
 [9] SpatialExperiment_1.12.0    SingleCellExperiment_1.24.0
[11] SummarizedExperiment_1.32.0 Biobase_2.62.0             
[13] GenomicRanges_1.54.1        MatrixGenerics_1.14.0      
[15] matrixStats_1.2.0           gridExtra_2.3              
[17] ggplot2_3.4.4               Biostrings_2.70.2          
[19] GenomeInfoDb_1.38.5         XVector_0.42.0             
[21] IRanges_2.36.0              S4Vectors_0.40.2           
[23] BiocGenerics_0.48.1         data.table_1.14.10         
[25] rlang_1.1.3                

loaded via a namespace (and not attached):
  [1] later_1.3.2                   BiocIO_1.12.0                
  [3] bitops_1.0-7                  filelock_1.0.3               
  [5] fields_15.2                   tibble_3.2.1                 
  [7] XML_3.99-0.16.1               lifecycle_1.0.4              
  [9] edgeR_4.0.14                  doParallel_1.0.17            
 [11] rprojroot_2.0.4               lattice_0.22-5               
 [13] magrittr_2.0.3                limma_3.58.1                 
 [15] plotly_4.10.4                 sass_0.4.8                   
 [17] rmarkdown_2.25                jquerylib_0.1.4              
 [19] yaml_2.3.8                    httpuv_1.6.13                
 [21] spam_2.10-0                   sessioninfo_1.2.2            
 [23] cowplot_1.1.3                 DBI_1.2.1                    
 [25] RColorBrewer_1.1-3            golem_0.4.1                  
 [27] maps_3.4.2                    abind_1.4-5                  
 [29] zlibbioc_1.48.0               purrr_1.0.2                  
 [31] RCurl_1.98-1.14               rappdirs_0.3.3               
 [33] GenomeInfoDbData_1.2.11       ggrepel_0.9.5                
 [35] irlba_2.3.5.1                 commonmark_1.9.1             
 [37] DelayedMatrixStats_1.24.0     codetools_0.2-19             
 [39] DelayedArray_0.28.0           xml2_1.3.6                   
 [41] DT_0.31                       scuttle_1.12.0               
 [43] tidyselect_1.2.0              farver_2.1.1                 
 [45] ScaledMatrix_1.10.0           shinyWidgets_0.8.1           
 [47] BiocFileCache_2.10.1          GenomicAlignments_1.38.2     
 [49] jsonlite_1.8.8                BiocNeighbors_1.20.2         
 [51] ellipsis_0.3.2                scater_1.30.1                
 [53] iterators_1.0.14              foreach_1.5.2                
 [55] tools_4.3.2                   Rcpp_1.0.12                  
 [57] glue_1.7.0                    SparseArray_1.2.3            
 [59] xfun_0.41                     here_1.0.1                   
 [61] dplyr_1.1.4                   withr_3.0.0                  
 [63] BiocManager_1.30.22           fastmap_1.1.1                
 [65] fansi_1.0.6                   digest_0.6.34                
 [67] rsvd_1.0.5                    R6_2.5.1                     
 [69] mime_0.12                     colorspace_2.1-0             
 [71] markdown_1.12                 RSQLite_2.3.5                
 [73] config_0.3.2                  utf8_1.2.4                   
 [75] tidyr_1.3.1                   generics_0.1.3               
 [77] rtracklayer_1.62.0            httr_1.4.7                   
 [79] htmlwidgets_1.6.4             S4Arrays_1.2.0               
 [81] pkgconfig_2.0.3               gtable_0.3.4                 
 [83] blob_1.2.4                    htmltools_0.5.7              
 [85] dotCall64_1.1-1               scales_1.3.0                 
 [87] png_0.1-8                     attempt_0.3.1                
 [89] knitr_1.45                    rstudioapi_0.15.0            
 [91] rjson_0.2.21                  curl_5.2.0                   
 [93] cachem_1.0.8                  stringr_1.5.1                
 [95] BiocVersion_3.18.1            parallel_4.3.2               
 [97] vipor_0.4.7                   AnnotationDbi_1.64.1         
 [99] restfulr_0.0.15               pillar_1.9.0                 
[101] grid_4.3.2                    vctrs_0.6.5                  
[103] promises_1.2.1                BiocSingular_1.18.0          
[105] dbplyr_2.4.0                  beachmat_2.18.0              
[107] xtable_1.8-4                  beeswarm_0.4.0               
[109] paletteer_1.6.0               evaluate_0.23                
[111] magick_2.8.2                  cli_3.6.2                    
[113] locfit_1.5-9.8                compiler_4.3.2               
[115] Rsamtools_2.18.0              crayon_1.5.2                 
[117] labeling_0.4.3                rematch2_2.1.2               
[119] ggbeeswarm_0.7.2              stringi_1.8.3                
[121] munsell_0.5.0                 lazyeval_0.2.2               
[123] Matrix_1.6-5                  ExperimentHub_2.10.0         
[125] benchmarkme_1.0.8             sparseMatrixStats_1.14.0     
[127] bit64_4.0.5                   KEGGREST_1.42.0              
[129] statmod_1.5.0                 shiny_1.8.0                  
[131] interactiveDisplayBase_1.40.0 AnnotationHub_3.10.0         
[133] gridtext_0.1.5                memoise_2.0.1                
[135] bslib_0.6.1                   benchmarkmeData_1.0.4        
[137] bit_4.0.5                    
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6.4
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-02-26
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package                * version     date (UTC) lib source
 abind                    1.4-5       2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1      2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0      2023-10-24 [1] Bioconductor
 attempt                  0.3.1       2020-05-03 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0      2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0       2021-06-01 [1] CRAN (R 4.3.2)
 benchmarkme              1.0.8       2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4       2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0      2023-10-24 [1] Bioconductor
 BiocFileCache            2.10.1      2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1      2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0      2023-10-26 [1] Bioconductor
 BiocManager              1.30.22     2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.2      2024-01-07 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel           * 1.36.0      2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0      2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1      2023-11-18 [1] Bioconductor 3.18 (R 4.3.2)
 Biostrings             * 2.70.2      2024-01-28 [1] Bioconductor 3.18 (R 4.3.2)
 bit                      4.0.5       2022-11-15 [1] CRAN (R 4.3.2)
 bit64                    4.0.5       2020-08-30 [1] CRAN (R 4.3.2)
 bitops                   1.0-7       2021-04-24 [1] CRAN (R 4.3.2)
 blob                     1.2.4       2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.6.1       2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8       2023-05-01 [1] CRAN (R 4.3.2)
 cli                      3.6.2       2023-12-11 [1] CRAN (R 4.3.2)
 codetools                0.2-19      2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0.1     2024-01-11 [1] local
 colorspace               2.1-0       2023-01-23 [1] CRAN (R 4.3.2)
 commonmark               1.9.1       2024-01-30 [1] CRAN (R 4.3.2)
 config                   0.3.2       2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.3       2024-01-22 [1] CRAN (R 4.3.1)
 crayon                   1.5.2       2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0       2023-12-08 [1] CRAN (R 4.3.2)
 data.table             * 1.14.10     2023-12-08 [1] CRAN (R 4.3.2)
 DBI                      1.2.1       2024-01-12 [1] CRAN (R 4.3.2)
 dbplyr                   2.4.0       2023-10-26 [1] CRAN (R 4.3.1)
 default                * 1.0.0       2017-08-07 [1] CRAN (R 4.3.0)
 DelayedArray             0.28.0      2023-10-24 [1] Bioconductor
 DelayedMatrixStats       1.24.0      2023-11-06 [1] Bioconductor
 digest                   0.6.34      2024-01-11 [1] CRAN (R 4.3.2)
 doParallel               1.0.17      2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1       2023-11-28 [1] CRAN (R 4.3.2)
 dplyr                    1.1.4       2023-11-17 [1] CRAN (R 4.3.2)
 DT                       0.31        2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.14      2024-01-29 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2       2021-04-29 [1] CRAN (R 4.3.2)
 evaluate                 0.23        2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0      2023-10-26 [1] Bioconductor
 fansi                    1.0.6       2023-12-08 [1] CRAN (R 4.3.2)
 farver                   2.1.1       2022-07-06 [1] CRAN (R 4.3.2)
 fastmap                  1.1.1       2023-02-24 [1] CRAN (R 4.3.2)
 fields                   15.2        2023-08-17 [1] CRAN (R 4.3.2)
 filelock                 1.0.3       2023-12-11 [1] CRAN (R 4.3.2)
 foreach                  1.5.2       2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3       2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.5      2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData         1.2.11      2024-01-11 [1] Bioconductor
 GenomicAlignments        1.38.2      2024-01-16 [1] Bioconductor 3.18 (R 4.3.2)
 GenomicRanges          * 1.54.1      2023-10-29 [1] Bioconductor
 ggbeeswarm               0.7.2       2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4       2023-10-12 [1] CRAN (R 4.3.1)
 ggrepel                  0.9.5       2024-01-10 [1] CRAN (R 4.3.2)
 ggtext                 * 0.1.2       2022-09-16 [1] CRAN (R 4.3.0)
 glue                     1.7.0       2024-01-09 [1] CRAN (R 4.3.2)
 golem                    0.4.1       2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3         2017-09-09 [1] CRAN (R 4.3.0)
 gridtext                 0.1.5       2022-09-16 [1] CRAN (R 4.3.2)
 gtable                   0.3.4       2023-08-21 [1] CRAN (R 4.3.0)
 here                     1.0.1       2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7       2023-11-03 [1] CRAN (R 4.3.2)
 htmlwidgets              1.6.4       2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13      2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7       2023-08-15 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0      2023-10-26 [1] Bioconductor
 IRanges                * 2.36.0      2023-10-24 [1] Bioconductor
 irlba                    2.3.5.1     2022-10-03 [1] CRAN (R 4.3.2)
 iterators                1.0.14      2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4       2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8       2023-12-04 [1] CRAN (R 4.3.2)
 KEGGREST                 1.42.0      2023-10-26 [1] Bioconductor
 knitr                    1.45        2023-10-30 [1] CRAN (R 4.3.2)
 labeling                 0.4.3       2023-08-29 [1] CRAN (R 4.3.0)
 later                    1.3.2       2023-12-06 [1] CRAN (R 4.3.2)
 lattice                  0.22-5      2023-10-24 [1] CRAN (R 4.3.2)
 lazyeval                 0.2.2       2019-03-15 [1] CRAN (R 4.3.2)
 lifecycle                1.0.4       2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1      2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8     2023-06-11 [1] CRAN (R 4.3.2)
 magick                   2.8.2       2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3       2022-03-30 [1] CRAN (R 4.3.2)
 maps                     3.4.2       2023-12-15 [1] CRAN (R 4.3.1)
 markdown                 1.12        2023-12-06 [1] CRAN (R 4.3.1)
 Matrix                   1.6-5       2024-01-11 [1] CRAN (R 4.3.2)
 MatrixGenerics         * 1.14.0      2023-10-26 [1] Bioconductor
 matrixStats            * 1.2.0       2023-12-11 [1] CRAN (R 4.3.2)
 memoise                  2.0.1       2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12        2021-09-28 [1] CRAN (R 4.3.2)
 munsell                  0.5.0       2018-06-12 [1] CRAN (R 4.3.0)
 paletteer                1.6.0       2024-01-21 [1] CRAN (R 4.3.1)
 parallelly             * 1.36.0      2023-05-26 [1] CRAN (R 4.3.0)
 pillar                   1.9.0       2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3       2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.4      2024-01-13 [1] CRAN (R 4.3.1)
 png                      0.1-8       2022-11-29 [1] CRAN (R 4.3.2)
 promises                 1.2.1       2023-08-10 [1] CRAN (R 4.3.2)
 purrr                    1.0.2       2023-08-10 [1] CRAN (R 4.3.2)
 R6                       2.5.1       2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3       2021-01-31 [1] CRAN (R 4.3.2)
 RColorBrewer             1.1-3       2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.12      2024-01-09 [1] CRAN (R 4.3.2)
 RCurl                    1.98-1.14   2024-01-09 [1] CRAN (R 4.3.1)
 rematch2                 2.1.2       2020-05-01 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15      2022-06-16 [1] CRAN (R 4.3.2)
 rjson                    0.2.21      2022-01-09 [1] CRAN (R 4.3.2)
 rlang                  * 1.1.3       2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown                2.25        2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot                2.0.4       2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0      2023-10-24 [1] Bioconductor
 RSQLite                  2.3.5       2024-01-21 [1] CRAN (R 4.3.2)
 rstudioapi               0.15.0      2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5       2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0      2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0       2023-10-24 [1] Bioconductor
 S4Vectors              * 0.40.2      2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8       2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0      2023-11-06 [1] Bioconductor
 scales                   1.3.0       2023-11-28 [1] CRAN (R 4.3.2)
 scater                   1.30.1      2023-11-16 [1] Bioconductor
 scuttle                  1.12.0      2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2       2021-12-06 [1] CRAN (R 4.3.0)
 shiny                    1.8.0       2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.1       2024-01-10 [1] CRAN (R 4.3.1)
 SingleCellExperiment   * 1.24.0      2023-11-06 [1] Bioconductor
 spam                     2.10-0      2023-10-23 [1] CRAN (R 4.3.2)
 SparseArray              1.2.3       2023-12-25 [1] Bioconductor 3.18 (R 4.3.2)
 sparseMatrixStats        1.14.0      2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0      2023-10-26 [1] Bioconductor
 spatialLIBD            * 1.14.1      2023-11-30 [1] Bioconductor 3.18 (R 4.3.2)
 statmod                  1.5.0       2023-01-06 [1] CRAN (R 4.3.2)
 stringi                  1.8.3       2023-12-11 [1] CRAN (R 4.3.1)
 stringr                  1.5.1       2023-11-14 [1] CRAN (R 4.3.1)
 SummarizedExperiment   * 1.32.0      2023-11-06 [1] Bioconductor
 tibble                   3.2.1       2023-03-20 [1] CRAN (R 4.3.2)
 tidyr                    1.3.1       2024-01-24 [1] CRAN (R 4.3.1)
 tidyselect               1.2.0       2022-10-10 [1] CRAN (R 4.3.0)
 utf8                     1.2.4       2023-10-22 [1] CRAN (R 4.3.2)
 vctrs                    0.6.5       2023-12-01 [1] CRAN (R 4.3.2)
 vipor                    0.4.7       2023-12-18 [1] CRAN (R 4.3.1)
 viridis                * 0.6.5       2024-01-29 [1] CRAN (R 4.3.2)
 viridisLite            * 0.4.2       2023-05-02 [1] CRAN (R 4.3.0)
 withr                    3.0.0       2024-01-16 [1] CRAN (R 4.3.1)
 xfun                     0.41        2023-11-01 [1] CRAN (R 4.3.2)
 XML                      3.99-0.16.1 2024-01-22 [1] CRAN (R 4.3.1)
 xml2                     1.3.6       2023-12-04 [1] CRAN (R 4.3.2)
 xtable                   1.8-4       2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0      2023-10-24 [1] Bioconductor
 yaml                     2.3.8       2023-12-11 [1] CRAN (R 4.3.2)
 zlibbioc                 1.48.0      2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
