---
title: "02-Supp Plots of PCA, Harmony, UMAP, k15-31 identities"
author: "Bernard Mulvey"
date: "2024-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(ggspavis)
library(ggrastr) # otherwise these will be enormous unopenable files because 80 bajillion points 
require(colorout)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(gtools)

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

# load SPE; append demographic/technical data (age, sex, BMI, PMI, best RIN from cortex..). plot dimensionality reductions already appended to the SPE coded by variables of interest. we also need to add the BayesSpace clustering results reported in the manuscript.
```{r}
hyp2 <- readRDS("data/04-feature_selection/02-hypfiltered_4featureset-pca-umap-harmonydefault-harmonylambdanull-mnn30.RDS")

# load and append demographics and other misc data
demos <- fread("raw-data/demos.txt")
setnames(demos,c("Best RIN PFC","BMI (calculated)"),c("best_RIN_ctx","BMI"))
tmpcd <- as.data.table(colData(hyp2))
tmpcd <- merge.data.table(tmpcd,demos,by.x="brnum",by.y="BrNum")
tmpcd <- DataFrame(tmpcd)
rownames(tmpcd) <- tmpcd$key
tmpcd <- tmpcd[colnames(hyp2),]
colData(hyp2) <- tmpcd

## append cluster assignments
bscls <- list.files("data/06-BayesSpace/02-bayesspace60kiter_k15-20-31_out/",pattern="HARMONYlmbna_nnsvg10.txt",full.names=T)

bscl <- lapply(bscls,FUN=function(x){
    y <- fread(x)
    setnames(y,c("rn","cl"))
    y[,cl:=paste0("X",cl)]
    if(length(unique(y$cl))==15){
        y[cl=="X3",cl:="ARC.1"]
        y[cl=="X5",cl:="ARC.2"]
        y[cl=="X11",cl:="VMH.1"]
        y[cl=="X12",cl:="VMH.2"]
    }
    else if (length(unique(y$cl))==20){
        y[cl=="X7",cl:="VMH20.1"]
        y[cl=="X9",cl:="VMH20.2"]
        y[cl=="X13",cl:="VMH20.3"]
        y[cl=="X14",cl:="ARC20.1"]
        y[cl=="X18",cl:="ARC20.2"]
    }
    else {
        y[cl=="X3",cl:="VMH31.1"]
        y[cl=="X4",cl:="VMH31.2"]
        y[cl=="X6",cl:="VMH31.3"]
        y[cl=="X19",cl:="VMH31.4"]
        y[cl=="X20",cl:="VMH31.5"]
        y[cl=="X29",cl:="VMH31.6"]
        
        y[cl=="X1",cl:="ARC31.1"]
        y[cl=="X7",cl:="ARC31.2"]
        y[cl=="X8",cl:="ARC31.3"]
        y[cl=="X21",cl:="ARC31.4"]
    }
    y <- as.data.frame(y)
    rownames(y) <- y$rn
    y <- y[colnames(hyp2),]
    return(y)
})

## add a fourth one for collapsed VMH-arc
bscl[[4]] <- copy(bscl[[1]])
bscl[[4]][bscl[[4]]$cl %in% c("ARC.1","ARC.2"),"cl"] <- "ARC"
bscl[[4]][bscl[[4]]$cl %in% c("VMH.1","VMH.2"),"cl"] <- "VMH"

names(bscl) <- c("k15","k20","k31","k15_collapsed")
i<-1
for (i in c(1:4)){
    hyp2$newcol <- bscl[[i]]$cl
    colnames(colData(hyp2))[ncol(colData(hyp2))] <- names(bscl)[i]
}
```

# get variables we want to plot
```{r}
colnames(colData(hyp2))

# for dimensionality reduction at the spot level, we want to include brnum, agedeath, race, sex, PMI, best_rin_pfc, bmi, sample_id, sum_umi, sum_gene, expr_chrM, slide, and to be sure nothing too strange is happening, array_row and array_col
labvars <- c("brnum","AgeDeath","Sex","PrimaryDx","PMI","best_RIN_ctx","BMI","sample_id","sum_umi","sum_gene","expr_chrM","slide","array_row","array_col","k15","k20","k31","k15_collapsed")

# as for dimreds, we just want the nnsvg10 / HARMONY lmbna ones.
drs <- grep(reducedDimNames(hyp2),pattern="10x",value=T,invert=T)
drs <- grep(drs,pattern="nnsvg10",value=T)
drs <- grep(drs,pattern="HARMONYdflt|mnn30",value=T,invert=T)
```

### run 'em
```{r}
options(StringsAsFactors = FALSE)
lapply(drs,FUN=function(x){
    # dimred combinations to plot #
    if(length(grep(x,pattern="UMAP",value=T))>0){
        zq <- rbind(c(1,2))
    }
    else{
        zq <- rbind(c(1,2),c(1,3))
    }
    
    plts <- list()
    # iterate thru variables
    v <- 1
    for (v in c(1:length(labvars))){
        m <- 1
        for (m in c(1:nrow(zq))){
            z <- zq[m,1]
            q <- zq[m,2]
            pltdat <- cbind(reducedDims(hyp2)[[x]][,c(z,q)],colData(hyp2)[,labvars[v]])
            
            pltdat <- as.data.table(pltdat)
            
            # axis labels
            axlab2 <- ifelse(length(grep(x,pattern="UMAP"))>0,
                             "UMAP","PC")
            axlab1 <- ifelse(length(grep(x,pattern="HARMONY"))>0,
                             "Harmony ","Pre-Harmony ")
            axlab <- paste0(axlab1,axlab2)
            rm(axlab1,axlab2)
            
            setnames(pltdat,c(paste0(axlab,paste0(" Dim ",z)),paste0(axlab,paste0(" Dim ",q)),labvars[v]))
            xvar <- names(pltdat)[1]
            yvar <- names(pltdat)[2]
            
            # set dimensions to numeric, plotting variable to factor or numeric
            set(pltdat,j=xvar,value=as.numeric(pltdat[,get(xvar)]))
            set(pltdat,j=yvar,value=as.numeric(pltdat[,get(yvar)]))
            if(labvars[v] %in% c("AgeDeath","PMI","best_rin_PFC","BMI","sum_umi","sum_gene","expr_chrM","array_row","array_col")){
                set(pltdat,j=labvars[v],value=as.numeric(pltdat[,get(labvars[v])]))
            }
            else{
                set(pltdat,j=labvars[v],value=as.factor(pltdat[,get(labvars[v])]))
            }
    
            plts[[length(plts)+1]] <- ggplot(pltdat,aes(x=.data[[xvar]],y=.data[[yvar]],col=.data[[labvars[v]]]))+
                rasterise(geom_point(size=0.2),dpi=300,dev="cairo")
            
            if(!(labvars[v] %in% c("AgeDeath","PMI","best_rin_PFC","BMI","sum_umi","sum_gene","expr_chrM","array_row","array_col"))){
                plts[[length(plts)]] <- plts[[length(plts)]]+
                    guides(col=guide_legend(byrow=T))+
                    guides(col=guide_legend(override.aes= list(size=1.5)))
            }
                #scale_color_discrete(DiscretePalette_scCustomize(num_colors = 2 + length(unique(colData(hyp2.tmp)$label)), palette = "varibow")[3:(length(unique(colData(hyp2.tmp)$label)) + 2)])
            
            ## determine appropriate subfolder for pllots -- github copilot came out swinging and filled most of this in near-perfectly!
        }
    }
    
    plts <- lapply(plts,FUN=function(x){
        x+theme(axis.text=element_blank(),axis.title=element_text(size=8),legend.text = element_text(size=7),legend.margin = margin(0,0.1,0,-0.1,unit = "in"),legend.key.height = unit(0.04,"in"))
    })
    
    cairo_pdf(paste0("plots/Manuscript/Supp Figs-Dimred/Supp File-",axlab," by xpt vars and clusts.pdf"),width=10.75,height=8.25,onefile=T)
    i <- 1
    for (i in seq(1,length(plts),4)){
        j <- ifelse(i>length(plts)-4,length(plts),i+3)
        do.call("grid.arrange",c(plts[c(i:j)],ncol=2))
    }
    dev.off()
    
    rm(i,plts,pltdat,xvar,yvar)
})

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] ggrastr_1.0.2               Cairo_1.6-2                
 [3] BiocParallel_1.36.0         parallelly_1.36.0          
 [5] default_1.0.0               gtools_3.9.5               
 [7] colorout_1.3-0.1            ggspavis_1.8.0             
 [9] SpatialExperiment_1.12.0    SingleCellExperiment_1.24.0
[11] SummarizedExperiment_1.32.0 Biobase_2.62.0             
[13] GenomicRanges_1.54.1        MatrixGenerics_1.14.0      
[15] matrixStats_1.2.0           gridExtra_2.3              
[17] ggplot2_3.4.4               Biostrings_2.70.2          
[19] GenomeInfoDb_1.38.5         XVector_0.42.0             
[21] IRanges_2.36.0              S4Vectors_0.40.2           
[23] BiocGenerics_0.48.1         data.table_1.14.10         
[25] rlang_1.1.3                

loaded via a namespace (and not attached):
  [1] later_1.3.2                   BiocIO_1.12.0                
  [3] bitops_1.0-7                  filelock_1.0.3               
  [5] fields_15.2                   tibble_3.2.1                 
  [7] XML_3.99-0.16.1               lifecycle_1.0.4              
  [9] edgeR_4.0.14                  doParallel_1.0.17            
 [11] rprojroot_2.0.4               lattice_0.22-5               
 [13] magrittr_2.0.3                limma_3.58.1                 
 [15] plotly_4.10.4                 sass_0.4.8                   
 [17] rmarkdown_2.25                jquerylib_0.1.4              
 [19] yaml_2.3.8                    httpuv_1.6.13                
 [21] ggside_0.2.3                  spam_2.10-0                  
 [23] sessioninfo_1.2.2             cowplot_1.1.3                
 [25] DBI_1.2.1                     RColorBrewer_1.1-3           
 [27] golem_0.4.1                   maps_3.4.2                   
 [29] abind_1.4-5                   zlibbioc_1.48.0              
 [31] purrr_1.0.2                   RCurl_1.98-1.14              
 [33] rappdirs_0.3.3                circlize_0.4.15              
 [35] GenomeInfoDbData_1.2.11       ggrepel_0.9.5                
 [37] irlba_2.3.5.1                 pheatmap_1.0.12              
 [39] DelayedMatrixStats_1.24.0     codetools_0.2-19             
 [41] DelayedArray_0.28.0           DT_0.31                      
 [43] scuttle_1.12.0                shape_1.4.6                  
 [45] tidyselect_1.2.0              farver_2.1.1                 
 [47] ScaledMatrix_1.10.0           viridis_0.6.5                
 [49] shinyWidgets_0.8.1            BiocFileCache_2.10.1         
 [51] GenomicAlignments_1.38.2      jsonlite_1.8.8               
 [53] GetoptLong_1.0.5              BiocNeighbors_1.20.2         
 [55] ellipsis_0.3.2                scater_1.30.1                
 [57] iterators_1.0.14              foreach_1.5.2                
 [59] tools_4.3.2                   Rcpp_1.0.12                  
 [61] glue_1.7.0                    SparseArray_1.2.3            
 [63] xfun_0.41                     here_1.0.1                   
 [65] dplyr_1.1.4                   withr_3.0.0                  
 [67] BiocManager_1.30.22           fastmap_1.1.1                
 [69] fansi_1.0.6                   digest_0.6.34                
 [71] rsvd_1.0.5                    R6_2.5.1                     
 [73] mime_0.12                     colorspace_2.1-0             
 [75] RSQLite_2.3.5                 config_0.3.2                 
 [77] utf8_1.2.4                    tidyr_1.3.1                  
 [79] generics_0.1.3                rtracklayer_1.62.0           
 [81] httr_1.4.7                    htmlwidgets_1.6.4            
 [83] S4Arrays_1.2.0                pkgconfig_2.0.3              
 [85] gtable_0.3.4                  blob_1.2.4                   
 [87] ComplexHeatmap_2.18.0         htmltools_0.5.7              
 [89] dotCall64_1.1-1               clue_0.3-65                  
 [91] scales_1.3.0                  png_0.1-8                    
 [93] attempt_0.3.1                 knitr_1.45                   
 [95] rstudioapi_0.15.0             rjson_0.2.21                 
 [97] curl_5.2.0                    cachem_1.0.8                 
 [99] GlobalOptions_0.1.2           BiocVersion_3.18.1           
[101] parallel_4.3.2                vipor_0.4.7                  
[103] AnnotationDbi_1.64.1          restfulr_0.0.15              
[105] pillar_1.9.0                  grid_4.3.2                   
[107] vctrs_0.6.5                   promises_1.2.1               
[109] BiocSingular_1.18.0           dbplyr_2.4.0                 
[111] beachmat_2.18.0               cluster_2.1.6                
[113] xtable_1.8-4                  beeswarm_0.4.0               
[115] paletteer_1.6.0               evaluate_0.23                
[117] magick_2.8.2                  cli_3.6.2                    
[119] locfit_1.5-9.8                compiler_4.3.2               
[121] Rsamtools_2.18.0              crayon_1.5.2                 
[123] labeling_0.4.3                rematch2_2.1.2               
[125] ggbeeswarm_0.7.2              viridisLite_0.4.2            
[127] munsell_0.5.0                 lazyeval_0.2.2               
[129] Matrix_1.6-5                  ExperimentHub_2.10.0         
[131] benchmarkme_1.0.8             sparseMatrixStats_1.14.0     
[133] bit64_4.0.5                   KEGGREST_1.42.0              
[135] statmod_1.5.0                 shiny_1.8.0                  
[137] interactiveDisplayBase_1.40.0 AnnotationHub_3.10.0         
[139] memoise_2.0.1                 bslib_0.6.1                  
[141] benchmarkmeData_1.0.4         bit_4.0.5                    
[143] spatialLIBD_1.14.1           
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-02-15
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package                * version     date (UTC) lib source
 abind                    1.4-5       2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1      2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0      2023-10-24 [1] Bioconductor
 attempt                  0.3.1       2020-05-03 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0      2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0       2021-06-01 [1] CRAN (R 4.3.2)
 benchmarkme              1.0.8       2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4       2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0      2023-10-24 [1] Bioconductor
 BiocFileCache            2.10.1      2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1      2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0      2023-10-26 [1] Bioconductor
 BiocManager              1.30.22     2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.2      2024-01-07 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel           * 1.36.0      2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0      2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1      2023-11-18 [1] Bioconductor 3.18 (R 4.3.2)
 Biostrings             * 2.70.2      2024-01-28 [1] Bioconductor 3.18 (R 4.3.2)
 bit                      4.0.5       2022-11-15 [1] CRAN (R 4.3.2)
 bit64                    4.0.5       2020-08-30 [1] CRAN (R 4.3.2)
 bitops                   1.0-7       2021-04-24 [1] CRAN (R 4.3.2)
 blob                     1.2.4       2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.6.1       2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8       2023-05-01 [1] CRAN (R 4.3.2)
 Cairo                  * 1.6-2       2023-11-28 [1] CRAN (R 4.3.1)
 circlize                 0.4.15      2022-05-10 [1] CRAN (R 4.3.0)
 cli                      3.6.2       2023-12-11 [1] CRAN (R 4.3.2)
 clue                     0.3-65      2023-09-23 [1] CRAN (R 4.3.2)
 cluster                  2.1.6       2023-12-01 [1] CRAN (R 4.3.2)
 codetools                0.2-19      2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0.1     2024-01-11 [1] local
 colorspace               2.1-0       2023-01-23 [1] CRAN (R 4.3.2)
 ComplexHeatmap           2.18.0      2023-10-26 [1] Bioconductor
 config                   0.3.2       2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.3       2024-01-22 [1] CRAN (R 4.3.1)
 crayon                   1.5.2       2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0       2023-12-08 [1] CRAN (R 4.3.2)
 data.table             * 1.14.10     2023-12-08 [1] CRAN (R 4.3.2)
 DBI                      1.2.1       2024-01-12 [1] CRAN (R 4.3.2)
 dbplyr                   2.4.0       2023-10-26 [1] CRAN (R 4.3.1)
 default                * 1.0.0       2017-08-07 [1] CRAN (R 4.3.0)
 DelayedArray             0.28.0      2023-10-24 [1] Bioconductor
 DelayedMatrixStats       1.24.0      2023-11-06 [1] Bioconductor
 digest                   0.6.34      2024-01-11 [1] CRAN (R 4.3.2)
 doParallel               1.0.17      2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1       2023-11-28 [1] CRAN (R 4.3.2)
 dplyr                    1.1.4       2023-11-17 [1] CRAN (R 4.3.2)
 DT                       0.31        2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.14      2024-01-29 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2       2021-04-29 [1] CRAN (R 4.3.2)
 evaluate                 0.23        2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0      2023-10-26 [1] Bioconductor
 fansi                    1.0.6       2023-12-08 [1] CRAN (R 4.3.2)
 farver                   2.1.1       2022-07-06 [1] CRAN (R 4.3.2)
 fastmap                  1.1.1       2023-02-24 [1] CRAN (R 4.3.2)
 fields                   15.2        2023-08-17 [1] CRAN (R 4.3.2)
 filelock                 1.0.3       2023-12-11 [1] CRAN (R 4.3.2)
 foreach                  1.5.2       2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3       2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.5      2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData         1.2.11      2024-01-11 [1] Bioconductor
 GenomicAlignments        1.38.2      2024-01-16 [1] Bioconductor 3.18 (R 4.3.2)
 GenomicRanges          * 1.54.1      2023-10-29 [1] Bioconductor
 GetoptLong               1.0.5       2020-12-15 [1] CRAN (R 4.3.0)
 ggbeeswarm               0.7.2       2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4       2023-10-12 [1] CRAN (R 4.3.1)
 ggrastr                * 1.0.2       2023-06-01 [1] CRAN (R 4.3.0)
 ggrepel                  0.9.5       2024-01-10 [1] CRAN (R 4.3.2)
 ggside                   0.2.3       2023-12-10 [1] CRAN (R 4.3.1)
 ggspavis               * 1.8.0       2023-10-26 [1] Bioconductor
 GlobalOptions            0.1.2       2020-06-10 [1] CRAN (R 4.3.0)
 glue                     1.7.0       2024-01-09 [1] CRAN (R 4.3.2)
 golem                    0.4.1       2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3         2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4       2023-08-21 [1] CRAN (R 4.3.0)
 gtools                 * 3.9.5       2023-11-20 [1] CRAN (R 4.3.2)
 here                     1.0.1       2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7       2023-11-03 [1] CRAN (R 4.3.2)
 htmlwidgets              1.6.4       2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13      2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7       2023-08-15 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0      2023-10-26 [1] Bioconductor
 IRanges                * 2.36.0      2023-10-24 [1] Bioconductor
 irlba                    2.3.5.1     2022-10-03 [1] CRAN (R 4.3.2)
 iterators                1.0.14      2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4       2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8       2023-12-04 [1] CRAN (R 4.3.2)
 KEGGREST                 1.42.0      2023-10-26 [1] Bioconductor
 knitr                    1.45        2023-10-30 [1] CRAN (R 4.3.2)
 labeling                 0.4.3       2023-08-29 [1] CRAN (R 4.3.0)
 later                    1.3.2       2023-12-06 [1] CRAN (R 4.3.2)
 lattice                  0.22-5      2023-10-24 [1] CRAN (R 4.3.2)
 lazyeval                 0.2.2       2019-03-15 [1] CRAN (R 4.3.2)
 lifecycle                1.0.4       2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1      2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8     2023-06-11 [1] CRAN (R 4.3.2)
 magick                   2.8.2       2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3       2022-03-30 [1] CRAN (R 4.3.2)
 maps                     3.4.2       2023-12-15 [1] CRAN (R 4.3.1)
 Matrix                   1.6-5       2024-01-11 [1] CRAN (R 4.3.2)
 MatrixGenerics         * 1.14.0      2023-10-26 [1] Bioconductor
 matrixStats            * 1.2.0       2023-12-11 [1] CRAN (R 4.3.2)
 memoise                  2.0.1       2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12        2021-09-28 [1] CRAN (R 4.3.2)
 munsell                  0.5.0       2018-06-12 [1] CRAN (R 4.3.0)
 paletteer                1.6.0       2024-01-21 [1] CRAN (R 4.3.1)
 parallelly             * 1.36.0      2023-05-26 [1] CRAN (R 4.3.0)
 pheatmap                 1.0.12      2019-01-04 [1] CRAN (R 4.3.0)
 pillar                   1.9.0       2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3       2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.4      2024-01-13 [1] CRAN (R 4.3.1)
 png                      0.1-8       2022-11-29 [1] CRAN (R 4.3.2)
 promises                 1.2.1       2023-08-10 [1] CRAN (R 4.3.2)
 purrr                    1.0.2       2023-08-10 [1] CRAN (R 4.3.2)
 R6                       2.5.1       2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3       2021-01-31 [1] CRAN (R 4.3.2)
 RColorBrewer             1.1-3       2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.12      2024-01-09 [1] CRAN (R 4.3.2)
 RCurl                    1.98-1.14   2024-01-09 [1] CRAN (R 4.3.1)
 rematch2                 2.1.2       2020-05-01 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15      2022-06-16 [1] CRAN (R 4.3.2)
 rjson                    0.2.21      2022-01-09 [1] CRAN (R 4.3.2)
 rlang                  * 1.1.3       2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown                2.25        2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot                2.0.4       2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0      2023-10-24 [1] Bioconductor
 RSQLite                  2.3.5       2024-01-21 [1] CRAN (R 4.3.2)
 rstudioapi               0.15.0      2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5       2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0      2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0       2023-10-24 [1] Bioconductor
 S4Vectors              * 0.40.2      2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8       2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0      2023-11-06 [1] Bioconductor
 scales                   1.3.0       2023-11-28 [1] CRAN (R 4.3.2)
 scater                   1.30.1      2023-11-16 [1] Bioconductor
 scuttle                  1.12.0      2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2       2021-12-06 [1] CRAN (R 4.3.0)
 shape                    1.4.6       2021-05-19 [1] CRAN (R 4.3.0)
 shiny                    1.8.0       2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.1       2024-01-10 [1] CRAN (R 4.3.1)
 SingleCellExperiment   * 1.24.0      2023-11-06 [1] Bioconductor
 spam                     2.10-0      2023-10-23 [1] CRAN (R 4.3.2)
 SparseArray              1.2.3       2023-12-25 [1] Bioconductor 3.18 (R 4.3.2)
 sparseMatrixStats        1.14.0      2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0      2023-10-26 [1] Bioconductor
 spatialLIBD              1.14.1      2023-11-30 [1] Bioconductor 3.18 (R 4.3.2)
 statmod                  1.5.0       2023-01-06 [1] CRAN (R 4.3.2)
 SummarizedExperiment   * 1.32.0      2023-11-06 [1] Bioconductor
 tibble                   3.2.1       2023-03-20 [1] CRAN (R 4.3.2)
 tidyr                    1.3.1       2024-01-24 [1] CRAN (R 4.3.1)
 tidyselect               1.2.0       2022-10-10 [1] CRAN (R 4.3.0)
 utf8                     1.2.4       2023-10-22 [1] CRAN (R 4.3.2)
 vctrs                    0.6.5       2023-12-01 [1] CRAN (R 4.3.2)
 vipor                    0.4.7       2023-12-18 [1] CRAN (R 4.3.1)
 viridis                  0.6.5       2024-01-29 [1] CRAN (R 4.3.2)
 viridisLite              0.4.2       2023-05-02 [1] CRAN (R 4.3.0)
 withr                    3.0.0       2024-01-16 [1] CRAN (R 4.3.1)
 xfun                     0.41        2023-11-01 [1] CRAN (R 4.3.2)
 XML                      3.99-0.16.1 2024-01-22 [1] CRAN (R 4.3.1)
 xtable                   1.8-4       2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0      2023-10-24 [1] Bioconductor
 yaml                     2.3.8       2023-12-11 [1] CRAN (R 4.3.2)
 zlibbioc                 1.48.0      2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
