---
title: "01-Feature selection"
author: "Bernard Mulvey"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different:
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater)
library(nnSVG)
library(BiocParallel)
library(scran)
library(parallel)
library(harmony)
library(batchelor) ## MNN

# ggplot defaults:
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```


# Load pre-processed SPE and demographics; attach to colData
```{r}
hyp2 <- readRDS("data/01-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

# make sure colnames of the spe are unique (i.e., colData(hyp2)$key)
colnames(hyp2) <- hyp2$key

# remove 10x space ranger cluster columns, since they start with numerals which some packages don't like in colnames
colData(hyp2) <- colData(hyp2)[,!grepl("^\\d",colnames(colData(hyp2)))]

demos <- fread("raw-data/demos.txt")
demos <- demos[BrNum %in% colData(hyp2)$brnum]
demos[,sample_id:=NULL]
demos <- unique(demos)
setnames(demos,"BrNum","brnum")
tmp <- merge.data.table(demos,as.data.table(colData(hyp2)),by="brnum")
# k. also fix "Best Rin PFC" and "BMI (calculated)" to not have spaces
setnames(tmp,c("Best RIN PFC","BMI (calculated)"),c("best_rin_pfc","bmi"))

# ^ has no rownames because it's a data table; change to DataFrame and make rownames = key in the merged table. reorder to match the column order in hyp2.
tmp <- DataFrame(tmp)
rownames(tmp) <- tmp$key
tmp <- tmp[colnames(hyp2),]
stopifnot(sum(rownames(tmp)==colnames(hyp2))==ncol(hyp2))

### add the demographic'd and cleaned colData to the SPE
colData(hyp2) <- tmp

rm(tmp,demos)
```

### mean-variance relationships: blocking on sample_id (not donor)
```{r}
hyp2.vars <- modelGeneVar(hyp2,block=hyp2$sample_id)

varperblock <- hyp2.vars$per.block

# examine per-sample (block) variances by plotting each mean-variance curve and storing in a list to arrange later as a single output
plts <- list()

i <- 1
for (i in c(1:length(varperblock))){
    hyp2.vars.dt <- cbind(as.data.table(as.character(rownames(varperblock[[i]]))),varperblock[[i]]$total)
    hyp2.means.dt <- cbind(as.data.table(as.character(rownames(varperblock[[i]]))),varperblock[[i]]$mean)

    setnames(hyp2.vars.dt,2,"var_log_xpr")
    setnames(hyp2.means.dt,2,"mean_log_xpr")
    hyp2.mnvar.pltdat <- merge(hyp2.means.dt,hyp2.vars.dt,by="V1")
    plts[[i]] <- ggplot(hyp2.mnvar.pltdat, aes(x = mean_log_xpr, y = var_log_xpr)) +
        geom_point() +
        geom_function(fun=metadata(varperblock[[i]])$trend,col="blue")+
        ggtitle(paste0("Mean-var curve-",names(varperblock)[i],"\n sample_id blocking"))
    rm(hyp2.mnvar.pltdat,hyp2.vars.dt,hyp2.means.dt)
}

## arrange the per sample plots generated above into a single output
pdf("plots/02-featureselection-harmony-mnn/samplewiseblocking_meanvars.pdf",height=3*length(unique(hyp2$sample_id))+2,width=10)
do.call("grid.arrange",c(plts,ncol=2))
dev.off()

## remove clutter from workspace
rm(plts,varperblock,i)
```

# get top 10, 20 %ile HVGs from sample_id-blocking'd variances
```{r}
hyp2.hvg10 <- scran::getTopHVGs(hyp2.vars,prop = 0.1)
hyp2.hvg20 <- scran::getTopHVGs(hyp2.vars,prop=0.2)
```

#### run nnSVG: does not take batch variables nor does it take in reduced dims. run samplewise.

### the workhorse function nnSVG can use parallel processing through BiocParallel. Set up the preferred (i.e., compatible) param. the parallelization is within the nnsvg call, so no need to match cpu # to sample #
```{r}
# sbp <- MulticoreParam(10)
# register(sbp)

# SUPER CRITICAL IMPORTANT NOTES: 
# 1. RECALCULATE LOG COUNTS AFTER FILTERING ^
# 2.nnSVG works ONE SAMPLE (one capture area) at a time. So need to wrap this in a loop. 
# 3. as such, perform filtering PER SAMPLE. 
# filter_genes_ncounts (N) and filter_genes_pcspots (S) work together: genes will be filtered to those with at least N counts in at least S% of spots.  filter_genes_pcspots expects a %age as 0-100, NOT a fraction.
# tried N=2 (instead of 3) and S = 0.25 instead of 0.5, since most spots are pretty diverse and high-depth (and the cell type diversity is also unknown), but didn't get past the first sample in 3 hours on 10 CPUs. so now trying 3, 0.5 .

caps <- unique(hyp2$sample_id)
res_list <- as.list(rep(NA, length(caps)))
names(res_list) <- caps

i<-1
for (i in c(1:length(caps))){
    #setup
    cap <- caps[i]
    message(paste0("Starting sample ",i," of ",length(caps)))
    hyp2.samp <- hyp2[,colData(hyp2)$sample_id==cap]
    
    hyp2.samp <- filter_genes(hyp2.samp,filter_genes_ncounts = 3,filter_genes_pcspots = 0.5,filter_mito = F)
    
    #recalculate logcounts after filtering for nnsvg-able genes
    hyp2.samp <- computeLibraryFactors(hyp2.samp)
    hyp2.samp <- logNormCounts(hyp2.samp)
    
    # main call -- defaults to logcounts but shown for posterity; X if using covariates; n_neighbors defaults to 10, slower but potentially more granular with higher values eg 15, order = don't touch unless you have < 70 spots total; n_threads or BPPARAM for parallelization
    hyp2.samp <- nnSVG(hyp2.samp,assay_name = "logcounts",n_threads = 10)
    
    # extract results; get index of the last column of rowdata in the full spe to determine where nnSVG-appended columns begin; add nnSVG_ to those column names so we can tell them apart
    ind <- ncol(rowData(hyp2))+1
    colnames(rowData(hyp2.samp))[ind:ncol(rowData(hyp2.samp))] <- paste0("nnSVG_",colnames(rowData(hyp2.samp))[ind:ncol(rowData(hyp2.samp))])

    # extract the nnsvg results from the rowData and store it in the result list. we don't need the rownames for storage since they're redundant with the ensembl identifier column in rowdata.
    res_list[[i]] <- as.data.table(as.data.frame(rowData(hyp2.samp)),keep.rownames=F)
    names(res_list)[i] <- cap
    rm(hyp2.samp,cap,ind)
    ## clean up memory after each sample
    gc(full=T)
}

saveRDS(res_list,"data/02-feature_selection/01-hyp2n9_nnSVG_samplewise_n3s0.5.RDS")

## make an environment copy of the nnsvg results with a more descriptive name we can come back to
nnsvg_N3Spt5 <- res_list

## further cleanup
rm(i,caps,sbp,res_list)
gc(full=T)
```

### get the mean nnSVG rank for each gene among samples where the gene was nominally significant for spatial variability.
```{r}
i<-1
for (i in c(1:sum(unique(colData(hyp2)$sample_id)%in%names(nnsvg_N3Spt5)))){
    if(i==1){
        nnsvg_ranks <- nnsvg_N3Spt5[[i]][nnSVG_padj<0.05]
        nnsvg_ranks <- nnsvg_ranks[,.(gene_id,gene_name,nnSVG_rank)]
        setnames(nnsvg_ranks,"nnSVG_rank",names(nnsvg_N3Spt5)[i])
    }
    else{
        tmp <- nnsvg_N3Spt5[[i]][nnSVG_padj<0.05]
        tmp <- tmp[,.(gene_id,gene_name,nnSVG_rank)]
        setnames(tmp,"nnSVG_rank",names(nnsvg_N3Spt5)[i])
        nnsvg_ranks <- merge.data.table(nnsvg_ranks, tmp, by=c("gene_id","gene_name"), all.x=T, all.y=T, incomparables = NA)
        rm(tmp)
    }
}
rm(i)

nnsvg_ranks[,meanrank:=rowMeans(.SD,na.rm=T),.SDcols=c(3:ncol(nnsvg_ranks))]
nnsvg_ranks <- nnsvg_ranks[,samples_signif:=rowSums(!is.na(.SD)),.SDcols=c(3:ncol(nnsvg_ranks))]

fwrite(nnsvg_ranks,"analysis/data/02-feature_selection/01-hyp2n9_meangeneranks_nnSVG_samplewise_n3s0.5.txt",sep='\t',quote=F)
```

### retrieve top M or N nnSVG genes (where M = number of genes in the HVG 10%ile set and N = number in the top 20%ile HVG set); save each vector of variable genes.
```{r}
nnsvg_ranks <- nnsvg_ranks[samples_signif>3]
setorderv(nnsvg_ranks,"meanrank",1) # ascending order
nnsvg10 <- nnsvg_ranks[c(1:length(hyp2.hvg10)),gene_id]
nnsvg20 <- nnsvg_ranks[c(1:length(hyp2.hvg20)),gene_id]

saveRDS(hyp2.hvg10,"analysis/data/02-feature_selection/01-hyp2n9_HVG10pctile_sample-blocked.RDS")
saveRDS(hyp2.hvg20,"analysis/data/02-feature_selection/01-hyp2n9_HVG20pctile_sample-blocked.RDS")
saveRDS(nnsvg10,"analysis/data/02-feature_selection/01-hyp2n9_nnsvg10pctile.RDS")
saveRDS(nnsvg20,"analysis/data/02-feature_selection/01-hyp2n9_nnsvg20pctile.RDS")
```

# session info follows call #
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.1 (2023-06-16)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets 
[8] methods   base     

other attached packages:
 [1] batchelor_1.18.0            harmony_1.2.0              
 [3] Rcpp_1.0.11                 nnSVG_1.6.0                
 [5] colorout_1.3-0              Biostrings_2.70.1          
 [7] XVector_0.42.0              gridExtra_2.3              
 [9] scran_1.30.0                BiocParallel_1.36.0        
[11] scater_1.30.1               scuttle_1.12.0             
[13] spatialLIBD_1.13.4          ggspavis_1.8.0             
[15] ggplot2_3.4.4               SpatialExperiment_1.12.0   
[17] SingleCellExperiment_1.24.0 SummarizedExperiment_1.32.0
[19] Biobase_2.62.0              GenomicRanges_1.54.1       
[21] GenomeInfoDb_1.38.1         IRanges_2.36.0             
[23] S4Vectors_0.40.1            BiocGenerics_0.48.1        
[25] MatrixGenerics_1.14.0       matrixStats_1.1.0          
[27] data.table_1.14.8           rlang_1.1.2                

loaded via a namespace (and not attached):
  [1] later_1.3.1                   BiocIO_1.12.0                
  [3] bitops_1.0-7                  filelock_1.0.2               
  [5] fields_15.2                   tibble_3.2.1                 
  [7] XML_3.99-0.15                 lifecycle_1.0.4              
  [9] edgeR_4.0.1                   doParallel_1.0.17            
 [11] rprojroot_2.0.4               lattice_0.22-5               
 [13] magrittr_2.0.3                limma_3.58.1                 
 [15] plotly_4.10.3                 sass_0.4.7                   
 [17] rmarkdown_2.25                jquerylib_0.1.4              
 [19] yaml_2.3.7                    metapod_1.10.0               
 [21] httpuv_1.6.12                 spam_2.10-0                  
 [23] ggside_0.2.2                  sessioninfo_1.2.2            
 [25] pbapply_1.7-2                 cowplot_1.1.1                
 [27] DBI_1.1.3                     RColorBrewer_1.1-3           
 [29] ResidualMatrix_1.12.0         golem_0.4.1                  
 [31] maps_3.4.1.1                  abind_1.4-5                  
 [33] zlibbioc_1.48.0               purrr_1.0.2                  
 [35] RCurl_1.98-1.13               rappdirs_0.3.3               
 [37] GenomeInfoDbData_1.2.11       ggrepel_0.9.4                
 [39] irlba_2.3.5.1                 dqrng_0.3.1                  
 [41] DelayedMatrixStats_1.24.0     codetools_0.2-19             
 [43] DelayedArray_0.28.0           DT_0.30                      
 [45] tidyselect_1.2.0              ScaledMatrix_1.10.0          
 [47] viridis_0.6.4                 shinyWidgets_0.8.0           
 [49] BiocFileCache_2.10.1          GenomicAlignments_1.38.0     
 [51] jsonlite_1.8.7                BiocNeighbors_1.20.0         
 [53] ellipsis_0.3.2                iterators_1.0.14             
 [55] foreach_1.5.2                 tools_4.3.1                  
 [57] glue_1.6.2                    SparseArray_1.2.2            
 [59] xfun_0.41                     here_1.0.1                   
 [61] dplyr_1.1.4                   withr_2.5.2                  
 [63] BiocManager_1.30.22           fastmap_1.1.1                
 [65] bluster_1.12.0                fansi_1.0.5                  
 [67] digest_0.6.33                 rsvd_1.0.5                   
 [69] R6_2.5.1                      mime_0.12                    
 [71] colorspace_2.1-0              RSQLite_2.3.3                
 [73] config_0.3.2                  utf8_1.2.4                   
 [75] tidyr_1.3.0                   generics_0.1.3               
 [77] rtracklayer_1.62.0            httr_1.4.7                   
 [79] htmlwidgets_1.6.2             S4Arrays_1.2.0               
 [81] pkgconfig_2.0.3               gtable_0.3.4                 
 [83] rdist_0.0.5                   blob_1.2.4                   
 [85] BRISC_1.0.5                   htmltools_0.5.7              
 [87] dotCall64_1.1-0               scales_1.2.1                 
 [89] png_0.1-8                     attempt_0.3.1                
 [91] knitr_1.45                    rstudioapi_0.15.0            
 [93] rjson_0.2.21                  curl_5.1.0                   
 [95] cachem_1.0.8                  BiocVersion_3.18.1           
 [97] vipor_0.4.5                   AnnotationDbi_1.64.1         
 [99] restfulr_0.0.15               pillar_1.9.0                 
[101] grid_4.3.1                    vctrs_0.6.4                  
[103] RANN_2.6.1                    promises_1.2.1               
[105] BiocSingular_1.18.0           dbplyr_2.4.0                 
[107] beachmat_2.18.0               cluster_2.1.4                
[109] xtable_1.8-4                  beeswarm_0.4.0               
[111] paletteer_1.5.0               evaluate_0.23                
[113] magick_2.8.1                  cli_3.6.1                    
[115] locfit_1.5-9.8                compiler_4.3.1               
[117] Rsamtools_2.18.0              crayon_1.5.2                 
[119] rematch2_2.1.2                ggbeeswarm_0.7.2             
[121] viridisLite_0.4.2             munsell_0.5.0                
[123] lazyeval_0.2.2                Matrix_1.6-3                 
[125] ExperimentHub_2.10.0          benchmarkme_1.0.8            
[127] sparseMatrixStats_1.14.0      bit64_4.0.5                  
[129] KEGGREST_1.42.0               statmod_1.5.0                
[131] shiny_1.8.0                   interactiveDisplayBase_1.40.0
[133] AnnotationHub_3.10.0          igraph_1.5.1                 
[135] memoise_2.0.1                 bslib_0.5.1                  
[137] benchmarkmeData_1.0.4         bit_4.0.5                    

##

─ Session info ────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.1 (2023-06-16)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2023-11-21
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────
 package                * version   date (UTC) lib source
 abind                    1.4-5     2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1    2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0    2023-10-24 [1] Bioconductor
 attempt                  0.3.1     2020-05-03 [1] CRAN (R 4.3.0)
 batchelor              * 1.18.0    2023-10-24 [1] Bioconductor
 beachmat                 2.18.0    2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0     2021-06-01 [1] CRAN (R 4.3.0)
 benchmarkme              1.0.8     2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4     2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0    2023-10-26 [1] Bioconductor
 BiocFileCache            2.10.1    2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1    2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0    2023-10-24 [1] Bioconductor
 BiocManager              1.30.22   2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.0    2023-10-24 [1] Bioconductor
 BiocParallel           * 1.36.0    2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0    2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1    2023-11-15 [1] Bioconductor
 Biostrings             * 2.70.1    2023-10-26 [1] Bioconductor
 bit                      4.0.5     2022-11-15 [1] CRAN (R 4.3.0)
 bit64                    4.0.5     2020-08-30 [1] CRAN (R 4.3.0)
 bitops                   1.0-7     2021-04-24 [1] CRAN (R 4.3.0)
 blob                     1.2.4     2023-03-17 [1] CRAN (R 4.3.0)
 bluster                  1.12.0    2023-10-24 [1] Bioconductor
 BRISC                    1.0.5     2022-04-29 [1] CRAN (R 4.3.0)
 bslib                    0.5.1     2023-08-11 [1] CRAN (R 4.3.0)
 cachem                   1.0.8     2023-05-01 [1] CRAN (R 4.3.0)
 cli                      3.6.1     2023-03-23 [1] CRAN (R 4.3.0)
 cluster                  2.1.4     2022-08-22 [1] CRAN (R 4.3.0)
 codetools                0.2-19    2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0     2023-11-01 [1] local (/Users/bmulvey/Desktop/colorout)
 colorspace               2.1-0     2023-01-23 [1] CRAN (R 4.3.0)
 config                   0.3.2     2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.1     2020-12-30 [1] CRAN (R 4.3.0)
 crayon                   1.5.2     2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.1.0     2023-10-02 [1] CRAN (R 4.3.1)
 data.table             * 1.14.8    2023-02-17 [1] CRAN (R 4.3.0)
 DBI                      1.1.3     2022-06-18 [1] CRAN (R 4.3.0)
 dbplyr                   2.4.0     2023-10-26 [1] CRAN (R 4.3.1)
 DelayedArray             0.28.0    2023-11-06 [1] Bioconductor
 DelayedMatrixStats       1.24.0    2023-10-24 [1] Bioconductor
 digest                   0.6.33    2023-07-07 [1] CRAN (R 4.3.0)
 doParallel               1.0.17    2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-0     2023-10-17 [1] CRAN (R 4.3.1)
 dplyr                    1.1.4     2023-11-17 [1] CRAN (R 4.3.1)
 dqrng                    0.3.1     2023-08-30 [1] CRAN (R 4.3.0)
 DT                       0.30      2023-10-05 [1] CRAN (R 4.3.1)
 edgeR                    4.0.1     2023-10-29 [1] Bioconductor
 ellipsis                 0.3.2     2021-04-29 [1] CRAN (R 4.3.0)
 evaluate                 0.23      2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0    2023-10-24 [1] Bioconductor
 fansi                    1.0.5     2023-10-08 [1] CRAN (R 4.3.1)
 fastmap                  1.1.1     2023-02-24 [1] CRAN (R 4.3.0)
 fields                   15.2      2023-08-17 [1] CRAN (R 4.3.0)
 filelock                 1.0.2     2018-10-05 [1] CRAN (R 4.3.0)
 foreach                  1.5.2     2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3     2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.1    2023-11-08 [1] Bioconductor
 GenomeInfoDbData         1.2.11    2023-11-10 [1] Bioconductor
 GenomicAlignments        1.38.0    2023-10-24 [1] Bioconductor
 GenomicRanges          * 1.54.1    2023-10-30 [1] Bioconductor
 ggbeeswarm               0.7.2     2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4     2023-10-12 [1] CRAN (R 4.3.1)
 ggrepel                  0.9.4     2023-10-13 [1] CRAN (R 4.3.1)
 ggside                   0.2.2     2022-12-04 [1] CRAN (R 4.3.0)
 ggspavis               * 1.8.0     2023-10-24 [1] Bioconductor
 glue                     1.6.2     2022-02-24 [1] CRAN (R 4.3.0)
 golem                    0.4.1     2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra              * 2.3       2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4     2023-08-21 [1] CRAN (R 4.3.0)
 harmony                * 1.2.0     2023-11-16 [1] Github (immunogenomics/harmony@f054b03)
 here                     1.0.1     2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7     2023-11-03 [1] CRAN (R 4.3.1)
 htmlwidgets              1.6.2     2023-03-17 [1] CRAN (R 4.3.0)
 httpuv                   1.6.12    2023-10-23 [1] CRAN (R 4.3.1)
 httr                     1.4.7     2023-08-15 [1] CRAN (R 4.3.0)
 igraph                   1.5.1     2023-08-10 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0    2023-10-24 [1] Bioconductor
 IRanges                * 2.36.0    2023-10-26 [1] Bioconductor
 irlba                    2.3.5.1   2022-10-03 [1] CRAN (R 4.3.1)
 iterators                1.0.14    2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4     2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.7     2023-06-29 [1] CRAN (R 4.3.0)
 KEGGREST                 1.42.0    2023-10-26 [1] Bioconductor
 knitr                    1.45      2023-10-30 [1] CRAN (R 4.3.1)
 later                    1.3.1     2023-05-02 [1] CRAN (R 4.3.0)
 lattice                  0.22-5    2023-10-24 [1] CRAN (R 4.3.1)
 lazyeval                 0.2.2     2019-03-15 [1] CRAN (R 4.3.0)
 lifecycle                1.0.4     2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1    2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8   2023-06-11 [1] CRAN (R 4.3.0)
 magick                   2.8.1     2023-10-22 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3     2022-03-30 [1] CRAN (R 4.3.0)
 maps                     3.4.1.1   2023-11-03 [1] CRAN (R 4.3.1)
 Matrix                   1.6-3     2023-11-14 [1] CRAN (R 4.3.1)
 MatrixGenerics         * 1.14.0    2023-10-26 [1] Bioconductor
 matrixStats            * 1.1.0     2023-11-07 [1] CRAN (R 4.3.1)
 memoise                  2.0.1     2021-11-26 [1] CRAN (R 4.3.0)
 metapod                  1.10.0    2023-10-24 [1] Bioconductor
 mime                     0.12      2021-09-28 [1] CRAN (R 4.3.0)
 munsell                  0.5.0     2018-06-12 [1] CRAN (R 4.3.0)
 nnSVG                  * 1.6.0     2023-10-24 [1] Bioconductor
 paletteer                1.5.0     2022-10-19 [1] CRAN (R 4.3.0)
 pbapply                  1.7-2     2023-06-27 [1] CRAN (R 4.3.0)
 pillar                   1.9.0     2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3     2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.3    2023-10-21 [1] CRAN (R 4.3.1)
 png                      0.1-8     2022-11-29 [1] CRAN (R 4.3.0)
 promises                 1.2.1     2023-08-10 [1] CRAN (R 4.3.0)
 purrr                    1.0.2     2023-08-10 [1] CRAN (R 4.3.0)
 R6                       2.5.1     2021-08-19 [1] CRAN (R 4.3.0)
 RANN                     2.6.1     2019-01-08 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3     2021-01-31 [1] CRAN (R 4.3.0)
 RColorBrewer             1.1-3     2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                   * 1.0.11    2023-07-06 [1] CRAN (R 4.3.1)
 RCurl                    1.98-1.13 2023-11-02 [1] CRAN (R 4.3.1)
 rdist                    0.0.5     2020-05-04 [1] CRAN (R 4.3.0)
 rematch2                 2.1.2     2020-05-01 [1] CRAN (R 4.3.0)
 ResidualMatrix           1.12.0    2023-10-24 [1] Bioconductor
 restfulr                 0.0.15    2022-06-16 [1] CRAN (R 4.3.0)
 rjson                    0.2.21    2022-01-09 [1] CRAN (R 4.3.0)
 rlang                  * 1.1.2     2023-11-04 [1] CRAN (R 4.3.1)
 rmarkdown                2.25      2023-09-18 [1] CRAN (R 4.3.1)
 rprojroot                2.0.4     2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0    2023-10-24 [1] Bioconductor
 RSQLite                  2.3.3     2023-11-04 [1] CRAN (R 4.3.1)
 rstudioapi               0.15.0    2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5     2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0    2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0     2023-10-26 [1] Bioconductor
 S4Vectors              * 0.40.1    2023-10-26 [1] Bioconductor
 sass                     0.4.7     2023-07-15 [1] CRAN (R 4.3.0)
 ScaledMatrix             1.10.0    2023-10-24 [1] Bioconductor
 scales                   1.2.1     2022-08-20 [1] CRAN (R 4.3.0)
 scater                 * 1.30.1    2023-11-16 [1] Bioconductor
 scran                  * 1.30.0    2023-10-24 [1] Bioconductor
 scuttle                * 1.12.0    2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2     2021-12-06 [1] CRAN (R 4.3.0)
 shiny                    1.8.0     2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.0     2023-08-30 [1] CRAN (R 4.3.0)
 SingleCellExperiment   * 1.24.0    2023-11-06 [1] Bioconductor
 spam                     2.10-0    2023-10-23 [1] CRAN (R 4.3.1)
 SparseArray              1.2.2     2023-11-08 [1] Bioconductor
 sparseMatrixStats        1.14.0    2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0    2023-10-24 [1] Bioconductor
 spatialLIBD            * 1.13.4    2023-05-25 [1] Bioconductor
 statmod                  1.5.0     2023-01-06 [1] CRAN (R 4.3.0)
 SummarizedExperiment   * 1.32.0    2023-11-06 [1] Bioconductor
 tibble                   3.2.1     2023-03-20 [1] CRAN (R 4.3.0)
 tidyr                    1.3.0     2023-01-24 [1] CRAN (R 4.3.0)
 tidyselect               1.2.0     2022-10-10 [1] CRAN (R 4.3.0)
 utf8                     1.2.4     2023-10-22 [1] CRAN (R 4.3.1)
 vctrs                    0.6.4     2023-10-12 [1] CRAN (R 4.3.1)
 vipor                    0.4.5     2017-03-22 [1] CRAN (R 4.3.0)
 viridis                  0.6.4     2023-07-22 [1] CRAN (R 4.3.0)
 viridisLite              0.4.2     2023-05-02 [1] CRAN (R 4.3.0)
 withr                    2.5.2     2023-10-30 [1] CRAN (R 4.3.1)
 xfun                     0.41      2023-11-01 [1] CRAN (R 4.3.1)
 XML                      3.99-0.15 2023-11-02 [1] CRAN (R 4.3.1)
 xtable                   1.8-4     2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0    2023-10-26 [1] Bioconductor
 yaml                     2.3.7     2023-01-23 [1] CRAN (R 4.3.0)
 zlibbioc                 1.48.0    2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

───────────────────────────────────────────────────────────────────────────
