---
title: "02-Cluster on Several Feature Sets"
author: "Bernie Mulvey"
date: "2023-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater) # addPerCellQC
# library(nnSVG)
library(BiocParallel)
library(scran)
library(parallel)
library(Polychrome)
# library(fasthplus)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm.RDS")
hyp2.vgs.list <- readRDS("analysis/data/01-feature_selection/nnSVG_HVG_and_combo_featurelists.RDS")

sbp <- MulticoreParam(workers=7)
register(sbp)

pcalist <- bpmapply(X=hyp2.vgs.list,Y=as.list(names(hyp2.vgs.list)),BPPARAM = sbp,SIMPLIFY=FALSE,FUN = function(X,Y){
  set.seed(42)
    # 3 reduced dim tables are already in the object, so grab #4 (newly created)
  tmp <- scater::runPCA(hyp2,name = paste0("PCA"),subset_row=X)
  return(reducedDims(tmp)[[4]])
})
stopifnot(sum(rownames(pcalist[[1]])==colnames(hyp2))==ncol(hyp2))

# saveRDS(pcalist,"analysis/data/02-clustering/nnsvg_HVG_and_combos_PCA_NOcovariates.RDS")

#clean up
gc(full=T)
```

### UMAP
```{r}
umaplist <- bpmapply(X=hyp2.vgs.list,Y=as.list(names(hyp2.vgs.list)),Z=pcalist,BPPARAM = sbp,SIMPLIFY=FALSE,FUN = function(X,Y,Z){
    set.seed(42)
    # pass the corresp. pca in to speed alog UMAP generation
    hyp2.tmp <- copy(hyp2)
    reducedDims(hyp2.tmp)$PCA <- Z 
    # 3 reduced dim tables are already in the object, #4 is the appended PCA, so grab #5 (newly created)
    tmp <- scater::runUMAP(hyp2.tmp,name = paste0(Y,"_UMAP"),dimred="PCA",subset_row=X)
  return(reducedDims(tmp)[[5]])
})
# sanity check-UMAP coords came out in the same row order as the spe obj
stopifnot(sum(rownames(umaplist[[1]])==colnames(hyp2))==ncol(hyp2))
# change colnames to those expected by ggspavis for plotting later
umaplist <- lapply(umaplist,FUN=function(x){colnames(x) <- c("UMAP1","UMAP2")})

# saveRDS(umaplist,"analysis/data/02-clustering/nnsvg_HVG_and_combos_UMAPs_NOcovariates.RDS")

#clean up
gc(full=T)
```

# walktrap (using PCA space): spatially UNAWARE clustering for all 7 sets across 5 initial values of k
# in the walktrap algorithm, k does NOT specify a number of resultant or expected clusters; rather, k is just number of nearest neighbors utilized for calculating the relatedness SNNgraph in the first step (which is then used to generate any number of clusters via igraph::cluster_walktrap processing of the SNNgraph)
```{r}
ks <- c(10,15,20,25,30)
glists <- list()
for (i in c(1:5)){
    curk <- ks[i]
    ### cannot directly assign the bplapply output to a position in the parent list for whatever reason. but storing the list of bplapply results and then passing that to a slot in a parent list works just fine. 
    curlist <- bplapply(pcalist,BPPARAM = sbp,FUN=function(x){
        hyp2.tmp <- copy(hyp2)
        reducedDims(hyp2.tmp)$PCA <- x
        scran::buildSNNGraph(hyp2.tmp,k=curk,use.dimred="PCA")
    })
    glists[[i]] <- curlist
    rm(curlist)
    gc(full=T)
}

# make sure names are retained in result lists
names(glists) <- paste0("k",ks)

# not needed:
# lapply(glists,FUN=function(x){names(x)<-names(pcalist)})

# clean up
rm(curk,i)
gc(full=T)

```


### walktrap, second step--actual cluster assignments ###
### this is a RAM killer, so throttle down the parallelization to 4 cpus at a time (2 cycles per feature sets)

## coded this stupidly, only got through k=10,15,20 before hitting mem ceiling. rerun for glists 4+5 -> SPLIT RUN into K 10,15,20 and K 25,30 --> bound after the fact into the intended list obj
```{r}
sbp <- MulticoreParam(workers=4)
register(sbp)
walkclusts <- list()
walkclusts.out <- list()

i<-1
for (i in c(1:5)){ # really 1:3
    tmpwalks <- bplapply(glists[[i]],igraph::cluster_walktrap,BPPARAM = sbp)
    walkclusts[[i]] <- tmpwalks
    walkclusts.tmpout <- lapply(walkclusts[[i]],FUN=function(x){x$membership})
    walkclusts.out[[i]] <- walkclusts.tmpout
    names(walkclusts[[i]]) <- names(pcalist)
    names(walkclusts.out[[i]]) <- names(pcalist)
    rm(tmpwalks) # bunch of many GB crap hanging around til the next run
    gc(full=T)
}

### split run only
# names(walkclusts.out)[1:3] <- names(glists)[1:3]
# saveRDS(walkclusts.out,"analysis/code/02-clustering/7featureset_k10-30by5s_walktrap_clusters_NOcovars_k10-20only.RDS")
```

### jhpce portion -- get rid of glists used above to conserve memory, save k25 and k30 as temporary RDS object to pass to JHPCE
```{r}
glists2 <- list(glists[[4]],glists[[5]])
names(glists2) <- names(glists)[4:5]
names(glists2)
saveRDS(glists2,"~/Desktop/SNNgraphs_svgsAndHvgs_k25_k30_WITHOUTcovariates_052623.RDS")
rm(glists)
gc(full=T)
```


### k25 and k30 walktrap: Each feature set is trying to use about 20 GB for the k=25 (so ~140 GB total for k25), and probably 25% more for the k=30 (175 GB). glists2[[]] is just the k=25 and k=30 SNNgraphs so run the commented out block below accordingly.
#### see analysis/code/02-clustering/walktrap_k25_k30_nocov_batched.sh and .R
