---
title: "02-M0l0_kgeom6_leidenMarkers"
output: html_document
date: "2024-06-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(ggrastr)
library(gridExtra)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(spatialLIBD)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load the exact SFE that was used for banksy, just to be safe (it's just the original one with spatial coords rejiggered, but for max sanity)
```{r}
hypb <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

# read in the clustering results
bcl <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")


# map labels to sfe
bcl <- DataFrame(bcl,row.names = bcl$rn)
colLabels(hypb) <- bcl$M0lam0_leiden_multisamp_res2
```

## use registration wrapper to get markers
```{r}
hypb.mk <- registration_wrapper(hypb,var_registration = "label",var_sample_id = "sample_id",gene_ensembl = "ID",gene_name = "Symbol")

saveRDS(hypb.mk,"processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_spLIBD_markers.RDS")

hypb.mk2 <- hypb.mk[["enrichment"]]
hypb.mk2 <- as.data.table(hypb.mk2)

hypb.mk2[,ensembl:=NULL]
hypb.mk2 <- melt(hypb.mk2,id.vars="gene")

hypb.mk2[,clus:=gsub(variable,pattern="^.*_(X.*)$",replacement="\\1")]
hypb.mk2[,stat:=gsub(variable,pattern="^(.*)_X.*$",replacement="\\1")]

hypb.mk3 <- hypb.mk2[stat %in% c("fdr","logFC")]
hypb.mk3.1 <- hypb.mk3[stat=="fdr"&value<0.1]
setnames(hypb.mk3.1,"value","fdr")
hypb.mk3.1[,variable:=NULL]
hypb.mk3.1[,stat:=NULL]
hypb.mk3.2 <- hypb.mk3[stat=="logFC"&value>0]
setnames(hypb.mk3.2,"value","logFC")
hypb.mk3.2[,variable:=NULL]
hypb.mk3.2[,stat:=NULL]

hypbmk4 <- merge.data.table(hypb.mk3.1,hypb.mk3.2,by=c("gene","clus"))

fwrite(hypbmk4,"processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers.txt",sep='\t',quote=F)

rm(hypb.mk2,hypb.mk3,hypb.mk3.1,hypb.mk3.2,bcl)
gc(full=T)
```


## reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.37.1             
 [3] colorout_1.3-0.2               spatialLIBD_1.16.2            
 [5] SpatialFeatureExperiment_1.6.1 SpatialExperiment_1.14.0      
 [7] SingleCellExperiment_1.26.0    SummarizedExperiment_1.34.0   
 [9] Biobase_2.64.0                 GenomicRanges_1.56.1          
[11] MatrixGenerics_1.16.0          matrixStats_1.3.0             
[13] gridExtra_2.3                  ggrastr_1.0.2                 
[15] ggplot2_3.5.1                  Biostrings_2.72.1             
[17] GenomeInfoDb_1.40.1            XVector_0.44.0                
[19] IRanges_2.38.0                 S4Vectors_0.42.0              
[21] BiocGenerics_0.50.0            data.table_1.15.4             
[23] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0            
  [3] fields_15.2               bitops_1.0-7             
  [5] filelock_1.0.3            tibble_3.2.1             
  [7] R.oo_1.26.0               XML_3.99-0.16.1          
  [9] lifecycle_1.0.4           sf_1.0-16                
 [11] edgeR_4.2.0               doParallel_1.0.17        
 [13] rprojroot_2.0.4           lattice_0.22-6           
 [15] magrittr_2.0.3            sass_0.4.9               
 [17] limma_3.60.3              plotly_4.10.4            
 [19] rmarkdown_2.27            jquerylib_0.1.4          
 [21] yaml_2.3.8                httpuv_1.6.15            
 [23] spam_2.10-0               sessioninfo_1.2.2        
 [25] sp_2.1-4                  cowplot_1.1.3            
 [27] DBI_1.2.3                 RColorBrewer_1.1-3       
 [29] golem_0.4.1               maps_3.4.2               
 [31] abind_1.4-5               zlibbioc_1.50.0          
 [33] purrr_1.0.2               R.utils_2.12.3           
 [35] RCurl_1.98-1.14           rappdirs_0.3.3           
 [37] GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [39] irlba_2.3.5.1             terra_1.7-78             
 [41] units_0.8-5               dqrng_0.4.1              
 [43] DelayedMatrixStats_1.26.0 codetools_0.2-20         
 [45] DropletUtils_1.24.0       DelayedArray_0.30.1      
 [47] DT_0.33                   scuttle_1.14.0           
 [49] tidyselect_1.2.1          UCSC.utils_1.0.0         
 [51] viridis_0.6.5             ScaledMatrix_1.12.0      
 [53] shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [55] GenomicAlignments_1.40.0  jsonlite_1.8.8           
 [57] BiocNeighbors_1.22.0      e1071_1.7-14             
 [59] scater_1.32.0             iterators_1.0.14         
 [61] foreach_1.5.2             tools_4.4.1              
 [63] Rcpp_1.0.12               glue_1.7.0               
 [65] SparseArray_1.4.8         xfun_0.45                
 [67] here_1.0.1                EBImage_4.46.0           
 [69] dplyr_1.1.4               HDF5Array_1.32.0         
 [71] withr_3.0.0               BiocManager_1.30.23      
 [73] fastmap_1.2.0             boot_1.3-30              
 [75] rhdf5filters_1.16.0       fansi_1.0.6              
 [77] spData_2.3.1              rsvd_1.0.5               
 [79] digest_0.6.35             R6_2.5.1                 
 [81] mime_0.12                 colorspace_2.1-0         
 [83] wk_0.9.1                  jpeg_0.1-10              
 [85] RSQLite_2.3.7             R.methodsS3_1.8.2        
 [87] config_0.3.2              utf8_1.2.4               
 [89] tidyr_1.3.1               generics_0.1.3           
 [91] rtracklayer_1.64.0        class_7.3-22             
 [93] httr_1.4.7                htmlwidgets_1.6.4        
 [95] S4Arrays_1.4.1            spdep_1.3-5              
 [97] pkgconfig_2.0.3           gtable_0.3.5             
 [99] blob_1.2.4                htmltools_0.5.8.1        
[101] dotCall64_1.1-1           fftwtools_0.9-11         
[103] scales_1.3.0              attempt_0.3.1            
[105] png_0.1-8                 knitr_1.47               
[107] rstudioapi_0.16.0         rjson_0.2.21             
[109] curl_5.2.1                proxy_0.4-27             
[111] cachem_1.1.0              rhdf5_2.48.0             
[113] BiocVersion_3.19.1        KernSmooth_2.23-24       
[115] parallel_4.4.1            vipor_0.4.7              
[117] AnnotationDbi_1.66.0      restfulr_0.0.15          
[119] s2_1.1.6                  pillar_1.9.0             
[121] grid_4.4.1                vctrs_0.6.5              
[123] promises_1.3.0            BiocSingular_1.20.0      
[125] dbplyr_2.5.0              beachmat_2.20.0          
[127] xtable_1.8-4              sfheaders_0.4.4          
[129] beeswarm_0.4.0            paletteer_1.6.0          
[131] evaluate_0.24.0           zeallot_0.1.0            
[133] magick_2.8.3              Rsamtools_2.20.0         
[135] cli_3.6.2                 locfit_1.5-9.9           
[137] compiler_4.4.1            crayon_1.5.3             
[139] classInt_0.4-10           rematch2_2.1.2           
[141] ggbeeswarm_0.7.2          viridisLite_0.4.2        
[143] deldir_2.0-4              munsell_0.5.1            
[145] lazyeval_0.2.2            tiff_0.1-12              
[147] Matrix_1.7-0              ExperimentHub_2.12.0     
[149] benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[151] bit64_4.0.5               Rhdf5lib_1.26.0          
[153] KEGGREST_1.44.1           statmod_1.5.0            
[155] shiny_1.8.1.1             AnnotationHub_3.12.0     
[157] memoise_2.0.1             bslib_0.7.0              
[159] benchmarkmeData_1.0.4     bit_4.0.5                
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version     date (UTC) lib source
   abind                      1.4-5       2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0      2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1       2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0       2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8       2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4       2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache              2.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23     2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular               1.20.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1      2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings               * 2.72.1      2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5       2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5       2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7       2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4       2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30      2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0       2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0       2024-05-16 [1] CRAN (R 4.4.0)
   class                      7.3-22      2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10      2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.2       2023-12-11 [2] CRAN (R 4.4.0)
   codetools                  0.2-20      2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2     2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0       2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2       2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3       2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3       2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1       2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4      2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3       2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0       2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1      2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4       2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.35      2024-03-11 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17      2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64                  1.1-1       2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4       2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1       2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33        2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14      2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0       2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0      2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6       2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0       2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11      2021-03-01 [1] CRAN (R 4.4.0)
   fields                     15.2        2023-08-17 [1] CRAN (R 4.4.0)
   filelock                   1.0.3       2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2       2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3       2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1      2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12      2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1      2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2       2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1       2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2       2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5       2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0       2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1       2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3         2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5       2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1       2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1     2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4       2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15      2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7       2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                      2.3.5.1     2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14      2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10      2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4       2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8       2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1      2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24     2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47        2024-05-29 [1] CRAN (R 4.4.0)
   later                      1.3.2       2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6      2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2       2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4       2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3      2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.9     2024-03-01 [1] CRAN (R 4.4.0)
   magick                     2.8.3       2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3       2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2       2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0       2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0       2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1       2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12        2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1       2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0       2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1      2024-02-29 [1] CRAN (R 4.4.0)
   pillar                     1.9.0       2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3       2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4      2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8       2022-11-29 [1] CRAN (R 4.4.0)
   promises                   1.3.0       2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27      2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2       2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2       2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0      2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3      2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1       2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3       2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3       2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12      2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14   2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2       2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15      2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21      2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4       2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27        2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4       2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7       2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0      2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5       2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6       2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1       2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sass                       0.4.9       2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0       2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                    1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2       2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16      2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4       2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1     2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6       2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4       2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0      2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8       2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1       2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2      2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1       2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5       2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0       2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78      2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1       2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1       2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1       2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12      2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0       2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5       2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4       2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5       2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7       2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5       2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2       2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0       2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.1       2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45        2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.16.1 2024-01-22 [1] CRAN (R 4.4.0)
   xtable                     1.8-4       2019-04-21 [1] CRAN (R 4.4.0)
   XVector                  * 0.44.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8       2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0       2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
