---
title: "Disgenet queries for gene resource"
author: "Bernard Mulvey"
date: "2023-09-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(disgenet2r)
require(colorout)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

#### Diseases to collect ######
ASD
ADHD
AUD
Cannabis use
Alzheimer's
MDD (disease ontology #s 12294 (atypical depression, which is included in virtually any study) and depressive disorders parent, 1596)
PTSD
Anorexia
BIP
SCZ
BMI
Lefthanded
Opioid use
Panic
OCD
Tourettes
#####

### MDD ### 
We'll do this one first, because there's genes I know that disgene lists as "curated" for MDD despite a publication in 2019 (Border, Johnson '19 Am J Psy) demonstrating beyond a doubt these early candidates from genetic studies of tiny handfuls of cases and controls are garbage. So we will be filtering out the following bullshit genes:
1. SLC6A4
2. BDNF
3. COMT
4. HTR2A
5. TPH1, TPH2
6. DRD2, DRD3, DRD4
7. MAOA
8. MTHFR
9. APOE
10. SLC6A3
11. CLOCK
12. DTNBP1
13. ACE
14. DBH
15. ANCB1

SLC6A4 BDNF COMT HTR2A TPH1 TPH2 DRD2 DRD3 DRD4 MAOA MTHFR APOE SLC6A3 CLOCK DTNBP1 ACE DBH ANCB1

```{r}
Sys.setenv(DISGENET_API_KEY="67559ece77977094389864b58d27185e8ac48ad3")

mdd <- disgenet2r::disease2gene(database = "CURATED",vocabulary = "DO",disease = c("12294", "1596"))

mdd2 <- as.data.table(mdd@qresult)
bullshit <- c("SLC6A4", "BDNF", "COMT", "HTR2A", "TPH1", "TPH2", "DRD2", "DRD3", "DRD4", "MAOA", "MTHFR", "APOE", "SLC6A3", "CLOCK", "DTNBP1", "ACE", "DBH", "ANCB1")

mdd2 <- mdd2[!(gene_symbol %in% bullshit)]

mdd3 <- mdd2[,.(geneid,gene_symbol,uniprotid)]
mdd3[,dx:="MDD"]
mdd3[,query:="DisOntol_12294_1596"]
restab <- copy(mdd3)
rm(mdd,mdd2,mdd3,bullshit)
```

#### ASD ####
Parent-most DO term 0060041
```{r}
asd <- disgenet2r::disease2gene(database = "CURATED",vocabulary = "DO",disease = "0060041")

asd <- as.data.table(asd@qresult)

asd <- asd[,.(geneid,gene_symbol,uniprotid)]
asd[,dx:="ASD"]
asd[,query:="DisOntol_0060041"]
restab <- rbind(restab,asd)
```


## ok this is silly nonsense. queue the loop.
### NO CYCLOTHYMIA for bipolar query
### NDD query = intellectual disability parent term 1059 (includes autosomal, non-syndromic, syndromic forms; EXcludes ADHD, specific learning disabilities (e.g. reading/writing/dyscalculia, ADHD, communication-domain-specific disorders))
# Cannabis use disorder 9505
# Alzheimer's 10652
# PTSD 2055
# Anorexia 8689
# BIP 14042 0060166 (BIP 1 and 2 only)
# SCZ 5419
# BMI (obesity) 9970
# Opioid use disorder 11206
# Panic 594
# OCD 10933
# Tourettes 11119
# GAD 14320
# NDD 1059
```{r}
# disq <- list(ADHD="1094",AUD="1574",Cannabis="9505",AD="10652",PTSD="2055",Anorexia="8689",BIP=c("14042","0060166"),SCZ="5419",BMI="9970",OUD="11206",Panic="594",OCD="10933",TS="11119",GAD="14320",NDD="1059")

## test thru loop to find any that won't work
# i<-8
# for (i in c(8:length(disq))){
#   x <- disq[[i]]
#   res <- disease2gene(disease = x,vocabulary = "DO",database = "CURATED")
#   res <- as.data.table(res@qresult)
#   res <- res[,.(geneid,gene_symbol,uniprotid)]
#   res[,query:=paste0("DisOntol_",x)]
# }

# errors on these, which we'll have to fetch a different way
# PTSD disq[[5]]
# BIP disq[[7]] 
### disgenet web only shows UMLS IDs for BIP: C0853193 (BIP I) and C0236788 (II)
### disgenet web has DO2055 for PTSD, but doesn't list it if you try and search it? UMLS works though: C0038436

disq <- list(ADHD="1094",AUD="1574",Cannabis="9505",AD="10652",Anorexia="8689",SCZ="5419",BMI="9970",OUD="11206",Panic="594",OCD="10933",TS="11119",GAD="14320",NDD="1059")

reslist <- lapply(disq,FUN=function(x){
  res <- disease2gene(disease = x,vocabulary = "DO",database = "CURATED")
  res <- as.data.table(res@qresult)
  res <- res[,.(geneid,gene_symbol,uniprotid)]
  res[,query:=paste0("DisOntol_",x)]
  return(res)
})

i <- 1
for (i in c(1:length(reslist))){
  tmp <- reslist[[i]]
  tmp[,dx:=names(reslist)[i]]
  if(i==1){restab2 <- copy(tmp)}
  else{restab2 <- rbind(restab2,tmp)}
  rm(tmp)
}

restab2 <- rbind(restab,restab2)
rm(restab)

```

### fetch the two goofy ones
```{r}
disq <- list(BIP=c("C0853193","C0236788"),PTSD="C0038436")
restab3 <- lapply(disq,FUN=function(x){
  res <- disease2gene(disease = x,vocabulary = "UMLS",database = "CURATED")
  res <- as.data.table(res@qresult)
  res <- res[,.(geneid,gene_symbol,uniprotid)]
  return(res)
})

### WTF?
restab3 <- disease2gene(disease = "C0853193",vocabulary = "UMLS",database = "CURATED")
restab3 <- as.data.table(restab3@qresult)
restab4 <- disease2gene(disease="C0236788",vocabulary="UMLS",database="CURATED")
restab4 <- as.data.table(restab4@qresult)
restab3 <- rbind(restab3,restab4)
restab3 <- restab3[,.(geneid,gene_symbol,uniprotid)]

restab3[,dx:="BIP"]
restab3[,query:="UMLS_C0236788_UMLS_C0853193"]
restab2 <- rbind(restab2,restab3)
rm(restab3,restab4)

# ptsd = 
ptsd <- disease2gene(disease = "C0038436",vocabulary = "UMLS",database = "CURATED")
 

i <- 1
for (i in c(1:length(restab3))){
  tmp <- reslist[[i]]
  tmp[,dx:=names(reslist)[i]]
  if(i==1){restab2 <- copy(tmp)}
  else{restab2 <- rbind(restab2,tmp)}
  rm(tmp)
}


```