---
title: "01-Prep Input Lists for GSEA of VMH, ARC sex DEGs"
author: "Bernard Mulvey"
date: "2023-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
require(colorout)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

```{r}
arcde <- readRDS("analysis/data/spe_053123/H05-bspace15-nnsvg10_ARC_clusters/ARC_pseudobulk_sexDE_noSpRep-adjs_V12D07075B1-cropped.RDS")

arcde <- arcde[["cl_both_ARC"]]

vmhde <- fread("analysis/data/spe_053123/H04-2cluster-consensus-VMH/pseudobulk-SEX-DE_Bspace-VMH-nonvmh-clusters_postcluster-crop_V12D07075B1.txt")
vmhde <- vmhde[group_name=="VMH"]

write.table(vmhde[,.(gene_name,logFC)],"analysis/data/spe_053123/H08-GSEA, GO etc/01-VMH2clust-sexDE_prerank.rnk",sep='\t',quote=F,row.names=F,col.names=F)

write.table(arcde[,.(gene_name,logFC)],"analysis/data/spe_053123/H08-GSEA, GO etc/01-ARCboth-sexDE_prerank.rnk",sep='\t',quote=F,row.names=F,col.names=F)

```

GSEA preranked (i.e. by LFC) run notes (same settings for VMH and ARC), 8/20/23:
1. mac app 4.3.2
2. Tested against msigdb sets c1-c8: ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c1.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c2.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c3.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c4.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c5.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c6.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c7.all.v2023.1.Hs.symbols.gmt,ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c8.all.v2023.1.Hs.symbols.gmt
3. Collapse or remap? Collapse
4. Symbol mapping "chip" ftp.broadinstitute.org://pub/gsea/msigdb/human/annotations/Human_Gene_Symbol_with_Remapping_MSigDB.v2023.1.Hs.chip
5. Permutations: 1000
6. positive logFC = male > female


## R session info output:
R version 4.3.0 (2023-04-21)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] colorout_1.2-2              spatialLIBD_1.12.0         
 [3] SpatialExperiment_1.10.0    SingleCellExperiment_1.22.0
 [5] SummarizedExperiment_1.30.2 Biobase_2.60.0             
 [7] GenomicRanges_1.52.0        MatrixGenerics_1.12.3      
 [9] matrixStats_1.0.0           gridExtra_2.3              
[11] ggplot2_3.4.3               Biostrings_2.68.1          
[13] GenomeInfoDb_1.36.1         XVector_0.40.0             
[15] IRanges_2.34.1              S4Vectors_0.38.1           
[17] BiocGenerics_0.46.0         data.table_1.14.8          
[19] rlang_1.1.1                

loaded via a namespace (and not attached):
  [1] later_1.3.1                   BiocIO_1.10.0                
  [3] bitops_1.0-7                  filelock_1.0.2               
  [5] fields_14.1                   tibble_3.2.1                 
  [7] R.oo_1.25.0                   XML_3.99-0.14                
  [9] lifecycle_1.0.3               edgeR_3.42.4                 
 [11] doParallel_1.0.17             rprojroot_2.0.3              
 [13] lattice_0.21-8                magrittr_2.0.3               
 [15] sass_0.4.7                    limma_3.56.2                 
 [17] plotly_4.10.2                 rmarkdown_2.24               
 [19] jquerylib_0.1.4               yaml_2.3.7                   
 [21] httpuv_1.6.11                 spam_2.9-1                   
 [23] sessioninfo_1.2.2             cowplot_1.1.1                
 [25] DBI_1.1.3                     RColorBrewer_1.1-3           
 [27] golem_0.4.1                   maps_3.4.1                   
 [29] abind_1.4-5                   zlibbioc_1.46.0              
 [31] purrr_1.0.2                   R.utils_2.12.2               
 [33] RCurl_1.98-1.12               rappdirs_0.3.3               
 [35] GenomeInfoDbData_1.2.10       ggrepel_0.9.3                
 [37] irlba_2.3.5.1                 dqrng_0.3.0                  
 [39] DelayedMatrixStats_1.22.5     codetools_0.2-19             
 [41] DropletUtils_1.20.0           DelayedArray_0.26.7          
 [43] DT_0.28                       scuttle_1.9.4                
 [45] tidyselect_1.2.0              ScaledMatrix_1.8.1           
 [47] viridis_0.6.4                 shinyWidgets_0.7.6           
 [49] BiocFileCache_2.8.0           GenomicAlignments_1.36.0     
 [51] jsonlite_1.8.7                BiocNeighbors_1.18.0         
 [53] ellipsis_0.3.2                scater_1.28.0                
 [55] iterators_1.0.14              foreach_1.5.2                
 [57] tools_4.3.0                   Rcpp_1.0.11                  
 [59] glue_1.6.2                    xfun_0.40                    
 [61] here_1.0.1                    dplyr_1.1.2                  
 [63] HDF5Array_1.28.1              withr_2.5.0                  
 [65] BiocManager_1.30.22           fastmap_1.1.1                
 [67] rhdf5filters_1.12.1           fansi_1.0.4                  
 [69] digest_0.6.33                 rsvd_1.0.5                   
 [71] R6_2.5.1                      mime_0.12                    
 [73] colorspace_2.1-0              RSQLite_2.3.1                
 [75] R.methodsS3_1.8.2             config_0.3.1                 
 [77] utf8_1.2.3                    tidyr_1.3.0                  
 [79] generics_0.1.3                rtracklayer_1.60.0           
 [81] httr_1.4.7                    htmlwidgets_1.6.2            
 [83] S4Arrays_1.0.5                pkgconfig_2.0.3              
 [85] gtable_0.3.3                  blob_1.2.4                   
 [87] htmltools_0.5.6               dotCall64_1.0-2              
 [89] scales_1.2.1                  png_0.1-8                    
 [91] attempt_0.3.1                 knitr_1.43                   
 [93] rstudioapi_0.15.0             rjson_0.2.21                 
 [95] curl_5.0.2                    cachem_1.0.8                 
 [97] rhdf5_2.44.0                  BiocVersion_3.17.1           
 [99] vipor_0.4.5                   parallel_4.3.0               
[101] AnnotationDbi_1.62.2          restfulr_0.0.15              
[103] pillar_1.9.0                  grid_4.3.0                   
[105] vctrs_0.6.3                   promises_1.2.1               
[107] BiocSingular_1.16.0           dbplyr_2.3.3                 
[109] beachmat_2.16.0               xtable_1.8-4                 
[111] beeswarm_0.4.0                paletteer_1.5.0              
[113] evaluate_0.21                 magick_2.7.5                 
[115] cli_3.6.1                     locfit_1.5-9.8               
[117] compiler_4.3.0                Rsamtools_2.16.0             
[119] crayon_1.5.2                  rematch2_2.1.2               
[121] ggbeeswarm_0.7.2              viridisLite_0.4.2            
[123] BiocParallel_1.34.2           munsell_0.5.0                
[125] lazyeval_0.2.2                Matrix_1.6-1                 
[127] ExperimentHub_2.8.1           benchmarkme_1.0.8            
[129] sparseMatrixStats_1.12.2      bit64_4.0.5                  
[131] Rhdf5lib_1.22.0               statmod_1.5.0                
[133] KEGGREST_1.40.0               shiny_1.7.5                  
[135] interactiveDisplayBase_1.38.0 AnnotationHub_3.8.0          
[137] memoise_2.0.1                 bslib_0.5.1                  
[139] benchmarkmeData_1.0.4         bit_4.0.5      

