---
title: "Plot_mouseVMH-APgenes_on_hypomap_spatial"
author: "Bernie Mulvey"
date: "2025-03-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(ggtext)
library(gridExtra)
library(SpatialExperiment)
library(Seurat)
library(escheR)
library(ggrastr)
# library(SpatialFeatureExperiment)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9), axis.title.x = element_text(size = 8), axis.text.y = element_text(size = 8), axis.title.y = element_text(size =9), plot.title = element_markdown(size = 11,hjust=0.5), strip.text = element_text(size=11), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),plot.title.position="plot"))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

obtained file from https://www.repository.cam.ac.uk/bitstreams/a47399a8-2735-4b65-b6cb-3cd1a48b08d7/download
```{r}
hmap <- readRDS("~/Desktop/humanHYPOMAP_spatial.rds")

## convert to SPE using code from https://github.com/drighelli/SpatialExperiment/issues/115 
library(dplyr)
seurat_to_spe <- function(seu, sample_id, img_id) {
    
    ## Convert to SCE
    sce <- Seurat::as.SingleCellExperiment(seu)
    
    ## Extract spatial coordinates
    spatialCoords <- as.matrix(
        seu@images[[img_id]]@coordinates[, c("imagecol", "imagerow")])
    
    ## Extract and process image data
    img <- SpatialExperiment::SpatialImage(
        x = as.raster(seu@images[[img_id]]@image))
    
    imgData <- DataFrame(
        sample_id = sample_id,
        image_id = img_id,
        data = I(list(img)),
        scaleFactor = seu@images[[img_id]]@scale.factors$lowres)
    
    # Convert to SpatialExperiment
    spe <- SpatialExperiment(
        assays = assays(sce),
        rowData = rowData(sce),
        colData = colData(sce),
        metadata = metadata(sce),
        reducedDims = reducedDims(sce),
        altExps = altExps(sce),
        sample_id = sample_id,
        spatialCoords = spatialCoords,
        imgData = imgData
    )
    # indicate all spots are on the tissue
    spe$in_tissue <- 1
    spe$sample_id <- sample_id
    # Return Spatial Experiment object
    spe
}


## sample_ids and image_ids are permutations of the following:
## sampleid = hmap$captureArea (b2,a3,...)
## image id is a permutation of this (slice2B, slice3A..)


spat_metas <- list(c("b2","slice2B"),
                   c("a3","slice3A"),
                   c("b4","slice4B"),
                   c("a5","slice5A"),
                   c("b6","slice6B"),
                   c("a7","slice7A"),
                   c("b8","slice8B"),
                   c("b1_b","sliceB1B"),
                   c("c1_a","sliceC1A"))
names(spat_metas) <- unlist(lapply(spat_metas,function(x){x[2]}))

spe_ls <- lapply(spat_metas, function(m) {
    seurat_to_spe(seu = hmap[,hmap$captureArea==m[1]], sample_id = m[1], img_id = m[2])
})

spe <- Reduce(cbind, spe_ls)
rm(hmap,spat_metas,spe_ls)
gc(full=T)
```

```{r}
## load the mouse VMH AP genes
msap <- fread("spatial_HYP/processed-data/12-Mouse comparisons/06a-Top5mks_per_APmapped_mspop_with1to1_hgorths.txt")
msap[,gene:=toupper(gene)]
msap <- msap[gene %in% rownames(spe)]

spe <- spe[msap$gene,]
nrow(spe) # 23 / 25 genes are in this dataset
gc(full=T)
```


```{r}
samps <- lapply(unique(spe$sample_id),FUN=function(x){spe[,spe$sample_id==x]})
rm(spe)

keepsamps <- lapply(samps,FUN=function(x){sum(x$regional_clusters_named=="VMH")>0})
samps <- samps[unlist(keepsamps)]

samps <- lapply(samps,FUN=function(s){
    s$isvmh <- s$regional_clusters_named
    s$isvmh[s$isvmh!="VMH"] <- "other"
    return(s)
})

vpal <- c("purple",NA)
names(vpal) <- c("VMH","other")

lapply(c(1:nrow(msap)),function(x){
    plts <- lapply(samps,FUN=function(s){
        # s$isvmh <- ifelse(s$regional_clusters_named=="VMH","VMH","other")
        s$logct <- as.numeric(logcounts(s[msap[x,gene],]))
        p <- make_escheR(s)
        p <- p |> add_ground("isvmh",point_size = 1.5,stroke = 0.2)
        p <- p |> add_fill("logct",point_size=1.5)
        
        p <- p+ 
            scale_fill_continuous("log\ncounts",low= "#FFFFFF",high="#000000")+
            scale_color_manual(values=vpal,na.value = NA)
        
        p <- rasterize(p,layer="Points",dpi=450,dev="cairo_png")
        return(p)
    })
    pdf(paste0("manuscript_plots/Supplemental/hmAPcheck/",msap[x,variable],"_",msap[x,gene],".pdf"),height=20,width=20)
    do.call("grid.arrange",plts)
    dev.off()
})
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.3 (2025-02-28)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.7.4

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] ggrastr_1.0.2               dplyr_1.1.4                 BiocParallel_1.40.0        
 [4] parallelly_1.43.0           colorout_1.3-0.2            escheR_1.6.0               
 [7] Seurat_5.2.1                SeuratObject_5.0.2          sp_2.2-0                   
[10] SpatialExperiment_1.16.0    SingleCellExperiment_1.28.1 SummarizedExperiment_1.36.0
[13] Biobase_2.66.0              GenomicRanges_1.58.0        GenomeInfoDb_1.42.3        
[16] IRanges_2.40.1              S4Vectors_0.44.0            BiocGenerics_0.52.0        
[19] MatrixGenerics_1.18.1       matrixStats_1.5.0           gridExtra_2.3              
[22] ggtext_0.1.2                ggplot2_3.5.1               data.table_1.17.0          
[25] rlang_1.1.5                

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3      rstudioapi_0.17.1       jsonlite_1.9.1         
  [4] magrittr_2.0.3          ggbeeswarm_0.7.2        spatstat.utils_3.1-3   
  [7] magick_2.8.6            farver_2.1.2            rmarkdown_2.29         
 [10] zlibbioc_1.52.0         vctrs_0.6.5             ROCR_1.0-11            
 [13] spatstat.explore_3.4-2  htmltools_0.5.8.1       S4Arrays_1.6.0         
 [16] SparseArray_1.6.2       sctransform_0.4.1       KernSmooth_2.23-26     
 [19] htmlwidgets_1.6.4       ica_1.0-3               plyr_1.8.9             
 [22] plotly_4.10.4           zoo_1.8-13              igraph_2.1.4           
 [25] mime_0.13               lifecycle_1.0.4         pkgconfig_2.0.3        
 [28] Matrix_1.7-3            R6_2.6.1                fastmap_1.2.0          
 [31] GenomeInfoDbData_1.2.13 fitdistrplus_1.2-2      future_1.34.0          
 [34] shiny_1.10.0            digest_0.6.37           colorspace_2.1-1       
 [37] patchwork_1.3.0         tensor_1.5              rprojroot_2.0.4        
 [40] RSpectra_0.16-2         irlba_2.3.5.1           labeling_0.4.3         
 [43] progressr_0.15.1        spatstat.sparse_3.1-0   polyclip_1.10-7        
 [46] httr_1.4.7              abind_1.4-8             compiler_4.4.3         
 [49] here_1.0.1              withr_3.0.2             fastDummies_1.7.5      
 [52] MASS_7.3-65             DelayedArray_0.32.0     rjson_0.2.23           
 [55] tools_4.4.3             vipor_0.4.7             lmtest_0.9-40          
 [58] beeswarm_0.4.0          httpuv_1.6.15           future.apply_1.11.3    
 [61] goftest_1.2-3           glue_1.8.0              nlme_3.1-167           
 [64] promises_1.3.2          gridtext_0.1.5          grid_4.4.3             
 [67] Rtsne_0.17              cluster_2.1.8.1         reshape2_1.4.4         
 [70] generics_0.1.3          gtable_0.3.6            spatstat.data_3.1-6    
 [73] tidyr_1.3.1             xml2_1.3.8              XVector_0.46.0         
 [76] spatstat.geom_3.3-6     RcppAnnoy_0.0.22        ggrepel_0.9.6          
 [79] RANN_2.6.2              pillar_1.10.1           stringr_1.5.1          
 [82] spam_2.11-1             RcppHNSW_0.6.0          later_1.4.1            
 [85] splines_4.4.3           lattice_0.22-6          deldir_2.0-4           
 [88] survival_3.8-3          tidyselect_1.2.1        miniUI_0.1.1.1         
 [91] pbapply_1.7-2           knitr_1.50              scattermore_1.2        
 [94] xfun_0.51               stringi_1.8.4           UCSC.utils_1.2.0       
 [97] lazyeval_0.2.2          yaml_2.3.10             evaluate_1.0.3         
[100] codetools_0.2-20        tibble_3.2.1            cli_3.6.4              
[103] uwot_0.2.3              xtable_1.8-4            reticulate_1.41.0.1    
[106] munsell_0.5.1           Rcpp_1.0.14             spatstat.random_3.3-3  
[109] globals_0.16.3          png_0.1-8               spatstat.univar_3.1-2  
[112] parallel_4.4.3          dotCall64_1.2           listenv_0.9.1          
[115] viridisLite_0.4.2       scales_1.3.0            ggridges_0.5.6         
[118] purrr_1.0.4             crayon_1.5.3            cowplot_1.1.3          
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.3 (2025-02-28)
 os       macOS Sonoma 14.7.4
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2025-03-31
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)
 quarto   1.4.555 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/quarto

─ Packages ───────────────────────────────────────────────────────────────────────────
 ! package              * version  date (UTC) lib source
   abind                  1.4-8    2024-09-12 [1] CRAN (R 4.4.1)
   beeswarm               0.4.0    2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.66.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocGenerics         * 0.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   BiocParallel         * 1.40.0   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
 P cli                    3.6.4    2025-02-13 [2] CRAN (R 4.4.1)
   cluster                2.1.8.1  2025-03-12 [1] CRAN (R 4.4.1)
   codetools              0.2-20   2024-03-31 [1] CRAN (R 4.4.3)
   colorout             * 1.3-0.2  2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-1    2024-07-26 [1] CRAN (R 4.4.0)
   cowplot                1.1.3    2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3    2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.17.0   2025-02-22 [1] CRAN (R 4.4.1)
   DelayedArray           0.32.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   deldir                 2.0-4    2024-02-28 [1] CRAN (R 4.4.0)
   digest                 0.6.37   2024-08-19 [1] CRAN (R 4.4.1)
   dotCall64              1.2      2024-10-04 [1] CRAN (R 4.4.1)
   dplyr                * 1.1.4    2023-11-17 [1] CRAN (R 4.4.0)
   escheR               * 1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   evaluate               1.0.3    2025-01-10 [1] CRAN (R 4.4.1)
   farver                 2.1.2    2024-05-13 [1] CRAN (R 4.4.0)
   fastDummies            1.7.5    2025-01-20 [1] CRAN (R 4.4.1)
   fastmap                1.2.0    2024-05-15 [1] CRAN (R 4.4.0)
   fitdistrplus           1.2-2    2025-01-07 [1] CRAN (R 4.4.1)
   future                 1.34.0   2024-07-29 [1] CRAN (R 4.4.0)
   future.apply           1.11.3   2024-10-27 [1] CRAN (R 4.4.1)
   generics               0.1.3    2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.42.3   2025-01-27 [1] Bioconductor 3.20 (R 4.4.2)
   GenomeInfoDbData       1.2.13   2024-12-12 [1] Bioconductor
   GenomicRanges        * 1.58.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   ggbeeswarm             0.7.2    2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1    2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr              * 1.0.2    2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                0.9.6    2024-09-07 [1] CRAN (R 4.4.1)
   ggridges               0.5.6    2024-01-23 [1] CRAN (R 4.4.0)
   ggtext               * 0.1.2    2022-09-16 [1] CRAN (R 4.4.0)
   globals                0.16.3   2024-03-08 [1] CRAN (R 4.4.0)
   glue                   1.8.0    2024-09-30 [1] CRAN (R 4.4.1)
   goftest                1.2-3    2021-10-07 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3      2017-09-09 [1] CRAN (R 4.4.0)
   gridtext               0.1.5    2022-09-16 [1] CRAN (R 4.4.0)
   gtable                 0.3.6    2024-10-25 [1] CRAN (R 4.4.1)
   here                   1.0.1    2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1  2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4    2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15   2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7    2023-08-15 [1] CRAN (R 4.4.0)
   ica                    1.0-3    2022-07-08 [1] CRAN (R 4.4.0)
   igraph                 2.1.4    2025-01-23 [1] CRAN (R 4.4.1)
   IRanges              * 2.40.1   2024-12-05 [1] Bioconductor 3.20 (R 4.4.2)
   irlba                  2.3.5.1  2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.9.1    2025-03-03 [1] CRAN (R 4.4.1)
   KernSmooth             2.23-26  2025-01-01 [1] CRAN (R 4.4.3)
   knitr                  1.50     2025-03-16 [1] CRAN (R 4.4.2)
   labeling               0.4.3    2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.4.1    2024-11-27 [1] CRAN (R 4.4.1)
   lattice                0.22-6   2024-03-20 [1] CRAN (R 4.4.3)
   lazyeval               0.2.2    2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4    2023-11-07 [1] CRAN (R 4.4.0)
   listenv                0.9.1    2024-01-29 [1] CRAN (R 4.4.0)
   lmtest                 0.9-40   2022-03-21 [1] CRAN (R 4.4.0)
   magick                 2.8.6    2025-03-23 [1] CRAN (R 4.4.1)
   magrittr               2.0.3    2022-03-30 [1] CRAN (R 4.4.0)
   MASS                   7.3-65   2025-02-28 [1] CRAN (R 4.4.1)
   Matrix                 1.7-3    2025-03-11 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.18.1   2025-01-09 [1] Bioconductor 3.20 (R 4.4.2)
   matrixStats          * 1.5.0    2025-01-07 [1] CRAN (R 4.4.1)
   mime                   0.13     2025-03-17 [1] CRAN (R 4.4.1)
   miniUI                 0.1.1.1  2018-05-18 [1] CRAN (R 4.4.0)
   munsell                0.5.1    2024-04-01 [1] CRAN (R 4.4.0)
   nlme                   3.1-167  2025-01-27 [1] CRAN (R 4.4.3)
   parallelly           * 1.43.0   2025-03-24 [1] CRAN (R 4.4.1)
   patchwork              1.3.0    2024-09-16 [1] CRAN (R 4.4.1)
   pbapply                1.7-2    2023-06-27 [1] CRAN (R 4.4.0)
   pillar                 1.10.1   2025-01-07 [1] CRAN (R 4.4.1)
   pkgconfig              2.0.3    2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4   2024-01-13 [1] CRAN (R 4.4.0)
   plyr                   1.8.9    2023-10-02 [1] CRAN (R 4.4.0)
   png                    0.1-8    2022-11-29 [1] CRAN (R 4.4.0)
   polyclip               1.10-7   2024-07-23 [1] CRAN (R 4.4.0)
   progressr              0.15.1   2024-11-22 [1] CRAN (R 4.4.1)
   promises               1.3.2    2024-11-28 [1] CRAN (R 4.4.1)
   purrr                  1.0.4    2025-02-05 [1] CRAN (R 4.4.1)
   R6                     2.6.1    2025-02-15 [1] CRAN (R 4.4.1)
   RANN                   2.6.2    2024-08-25 [1] CRAN (R 4.4.1)
   RColorBrewer           1.1-3    2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.14   2025-01-12 [1] CRAN (R 4.4.1)
   RcppAnnoy              0.0.22   2024-01-23 [1] CRAN (R 4.4.0)
   RcppHNSW               0.6.0    2024-02-04 [1] CRAN (R 4.4.0)
   reshape2               1.4.4    2020-04-09 [1] CRAN (R 4.4.0)
   reticulate             1.41.0.1 2025-03-09 [1] CRAN (R 4.4.1)
   rjson                  0.2.23   2024-09-16 [1] CRAN (R 4.4.1)
 P rlang                * 1.1.5    2025-01-17 [2] CRAN (R 4.4.1)
   rmarkdown              2.29     2024-11-04 [1] CRAN (R 4.4.1)
   ROCR                   1.0-11   2020-05-02 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4    2023-11-05 [1] CRAN (R 4.4.0)
   RSpectra               0.16-2   2024-07-18 [1] CRAN (R 4.4.0)
   rstudioapi             0.17.1   2024-10-22 [1] CRAN (R 4.4.1)
   Rtsne                  0.17     2023-12-07 [1] CRAN (R 4.4.0)
   S4Arrays               1.6.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   S4Vectors            * 0.44.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   scales                 1.3.0    2023-11-28 [1] CRAN (R 4.4.0)
   scattermore            1.2      2023-06-12 [1] CRAN (R 4.4.0)
   sctransform            0.4.1    2023-10-19 [1] CRAN (R 4.4.0)
   sessioninfo            1.2.3    2025-02-05 [1] CRAN (R 4.4.1)
   Seurat               * 5.2.1    2025-01-24 [1] CRAN (R 4.4.1)
   SeuratObject         * 5.0.2    2024-05-08 [1] CRAN (R 4.4.0)
   shiny                  1.10.0   2024-12-14 [1] CRAN (R 4.4.1)
   SingleCellExperiment * 1.28.1   2024-11-23 [1] Bioconductor 3.20 (R 4.4.2)
   sp                   * 2.2-0    2025-02-01 [1] CRAN (R 4.4.1)
   spam                   2.11-1   2025-01-20 [1] CRAN (R 4.4.1)
   SparseArray            1.6.2    2025-02-20 [1] Bioconductor 3.20 (R 4.4.2)
   SpatialExperiment    * 1.16.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   spatstat.data          3.1-6    2025-03-17 [1] CRAN (R 4.4.2)
   spatstat.explore       3.4-2    2025-03-21 [1] CRAN (R 4.4.1)
   spatstat.geom          3.3-6    2025-03-18 [1] CRAN (R 4.4.1)
   spatstat.random        3.3-3    2025-03-19 [1] CRAN (R 4.4.1)
   spatstat.sparse        3.1-0    2024-06-21 [1] CRAN (R 4.4.0)
   spatstat.univar        3.1-2    2025-03-05 [1] CRAN (R 4.4.1)
   spatstat.utils         3.1-3    2025-03-15 [1] CRAN (R 4.4.1)
   stringi                1.8.4    2024-05-06 [1] CRAN (R 4.4.0)
   stringr                1.5.1    2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.36.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   survival               3.8-3    2024-12-17 [1] CRAN (R 4.4.3)
   tensor                 1.5      2012-05-05 [1] CRAN (R 4.4.0)
   tibble                 3.2.1    2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1    2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1    2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.2.0    2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   uwot                   0.2.3    2025-02-24 [1] CRAN (R 4.4.1)
   vctrs                  0.6.5    2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7    2023-12-18 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2    2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.2    2024-10-28 [1] CRAN (R 4.4.1)
   xfun                   0.51     2025-02-19 [1] CRAN (R 4.4.1)
   xml2                   1.3.8    2025-03-14 [1] CRAN (R 4.4.1)
   xtable                 1.8-4    2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.46.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   yaml                   2.3.10   2024-07-26 [1] CRAN (R 4.4.0)
   zlibbioc               1.52.0   2024-11-08 [1] Bioconductor 3.20 (R 4.4.1)
   zoo                    1.8-13   2025-02-22 [1] CRAN (R 4.4.1)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 * ── Packages attached to the search path.
 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────────