---
title: "02b-BSpace hvg, svg q9-15-20-31 VMH, ARC subcluster marker genes"
author: "Bernie Mulvey"
date: "2023-06-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater) # addPerCellQC
library(BiocParallel)
library(scran)
library(parallel)
# library(scCustomize)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

load required files
```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")
colnames(hyp2) <- colData(hyp2)$key

bscl <- list.files("analysis/data/H02-clustering/bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out",pattern=".txt",full.names = T)

bscls <- lapply(bscl,FUN=fread)
names(bscls) <- gsub(bscl,pattern="^.*9-15-20-31_out\\/BSpace_Harmony_(.*)\\.txt$",replacement="\\1")

bs_markers <- readRDS("analysis/data/H03-marker_genes/BSpace_HVG1020-SVG1020_q9-15-20-31_findmarkers_out.RDS")
```

## Ran scran::scoreMarkers(...,full.stats=T) to get results compatabile w finding markers to separate highly related clusters e.g. nnsvg10_31 clusters 4 5 11 17
## following OSCA: http://bioconductor.org/books/3.17/OSCA.basic/marker-detection.html#obtaining-the-full-effects

### What distinguishes, e.g. 4, 5, 11, 17? 11 has a couple (GABRA5, HRH3) that do decently, but 4 5 and 17 are all FEZF1, SLC17A6, and ADCYAP1 as top markers by binomial analysis.
### I'm actually very confused about how to read the AUC tables here, so not going to explore further atm.
```{r}
vmhish <- c("4","5","11","17")
svg10_31_vmh <- bs_markers[["nnsvg10pctile_31"]][c(4,5,11,17)]
colnames(as.data.table(svg10_31_vmh[[1]]))

svg10_31_vmh_aucs <- lapply(svg10_31_vmh,FUN=function(x){
    y <- x$full.AUC[,colnames(x$full.AUC) %in% vmhish]
    z <- y[computeMinRank(y) <= 10,]
    as.data.table(z,keep.rownames=T)
})

```

### and using the ranked cohen's d for logFC, as OSCA explains: "If we select all genes with min-ranks less than or equal to T, the resulting set is the union of the top T genes from all pairwise comparisons. This guarantees that our set contains at least T genes that can distinguish our cluster of interest from any other cluster, which permits a comprehensive determination of a cluster’s identity."

```{r}
svg10_31_vmh_cod <- lapply(svg10_31_vmh,FUN=function(x){
    as.data.table(x[x$rank.logFC.cohen <= 4,],keep.rownames=T)
})

svg10_31_cl <- as.data.frame(bscls[["nnsvg20pctile_31"]])
rownames(svg10_31_cl) <- svg10_31_cl$rn
colLabels(hyp2) <- factor(svg10_31_cl[,2])


rownames(hyp2) <- rowData(hyp2)$gene_name

pdf("plots/H03-marker_genes/BSpace_nnsvg10_q31_cls4,5,11,17_cohen-fc-rank4_heat.pdf")
plotGroupedHeatmap(hyp2[,hyp2$label %in% c(4,5,11,17)],features = unique(c(svg10_31_vmh_cod[[1]]$rn,svg10_31_vmh_cod[[2]]$rn,svg10_31_vmh_cod[[3]]$rn,svg10_31_vmh_cod[[4]]$rn)),center=T,group="label")
dev.off()

```

Let's look at nnsvg10, q=15 by comparison, which seems to separate the inner VMH populations (above, 4 and 11) differently (5,6 interior, 9 post-curved-face, 1 outer margin).
```{r}

vmhish <- c("5","6","9","1")
svg10_15_vmh <- bs_markers[["nnsvg10pctile_15"]][c(1,5,6,9)]

svg10_15_vmh_aucs <- lapply(svg10_15_vmh,FUN=function(x){
    y <- x$full.AUC[,colnames(x$full.AUC) %in% vmhish]
    z <- y[computeMinRank(y) <= 10,]
    as.data.table(z,keep.rownames=T)
})

svg10_15_cl <- as.data.frame(bscls[["nnsvg20pctile_15"]])
rownames(svg10_15_cl) <- svg10_15_cl$rn
colLabels(hyp2) <- factor(svg10_15_cl[,2])

svg10_15_vmh_cod <- lapply(svg10_15_vmh,FUN=function(x){
    as.data.table(x[x$rank.logFC.cohen <= 4,],keep.rownames=T)
})

plotGroupedHeatmap(hyp2[,hyp2$label %in% c(1,5,6,9)],features = unique(c(svg10_15_vmh_cod[[1]]$rn,svg10_15_vmh_cod[[2]]$rn,svg10_15_vmh_cod[[3]]$rn,svg10_15_vmh_cod[[4]]$rn)),center=T,group="label")

### basically the same very limited scope of markers separating these clusters. lets look at this a second way, esp since we could use a few more genes for xenium. based on the two-cluster approach, only clusters 5 (inner cells)+6(exterior vmh) in the nnsvg10/k=15 set are really bonafide VMH, I think.
bscl <- fread("analysis/data/spe_053123/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/BSpace_Harmony_nnsvg10pctile_15.txt")

bscl <- as.data.frame(bscl)
rownames(bscl) <- bscl$rn
bscl <- DataFrame(bscl[,2])
colnames(bscl) <- "cl"
colLabels(hyp2) <- as.factor(bscl$cl)
rownames(hyp2) <- rowData(hyp2)$gene_name

n10k15 <- scran::scoreMarkers(hyp2,full.stats=T)

### important! we want CLUSTERS 5/6, not the 5th item of the result list (which is not 5) --> put in quotes
n10k15.5 <- as.data.table(n10k15[["5"]],keep.rownames=T)
n10k15.6 <- as.data.table(n10k15[["6"]],keep.rownames=T)
keepnames5 <- grep(names(n10k15.5),pattern="rn|X6|self|other|mean|median|min|max",value=T)
keepnames6 <- grep(names(n10k15.6),pattern="rn|X5|self|other|mean|median|min|max",value=T)
n10k15.5 <- n10k15.5[,..keepnames5]
n10k15.6 <- n10k15.6[,..keepnames6]

write.table(n10k15.5,"analysis/data/spe_053123/H03-marker_genes/VMH cl5 markers vs VMH clust 6 in harm-bspace nnsvg10 k15.txt",row.names=F,col.names=T,sep='\t',quote=F)
write.table(n10k15.6,"analysis/data/spe_053123/H03-marker_genes/VMH cl6 markers vs VMH clust 5 in harm-bspace nnsvg10 k15.txt",row.names=F,col.names=T,sep='\t',quote=F)

```


#### separating putative ARC clusters in harmony-bayesspace nnsvg10 k=15 (clusters #8 nad #13)
```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")
colnames(hyp2) <- colData(hyp2)$key
rownames(hyp2) <- rowData(hyp2)$gene_name

bscl <- fread("analysis/data/spe_053123/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/BSpace_Harmony_nnsvg10pctile_15.txt")

bscl <- as.data.frame(bscl)
rownames(bscl) <- bscl$rn
bscl <- bscl[colnames(hyp2),]
bscl <- DataFrame(bscl)
colnames(bscl)[2] <- "cl"
colLabels(hyp2) <- as.character(bscl[,2])

n10k15 <- scoreMarkers(hyp2,full.stats=T)

n10k15.13 <- as.data.table(n10k15[["13"]],keep.rownames=T)
n10k15.8 <- as.data.table(n10k15[["8"]],keep.rownames=T)

keepnames13 <- grep(names(n10k15.13),pattern="rn|X8|self|other|mean|median|min|max",value=T)
keepnames8 <- grep(names(n10k15.8),pattern="rn|X13|self|other|mean|median|min|max",value=T)
n10k15.13 <- n10k15.13[,..keepnames13]
n10k15.8 <- n10k15.8[,..keepnames8]

write.table(n10k15.8,"analysis/data/spe_053123/H03-marker_genes/ARC cluster 8 markers vs ARC clust 13 in harm-bspace nnsvg10 k15.txt",row.names=F,col.names=T,sep='\t',quote=F)

write.table(n10k15.13,"analysis/data/spe_053123/H03-marker_genes/ARC cluster 13 markers vs ARC clust 8 in harm-bspace nnsvg10 k15.txt",row.names=F,col.names=T,sep='\t',quote=F)
```
