---
title: "02-Tabulate various pre and post filter data factoids"
author: "Bernard Mulvey"
date: "2024-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(BiocParallel)
library(parallel)
library(magrittr)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)
```

get SPEs
```{r}
# original SPE, all spots
load("processed-data/02_build_spe/spe_raw_n10.Rdata")
unfilt <- spe
rm(spe)

# all spots overlapping tissue after alignment
alltissu <- unfilt[,unfilt$in_tissue]

# filtered, analyzed data
hyp2 <- readRDS("processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_071924.RDS")
```


```{r}
spelist <- list(unfilt,alltissu,hyp2)
names(spelist) <- c("raw","tiss_spots","final")

rm(hyp2,alltissu,unfilt)
gc(full=T)

sbp <- MulticoreParam(3)
register(sbp)
spemetrics <- bplapply(spelist,BPPARAM = sbp,FUN=function(x){
    # coldata to d.t
    tmp1 <- as.data.table(colData(x))
    
    # result table holder
    outtab <- as.data.table(unique(colData(x)[,c("sample_id","brnum")]))
    
    # the next several lines work because we haven't shuffled around the colData at all. a skeptic could confirm that by making each of these tables as above and merging them all, which is just clunkier
    tmp1[,nspots:=.N,by="sample_id"]
    tmp1[,mean_umis:=mean(sum_umi,na.rm=T),by="sample_id"]
    tmp1[,mean_genes:=mean(sum_gene,na.rm=T),by="sample_id"]
    tmp1[,median_umis:=median(sum_umi,na.rm=T),by="sample_id"]
    tmp1[,median_genes:=median(sum_gene,na.rm=T),by="sample_id"]
    outtab <- merge.data.table(outtab,tmp1[,.(sample_id,brnum,nspots,mean_umis,mean_genes,median_umis,median_genes)],by=c("sample_id","brnum"))
    outtab <- unique(outtab)
    return(outtab)
})
rm(spelist)
gc(full=T)

### tabulate
i<-1
for (i in c(1:length(spemetrics))){
    setnames(spemetrics[[i]],c(3:ncol(spemetrics[[i]])),paste0(
        names(spemetrics)[i],
        "_",
        names(spemetrics[[i]])[c(3:ncol(spemetrics[[i]]))]
        ))
    if(i==1){restab <- copy(spemetrics[[i]])}
    else{restab <- merge.data.table(restab,spemetrics[[i]],by=c("sample_id","brnum"))}
}
rm(i)

## calculate changes in each metric (tiss vs raw, final vs raw, final vs tiss)
val1 <- c("tiss_spots_","final_","final_")
val2 <- c("raw_","raw_","tiss_spots_")
mets <- c("nspots","mean_genes","median_genes","mean_umis","median_umis")

i<-1
for (i in c(1:length(val1))){
    curv1 <- val1[i]
    curv2 <- val2[i]
    
    j <- 1
    for (j in c(1:length(mets))){
        getv1 <- paste0(curv1,mets[j])
        getv2 <- paste0(curv2,mets[j])

        restab[,newcol:=get(getv1)-get(getv2)]
        newname <- paste0(mets[j],"_",curv1,"vs_",curv2)
        newname %<>% gsub(x=.,pattern="^(.*)_$",replacement="\\1")
        setnames(restab,"newcol",newname)
        rm(getv1,getv2,newname)
    }
    rm(j,curv1,curv2)
}
rm(i,val1,val2,mets)
```

### now, get the additional sample metrics output by spaceranger and bind them in
```{r}
samps <- restab$sample_id
samps # examine and append approp output dirs; different for the first 7 samples
samps[1:7] <- paste0("spaceranger_230511/",samps[1:7])
i<-1
for (i in c(1:length(samps))){
    curmd <- fread(paste0("processed-data/01_spaceranger/",samps[i],"/outs/metrics_summary.csv"))
    curmd[,sample_id:=gsub(samps[i],pattern="^spaceranger_230511/(V.*)$",replacement="\\1")]
    
    if(i==1){addnlmd <- copy(curmd)}
    else{addnlmd <- rbind(addnlmd,curmd)}
    rm(curmd)
}



restab <- merge.data.table(addnlmd,restab,by="sample_id")
```

# save (will be a supp tab)
```{r}
fwrite(restab,"processed-data/03-QC_filters/02-UMI_gene_nspot_metrics_pre-and-post-filter.txt",sep='\t',quote=F)
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] SpotSweeper_1.0.2           gridExtra_2.3              
 [3] scran_1.32.0                BiocParallel_1.38.0        
 [5] scater_1.32.0               scuttle_1.14.0             
 [7] spatialLIBD_1.16.2          ggspavis_1.10.0            
 [9] ggplot2_3.5.1               data.table_1.15.4          
[11] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
[13] SummarizedExperiment_1.34.0 Biobase_2.64.0             
[15] GenomicRanges_1.56.1        GenomeInfoDb_1.40.1        
[17] IRanges_2.38.1              S4Vectors_0.42.1           
[19] BiocGenerics_0.50.0         MatrixGenerics_1.16.0      
[21] matrixStats_1.3.0           rlang_1.1.4                

loaded via a namespace (and not attached):
  [1] later_1.3.2               BiocIO_1.14.0             bitops_1.0-7             
  [4] filelock_1.0.3            fields_16.2               tibble_3.2.1             
  [7] XML_3.99-0.17             lifecycle_1.0.4           edgeR_4.2.0              
 [10] doParallel_1.0.17         rprojroot_2.0.4           MASS_7.3-61              
 [13] lattice_0.22-6            magrittr_2.0.3            limma_3.60.3             
 [16] plotly_4.10.4             sass_0.4.9                rmarkdown_2.27           
 [19] jquerylib_0.1.4           yaml_2.3.9                metapod_1.12.0           
 [22] httpuv_1.6.15             spam_2.10-0               ggside_0.3.1             
 [25] sessioninfo_1.2.2         cowplot_1.1.3             DBI_1.2.3                
 [28] RColorBrewer_1.1-3        golem_0.4.1               maps_3.4.2               
 [31] abind_1.4-5               zlibbioc_1.50.0           purrr_1.0.2              
 [34] RCurl_1.98-1.14           rappdirs_0.3.3            GenomeInfoDbData_1.2.12  
 [37] ggrepel_0.9.5             irlba_2.3.5.1             terra_1.7-78             
 [40] dqrng_0.4.1               DelayedMatrixStats_1.26.0 codetools_0.2-20         
 [43] DelayedArray_0.30.1       DT_0.33                   tidyselect_1.2.1         
 [46] farver_2.1.2              UCSC.utils_1.0.0          ScaledMatrix_1.12.0      
 [49] viridis_0.6.5             shinyWidgets_0.8.6        BiocFileCache_2.12.0     
 [52] GenomicAlignments_1.40.0  jsonlite_1.8.8            escheR_1.4.0             
 [55] BiocNeighbors_1.22.0      iterators_1.0.14          foreach_1.5.2            
 [58] tools_4.4.1               Rcpp_1.0.13               glue_1.7.0               
 [61] SparseArray_1.4.8         xfun_0.45                 here_1.0.1               
 [64] dplyr_1.1.4               withr_3.0.0               BiocManager_1.30.23      
 [67] fastmap_1.2.0             bluster_1.14.0            fansi_1.0.6              
 [70] digest_0.6.36             rsvd_1.0.5                R6_2.5.1                 
 [73] mime_0.12                 colorspace_2.1-0          RSQLite_2.3.7            
 [76] config_0.3.2              utf8_1.2.4                tidyr_1.3.1              
 [79] generics_0.1.3            rtracklayer_1.64.0        httr_1.4.7               
 [82] htmlwidgets_1.6.4         S4Arrays_1.4.1            pkgconfig_2.0.3          
 [85] gtable_0.3.5              blob_1.2.4                XVector_0.44.0           
 [88] htmltools_0.5.8.1         dotCall64_1.1-1           scales_1.3.0             
 [91] png_0.1-8                 attempt_0.3.1             knitr_1.48               
 [94] rstudioapi_0.16.0         rjson_0.2.21              curl_5.2.1               
 [97] cachem_1.1.0              BiocVersion_3.19.1        parallel_4.4.1           
[100] vipor_0.4.7               AnnotationDbi_1.66.0      restfulr_0.0.15          
[103] pillar_1.9.0              grid_4.4.1                vctrs_0.6.5              
[106] promises_1.3.0            BiocSingular_1.20.0       dbplyr_2.5.0             
[109] beachmat_2.20.0           xtable_1.8-4              cluster_2.1.6            
[112] beeswarm_0.4.0            paletteer_1.6.0           evaluate_0.24.0          
[115] magick_2.8.3              cli_3.6.3                 locfit_1.5-9.10          
[118] compiler_4.4.1            Rsamtools_2.20.0          crayon_1.5.3             
[121] labeling_0.4.3            rematch2_2.1.2            spatialEco_2.0-2         
[124] ggbeeswarm_0.7.2          viridisLite_0.4.2         munsell_0.5.1            
[127] Biostrings_2.72.1         lazyeval_0.2.2            Matrix_1.7-0             
[130] ExperimentHub_2.12.0      benchmarkme_1.0.8         sparseMatrixStats_1.16.0 
[133] bit64_4.0.5               KEGGREST_1.44.1           statmod_1.5.0            
[136] shiny_1.8.1.1             AnnotationHub_3.12.0      igraph_2.0.3             
[139] memoise_2.0.1             bslib_0.7.0               benchmarkmeData_1.0.4    
[142] bit_4.0.5                
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-22
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package              * version   date (UTC) lib source
   abind                  1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi          1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub          3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat               2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme            1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData        1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager            1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors          1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular           1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion            3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings             2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                    4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                  4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                 1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                   1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   bluster                1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bslib                  0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                 1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 P cli                    3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   cluster                2.1.6     2023-12-01 [1] CRAN (R 4.4.1)
   codetools              0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorspace             2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   config                 0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                 1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                   5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                    1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                 2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel             1.0.17    2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64              1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                  0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DT                     0.33      2024-04-04 [1] CRAN (R 4.4.0)
   edgeR                  4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   escheR                 1.4.0     2024-05-16 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate               0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub          2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                  1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                 2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fields                 16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock               1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   foreach                1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics               0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments      1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges        * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   ggside                 0.3.1     2024-03-01 [1] CRAN (R 4.4.0)
   ggspavis             * 1.10.0    2024-05-16 [1] Bioconductor 3.19 (R 4.4.0)
   glue                   1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                  0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets            1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                 1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                   1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   igraph                 2.0.3     2024-03-13 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators              1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jquerylib              0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST               1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   knitr                  1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling               0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                  1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval               0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle              1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                  3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                 1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                 2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                   3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   MASS                   7.3-61    2024-06-13 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   metapod                1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   mime                   0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer              1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   pillar                 1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                 4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                    0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   promises               1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   purrr                  1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R6                     2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs               0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer           1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13    2024-07-17 [1] CRAN (R 4.4.1)
   RCurl                  1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rematch2               2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr               0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rjson                  0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools              2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer            1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   S4Arrays               1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                   0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix           1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scran                * 1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   shiny                  1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets           0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spam                   2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray            1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialEco             2.0-2     2023-11-17 [1] CRAN (R 4.4.0)
   SpatialExperiment    * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD          * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   SpotSweeper          * 1.0.2     2024-07-14 [1] Bioconductor 3.19 (R 4.4.1)
   statmod                1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                  1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                  1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                    3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xtable                 1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
