---
title: "01-Mouse subclass sexDE t stats to spatialHYP"
author: "Bernard Mulvey"
date: "2024-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(pheatmap)
```

```{r}
correls <- readRDS("processed-data/12-Mouse comparisons/03-Yao23-subclass-MvFsexDE-tstat_correls_to_k15-20-31-15collapsed_bytopNhumanDEGs.RDS")

correls <- correls[c("BSpace_k15_HARMONYlmbna_nnsvg10","BSpace_k20_HARMONYlmbna_nnsvg10","BSpace_k31_HARMONYlmbna_nnsvg10","BSpace_k15_HARMONYlmbna_nnsvg10_clpsdARC_clpsdVMH")]

## 
correls[[1]][human_domain=="X7",human_domain:="VMH.1"]
correls[[1]][human_domain=="X12",human_domain:="VMH.2"]
correls[[1]][human_domain=="X4",human_domain:="ARC.1"]
correls[[1]][human_domain=="X6",human_domain:="ARC.2"]

correls[[2]][human_domain=="X2",human_domain:="VMH20.1"]
correls[[2]][human_domain=="X3",human_domain:="VMH20.2"]
correls[[2]][human_domain=="X1",human_domain:="ARC20.1"]
correls[[2]][human_domain=="X4",human_domain:="ARC20.2"]

#k31
correls[[3]][human_domain=="X6",human_domain:="VMH31.1"]
correls[[3]][human_domain=="X21",human_domain:="VMH31.2"]
correls[[3]][human_domain=="X25",human_domain:="VMH31.3"]
correls[[3]][human_domain=="X26",human_domain:="VMH31.4"]
correls[[3]][human_domain=="X31",human_domain:="VMH31.5"]
correls[[3]][human_domain=="X2",human_domain:="ARC31.1"]
correls[[3]][human_domain=="X4",human_domain:="ARC31.2"]
correls[[3]][human_domain=="X11",human_domain:="ARC31.3"]
correls[[3]][human_domain=="X22",human_domain:="ARC31.4"]

correls <- lapply(correls,FUN=function(x){x[human_domain %in% grep(human_domain,pattern="VMH|ARC",value=T)]})

## change the top N that used all orthologs for a given cell type-spot pair to 10000 for the all case, so that these all get plotted together
correls <- lapply(correls,FUN=function(x){x[top_n_human>1000,top_n_human:=10000]})

mapply(X=correls,Y=names(correls),SIMPLIFY=FALSE,FUN=function(X,Y){
        i<-1
        for (i in unique(X$top_n_human)){
                pccs <- dcast(X[top_n_human==i], mouse_subclass ~ human_domain, value.var = "pcc")
                spears <- dcast(X[top_n_human==i], mouse_subclass ~ human_domain, value.var="spearmancor")
                pccs <- as.data.frame(pccs)
                rownames(pccs) <- pccs$mouse_subclass
                pccs$mouse_subclass <- NULL
        
                spears <- as.data.frame(spears)
                rownames(spears) <- spears$mouse_subclass
                spears$mouse_subclass <- NULL
                
                pdf(paste0("plots/12-Comparisons to other mouse and human datasets/01-Yaosubclass_sexDE_vs_",Y,"_top_",i,"genes_PCCs.pdf"),height=10,width=8)
                pheatmap(as.matrix(pccs),treeheight_col=0,treeheight_row = 0)
                dev.off()
        
                pdf(paste0("plots/12-Comparisons to other mouse and human datasets/01-Yaosubclass_sexDE_vs_",Y,"_top_",i,"genes_Spearman.pdf"),height=10,width=7)
                pheatmap(as.matrix(spears),treeheight_col=0,treeheight_row = 0)
                dev.off()
        }
        
})


```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] pheatmap_1.0.12   default_1.0.0     colorout_1.3-0.2  data.table_1.15.4
[5] rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] vctrs_0.6.5        cli_3.6.3          knitr_1.48         xfun_0.45         
 [5] stringi_1.8.4      glue_1.7.0         colorspace_2.1-0   plyr_1.8.9        
 [9] rprojroot_2.0.4    htmltools_0.5.8.1  rsconnect_1.3.1    scales_1.3.0      
[13] rmarkdown_2.27     grid_4.4.1         evaluate_0.24.0    munsell_0.5.1     
[17] fastmap_1.2.0      reshape2_1.4.4     yaml_2.3.9         lifecycle_1.0.4   
[21] stringr_1.5.1      compiler_4.4.1     sessioninfo_1.2.2  RColorBrewer_1.1-3
[25] Rcpp_1.0.13        here_1.0.1         rstudioapi_0.16.0  R.oo_1.26.0       
[29] R.utils_2.12.3     digest_0.6.36      R6_2.5.1           magrittr_2.0.3    
[33] R.methodsS3_1.8.2  tools_4.4.1        gtable_0.3.5      
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-07
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package      * version date (UTC) lib source
 P cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table   * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   default      * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
   digest         0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   evaluate       0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   glue           1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gtable         0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   pheatmap     * 1.0.12  2019-01-04 [1] CRAN (R 4.4.0)
   plyr           1.8.9   2023-10-02 [1] CRAN (R 4.4.0)
   R.methodsS3    1.8.2   2022-06-13 [1] CRAN (R 4.4.0)
   R.oo           1.26.0  2024-01-24 [1] CRAN (R 4.4.0)
   R.utils        2.12.3  2023-11-18 [1] CRAN (R 4.4.0)
   R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   RColorBrewer   1.1-3   2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp           1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   reshape2       1.4.4   2020-04-09 [1] CRAN (R 4.4.0)
 P rlang        * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rsconnect      1.3.1   2024-06-04 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   stringi        1.8.4   2024-05-06 [1] CRAN (R 4.4.0)
   stringr        1.5.1   2023-11-14 [1] CRAN (R 4.4.0)
   vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   xfun           0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
