---
title: "Fig2_Atlasing_Novelties"
output: html_document
date: "2024-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table) # Preferred data manipulation package
library(ggplot2) # Dependency for several plotting functions
library(ggtext) # more character types in ggplots
library(ggrastr) # avoid making raster images one can't even open
library(SpatialExperiment) # Visium data framework
library(SpatialFeatureExperiment) # Xenium data framework
library(spatialLIBD) # one option for plotting cluster assignments, but breaks when the in-tissue portion of a visium area is very un-square.
library(escheR) # alternative spotplotting function, at least for visium
library(viridis) # palettes
library(Polychrome) # better palettes
require(colorout) # Utility for RStudio
library(sf) # define the polygonal boundaries of xenium domains
ColorOut()

# code reformatting in Rstudio
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

# set plotting defaults for ggplot
theme_set(theme_bw() + theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 9), plot.title = element_blank(), strip.text = element_text(size = 10), legend.text = element_text(size = 8), legend.title = element_text(size = 8, hjust = 0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

#### VISIUM setup ####
```{r}
hyp2 <- readRDS("spatial_HYP/processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

bscl <- fread("spatial_HYP/processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt") # main visium clustering (bs=bayesspace)

setnames(bscl,2,"cl")
bscl[,cl:=paste0("Vis",cl)]
bscl[cl=="Vis7",cl:="VMH.1"]
bscl[cl=="Vis12",cl:="VMH.2"]
bscl[cl=="Vis4",cl:="ARC.1"]
bscl[cl=="Vis6",cl:="ARC.2"]

bscl[,dom:=cl]
bscl[dom %in% c("VMH.1","VMH.2"),dom:="VMH"]
bscl[dom %in% c("ARC.1","ARC.2"),dom:="ARC"]
bscl[!(dom %in% c("VMH","ARC")),dom:="Other"]

bscl<-DataFrame(bscl,row.names=bscl$rn)[colnames(hyp2),]
hyp2$k15dom <- bscl$cl
hyp2$`Visium Domain` <- factor(bscl$dom,levels=c("ARC","VMH","Other"))
```

### XENIUM setup ###
```{r}
hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

bksmooth <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt") ## xenium domain assignments after smoothing, by xenium cell

## but drop discarded cell clusters, too
bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"cl")

bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

## drop sample-specific clusters, append cluster annotations 
bkanno[,clus:=paste0("X",clus)]
bkanno <- unique(bkanno[,.(clus,bjm_annot)])
bkanno <- bkanno[bjm_annot!="DISCARD"&bjm_annot!="VMH_4_DISCARD"]

bkcl <- merge.data.table(bkcl,bkanno,by.x="cl",by.y="clus")
## drop excluded clusters from xenium spe
hypx <- hypx[,colnames(hypx) %in% bkcl$rn]

## append cluster annots to xenium spe
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
bkcl$rn <- NULL
hypx$banksyclus <- bkcl$bjm_annot

## assign the domain labels to the xenium cells
bksmooth[dualVMHARC4=="other",dualVMHARC4:="Other"]
bksmooth <- DataFrame(bksmooth,row.names=bksmooth$rn)[colnames(hypx),]
hypx$dom <- bksmooth$dualVMHARC4
```

## palette set up (shared)
```{r}
pal <- c("#8c63cf","#5dd959","#A9A7A9")
names(pal) <- c("VMH","ARC","Other")
```

Tentatively, the plan here is to have the top row be two novel markers each for VMH and ARC identified in Visium, their respective followup results in Xenium, then a row with LAMP5-SLC17A6-GAD-NR5A1 RNAscope, then a row with two Visium ARC SVGs that we subsequently examine in Xenium.

To show the robustness of the dataset, a different sample will be used on a per-figure basis. So here, let's use the classic V12Y31-080_A1 and its xenium counterpart, X86 reg1.

Panel A: LAMP5 in visium VMH
```{r}
### extract sample to plot ###
lamp <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(lamp) <- rowData(lamp)$gene_name
colData(lamp)$`log counts` <- logcounts(lamp)["LAMP5",]

### make spotplots
p <- make_escheR(lamp)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Visium Domain",stroke=0.15,point_size = 0.6)

pdf("manuscript_plots/Fig2/2A-LAMP5_Visium.pdf",height=1.6,width=1.75)
p+
    scale_color_manual(values=pal,na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("*LAMP5* (Visium)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(lamp,p)

```

Panel B: ANKRD34B in visium VMH
```{r}
### extract sample to plot ###
ankrd <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(ankrd) <- rowData(ankrd)$gene_name
colData(ankrd)$`log counts` <- logcounts(ankrd)["ANKRD34B",]

### make spotplots
p <- make_escheR(ankrd)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Visium Domain",stroke=0.15,point_size = 0.6)

pdf("manuscript_plots/Fig2/2B-ANKRD34B_Visium.pdf",height=1.6,width=1.75)
p+
    scale_color_manual(values=pal,na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("*ANKRD34B* (Visium)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(ankrd,p)

```

Panel C: GABRE in visium ARC
```{r}
### extract sample to plot ###
gabre <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(gabre) <- rowData(gabre)$gene_name
colData(gabre)$`log counts` <- logcounts(gabre)["GABRE",]

### make spotplots
p <- make_escheR(gabre)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Visium Domain",stroke=0.15,point_size = 0.6)

pdf("manuscript_plots/Fig2/2C-GABRE_Visium.pdf",height=1.6,width=1.75)
p+
    scale_color_manual(values=pal,na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("*GABRE* (Visium)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(gabre,p)

```


Panel D: CYP26A1 in visium ARC
```{r}
### extract sample to plot ###
cyp <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(cyp) <- rowData(cyp)$gene_name
colData(cyp)$`log counts` <- logcounts(cyp)["CYP26A1",]

### make spotplots
p <- make_escheR(cyp)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Visium Domain",stroke=0.15,point_size = 0.6)

pdf("manuscript_plots/Fig2/2D-CYP26A1_Visium.pdf",height=1.6,width=1.75)
p+
    scale_color_manual(values=pal,na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("*CYP26A1* (Visium)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(cyp,p)

```

For Xenium panels (E-G), we want to again have just the polygonal boundaries of the VMH and ARC shown. we previously code this up in the code for making figure 1 panels. we only need to do determine the points forming this polygon once to make all the panels.
```{r}
pane <- hypx["LAMP5",hypx$sample_id=="X86_reg1"]
# 
## attach the coordinate info to colData so we can make an sf object out of this for hull extraction. see well-annotated code from DE benchmarking of domain boundaries in xenium_HYP/code/07/03c-DE benchmarks...
tmpcoldata <- cbind(colData(pane),spatialCoords(pane))
panev <- st_as_sf(as.data.frame(tmpcoldata[tmpcoldata$dom=="VMH",]),coords = c("sdimx","sdimy"))
panev <- st_convex_hull(st_union(panev))

panea <- st_as_sf(as.data.frame(tmpcoldata[tmpcoldata$dom=="ARC",]),coords=c("sdimx","sdimy"))
panea <- st_concave_hull(st_union(panea),ratio = 0.1)

## create the data table we will use for overlaying the boundary polygons
dompoly <- cbind(panev[[1]][[1]],rep("VMH",nrow(panev[[1]][[1]])))
dompoly <- rbind(dompoly,cbind(panea[[1]][[1]],rep("ARC",nrow(panea[[1]][[1]]))))
dompoly <- as.data.table(dompoly)
setnames(dompoly,c("xpol","ypol","dompol"))
dompoly[,xpol:=as.numeric(xpol)]
dompoly[,ypol:=as.numeric(ypol)]

rm(panea,panev)
```

### Make the Xenium panels
```{r}
colData(pane)$`log counts` <- as.numeric(logcounts(pane)["LAMP5",])

p <- make_escheR(pane,y_reverse=FALSE)
p <- p |> add_fill("log counts",size=0.125,point_size = 0.125)

pdf("manuscript_plots/Fig2/2E-X86_reg1_LAMP5.pdf",height=1.6,width=1.75)
p+
    scale_fill_continuous("log\ncounts",low= "#FFFFFF",high="#000000")+
    geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol),linewidth = 0.25)+
    scale_color_manual(values=pal,na.value = NA)+
    ggtitle("*LAMP5* (Xenium)")+
    guides(color="none")+
    labs(col="Xenium\nDomain")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=9,hjust=0.5,margin=margin(0.15,0,-0.15,0,"in")),legend.text = element_text(size=7),legend.key.size = ggplot2::unit(0.125,"in"),legend.title=element_text(size=7.5,hjust=0.5),plot.margin = margin(-0.07,0,-0.07,-0.1,unit = "in"),legend.margin=margin(0,0,0,-0.15,"in"))
dev.off()

rm(pane,p)
```

panel F: ankrd34b
```{r}
panf <- hypx["ANKRD34B",hypx$sample_id=="X86_reg1"]
colData(panf)$`log counts` <- as.numeric(logcounts(panf)["ANKRD34B",])

p <- make_escheR(panf,y_reverse=FALSE)
p <- p |> add_fill("log counts",size=0.125,point_size = 0.125)

pdf("manuscript_plots/Fig2/2F-X86_reg1_ANKRD34B.pdf",height=1.6,width=1.75)
p+
    scale_fill_continuous("log\ncounts",low= "#FFFFFF",high="#000000")+
    geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol),linewidth = 0.25)+
    scale_color_manual(values=pal,na.value = NA)+
    ggtitle("*ANKRD34B* (Xenium)")+
    guides(color="none")+
    labs(col="Xenium\nDomain")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=9,hjust=0.5,margin=margin(0.15,0,-0.15,0,"in")),legend.text = element_text(size=7),legend.key.size = ggplot2::unit(0.125,"in"),legend.title=element_text(size=7.5,hjust=0.5),plot.margin = margin(-0.07,0,-0.07,-0.1,unit = "in"),legend.margin=margin(0,0,0,-0.15,"in"))
dev.off()

rm(panf,p)
```


panel G: GABRE
```{r}
pang <- hypx["GABRE",hypx$sample_id=="X86_reg1"]
colData(pang)$`log counts` <- as.numeric(logcounts(pang)["GABRE",])

p <- make_escheR(pang,y_reverse=FALSE)
p <- p |> add_fill("log counts",size=0.125,point_size = 0.125)

pdf("manuscript_plots/Fig2/2G-X86_reg1_GABRE.pdf",height=1.6,width=1.75)
p+
    scale_fill_continuous("log\ncounts",low= "#FFFFFF",high="#000000")+
    geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol),linewidth = 0.25)+
    scale_color_manual(values=pal,na.value = NA)+
    ggtitle("*GABRE* (Xenium)")+
    guides(color="none")+
    labs(col="Xenium\nDomain")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=9,hjust=0.5,margin=margin(0.15,0,-0.15,0,"in")),legend.text = element_text(size=7),legend.key.size = ggplot2::unit(0.125,"in"),legend.title=element_text(size=7.5,hjust=0.5),plot.margin = margin(-0.07,0,-0.07,-0.1,unit = "in"),legend.margin=margin(0,0,0,-0.15,"in"))
dev.off()

rm(pang,p)
```

panel H: CYP26A1
```{r}
panh <- hypx["CYP26A1",hypx$sample_id=="X86_reg1"]
colData(panh)$`log counts` <- as.numeric(logcounts(panh)["CYP26A1",])

p <- make_escheR(panh,y_reverse=FALSE)
p <- p |> add_fill("log counts",size=0.125,point_size = 0.125)

pdf("manuscript_plots/Fig2/2H-X86_reg1_CYP26A1.pdf",height=1.6,width=1.75)
p+
    scale_fill_continuous("log\ncounts",low= "#FFFFFF",high="#000000")+
    geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol),linewidth = 0.25)+
    scale_color_manual(values=pal,na.value = NA)+
    ggtitle("*CYP26A1* (Xenium)")+
    guides(color="none")+
    labs(col="Xenium\nDomain")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=9,hjust=0.5,margin=margin(0.15,0,-0.15,0,"in")),legend.text = element_text(size=7),legend.key.size = ggplot2::unit(0.125,"in"),legend.title=element_text(size=7.5,hjust=0.5),plot.margin = margin(-0.07,0,-0.07,-0.1,unit = "in"),legend.margin=margin(0,0,0,-0.15,"in"))
dev.off()

rm(panh,p)
```

I J K L M: RNAscope.

N: Visium ARC svg #1: GHRH
O: Visium ARC svg #2: TAC3
P: dual plot of the two on xenium?

Panel N: Visium GHRH
```{r}
pann <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(pann) <- rowData(pann)$gene_name
colData(pann)$`log counts` <- as.numeric(logcounts(pann)["GHRH",])

p <- make_escheR(pann)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Visium Domain",stroke=0.15,point_size = 0.6)

pdf("manuscript_plots/Fig2/2N-GHRH_visium.pdf",height=1.6,width=1.75)
p+
    scale_color_manual(values=pal,na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("*GHRH* (Visium)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(pann,p)
```

Panel O: Visium TAC3
```{r}
pano <- hyp2[,hyp2$sample_id=="V12Y31-080_A1"]
rownames(pano) <- rowData(pano)$gene_name
colData(pano)$`log counts` <- as.numeric(logcounts(pano)["TAC3",])

p <- make_escheR(pano)
p <- p |> add_fill("log counts",size=0.25,point_size = 0.6)
p <- p |> add_ground(var = "Visium Domain",stroke=0.15,point_size = 0.6)

pdf("manuscript_plots/Fig2/2O-TAC3_visium.pdf",height=1.6,width=1.75)
p+
    scale_color_manual(values=pal,na.value = NA)+
    scale_fill_gradient(low="white",high="black")+
    ggtitle("*TAC3* (Visium)")+
    guides(color="none")+
    labs(fill="log\ncounts")+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=10,hjust=0.5),legend.text = element_text(size=8),legend.key.size = ggplot2::unit(0.125,"in") ,legend.title=element_text(size=9,hjust=0.5),plot.margin = margin(-0.05,0.03,-0.05,0,unit = "in"))
dev.off()

rm(pano,p)
```


Let's see if we can plot both in Xenium in one panel.
```{r}
pano <- hypx[c("GHRH","TAC3"),hypx$sample_id=="X86_reg1"]
colData(pano)$`log counts GHRH` <- as.numeric(logcounts(pano)[c("GHRH"),])
colData(pano)$`log counts TAC3` <- as.numeric(logcounts(pano)[c("TAC3"),])

### adapting code from Boyi Guo/the "bivariate colorizer" code sandbox
RBG_norm_GHRH <- scales::rescale(
logcounts(pano)["GHRH", ] |>
  base::scale(center = TRUE, scale = FALSE) |>
  sapply(max, 0.2),
to = c(0.2,1)
)
RBG_norm_TAC3 <- scales::rescale(
logcounts(pano)["TAC3", ] |>
  base::scale(center = TRUE, scale = FALSE) |>
  sapply(max, 0.2),
to = c(0.2, 1)
)

### assign each cell a single color based on its combo of these features
pano$rbg_val <- rgb(
red = RBG_norm_GHRH,
green = 0,
blue= RBG_norm_TAC3
)

### find the color corresponding to most spots and blank those out
as.data.table(pano$rbg_val)[,.N,by="V1"][max(N)] # #330033
pano$rbg_val[which(pano$rbg_val=="#330033")] <- NA
####

p <- make_escheR(pano,y_reverse=FALSE)
p <- p |> add_fill("rbg_val",size=0.125,point_size = 0.125)

p <- p+
    scale_fill_identity()+
    geom_path(data=dompoly,aes(x=xpol,y=ypol,group=dompol,col=dompol),linewidth = 0.25)+scale_color_manual(values=pal,na.value = NA)+
    theme(plot.title.position = "plot",plot.title = element_markdown(size=9,hjust=0.5,margin=margin(0.15,0,-0.15,0,"in")),legend.text = element_text(size=7),legend.key.size = ggplot2::unit(0.125,"in"),legend.title=element_text(size=7.5,hjust=0.5),plot.margin = margin(-0.07,0,-0.07,-0.1,unit = "in"),legend.margin=margin(0,0,0,-0.15,"in"))

pdf("manuscript_plots/Fig2/2P-X86_reg1_GHRH-TAC3.pdf",height=1.6,width=1.75)
p
dev.off()



dev.off()

rm(pano,p)
```

### reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] lubridate_1.9.3                forcats_1.0.0                 
 [3] stringr_1.5.1                  dplyr_1.1.4                   
 [5] purrr_1.0.2                    readr_2.1.5                   
 [7] tidyr_1.3.1                    tibble_3.2.1                  
 [9] tidyverse_2.0.0                gridExtra_2.3                 
[11] sf_1.0-16                      colorout_1.3-0.2              
[13] Polychrome_1.5.1               viridis_0.6.5                 
[15] viridisLite_0.4.2              escheR_1.4.0                  
[17] spatialLIBD_1.16.2             SpatialFeatureExperiment_1.6.1
[19] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[21] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[23] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
[25] IRanges_2.38.1                 S4Vectors_0.42.1              
[27] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[29] matrixStats_1.3.0              ggrastr_1.0.2                 
[31] ggtext_0.1.2                   ggplot2_3.5.1                 
[33] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] bitops_1.0-7              EBImage_4.46.0            httr_1.4.7               
  [4] RColorBrewer_1.1-3        doParallel_1.0.17         tools_4.4.1              
  [7] utf8_1.2.4                R6_2.5.1                  DT_0.33                  
 [10] HDF5Array_1.32.0          lazyeval_0.2.2            rhdf5filters_1.16.0      
 [13] withr_3.0.0               sp_2.1-4                  cli_3.6.3                
 [16] labeling_0.4.3            sass_0.4.9                proxy_0.4-27             
 [19] Rsamtools_2.20.0          commonmark_1.9.1          R.utils_2.12.3           
 [22] scater_1.32.0             sessioninfo_1.2.2         attempt_0.3.1            
 [25] maps_3.4.2                limma_3.60.3              rstudioapi_0.16.0        
 [28] RSQLite_2.3.7             generics_0.1.3            BiocIO_1.14.0            
 [31] spdep_1.3-5               Matrix_1.7-0              ggbeeswarm_0.7.2         
 [34] fansi_1.0.6               abind_1.4-5               R.methodsS3_1.8.2        
 [37] terra_1.7-78              lifecycle_1.0.4           scatterplot3d_0.3-44     
 [40] yaml_2.3.9                edgeR_4.2.0               rhdf5_2.48.0             
 [43] SparseArray_1.4.8         BiocFileCache_2.12.0      paletteer_1.6.0          
 [46] grid_4.4.1                blob_1.2.4                promises_1.3.0           
 [49] dqrng_0.4.1               ExperimentHub_2.12.0      crayon_1.5.3             
 [52] lattice_0.22-6            beachmat_2.20.0           cowplot_1.1.3            
 [55] KEGGREST_1.44.1           magick_2.8.3              zeallot_0.1.0            
 [58] pillar_1.9.0              knitr_1.48                rjson_0.2.21             
 [61] boot_1.3-30               codetools_0.2-20          wk_0.9.2                 
 [64] glue_1.7.0                vctrs_0.6.5               png_0.1-8                
 [67] spam_2.10-0               gtable_0.3.5              rematch2_2.1.2           
 [70] cachem_1.1.0              xfun_0.45                 S4Arrays_1.4.1           
 [73] mime_0.12                 DropletUtils_1.24.0       sfheaders_0.4.4          
 [76] iterators_1.0.14          units_0.8-5               fields_16.2              
 [79] statmod_1.5.0             bit64_4.0.5               filelock_1.0.3           
 [82] rprojroot_2.0.4           bslib_0.7.0               irlba_2.3.5.1            
 [85] vipor_0.4.7               KernSmooth_2.23-24        colorspace_2.1-0         
 [88] spData_2.3.1              DBI_1.2.3                 tidyselect_1.2.1         
 [91] bit_4.0.5                 compiler_4.4.1            curl_5.2.1               
 [94] BiocNeighbors_1.22.0      xml2_1.3.6                DelayedArray_0.30.1      
 [97] plotly_4.10.4             rtracklayer_1.64.0        scales_1.3.0             
[100] classInt_0.4-10           rappdirs_0.3.3            tiff_0.1-12              
[103] digest_0.6.36             fftwtools_0.9-11          rmarkdown_2.27           
[106] benchmarkmeData_1.0.4     XVector_0.44.0            htmltools_0.5.8.1        
[109] pkgconfig_2.0.3           jpeg_0.1-10               sparseMatrixStats_1.16.0 
[112] dbplyr_2.5.0              fastmap_1.2.0             htmlwidgets_1.6.4        
[115] UCSC.utils_1.0.0          shiny_1.8.1.1             DelayedMatrixStats_1.26.0
[118] farver_2.1.2              jquerylib_0.1.4           jsonlite_1.8.8           
[121] BiocParallel_1.38.0       config_0.3.2              R.oo_1.26.0              
[124] BiocSingular_1.20.0       RCurl_1.98-1.14           magrittr_2.0.3           
[127] scuttle_1.14.0            GenomeInfoDbData_1.2.12   s2_1.1.6                 
[130] dotCall64_1.1-1           Rhdf5lib_1.26.0           munsell_0.5.1            
[133] Rcpp_1.0.13               stringi_1.8.4             zlibbioc_1.50.0          
[136] AnnotationHub_3.12.0      parallel_4.4.1            ggrepel_0.9.5            
[139] deldir_2.0-4              Biostrings_2.72.1         gridtext_0.1.5           
[142] hms_1.1.3                 locfit_1.5-9.10           markdown_1.13            
[145] ScaledMatrix_1.12.0       BiocVersion_3.19.1        XML_3.99-0.17            
[148] evaluate_0.24.0           golem_0.4.1               BiocManager_1.30.23      
[151] tzdb_0.4.0                foreach_1.5.2             httpuv_1.6.15            
[154] benchmarkme_1.0.8         rsvd_1.0.5                xtable_1.8-4             
[157] restfulr_0.0.15           e1071_1.7-14              later_1.3.2              
[160] class_7.3-22              memoise_2.0.1             beeswarm_0.4.0           
[163] AnnotationDbi_1.66.0      GenomicAlignments_1.40.0  timechange_0.3.0         
[166] shinyWidgets_0.8.6        here_1.0.1               
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-08
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0    2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1     2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8     2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4     2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23   2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel               1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular               1.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1    2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4     2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0     2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   commonmark                 1.9.1     2024-01-30 [1] CRAN (R 4.4.0)
   config                     0.3.2     2023-08-30 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3     2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1     2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0     2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17    2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64                  1.1-1     2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                    * 1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33      2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   escheR                   * 1.4.0     2024-05-16 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   fields                     16.2      2024-06-27 [1] CRAN (R 4.4.0)
   filelock                   1.0.3     2023-12-11 [1] CRAN (R 4.4.0)
   forcats                  * 1.0.0     2023-01-29 [1] CRAN (R 4.4.0)
   foreach                    1.5.2     2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5     2024-01-10 [1] CRAN (R 4.4.0)
   ggtext                   * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1     2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext                   0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   hms                        1.1.3     2023-03-21 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15    2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14    2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4     2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.1    2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   later                      1.3.2     2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2     2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   lubridate                * 1.9.3     2023-09-27 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2     2023-12-15 [1] CRAN (R 4.4.0)
   markdown                   1.13      2024-06-04 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1     2021-11-26 [1] CRAN (R 4.4.0)
   mime                       0.12      2021-09-28 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0     2024-01-21 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4    2024-01-13 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   Polychrome               * 1.5.1     2022-05-03 [1] CRAN (R 4.4.0)
   promises                   1.3.0     2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   purrr                    * 1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3     2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3     2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   readr                    * 2.1.5     2024-01-10 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2     2020-05-01 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15    2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7     2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5     2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                       0.4.9     2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scatterplot3d              0.3-44    2023-05-05 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1   2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6     2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0    2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2    2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   stringi                    1.8.4     2024-05-06 [1] CRAN (R 4.4.0)
   stringr                  * 1.5.1     2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                   * 3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                    * 1.3.1     2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tidyverse                * 2.0.0     2023-02-22 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   timechange                 0.3.0     2024-01-18 [1] CRAN (R 4.4.0)
   tzdb                       0.4.0     2023-05-12 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   viridis                  * 0.6.5     2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite              * 0.4.2     2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17 2024-06-25 [1] CRAN (R 4.3.3)
   xml2                       1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   xtable                     1.8-4     2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────