---
title: "Fig4_Attributing_SexDE_to_celltypes"
output: html_document
date: "2024-08-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table) # Preferred data manipulation package
library(ggplot2) # Dependency for several plotting functions
library(ggtext) # more character types in ggplots
library(ggrepel) # text labels in volcano plots
library(ggrastr) # avoid making raster images one can't even open
library(SpatialExperiment) # Visium data framework
library(SpatialFeatureExperiment) # Xenium data framework
library(spatialLIBD) # one option for plotting cluster assignments, but breaks when the in-tissue portion of a visium area is very un-square.
library(escheR) # alternative spotplotting function, at least for visium
library(viridis) # palettes
library(Polychrome) # better palettes
library(ggstance) # for y axis dodge
library(sf) # define the polygonal boundaries of xenium domains

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw() + theme(axis.text.x = element_text(size = 8), axis.title.x = element_text(size = 9), axis.text.y = element_text(size = 8), axis.title.y = element_text(size = 9), plot.title = element_blank(), strip.text = element_text(size = 10), legend.text = element_text(size = 8), legend.title = element_text(size = 8, hjust = 0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```



#### VISIUM setup ####
```{r}
hyp2 <- readRDS("spatial_HYP/processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_072224.RDS")

bscl <- fread("spatial_HYP/processed-data/06-BayesSpace/01-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt") # main visium clustering (bs=bayesspace)

setnames(bscl,2,"cl")
bscl[,cl:=paste0("Vis",cl)]
bscl[cl=="Vis7",cl:="VMH.1"]
bscl[cl=="Vis12",cl:="VMH.2"]
bscl[cl=="Vis4",cl:="ARC.1"]
bscl[cl=="Vis6",cl:="ARC.2"]

bscl[,dom:=cl]
bscl[dom %in% c("VMH.1","VMH.2"),dom:="VMH"]
bscl[dom %in% c("ARC.1","ARC.2"),dom:="ARC"]
bscl[!(dom %in% c("VMH","ARC")),dom:="Other"]

bscl<-DataFrame(bscl,row.names=bscl$rn)[colnames(hyp2),]
hyp2$k15dom <- bscl$cl
hyp2$`Visium Domain` <- factor(bscl$dom,levels=c("ARC","VMH","Other"))
```

### XENIUM setup ###
```{r}
hypx <- readRDS("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

bksmooth <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt") ## xenium domain assignments after smoothing, by xenium cell

## but drop discarded cell clusters, too
bkcl <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")
setnames(bkcl,2,"cl")

bkanno <- fread("xenium_HYP/processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")

## drop sample-specific clusters, append cluster annotations 
bkanno[,clus:=paste0("X",clus)]
bkanno <- unique(bkanno[,.(clus,bjm_annot)])
bkanno <- bkanno[bjm_annot!="DISCARD"&bjm_annot!="VMH_4_DISCARD"]

## convert ARC_1..etc to more descriptive cluster IDs defined based on ARC/VMH specific cluster marker analyses
arcvmhanno <- fread("xenium_HYP/processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/04e_ARC_and_VMH_cluster_identities_based_on_cellgroup_1vall.txt")
bkanno <- merge.data.table(bkanno,arcvmhanno,by="bjm_annot",all.x=T)
bkanno[!is.na(subclus_annot),bjm_annot:=subclus_annot]

## this drops rows that were assigned one of the discarded types (ie no longer in the cell type annotations)
bkcl <- merge.data.table(bkcl,bkanno,by.x="cl",by.y="clus")
## and now we can drop excluded clusters from xenium spe
hypx <- hypx[,colnames(hypx) %in% bkcl$rn]

## append cluster annots to xenium spe
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)[colnames(hypx),]
bkcl$rn <- NULL
hypx$banksyclus <- bkcl$bjm_annot

## assign the domain labels to the xenium cells
bksmooth[dualVMHARC4=="other",dualVMHARC4:="Other"]
bksmooth <- DataFrame(bksmooth,row.names=bksmooth$rn)[colnames(hypx),]
hypx$dom <- bksmooth$dualVMHARC4
```

### sex DE results
```{r}
visde <- fread("spatial_HYP/processed-data/09-Sex DE/01c-voomLmFit_nnsvg10-HmnylmbNA-BS-15-VMHARCclpsd.txt")
### exclude sex chromosomal, mito genes (need to append this info)
genelut <- fread("spatial_HYP/raw-data/USEFOR10X_ensg_lut_hg38_fromEns98_111123.txt")
genelut <- unique(genelut[,.(ensembl_gene_id,chromosome_name)])
setnames(genelut,c("ens","chr"))
visde <- merge.data.table(visde,genelut,by.x="gene_id",by.y="ens")
visde <- visde[!(chr %in% c("MT","X","Y"))]

### load by cell type and full-domains xenium results from 4 ARC-cell-type-smoothing domain
xenarcde <- fread("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/02a-4ARCtypesmoothed_withinARC_celltypeSexDE.txt")
xenvmhde <- fread("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/02b-4ARCtypesmoothed_withinVMH_celltypeSexDE.txt")

xendomde <- fread("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/04-4typeARC-and-dualassignVMH_domainwise_sexDE.txt")

### for correlation plot, merge xenium domain and visium arc/vmh stats
vistmp <- copy(visde[assay %in% c("VMH","ARC")])
setnames(vistmp,"logFC","logFC_Visium")

setnames(xendomde,"logFC_MvF","logFC_Xenium")

xendomde <- merge.data.table(xendomde[,.(domain,gene_name,logFC_Xenium)],vistmp[,.(assay,gene_name,logFC_Visium)],by.x=c("domain","gene_name"),by.y=c("assay","gene_name"))

## clean up
rm(genelut,vistmp)

# also load pseudobulk ELists used in the DE analysis for plotting DE from pseudobulk results
arcelist <- readRDS("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/02c-pseudobulkARC_voomLmFitELists.RDS")
vmhelist <- readRDS("xenium_HYP/processed-data/08_VMH-ARC cell type sex DE within domains/02d-pseudobulkVMH_voomLmFitELists.RDS")

## subset the pseudobulk tables to the ARC clusters for the arc domain and VMH for vmh
arcelist <- arcelist[bkanno[bjm_annot %in% grep(bjm_annot,pattern= "ARC",value=T),clus]]
vmhelist <- vmhelist[bkanno[bjm_annot %in% grep(bjm_annot,pattern= "VMH",value=T),clus]]

## change the pseudobulk entry names to be the more granularly annotated cell types
names(vmhelist) <-  bkanno[match(names(vmhelist),bkanno$clus),bjm_annot]
names(arcelist) <-  bkanno[match(names(arcelist),bkanno$clus),bjm_annot]
```

cell types palettes
```{r}
vmhpal <- c("#8c63cf", "#6A0D53", "#D193FE", "#AD260D")
names(vmhpal) <-  names(vmhelist)

arcpal <- c("#5dd959","#4CAF50", "#0DAD26","#2E8B57","#0D6A0D")
names(arcpal) <-  names(arcelist)

```

Panel A: plot the VMH cells on an exemplary sample, showing only the VMH area
```{r}
pana <- hypx[,hypx$sample_id=="X99_1225A"&hypx$dom=="VMH"]
pana <- as.data.table(cbind(spatialCoords(pana),as.data.table(colData(pana))))[,.(sdimx,sdimy,banksyclus)]

pana <- pana[banksyclus %in% grep(banksyclus,pattern="VMH",value=T)]
pana[,banksyclus:=gsub(banksyclus,pattern="_",replacement=" ")]

pdf("manuscript_plots/Fig4/4A-VMH_celltypes_1225A.pdf",height=3.75,width=3.75)
ggplot(pana[sdimy<6500],aes(x=sdimx,y=sdimy,color=banksyclus))+
    geom_point(size=0.075)+
    scale_color_manual(values=vmhpal)+
    guides(color=guide_legend(override.aes = list(size=1)))+
    labs(color="VMH Cluster")+
    theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank(),axis.title.x=element_blank(),axis.ticks = element_blank(),legend.background = element_blank(),legend.box=element_blank(),legend.margin=margin(0,0,0,-0.15,unit="in"),plot.margin = margin(0.01,0,0,0,unit="in"))
dev.off()


rm(pana)
```

Panel B: same sample, showing ARC types in ARC domain.
```{r}
panb <- hypx[,hypx$sample_id=="X99_1225A"&hypx$dom=="ARC"]

panb <- as.data.table(cbind(spatialCoords(panb),as.data.table(colData(panb))))[,.(sdimx,sdimy,banksyclus)]

panb <- panb[banksyclus %in% grep(banksyclus,pattern="ARC",value=T)]
panb[,banksyclus:=gsub(banksyclus,pattern="_",replacement=" ")]

arcpal["ARC TAC3-AR-ESR1"] <- "#D0B"

pdf("manuscript_plots/Fig4/4B-ARC_celltypes_1225A.pdf",height=3.75,width=3.75)
ggplot(panb,aes(x=sdimx,y=sdimy,color=banksyclus))+
    geom_point(size=0.075)+
    scale_color_manual(values=arcpal)+
    guides(color=guide_legend(override.aes = list(size=1)))+
    labs(color="ARC Cluster")+
    theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.title.y=element_blank(),axis.title.x=element_blank(),axis.ticks = element_blank(),legend.background = element_blank(),legend.box=element_blank(),legend.margin=margin(0,0,0,-0.15,unit="in"),plot.margin = margin(0.01,0,0,0,unit="in"))
dev.off()








```


Panel C: Show a DEG per VMH cluster, shown for all VMH xenium clusters in a violin plot datawide (at the level of pseudobulked samples). VMH_1: MC4R, VMH_2 (and 3): TAC1, VMH_3: TMEM233
```{r}
vmhe <- rbindlist(vmhelist,idcol = "celltype",fill=T)
# removing leading "elist." from colnames
setnames(vmhe,gsub(names(vmhe),pattern="^elist\\.",replacement=""))

# melt for genes of interest
vmhe.m <- melt(vmhe[rn %in% c("MC4R","TAC1","TMEM233")],id.vars=c("rn","celltype"))

# append sex info
vmhe.m <- merge.data.table(vmhe.m,unique(as.data.table(colData(hypx))[,.(sample_id,Sex)]),by.x="variable",by.y="sample_id")
vmhe.m[,Sex:=gsub(Sex,pattern="^(.).*$",replacement="\\1")]

## tidy up for plotting -- also do this for the palette or they won't match up
vmhe.m[,celltype:=gsub(celltype,pattern="_",replacement=" ")]
names(vmhpal) <- gsub(names(vmhpal),pattern="_",replacement=" ")
setnames(vmhe.m,"celltype","Xenium Cluster")

# make the gene names markdown'ed so they appear in italics
vmhe.m[,rn:=paste0("*",rn,"*")]

pdf("manuscript_plots/Fig4/4C-VMH_DEGs_boxplot.pdf",height=3,width=3.5)
ggplot(vmhe.m,aes(x=`Xenium Cluster`,y=value,color=`Xenium Cluster`,fill = Sex))+
    geom_boxplot(outliers = FALSE,linewidth = 0.25)+
    geom_point(size=0.35,position=position_jitterdodge(jitter.width = 0.15))+
    scale_fill_manual(values=c("#FFFFFF","#000000"))+
    scale_color_manual(values=vmhpal)+
    guides(color="none")+
    ylab("Pseudobulk log CPM")+
    facet_wrap(~rn,scales = "free_y")+
    theme(strip.text = element_markdown(size = 8,margin=margin(0.1,0,0,0,unit="in")),strip.background = element_blank(),axis.text.x=element_text(angle=90,size=7,hjust=1,vjust=0.5),axis.title.x=element_text(size=8),axis.title.y=element_text(size=8),axis.text.y=element_text(size=7),plot.margin = margin(-0.065,-0.05,0,0.02,unit="in"),legend.background = element_blank(),legend.box=element_blank(),legend.margin=margin(0,0.1,0,-0.075,unit="in"))
dev.off()


# clean up
rm(vmhe.m,vmhe)
```

## for panel D, we want to do the same for three ARC genes of interest. For ARC_4 we have the pick of the litter--we'll go with CDH13, which has known associations to weight. For ARC_3, RXRG is pretty high up there signif wise. And ARC_1 has signif CYP26A1 DE.  
```{r}
arce <- rbindlist(arcelist,idcol = "celltype",fill=T)
# removing leading "elist." from colnames
setnames(arce,gsub(names(arce),pattern="^elist\\.",replacement=""))

# melt for genes of interest
arce.m <- melt(arce[rn %in% c("CDH13","RXRG","CYP26A1")],id.vars=c("rn","celltype"))

# append sex info
arce.m <- merge.data.table(arce.m,unique(as.data.table(colData(hypx))[,.(sample_id,Sex)]),by.x="variable",by.y="sample_id")
arce.m[,Sex:=gsub(Sex,pattern="^(.).*$",replacement="\\1")]

## tidy up for plotting -- also do this for the palette or they won't match up
arce.m[,celltype:=gsub(celltype,pattern="_",replacement=" ")]
names(arcpal) <- gsub(names(arcpal),pattern="_",replacement=" ")
setnames(arce.m,"celltype","Xenium Cluster")

# make the gene names markdown'ed so they appear in italics
arce.m[,rn:=paste0("*",rn,"*")]

pdf("manuscript_plots/Fig4/4D-ARC_DEGs_boxplot.pdf",height=3,width=4)
ggplot(arce.m,aes(x=`Xenium Cluster`,y=value,color=`Xenium Cluster`,fill = Sex))+
    geom_boxplot(outliers = FALSE,linewidth = 0.25)+
    geom_point(size=0.35,position=position_jitterdodge(jitter.width = 0.15))+
    scale_fill_manual(values=c("#FFFFFF","#000000"))+
    scale_color_manual(values=arcpal)+
    guides(color="none")+
    ylab("Pseudobulk log CPM")+
    facet_wrap(~rn,scales = "free_y")+
    theme(strip.text = element_markdown(size = 8,margin=margin(0.1,0,0,0,unit="in")),strip.background = element_blank(),axis.text.x=element_text(angle=90,size=7,hjust=1,vjust=0.5),axis.title.x=element_text(size=8),axis.title.y=element_text(size=8),axis.text.y=element_text(size=7),plot.margin = margin(-0.065,-0.05,0,0.02,unit="in"),legend.background = element_blank(),legend.box=element_blank(),legend.margin=margin(0,0.1,0,-0.075,unit="in"))
dev.off()


# clean up
rm(arce.m,arce)
```