---
title: "02-Marker TF-Target GSEA"
author: "Bernard Mulvey"
date: "2023-12-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(BiocParallel)
library(parallel)
library(parallelly)
library(fgsea)

require(colorout)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
options(bphost="localhost")

```

load SpatialLIBD marker enrichment stats and subset to clusterings of interest
```{r}
mks <- readRDS("data/07-Marker_genes/01b-BSpace_allruns_spatialLIBD-regwrap-enrichment_full.RDS")

mks <- mks[grep(names(mks),pattern="HARMONYlmbna_nnsvg10_60kiter")]
# check:
# names(mks)
# [1] "k15_HARMONYlmbna_nnsvg10_60kiter"          
# [2] "k20_HARMONYlmbna_nnsvg10_60kiter"          
# [3] "k31_HARMONYlmbna_nnsvg10_60kiter"          
# [4] "k15_HARMONYlmbna_nnsvg10_60kiter_collapsed"

```

### Enrichr TF-target databases gene sets retrieval
```{r}
tftargsets <- readRDS("data/11-GSEA/00a-enrichrTFsets_forfgseainput.RDS")
```

# MARKER gsea for TF-target sets
# note that we need SYMBOLS ("gene" column), not ens IDs in this case. the TF-target lists are only available as symbols
```{r}
### drop mito genes
mks <- lapply(mks,FUN=function(x){x[!(gene %in% grep(gene,pattern="MT-",value=T))]})

### get named vectors of marker t-stats for each cluster in each result set (max 15+20+31+13)=79; probably less for clusters that dropped from spatiallibd due to poor representation
mklist <- list()
i<-1
for (i in c(1:length(mks))){
    sdoms <- grep(names(mks[[i]]),pattern="logFC_",value=T)
    sdoms <- gsub(sdoms,pattern="^logFC_(.*)$",replacement="\\1")
    j<-1
    for (j in c(1:length(sdoms))){
        getcol <- paste0("t_stat_",sdoms[j])
        tmp <- copy(mks[[i]])[,c("gene",..getcol)]
        setorderv(tmp,getcol,-1)
        ## get the stat column as a vector and give it the gene ids as names for fgsea
        ## note that since data.table (invisibly) uses list() for column data, its cleaner to just make the working table a data frame for this last bit
        setnames(tmp,getcol,"stat")
        tmp <- as.data.frame(tmp)
        v <- tmp$stat
        names(v) <- tmp$gene
        mklist[[length(mklist)+1]] <- v
        names(mklist)[length(mklist)] <- paste0(names(mks)[i],"_",sdoms[j])
        rm(getcol,v,tmp)
    }
    rm(j,sdoms)
}
rm(i)
```

### run fgsea
### looping through each domain (each run is very fast with 10 cpus in biocparallel, so just serially run 10cpu GSEA for each cluster)

```{r}
sbp <- MulticoreParam(10)
register(sbp)
mkres <- mapply(X=mklist,Y=names(mklist),SIMPLIFY=FALSE,FUN=function(X,Y){
    z<-as.data.table(fgseaMultilevel(pathways=tftargsets,stats = X,minSize = 15,maxSize = 500,eps=0.0,BPPARAM = sbp,nPermSimple = 50000))
    z[,spdom:=Y]
    return(z)
})
names(mkres) <- names(mklist)
# saveRDS(mkres,"data/11-GSEA/02a-Marker TF-target GSEA svg10-hmnyLmdaNA_BS60k-k15-k20-k31-k15clpsd.RDS")

### collapse pathways-this part is WAY WAY slower than the GSEA itself and serialized, so bplapply it
sbp <- MulticoreParam(6)
register(sbp)

mkres2 <- bpmapply(X=mkres,Y=mklist,SIMPLIFY = FALSE, BPPARAM=sbp,FUN=function(X,Y){
    tmpres <- fgsea::collapsePathways(fgseaRes = X,stats=Y,pathways = tftargsets)
    return(tmpres)
})
names(mkres2) <- names(mkres)

saveRDS(mkres2,"data/11-GSEA/02b-Marker TF-target fGSEA-collapsePathways output.RDS")

rm(reslist,mklist,mks)
gc(full=T)

### subset to main enrichments
mkres3 <- mapply(X=mkres,Y=mkres2,SIMPLIFY = FALSE,FUN=function(X,Y){
    X[pathway %in% Y[[1]]]
})
saveRDS(mkres3,"data/11-GSEA/02c-Marker TF-target GSEA clpspath mainPathways-only.RDS")
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets 
[8] methods   base     

other attached packages:
 [1] default_1.0.0       colorout_1.3-0.1    fgsea_1.28.0       
 [4] parallelly_1.36.0   BiocParallel_1.36.0 Biostrings_2.70.1  
 [7] GenomeInfoDb_1.38.5 XVector_0.42.0      IRanges_2.36.0     
[10] S4Vectors_0.40.2    BiocGenerics_0.48.1 data.table_1.14.10 
[13] rlang_1.1.3        

loaded via a namespace (and not attached):
 [1] utf8_1.2.4              generics_0.1.3          bitops_1.0-7           
 [4] lattice_0.22-5          digest_0.6.34           magrittr_2.0.3         
 [7] evaluate_0.23           grid_4.3.2              fastmap_1.1.1          
[10] rprojroot_2.0.4         Matrix_1.6-5            sessioninfo_1.2.2      
[13] fansi_1.0.6             scales_1.3.0            codetools_0.2-19       
[16] cli_3.6.2               crayon_1.5.2            cowplot_1.1.3          
[19] munsell_0.5.0           yaml_2.3.8              tools_4.3.2            
[22] dplyr_1.1.4             colorspace_2.1-0        ggplot2_3.4.4          
[25] fastmatch_1.1-4         GenomeInfoDbData_1.2.11 here_1.0.1             
[28] vctrs_0.6.5             R6_2.5.1                lifecycle_1.0.4        
[31] zlibbioc_1.48.0         pkgconfig_2.0.3         pillar_1.9.0           
[34] gtable_0.3.4            glue_1.7.0              Rcpp_1.0.12            
[37] xfun_0.41               tibble_3.2.1            tidyselect_1.2.0       
[40] rstudioapi_0.15.0       knitr_1.45              htmltools_0.5.7        
[43] rmarkdown_2.25          compiler_4.3.2          RCurl_1.98-1.14        
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-01-28
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────
 package          * version   date (UTC) lib source
 BiocGenerics     * 0.48.1    2023-11-02 [1] Bioconductor
 BiocParallel     * 1.36.0    2023-10-26 [1] Bioconductor
 Biostrings       * 2.70.1    2023-10-25 [1] Bioconductor
 bitops             1.0-7     2021-04-24 [1] CRAN (R 4.3.2)
 cli                3.6.2     2023-12-11 [1] CRAN (R 4.3.2)
 codetools          0.2-19    2023-02-01 [1] CRAN (R 4.3.0)
 colorout         * 1.3-0.1   2024-01-11 [1] local
 colorspace         2.1-0     2023-01-23 [1] CRAN (R 4.3.2)
 cowplot            1.1.3     2024-01-22 [1] CRAN (R 4.3.1)
 crayon             1.5.2     2022-09-29 [1] CRAN (R 4.3.0)
 data.table       * 1.14.10   2023-12-08 [1] CRAN (R 4.3.2)
 default          * 1.0.0     2017-08-07 [1] CRAN (R 4.3.0)
 digest             0.6.34    2024-01-11 [1] CRAN (R 4.3.2)
 dplyr              1.1.4     2023-11-17 [1] CRAN (R 4.3.2)
 evaluate           0.23      2023-11-01 [1] CRAN (R 4.3.1)
 fansi              1.0.6     2023-12-08 [1] CRAN (R 4.3.2)
 fastmap            1.1.1     2023-02-24 [1] CRAN (R 4.3.2)
 fastmatch          1.1-4     2023-08-18 [1] CRAN (R 4.3.2)
 fgsea            * 1.28.0    2023-10-24 [1] Bioconductor
 generics           0.1.3     2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb     * 1.38.5    2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData   1.2.11    2024-01-11 [1] Bioconductor
 ggplot2            3.4.4     2023-10-12 [1] CRAN (R 4.3.1)
 glue               1.7.0     2024-01-09 [1] CRAN (R 4.3.2)
 gtable             0.3.4     2023-08-21 [1] CRAN (R 4.3.0)
 here               1.0.1     2020-12-13 [1] CRAN (R 4.3.0)
 htmltools          0.5.7     2023-11-03 [1] CRAN (R 4.3.2)
 IRanges          * 2.36.0    2023-10-24 [1] Bioconductor
 knitr              1.45      2023-10-30 [1] CRAN (R 4.3.2)
 lattice            0.22-5    2023-10-24 [1] CRAN (R 4.3.2)
 lifecycle          1.0.4     2023-11-07 [1] CRAN (R 4.3.1)
 magrittr           2.0.3     2022-03-30 [1] CRAN (R 4.3.2)
 Matrix             1.6-5     2024-01-11 [1] CRAN (R 4.3.2)
 munsell            0.5.0     2018-06-12 [1] CRAN (R 4.3.0)
 parallelly       * 1.36.0    2023-05-26 [1] CRAN (R 4.3.0)
 pillar             1.9.0     2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig          2.0.3     2019-09-22 [1] CRAN (R 4.3.0)
 R6                 2.5.1     2021-08-19 [1] CRAN (R 4.3.0)
 Rcpp               1.0.12    2024-01-09 [1] CRAN (R 4.3.2)
 RCurl              1.98-1.14 2024-01-09 [1] CRAN (R 4.3.1)
 rlang            * 1.1.3     2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown          2.25      2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot          2.0.4     2023-11-05 [1] CRAN (R 4.3.1)
 rstudioapi         0.15.0    2023-07-07 [1] CRAN (R 4.3.0)
 S4Vectors        * 0.40.2    2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 scales             1.3.0     2023-11-28 [1] CRAN (R 4.3.2)
 sessioninfo        1.2.2     2021-12-06 [1] CRAN (R 4.3.0)
 tibble             3.2.1     2023-03-20 [1] CRAN (R 4.3.2)
 tidyselect         1.2.0     2022-10-10 [1] CRAN (R 4.3.0)
 utf8               1.2.4     2023-10-22 [1] CRAN (R 4.3.2)
 vctrs              0.6.5     2023-12-01 [1] CRAN (R 4.3.2)
 xfun               0.41      2023-11-01 [1] CRAN (R 4.3.2)
 XVector          * 0.42.0    2023-10-24 [1] Bioconductor
 yaml               2.3.8     2023-12-11 [1] CRAN (R 4.3.2)
 zlibbioc           1.48.0    2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

───────────────────────────────────────────────────────────────────────────
