---
title: "04-Save Yao 23 Sex DE Comparison Results to Supp Tab Sheets"
author: "Bernard Mulvey"
date: "2024-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

```

load dreamlet results
```{r}
dlt <- fread("processed-data/12-Mouse comparisons/02-Yao23-HYP-subclass-sexDE_dreamlet_donormouse-ranfx.txt")

setnames(dlt,c(1,2),c("Yao23 Cell Subclass","Mouse Gene Symbol"))
dlt <- dlt[,c(1,2,3,7,5,9,8,6,4)]

saveRDS(dlt,"local_supptab_assembly/Supp Data-Yao23 HYP sex DE with random effect of donor.RDS")

rm(dlt)
```

load -> write registration results
```{r}
sexreg <- readRDS("processed-data/12-Mouse comparisons/03-Yao23-subclass-MvFsexDE-tstat_correls_to_k15-20-31-15collapsed_bytopNhumanDEGs.RDS")

## name ARC and VMH clusters for the nnsvg10 harmony lambda=NA bayesspaces
sexreg[[2]][human_domain=="X4",human_domain:="ARC.1"]
sexreg[[2]][human_domain=="X6",human_domain:="ARC.2"]
sexreg[[2]][human_domain=="X7",human_domain:="VMH.1"]
sexreg[[2]][human_domain=="X12",human_domain:="VMH.2"]

sexreg[[8]][human_domain=="X2",human_domain:="VMH20.1"]
sexreg[[8]][human_domain=="X3",human_domain:="VMH20.2"]
sexreg[[8]][human_domain=="X1",human_domain:="ARC20.1"]
sexreg[[8]][human_domain=="X4",human_domain:="ARC20.2"]

sexreg[[14]][human_domain=="X6",human_domain:="VMH31.1"]
sexreg[[14]][human_domain=="X21",human_domain:="VMH31.2"]
sexreg[[14]][human_domain=="X25",human_domain:="VMH31.3"]
sexreg[[14]][human_domain=="X26",human_domain:="VMH31.4"]
sexreg[[14]][human_domain=="X31",human_domain:="VMH31.5"]
sexreg[[14]][human_domain=="X2",human_domain:="ARC31.1"]
sexreg[[14]][human_domain=="X4",human_domain:="ARC31.2"]
sexreg[[14]][human_domain=="X11",human_domain:="ARC31.3"]
sexreg[[14]][human_domain=="X22",human_domain:="ARC31.4"]

names(sexreg)

fn <- names(sexreg)
mapply(x=sexreg,y=fn,FUN=function(x,y){fwrite(x,paste0("local_supptab_assembly/Yao Sex DE correls ",y,".txt"),quote=F,sep='\t')})
rm(sexreg,fn)
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] default_1.0.0     colorout_1.3-0.2  data.table_1.15.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] digest_0.6.36     fastmap_1.2.0     xfun_0.45         knitr_1.48       
 [5] htmltools_0.5.8.1 rmarkdown_2.27    cli_3.6.3         compiler_4.4.1   
 [9] rprojroot_2.0.4   here_1.0.1        rstudioapi_0.16.0 tools_4.4.1      
[13] evaluate_0.24.0   yaml_2.3.9       
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-07
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package     * version date (UTC) lib source
 P cli           3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout    * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   data.table  * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   default     * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
   digest        0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   evaluate      0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   here          1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools     0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr         1.48    2024-07-07 [1] CRAN (R 4.4.1)
 P rlang       * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown     2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot     2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi    0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   sessioninfo   1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   xfun          0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml          2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
