---
title: "spe to Seurat to samplewise PRECAST"
author: "Bernie Mulvey"
date: "2023-06-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(Seurat)
library(PRECAST)
library(BiocParallel)
library(parallel)
library(parallelly)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load up everything we'll need--filtered SPE object, hvg/nnsvg lists
```{r}
spe <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")
# set coldata rownames / main colnames to be unique
rownames(colData(spe)) <- spe$key
colnames(spe) <- spe$key
stopifnot(sum(rownames(colData(spe))==colnames(spe))==ncol(spe))

hvg10 <- readRDS("analysis/data/H01-feature_selection/HVG10pctile_sample-blocked.RDS")
hvg20 <- readRDS("analysis/data/H01-feature_selection/HVG20pctile_sample-blocked.RDS")

# get the 10 and 20pctile sets of mean ranked nnsvgs
svgs <- fread("analysis/data/01-feature_selection/meangeneranks_nnSVG_samplewise_n3s0.5.txt")
setorderv(svgs,"meanrank")
svg10 <- svgs[c(1:length(hvg10)),gene_id]
svg20 <- svgs[c(1:length(hvg20)),gene_id]

featlist <- list(hvg10,hvg20,svg10,svg20)
names(featlist) <- c("hvg10","hvg20","svg10","svg20")
rm(hvg10,hvg20,svgs,svg10,svg20)
```


### Following Tony R's conversion script https://github.com/LieberInstitute/spatial_DG_lifespan/blob/main/code/EDA/spe_to_Seurat.R


### Test running here--just one feature set (otherwise we would need to iteratively or parallely make four seurat objects to separate out by sample, which would get ugly RAMwise)
```{r}

# omfg you could access coldata just by using the $ instead of colData(spe)$column. well that saves some time from now on
spe$row <- spe$array_row
spe$col <- spe$array_col

spe_srat <- CreateSeuratObject(counts=as.matrix(counts(spe)),meta.data=as.data.frame(colData(spe)),project="hype")

seulist <- list()
samples <- unique(spe$sample_id)

i <-1
for (i in c(1:length(samples))){
    seulist[[i]] <- subset(spe_srat,subset=sample_id==samples[i])
    names(seulist)[i] <- samples[i]
}

# saveRDS(seulist,file="analysis/data/H02-clustering/seurat_list_forprecast.RDS")
```

### And using PRECAST, with more of tony's code: https://github.com/LieberInstitute/spatial_DG_lifespan/blob/main/code/EDA/01_PRECAST.R
### as well as Erik's code for bypassing SPARK-X and passing in other genes
```{r}
# providing a custom gene list suppresses SPARK-X.
preobj = CreatePRECASTObject(seuList=seulist,project="hype",customGenelist = featlist[[1]])
preobj@seulist

PRECASTObj <- AddAdjList(preobj, platform = "Visium")
## Add a model setting in advance for a PRECASTObj object. verbose =TRUE helps outputing the information in the algorithm.

## MPI on a mac is like, impossible to set up with anything besides openMPI, and even that was failing when I had it installed so we won't parallelize for the test run.
PRECASTObj <- AddParSetting(PRECASTObj, Sigma_equal = FALSE, coreNum = 1, maxIter = 30, verbose = TRUE)


PRECASTObj <- PRECAST(PRECASTObj, K = 15)
saveRDS(PRECASTObj,paste0(
    "~/Desktop/",names(feats)[fks[i,1]],"_k",fks[i,2],"_precast_clusts.RDS")
    )
saveRDS(PRECASTObj,"~/Desktop/testsave.RDS")
```
