---
title: "01-Plot KNN domain first-run smoothings-VMH k10, ARC k50"
output: html_document
date: "2024-06-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(ggrastr)
library(ggtext) # for newlines in titles

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```


read in tables of VMH/ARC probabilities per cell determined by KNN smoothing 
```{r}
vmhknn1 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01a-step1smoothing_VMH_knn-k10.txt")

arcknn1 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01b-step1smoothing_ARC_knn-k50.txt")
```

## conveniently, we don't even need to read in the SFE to plot from these, since we have X, Y and KNN prediction outputs.
```{r}
threshes <- seq(0.0,1,by=0.1)

## parallelize using SOCK with snowparam, which requires us to manually specify contents to clone to each worker. this avoids colliding changes to the source data.tables by spawning totally siloed processes

bplapply(threshes,BPPARAM=SnowParam(11,type = "SOCK"),BPOPTIONS = bpoptions(exportglobals = T,packages = c("data.table","ggplot2","ggrastr","ggtext","gridExtra"),exports = c("arcknn1","vmhknn1")),FUN=function(t){
    ## make a copy of the data.table since the in-place modifications will collide across multicoreparam threads and error out
    vmhknn1 <- copy(vmhknn1)
    arcknn1 <- copy(arcknn1)
    
    vmhknn1[,newcol:=ifelse(vmhpred.smooth1.k10>=t,yes="VMH",no="other")]
    vmhthreshname <- paste0("VMH KNN pred min ",t)
    setnames(vmhknn1,"newcol",vmhthreshname)
    
    arcknn1[,newcol:=ifelse(arcpred.smooth1.k50>=t,yes="ARC",no="other")]
    arcthreshname <- paste0("ARC KNN pred min ",t)
    setnames(arcknn1,"newcol",arcthreshname)
    
    vmhpal <- c("#22dd00","#888888")
    names(vmhpal) <- c("VMH","other")
    
    arcpal <- c("#890090","#888888")
    names(arcpal) <- c("ARC","other")
    
    ## iterate over samples -> list of vmh plots per thresh -> gridExtra that
    vmhplots <- lapply(unique(vmhknn1$sample_id),function(s){
        xl <- min(vmhknn1[sample_id==s,xcoord])
        xh <- max(vmhknn1[sample_id==s,xcoord])
        yl <- min(vmhknn1[sample_id==s,ycoord])
        yh <- max(vmhknn1[sample_id==s,ycoord])
        plt <- ggplot(vmhknn1[sample_id==s], aes(x=xcoord, y=ycoord, color=.data[[vmhthreshname]])) +
            rasterize(geom_point(size=0.575)) + 
            scale_color_manual(values = vmhpal) +
            ggtitle(paste0(s," VMH\ndomain, k10, thresh ",t))+
            theme_minimal()+
            theme(title = element_markdown(size=20,hjust=0.5))+
            xlim(xl,xh)+
            ylim(yl,yh)
        return(plt)
    })
    png(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01a-VMH_k10_minscore_",t,".png"),width=4000,height=4000)
    do.call("grid.arrange",c(vmhplots,ncol=4))
    dev.off()
    
    
    ### plot ARC the same way
    arcplots <- lapply(unique(arcknn1$sample_id),function(s){
        xl <- min(arcknn1[sample_id==s,xcoord])
        xh <- max(arcknn1[sample_id==s,xcoord])
        yl <- min(arcknn1[sample_id==s,ycoord])
        yh <- max(arcknn1[sample_id==s,ycoord])
        plt <- ggplot(arcknn1[sample_id==s], aes(x=xcoord, y=ycoord, color=.data[[arcthreshname]])) +
            rasterize(geom_point(size=0.575)) + 
            scale_color_manual(values = arcpal) +
            ggtitle(paste0(s," ARC\ndomain, k50, thresh ",t))+
            theme_minimal()+
            theme(title = element_markdown(size=20,hjust=0.5))+
            xlim(xl,xh)+
            ylim(yl,yh)
        return(plt)
    })
    png(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01b-ARC_k50_minscore_",t,".png"),width=4000,height=4000)
    do.call("grid.arrange",c(arcplots,ncol=4))
    dev.off()
    
})
```


### make an additional plot each with the spots color coded by predictoin value, rather than thresholded.
```{r}
vmhplots <- lapply(unique(vmhknn1$sample_id),function(s){
    xl <- min(vmhknn1[sample_id==s,xcoord])
    xh <- max(vmhknn1[sample_id==s,xcoord])
    yl <- min(vmhknn1[sample_id==s,ycoord])
    yh <- max(vmhknn1[sample_id==s,ycoord])
    plt <- ggplot(vmhknn1[sample_id==s], aes(x=xcoord, y=ycoord, color=vmhpred.smooth1.k10)) +
        rasterize(geom_point(size=2.75)) + 
        scale_color_continuous(type = "viridis") +
        ggtitle(paste0(s," VMH\n smoothing scores, k10"))+
        theme_minimal()+
        theme(title = element_markdown(size=20,hjust=0.5))+
        xlim(xl,xh)+
        ylim(yl,yh)+
        labs(color="VMH KNN Score")
    return(plt)
})

png(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01a-VMH_k10_colored-by-KNNpredVal",".png"),width=4000,height=4000)
do.call("grid.arrange",c(vmhplots,ncol=4))
dev.off()


### aaand for ARC
arcplots <- lapply(unique(arcknn1$sample_id),function(s){
    xl <- min(arcknn1[sample_id==s,xcoord])
    xh <- max(arcknn1[sample_id==s,xcoord])
    yl <- min(arcknn1[sample_id==s,ycoord])
    yh <- max(arcknn1[sample_id==s,ycoord])
    plt <- ggplot(arcknn1[sample_id==s], aes(x=xcoord, y=ycoord, color=arcpred.smooth1.k50)) +
        rasterize(geom_point(size=2.75)) + 
        scale_color_continuous(type = "viridis") +
        ggtitle(paste0(s," ARC\n smoothing scores, k50"))+
        theme_minimal()+
        theme(title = element_markdown(size=20,hjust=0.5))+
        xlim(xl,xh)+
        ylim(yl,yh)+
        labs(color="ARC KNN Score")
    return(plt)
})

png(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/01b-ARC_k50_colored-by-KNNpredVal",".png"),width=4000,height=4000)
do.call("grid.arrange",c(arcplots,ncol=4))
dev.off()
```


## reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0 parallelly_1.37.1   colorout_1.3-0.2   
 [4] ggtext_0.1.2        ggrastr_1.0.2       gridExtra_2.3      
 [7] ggplot2_3.5.1       Biostrings_2.72.1   GenomeInfoDb_1.40.1
[10] XVector_0.44.0      IRanges_2.38.0      S4Vectors_0.42.0   
[13] BiocGenerics_0.50.0 data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] utf8_1.2.4              generics_0.1.3          xml2_1.3.6             
 [4] magrittr_2.0.3          grid_4.4.1              rprojroot_2.0.4        
 [7] jsonlite_1.8.8          httr_1.4.7              fansi_1.0.6            
[10] UCSC.utils_1.0.0        scales_1.3.0            codetools_0.2-20       
[13] cli_3.6.3               crayon_1.5.3            munsell_0.5.1          
[16] withr_3.0.0             ggbeeswarm_0.7.2        tools_4.4.1            
[19] parallel_4.4.1          dplyr_1.1.4             colorspace_2.1-0       
[22] GenomeInfoDbData_1.2.12 here_1.0.1              vctrs_0.6.5            
[25] R6_2.5.1                lifecycle_1.0.4         zlibbioc_1.50.0        
[28] vipor_0.4.7             pkgconfig_2.0.3         beeswarm_0.4.0         
[31] pillar_1.9.0            gtable_0.3.5            glue_1.7.0             
[34] Rcpp_1.0.12             xfun_0.45               tibble_3.2.1           
[37] tidyselect_1.2.1        rstudioapi_0.16.0       knitr_1.47             
[40] compiler_4.4.1          gridtext_0.1.5         
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-28
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   NA

─ Packages ─────────────────────────────────────────────────────────────────────
 ! package          * version date (UTC) lib source
   beeswarm           0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   BiocGenerics     * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel     * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings       * 2.72.1  2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools          0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout         * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace         2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon             1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table       * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   dplyr              1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   fansi              1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   generics           0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb     * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData   1.2.12  2024-05-01 [1] Bioconductor
   ggbeeswarm         0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2          * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr          * 1.0.2   2023-06-01 [1] CRAN (R 4.4.0)
   ggtext           * 0.1.2   2022-09-16 [1] CRAN (R 4.4.0)
   glue               1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra        * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gridtext           0.1.5   2022-09-16 [1] CRAN (R 4.4.0)
   gtable             0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here               1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   httr               1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges          * 2.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   jsonlite           1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr              1.47    2024-05-29 [1] CRAN (R 4.4.0)
   lifecycle          1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr           2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell            0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly       * 1.37.1  2024-02-29 [1] CRAN (R 4.4.0)
   pillar             1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig          2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6                 2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp               1.0.12  2024-01-09 [1] CRAN (R 4.4.0)
 P rlang            * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rprojroot          2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi         0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors        * 0.42.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales             1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo        1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble             3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect         1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils         1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8               1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs              0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor              0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   withr              3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun               0.45    2024-06-16 [1] CRAN (R 4.4.0)
   xml2               1.3.6   2023-12-04 [1] CRAN (R 4.4.0)
   XVector          * 0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   zlibbioc           1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

────────────────────────────────────────────────────────────────────────────────
