---
title: "02b_02c-Compare marker DE at boundaries of VMH and ARC defined at various
  knn thresholds"
output: html_document
date: "2024-07-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### read in results from yi's parameters and from tested parameter sets
```{r}
yiparamde.bdry <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b-VMHARC_domains_yi-threshes_boundaryarea-VMH-ARC-DE.txt")
yiparamde.full <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02a-VMHARC_domains_yi-thresholds_pseudobulkDE.txt")

paramtestde <- readRDS("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02c-fullDom_4pctMediDiamBdryDom_DE_by_threshpairs.RDS")
```

### concatenate into giant table for plotting
```{r}
## for the parameter tests, drop results where there 
### a) were no boundary regions detectable after expanding the size of VMH/ARC domains (we know they're right against one another from RNAscope): list entries of size 1 
### b) inadequate cells in VMH, ARC, or virtually all cells were VMH or ARC, so no DE OR boundary testing was done (list entries of size 1)
### c) mystery total-crash errors (list entries of length 0)

paramtestde <- lapply(paramtestde,FUN=function(x){
    x <- x[!sapply(x,function(y) length(y)==0|length(y)==1)]
    x <- x[!sapply(x,function(y) length(grep(y[[2]],pattern="under 150 boundary cells for VMH or ARC found for",value=T)==1))]
})

### concat all the bulk DE results into a single one-level list
paramtestde <- paramtestde[!sapply(paramtestde,function(x) length(x)==0)]

bulkdelist <- list()
i<-1
for (i in c(1:length(paramtestde))){
    j<-1
    for (j in c(1:length(paramtestde[[i]]))){
        bulkdelist <- c(bulkdelist,paramtestde[[i]][[j]][[1]][1])
        names(bulkdelist)[length(bulkdelist)] <- names(paramtestde[[i]][[j]][[1]])[1]
    }
}
rm(i,j)

### concat all the boundary DE results into a single one-level list
bdrydelist <- list()
i<-1
for (i in c(1:length(paramtestde))){
    j<-1
    for (j in c(1:length(paramtestde[[i]]))){
        bdrydelist <- c(bdrydelist,paramtestde[[i]][[j]][[2]][1])
        names(bdrydelist)[length(bdrydelist)] <- names(paramtestde[[i]][[j]][[2]])[1]
    }
}
rm(i,j)
### append cutoffs info to both lists
bulkdelist <- mapply(X=bulkdelist,Y=names(bulkdelist),SIMPLIFY = FALSE,FUN=function(X,Y){
    X[,thresh1:=as.numeric(gsub(Y,pattern="^.*_thresh1_(.*)_thr2ARC.*$",replacement="\\1"))]
    X[,thresh2ARC:=as.numeric(gsub(Y,pattern="^.*_thr2ARC_(.*)_thr2.*$",replacement="\\1"))]
    X[,thresh2VMH:=as.numeric(gsub(Y,pattern="^.*_thr2VMH_(.*)$",replacement="\\1"))]
})

bdrydelist <- mapply(X=bdrydelist,Y=names(bdrydelist),SIMPLIFY = FALSE,FUN=function(X,Y){
    X[,thresh1:=as.numeric(gsub(Y,pattern="^.*_thr1_(.*)_thr2ARC.*$",replacement="\\1"))]
    X[,thresh2ARC:=as.numeric(gsub(Y,pattern="^.*_thr2ARC_(.*)_thr2.*$",replacement="\\1"))]
    X[,thresh2VMH:=as.numeric(gsub(Y,pattern="^.*_thr2VMH_(.*)$",replacement="\\1"))]
})

### collapse to gigundo data tables
bulkdelist <- rbindlist(bulkdelist)
bdrydelist <- rbindlist(bdrydelist)

### bulkdelist has two gene columns, drop the 2nd one (col 7)
bulkdelist <- bulkdelist[,c(1:6,8:10)]
```

### plot the results for each of a few genes per thresh1 cutoff
```{r}
t1s <- unique(bulkdelist$thresh1)

lapply(t1s,FUN=function(x){
    tmpbulk <- bulkdelist[thresh1==x]
    
    mks <- c("POMC","AGRP","ESR1","GAD1","GAD2","GHRH","TAC3","SLC17A6","FEZF1","NR5A1","LAMP5","ANKRD34B","ADCYAP1","NRGN")
    tmpbulk <- tmpbulk[gene%in%mks]
    tmpbulk[,gene:=factor(gene,levels=mks)]
    
    tmpbulka <- copy(tmpbulk)
    tmpbulka[,thresh2VMH:=as.character(thresh2VMH)]

    arcplt <- lapply(mks,FUN=function(g){
        ggplot(tmpbulka[gene==g],aes(x=thresh2ARC,y=`logFC_ARC-VMH`,fill=thresh2VMH))+
            geom_col(position="dodge")+
            xlab("ARC knn round 2 threshold")+
            ylab("Full domain ARC vs VMH logFC")+
            geom_hline(yintercept=yiparamde.full[gene==g,`logFC_ARC_yi_2stage-VMH_yi_2stage`],linetype="dashed")+
            ggtitle(g)+
            theme(axis.text.x = element_text(size=9),axis.text.y=element_text(size=9),axis.title.x = element_text(size=9),axis.title.y=element_text(size=9),strip.text = element_text(size=9),legend.text = element_text(size=8),plot.title = element_text(size=11,hjust=0.5))
    })
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b_02c-ARCVMH_smoothing_refinement_by_DE/02b_02c-ARCthresh2-wise_bulkDE_thresh1_",x,".pdf"),width=18,height=28)
    do.call("grid.arrange",c(arcplt,ncol=2))
    dev.off()
    
    
    tmpbulkv <- copy(tmpbulk)
    tmpbulkv[,thresh2ARC:=as.character(thresh2ARC)]
    
    vmhplt <- lapply(mks,FUN=function(g){
        ggplot(tmpbulkv[gene==g],aes(x=thresh2VMH,y=`logFC_ARC-VMH`,fill=thresh2ARC))+
            geom_col(position="dodge")+
            xlab("VMH knn round 2 threshold")+
            ylab("Full domain ARC vs VMH logFC")+
            geom_hline(yintercept=yiparamde.full[gene==g,`logFC_ARC_yi_2stage-VMH_yi_2stage`],linetype="dashed")+
            ggtitle(g)+
            theme(axis.text.x = element_text(size=9),axis.text.y=element_text(size=9),axis.title.x = element_text(size=9),axis.title.y=element_text(size=9),strip.text = element_text(size=9),legend.text = element_text(size=8),plot.title = element_text(size=11,hjust=0.5))
    })
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b_02c-ARCVMH_smoothing_refinement_by_DE/02b_02c-VMHthresh2-wise_bulkDE_thresh1_",x,".pdf"),width=18,height=28)
    do.call("grid.arrange",c(vmhplt,ncol=2))
    dev.off()
})

```

### same for boundary DE
```{r}
t1s <- unique(bdrydelist$thresh1)

lapply(t1s,FUN=function(x){
    tmpbdry <- bdrydelist[thresh1==x]
    
    mks <- c("POMC","AGRP","ESR1","GAD1","GAD2","GHRH","TAC3","SLC17A6","FEZF1","NR5A1","LAMP5","ANKRD34B","ADCYAP1","NRGN")
    tmpbdry <- tmpbdry[gene%in%mks]
    tmpbdry[,gene:=factor(gene,levels=mks)]

    tmpbdrya <- copy(tmpbdry)
    tmpbdrya[,thresh2VMH:=as.character(thresh2VMH)]
    
    arcplt <- lapply(mks,FUN=function(g){
        ggplot(tmpbdrya[gene==g],aes(x=thresh2ARC,y=`logFC_ARCboundary-VMHboundary`,fill=thresh2VMH))+
            geom_col(position="dodge")+
            xlab("ARC knn round 2 threshold")+
            ylab("Boundary region ARC vs VMH logFC")+
            geom_hline(yintercept=yiparamde.bdry[gene==g,`logFC_ARCboundary-VMHboundary`],linetype="dashed")+
            ggtitle(g)+
            theme(axis.text.x = element_text(size=9),axis.text.y=element_text(size=9),axis.title.x = element_text(size=9),axis.title.y=element_text(size=9),strip.text = element_text(size=9),legend.text = element_text(size=8),plot.title = element_text(size=11,hjust=0.5))
    })
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b_02c-ARCVMH_smoothing_refinement_by_DE/02b_02c-ARCthresh2-wise_boundaryDE_thresh1_",x,".pdf"),width=18,height=28)
    do.call("grid.arrange",c(arcplt,ncol=2))
    dev.off()
    
    tmpbdryv <- copy(tmpbdry)
    tmpbdryv[,thresh2ARC:=as.character(thresh2ARC)]
    
    vmhplt <- lapply(mks,FUN=function(g){
        ggplot(tmpbdryv[gene==g],aes(x=thresh2VMH,y=`logFC_ARCboundary-VMHboundary`,fill=thresh2ARC))+
            geom_col(position="dodge")+
            xlab("VMH knn round 2 threshold")+
            ylab("Boundary ARC vs VMH logFC")+
            geom_hline(yintercept=yiparamde.bdry[gene==g,`logFC_ARCboundary-VMHboundary`],linetype="dashed")+
            ggtitle(g)+
            theme(axis.text.x = element_text(size=9),axis.text.y=element_text(size=9),axis.title.x = element_text(size=9),axis.title.y=element_text(size=9),strip.text = element_text(size=9),legend.text = element_text(size=8),plot.title = element_text(size=11,hjust=0.5))
    })
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02b_02c-ARCVMH_smoothing_refinement_by_DE/02b_02c-VMHthresh2-wise_boundaryDE_thresh1_",x,".pdf"),width=18,height=28)
    do.call("grid.arrange",c(vmhplt,ncol=2))
    dev.off()
})

```

### reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0 parallelly_1.37.1   colorout_1.3-0.2   
 [4] gridExtra_2.3       ggplot2_3.5.1       Biostrings_2.72.1  
 [7] GenomeInfoDb_1.40.1 XVector_0.44.0      IRanges_2.38.0     
[10] S4Vectors_0.42.0    BiocGenerics_0.50.0 data.table_1.15.4  
[13] rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] utf8_1.2.4              generics_0.1.3          digest_0.6.36          
 [4] magrittr_2.0.3          evaluate_0.24.0         grid_4.4.1             
 [7] fastmap_1.2.0           rprojroot_2.0.4         jsonlite_1.8.8         
[10] sessioninfo_1.2.2       httr_1.4.7              fansi_1.0.6            
[13] UCSC.utils_1.0.0        scales_1.3.0            codetools_0.2-20       
[16] cli_3.6.3               crayon_1.5.3            munsell_0.5.1          
[19] withr_3.0.0             yaml_2.3.8              tools_4.4.1            
[22] parallel_4.4.1          dplyr_1.1.4             colorspace_2.1-0       
[25] GenomeInfoDbData_1.2.12 here_1.0.1              vctrs_0.6.5            
[28] R6_2.5.1                lifecycle_1.0.4         zlibbioc_1.50.0        
[31] pkgconfig_2.0.3         pillar_1.9.0            gtable_0.3.5           
[34] glue_1.7.0              xfun_0.45               tibble_3.2.1           
[37] tidyselect_1.2.1        rstudioapi_0.16.0       knitr_1.47             
[40] farver_2.1.2            htmltools_0.5.8.1       rmarkdown_2.27         
[43] labeling_0.4.3          compiler_4.4.1         
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-01
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package          * version date (UTC) lib source
   BiocGenerics     * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel     * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings       * 2.72.1  2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools          0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout         * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace         2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon             1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table       * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest             0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr              1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate           0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi              1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver             2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap            1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics           0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb     * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData   1.2.12  2024-05-01 [1] Bioconductor
   ggplot2          * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue               1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra        * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable             0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here               1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools          0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr               1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges          * 2.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   jsonlite           1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr              1.47    2024-05-29 [1] CRAN (R 4.4.0)
   labeling           0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle          1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr           2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell            0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly       * 1.37.1  2024-02-29 [1] CRAN (R 4.4.0)
   pillar             1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig          2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6                 2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
 P rlang            * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown          2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot          2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi         0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Vectors        * 0.42.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales             1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo        1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble             3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect         1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils         1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8               1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs              0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr              3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun               0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector          * 0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml               2.3.8   2023-12-11 [1] CRAN (R 4.4.0)
   zlibbioc           1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
