---
title: "03a-Plot ARC4type-VMH"
output: html_document
date: "2024-07-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(ggrastr)
library(ggtext) # for newlines in titles

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```


read in tables of VMH/ARC probabilities per cell determined by 2 step KNN smoothing with the three core ARC cell types considered alone
```{r}
arcvmh4 <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03a-Alt_2stepKNN-ARCVMH_ARC-no-X23.txt")
```

## conveniently, we don't even need to read in the SFE to plot from these, since we have X, Y and KNN prediction outputs.
```{r}
threshes <- seq(0.0,1,by=0.1)
### make combinations of thresholds
threshtab <- expand.grid(threshes,threshes)
threshtab <- as.data.table(threshtab)
setnames(threshtab,c("ARCthresh","VMHthresh"))
## parallelize using SOCK with snowparam, which requires us to manually specify contents to clone to each worker. this avoids colliding changes to the source data.tables by spawning totally siloed processes

bplapply(1:nrow(threshtab),BPPARAM=SnowParam(11,type = "SOCK"),BPOPTIONS = bpoptions(exportglobals = T,packages = c("data.table","ggplot2","ggrastr","ggtext","gridExtra"),exports = c("threshtab","arcvmh4")),FUN=function(t){
    ## make a copy of the data.table since the in-place modifications will collide across multicoreparam threads and error out
    arcvmh4 <- copy(arcvmh4)
    thra <- threshtab[t,ARCthresh]
    thrv <- threshtab[t,VMHthresh]
    
    ## assign based on current thresholds
    arcvmh4[,newcol:=ifelse(VMHpred_thresh1_0.2_smooth2.k200>thrv,yes="VMH",no="other")]
    arcvmh4[,newcol2:=ifelse(ARCpred_thresh1_0.1_smooth2.k500>thra,yes="ARC",no="other")]
    ## resolve dual labels
    arcvmh4[,newcol3:="other"]
    arcvmh4[newcol=="VMH",newcol3:="VMH"]
    arcvmh4[newcol2=="ARC",newcol3:="ARC"]
    arcvmh4[newcol=="VMH"&newcol2=="ARC",newcol3:=ifelse(VMHpred_thresh1_0.2_smooth2.k200>ARCpred_thresh1_0.1_smooth2.k500,yes="VMH",no="ARC")]
    ##
    
    dualassignname <- paste0("VMH_",thrv,"_ARC3type_",thra)
    setnames(arcvmh4,"newcol3",dualassignname)
    
    dompal <- c("#890090","#22dd00","#888888")
    
    names(dompal) <- c("ARC","VMH","other")
    
    ## iterate over samples -> list of vmh plots per thresh -> gridExtra that
    plts <- lapply(unique(arcvmh4$sample_id),function(s){
        xl <- min(arcvmh4[sample_id==s,xcoord])
        xh <- max(arcvmh4[sample_id==s,xcoord])
        yl <- min(arcvmh4[sample_id==s,ycoord])
        yh <- max(arcvmh4[sample_id==s,ycoord])
        plt <- ggplot(arcvmh4[sample_id==s], aes(x=xcoord, y=ycoord, color=.data[[dualassignname]])) +
            rasterize(geom_point(size=0.575)) + 
            scale_color_manual(values = dompal) +
            ggtitle(paste0(s," ",dualassignname))+
            theme_minimal()+
            theme(title = element_markdown(size=20,hjust=0.5))+
            xlim(xl,xh)+
            ylim(yl,yh)
        return(plt)
    })
    png(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03a-ARC-VMH_domains_4arcCelltypes_by_thresh/03a-VMH_k200_grthn_",thrv,"_ARC3type_k500_grthn_",thra,".png"),width=4000,height=4000)
    do.call("grid.arrange",c(plts,ncol=4))
    dev.off()
    
    
})
```


## reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] ggtext_0.1.2                   ggrastr_1.0.2                 
 [3] gridExtra_2.3                  ggplot2_3.5.1                 
 [5] BiocParallel_1.38.0            parallelly_1.37.1             
 [7] colorout_1.3-0.2               SpatialFeatureExperiment_1.6.1
 [9] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[11] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[13] GenomicRanges_1.56.1           MatrixGenerics_1.16.0         
[15] matrixStats_1.3.0              FNN_1.1.4                     
[17] caTools_1.18.2                 FastKNN_0.0.1                 
[19] Biostrings_2.72.1              GenomeInfoDb_1.40.1           
[21] XVector_0.44.0                 IRanges_2.38.0                
[23] S4Vectors_0.42.0               BiocGenerics_0.50.0           
[25] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] DBI_1.2.3                 bitops_1.0-7             
  [3] deldir_2.0-4              s2_1.1.6                 
  [5] magrittr_2.0.3            e1071_1.7-14             
  [7] compiler_4.4.1            DelayedMatrixStats_1.26.0
  [9] vctrs_0.6.5               png_0.1-8                
 [11] sfheaders_0.4.4           fftwtools_0.9-11         
 [13] pkgconfig_2.0.3           wk_0.9.1                 
 [15] crayon_1.5.3              fastmap_1.2.0            
 [17] magick_2.8.3              scuttle_1.14.0           
 [19] utf8_1.2.4                rmarkdown_2.27           
 [21] ggbeeswarm_0.7.2          UCSC.utils_1.0.0         
 [23] xfun_0.45                 zlibbioc_1.50.0          
 [25] beachmat_2.20.0           jsonlite_1.8.8           
 [27] rhdf5filters_1.16.0       DelayedArray_0.30.1      
 [29] Rhdf5lib_1.26.0           jpeg_0.1-10              
 [31] pdist_1.2.1               tiff_0.1-12              
 [33] terra_1.7-78              parallel_4.4.1           
 [35] R6_2.5.1                  limma_3.60.3             
 [37] boot_1.3-30               Rcpp_1.0.12              
 [39] assertthat_0.2.1          knitr_1.47               
 [41] snow_0.4-4                R.utils_2.12.3           
 [43] tidyselect_1.2.1          Matrix_1.7-0             
 [45] rstudioapi_0.16.0         abind_1.4-5              
 [47] yaml_2.3.8                EBImage_4.46.0           
 [49] codetools_0.2-20          tibble_3.2.1             
 [51] lattice_0.22-6            withr_3.0.0              
 [53] evaluate_0.24.0           sf_1.0-16                
 [55] units_0.8-5               spData_2.3.1             
 [57] proxy_0.4-27              xml2_1.3.6               
 [59] pillar_1.9.0              KernSmooth_2.23-24       
 [61] generics_0.1.3            rprojroot_2.0.4          
 [63] sp_2.1-4                  RCurl_1.98-1.14          
 [65] sparseMatrixStats_1.16.0  munsell_0.5.1            
 [67] scales_1.3.0              class_7.3-22             
 [69] glue_1.7.0                tools_4.4.1              
 [71] BiocNeighbors_1.22.0      locfit_1.5-9.10          
 [73] rhdf5_2.48.0              grid_4.4.1               
 [75] spdep_1.3-5               DropletUtils_1.24.0      
 [77] edgeR_4.2.0               colorspace_2.1-0         
 [79] GenomeInfoDbData_1.2.12   beeswarm_0.4.0           
 [81] HDF5Array_1.32.0          vipor_0.4.7              
 [83] cli_3.6.3                 fansi_1.0.6              
 [85] S4Arrays_1.4.1            dplyr_1.1.4              
 [87] gtable_0.3.5              R.methodsS3_1.8.2        
 [89] zeallot_0.1.0             digest_0.6.36            
 [91] classInt_0.4-10           SparseArray_1.4.8        
 [93] dqrng_0.4.1               rjson_0.2.21             
 [95] htmlwidgets_1.6.4         htmltools_0.5.8.1        
 [97] R.oo_1.26.0               lifecycle_1.0.4          
 [99] httr_1.4.7                here_1.0.1               
[101] statmod_1.5.0             gridtext_0.1.5           
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-05
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   assertthat                 0.2.1     2019-03-21 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0     2021-06-01 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings               * 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   caTools                  * 1.18.2    2021-03-28 [1] CRAN (R 4.4.0)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   FastKNN                  * 0.0.1     2015-02-12 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   FNN                      * 1.1.4     2024-01-12 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2     2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2     2023-06-01 [1] CRAN (R 4.4.0)
   ggtext                   * 0.1.2     2022-09-16 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gridtext                   0.1.5     2022-09-16 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47      2024-05-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pdist                      1.2.1     2022-05-02 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   snow                       0.4-4     2021-10-27 [1] CRAN (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7     2023-12-18 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   xml2                       1.3.6     2023-12-04 [1] CRAN (R 4.4.0)
   XVector                  * 0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8     2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
