---
title: "05b-Plot_proportions_of_celltypes_neighboring_VMHARCcluster_cells"
output: html_document
date: "2024-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 12,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank()))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load the RDS with number, proportion of neighboring cells falling into each given cluster for VMH clusters or ARC clusters (within respective domain boundaries only)
```{r}
vmhcelldists <- readRDS("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-VMHdomain_VMHcells_NandProp_pertype_within_100-500u_and_50u_radii.RDS")

arccelldists <- readRDS("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-ARCdomain_ARCcells_NandProp_pertype_within_100-500u_and_50u_radii.RDS")


### add 'cell group'-level info (e.g., all oligo clusters, all microglia clusters, etc)
bkcl <- unique(unlist(
    lapply(c(arccelldists,vmhcelldists),FUN=function(d){
        unlist(lapply(d,function(x){
                unique(x$label2)
            }))
        }))
)

bkcl <- as.data.table(bkcl)
setnames(bkcl,1,"bjm_annot")

bkcl[bjm_annot %in% grep(bjm_annot,pattern="VMH",value=T),supercl:="VMH (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="ARC",value=T),supercl:="ARC (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Tanyc|Alar",value=T),supercl:="Tanycytes, Portal Vasc. (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Oligo",value=T),supercl:="Oligo (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Astro",value=T),supercl:="Astro (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Microg",value=T),supercl:="Microglia (3)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Supraopt",value=T),supercl:="SON (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Unsure_AVP",value=T),supercl:="Non-SON AVP+OXT+ (1)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Macrop|Periph|Vascul|Endothel",value=T),supercl:="Vascular and Peripheral Immune (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="GABA",value=T),supercl:="Other GABAergic Neurons (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Periven",value=T),supercl:="PeriVN (Excitatory) (1)"]

## merge this info into the arccell and vmhcell data and calculate supercl N and props
arccelldists2 <- lapply(arccelldists,FUN=function(x){
    lapply(x,FUN=function(y){
        z <- merge.data.table(y,bkcl,by.x="label2",by.y="bjm_annot",all.x=T)
        z[,N_supercl_neighboring_cell1:=sum(N_label2_neighboring_cell1),by=c("cell1","supercl")]
        z[,prop_supercl_neighboring_cell1:=N_supercl_neighboring_cell1/total_cell1_neighbors,by="cell1"]
        return(z)
    })
})

vmhcelldists2 <- lapply(vmhcelldists,FUN=function(x){
    lapply(x,FUN=function(y){
        z <- merge.data.table(y,bkcl,by.x="label2",by.y="bjm_annot",all.x=T)
        z[,N_supercl_neighboring_cell1:=sum(N_label2_neighboring_cell1),by=c("cell1","supercl")]
        z[,prop_supercl_neighboring_cell1:=N_supercl_neighboring_cell1/total_cell1_neighbors,by="cell1"]
        return(z)
    })
})
```

before plotting, rearrange data into one big table per radius tested
```{r}
rads <- names(arccelldists2[[1]])

arccelldists3 <- list()
i<-1
for (i in c(1:length(rads))){
    arccelldists3[[i]] <- rbindlist(lapply(arccelldists2,FUN=function(x){x[[i]]}))
}
names(arccelldists3) <- rads

vmhcelldists3 <- list()
i<-1
for (i in c(1:length(rads))){
    vmhcelldists3[[i]] <- rbindlist(lapply(vmhcelldists2,FUN=function(x){x[[i]]}))
}
names(vmhcelldists3) <- rads
```

now make violin plots faceted on cell1 label with supercl on the x axis and proportion of neighbors on the y axis. one plot per radius per VMH/ARC.
```{r}
lapply(c(1:length(rads)),FUN=function(i){
    plts <- list()
    
    plts[[1]] <- ggplot(arccelldists3[[i]],aes(x=supercl,y=prop_supercl_neighboring_cell1))+
        geom_violin(aes(fill=supercl),scale="width",trim=FALSE)+
        geom_boxplot(width=0.1)+
        theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 12,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank())+
        facet_wrap(~label1,scales="free",nrow=5)+
        ggtitle(paste("ARC clusters, ",rads[i]," unit radius"))
    
    plts[[2]] <- ggplot(vmhcelldists3[[i]],aes(x=supercl,y=prop_supercl_neighboring_cell1))+
        geom_violin(aes(fill=supercl),scale="width",trim=FALSE)+
        geom_boxplot(width=0.1)+
        facet_wrap(~label1,scales="free",nrow = 4)+
        ggtitle(paste("VMH cells, ",rads[i]," radius"))
        theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 12,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank())+
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-Celltype_proportions_in_various_radii_from_eachARCVMH_clustertype/supercluster_props_in_",
               gsub(names(vmhcelldists3)[i],pattern="d(.*)",replacement="\\1"),
               "unit_radius.pdf"),
        height=40,width=15
    )
    do.call("grid.arrange",c(plts,nrow=2))
    dev.off()
})

```


```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] sf_1.0-16                      SpatialFeatureExperiment_1.6.1
 [3] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [5] SummarizedExperiment_1.34.0    Biobase_2.64.0                
 [7] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
 [9] IRanges_2.38.1                 S4Vectors_0.42.1              
[11] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[13] matrixStats_1.3.0              BiocParallel_1.38.0           
[15] parallelly_1.38.0              colorout_1.3-0.2              
[17] gridExtra_2.3                  ggplot2_3.5.1                 
[19] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
 [1] DBI_1.2.3                 bitops_1.0-7              deldir_2.0-4             
 [4] s2_1.1.6                  magrittr_2.0.3            e1071_1.7-14             
 [7] compiler_4.4.1            DelayedMatrixStats_1.26.0 sfheaders_0.4.4          
[10] png_0.1-8                 fftwtools_0.9-11          vctrs_0.6.5              
[13] wk_0.9.2                  pkgconfig_2.0.3           crayon_1.5.3             
[16] fastmap_1.2.0             magick_2.8.3              XVector_0.44.0           
[19] labeling_0.4.3            scuttle_1.14.0            utf8_1.2.4               
[22] rmarkdown_2.27            sessioninfo_1.2.2         UCSC.utils_1.0.0         
[25] xfun_0.45                 zlibbioc_1.50.0           beachmat_2.20.0          
[28] jsonlite_1.8.8            rhdf5filters_1.16.0       DelayedArray_0.30.1      
[31] Rhdf5lib_1.26.0           terra_1.7-78              jpeg_0.1-10              
[34] tiff_0.1-12               R6_2.5.1                  limma_3.60.3             
[37] boot_1.3-30               Rcpp_1.0.13               knitr_1.48               
[40] R.utils_2.12.3            Matrix_1.7-0              tidyselect_1.2.1         
[43] rstudioapi_0.16.0         abind_1.4-5               yaml_2.3.9               
[46] EBImage_4.46.0            codetools_0.2-20          lattice_0.22-6           
[49] tibble_3.2.1              withr_3.0.0               evaluate_0.24.0          
[52] spData_2.3.1              units_0.8-5               proxy_0.4-27             
[55] pillar_1.9.0              KernSmooth_2.23-24        generics_0.1.3           
[58] sp_2.1-4                  rprojroot_2.0.4           RCurl_1.98-1.14          
[61] sparseMatrixStats_1.16.0  munsell_0.5.1             scales_1.3.0             
[64] class_7.3-22              glue_1.7.0                tools_4.4.1              
[67] BiocNeighbors_1.22.0      locfit_1.5-9.10           rhdf5_2.48.0             
[70] grid_4.4.1                spdep_1.3-5               DropletUtils_1.24.0      
[73] edgeR_4.2.0               colorspace_2.1-0          GenomeInfoDbData_1.2.12  
[76] HDF5Array_1.32.0          cli_3.6.3                 fansi_1.0.6              
[79] S4Arrays_1.4.1            dplyr_1.1.4               gtable_0.3.5             
[82] R.methodsS3_1.8.2         zeallot_0.1.0             digest_0.6.36            
[85] classInt_0.4-10           SparseArray_1.4.8         dqrng_0.4.1              
[88] farver_2.1.2              rjson_0.2.21              htmlwidgets_1.6.4        
[91] htmltools_0.5.8.1         R.oo_1.26.0               lifecycle_1.0.4          
[94] httr_1.4.7                here_1.0.1                statmod_1.5.0            
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-25
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3     2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36    2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2     2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggplot2                  * 3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3       2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48      2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3     2023-08-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.10  2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0    2024-07-27 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13    2024-07-17 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1    2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2     2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XVector                    0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9     2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
