---
title: "05b-Plot_proportions_of_celltypes_neighboring_VMHARCcluster_cells"
output: html_document
date: "2024-09-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
library(parallel)

## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 12,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank()))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

load the RDS with number, proportion of neighboring cells falling into each given cluster for VMH clusters or ARC clusters (within respective domain boundaries only)
```{r}
vmhcelldists <- readRDS("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-VMHdomain_VMHcells_NandProp_pertype_within_100-500u_and_50u_radii.RDS")

arccelldists <- readRDS("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-ARCdomain_ARCcells_NandProp_pertype_within_100-500u_and_50u_radii.RDS")


### add 'cell group'-level info (e.g., all oligo clusters, all microglia clusters, etc)
bkcl <- unique(unlist(
    lapply(c(arccelldists,vmhcelldists),FUN=function(d){
        unlist(lapply(d,function(x){
                unique(x$label2)
            }))
        }))
)

bkcl <- as.data.table(bkcl)
setnames(bkcl,1,"bjm_annot")

bkcl[bjm_annot %in% grep(bjm_annot,pattern="VMH",value=T),supercl:="VMH (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="ARC",value=T),supercl:="ARC (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Tanyc|Alar",value=T),supercl:="Tanycytes, Portal Vasc. (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Oligo",value=T),supercl:="Oligo (4)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Astro",value=T),supercl:="Astro (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Microg",value=T),supercl:="Microglia (3)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Supraopt",value=T),supercl:="SON (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Unsure_AVP",value=T),supercl:="Non-SON AVP+OXT+ (1)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Macrop|Periph|Vascul|Endothel",value=T),supercl:="Vascular and Peripheral Immune (5)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="GABA",value=T),supercl:="Other GABAergic Neurons (2)"]
bkcl[bjm_annot %in% grep(bjm_annot,pattern="Periven",value=T),supercl:="PeriVN (Excitatory) (1)"]

## merge this info into the arccell and vmhcell data and calculate supercl N and props
arccelldists2 <- lapply(arccelldists,FUN=function(x){
    lapply(x,FUN=function(y){
        z <- merge.data.table(y,bkcl,by.x="label2",by.y="bjm_annot",all.x=T)
        z[,N_supercl_neighboring_cell1:=sum(N_label2_neighboring_cell1),by=c("cell1","supercl")]
        z[,prop_supercl_neighboring_cell1:=N_supercl_neighboring_cell1/total_cell1_neighbors,by="cell1"]
        return(z)
    })
})

vmhcelldists2 <- lapply(vmhcelldists,FUN=function(x){
    lapply(x,FUN=function(y){
        z <- merge.data.table(y,bkcl,by.x="label2",by.y="bjm_annot",all.x=T)
        z[,N_supercl_neighboring_cell1:=sum(N_label2_neighboring_cell1),by=c("cell1","supercl")]
        z[,prop_supercl_neighboring_cell1:=N_supercl_neighboring_cell1/total_cell1_neighbors,by="cell1"]
        return(z)
    })
})
```

before plotting, rearrange data into one big table per radius tested
```{r}
rads <- names(arccelldists2[[1]])

arccelldists3 <- list()
i<-1
for (i in c(1:length(rads))){
    arccelldists3[[i]] <- rbindlist(lapply(arccelldists2,FUN=function(x){x[[i]]}))
}
names(arccelldists3) <- rads

vmhcelldists3 <- list()
i<-1
for (i in c(1:length(rads))){
    vmhcelldists3[[i]] <- rbindlist(lapply(vmhcelldists2,FUN=function(x){x[[i]]}))
}
names(vmhcelldists3) <- rads
```

now make violin plots faceted on cell1 label with supercl on the x axis and proportion of neighbors on the y axis. one plot per radius per VMH/ARC.
```{r}
lapply(c(1:length(rads)),FUN=function(i){
    plts <- list()
    
    plts[[1]] <- ggplot(arccelldists3[[i]],aes(x=supercl,y=prop_supercl_neighboring_cell1))+
        geom_violin(aes(fill=supercl),scale="width",trim=FALSE)+
        geom_boxplot(width=0.1)+
        facet_wrap(~label1,scales="free",nrow=5)+
        ggtitle(paste("ARC clusters, ",rads[i]," unit radius\nMean+-SD ",round(mean(arccelldists3[[i]]$total_cell1_neighbors))," +- ",round(sd(arccelldists3[[i]]$total_cell1_neighbors))," total neighbors"))+
        theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 18,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank())

    
    plts[[2]] <- ggplot(vmhcelldists3[[i]],aes(x=supercl,y=prop_supercl_neighboring_cell1))+
        geom_violin(aes(fill=supercl),scale="width",trim=FALSE)+
        geom_boxplot(width=0.1)+
        facet_wrap(~label1,scales="free",nrow = 4)+
        ggtitle(paste("VMH clusters, ",rads[i]," unit radius\nMean+-SD ",round(mean(vmhcelldists3[[i]]$total_cell1_neighbors))," +- ",round(sd(vmhcelldists3[[i]]$total_cell1_neighbors))," total neighbors"))+        theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 18,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank())+
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-Celltype_proportions_in_various_radii_from_eachARCVMH_clustertype/supercluster_props_in_",
               gsub(names(vmhcelldists3)[i],pattern="d(.*)",replacement="\\1"),
               "unit_radius.pdf"),
        height=40,width=15
    )
    do.call("grid.arrange",c(plts,nrow=2))
    dev.off()
})

```


### make a second set faceted by neighboring cell type so that the x axis is all of the VMH or ARC clusters. this one can be a couple three plots wide per row.

```{r}
lapply(c(1:length(rads)),FUN=function(i){
    plts <- list()
    
    plts[[1]] <- ggplot(arccelldists3[[i]],aes(x=label1,y=prop_supercl_neighboring_cell1))+
        geom_violin(aes(fill=label1),scale="width",trim=FALSE)+
        geom_boxplot(width=0.1)+
        facet_wrap(~supercl,scales="free",ncol=3)+
        ggtitle(paste("ARC clusters, ",rads[i]," unit radius\nMean+-SD ",round(mean(arccelldists3[[i]]$total_cell1_neighbors))," +- ",round(sd(arccelldists3[[i]]$total_cell1_neighbors))," total neighbors"))+
        theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 18,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank())
    
    plts[[2]] <- ggplot(vmhcelldists3[[i]],aes(x=label1,y=prop_supercl_neighboring_cell1))+
        geom_violin(aes(fill=label1),scale="width",trim=FALSE)+
        geom_boxplot(width=0.1)+
        facet_wrap(~supercl,scales="free",ncol = 3)+
        ggtitle(paste("VMH clusters, ",rads[i]," unit radius\nMean+-SD ",round(mean(vmhcelldists3[[i]]$total_cell1_neighbors))," +- ",round(sd(vmhcelldists3[[i]]$total_cell1_neighbors))," total neighbors"))+        theme(axis.text.x = element_text(size = 9,angle=90,vjust=0.5), axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size =10), plot.title = element_text(size = 18,hjust=0.5), strip.text = element_text(size=12), legend.text = element_text(size=8), legend.title = element_text(size=9,hjust=0.5),strip.background = element_blank())
    
    pdf(paste0("plots/07_Domains-subdomains_by_knnSmoothing_Banksytypes/05b-Celltype_proportions_in_various_radii_from_eachARCVMH_clustertype/supercluster_props_in_",
               gsub(names(vmhcelldists3)[i],pattern="d(.*)",replacement="\\1"),
               "unit_radius_by_neighbortype.pdf"),
        height=40,width=15
    )
    do.call("grid.arrange",c(plts,nrow=2))
    dev.off()
})
         



```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.38.0 parallelly_1.38.0   colorout_1.3-0.2    gridExtra_2.3      
[5] ggplot2_3.5.1       data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.5      dplyr_1.1.4       compiler_4.4.1    tidyselect_1.2.1 
 [5] scales_1.3.0      yaml_2.3.9        fastmap_1.2.0     here_1.0.1       
 [9] R6_2.5.1          labeling_0.4.3    generics_0.1.3    knitr_1.48       
[13] datasets_4.4.1    tibble_3.2.1      munsell_0.5.1     rprojroot_2.0.4  
[17] pillar_1.9.0      utf8_1.2.4        xfun_0.45         cli_3.6.3        
[21] withr_3.0.0       magrittr_2.0.3    digest_0.6.36     grid_4.4.1       
[25] rstudioapi_0.16.0 lifecycle_1.0.4   vctrs_0.6.5       evaluate_0.24.0  
[29] glue_1.7.0        farver_2.1.2      codetools_0.2-20  fansi_1.0.6      
[33] colorspace_2.1-0  rmarkdown_2.27    tools_4.4.1       pkgconfig_2.0.3  
[37] htmltools_0.5.8.1
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-09-25
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package      * version date (UTC) lib source
   BiocParallel * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table   * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest         0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate       0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi          1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver         2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics       0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggplot2      * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue           1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable         0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling       0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly   * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar         1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
 P rlang        * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8           1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr          3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun           0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
