---
title: "02-plot Bayesspace q9-15-20-31 clusters from harmonyd hvg10-20 nnsvg10-20"
author: "Bernie Mulvey"
date: "2023-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
# library(scater) # addPerCellQC
# library(nnSVG)
library(BiocParallel)
library(scran)
library(parallel)
library(Polychrome)
library(spatialLIBD)
# library(fasthplus)

# theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load object with the harmonyPCA and harmonyUMAPs for these four feature sets for plotting purposes in those spaces. load bayesspace cluster assignments.
```{r}
pcas <- readRDS("analysis/data/H01-feature_selection/SPEs_sampleBlock-hvgs-10or20pctile_nnsvg10or20pctile_Harmony-PCA-UMAP.RDS")

### clone an spe out of this ^ for the spotplotting
hyp2.tmp <- pcas[[1]]

### set colnames(hyp2.tmp) to key for unique rownames etc; these are the rownames used with the bayesspace cluster results
colnames(hyp2.tmp) <- hyp2.tmp$key

bs_clusts <- list.files("analysis/data/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",pattern=".txt")
```

helper function to extract a legend and pipe back to a diff plot: from https://statisticsglobe.com/add-common-legend-to-combined-ggplot2-plots-in-r/#example-2-add-shared-legend-to-ggplot2-plots-using-gridextra-package 
```{r}
extract_legend <- function(my_ggp) {
  step1 <- ggplot_gtable(ggplot_build(my_ggp))
  step2 <- which(sapply(step1$grobs, function(x) x$name) == "guide-box")
  step3 <- step1$grobs[[step2]]
  return(step3)
}


```

spotplots
```{r}

i <- 1
for (i in c(1:length(bs_clusts))) {
    curclusts <-
        fread(
            paste0(
                "analysis/data/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",
                bs_clusts[i]
            )
        )
    
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    
    if (sum(rownames(curclusts) == rownames(colData(hyp2.tmp))) != nrow(colData(hyp2.tmp))) {
        curclusts <- curclusts[rownames(colData(hyp2.tmp)),]
    }
    
    colLabels(hyp2.tmp) <- as.factor(curclusts[, 2])
    pal <- palette36.colors(n = length(unique(curclusts[, 2])))
    # next line is needed for consistency in color-cluster assignments between libdspatial and the generated legend below
    names(pal) <- c(1:length(unique(hyp2.tmp$label)))
    
    ## create a unified legend spanning all possible cluster values (regardless of presence in a given sample)
    leg.tmp <- ggplot(as.data.table(colData(hyp2.tmp))[,.(label,array_row,array_col)],aes(x=array_row,y=array_col,col=label))+geom_point()+
        scale_color_manual(values=as.character(pal))+theme(
                    legend.text = element_text(size = 10),
                    legend.title = element_text(size = 10, hjust = 0.5),
                    legend.spacing.y = unit(0, "cm"),
                    legend.spacing.x = unit(0.1, "cm"))
    leg.tmp2 <- extract_legend(leg.tmp)
    dev.off()
    ### make spotplots
    plt <-
        spatialLIBD::vis_grid_clus(
            hyp2.tmp,
            clustervar = "label",
            pdf_file = NULL,
            colors = pal,
            spatial = F,
            auto_crop = T,
            return_plots = T,
            point_size = 2,
            sort_clust = F
        )
    
    ### adjust font sizes, drop individual plot legends
    plt <- lapply(
        plt,
        FUN = function(x) {
            x +
                theme(
                    plot.title = element_text(size = 11, hjust = 0.5),
                    strip.text = element_text(size = 10),
                )+
                guides(fill="none")
        }
    )
    
    # save
    pdf(
        paste0(
            "plots/H02-clustering/02-bayesspace_q9-15-20-31_HVGorSVG_spotplots/",
            gsub(bs_clusts[i], pattern = "^(.*)\\.txt$", replacement = "\\1"),
            ".pdf"
        ),
        width = 20,
        height = 16
    )
    ### arrangeGrob() on the spotplot list --> 3 x 3 of plots from that list only, constituting the "first column" of the parent; leg.tmp2 is an additional thing being plotted in the second column of the parent. width of the plot grid is set to 10, width of the second column (legend only) set to 2.
    do.call("grid.arrange",list(arrangeGrob(grobs=plt,ncol=3,nrow=3),leg.tmp2,ncol=2,widths=c(18,2)))
dev.off()
    rm(plt, curclusts)
}

rm(hyp2.tmp)
```

## for PCA and UMAP plotting below: set unadjusted data PCA, UMAP dimred names in the list of SPEs to uncor_PCA, uncor_UMAP so that we can trick plotDimRed into using the HARMONY results as in the plotting loop
```{r}
pcas <- lapply(
    pcas,
    FUN = function(x) {
        reducedDimNames(x)[which(reducedDimNames(x) == "PCA")] <-
            "uncor_PCA"
        reducedDimNames(x)[which(reducedDimNames(x) == "UMAP")] <-
            "uncor_UMAP"
        x
    }
)
```

Plot on feature-set-matched harmonyPCA

the BS_clusts file list conveniently is sorted in the same order as the PCA table (hvg10, hvg20, svg10, svg20)
```{r}
i <- 1
for (i in c(1:length(bs_clusts))) {
    k <- ceiling(i / 4)
    curclusts <-
        fread(
            paste0(
                "analysis/data/H02-clustering/bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",
                bs_clusts[i]
            )
        )
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    if (sum(rownames(curclusts) == rownames(colData(pcas[[k]]))) != nrow(colData(pcas[[k]]))) {
        curclusts <- curclusts[rownames(colData(pcas[[k]])), ]
    }
    
    colLabels(pcas[[k]]) <- factor(curclusts[, 2])
    reducedDimNames(pcas[[k]])[which(reducedDimNames(pcas[[k]]) == "HARMONY")] <-
        "PCA"
    colnames(pcas[[k]]@int_colData@listData$reducedDims@listData$PCA)[c(1,2)] <- c("PC1","PC2") 
    
    plt <-
        plotDimRed(
            pcas[[k]],
            type = "PCA",
            annotate = "label",
            palette = paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts[, 2])
            )))
        ) +
        ggtitle(paste0(
            "BSpace ",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_(.*)_.*\\.txt$", replacement =
                     "\\1"),
            " q=",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_.*_(.*)\\.txt$", replacement =
                     "\\1")
        ))
    pdf(
        paste0(
            "plots/H02-clustering/02-bayesspace_q9-15-20-31_HVGorSVG_harmonyPCA/",
            gsub(bs_clusts[i], pattern = "^(.*)\\.txt$", replacement = "\\1"),
            ".pdf"
        ),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    rm(plt, curclusts)
}

```


plot onto respective UMAPs
```{r}

i <- 1
for (i in c(1:length(bs_clusts))) {
    k <- ceiling(i / 4)
    
    curclusts <- fread(
        paste0("analysis/data/H02-clustering/bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",
               bs_clusts[i]))
    
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    if (sum(rownames(curclusts) == rownames(colData(pcas[[k]]))) != nrow(colData(pcas[[k]]))) {
        curclusts <- curclusts[rownames(colData(pcas[[k]])), ]
    }
    
    colLabels(pcas[[k]]) <- factor(curclusts[, 2])
    reducedDimNames(pcas[[k]])[which(reducedDimNames(pcas[[k]]) == "HARMONY_UMAP")] <- "UMAP"
    colnames(pcas[[k]]@int_colData@listData$reducedDims@listData$UMAP) <- c("UMAP1","UMAP2") 
    
    plt <-
        plotDimRed(
            pcas[[k]],
            type = "UMAP",
            annotate = "label",
            palette = paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts[, 2]))))) +
        ggtitle(paste0("BSpace ",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_(.*)_.*\\.txt$", replacement =
                     "\\1"),
            " q=",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_.*_(.*)\\.txt$", replacement =
                     "\\1")))
    
    pdf(paste0("plots/H02-clustering/02-bayesspace_q9-15-20-31_HVGorSVG_harmonyUMAP/",
            gsub(bs_clusts[i], pattern = "^(.*)\\.txt$", replacement = "\\1"),
            ".pdf"),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    
    rm(plt, curclusts)
}
```
