---
title: "05-AutosomalSex DEG GSEA Setup"
author: "Bernard Mulvey"
date: "2023-10-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(spatialLIBD)
require(colorout)
source("~/.R/calc_tau_for_cellOrTissueSpecificity_metric.R")
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

# try this first with dreamlet ran fx results
# original call-dreamlet::processAssays(proc,min.cells = 10,min.count = 1,min.samples = 4,min.prop = 0.4,normalize.method = "TMM",useCountsWeights = T,BPPARAM = sbp,formula = ~Sex+(1|brnum))
# coef: SexM
### ^ logFC>0 = M>F ###
```{r}
sde <- readRDS("analysis/data/spe_053123/H07-Outside Dataset Analyses and Comparisons/spatialHYP_dreamletSexDE_brnum-as-random-effect-inDEmodel.RDS")
sde <- as.data.table(sde)

i<-1
for (i in c(1:15)){
  curdom <- paste0("x",i)
  curdat <- copy(sde)[assay==curdom]
  setorder(curdat,-logFC)
  write.table(curdat[,.(ID,logFC)],paste0("analysis/data/spe_053123/H08-GSEA, GO etc/H08.05-Sex DEG GSEA/Domain dreamlet ranef lfcs for GSEA prerank input/",curdom,".rnk"),sep='\t',quote=F,row.names=F,col.names=F)
  rm(curdom,curdat)
}
```

run parameters:
#### gene id "chip"
ftp.broadinstitute.org://pub/gsea/msigdb/human/annotations/Human_Gene_Symbol_with_Remapping_MSigDB.v2023.1.Hs.chip

#### settings (all defaults except collapse)
collapse
nperm 1000
norm meandiv
scoring_scheme weighted
min 15 max 500

#### datasets
h.all.v2023.1.Hs.symbols.gmt
ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c1.all.v2023.1.Hs.symbols.gmt
ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c2.all.v2023.1.Hs.symbols.gmt
ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c3.all.v2023.1.Hs.symbols.gmt,
ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c5.all.v2023.1.Hs.symbols.gmt,
ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c6.all.v2023.1.Hs.symbols.gmt,
ftp.broadinstitute.org://pub/gsea/msigdb/human/gene_sets/c8.all.v2023.1.Hs.symbols.gmt
