---
title: "04-Save Yao 23 Sex DE Comparison Results to Supp Tab Sheets"
author: "Bernard Mulvey"
date: "2024-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

```

load dreamlet results
```{r}
dlt <- fread("data/12-Mouse comparisons/02-Yao23-HYP-subclass-sexDE_dreamlet_donormouse-ranfx.txt")

setnames(dlt,c(1,2),c("Yao23 Cell Subclass","Mouse Gene Symbol"))
dlt <- dlt[,c(1,2,3,7,5,9,8,6,4)]

saveRDS(dlt,"local_supptab_assembly/Supp Data-Yao23 HYP sex DE with random effect of donor.RDS")

rm(dlt)
```

load -> write registration results
```{r}
sexreg <- readRDS("data/12-Mouse comparisons/03-Yao23-subclass-MvFsexDE-tstat_correls_to_k15-20-31-15collapsed_bytopNhumanDEGs.RDS")


names(sexreg)

fn <- names(sexreg)
mapply(x=sexreg,y=fn,FUN=function(x,y){fwrite(x,paste0("local_supptab_assembly/Yao Sex DE correls ",y,".txt"),quote=F,sep='\t')})
rm(sexreg,fn)
```

```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] default_1.0.0      colorout_1.3-0.1   data.table_1.14.10
[4] rlang_1.1.3       

loaded via a namespace (and not attached):
  [1] RColorBrewer_1.1-3            spatialLIBD_1.14.1           
  [3] rstudioapi_0.15.0             jsonlite_1.8.8               
  [5] magrittr_2.0.3                ggbeeswarm_0.7.2             
  [7] magick_2.8.2                  rmarkdown_2.25               
  [9] BiocIO_1.12.0                 fields_15.2                  
 [11] zlibbioc_1.48.0               vctrs_0.6.5                  
 [13] DelayedMatrixStats_1.24.0     memoise_2.0.1                
 [15] config_0.3.2                  Rsamtools_2.18.0             
 [17] paletteer_1.6.0               RCurl_1.98-1.14              
 [19] benchmarkme_1.0.8             htmltools_0.5.7              
 [21] S4Arrays_1.2.0                AnnotationHub_3.10.0         
 [23] curl_5.2.0                    BiocNeighbors_1.20.2         
 [25] SparseArray_1.2.3             sass_0.4.8                   
 [27] bslib_0.6.1                   htmlwidgets_1.6.4            
 [29] plotly_4.10.4                 cachem_1.0.8                 
 [31] GenomicAlignments_1.38.2      mime_0.12                    
 [33] lifecycle_1.0.4               iterators_1.0.14             
 [35] pkgconfig_2.0.3               rsvd_1.0.5                   
 [37] Matrix_1.6-5                  R6_2.5.1                     
 [39] fastmap_1.1.1                 GenomeInfoDbData_1.2.11      
 [41] MatrixGenerics_1.14.0         shiny_1.8.0                  
 [43] digest_0.6.34                 colorspace_2.1-0             
 [45] rematch2_2.1.2                AnnotationDbi_1.64.1         
 [47] S4Vectors_0.40.2              scater_1.30.1                
 [49] rprojroot_2.0.4               irlba_2.3.5.1                
 [51] ExperimentHub_2.10.0          GenomicRanges_1.54.1         
 [53] RSQLite_2.3.5                 beachmat_2.18.0              
 [55] filelock_1.0.3                fansi_1.0.6                  
 [57] httr_1.4.7                    abind_1.4-5                  
 [59] compiler_4.3.2                here_1.0.1                   
 [61] bit64_4.0.5                   doParallel_1.0.17            
 [63] attempt_0.3.1                 BiocParallel_1.36.0          
 [65] viridis_0.6.5                 DBI_1.2.1                    
 [67] maps_3.4.2                    sessioninfo_1.2.2            
 [69] rappdirs_0.3.3                DelayedArray_0.28.0          
 [71] rjson_0.2.21                  tools_4.3.2                  
 [73] vipor_0.4.7                   beeswarm_0.4.0               
 [75] interactiveDisplayBase_1.40.0 httpuv_1.6.13                
 [77] glue_1.7.0                    restfulr_0.0.15              
 [79] promises_1.2.1                grid_4.3.2                   
 [81] generics_0.1.3                gtable_0.3.4                 
 [83] tidyr_1.3.1                   ScaledMatrix_1.10.0          
 [85] BiocSingular_1.18.0           utf8_1.2.4                   
 [87] XVector_0.42.0                BiocGenerics_0.48.1          
 [89] ggrepel_0.9.5                 BiocVersion_3.18.1           
 [91] foreach_1.5.2                 pillar_1.9.0                 
 [93] spam_2.10-0                   limma_3.58.1                 
 [95] later_1.3.2                   benchmarkmeData_1.0.4        
 [97] dplyr_1.1.4                   BiocFileCache_2.10.1         
 [99] lattice_0.22-5                rtracklayer_1.62.0           
[101] bit_4.0.5                     tidyselect_1.2.0             
[103] SingleCellExperiment_1.24.0   locfit_1.5-9.8               
[105] scuttle_1.12.0                Biostrings_2.70.2            
[107] knitr_1.45                    gridExtra_2.3                
[109] IRanges_2.36.0                edgeR_4.0.14                 
[111] SummarizedExperiment_1.32.0   stats4_4.3.2                 
[113] xfun_0.41                     Biobase_2.62.0               
[115] statmod_1.5.0                 matrixStats_1.2.0            
[117] DT_0.31                       lazyeval_0.2.2               
[119] yaml_2.3.8                    shinyWidgets_0.8.1           
[121] evaluate_0.23                 codetools_0.2-19             
[123] tibble_3.2.1                  BiocManager_1.30.22          
[125] cli_3.6.2                     xtable_1.8-4                 
[127] jquerylib_0.1.4               munsell_0.5.0                
[129] golem_0.4.1                   Rcpp_1.0.12                  
[131] GenomeInfoDb_1.38.5           dbplyr_2.4.0                 
[133] png_0.1-8                     XML_3.99-0.16.1              
[135] parallel_4.3.2                ellipsis_0.3.2               
[137] ggplot2_3.4.4                 blob_1.2.4                   
[139] dotCall64_1.1-1               sparseMatrixStats_1.14.0     
[141] bitops_1.0-7                  SpatialExperiment_1.12.0     
[143] viridisLite_0.4.2             scales_1.3.0                 
[145] purrr_1.0.2                   crayon_1.5.2                 
[147] cowplot_1.1.3                 KEGGREST_1.42.0              
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-02-14
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package                * version     date (UTC) lib source
 abind                    1.4-5       2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1      2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0      2023-10-24 [1] Bioconductor
 attempt                  0.3.1       2020-05-03 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0      2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0       2021-06-01 [1] CRAN (R 4.3.2)
 benchmarkme              1.0.8       2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4       2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                  2.62.0      2023-10-24 [1] Bioconductor
 BiocFileCache            2.10.1      2023-10-26 [1] Bioconductor
 BiocGenerics             0.48.1      2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0      2023-10-26 [1] Bioconductor
 BiocManager              1.30.22     2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.2      2024-01-07 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel             1.36.0      2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0      2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1      2023-11-18 [1] Bioconductor 3.18 (R 4.3.2)
 Biostrings               2.70.2      2024-01-28 [1] Bioconductor 3.18 (R 4.3.2)
 bit                      4.0.5       2022-11-15 [1] CRAN (R 4.3.2)
 bit64                    4.0.5       2020-08-30 [1] CRAN (R 4.3.2)
 bitops                   1.0-7       2021-04-24 [1] CRAN (R 4.3.2)
 blob                     1.2.4       2023-03-17 [1] CRAN (R 4.3.0)
 bslib                    0.6.1       2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8       2023-05-01 [1] CRAN (R 4.3.2)
 cli                      3.6.2       2023-12-11 [1] CRAN (R 4.3.2)
 codetools                0.2-19      2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0.1     2024-01-11 [1] local
 colorspace               2.1-0       2023-01-23 [1] CRAN (R 4.3.2)
 config                   0.3.2       2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.3       2024-01-22 [1] CRAN (R 4.3.1)
 crayon                   1.5.2       2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0       2023-12-08 [1] CRAN (R 4.3.2)
 data.table             * 1.14.10     2023-12-08 [1] CRAN (R 4.3.2)
 DBI                      1.2.1       2024-01-12 [1] CRAN (R 4.3.2)
 dbplyr                   2.4.0       2023-10-26 [1] CRAN (R 4.3.1)
 default                * 1.0.0       2017-08-07 [1] CRAN (R 4.3.0)
 DelayedArray             0.28.0      2023-10-24 [1] Bioconductor
 DelayedMatrixStats       1.24.0      2023-11-06 [1] Bioconductor
 digest                   0.6.34      2024-01-11 [1] CRAN (R 4.3.2)
 doParallel               1.0.17      2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1       2023-11-28 [1] CRAN (R 4.3.2)
 dplyr                    1.1.4       2023-11-17 [1] CRAN (R 4.3.2)
 DT                       0.31        2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.14      2024-01-29 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2       2021-04-29 [1] CRAN (R 4.3.2)
 evaluate                 0.23        2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0      2023-10-26 [1] Bioconductor
 fansi                    1.0.6       2023-12-08 [1] CRAN (R 4.3.2)
 fastmap                  1.1.1       2023-02-24 [1] CRAN (R 4.3.2)
 fields                   15.2        2023-08-17 [1] CRAN (R 4.3.2)
 filelock                 1.0.3       2023-12-11 [1] CRAN (R 4.3.2)
 foreach                  1.5.2       2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3       2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb             1.38.5      2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData         1.2.11      2024-01-11 [1] Bioconductor
 GenomicAlignments        1.38.2      2024-01-16 [1] Bioconductor 3.18 (R 4.3.2)
 GenomicRanges            1.54.1      2023-10-29 [1] Bioconductor
 ggbeeswarm               0.7.2       2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                  3.4.4       2023-10-12 [1] CRAN (R 4.3.1)
 ggrepel                  0.9.5       2024-01-10 [1] CRAN (R 4.3.2)
 glue                     1.7.0       2024-01-09 [1] CRAN (R 4.3.2)
 golem                    0.4.1       2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra                2.3         2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4       2023-08-21 [1] CRAN (R 4.3.0)
 here                     1.0.1       2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7       2023-11-03 [1] CRAN (R 4.3.2)
 htmlwidgets              1.6.4       2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13      2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7       2023-08-15 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0      2023-10-26 [1] Bioconductor
 IRanges                  2.36.0      2023-10-24 [1] Bioconductor
 irlba                    2.3.5.1     2022-10-03 [1] CRAN (R 4.3.2)
 iterators                1.0.14      2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4       2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8       2023-12-04 [1] CRAN (R 4.3.2)
 KEGGREST                 1.42.0      2023-10-26 [1] Bioconductor
 knitr                    1.45        2023-10-30 [1] CRAN (R 4.3.2)
 later                    1.3.2       2023-12-06 [1] CRAN (R 4.3.2)
 lattice                  0.22-5      2023-10-24 [1] CRAN (R 4.3.2)
 lazyeval                 0.2.2       2019-03-15 [1] CRAN (R 4.3.2)
 lifecycle                1.0.4       2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1      2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8     2023-06-11 [1] CRAN (R 4.3.2)
 magick                   2.8.2       2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3       2022-03-30 [1] CRAN (R 4.3.2)
 maps                     3.4.2       2023-12-15 [1] CRAN (R 4.3.1)
 Matrix                   1.6-5       2024-01-11 [1] CRAN (R 4.3.2)
 MatrixGenerics           1.14.0      2023-10-26 [1] Bioconductor
 matrixStats              1.2.0       2023-12-11 [1] CRAN (R 4.3.2)
 memoise                  2.0.1       2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12        2021-09-28 [1] CRAN (R 4.3.2)
 munsell                  0.5.0       2018-06-12 [1] CRAN (R 4.3.0)
 paletteer                1.6.0       2024-01-21 [1] CRAN (R 4.3.1)
 pillar                   1.9.0       2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3       2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.4      2024-01-13 [1] CRAN (R 4.3.1)
 png                      0.1-8       2022-11-29 [1] CRAN (R 4.3.2)
 promises                 1.2.1       2023-08-10 [1] CRAN (R 4.3.2)
 purrr                    1.0.2       2023-08-10 [1] CRAN (R 4.3.2)
 R6                       2.5.1       2021-08-19 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3       2021-01-31 [1] CRAN (R 4.3.2)
 RColorBrewer             1.1-3       2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.12      2024-01-09 [1] CRAN (R 4.3.2)
 RCurl                    1.98-1.14   2024-01-09 [1] CRAN (R 4.3.1)
 rematch2                 2.1.2       2020-05-01 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15      2022-06-16 [1] CRAN (R 4.3.2)
 rjson                    0.2.21      2022-01-09 [1] CRAN (R 4.3.2)
 rlang                  * 1.1.3       2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown                2.25        2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot                2.0.4       2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0      2023-10-24 [1] Bioconductor
 RSQLite                  2.3.5       2024-01-21 [1] CRAN (R 4.3.2)
 rstudioapi               0.15.0      2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5       2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0      2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0       2023-10-24 [1] Bioconductor
 S4Vectors                0.40.2      2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8       2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0      2023-11-06 [1] Bioconductor
 scales                   1.3.0       2023-11-28 [1] CRAN (R 4.3.2)
 scater                   1.30.1      2023-11-16 [1] Bioconductor
 scuttle                  1.12.0      2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2       2021-12-06 [1] CRAN (R 4.3.0)
 shiny                    1.8.0       2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.1       2024-01-10 [1] CRAN (R 4.3.1)
 SingleCellExperiment     1.24.0      2023-11-06 [1] Bioconductor
 spam                     2.10-0      2023-10-23 [1] CRAN (R 4.3.2)
 SparseArray              1.2.3       2023-12-25 [1] Bioconductor 3.18 (R 4.3.2)
 sparseMatrixStats        1.14.0      2023-10-24 [1] Bioconductor
 SpatialExperiment        1.12.0      2023-10-26 [1] Bioconductor
 spatialLIBD              1.14.1      2023-11-30 [1] Bioconductor 3.18 (R 4.3.2)
 statmod                  1.5.0       2023-01-06 [1] CRAN (R 4.3.2)
 SummarizedExperiment     1.32.0      2023-11-06 [1] Bioconductor
 tibble                   3.2.1       2023-03-20 [1] CRAN (R 4.3.2)
 tidyr                    1.3.1       2024-01-24 [1] CRAN (R 4.3.1)
 tidyselect               1.2.0       2022-10-10 [1] CRAN (R 4.3.0)
 utf8                     1.2.4       2023-10-22 [1] CRAN (R 4.3.2)
 vctrs                    0.6.5       2023-12-01 [1] CRAN (R 4.3.2)
 vipor                    0.4.7       2023-12-18 [1] CRAN (R 4.3.1)
 viridis                  0.6.5       2024-01-29 [1] CRAN (R 4.3.2)
 viridisLite              0.4.2       2023-05-02 [1] CRAN (R 4.3.0)
 xfun                     0.41        2023-11-01 [1] CRAN (R 4.3.2)
 XML                      3.99-0.16.1 2024-01-22 [1] CRAN (R 4.3.1)
 xtable                   1.8-4       2019-04-21 [1] CRAN (R 4.3.0)
 XVector                  0.42.0      2023-10-24 [1] Bioconductor
 yaml                     2.3.8       2023-12-11 [1] CRAN (R 4.3.2)
 zlibbioc                 1.48.0      2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
