---
title: "01-BayesSpace domain registration to ABA adult mouse ISH brain atlas HYP"
output: html_document
date: "2024-08-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure. part of this is adding the line 
# OBJC_DISABLE_INITIALIZE_FORK_SAFETY=YES
# to Renviron.site. see e.g. top response on https://stackoverflow.com/questions/73638290/python-on-mac-is-it-safe-to-set-objc-disable-initialize-fork-safety-yes-globall 
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))

## last of all, unload the base package datasets, whose data keep getting in the way of autocompletions (e.g., Theoph is priortized over TRUE)
unloadNamespace("datasets")
```

### make scCoco result plots-heatmap format
```{r}
cocores <- readRDS("processed-data/10-Spatial Registration/01b-scCoco query of top 150 markers per k15-20-31-collapsedK15 domain in ABA ms hyp CCF areas.RDS")

cocores <- lapply(cocores,function(x){
    y <- as.data.table(x[["scores_per_target_level_region_all"]])
    setnames(y,c(3:ncol(y)),gsub(names(y)[3:ncol(y)],pattern="k.*_HARMONYlmbna_nnsvg10_60kiter.*_(.*)",replacement="\\1"))
    return(y)
})

setnames(cocores,c("XVMH","XARC"),c("VMH","ARC"))
cocores <- cocores[,.(topnode,topname,VMH,ARC,X1, X2, X4, X6, X7, X8, X9, X10, X13, X14, X15)]

### only keep regions that are in the top 3 for any cluster
cocokeep <- lapply(cocores,function(x){
    keepres <- c("")
    i <- 3
    for (i in c(3:ncol(x))){
        curvar <- names(x)[i]
        # order the table by the current visium cluster of interest in desc. order
        setorderv(x,curvar,-1)
        # get the mouse brain anatomic area for these top 3 areas
        keepres <- c(keepres,x[c(1:3),topname])
    }
    return(keepres)
})

cocores <- mapply(X=cocores,Y=cocokeep,SIMPLIFY=FALSE,function(X,Y){
    Z <- X[topname %in% Y]
    return(Z)
})

## melt into plottable-er format
mapply(X=cocores,Y=names(cocores),SIMPLIFY=F,FUN=function(X,Y){
    Z <- melt(X,id.vars=c("topnode","topname"),variable.name="Visium Cluster",value.name="score")
    
    setnames(Z,"score","scCoco Score (top\n150 Visium Markers)")
    if(nrow(Z[topname=="Periventricular hypothalamic nucleus, posterior part"])>1){
        Z[topname=="Periventricular hypothalamic nucleus, posterior part",topname:="Posterior periventricular nu."]
    }
    
    ## relevel visium clusters to put VMH/ARC clusters leftmost along the x
    Z[,`Visium Cluster`:=as.character(`Visium Cluster`)]
    arcclusts <- unique(grep(Z$`Visium Cluster`,pattern="ARC",value=T))
    vmhclusts <- unique(grep(Z$`Visium Cluster`,pattern="VMH",value=T))
    vmharcclusts <- c(arcclusts,vmhclusts)
    others <- unique(Z[!(`Visium Cluster` %in% vmharcclusts),`Visium Cluster`])
    Z[,`Visium Cluster`:=factor(`Visium Cluster`,levels=c(arcclusts,vmhclusts,others))]
    ## relevel mouse VMH/ARC to put them in the top two rows of the Y axis
    otherlevs <- unique(Z$topname[!Z$topname %in% c("Ventromedial hypothalamic nucleus","Arcuate hypothalamic nucleus")])
    Z[,topname:=factor(topname,levels=c(otherlevs,"Ventromedial hypothalamic nucleus","Arcuate hypothalamic nucleus"))]
    # return(Z)
    
    plt <- ggplot(Z,aes(x=`Visium Cluster`,y=topname,fill=`scCoco Score (top\n150 Visium Markers)`))+
        geom_tile()+
        scale_fill_gradient2(low="white",mid="yellow",high="red",midpoint=0.5)+
        theme(axis.text.x = element_text(size = 9,angle = 90, hjust = 1))+
        scale_x_discrete(expand = c(0,0))+
        scale_y_discrete(expand=c(0,0))+
        ylab("Allen Mouse Atlas Region Name")+
        xlab(paste0("Visium Domain"))+
        ggtitle(gsub(gsub(Y,pattern="_",replacement=" "),pattern="lmbna",replacement=""))+
    theme(axis.title.x = element_text(size = 10), axis.text.y = element_text(size = 9), axis.title.y = element_text(size=11),legend.title = element_text(size=10,hjust=0.5),legend.text=element_text(size=9))
    
    pdf(paste0("plots/10-Spatial Registration/01-scCoco registration of ",Y," to ABA adult mouse ISH hyp areas.pdf"),width=9,height=6)
    print(plt)
dev.off()
    
})


```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     methods   base     

other attached packages:
[1] BiocParallel_1.38.0 parallelly_1.38.0   colorout_1.3-0.2    gridExtra_2.3      
[5] ggplot2_3.5.1       data.table_1.15.4   rlang_1.1.4        

loaded via a namespace (and not attached):
 [1] gtable_0.3.5      dplyr_1.1.4       compiler_4.4.1    tidyselect_1.2.1 
 [5] parallel_4.4.1    scales_1.3.0      yaml_2.3.9        fastmap_1.2.0    
 [9] here_1.0.1        R6_2.5.1          labeling_0.4.3    generics_0.1.3   
[13] knitr_1.48        tibble_3.2.1      munsell_0.5.1     rprojroot_2.0.4  
[17] pillar_1.9.0      utf8_1.2.4        xfun_0.45         cli_3.6.3        
[21] withr_3.0.0       magrittr_2.0.3    digest_0.6.36     grid_4.4.1       
[25] rstudioapi_0.16.0 lifecycle_1.0.4   vctrs_0.6.5       evaluate_0.24.0  
[29] glue_1.7.0        farver_2.1.2      sessioninfo_1.2.2 codetools_0.2-20 
[33] fansi_1.0.6       colorspace_2.1-0  rmarkdown_2.27    tools_4.4.1      
[37] pkgconfig_2.0.3   htmltools_0.5.8.1
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-04
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package      * version date (UTC) lib source
   BiocParallel * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 P cli            3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools      0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout     * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace     2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   data.table   * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest         0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr          1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate       0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi          1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   farver         2.1.2   2024-05-13 [1] CRAN (R 4.4.0)
   fastmap        1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics       0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   ggplot2      * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue           1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra    * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable         0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here           1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools      0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr          1.48    2024-07-07 [1] CRAN (R 4.4.1)
   labeling       0.4.3   2023-08-29 [1] CRAN (R 4.4.0)
   lifecycle      1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magrittr       2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   munsell        0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly   * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pillar         1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig      2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6             2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
 P rlang        * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown      2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot      2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi     0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   scales         1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo    1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   tibble         3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect     1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   utf8           1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs          0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr          3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun           0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml           2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
