---
title: "02-spatialLIBD Registration to Yao HYP subclasses or Yao HYP Subclasses with unitary VMH and ARC"
author: "Bernard Mulvey"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(ggplot2)
library(gridExtra)
require(colorout)
library(BiocParallel)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

Load cross-correlation results 
```{r}
spatregs <- readRDS("data/10-Spatial Registration/02g-sphyp-k15-20-31-15clps_ABAms-47subclass-47withVMHARCclps.RDS")
```

manually plot each; collapse all lists of results into one big list
## identations in this script are messed up, apologies to anyone reading
```{r}
plotlist <- c(
        spatregs[[1]],
        spatregs[[2]],
        spatregs[[3]],
        spatregs[[4]],
        spatregs[[5]],
        spatregs[[6]],
        spatregs[[7]],
        spatregs[[8]]
)
stopifnot(length(unique(names(plotlist)))==length(spatregs)*length(spatregs[[1]]))

sbp <- MulticoreParam(8)
register(sbp)

# plots
bpmapply(m=plotlist,n=as.list(names(plotlist)),SIMPLIFY = F,BPPARAM = sbp,FUN = function(m,n){
        plttab <- as.data.table(m,keep.rownames=T)
        plttab <- melt(plttab,id.vars="sphyp_domain")
        plt <- ggplot(plttab,aes(x=sphyp_domain,y=variable,fill=value))+
                geom_tile()+
                ggtitle(gsub(n,pattern="_",replacement=" "))+
                theme(axis.text.x = element_text(angle=90,size=9,hjust=1),axis.text.y=element_text(size=9),title = element_text(size=12))+
                scale_fill_viridis_c()
        n <-  ifelse(as.numeric(gsub(
                n,pattern="^.*_top_(.*)_sp.*$",replacement="\\1"))>1000,
                yes=gsub(n,pattern="^(.*)_top_.*_(sp.*)$",replacement=paste0("\\1","_allgenes_","\\2")),
                no=n)
        pdf(paste0("plots/10-Spatial Registration/02-",n,".pdf"),height = 0.2*ncol(m),width=10)
        print(plt)
        dev.off()
})
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.2 (2023-10-31)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/New_York
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] orthogene_1.8.0     BiocParallel_1.36.0 colorout_1.3-0.1   
 [4] gridExtra_2.3       ggplot2_3.4.4       Biostrings_2.70.2  
 [7] GenomeInfoDb_1.38.5 XVector_0.42.0      IRanges_2.36.0     
[10] S4Vectors_0.40.2    BiocGenerics_0.48.1 data.table_1.14.10 
[13] rlang_1.1.3        

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.0          viridisLite_0.4.2        
 [3] dplyr_1.1.4               bitops_1.0-7             
 [5] fastmap_1.1.1             RCurl_1.98-1.14          
 [7] lazyeval_0.2.2            homologene_1.4.68.19.3.27
 [9] digest_0.6.34             lifecycle_1.0.4          
[11] tidytree_0.4.6            magrittr_2.0.3           
[13] compiler_4.3.2            tools_4.3.2              
[15] utf8_1.2.4                yaml_2.3.8               
[17] knitr_1.45                ggsignif_0.6.4           
[19] htmlwidgets_1.6.4         here_1.0.1               
[21] aplot_0.2.2               abind_1.4-5              
[23] babelgene_22.9            withr_3.0.0              
[25] purrr_1.0.2               grid_4.3.2               
[27] fansi_1.0.6               ggpubr_0.6.0             
[29] colorspace_2.1-0          scales_1.3.0             
[31] cli_3.6.2                 rmarkdown_2.25           
[33] crayon_1.5.2              treeio_1.26.0            
[35] generics_0.1.3            rstudioapi_0.15.0        
[37] ggtree_3.10.0             httr_1.4.7               
[39] gprofiler2_0.2.2          sessioninfo_1.2.2        
[41] ape_5.7-1                 cachem_1.0.8             
[43] zlibbioc_1.48.0           parallel_4.3.2           
[45] ggplotify_0.1.2           vctrs_0.6.5              
[47] yulab.utils_0.1.4         Matrix_1.6-5             
[49] jsonlite_1.8.8            carData_3.0-5            
[51] car_3.1-2                 gridGraphics_0.5-1       
[53] patchwork_1.2.0           rstatix_0.7.2            
[55] plotly_4.10.4             grr_0.9.5                
[57] tidyr_1.3.1               glue_1.7.0               
[59] codetools_0.2-19          gtable_0.3.4             
[61] munsell_0.5.0             tibble_3.2.1             
[63] pillar_1.9.0              htmltools_0.5.7          
[65] GenomeInfoDbData_1.2.11   R6_2.5.1                 
[67] rprojroot_2.0.4           evaluate_0.23            
[69] lattice_0.22-5            backports_1.4.1          
[71] memoise_2.0.1             broom_1.0.5              
[73] ggfun_0.1.4               Rcpp_1.0.12              
[75] nlme_3.1-164              xfun_0.41                
[77] fs_1.6.3                  pkgconfig_2.0.3   

> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.2 (2023-10-31)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/New_York
 date     2024-02-07
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package          * version        date (UTC) lib source
 abind              1.4-5          2016-07-21 [1] CRAN (R 4.3.0)
 ape                5.7-1          2023-03-13 [1] CRAN (R 4.3.2)
 aplot              0.2.2          2023-10-06 [1] CRAN (R 4.3.1)
 babelgene          22.9           2022-09-29 [1] CRAN (R 4.3.0)
 backports          1.4.1          2021-12-13 [1] CRAN (R 4.3.2)
 BiocGenerics     * 0.48.1         2023-11-02 [1] Bioconductor
 BiocParallel     * 1.36.0         2023-10-26 [1] Bioconductor
 Biostrings       * 2.70.2         2024-01-28 [1] Bioconductor 3.18 (R 4.3.2)
 bitops             1.0-7          2021-04-24 [1] CRAN (R 4.3.2)
 broom              1.0.5          2023-06-09 [1] CRAN (R 4.3.0)
 cachem             1.0.8          2023-05-01 [1] CRAN (R 4.3.2)
 car                3.1-2          2023-03-30 [1] CRAN (R 4.3.0)
 carData            3.0-5          2022-01-06 [1] CRAN (R 4.3.0)
 cli                3.6.2          2023-12-11 [1] CRAN (R 4.3.2)
 codetools          0.2-19         2023-02-01 [1] CRAN (R 4.3.0)
 colorout         * 1.3-0.1        2024-01-11 [1] local
 colorspace         2.1-0          2023-01-23 [1] CRAN (R 4.3.2)
 crayon             1.5.2          2022-09-29 [1] CRAN (R 4.3.0)
 data.table       * 1.14.10        2023-12-08 [1] CRAN (R 4.3.2)
 digest             0.6.34         2024-01-11 [1] CRAN (R 4.3.2)
 dplyr              1.1.4          2023-11-17 [1] CRAN (R 4.3.2)
 evaluate           0.23           2023-11-01 [1] CRAN (R 4.3.1)
 fansi              1.0.6          2023-12-08 [1] CRAN (R 4.3.2)
 fastmap            1.1.1          2023-02-24 [1] CRAN (R 4.3.2)
 fs                 1.6.3          2023-07-20 [1] CRAN (R 4.3.0)
 generics           0.1.3          2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb     * 1.38.5         2023-12-30 [1] Bioconductor 3.18 (R 4.3.2)
 GenomeInfoDbData   1.2.11         2024-01-11 [1] Bioconductor
 ggfun              0.1.4          2024-01-19 [1] CRAN (R 4.3.2)
 ggplot2          * 3.4.4          2023-10-12 [1] CRAN (R 4.3.1)
 ggplotify          0.1.2          2023-08-09 [1] CRAN (R 4.3.0)
 ggpubr             0.6.0          2023-02-10 [1] CRAN (R 4.3.0)
 ggsignif           0.6.4          2022-10-13 [1] CRAN (R 4.3.0)
 ggtree             3.10.0         2023-11-06 [1] Bioconductor
 glue               1.7.0          2024-01-09 [1] CRAN (R 4.3.2)
 gprofiler2         0.2.2          2023-06-14 [1] CRAN (R 4.3.0)
 gridExtra        * 2.3            2017-09-09 [1] CRAN (R 4.3.0)
 gridGraphics       0.5-1          2020-12-13 [1] CRAN (R 4.3.0)
 grr                0.9.5          2016-08-26 [1] CRAN (R 4.3.2)
 gtable             0.3.4          2023-08-21 [1] CRAN (R 4.3.0)
 here               1.0.1          2020-12-13 [1] CRAN (R 4.3.0)
 homologene         1.4.68.19.3.27 2019-03-28 [1] CRAN (R 4.3.0)
 htmltools          0.5.7          2023-11-03 [1] CRAN (R 4.3.2)
 htmlwidgets        1.6.4          2023-12-06 [1] CRAN (R 4.3.1)
 httr               1.4.7          2023-08-15 [1] CRAN (R 4.3.0)
 IRanges          * 2.36.0         2023-10-24 [1] Bioconductor
 jsonlite           1.8.8          2023-12-04 [1] CRAN (R 4.3.2)
 knitr              1.45           2023-10-30 [1] CRAN (R 4.3.2)
 lattice            0.22-5         2023-10-24 [1] CRAN (R 4.3.2)
 lazyeval           0.2.2          2019-03-15 [1] CRAN (R 4.3.2)
 lifecycle          1.0.4          2023-11-07 [1] CRAN (R 4.3.1)
 magrittr           2.0.3          2022-03-30 [1] CRAN (R 4.3.2)
 Matrix             1.6-5          2024-01-11 [1] CRAN (R 4.3.2)
 memoise            2.0.1          2021-11-26 [1] CRAN (R 4.3.0)
 munsell            0.5.0          2018-06-12 [1] CRAN (R 4.3.0)
 nlme               3.1-164        2023-11-27 [1] CRAN (R 4.3.2)
 orthogene        * 1.8.0          2023-11-06 [1] Bioconductor
 patchwork          1.2.0          2024-01-08 [1] CRAN (R 4.3.1)
 pillar             1.9.0          2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig          2.0.3          2019-09-22 [1] CRAN (R 4.3.0)
 plotly             4.10.4         2024-01-13 [1] CRAN (R 4.3.1)
 purrr              1.0.2          2023-08-10 [1] CRAN (R 4.3.2)
 R6                 2.5.1          2021-08-19 [1] CRAN (R 4.3.0)
 Rcpp               1.0.12         2024-01-09 [1] CRAN (R 4.3.2)
 RCurl              1.98-1.14      2024-01-09 [1] CRAN (R 4.3.1)
 rlang            * 1.1.3          2024-01-10 [1] CRAN (R 4.3.2)
 rmarkdown          2.25           2023-09-18 [1] CRAN (R 4.3.2)
 rprojroot          2.0.4          2023-11-05 [1] CRAN (R 4.3.1)
 rstatix            0.7.2          2023-02-01 [1] CRAN (R 4.3.0)
 rstudioapi         0.15.0         2023-07-07 [1] CRAN (R 4.3.0)
 S4Vectors        * 0.40.2         2023-11-23 [1] Bioconductor 3.18 (R 4.3.2)
 scales             1.3.0          2023-11-28 [1] CRAN (R 4.3.2)
 sessioninfo        1.2.2          2021-12-06 [1] CRAN (R 4.3.0)
 tibble             3.2.1          2023-03-20 [1] CRAN (R 4.3.2)
 tidyr              1.3.1          2024-01-24 [1] CRAN (R 4.3.1)
 tidyselect         1.2.0          2022-10-10 [1] CRAN (R 4.3.0)
 tidytree           0.4.6          2023-12-12 [1] CRAN (R 4.3.1)
 treeio             1.26.0         2023-11-06 [1] Bioconductor
 utf8               1.2.4          2023-10-22 [1] CRAN (R 4.3.2)
 vctrs              0.6.5          2023-12-01 [1] CRAN (R 4.3.2)
 viridisLite        0.4.2          2023-05-02 [1] CRAN (R 4.3.0)
 withr              3.0.0          2024-01-16 [1] CRAN (R 4.3.1)
 xfun               0.41           2023-11-01 [1] CRAN (R 4.3.2)
 XVector          * 0.42.0         2023-10-24 [1] Bioconductor
 yaml               2.3.8          2023-12-11 [1] CRAN (R 4.3.2)
 yulab.utils        0.1.4          2024-01-28 [1] CRAN (R 4.3.2)
 zlibbioc           1.48.0         2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
