---
title: "01-VMH nnSVG"
author: "Bernard Mulvey"
date: "2023-12-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(SpatialExperiment)
library(spatialLIBD)
library(nnSVG)
library(scater)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

```

### Setup: load data, cluster assignments, and rename from generic to appropriate region (VMH/ARC/WM)
```{r}
hyp2 <- readRDS("data/03-QC_filters/hypN9_umi275_gene166_chrm50_lognorm_111723.RDS")

bscl <- fread("data/06-BayesSpace/02-bayesspace60kiter_k15-20-31_out/BSpace_k15_HARMONYlmbna_nnsvg10.txt")

setnames(bscl,c("rn","cl"))
bscl[,cl:=paste0("X",cl)]
bscl <- DataFrame(bscl)
rownames(bscl) <- bscl$rn
bscl <- bscl[colnames(hyp2),]

colLabels(hyp2) <- factor(bscl$cl,levels=paste0("X",c(1:15)))

### relabel as VMH/ARC/WM for spotplottings
tmpcd <- as.data.table(colData(hyp2),keep.rownames=T)
tmpcd[label %in% c("X11","X12"),label:="VMH"]
tmpcd[label %in% c("X3","X5"),label:="ARC"]
tmpcd[label %in% c("X1","X9"),label:="WM"]

tmpcd <- DataFrame(tmpcd)
rownames(tmpcd) <- tmpcd$rn
tmpcd <- tmpcd[colnames(hyp2),]
colData(hyp2) <- tmpcd
hyp2$label <- factor(hyp2$label,levels=c("VMH","ARC","WM",paste0("X",c(2,4,6:8,10,13:15))))

# clean up
rm(tmpcd,bscl)
gc(full=T)
```

### make subsetted SPE that only includes VMH
```{r}
vmh <- hyp2[,hyp2$label=="VMH"]
### 
sbp <- MulticoreParam(5)
register(sbp)

### iterate nnSVG over each sample, 5 samples at a time x 2 threads per samp
### at least 1 counts in at least 5% of spots
caps <- as.list(unique(colData(vmh)$sample_id))
res_list2 <- bplapply(caps,BPPARAM = sbp,FUN=function(x){
    vmh.samp <- vmh[,colData(vmh)$sample_id==x]
    vmh.samp <- filter_genes(vmh.samp,filter_genes_ncounts = 1,filter_genes_pcspots = 5,filter_mito = F)
    vmh.samp <- computeLibraryFactors(vmh.samp)
    vmh.samp <- logNormCounts(vmh.samp)
    out <- nnSVG(vmh.samp,assay_name = "logcounts",n_threads = 2)
    out
})
names(res_list2) <- caps

res_list <- lapply(res_list2,FUN=function(x){as.data.table(rowData(x))})
res_list <- lapply(res_list,FUN=function(x){setnames(x,c(8:ncol(x)),paste0("nnsvg_",names(x)[8:ncol(x)]))})

saveRDS(res_list,"data/08-Nucleus-specific SVGs/01a-VMH_1count-5pctspot_fulloutput.RDS")
```

### tabulate significant genes per sample
```{r}
i <- 1
for (i in c(1:length(res_list))){
    tmp <- copy(res_list[[i]])
    tmp <- tmp[nnsvg_pval<0.05,.(gene_id,gene_name,nnsvg_rank)]
    setnames(tmp,"nnsvg_rank",names(res_list)[i])
    if(i==1){vmhsvgranks <- copy(tmp)}
    else{vmhsvgranks <- merge.data.table(vmhsvgranks,tmp,by=c("gene_id","gene_name"),all.x=T,all.y=T)}
    rm(tmp)
}

vmhsvgranks[,nasamps:=rowSums(is.na(vmhsvgranks[,c(3:11)]))]
#
vmhsvgranks[,nsigsamps:=9-nasamps]
vmhsvgranks[,meanrank:=rowMeans(vmhsvgranks[,c(3:11)],na.rm=T)]
vmhsvgranks[,nasamps:=NULL]

fwrite(vmhsvgranks,"data/08-Nucleus-specific SVGs/01b-VMH nnSVG nomsig genes and ranks per sample.txt",sep='\t',quote=F)
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.3.1 (2023-06-16)
Platform: aarch64-apple-darwin20 (64-bit)
Running under: macOS Ventura 13.6

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.11.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] scater_1.30.1               ggplot2_3.4.4              
 [3] scuttle_1.12.0              BiocParallel_1.36.0        
 [5] parallelly_1.36.0           default_1.0.0              
 [7] colorout_1.3-0              nnSVG_1.6.0                
 [9] spatialLIBD_1.14.1          SpatialExperiment_1.12.0   
[11] SingleCellExperiment_1.24.0 SummarizedExperiment_1.32.0
[13] Biobase_2.62.0              GenomicRanges_1.54.1       
[15] MatrixGenerics_1.14.0       matrixStats_1.2.0          
[17] Biostrings_2.70.1           GenomeInfoDb_1.38.2        
[19] XVector_0.42.0              IRanges_2.36.0             
[21] S4Vectors_0.40.2            BiocGenerics_0.48.1        
[23] data.table_1.14.10          rlang_1.1.2                

loaded via a namespace (and not attached):
  [1] later_1.3.2                   BiocIO_1.12.0                
  [3] bitops_1.0-7                  filelock_1.0.3               
  [5] fields_15.2                   tibble_3.2.1                 
  [7] XML_3.99-0.16                 lifecycle_1.0.4              
  [9] edgeR_4.0.3                   doParallel_1.0.17            
 [11] rprojroot_2.0.4               lattice_0.22-5               
 [13] magrittr_2.0.3                limma_3.58.1                 
 [15] plotly_4.10.3                 sass_0.4.8                   
 [17] rmarkdown_2.25                jquerylib_0.1.4              
 [19] yaml_2.3.8                    httpuv_1.6.13                
 [21] spam_2.10-0                   sessioninfo_1.2.2            
 [23] pbapply_1.7-2                 cowplot_1.1.2                
 [25] DBI_1.2.0                     RColorBrewer_1.1-3           
 [27] golem_0.4.1                   maps_3.4.2                   
 [29] abind_1.4-5                   zlibbioc_1.48.0              
 [31] purrr_1.0.2                   RCurl_1.98-1.13              
 [33] rappdirs_0.3.3                GenomeInfoDbData_1.2.11      
 [35] ggrepel_0.9.4                 irlba_2.3.5.1                
 [37] DelayedMatrixStats_1.24.0     codetools_0.2-19             
 [39] DelayedArray_0.28.0           DT_0.31                      
 [41] tidyselect_1.2.0              ScaledMatrix_1.10.0          
 [43] viridis_0.6.4                 shinyWidgets_0.8.0           
 [45] BiocFileCache_2.10.1          GenomicAlignments_1.38.0     
 [47] jsonlite_1.8.8                BiocNeighbors_1.20.1         
 [49] ellipsis_0.3.2                iterators_1.0.14             
 [51] foreach_1.5.2                 tools_4.3.1                  
 [53] Rcpp_1.0.11                   glue_1.6.2                   
 [55] gridExtra_2.3                 SparseArray_1.2.2            
 [57] xfun_0.41                     here_1.0.1                   
 [59] dplyr_1.1.4                   withr_2.5.2                  
 [61] BiocManager_1.30.22           fastmap_1.1.1                
 [63] fansi_1.0.6                   digest_0.6.33                
 [65] rsvd_1.0.5                    R6_2.5.1                     
 [67] mime_0.12                     colorspace_2.1-0             
 [69] RSQLite_2.3.4                 config_0.3.2                 
 [71] utf8_1.2.4                    tidyr_1.3.0                  
 [73] generics_0.1.3                rtracklayer_1.62.0           
 [75] httr_1.4.7                    htmlwidgets_1.6.4            
 [77] S4Arrays_1.2.0                scatterplot3d_0.3-44         
 [79] pkgconfig_2.0.3               gtable_0.3.4                 
 [81] rdist_0.0.5                   blob_1.2.4                   
 [83] BRISC_1.0.5                   htmltools_0.5.7              
 [85] dotCall64_1.1-1               scales_1.3.0                 
 [87] png_0.1-8                     attempt_0.3.1                
 [89] knitr_1.45                    rstudioapi_0.15.0            
 [91] rjson_0.2.21                  curl_5.2.0                   
 [93] cachem_1.0.8                  Polychrome_1.5.1             
 [95] BiocVersion_3.18.1            parallel_4.3.1               
 [97] vipor_0.4.7                   AnnotationDbi_1.64.1         
 [99] restfulr_0.0.15               pillar_1.9.0                 
[101] grid_4.3.1                    vctrs_0.6.5                  
[103] RANN_2.6.1                    promises_1.2.1               
[105] BiocSingular_1.18.0           dbplyr_2.4.0                 
[107] beachmat_2.18.0               xtable_1.8-4                 
[109] beeswarm_0.4.0                paletteer_1.5.0              
[111] evaluate_0.23                 magick_2.8.2                 
[113] cli_3.6.2                     locfit_1.5-9.8               
[115] compiler_4.3.1                Rsamtools_2.18.0             
[117] crayon_1.5.2                  rematch2_2.1.2               
[119] ggbeeswarm_0.7.2              viridisLite_0.4.2            
[121] munsell_0.5.0                 lazyeval_0.2.2               
[123] Matrix_1.6-4                  ExperimentHub_2.10.0         
[125] benchmarkme_1.0.8             sparseMatrixStats_1.14.0     
[127] bit64_4.0.5                   KEGGREST_1.42.0              
[129] statmod_1.5.0                 shiny_1.8.0                  
[131] interactiveDisplayBase_1.40.0 AnnotationHub_3.10.0         
[133] memoise_2.0.1                 bslib_0.6.1                  
[135] benchmarkmeData_1.0.4         bit_4.0.5                    
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────
 setting  value
 version  R version 4.3.1 (2023-06-16)
 os       macOS Ventura 13.6
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2023-12-25
 rstudio  2023.09.1+494 Desert Sunflower (desktop)
 pandoc   3.1.1 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────
 package                * version   date (UTC) lib source
 abind                    1.4-5     2016-07-21 [1] CRAN (R 4.3.0)
 AnnotationDbi            1.64.1    2023-11-02 [1] Bioconductor
 AnnotationHub            3.10.0    2023-10-24 [1] Bioconductor
 attempt                  0.3.1     2020-05-03 [1] CRAN (R 4.3.0)
 beachmat                 2.18.0    2023-10-24 [1] Bioconductor
 beeswarm                 0.4.0     2021-06-01 [1] CRAN (R 4.3.0)
 benchmarkme              1.0.8     2022-06-12 [1] CRAN (R 4.3.0)
 benchmarkmeData          1.0.4     2020-04-23 [1] CRAN (R 4.3.0)
 Biobase                * 2.62.0    2023-10-26 [1] Bioconductor
 BiocFileCache            2.10.1    2023-10-26 [1] Bioconductor
 BiocGenerics           * 0.48.1    2023-11-02 [1] Bioconductor
 BiocIO                   1.12.0    2023-10-24 [1] Bioconductor
 BiocManager              1.30.22   2023-08-08 [1] CRAN (R 4.3.0)
 BiocNeighbors            1.20.1    2023-12-19 [1] Bioconductor 3.18 (R 4.3.2)
 BiocParallel           * 1.36.0    2023-10-26 [1] Bioconductor
 BiocSingular             1.18.0    2023-10-24 [1] Bioconductor
 BiocVersion              3.18.1    2023-11-15 [1] Bioconductor
 Biostrings             * 2.70.1    2023-10-26 [1] Bioconductor
 bit                      4.0.5     2022-11-15 [1] CRAN (R 4.3.0)
 bit64                    4.0.5     2020-08-30 [1] CRAN (R 4.3.0)
 bitops                   1.0-7     2021-04-24 [1] CRAN (R 4.3.0)
 blob                     1.2.4     2023-03-17 [1] CRAN (R 4.3.0)
 BRISC                    1.0.5     2022-04-29 [1] CRAN (R 4.3.0)
 bslib                    0.6.1     2023-11-28 [1] CRAN (R 4.3.1)
 cachem                   1.0.8     2023-05-01 [1] CRAN (R 4.3.0)
 cli                      3.6.2     2023-12-11 [1] CRAN (R 4.3.1)
 codetools                0.2-19    2023-02-01 [1] CRAN (R 4.3.0)
 colorout               * 1.3-0     2023-11-01 [1] local (/Users/bmulvey/Desktop/colorout)
 colorspace               2.1-0     2023-01-23 [1] CRAN (R 4.3.0)
 config                   0.3.2     2023-08-30 [1] CRAN (R 4.3.0)
 cowplot                  1.1.2     2023-12-15 [1] CRAN (R 4.3.1)
 crayon                   1.5.2     2022-09-29 [1] CRAN (R 4.3.0)
 curl                     5.2.0     2023-12-08 [1] CRAN (R 4.3.1)
 data.table             * 1.14.10   2023-12-08 [1] CRAN (R 4.3.1)
 DBI                      1.2.0     2023-12-21 [1] CRAN (R 4.3.1)
 dbplyr                   2.4.0     2023-10-26 [1] CRAN (R 4.3.1)
 default                * 1.0.0     2017-08-07 [1] CRAN (R 4.3.0)
 DelayedArray             0.28.0    2023-11-06 [1] Bioconductor
 DelayedMatrixStats       1.24.0    2023-10-24 [1] Bioconductor
 digest                   0.6.33    2023-07-07 [1] CRAN (R 4.3.0)
 doParallel               1.0.17    2022-02-07 [1] CRAN (R 4.3.0)
 dotCall64                1.1-1     2023-11-28 [1] CRAN (R 4.3.1)
 dplyr                    1.1.4     2023-11-17 [1] CRAN (R 4.3.1)
 DT                       0.31      2023-12-09 [1] CRAN (R 4.3.1)
 edgeR                    4.0.3     2023-12-09 [1] Bioconductor 3.18 (R 4.3.2)
 ellipsis                 0.3.2     2021-04-29 [1] CRAN (R 4.3.0)
 evaluate                 0.23      2023-11-01 [1] CRAN (R 4.3.1)
 ExperimentHub            2.10.0    2023-10-24 [1] Bioconductor
 fansi                    1.0.6     2023-12-08 [1] CRAN (R 4.3.1)
 fastmap                  1.1.1     2023-02-24 [1] CRAN (R 4.3.0)
 fields                   15.2      2023-08-17 [1] CRAN (R 4.3.0)
 filelock                 1.0.3     2023-12-11 [1] CRAN (R 4.3.1)
 foreach                  1.5.2     2022-02-02 [1] CRAN (R 4.3.0)
 generics                 0.1.3     2022-07-05 [1] CRAN (R 4.3.0)
 GenomeInfoDb           * 1.38.2    2023-12-13 [1] Bioconductor 3.18 (R 4.3.1)
 GenomeInfoDbData         1.2.11    2023-11-10 [1] Bioconductor
 GenomicAlignments        1.38.0    2023-10-24 [1] Bioconductor
 GenomicRanges          * 1.54.1    2023-10-30 [1] Bioconductor
 ggbeeswarm               0.7.2     2023-04-29 [1] CRAN (R 4.3.0)
 ggplot2                * 3.4.4     2023-10-12 [1] CRAN (R 4.3.1)
 ggrepel                  0.9.4     2023-10-13 [1] CRAN (R 4.3.1)
 glue                     1.6.2     2022-02-24 [1] CRAN (R 4.3.0)
 golem                    0.4.1     2023-06-05 [1] CRAN (R 4.3.0)
 gridExtra                2.3       2017-09-09 [1] CRAN (R 4.3.0)
 gtable                   0.3.4     2023-08-21 [1] CRAN (R 4.3.0)
 here                     1.0.1     2020-12-13 [1] CRAN (R 4.3.0)
 htmltools                0.5.7     2023-11-03 [1] CRAN (R 4.3.1)
 htmlwidgets              1.6.4     2023-12-06 [1] CRAN (R 4.3.1)
 httpuv                   1.6.13    2023-12-06 [1] CRAN (R 4.3.1)
 httr                     1.4.7     2023-08-15 [1] CRAN (R 4.3.0)
 interactiveDisplayBase   1.40.0    2023-10-24 [1] Bioconductor
 IRanges                * 2.36.0    2023-10-26 [1] Bioconductor
 irlba                    2.3.5.1   2022-10-03 [1] CRAN (R 4.3.1)
 iterators                1.0.14    2022-02-05 [1] CRAN (R 4.3.0)
 jquerylib                0.1.4     2021-04-26 [1] CRAN (R 4.3.0)
 jsonlite                 1.8.8     2023-12-04 [1] CRAN (R 4.3.1)
 KEGGREST                 1.42.0    2023-10-26 [1] Bioconductor
 knitr                    1.45      2023-10-30 [1] CRAN (R 4.3.1)
 later                    1.3.2     2023-12-06 [1] CRAN (R 4.3.1)
 lattice                  0.22-5    2023-10-24 [1] CRAN (R 4.3.1)
 lazyeval                 0.2.2     2019-03-15 [1] CRAN (R 4.3.0)
 lifecycle                1.0.4     2023-11-07 [1] CRAN (R 4.3.1)
 limma                    3.58.1    2023-10-31 [1] Bioconductor
 locfit                   1.5-9.8   2023-06-11 [1] CRAN (R 4.3.0)
 magick                   2.8.2     2023-12-20 [1] CRAN (R 4.3.1)
 magrittr                 2.0.3     2022-03-30 [1] CRAN (R 4.3.0)
 maps                     3.4.2     2023-12-15 [1] CRAN (R 4.3.1)
 Matrix                   1.6-4     2023-11-30 [1] CRAN (R 4.3.1)
 MatrixGenerics         * 1.14.0    2023-10-26 [1] Bioconductor
 matrixStats            * 1.2.0     2023-12-11 [1] CRAN (R 4.3.1)
 memoise                  2.0.1     2021-11-26 [1] CRAN (R 4.3.0)
 mime                     0.12      2021-09-28 [1] CRAN (R 4.3.0)
 munsell                  0.5.0     2018-06-12 [1] CRAN (R 4.3.0)
 nnSVG                  * 1.6.0     2023-10-24 [1] Bioconductor
 paletteer                1.5.0     2022-10-19 [1] CRAN (R 4.3.0)
 parallelly             * 1.36.0    2023-05-26 [1] CRAN (R 4.3.0)
 pbapply                  1.7-2     2023-06-27 [1] CRAN (R 4.3.0)
 pillar                   1.9.0     2023-03-22 [1] CRAN (R 4.3.0)
 pkgconfig                2.0.3     2019-09-22 [1] CRAN (R 4.3.0)
 plotly                   4.10.3    2023-10-21 [1] CRAN (R 4.3.1)
 png                      0.1-8     2022-11-29 [1] CRAN (R 4.3.0)
 Polychrome               1.5.1     2022-05-03 [1] CRAN (R 4.3.0)
 promises                 1.2.1     2023-08-10 [1] CRAN (R 4.3.0)
 purrr                    1.0.2     2023-08-10 [1] CRAN (R 4.3.0)
 R6                       2.5.1     2021-08-19 [1] CRAN (R 4.3.0)
 RANN                     2.6.1     2019-01-08 [1] CRAN (R 4.3.0)
 rappdirs                 0.3.3     2021-01-31 [1] CRAN (R 4.3.0)
 RColorBrewer             1.1-3     2022-04-03 [1] CRAN (R 4.3.0)
 Rcpp                     1.0.11    2023-07-06 [1] CRAN (R 4.3.1)
 RCurl                    1.98-1.13 2023-11-02 [1] CRAN (R 4.3.1)
 rdist                    0.0.5     2020-05-04 [1] CRAN (R 4.3.0)
 rematch2                 2.1.2     2020-05-01 [1] CRAN (R 4.3.0)
 restfulr                 0.0.15    2022-06-16 [1] CRAN (R 4.3.0)
 rjson                    0.2.21    2022-01-09 [1] CRAN (R 4.3.0)
 rlang                  * 1.1.2     2023-11-04 [1] CRAN (R 4.3.1)
 rmarkdown                2.25      2023-09-18 [1] CRAN (R 4.3.1)
 rprojroot                2.0.4     2023-11-05 [1] CRAN (R 4.3.1)
 Rsamtools                2.18.0    2023-10-24 [1] Bioconductor
 RSQLite                  2.3.4     2023-12-08 [1] CRAN (R 4.3.1)
 rstudioapi               0.15.0    2023-07-07 [1] CRAN (R 4.3.0)
 rsvd                     1.0.5     2021-04-16 [1] CRAN (R 4.3.0)
 rtracklayer              1.62.0    2023-10-26 [1] Bioconductor
 S4Arrays                 1.2.0     2023-10-26 [1] Bioconductor
 S4Vectors              * 0.40.2    2023-11-25 [1] Bioconductor 3.18 (R 4.3.2)
 sass                     0.4.8     2023-12-06 [1] CRAN (R 4.3.1)
 ScaledMatrix             1.10.0    2023-10-24 [1] Bioconductor
 scales                   1.3.0     2023-11-28 [1] CRAN (R 4.3.1)
 scater                 * 1.30.1    2023-11-16 [1] Bioconductor
 scatterplot3d            0.3-44    2023-05-05 [1] CRAN (R 4.3.0)
 scuttle                * 1.12.0    2023-10-24 [1] Bioconductor
 sessioninfo              1.2.2     2021-12-06 [1] CRAN (R 4.3.0)
 shiny                    1.8.0     2023-11-17 [1] CRAN (R 4.3.1)
 shinyWidgets             0.8.0     2023-08-30 [1] CRAN (R 4.3.0)
 SingleCellExperiment   * 1.24.0    2023-11-06 [1] Bioconductor
 spam                     2.10-0    2023-10-23 [1] CRAN (R 4.3.1)
 SparseArray              1.2.2     2023-11-08 [1] Bioconductor
 sparseMatrixStats        1.14.0    2023-10-24 [1] Bioconductor
 SpatialExperiment      * 1.12.0    2023-10-24 [1] Bioconductor
 spatialLIBD            * 1.14.1    2023-11-30 [1] Bioconductor 3.18 (R 4.3.1)
 statmod                  1.5.0     2023-01-06 [1] CRAN (R 4.3.0)
 SummarizedExperiment   * 1.32.0    2023-11-06 [1] Bioconductor
 tibble                   3.2.1     2023-03-20 [1] CRAN (R 4.3.0)
 tidyr                    1.3.0     2023-01-24 [1] CRAN (R 4.3.0)
 tidyselect               1.2.0     2022-10-10 [1] CRAN (R 4.3.0)
 utf8                     1.2.4     2023-10-22 [1] CRAN (R 4.3.1)
 vctrs                    0.6.5     2023-12-01 [1] CRAN (R 4.3.1)
 vipor                    0.4.7     2023-12-18 [1] CRAN (R 4.3.1)
 viridis                  0.6.4     2023-07-22 [1] CRAN (R 4.3.0)
 viridisLite              0.4.2     2023-05-02 [1] CRAN (R 4.3.0)
 withr                    2.5.2     2023-10-30 [1] CRAN (R 4.3.1)
 xfun                     0.41      2023-11-01 [1] CRAN (R 4.3.1)
 XML                      3.99-0.16 2023-11-29 [1] CRAN (R 4.3.1)
 xtable                   1.8-4     2019-04-21 [1] CRAN (R 4.3.0)
 XVector                * 0.42.0    2023-10-26 [1] Bioconductor
 yaml                     2.3.8     2023-12-11 [1] CRAN (R 4.3.1)
 zlibbioc                 1.48.0    2023-10-26 [1] Bioconductor

 [1] /Users/bmulvey/Library/R/arm64/4.3/library
 [2] /Library/Frameworks/R.framework/Versions/4.3-arm64/Resources/library

──────────────────────────────────────────────────────────────────────────
