---
title: "04-Output Region-Specific nnSVG Results for Supp Tabs"
author: "Bernard Mulvey"
date: "2024-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)
```

read in data, write out as tsv
```{r}
vmhsvg <- readRDS("processed-data/08-Nucleus-specific SVGs/01a-VMH_1count-5pctspot_fulloutput.RDS")

arcsvg <- readRDS("processed-data/08-Nucleus-specific SVGs/02a-ARC_1count-5pctspot_fulloutput.RDS")

i<-1
for (i in c(1:length(vmhsvg))){
    tmp <- vmhsvg[[i]]
    getnames <- c("gene_id","gene_name",grep(names(tmp),pattern="^nnsvg",value=T))
    tmp <- tmp[,..getnames]
    
    tmp[,sample:=names(vmhsvg)[i]]

    if (i==1){
        vmhtab <- tmp
    } else {
        vmhtab <- rbind(vmhtab,tmp)
    }
}

i<-1
for (i in c(1:length(arcsvg))){
    tmp <- arcsvg[[i]]
    getnames <- c("gene_id","gene_name",grep(names(tmp),pattern="^nnsvg",value=T))
    tmp <- tmp[,..getnames]
    
    tmp[,sample:=names(vmhsvg)[i]]

    if (i==1){
        arctab <- tmp
    } else {
        arctab <- rbind(arctab,tmp)
    }
}

fwrite(vmhtab,"local_supptab_assembly/VMHonly_nnsvg_raw.txt",sep='\t',quote=F)
fwrite(arctab,"local_supptab_assembly/ARConly_nnsvg_raw.txt",sep='\t',quote=F)

# clone aggregated results to supptab folder
vmhsum <- fread("processed-data/08-Nucleus-specific SVGs/01b-VMH nnSVG nomsig genes and ranks per sample.txt")
arcsum <- fread("processed-data/08-Nucleus-specific SVGs/02b-ARC nnSVG nomsig genes and ranks per sample.txt")

fwrite(vmhsum,"local_supptab_assembly/VMHonly_nnsvg_summarized.txt",sep='\t',quote=F)
fwrite(arcsum,"local_supptab_assembly/ARConly_nnsvg_summarized.txt",sep='\t',quote=F)

rm(list=ls())
```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] default_1.0.0               BiocParallel_1.38.0        
 [3] parallelly_1.38.0           colorout_1.3-0.2           
 [5] scater_1.32.0               ggplot2_3.5.1              
 [7] scuttle_1.14.0              nnSVG_1.8.0                
 [9] SpatialExperiment_1.14.0    SingleCellExperiment_1.26.0
[11] SummarizedExperiment_1.34.0 Biobase_2.64.0             
[13] GenomicRanges_1.56.1        GenomeInfoDb_1.40.1        
[15] IRanges_2.38.1              S4Vectors_0.42.1           
[17] BiocGenerics_0.50.0         MatrixGenerics_1.16.0      
[19] matrixStats_1.3.0           data.table_1.15.4          
[21] rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] tidyselect_1.2.1          viridisLite_0.4.2         dplyr_1.1.4              
 [4] vipor_0.4.7               BRISC_1.0.5               viridis_0.6.5            
 [7] fastmap_1.2.0             RANN_2.6.1                digest_0.6.36            
[10] rsvd_1.0.5                lifecycle_1.0.4           magrittr_2.0.3           
[13] compiler_4.4.1            tools_4.4.1               utf8_1.2.4               
[16] yaml_2.3.9                knitr_1.48                S4Arrays_1.4.1           
[19] here_1.0.1                DelayedArray_0.30.1       rdist_0.0.5              
[22] abind_1.4-5               withr_3.0.0               grid_4.4.1               
[25] fansi_1.0.6               beachmat_2.20.0           colorspace_2.1-0         
[28] scales_1.3.0              cli_3.6.3                 rmarkdown_2.27           
[31] crayon_1.5.3              generics_0.1.3            rstudioapi_0.16.0        
[34] httr_1.4.7                rjson_0.2.21              sessioninfo_1.2.2        
[37] DelayedMatrixStats_1.26.0 pbapply_1.7-2             ggbeeswarm_0.7.2         
[40] zlibbioc_1.50.0           XVector_0.44.0            vctrs_0.6.5              
[43] Matrix_1.7-0              jsonlite_1.8.8            BiocSingular_1.20.0      
[46] BiocNeighbors_1.22.0      ggrepel_0.9.5             irlba_2.3.5.1            
[49] beeswarm_0.4.0            magick_2.8.3              glue_1.7.0               
[52] codetools_0.2-20          gtable_0.3.5              UCSC.utils_1.0.0         
[55] ScaledMatrix_1.12.0       munsell_0.5.1             tibble_3.2.1             
[58] pillar_1.9.0              htmltools_0.5.8.1         GenomeInfoDbData_1.2.12  
[61] R6_2.5.1                  sparseMatrixStats_1.16.0  rprojroot_2.0.4          
[64] evaluate_0.24.0           lattice_0.22-6            Rcpp_1.0.13              
[67] gridExtra_2.3             SparseArray_1.4.8         xfun_0.45                
[70] pkgconfig_2.0.3          
> sessioninfo::session_info()
─ Session info ───────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-30
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ───────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                  1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   beachmat               2.20.0  2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm               0.4.0   2021-06-01 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors          1.22.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular           1.20.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BRISC                  1.0.5   2022-04-29 [1] CRAN (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   default              * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats     1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi                  1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm             0.7.2   2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                0.9.5   2024-01-10 [1] CRAN (R 4.4.0)
   glue                   1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra              2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                  2.3.5.1 2022-10-03 [1] CRAN (R 4.4.0)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   nnSVG                * 1.8.0   2024-05-16 [1] Bioconductor 3.19 (R 4.4.0)
   parallelly           * 1.38.0  2024-07-27 [1] CRAN (R 4.4.0)
   pbapply                1.7-2   2023-06-27 [1] CRAN (R 4.4.0)
   pillar                 1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   RANN                   2.6.1   2019-01-08 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13  2024-07-17 [1] CRAN (R 4.4.0)
   rdist                  0.0.5   2020-05-04 [1] CRAN (R 4.4.0)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                   1.0.5   2021-04-16 [1] CRAN (R 4.4.0)
   S4Arrays               1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   ScaledMatrix           1.12.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   scater               * 1.32.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle              * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray            1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats      1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   vipor                  0.4.7   2023-12-18 [1] CRAN (R 4.4.0)
   viridis                0.6.5   2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite            0.4.2   2023-05-02 [1] CRAN (R 4.4.0)
   withr                  3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector                0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

──────────────────────────────────────────────────────────────────────────────────
