---
title: "02-pair resegmented nuclei to prior segs, plot nucleus-cell pair gene expr per sample.Rmd"
author: "Bernard Mulvey"
date: "2024-06-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Voyager)
library(patchwork)
library(SpatialFeatureExperiment)
library(SingleCellExperiment)
library(SpatialExperiment)
library(ggplot2)
library(gridExtra)
library(stringr)
library(scuttle)
library(BiocSingular)
library(scater)
library(Matrix)
library(DropletUtils)
library(vroom)
library(sf)
library(here)

library(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## github copilot chat in the rstudio Viewer pane
library(chattr) # development branch @copilot installed 031324
# call chattr internal to disable debug mode so that this works
chattr:::ch_debug_set_false()
chattr_use("copilot")
chattr_defaults(model = "copilot",
                type = "chat",
                max_data_files = 20,
                max_data_frames = 20,
                include_doc_contents = TRUE,
                include_history = TRUE,
                prompt = "Use the 'R for Data Science' (https://r4ds.had.co.nz/) book as main reference
                        You can provide information about statistical methods and appropriateness of different statistical tests using material from https://compgenomr.github.io/book/, http://www.cookbook-r.com/, https://bookdown.org/rdpeng/advstatcomp/
                            Provide answers using data.table syntax as documented in https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html and  http://brooksandrew.github.io/simpleblog/articles/advanced-data-table/
                            Use packages: data.table, ggplot2, magrittr
                        Provide explanations with code
                        For code output, use RMarkdown code chunks
                        Avoid all code chunk options")
chattr::chattr_app("viewer",as_job=T,as_job_port=sample(c(4101:4199),size=1,replace=F))

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc on M processor macs
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
Sys.setenv("OBJC_DISABLE_INITIALIZE_FORK_SAFETY"="YES")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### Define Xenium data importer to spatialfeatureexperiment format (courtesy Cindy Fang)
### edit this function to calculate log counts, because trying to do so once in spatialfeatureexperiment won't work
```{r}
# Reference: https://github.com/pachterlab/SFEData/blob/main/inst/scripts/make-data.R

# Read in Xenium data and create an SFE object
# give this function a different name than readXenium, which is imported 
# from SpatialFeatureExperiment
readXenium_manual <- function(dir_name){

    counts_path <- here(dir_name, "cell_feature_matrix.h5")
    cell_info_path <- here(dir_name, "cells.csv.gz")
    cell_poly_path <- here(dir_name, "cell_boundaries.csv.gz")
    nuc_poly_path <- here(dir_name, "nucleus_boundaries.csv.gz")
    # 
    # Read in the data
    sce <- read10xCounts(counts_path)
    counts(sce) <- as(realize(counts(sce)), "dgCMatrix")
    cell_info <- vroom(cell_info_path)
    
    print("Read in files")
    
    cell_poly <- vroom(cell_poly_path)
    nuc_poly <- vroom(nuc_poly_path)
    
    names(cell_poly)[1] <- "ID"
    names(nuc_poly)[1] <- "ID"
    
    # change df to sf object for cell/nuc images
    cells_sf <- df2sf(cell_poly, c("vertex_x", "vertex_y"), geometryType = "POLYGON")
    nuc_sf <- df2sf(nuc_poly, c("vertex_x", "vertex_y"), geometryType = "POLYGON")
    
    # QC check
    all(st_is_valid(cells_sf))
    all(st_is_valid(nuc_sf))
    
    # get rid if invalid cells/nucs
    ind_invalid <- !st_is_valid(nuc_sf)
    nuc_sf[ind_invalid,] <- nngeo::st_remove_holes(st_buffer(nuc_sf[ind_invalid,], 0))
    
    # add cell info
    colData(sce) <- cbind(colData(sce), cell_info)
    print(cell_info)
    
    # make spatial objects
    spe <- toSpatialExperiment(sce, spatialCoordsNames = c("x_centroid", "y_centroid"))
    
    ###  log normalize the counts data, mostly just so that the plot doesn't look like its comprised of just a handful of points
    ### and to do that, we'll need to drop zero-count cells
    spe <- spe[,colSums(counts(spe))>0]
    spe <- scater::computeLibraryFactors(spe)
    spe <- scater::logNormCounts(spe)
    sfe <- toSpatialFeatureExperiment(spe)
    
    # DONT add segmented cells/nuc to spatial object-something's weird with some resulting SFEs from the 2024 slides (processed in xeniumranger1.9) coming b ack with cell and nucleus geoms in colGeometries, but others only coming back with centroids. we don't need these for the moment anyhow (we'll work with them manually later in the script)
    # cellSeg(sfe, withDimnames = FALSE) <- cells_sf
    # nucSeg(sfe, withDimnames = FALSE) <- nuc_sf
    
    
    # Add some QC Metrics
    colData(sfe)$nCounts <- colSums(counts(sfe))
    colData(sfe)$nGenes <- colSums(counts(sfe) > 0)
    
    is_blank <- str_detect(rownames(sfe), "^BLANK_")
    is_neg <- str_detect(rownames(sfe), "^NegControlProbe")
    is_neg2 <- str_detect(rownames(sfe), "^NegControlCodeword")
    is_anti <- str_detect(rownames(sfe), "^antisense")
    is_depr <- str_detect(rownames(sfe), "^DeprecatedCodeword")
    is_unassigned <- str_detect(rownames(sfe), "^Unassigned")
    
    is_any_neg <- is_blank | is_neg | is_neg2 | is_anti | is_depr | is_unassigned
    rowData(sfe)$is_neg <- is_any_neg
    
    n_panel <- nrow(sfe) - sum(is_any_neg)
    #print(n_panel)
    
    # normalize counts after QC
    colData(sfe)$nCounts_normed <- sfe$nCounts/n_panel
    colData(sfe)$nGenes_normed <- sfe$nGenes/n_panel
    colData(sfe)$prop_nuc <- sfe$nucleus_area / sfe$cell_area
    
    # add QC columns
    sfe <- addPerCellQCMetrics(sfe, subsets = list(blank = is_blank,
    negProbe = is_neg,
    negCodeword = is_neg2,
    anti = is_anti,
    depr = is_depr,
    unassigned = is_unassigned,
    any_neg = is_any_neg))

    # add features
    rowData(sfe)$means <- rowMeans(counts(sfe))
    rowData(sfe)$vars <- rowVars(counts(sfe))
    rowData(sfe)$cv2 <- rowData(sfe)$vars/rowData(sfe)$means^2

    
    # Add cell ids and make gene names unique
    #colnames(sfe) <- seq_len(ncol(sfe))
    rownames(sfe) <- uniquifyFeatureNames(ID=rownames(sfe),  names=rowData(sfe)$Symbol)
    
    return(sfe)
}
```

### load resegmentations and original segmentations as separate SpatialFeatureExperiment objects to start with
```{r}
resegs <- list.files("processed-data/01_xeniumranger1.7-resegment",pattern="03",full.names = T)
resegsampname <- gsub(resegs,pattern="^.*1.7-resegment/03_(.*)_wnuclei$",replacement="\\1")
resegs <- paste0(resegs,"/outs")

# outs directory files were previously placed directly inside the original samplewise directories
origsegs <- list.files("raw-data/xenium/20240223__174743__022324_KMon120823",full.names=T)
origsegsampname <- gsub(origsegs,pattern="^.*output-(.*)$",replacement="\\1")
origsegsampname <- gsub(origsegsampname,pattern="^XETG.*__00[[:digit:]]{3}(..)__[[:digit:]]{5}X_(.*)_HYP.*$",replacement=paste0("\\1","_","\\2","_","origseg"))
i<-1
for (i in c(1:length(origsegs))){
    if(i==1){
        resegsamps <- readXenium_manual(resegs[i])
        origsamps <- readXenium_manual(origsegs[i])
        
        resegsamps$Sample <- resegsampname[i]
        resegsamps$sample_id <- resegsampname[i]
        
        origsamps$Sample <- origsegsampname[i]
        origsamps$sample_id <- origsegsampname[i]
        # make unique colnames
        resegsamps$key <- paste0(resegsamps$Sample,"_",resegsamps$cell_id)
        colnames(resegsamps) <- resegsamps$key
        
        origsamps$key <- paste0(origsamps$Sample,"_",origsamps$cell_id)
        colnames(origsamps) <- origsamps$key
        
        # make unique colnames in rowdata
        colnames(rowData(origsamps))[c(5:7)] <- paste0(origsegsampname[i],"_",colnames(rowData(origsamps))[5:7]) 
        
        colnames(rowData(resegsamps))[c(5:7)] <- paste0(resegsampname[i],"_",colnames(rowData(resegsamps))[5:7]) 

    }
    else{
        tmpresegsamp <- readXenium_manual(resegs[i])
        tmpresegsamp$Sample <- resegsampname[i]
        tmpresegsamp$sample_id <- resegsampname[i]
        tmpresegsamp$key <- paste0(tmpresegsamp$Sample,"_",tmpresegsamp$cell_id)
        colnames(tmpresegsamp) <- tmpresegsamp$key
        
        ## remove the gene identifier/etc information here since we got it when i==1; make remaining colnames sample-specific 
        rowData(tmpresegsamp) <- rowData(tmpresegsamp)[,c(5:7)]
        colnames(rowData(tmpresegsamp)) <- paste0(resegsampname[i],"_",colnames(rowData(tmpresegsamp)))
        
        resegsamps <- cbind(resegsamps,tmpresegsamp)
        
        tmporigsamp <- readXenium_manual(origsegs[i])
        tmporigsamp$Sample <- origsegsampname[i]
        tmporigsamp$sample_id <- origsegsampname[i]
        tmporigsamp$key <- paste0(tmporigsamp$Sample,"_",tmporigsamp$cell_id)
        colnames(tmporigsamp) <- tmporigsamp$key
        
        ## remove the gene identifier/etc information here since we got it when i==1; make remaining colnames sample-specific 
        rowData(tmporigsamp) <- rowData(tmporigsamp)[,c(5:7)]
        colnames(rowData(tmporigsamp)) <- paste0(origsegsampname[i],"_",colnames(rowData(tmporigsamp)))
        
        origsamps <- cbind(origsamps,tmporigsamp)
        
        rm(tmpresegsamp,tmporigsamp)
    }
}
# cleanup
rm(i)
gc(full=T)
```


### getting the data paired back up for plotting here may not be easy, since we resegmented nuclei.

# by coercing from SPE to SFE in the custom loader defined above, we're preventing return of both nuclear and cell centroids (and seems like from SFE documentation what SPE -> SFE returns for centroids are cell, not nuc). so we need to circle back and pull in the nuclear boundaries + calc centroids for both original and resegmented nuclei.
conveniently, we make a lookup table between original and reseg nucleus (==cell) identifiers doing things this way!

```{r}
bpp <- MulticoreParam(6)
## get nuclear boundaries from v1.5 and calculate centroids
orignuccentroids <- bplapply(origsegs,FUN=function(x){
    nuc <- fread(paste0(x,"/nucleus_boundaries.csv.gz"))
    nuc[,centroidx:=mean(vertex_x),by="cell_id"]
    nuc[,centroidy:=mean(vertex_y),by="cell_id"]
    nuc <- unique(nuc[,.(cell_id,centroidx,centroidy)])
    return(nuc)
},BPPARAM=bpp)
names(orignuccentroids) <- gsub(origsegsampname,pattern="^(.*)_origseg$",replacement="\\1")

## get nuclear boundaries from v1.7 and calculate centroids
resegnuccentroids <- bplapply(resegs,FUN=function(x){
    nuc <- fread(paste0(x,"/nucleus_boundaries.csv.gz"))
    nuc[,centroidx:=mean(vertex_x),by="cell_id"]
    nuc[,centroidy:=mean(vertex_y),by="cell_id"]
    nuc <- unique(nuc[,.(cell_id,centroidx,centroidy)])
    return(nuc)
},BPPARAM=bpp)
names(resegnuccentroids) <- gsub(resegsampname,pattern="^(.*)_.*_reseg$",replacement="\\1")

## make sure both lists are in the same order for mapply
resegnuccentroids <- resegnuccentroids[names(orignuccentroids)]

## merge to subset to nuclei with the same centroid in 1.5 and 1.7, and to get their orig/reseg nucleus (==cell) identifiers
nuclut <- bpmapply(o=orignuccentroids,r=resegnuccentroids,SIMPLIFY = FALSE,BPPARAM = bpp,FUN=function(o,r){
    out <- merge.data.table(o,r,by=c("centroidx","centroidy"))
    setnames(out,c("cell_id.x","cell_id.y"),c("orig_cellid","reseg_cellid"))
    return(out)
})
names(nuclut) <- names(resegnuccentroids)

# check how many nuclei were in the original, reseg, and centroid-merged tables
# mapply(X=orignuccentroids,Y=resegnuccentroids,Z=nuclut,FUN=function(X,Y,Z){return(c(nrow(X),nrow(Y),nrow(Z)))})
rm(resegnuccentroids,orignuccentroids,bpp)
gc(full=T)

# now concatenate the nuclear centroids info into one big table to map between original and resegmented nuclei for plotting
nuclut2 <- rbindlist(nuclut,idcol = "samp")
rm(nuclut)
```
# >500k nuclei have the same centroids after either segmentation algo in every sample.

now we can proceed with getting the logcounts matrices into a format for plotting
```{r}
origcounts <- as.data.table(as.data.frame(as.matrix(logcounts(origsamps))),keep.rownames=T)

resegcounts <- as.data.table(as.data.frame(as.matrix(logcounts(resegsamps))),keep.rownames=T)

## melt so that each cell-gene-count value from a dataset is a row, i.e. so can have an x or y coord value for plotting
origcounts.m <- melt(origcounts,id.vars=c("rn"))
resegcounts.m <- melt(resegcounts,id.vars=c("rn"))

## make column names informative
setnames(origcounts.m,c("gene","cell","orig_value"))
setnames(resegcounts.m,c("gene","cell","reseg_value"))

## extract constant sample, cell identifiers for merging
origcounts.m[,sample:=gsub(cell,pattern="^(.*)_origseg.*$",replacement="\\1")]
origcounts.m[,cell:=gsub(cell,pattern="^.*origseg_(.*)$",replacement="\\1")]

resegcounts.m[,sample:=gsub(cell,pattern="^(.*)_xr17.*$",replacement="\\1")]
resegcounts.m[,cell:=gsub(cell,pattern="^.*_xr17_reseg_(.*)$",replacement="\\1")]

## append identical-centroid-to-nucleus/cell-id lookup table to both of these so we can match nuclei up between the two runs for plotting
origcounts.m <- merge.data.table(nuclut2,origcounts.m,by.x="orig_cellid",by.y="cell")

resegcounts.m <- merge.data.table(nuclut2,resegcounts.m,by.x="reseg_cellid",by.y="cell")

### remove excess columns
origcounts.m <- origcounts.m[,.(sample,orig_cellid,gene,orig_value)]
resegcounts.m <- resegcounts.m[,.(sample,orig_cellid,reseg_cellid,gene,reseg_value)]

### remove some old intermediate objects to conserve mem
rm(origcounts,resegcounts)
gc(full=T)

## merge to one table for plotting. we only need the columns from the nucleus id lookup table once (besides orig_cellid, which we'll use to merge by)
comparecounts <- merge.data.table(origcounts.m,resegcounts.m,by=c("sample","orig_cellid","gene"))

## ^ this table ^ should be > 100 million rows, since most nuclei (and thus parent cells) can be mapped between datasets (using the same nuclear segmentations with 1.5 and 1.7 with one fewer samples than we have here gave us a ~229 million row table at this step)
stopifnot(nrow(comparecounts)>10^8 & nrow(comparecounts)<10^9)
# spot check a row to make sure we got the correct table structure (e.g., cell aaaaaaa-1 counts of AABC1 in sample 86_reg1 are in a row)
comparecounts[1,]
## cleanup workspace/conserve memory (just about everything besides the logcount-logcount table)
rm(origcounts.m,resegcounts.m,nclut2,i,origsegs,origsegsampname,resegs,resegsampname,nuclut2,i,origsamps,resegsamps)
gc(full=T)
```

add some gene info (base panel vs. custom genes)
```{r}
genepanel <- fread("raw-data/humanbrain_basepanel_genes.txt")
comparecounts <- merge.data.table(comparecounts,genepanel,by.x="gene",by.y="Gene",all.x=T)
setnames(comparecounts,"Ensembl ID","ensembl")
comparecounts[!is.na(ensembl),gene_set:="Brain Base Panel"]
comparecounts[gene %in% grep(unique(comparecounts$gene),pattern="BLANK",value=T),gene_set:="BLANK"]
comparecounts[gene %in% grep(unique(comparecounts$gene),pattern="NegControl",value=T),gene_set:="Neg Ctrl"]
comparecounts[gene %in% grep(unique(comparecounts$gene),pattern="Deprecated",value=T),gene_set:="Deprecated Codeword"]
comparecounts[is.na(gene_set),gene_set:="HYP Custom Panel"]
```

### check whether any values changed in nuclei with identical centroids
```{r}
nrow(comparecounts[orig_value!=reseg_value])/nrow(comparecounts)
```
## ^ Yes, 18.03% across the 2024 dataset

### plot from here instead of saving this 220 million row table to load into a separate plotting script
### one pdf = one sample, one page = one gene
```{r}
samps <- as.list(unique(comparecounts$sample))

### divy up data before bplapply so each worker only loads relevant data
singsampdat <- lapply(samps,FUN=function(x){comparecounts[sample==x]})
names(singsampdat) <- unlist(samps)
rm(samps)

### Make a 5 color palette so points are colored by gene type
pals <- Polychrome::createPalette(5,seedcolors = "#43AE30")
pals <- as.data.table(cbind(unique(comparecounts$gene_set),pals))
setnames(pals,c("gene_set","pal"))

## delete the all-samples table (since we have a copy of it split by sample now for parallelized plotting)
rm(comparecounts)
gc(full=T)
```

### first, make giant pngs showing all genes in all cells per sample
```{r}
plts <- lapply(singsampdat,FUN=function(d){
  ggplot(d,aes(x=orig_value,y=reseg_value,col=gene_set))+
    geom_point(size = 1.75)+
    ggtitle(paste0(unique(d$sample),": logcounts"))+
    xlim(0,max(c(d$orig_value,d$reseg_value)))+
    ylim(0,max(c(d$orig_value,d$reseg_value)))+
    geom_abline(intercept = 0,slope = 2,linetype="dashed",col="black",alpha=0.4)+
    xlab("v1.9 cells: w/in samp norm logcounts")+
    ylab("v1.7 cells: w/in samp norm logcounts")+
    theme(axis.text.x = element_text(size = 16), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=14), legend.title = element_text(size=14,hjust=0.5))
})
names(plts) <- names(singsampdat)

mapply(X=plts,Y=names(plts),SIMPLIFY = F,FUN = function(X,Y){
  png(paste0("~/Desktop/03-",Y,".png"),width=8.5,height=7.5,units="in",res=150)
  print(X)
  dev.off()
})

```

### reset biocparallel default to avoid errors (multicore above, snowparam here)
### output plots in parallel
```{r}
register(SerialParam())

## restart biocparallel as snowparam
# bpp <- SnowParam(3)
# register(bpp)
# lapply(singsampdat,FUN=function(d){
#     genes <- unique(d$gene)
#     plts <- bplapply(genes,BPPARAM=bpp,BPOPTIONS = bpoptions(exportglobals =TRUE,packages=c("data.table","ggplot2","gridExtra"),exports = c("pals")),FUN=function(g){
#         curcolor <- pals[gene_set==unique(d[gene==g,gene_set]),pal]
#         ggplot(d[gene==g],aes(x=orig_value,y=reseg_value,col=curcolor))+
#             geom_point(size = 0.5)+
#             ggtitle(paste0(unique(d$sample),": ",g, " logcounts"))+
#             xlim(0,max(c(d$orig_value,d$reseg_value)))+
#             ylim(0,max(c(d$orig_value,d$reseg_value)))+
#             geom_abline(intercept = 0,slope = 1,linetype="dashed",col="black",alpha=0.4)+
#             xlab("v1.5 cells: w/in samp norm logcounts")+
#             ylab("v1.7 cells: w/in samp norm logcounts")
#     })
#     ## add legend to front of list so it's on the first page
#     plotsout <- marrangeGrob(plts,nrow=1,ncol=1)
#     ggsave(path="plots/02_compareCellSegmentations_1.5vs1.7",filename = paste0("02b-",unique(d$sample),".pdf"),plot = plotsout,width = 8.5,height=7.5)
# })

## stop parallel workers
register(SerialParam())

rm(list=ls())
gc(full=T)
```


Reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.0 Patched (2024-04-30 r86503)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.4.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.37.1             
 [3] chattr_0.1.0                   default_1.0.0                 
 [5] colorout_1.3-0.2               here_1.0.1                    
 [7] sf_1.0-16                      vroom_1.6.5                   
 [9] DropletUtils_1.24.0            Matrix_1.7-0                  
[11] scater_1.32.0                  BiocSingular_1.20.0           
[13] scuttle_1.14.0                 stringr_1.5.1                 
[15] gridExtra_2.3                  ggplot2_3.5.1                 
[17] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[19] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[21] GenomicRanges_1.56.0           GenomeInfoDb_1.40.1           
[23] IRanges_2.38.0                 S4Vectors_0.42.0              
[25] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[27] matrixStats_1.3.0              patchwork_1.2.0               
[29] Voyager_1.6.0                  SpatialFeatureExperiment_1.6.1
[31] data.table_1.15.99             rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               bitops_1.0-7             
  [3] tibble_3.2.1              R.oo_1.26.0              
  [5] lifecycle_1.0.4           httr2_1.0.1              
  [7] edgeR_4.2.0               rprojroot_2.0.4          
  [9] processx_3.8.4            lattice_0.22-6           
 [11] magrittr_2.0.3            limma_3.60.2             
 [13] sass_0.4.9                rmarkdown_2.27           
 [15] jquerylib_0.1.4           yaml_2.3.8               
 [17] httpuv_1.6.15             sp_2.1-4                 
 [19] DBI_1.2.3                 abind_1.4-5              
 [21] zlibbioc_1.50.0           purrr_1.0.2              
 [23] R.utils_2.12.3            nngeo_0.4.8              
 [25] RCurl_1.98-1.14           rappdirs_0.3.3           
 [27] GenomeInfoDbData_1.2.12   ggrepel_0.9.5            
 [29] irlba_2.3.5.1             terra_1.7-78             
 [31] units_0.8-5               RSpectra_0.16-1          
 [33] dqrng_0.4.1               DelayedMatrixStats_1.26.0
 [35] codetools_0.2-20          DelayedArray_0.30.1      
 [37] tidyselect_1.2.1          UCSC.utils_1.0.0         
 [39] memuse_4.2-3              farver_2.1.2             
 [41] ScaledMatrix_1.12.0       viridis_0.6.5            
 [43] jsonlite_1.8.8            BiocNeighbors_1.22.0     
 [45] e1071_1.7-14              systemfonts_1.1.0        
 [47] tools_4.4.0               ggnewscale_0.4.10        
 [49] ragg_1.3.2                snow_0.4-4               
 [51] Rcpp_1.0.12               glue_1.7.0               
 [53] SparseArray_1.4.8         xfun_0.44                
 [55] EBImage_4.46.0            dplyr_1.1.4              
 [57] HDF5Array_1.32.0          withr_3.0.0              
 [59] fastmap_1.2.0             boot_1.3-30              
 [61] rhdf5filters_1.16.0       bluster_1.14.0           
 [63] fansi_1.0.6               spData_2.3.1             
 [65] callr_3.7.6               digest_0.6.35            
 [67] rsvd_1.0.5                R6_2.5.1                 
 [69] mime_0.12                 textshaping_0.4.0        
 [71] colorspace_2.1-0          wk_0.9.1                 
 [73] jpeg_0.1-10               R.methodsS3_1.8.2        
 [75] utf8_1.2.4                generics_0.1.3           
 [77] class_7.3-22              httr_1.4.7               
 [79] htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [81] scatterplot3d_0.3-44      spdep_1.3-4              
 [83] pkgconfig_2.0.3           scico_1.5.0              
 [85] gtable_0.3.5              XVector_0.44.0           
 [87] htmltools_0.5.8.1         fftwtools_0.9-11         
 [89] scales_1.3.0              png_0.1-8                
 [91] knitr_1.47                rstudioapi_0.16.0        
 [93] tzdb_0.4.0                rjson_0.2.21             
 [95] proxy_0.4-27              cachem_1.1.0             
 [97] rhdf5_2.48.0              Polychrome_1.5.1         
 [99] KernSmooth_2.23-24        parallel_4.4.0           
[101] vipor_0.4.7               s2_1.1.6                 
[103] pillar_1.9.0              grid_4.4.0               
[105] vctrs_0.6.5               promises_1.3.0           
[107] beachmat_2.20.0           xtable_1.8-4             
[109] sfheaders_0.4.4           cluster_2.1.6            
[111] beeswarm_0.4.0            evaluate_0.23            
[113] zeallot_0.1.0             magick_2.8.3             
[115] cli_3.6.2                 locfit_1.5-9.9           
[117] compiler_4.4.0            crayon_1.5.2             
[119] labeling_0.4.3            classInt_0.4-10          
[121] ps_1.7.6                  fs_1.6.4                 
[123] ggbeeswarm_0.7.2          stringi_1.8.4            
[125] viridisLite_0.4.2         deldir_2.0-4             
[127] munsell_0.5.1             tiff_0.1-12              
[129] sparseMatrixStats_1.16.0  bit64_4.0.5              
[131] Rhdf5lib_1.26.0           statmod_1.5.0            
[133] shiny_1.8.1.1             clipr_0.8.0              
[135] igraph_2.0.3              bslib_0.7.0              
[137] bit_4.0.5                
> sessioninfo::session_info()
─ Session info ─────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.0 Patched (2024-04-30 r86503)
 os       macOS Sonoma 14.4.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-11
 rstudio  2024.04.2+764 Chocolate Cosmos (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ─────────────────────────────────────────────────────────────
 package                  * version   date (UTC) lib source
 abind                      1.4-5     2016-07-21 [2] CRAN (R 4.4.0)
 beachmat                   2.20.0    2024-05-06 [2] Bioconductor 3.19 (R 4.4.0)
 beeswarm                   0.4.0     2021-06-01 [2] CRAN (R 4.4.0)
 Biobase                  * 2.64.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocGenerics             * 0.50.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocNeighbors              1.22.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocParallel             * 1.38.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocSingular             * 1.20.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 bit                        4.0.5     2022-11-15 [1] CRAN (R 4.4.0)
 bit64                      4.0.5     2020-08-30 [1] CRAN (R 4.4.0)
 bitops                     1.0-7     2021-04-24 [2] CRAN (R 4.4.0)
 bluster                    1.14.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 boot                       1.3-30    2024-02-26 [2] CRAN (R 4.4.0)
 bslib                      0.7.0     2024-03-29 [2] CRAN (R 4.4.0)
 cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 callr                      3.7.6     2024-03-25 [2] CRAN (R 4.4.0)
 chattr                   * 0.1.0     2024-04-27 [2] CRAN (R 4.4.0)
 class                      7.3-22    2023-05-03 [2] CRAN (R 4.4.0)
 classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 cli                        3.6.2     2023-12-11 [1] CRAN (R 4.4.0)
 clipr                      0.8.0     2022-02-22 [2] CRAN (R 4.4.0)
 cluster                    2.1.6     2023-12-01 [2] CRAN (R 4.4.0)
 codetools                  0.2-20    2024-03-31 [2] CRAN (R 4.4.0)
 colorout                 * 1.3-0.2   2024-05-01 [2] Github (jalvesaq/colorout@c6113a2)
 colorspace                 2.1-0     2023-01-23 [2] CRAN (R 4.4.0)
 crayon                     1.5.2     2022-09-29 [2] CRAN (R 4.4.0)
 data.table               * 1.15.99   2024-05-01 [2] local
 DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
 default                  * 1.0.0     2017-08-07 [2] CRAN (R 4.4.0)
 DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
 DelayedMatrixStats         1.26.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 deldir                     2.0-4     2024-02-28 [2] CRAN (R 4.4.0)
 digest                     0.6.35    2024-03-11 [2] CRAN (R 4.4.0)
 dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
 dqrng                      0.4.1     2024-05-28 [2] CRAN (R 4.4.0)
 DropletUtils             * 1.24.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
 EBImage                    4.46.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 edgeR                      4.2.0     2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 evaluate                   0.23      2023-11-01 [2] CRAN (R 4.4.0)
 fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
 farver                     2.1.2     2024-05-13 [2] CRAN (R 4.4.0)
 fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
 fftwtools                  0.9-11    2021-03-01 [2] CRAN (R 4.4.0)
 fs                         1.6.4     2024-04-25 [2] CRAN (R 4.4.0)
 generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
 GenomeInfoDb             * 1.40.1    2024-05-24 [2] Bioconductor 3.19 (R 4.4.0)
 GenomeInfoDbData           1.2.12    2024-05-01 [2] Bioconductor
 GenomicRanges            * 1.56.0    2024-05-06 [2] Bioconductor 3.19 (R 4.4.0)
 ggbeeswarm                 0.7.2     2023-04-29 [2] CRAN (R 4.4.0)
 ggnewscale                 0.4.10    2024-02-08 [2] CRAN (R 4.4.0)
 ggplot2                  * 3.5.1     2024-04-23 [2] CRAN (R 4.4.0)
 ggrepel                    0.9.5     2024-01-10 [2] CRAN (R 4.4.0)
 glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
 gridExtra                * 2.3       2017-09-09 [2] CRAN (R 4.4.0)
 gtable                     0.3.5     2024-04-22 [2] CRAN (R 4.4.0)
 HDF5Array                  1.32.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 here                     * 1.0.1     2020-12-13 [2] CRAN (R 4.4.0)
 htmltools                  0.5.8.1   2024-04-04 [2] CRAN (R 4.4.0)
 htmlwidgets                1.6.4     2023-12-06 [2] CRAN (R 4.4.0)
 httpuv                     1.6.15    2024-03-26 [2] CRAN (R 4.4.0)
 httr                       1.4.7     2023-08-15 [2] CRAN (R 4.4.0)
 httr2                      1.0.1     2024-04-01 [2] CRAN (R 4.4.0)
 igraph                     2.0.3     2024-03-13 [2] CRAN (R 4.4.0)
 IRanges                  * 2.38.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 irlba                      2.3.5.1   2022-10-03 [2] CRAN (R 4.4.0)
 jpeg                       0.1-10    2022-11-29 [2] CRAN (R 4.4.0)
 jquerylib                  0.1.4     2021-04-26 [2] CRAN (R 4.4.0)
 jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
 KernSmooth                 2.23-24   2024-05-17 [2] CRAN (R 4.4.0)
 knitr                      1.47      2024-05-29 [2] CRAN (R 4.4.0)
 labeling                   0.4.3     2023-08-29 [2] CRAN (R 4.4.0)
 later                      1.3.2     2023-12-06 [2] CRAN (R 4.4.0)
 lattice                    0.22-6    2024-03-20 [2] CRAN (R 4.4.0)
 lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
 limma                      3.60.2    2024-05-19 [1] Bioconductor 3.19 (R 4.4.0)
 locfit                     1.5-9.9   2024-03-01 [2] CRAN (R 4.4.0)
 magick                     2.8.3     2024-02-18 [2] CRAN (R 4.4.0)
 magrittr                   2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
 Matrix                   * 1.7-0     2024-03-22 [2] CRAN (R 4.4.0)
 MatrixGenerics           * 1.16.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 matrixStats              * 1.3.0     2024-04-11 [2] CRAN (R 4.4.0)
 memuse                     4.2-3     2023-01-24 [2] CRAN (R 4.4.0)
 mime                       0.12      2021-09-28 [2] CRAN (R 4.4.0)
 munsell                    0.5.1     2024-04-01 [2] CRAN (R 4.4.0)
 nngeo                      0.4.8     2024-04-17 [2] CRAN (R 4.4.0)
 parallelly               * 1.37.1    2024-02-29 [2] CRAN (R 4.4.0)
 patchwork                * 1.2.0     2024-01-08 [2] CRAN (R 4.4.0)
 pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
 pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
 png                        0.1-8     2022-11-29 [2] CRAN (R 4.4.0)
 Polychrome                 1.5.1     2022-05-03 [2] CRAN (R 4.4.0)
 processx                   3.8.4     2024-03-16 [2] CRAN (R 4.4.0)
 promises                   1.3.0     2024-04-05 [2] CRAN (R 4.4.0)
 proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
 ps                         1.7.6     2024-01-18 [2] CRAN (R 4.4.0)
 purrr                      1.0.2     2023-08-10 [1] CRAN (R 4.4.0)
 R.methodsS3                1.8.2     2022-06-13 [2] CRAN (R 4.4.0)
 R.oo                       1.26.0    2024-01-24 [2] CRAN (R 4.4.0)
 R.utils                    2.12.3    2023-11-18 [2] CRAN (R 4.4.0)
 R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
 ragg                       1.3.2     2024-05-15 [2] CRAN (R 4.4.0)
 rappdirs                   0.3.3     2021-01-31 [2] CRAN (R 4.4.0)
 Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
 RCurl                      1.98-1.14 2024-01-09 [2] CRAN (R 4.4.0)
 rhdf5                      2.48.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 rhdf5filters               1.16.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 Rhdf5lib                   1.26.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 rjson                      0.2.21    2022-01-09 [2] CRAN (R 4.4.0)
 rlang                    * 1.1.4     2024-06-04 [1] CRAN (R 4.4.0)
 rmarkdown                  2.27      2024-05-17 [2] CRAN (R 4.4.0)
 rprojroot                  2.0.4     2023-11-05 [2] CRAN (R 4.4.0)
 RSpectra                   0.16-1    2022-04-24 [2] CRAN (R 4.4.0)
 rstudioapi                 0.16.0    2024-03-24 [2] CRAN (R 4.4.0)
 rsvd                       1.0.5     2021-04-16 [2] CRAN (R 4.4.0)
 s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
 S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
 S4Vectors                * 0.42.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 sass                       0.4.9     2024-03-15 [2] CRAN (R 4.4.0)
 ScaledMatrix               1.12.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 scales                     1.3.0     2023-11-28 [2] CRAN (R 4.4.0)
 scater                   * 1.32.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 scatterplot3d              0.3-44    2023-05-05 [2] CRAN (R 4.4.0)
 scico                      1.5.0     2023-08-14 [2] CRAN (R 4.4.0)
 scuttle                  * 1.14.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 sessioninfo                1.2.2     2021-12-06 [2] CRAN (R 4.4.0)
 sf                       * 1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
 sfheaders                  0.4.4     2024-01-17 [2] CRAN (R 4.4.0)
 shiny                      1.8.1.1   2024-04-02 [2] CRAN (R 4.4.0)
 SingleCellExperiment     * 1.26.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 snow                       0.4-4     2021-10-27 [2] CRAN (R 4.4.0)
 sp                         2.1-4     2024-04-30 [2] CRAN (R 4.4.0)
 SparseArray                1.4.8     2024-05-30 [2] Bioconductor 3.19 (R 4.4.0)
 sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 SpatialExperiment        * 1.14.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 SpatialFeatureExperiment * 1.6.1     2024-05-15 [2] Bioconductor 3.19 (R 4.4.0)
 spData                     2.3.1     2024-05-31 [2] CRAN (R 4.4.0)
 spdep                      1.3-4     2024-05-31 [2] CRAN (R 4.4.0)
 statmod                    1.5.0     2023-01-06 [2] CRAN (R 4.4.0)
 stringi                    1.8.4     2024-05-06 [2] CRAN (R 4.4.0)
 stringr                  * 1.5.1     2023-11-14 [2] CRAN (R 4.4.0)
 SummarizedExperiment     * 1.34.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 systemfonts                1.1.0     2024-05-15 [2] CRAN (R 4.4.0)
 terra                      1.7-78    2024-05-22 [2] CRAN (R 4.4.0)
 textshaping                0.4.0     2024-05-24 [2] CRAN (R 4.4.0)
 tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
 tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
 tiff                       0.1-12    2023-11-28 [2] CRAN (R 4.4.0)
 tzdb                       0.4.0     2023-05-12 [2] CRAN (R 4.4.0)
 UCSC.utils                 1.0.0     2024-05-06 [2] Bioconductor 3.19 (R 4.4.0)
 units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
 utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
 vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
 vipor                      0.4.7     2023-12-18 [2] CRAN (R 4.4.0)
 viridis                    0.6.5     2024-01-29 [2] CRAN (R 4.4.0)
 viridisLite                0.4.2     2023-05-02 [2] CRAN (R 4.4.0)
 Voyager                  * 1.6.0     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
 vroom                    * 1.6.5     2023-12-05 [2] CRAN (R 4.4.0)
 withr                      3.0.0     2024-01-16 [1] CRAN (R 4.4.0)
 wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
 xfun                       0.44      2024-05-15 [2] CRAN (R 4.4.0)
 xtable                     1.8-4     2019-04-21 [2] CRAN (R 4.4.0)
 XVector                    0.44.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 yaml                       2.3.8     2023-12-11 [2] CRAN (R 4.4.0)
 zeallot                    0.1.0     2018-01-28 [2] CRAN (R 4.4.0)
 zlibbioc                   1.50.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)

 [1] /Users/bmulvey/Library/R/arm64/4.4/library
 [2] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library

────────────────────────────────────────────────────────────────────────