---
title: "04-Compare ARC sex DE LFCs w each comb of single SpReps per Donor"
author: "Bernard Mulvey"
date: "2023-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater)
library(scran)
library(edgeR)
library(jaffelab)
library(PCAtools) # findElbowPoint
library(tidyr) # generate table of unique combinations of SpRep vectors
library(parallel)
library(BiocParallel)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

Load SPE, set spot names to coldata$key, add bayesspsace nnsvg10 k=15 labels. Crop V12D07-075-B1 down to the portion containing ISH-verified VMH. (This contains some spots of ARC clusters as well).
```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")

colnames(hyp2) <- colData(hyp2)$key

bscl <- fread("analysis/data/spe_053123/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/BSpace_Harmony_nnsvg10pctile_15.txt")
bscl <- as.data.frame(bscl)
rownames(bscl) <- bscl$rn

colLabels(hyp2) <- as.factor(bscl$BShar_nnsvg10pctile_15)

### crop V12D07-075-B1 as previously
coordstmp <- as.data.frame(spatialCoords(hyp2))
rownames(coordstmp) <- colnames(hyp2)
coordstmp <- as.data.table(coordstmp,keep.rownames=T)
coordstmp <- coordstmp[,diag:=pxl_col_in_fullres+pxl_row_in_fullres]
coordstmp <- coordstmp[!(rn %in% grep(rn,pattern="V12D07-075_B1",value=T)&diag>30000)]

hyp2 <- hyp2[,colnames(hyp2)%in%coordstmp$rn]
stopifnot(dim(hyp2)[2]==nrow(coordstmp))

rm(bscl,coordstmp)

## add demographics for sample sexes
demos <- fread("raw-data/demos.txt")
tmpcd <- as.data.table(colData(hyp2),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,demos,by.x="brnum",by.y="BrNum")
tmpcd <- as.data.frame(tmpcd)
rownames(tmpcd) <- tmpcd$rn
tmpcd <- tmpcd[colnames(hyp2),]
colData(hyp2) <- DataFrame(tmpcd) 
rm(tmpcd,demos)
# load chromosome info for genes for filtering by X/Y later. (queried rownames of hyp2 in biomart human genes v 109 for this.)
chrs <- fread("raw-data/ENSG_chromosomes.txt")
allg <- as.data.table(rowData(hyp2),keep.rownames=T)
allg <- merge.data.table(allg,chrs,by.x="rn",by.y="gene_id",all.x=T)
allg <- as.data.frame(allg)
rownames(allg) <- allg$rn
allg <- DataFrame(allg)
### CRITICAL TO REORDER HERE OR GENES SOMEHOW GET MIS-LABELED LATER (easy test for this is examining DDX3Y expression spotplot in hyp2 now vs after)
allg <- allg[rownames(hyp2),]
rowData(hyp2) <- allg
### spot checks confirm gene-chr pairs are correct. chrY genes are absent in female samples up to this point.
rm(chrs,allg)

## ad a columns to indicate whether or not the spot is in one of the ARC clustesr (8 OR 13)
hyp2$is.arc <- as.factor(colData(hyp2)$label %in% c(8,13))
hyp2$is.arc8 <- as.factor(colData(hyp2)$label==8)
hyp2$is.arc13 <- as.factor(colData(hyp2)$label==13)
```

### Comparative Approach 1: Each combination of data where only 1 spatial rep per donor is represented. Cross-correlate logFCs from the unique combinations of included samples.

Br8741 spatial reps: V12D05-348_C1 and V12D05-350_C1
Br6588 spatial reps: V12D05-348_D1 and V12D05-350_D1

### 
```{r}
unique(as.data.table(colData(hyp2))[,.(sample_id,brnum)])

# get table of unique combos of spatial reps that could be left out using tidyr::crossing(), which won't change the characters to factors (unlike expand.grid), per https://stackoverflow.com/a/50957414 . keep this table later so we add the LFCs etc
Br8741 <- c("V12D05-348_C1","V12D05-350_C1")
Br6588 <- c("V12D05-348_D1","V12D05-350_D1")
leaveout <- as.data.frame(tidyr::crossing(Br8741=Br8741,Br6588=Br6588))

# make which samples to exclude into lists to be passed and iterated thru in bpmapply
leaveoutA <- as.list(leaveout$Br8741)
leaveoutB <- as.list(leaveout$Br6588)
# and a third list for the cluster set(s) from ARC to compare
cllists <- list("is.arc8","is.arc13","is.arc")

rm(Br8741,Br6588,leaveout)
### set up for parallelized exeec
bpm <- MulticoreParam(workers=9)
register(bpm)
hyp2.tmp <- hyp2 # belt+suspenders: create copy of the full SPE to use for subsetting to avoid modification of the full object in case we need later 

### calculate sex DE results for each combination of singular spatial reps without additional adjustments
arc.lvout.res <- list()
i <- 1
for (i in c(1:4)){
    j <-1
    for (j in c(1:3)){
            hyp2.proc <- hyp2.tmp[,!(hyp2.tmp$sample_id==leaveoutA[[i]])&!(hyp2.tmp$sample_id==leaveoutB[[i]])]
            
            hyp2.pseu <- aggregateAcrossCells(hyp2.proc,
                       DataFrame(is.arc=colData(hyp2.proc)[,cllists[[j]]],
                                 sample=hyp2.proc$sample_id))
        
            colData(hyp2.pseu)[,cllists[[j]]] <- as.factor(colData(hyp2.pseu)[,cllists[[j]]])
    
    # find a good expression cutoff using edgeR::filterByExpr
        rowData(hyp2.pseu)$high_expr <- filterByExpr(hyp2.pseu)
        rowData(hyp2.pseu)$high_expr_cl <- filterByExpr(hyp2.pseu, group = colData(hyp2.pseu)[,cllists[[j]]])
        rowData(hyp2.pseu)$high_expr_sample <- filterByExpr(hyp2.pseu, group = hyp2.pseu$sample)
    
        rowData(hyp2.pseu)$high_expr_clorsam <-     rowSums(cbind(rowData(hyp2.pseu)$high_expr_sample,rowData(hyp2.pseu)$high_expr_cl))
        hyp2.pseu <- hyp2.pseu[rowData(hyp2.pseu)$high_expr_clorsam>=1]
        
        hyp2.pseu2 <- edgeR::cpm(edgeR::calcNormFactors(hyp2.pseu), log = TRUE, prior.count = 0.1)
        #### fix dimnames, which rigtmpcdht now are changed to "sample1, sample2...":
        # colnames(hyp2.pseu)
        colnames(hyp2.pseu2) <- paste0(hyp2.pseu$sample_id,"_",cllists[[j]])
        colnames(hyp2.pseu) <- paste0(hyp2.pseu$sample_id,"_",cllists[[j]])
        
        ## store norm counts on spe
        logcounts(hyp2.pseu) <- hyp2.pseu2
        
        ## double check
        stopifnot(rownames(hyp2.pseu)==rownames(hyp2.tmp)[which(rownames(hyp2.tmp) %in% rownames(hyp2.pseu))])
    
        ## dump unneeded objs
        rm(hyp2.pseu2)
        ## make sure variables of interest for DE modeling are factors
        colData(hyp2.pseu)$Sex <- factor(colData(hyp2.pseu)$Sex)
        
        ## de analyze
        deres <- pseudoBulkDGE(
            hyp2.pseu,
            col.data = colData(hyp2.pseu),
            label = colData(hyp2.pseu)[,cllists[[j]]],
            design = ~Sex,
            coef="SexM",
            row.data = rowData(hyp2.pseu),
            method = "voom",
            qualities = TRUE,
            robust = TRUE
            )
        
            tmpout <- as.data.table(deres)
            arc.lvout.res[[(length(arc.lvout.res))+1]] <- tmpout
            names(arc.lvout.res)[length(arc.lvout.res)] <- paste0(leaveoutA[[i]],"_",leaveoutB[[i]],"_",cllists[[j]])
            rm(hyp2.proc,hyp2.pseu,hyp2.tmp,tmpout)
    }
}
rm(hyp2.tmp)
# saveRDS(arc.lvout.res,"analysis/data/spe_053123/H05-bspace15-nnsvg10_ARC_clusters/singleSpReps_ARCclusters-sexDE_reslist_named-by-excluded-samples-and-tested-cluster.RDS")
```


Tabulate DEGs
```{r}

i <- 1
for (i in c(1:12)){
    setnames(arc.lvout.res[[i]],"logFC",paste0(names(arc.lvout.res)[i],"_logFC"))
    arc.lvout.res[[i]] <- arc.lvout.res[[i]][group_name==TRUE]
    arc.lvout.res[[i]]$group_name <- gsub(names(arc.lvout.res)[i],pattern="^.*_.*_.*_.*_(is\\.ar.*)$",replacement="\\1")}

### merge
getnames <- c("gene_name",grep(names(arc.lvout.res[[1]]),pattern="logFC",value=T))
alllfc <- arc.lvout.res[[1]][,..getnames]
rm(getnames)

i<-2
for (i in c(2:12)){
    getnames <- c("gene_name",grep(names(arc.lvout.res[[i]]),pattern="logFC",value=T))
    alllfc <- merge.data.table(alllfc,arc.lvout.res[[i]][,..getnames],by="gene_name",all.x=T,all.y=T)
    rm(getnames)
}

### paircors
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y,use = "pairwise.complete.obs"))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * r)
}

# png("plots/spe_053123/H05-bspace15-nnsvg10_ARC_clusters/02-Autosomal Sex DEGs-no SpRep adjustment, postcrop V12D07075B1/pbulk_sexDE-LFC_correls_onesprep-per-donor_eaARCgrouping.png",height=5000,width=5000)
pairs(as.matrix(alllfc[,c(2:13)]),use="pairwise.complete.obs",upper.panel=panel.cor)
dev.off()
####
```

### Get consensus DEGs (≥75% of tests FDR < 0.25 with matched dirs)
```{r}
getnames <- c("gene_name",grep(names(arc.lvout.res[[1]]),pattern="logFC",value=T))
alllfc2 <- arc.lvout.res[[1]][adj.P.Val<0.25,..getnames]
rm(getnames)

i<-2
for (i in c(2:12)){
    getnames <- c("gene_name",grep(names(arc.lvout.res[[i]]),pattern="logFC",value=T))
    alllfc2 <- merge.data.table(alllfc2,arc.lvout.res[[i]][adj.P.Val<0.25,..getnames],by="gene_name",all.x=T,all.y=T)
    rm(getnames)
}
lfcs.long <- melt(alllfc2,by="gene_name")
lfcs.long[,dir:=ifelse(value>0,yes="M",no="F")]
lfcs.long <- lfcs.long[!is.na(dir)]
lfcs.long[,gene_dir:=paste(gene_name,"_",dir)]
ans1<-lfcs.long[,.N,by=gene_dir][N>7]

### well shank is going to come out fo sho.

```
