---
title: "01-HVGs and nnSVG"
author: "Bernie Mulvey"
date: "2023-05-23"
output: html_document
---

```{r setup, include=FALSE}
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(ggspavis)
library(scater) # addPerCellQC
library(BiocParallel)
library(scran)
library(parallel)
library(nnSVG)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### Without batch, etc correction ###

### scran - variable genes (HVG), spatially agnostic
### OSTA verbatim: "If the biologically meaningful spatial information in this dataset mainly reflects spatial distributions of major cell types, then relying on HVGs for downstream analyses may be sufficient. But if there are additional important spatial features in the dataset, then it may be more meaningful to define spatially variable genes."
### We won't know with the hyp samples til we know. So let's start unsupervised.
```{r}
# ordinarily, we exclude mitochondrial genes as not of interest. but the LC spatial preprint highlighted that LC neurons had very high mitochondrial reads, so what the hey, we'll keep them in here and just grab extra top HVGs to account for the fact that the 13 mito genes might take up some of the list.

hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_D07-075_B1 crop 070223.RDS")
# first, we model gene-level mean-variance relationships
hyp2.vars <- modelGeneVar(hyp2)
hyp2.vars.fit <- metadata(hyp2.vars)

# # plot in base R:
# plot(splc2.vars.fit$mean, splc2.vars.fit$var, 
#      xlab = "mean of log-expression", ylab = "variance of log-expression")
# curve(splc2.vars.fit$trend(x), col = "blue", add = TRUE, lwd = 2)
# dev.off()
# plot in ggplot2
hyp2.vars.dt <- as.data.table(hyp2.vars.fit$var,keep.rownames=T)
hyp2.means.dt <- as.data.table(hyp2.vars.fit$mean,keep.rownames=T)
stopifnot(sum(hyp2.means.dt$V1==hyp2.vars.dt$V1)==nrow(hyp2.means.dt))
setnames(hyp2.vars.dt,2,"var_log_xpr")
setnames(hyp2.means.dt,2,"mean_log_xpr")
hyp2.mnvar.pltdat <- merge(hyp2.means.dt,hyp2.vars.dt,by="V1")

rm(hyp2.vars.dt,hyp2.means.dt)
# ggplot(hyp2.mnvar.pltdat,aes(x=mean_log_xpr,y=var_log_xpr))+
#   geom_point()+
# geom_line(data=environment(hyp2.vars.fit$trend),aes(x=x,y=y),inherit.aes = F)

# ^ "`data` must be a <data.frame>, or an object coercible by `fortify()`, not an environment."

pdf("plots/spe_070123/01-feature_selection/mean-var-relation_umi600_gene450_mito35_NOharmony.pdf",height=6,width=6)
ggplot(hyp2.mnvar.pltdat, aes(x = mean_log_xpr, y = var_log_xpr)) +
    geom_point() +
    geom_function(fun=hyp2.vars.fit$trend,col="blue")
dev.off()
rm(hyp2.mnvar.pltdat)
```

#### HVGs continued: pull out the top p % of variable genes (returns a vector of identifiers,no stats etc)
```{r}
hyp2.hvg10 <- scran::getTopHVGs(hyp2.vars,prop = 0.1)
hyp2.hvg20 <- scran::getTopHVGs(hyp2.vars,prop=0.2)
#saveRDS(hyp2.hvg10,"analysis/data/spe_070123/01-feature_selection/top10pct_HVGs_NOharmony.RDS")
#saveRDS(hyp2.hvg20,"analysis/data/spe_070123/01-feature_selection/top20pct_HVGs_NOharmony.RDS")
```


### Spatially variable genes--nnSVG, per sample
### the workhorse function nnSVG can use parallel processing through BiocParallel. Set up the preferred (i.e., compatible) param. the parallelization is within the nnsvg call, so no need to match cpu # to sample #
```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_D07-075_B1 crop 070223.RDS")
sbp <- MulticoreParam(10)
register(sbp)

# SUPER CRITICAL IMPORTANT NOTES: 
# 1. RECALCULATE LOG COUNTS AFTER FILTERING ^
# 2.nnSVG works ONE SAMPLE (one capture area) at a time. So need to wrap this in a loop. 
# 3. as such, perform filtering PER SAMPLE. 
# filter_genes_ncounts (N) and filter_genes_pcspots (S) work together: genes will be filtered to those with at least N counts in at least S% of spots.  filter_genes_pcspots expects a %age as 0-100, NOT a fraction.
# tried N=2 (instead of 3) and S = 0.25 instead of 0.5, since most spots are pretty diverse and high-depth (and the cell type diversity is also unknown), but didn't get past the first sample in 3 hours on 10 CPUs. so now trying 3, 0.5 .

# With vignette as a guide...:
caps <- unique(colData(hyp2)$sample_id)
res_list2 <- as.list(rep(NA, length(caps)))
names(res_list2) <- caps

i<-1
for (i in c(1:length(caps))){
    cap <- caps[i]
    message(paste0("Starting sample ",i," of ",length(caps)))
    hyp2.samp <- hyp2[,colData(hyp2)$sample_id==cap]
    # dim(hyp2.samp)
    hyp2.samp <- filter_genes(hyp2.samp,filter_genes_ncounts = 3,filter_genes_pcspots = 0.5,filter_mito = F)
    #recalculate logcounts
    hyp2.samp <- computeLibraryFactors(hyp2.samp)
    hyp2.samp <- logNormCounts(hyp2.samp)
    # main call -- defaults to logcounts but shown for posterity; X if using covariates; n_neighbors defaults to 10, slower but potentially more granular with higher values eg 15, order = don't touch unless you have < 70 spots total; n_threads or BPPARAM for parallelization
  res_list2[[i]] <- nnSVG(hyp2.samp,assay_name = "logcounts",BPPARAM = sbp)
  rm(hyp2.samp,cap)
  gc(full=T)
}

# extract SVG results
res_list3 <- lapply(res_list2,FUN=function(x){
    # this first line renders rownames invalid, so don't carry them along
    rowData(x) <- rowData(x)[order(rowData(x)$rank),]
    colnames(rowData(x))[8:21] <- paste0("nnSVG_",colnames(rowData(x))[8:21])
    as.data.table(as.data.frame(rowData(x)),keep.rownames=F)
})

saveRDS(res_list3,"analysis/data/spe_070123/01-feature_selection/nnSVG_samplewise_n3s0.5_070223.RDS")
nnsvg_N3Spt5 <- res_list3
rm(i,caps,sbp,res_list2,res_list3)
gc(full=T)
```

### To determine experiment-wide nnSVG-derived SVGs, average the ranks from each sample in our results list.

```{r}
i<-1
for (i in c(1:sum(unique(colData(hyp2)$sample_id)%in%names(nnsvg_N3Spt5)))){
  if(i==1){
  nnsvg_ranks <- nnsvg_N3Spt5[[i]][nnSVG_padj<0.05]
  nnsvg_ranks <- nnsvg_ranks[,.(gene_id,gene_name,nnSVG_rank)]
  setnames(nnsvg_ranks,3,names(nnsvg_N3Spt5)[i])
  }
  else{
  tmp <- nnsvg_N3Spt5[[i]][nnSVG_padj<0.05]
  tmp <- tmp[,.(gene_id,gene_name,nnSVG_rank)]
  setnames(tmp,3,names(nnsvg_N3Spt5)[i])
  nnsvg_ranks <- merge.data.table(nnsvg_ranks,tmp,by=c("gene_id","gene_name"),all.x=T,all.y=T,incomparables = NA)
  rm(tmp)
  }
}
rm(i)

nnsvg_ranks[,meanrank:=rowMeans(.SD,na.rm=T),.SDcols=c(3:10)]
write.table(nnsvg_ranks,"analysis/data/spe_070123/01-feature_selection/meangeneranks_nnSVG_samplewise_n3s0.5.txt",sep='\t',quote=F,row.names=F,col.names=T)
```

### LEFT OFF HERE ####

There are 1828 HVGs (top 10%ile), 3656 HVGs (top 20%ile); 7767 HVGs significant in at least 1 sample. let's compare what's unique and shared between the M or N HVGs and top (by mean rank) M or N SVGs.
```{r}
hyp2.hvg10 <- readRDS("analysis/data/01-feature_selection/top10pct_HVGs_NOharmony.RDS")
hyp2.hvg20 <- readRDS("analysis/data/01-feature_selection/top20pct_HVGs_NOharmony.RDS")
setorder(nnsvg_ranks,meanrank)
sum(hyp2.hvg10 %in% nnsvg_ranks[1:1828,gene_id]) 
sum(hyp2.hvg20 %in% nnsvg_ranks[1:3656,gene_id])
# 696/1828 of top 10% HVGs are top 1828 SVGs
# 1299/3656 top 20% HVGs are in top 3656 SVGs
# let's plot how many of the SVGs are in the two HVG sets over a step series of 25 genes
olap.plt <- as.data.frame(matrix(nrow=floor(length(hyp2.hvg20)/25),ncol=2))
olap.plt[,1] <- seq(from=26,to=nrow(olap.plt)*25+25,by=25)
i <- 1
for (i in c(1:nrow(olap.plt))){
  olap.plt[i,2] <- sum(hyp2.hvg20[c(i:olap.plt[i,1])] %in% nnsvg_ranks$gene_id[c(1:olap.plt[i,1])])/(olap.plt[i,1])
}
rm(i)

colnames(olap.plt) <- c("Top n SorH VGs","pct overlap")

ggplot(olap.plt,aes(x=`Top n SorH VGs`,y=`pct overlap`))+
  geom_col()
dev.off()

## lets make sets of the top M SVGs, top N SVGs, and the top M SVGs + 10%ile list of HVGs (length=N), and N SVGs + 20%ile HVGs (length=N) so we can visualize the different sets in clusters downstream. also the equivalent length of unique hvgs + top 1310 svgs in top svgs alone.
hyp2.vgs.list <- list(hyp10pct=hyp2.hvg10,
                      hvg20pct=hyp2.hvg20,
                      nnsvg2k=nnsvg_ranks[1:length(hyp2.hvg10),gene_id],
                      nnsvg3k=nnsvg_ranks[1:length(hyp2.hvg20),gene_id],
                      nnsvg2k_hvg10pct=unique(c(hyp2.hvg10,nnsvg_ranks
                               [1:length(hyp2.hvg10),gene_id])),
                      nnsvg3k_hvg20pct=unique(c(hyp2.hvg20,nnsvg_ranks
                               [1:length(hyp2.hvg20),gene_id])),
                      nnsvg_all=nnsvg_ranks$gene_id
                      )

saveRDS(hyp2.vgs.list,"analysis/data/spe_070123/01-feature_selection/nnSVG_HVG_and_combo_featurelists.RDS")
rm(olap.plt,nnsvg_ranks,hyp2.hvg10,hyp2.hvg20,nnsvg_N3Spt5)
gc(full=T)
```

