---
title: "01_check_dimreds_for_donor_slide_rundate_fx"
output: html_document
date: "2024-06-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(scater)
library(scran)
library(spatialLIBD)
library(ggrastr)
library(corral)
# for faster UMAP -- will be automatically used if detected
library(RcppHNSW)


## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")


## ggplot defaults
#theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

## read in filtered SFE
```{r}
hypx <- readRDS("processed-data/04_general_QC_and_normalization/01b-sfe_controlOutliers_and_giantNuclei_filtered.RDS")
```

## get all (targeting) genes
## get lognorm counts for running standard PCA / UMAP
```{r}
genetargeting <- grep(rownames(hypx),pattern="Unassigned|NegCon|Deprecated",value=T,invert=T)
hypx <- hypx[genetargeting,]

hypx <- computeLibraryFactors(hypx)
hypx <- logNormCounts(hypx)

## Banksy's documentation generally uses NON-log normalization, so we might as well get that too while we're at it.
assay(hypx,"normcounts") <- normalizeCounts(hypx,log=FALSE)
```

## run PCA on log counts using all (targeting) genes
## this is way faster without parallelization than with
```{r}
set.seed(42)
hypx <- scater::runPCA(hypx,subset_row=genetargeting,assay.type="logcounts")

set.seed(42)
hypx <- scater::runUMAP(hypx)

rastumap <- as.data.table(reducedDim(hypx,"UMAP"),keep.rownames=T)
rastumap <- merge.data.table(rastumap,as.data.table(colData(hypx),keep.rownames=T),by="rn")


```

## repeat as above using raw counts with correspondence analysis (CA) as executed via corral:: . this is described in https://bioconductor.org/books/3.19/OSCA.advanced/dimensionality-reduction-redux.html#count-based-dimensionality-reduction

OSCA explains the underlying procedure for CA for raw cuont data:
"For count matrices, correspondence analysis (CA) is a natural approach to dimensionality reduction. In this procedure, we compute an expected value for each entry in the matrix based on the per-gene abundance and size factors. Each count is converted into a standardized residual in a manner analogous to the calculation of the statistic in Pearson’s chi-squared tests, i.e., subtraction of the expected value and division by its square root. An SVD is then applied on this matrix of residuals to obtain the necessary low-dimensional coordinates for each cell. 

The major advantage of CA is that it avoids difficulties with the mean-variance relationship upon transformation [...] If two cells have the same expression profile but differences in their total counts, CA will return the same expected location for both cells; this avoids artifacts observed in PCA on log-transformed counts [...] However, CA is more sensitive to overdispersion in the random noise due to the nature of its standardization. This may cause some problems in some datasets where the CA factors may be driven by a few genes with random expression rather than the underlying biological structure."

```{r}
## When run on aSingleCellExperiment, returns a SCE with the embeddings (PCv from the full corral output) in the reducedDim slot "corral" (default)
## being a non SCE summarizedexperiment, we will get back a list object. and want to extract the PCv matrix for reducedDim
hypx.ca <- corral::corral_sce(hypx,subset_row=genetargeting,whichmat = "counts")

# oh, it did return an SFE. that works too.
rm(hypx)

## run UMAP on the CA matrix
set.seed(42)
hypx.ca <- runUMAP(hypx.ca,dimred="corral",name="UMAP_CA",n_threads=6)

# save the CA and both UMAP matrices bound to cell-level metadata as a table to plot from without needing to load up the whole SFE
rastumap <- as.data.table(reducedDim(hypx.ca,"UMAP"),keep.rownames=T)
rastumap <- merge.data.table(rastumap,as.data.table(colData(hypx.ca),keep.rownames=T),by="rn")
setnames(rastumap,c("UMAP1","UMAP2"),paste0(c("UMAP1","UMAP2"),"_logct"))

caumap <- as.data.table(reducedDim(hypx.ca,"UMAP_CA"),keep.rownames=T)
setnames(caumap,c("UMAP1","UMAP2"),paste0(c("UMAP1","UMAP2"),"_CA")
rastumap <- merge.data.table(rastumap,caumap,by="rn")

camat <- as.data.table(reducedDim(hypx.ca,"corral"),keep.rownames=T)
## ^ this doesnt have colnames so add
setnames(camat,c(2:ncol(camat))) <- paste("CA_dim",c(1:ncol(caumap)-1))

rastumap <- merge.data.table(rastumap,camat,by="rn")

## save for plotting
fwrite(rastumap,"processed-data/04_general_QC_and_normalization/02a-genetarg-lctUMAP_rawctCAandUMAP_w_metadata.txt.gz",sep='\t',quote=F)

# clean up enviro
rm(rastumap,caumap,camat)
gc()

```

## save SFE with normalized counts and dimreds
```{r}
saveRDS(hypx.ca,"processed-data/04_general_QC_and_normalization/02b-sfe_filtered_genetargonly_log-and-nonlog-norms.RDS")
```

# Reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.37.1             
 [3] colorout_1.3-0.2               RcppHNSW_0.6.0                
 [5] ggrastr_1.0.2                  spatialLIBD_1.16.2            
 [7] scran_1.32.0                   scater_1.32.0                 
 [9] scuttle_1.14.0                 SpatialFeatureExperiment_1.6.1
[11] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[13] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[15] GenomicRanges_1.56.1           MatrixGenerics_1.16.0         
[17] matrixStats_1.3.0              gridExtra_2.3                 
[19] ggplot2_3.5.1                  Biostrings_2.72.1             
[21] GenomeInfoDb_1.40.1            XVector_0.44.0                
[23] IRanges_2.38.0                 S4Vectors_0.42.0              
[25] BiocGenerics_0.50.0            data.table_1.15.4             
[27] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] bitops_1.0-7                sf_1.0-16                  
  [3] EBImage_4.46.0              httr_1.4.7                 
  [5] dimRed_0.2.6                RColorBrewer_1.1-3         
  [7] doParallel_1.0.17           tools_4.4.1                
  [9] utf8_1.2.4                  R6_2.5.1                   
 [11] DT_0.33                     HDF5Array_1.32.0           
 [13] lazyeval_0.2.2              uwot_0.2.2                 
 [15] rhdf5filters_1.16.0         withr_3.0.0                
 [17] sp_2.1-4                    cli_3.6.2                  
 [19] sass_0.4.9                  proxy_0.4-27               
 [21] Rsamtools_2.20.0            R.utils_2.12.3             
 [23] dichromat_2.0-0.1           sessioninfo_1.2.2          
 [25] attempt_0.3.1               maps_3.4.2                 
 [27] limma_3.60.3                pals_1.8                   
 [29] rstudioapi_0.16.0           RSQLite_2.3.7              
 [31] generics_0.1.3              BiocIO_1.14.0              
 [33] spdep_1.3-5                 dplyr_1.1.4                
 [35] Matrix_1.7-0                ggbeeswarm_0.7.2           
 [37] fansi_1.0.6                 abind_1.4-5                
 [39] R.methodsS3_1.8.2           terra_1.7-78               
 [41] lifecycle_1.0.4             yaml_2.3.8                 
 [43] edgeR_4.2.0                 rhdf5_2.48.0               
 [45] SparseArray_1.4.8           BiocFileCache_2.12.0       
 [47] paletteer_1.6.0             grid_4.4.1                 
 [49] blob_1.2.4                  promises_1.3.0             
 [51] dqrng_0.4.1                 ExperimentHub_2.12.0       
 [53] crayon_1.5.3                lattice_0.22-6             
 [55] beachmat_2.20.0             cowplot_1.1.3              
 [57] mapproj_1.2.11              KEGGREST_1.44.1            
 [59] magick_2.8.3                zeallot_0.1.0              
 [61] pillar_1.9.0                knitr_1.47                 
 [63] metapod_1.12.0              rjson_0.2.21               
 [65] boot_1.3-30                 codetools_0.2-20           
 [67] CVST_0.2-3                  wk_0.9.1                   
 [69] glue_1.7.0                  corral_1.14.0              
 [71] MultiAssayExperiment_1.30.2 vctrs_0.6.5                
 [73] png_0.1-8                   spam_2.10-0                
 [75] gtable_0.3.5                kernlab_0.9-32             
 [77] rematch2_2.1.2              cachem_1.1.0               
 [79] xfun_0.45                   S4Arrays_1.4.1             
 [81] mime_0.12                   DropletUtils_1.24.0        
 [83] sfheaders_0.4.4             iterators_1.0.14           
 [85] units_0.8-5                 fields_15.2                
 [87] statmod_1.5.0               bluster_1.14.0             
 [89] bit64_4.0.5                 RcppAnnoy_0.0.22           
 [91] filelock_1.0.3              rprojroot_2.0.4            
 [93] bslib_0.7.0                 irlba_2.3.5.1              
 [95] vipor_0.4.7                 KernSmooth_2.23-24         
 [97] colorspace_2.1-0            spData_2.3.1               
 [99] DBI_1.2.3                   tidyselect_1.2.1           
[101] bit_4.0.5                   compiler_4.4.1             
[103] curl_5.2.1                  BiocNeighbors_1.22.0       
[105] DelayedArray_0.30.1         plotly_4.10.4              
[107] rtracklayer_1.64.0          scales_1.3.0               
[109] classInt_0.4-10             rappdirs_0.3.3             
[111] stringr_1.5.1               tiff_0.1-12                
[113] digest_0.6.35               fftwtools_0.9-11           
[115] rmarkdown_2.27              benchmarkmeData_1.0.4      
[117] htmltools_0.5.8.1           pkgconfig_2.0.3            
[119] jpeg_0.1-10                 sparseMatrixStats_1.16.0   
[121] dbplyr_2.5.0                fastmap_1.2.0              
[123] ggthemes_5.1.0              htmlwidgets_1.6.4          
[125] UCSC.utils_1.0.0            shiny_1.8.1.1              
[127] DelayedMatrixStats_1.26.0   jquerylib_0.1.4            
[129] jsonlite_1.8.8              config_0.3.2               
[131] R.oo_1.26.0                 BiocSingular_1.20.0        
[133] RCurl_1.98-1.14             magrittr_2.0.3             
[135] GenomeInfoDbData_1.2.12     transport_0.15-2           
[137] s2_1.1.6                    dotCall64_1.1-1            
[139] Rhdf5lib_1.26.0             munsell_0.5.1              
[141] Rcpp_1.0.12                 viridis_0.6.5              
[143] stringi_1.8.4               zlibbioc_1.50.0            
[145] plyr_1.8.9                  AnnotationHub_3.12.0       
[147] parallel_4.4.1              ggrepel_0.9.5              
[149] deldir_2.0-4                locfit_1.5-9.9             
[151] igraph_2.0.3                reshape2_1.4.4             
[153] ScaledMatrix_1.12.0         BiocVersion_3.19.1         
[155] XML_3.99-0.16.1             evaluate_0.24.0            
[157] golem_0.4.1                 BiocManager_1.30.23        
[159] foreach_1.5.2               httpuv_1.6.15              
[161] tidyr_1.3.1                 purrr_1.0.2                
[163] benchmarkme_1.0.8           DRR_0.0.4                  
[165] rsvd_1.0.5                  xtable_1.8-4               
[167] restfulr_0.0.15             e1071_1.7-14               
[169] later_1.3.2                 viridisLite_0.4.2          
[171] class_7.3-22                tibble_3.2.1               
[173] memoise_2.0.1               beeswarm_0.4.0             
[175] AnnotationDbi_1.66.0        GenomicAlignments_1.40.0   
[177] cluster_2.1.6               shinyWidgets_0.8.6         
[179] here_1.0.1                 
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-22
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version     date (UTC) lib source
   abind                      1.4-5       2016-07-21 [1] CRAN (R 4.4.0)
   AnnotationDbi              1.66.0      2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   attempt                    0.3.1       2020-05-03 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0       2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8       2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4       2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocFileCache              2.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23     2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocSingular               1.20.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1      2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings               * 2.72.1      2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5       2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5       2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7       2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4       2023-03-17 [1] CRAN (R 4.4.0)
   bluster                    1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   boot                       1.3-30      2024-02-26 [1] CRAN (R 4.4.1)
   bslib                      0.7.0       2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0       2024-05-16 [1] CRAN (R 4.4.0)
   class                      7.3-22      2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10      2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.2       2023-12-11 [2] CRAN (R 4.4.0)
   cluster                    2.1.6       2023-12-01 [1] CRAN (R 4.4.1)
   codetools                  0.2-20      2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2     2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0       2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2       2023-08-30 [1] CRAN (R 4.4.0)
   corral                     1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   cowplot                    1.1.3       2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.2       2024-06-20 [1] CRAN (R 4.4.0) (on disk 1.5.3)
   curl                       5.2.1       2024-03-01 [1] CRAN (R 4.4.0)
   CVST                       0.2-3       2022-02-21 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4      2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3       2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0       2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1      2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4       2024-02-28 [1] CRAN (R 4.4.0)
   dichromat                  2.0-0.1     2022-05-02 [1] CRAN (R 4.4.0)
   digest                     0.6.35      2024-03-11 [1] CRAN (R 4.4.0)
   dimRed                     0.2.6       2022-07-11 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17      2022-02-07 [1] CRAN (R 4.4.0)
   dotCall64                  1.1-1       2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4       2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1       2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DRR                        0.0.4       2020-02-12 [1] CRAN (R 4.4.0)
   DT                         0.33        2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14      2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0       2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0      2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fansi                      1.0.6       2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0       2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11      2021-03-01 [1] CRAN (R 4.4.0)
   fields                     15.2        2023-08-17 [1] CRAN (R 4.4.0)
   filelock                   1.0.3       2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2       2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3       2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1      2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12      2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1      2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2       2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1       2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2       2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5       2024-01-10 [1] CRAN (R 4.4.0)
   ggthemes                   5.1.0       2024-02-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0       2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1       2023-06-05 [1] CRAN (R 4.4.0)
   gridExtra                * 2.3         2017-09-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5       2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1       2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1     2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4       2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15      2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7       2023-08-15 [1] CRAN (R 4.4.0)
   igraph                     2.0.3       2024-03-13 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                      2.3.5.1     2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14      2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10      2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4       2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8       2023-12-04 [1] CRAN (R 4.4.0)
   KEGGREST                   1.44.0      2024-06-19 [1] Bioconductor 3.19 (R 4.4.0) (on disk 1.44.1)
   kernlab                    0.9-32      2023-01-31 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24     2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47        2024-05-29 [1] CRAN (R 4.4.0)
   later                      1.3.2       2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6      2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2       2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4       2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.2      2024-06-16 [1] Bioconductor 3.19 (R 4.4.0) (on disk 3.60.3)
   locfit                     1.5-9.9     2024-03-01 [1] CRAN (R 4.4.0)
   magick                     2.8.3       2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3       2022-03-30 [1] CRAN (R 4.4.0)
   mapproj                    1.2.11      2023-01-12 [1] CRAN (R 4.4.0)
   maps                       3.4.2       2023-12-15 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0       2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0       2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1       2021-11-26 [1] CRAN (R 4.4.0)
   metapod                    1.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   mime                       0.12        2021-09-28 [1] CRAN (R 4.4.0)
   MultiAssayExperiment       1.30.2      2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   munsell                    0.5.1       2024-04-01 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0       2024-01-21 [1] CRAN (R 4.4.0)
   pals                       1.8         2023-08-23 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1      2024-02-29 [1] CRAN (R 4.4.0)
   pillar                     1.9.0       2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3       2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4      2024-01-13 [1] CRAN (R 4.4.0)
   plyr                       1.8.9       2023-10-02 [1] CRAN (R 4.4.0)
   png                        0.1-8       2022-11-29 [1] CRAN (R 4.4.0)
   promises                   1.3.0       2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27      2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2       2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2       2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0      2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3      2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1       2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3       2021-01-31 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3       2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12      2024-01-09 [1] CRAN (R 4.4.0)
   RcppAnnoy                  0.0.22      2024-01-23 [1] CRAN (R 4.4.0)
   RcppHNSW                 * 0.6.0       2024-02-04 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14   2024-01-09 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2       2020-05-01 [1] CRAN (R 4.4.0)
   reshape2                   1.4.4       2020-04-09 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15      2022-06-16 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21      2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4       2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27        2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4       2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7       2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0      2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5       2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6       2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1       2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sass                       0.4.9       2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0       2023-11-28 [1] CRAN (R 4.4.0)
   scater                   * 1.32.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scran                    * 1.32.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scuttle                  * 1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2       2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16      2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4       2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1     2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6       2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4       2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0      2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8       2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1       2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2      2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1       2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5       2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0       2023-01-06 [1] CRAN (R 4.4.0)
   stringi                    1.8.4       2024-05-06 [1] CRAN (R 4.4.0)
   stringr                    1.5.1       2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78      2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1       2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1       2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1       2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12      2023-11-28 [1] CRAN (R 4.4.0)
   transport                  0.15-2      2024-05-10 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0       2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5       2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4       2023-10-22 [1] CRAN (R 4.4.0)
   uwot                       0.2.2       2024-04-21 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5       2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7       2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5       2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2       2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0       2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.1       2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45        2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.16.1 2024-01-22 [1] CRAN (R 4.4.0)
   xtable                     1.8-4       2019-04-21 [1] CRAN (R 4.4.0)
   XVector                  * 0.44.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8       2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0       2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
