---
title: "01-Cell Type Sex DE within all-types-smoothed VMH and ARC Domains"
output: html_document
date: "2024-07-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(dreamlet)
library(limma)
library(edgeR)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)

```


### we will use Yi's previous parameter sets for VMH/ARC smoothed domain definitions and subset to cells within those domains for sex DE analysis. since dual VMH-ARC labeled cells have to be competitively labeled, a given smoothing set generates unique labels for both domains, so for consistency with this, we will write scripts to run sex DE in both domains of a given domain labeling run.

```{r}
## read SFE and domain labels
hypb <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

## we used yi's parameters and deconflicted assignments in the same script/output file.
smoovedom <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/02a-VMHARC_domains_yi-thresholds.txt")

## load banksy cell typing results
bkcl <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")

## load cell type annots, append to banksy clustering results; discard clusters labeled "DISCARD"
bkanno <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
bkanno <- unique(bkanno[,.(clus,bjm_annot)])

bkcl <- merge(bkcl,bkanno,by.x="M0lam0_leiden_multisamp_res2",by.y="clus",all.x=TRUE)

## drop the DISCARDs from the cell types and the SFE
bkcl <- bkcl[!(bjm_annot %in% grep(bjm_annot,pattern="DISCARD",value=T))]
hypb <- hypb[,hypb$key %in% bkcl$rn]
stopifnot(ncol(hypb)==nrow(bkcl))

## label SFE with dual assigned VMH/ARC/neither, subset to VMH and separately to ARC
smoovedom <- DataFrame(smoovedom,row.names=smoovedom$rn)
smoovedom <- smoovedom[colnames(hypb),]
hypb$dualassignedSmooth2 <- smoovedom$dualassignedSmooth2

hyp.vdom <- hypb[,hypb$dualassignedSmooth2=="VMH_yi_2stage"]
hyp.adom <- hypb[,hypb$dualassignedSmooth2=="ARC_yi_2stage"]

## label cell clusters from banksy
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)

bkcl.vdom <- bkcl[colnames(hyp.vdom),]
bkcl.adom <- bkcl[colnames(hyp.adom),]

hyp.vdom$banksy_clus <- factor(bkcl.vdom$M0lam0_leiden_multisamp_res2,levels=c(paste0("X",c(1:41))))
hyp.vdom$banksy_clus <- droplevels(hyp.vdom$banksy_clus)

hyp.adom$banksy_clus <- factor(bkcl.adom$M0lam0_leiden_multisamp_res2,levels=c(paste0("X",c(1:41))))
hyp.adom$banksy_clus <- droplevels(hyp.adom$banksy_clus)

## remove misc.
rm(bkcl,bkcl.adom,bkcl.vdom,hypb,smoovedom,bkcl)
gc()
```

### use dreamlet to pseudobulk, then edgeR's voomLmFit to run sex DE (mirroring the procedures used for visium) for ARC. control for slide run (fall or winter).
```{r}
hyp.adom.ag <- dreamlet::aggregateToPseudoBulk(hyp.adom,assay="counts",sample_id="sample_id",cluster_id = "banksy_clus",BPPARAM = MulticoreParam(8))

# extract colData to feed into DGEList as the "samples" data
agcd <- as.data.table(colData(hyp.adom.ag),keep.rownames=T)
agcd[Date_Experiment_Began=="2/21/24",batch:="second"]
agcd[Date_Experiment_Began=="10/17/23",batch:="first"]
agcd <- as.data.frame(agcd)
rownames(agcd) <- agcd$rn
agcd$rn <- NULL

## only retain clusters for which at least 6 samples (size of the smaller sex group in total samps (females))) are available after pseudobulking+filtering through dreamlet and have at least 10,000 pseudobulked counts for the cluster of interest (or normalization and low-expression filtering gets to be impossible):
ag <- as.list(assays(hyp.adom.ag))
agc <- lapply(ag,FUN=function(a){ncol(a[,colSums(a)>10000])})
# no samples pass for X30, which breaks this bit of code. substitute in 0 for that slot
agc["X30"] <- 0
ag <- ag[!(agc<6)]

## drop clusters where there are 6+ samples, but only 0 or 1 samples for one of the sexes:
agsc <- lapply(ag,FUN=function(a){
 cn <- colnames(a[,colSums(a)>10000])
 tmp <- as.data.table(agcd[cn,])
 if(nrow(tmp[,.N,by="Sex"][N<2])>0){
   return(0)
 } else {return(1)}
})

ag <- ag[agsc==1]

## make the DGELists per remaining cluster:
dgels <- lapply(ag,FUN=function(a){
   # filtering for 10,000 counts minimum as above
 X <- a[,colSums(a)>10000]
 s <- copy(agcd)[colnames(X),]
 if(length(unique(s$Sex))<2){
   return()
 } else {
 
   dge <- DGEList(counts=as.matrix(X),samples=s)
   dge <- normLibSizes(dge,method = "TMM")
   
   ## find the smaller group for this cluster
   smolgrp <- table(dge$samples$Sex)[which.min(table(dge$samples$Sex))]
   
   ## for the smallest group, determine low expression cutoff
   rowmn.m <- as.numeric(rowMeans(cpm(dge[,(dge$samples$Sex==names(smolgrp)[1])],log=TRUE)))
   
    ## get histogram distribution with 10 bins at the low end of the logcpm space, skipping to the second minimum in the distribution (to avoid including 0-count genes etc). this is tricky because with the small number of genes, the distribution is rather peaky. however, some spot-checking to tune this looks like using 8 bins for the lower norm-sized-logcpm range (second minima + 1.5 logcpm) retrieves an accurate last-minimum-before-the-normalish-part-of-the-norm-count-distribution.
   # additionally, most genes fall in a small normalized library size window (6-10ish) -- part of the 10x panel design process when providing seq data from a similar sample (as we did, providing visium), which they use to tweak the number of probes per gene for detection based on the tissue's expression of the targets.
   h <- hist(rowmn.m[which(rowmn.m>min(rowmn.m)&rowmn.m<(1.5+min(rowmn.m)))],breaks=8,plot=F)
   ## get the minimum in this low logcpm range and its corresponding bin (x axis value, i.e. log cpm). in case of repeat values, grab the first (lowest)
   minc <- h$breaks[which(h$counts==min(h$counts))][1]
   
   ## get the corresponding 33rd %ile-library count value for this minimum^
   smollib <- min(dge$samples[(dge$samples$Sex==names(smolgrp)[1]),"lib.size"])
   countthresh <- (minc^2)*(as.numeric(smollib)/10^6)
   
   ## and convert that smallest count to the logcpm value in said smallest library to use as a low-expression threshold. for conservativeness, round up, since we're dealing with relatively small numbers of counts here
   cpmthresh <- log2(ceiling(countthresh)/(as.numeric(smollib)*10^-6))
   
   ## filter for genes with at least thresh counts in at least (smaller sex group size) samples
   keep <- rowSums(cpm(dge,log=T)>cpmthresh) >= as.numeric(smolgrp)
   
   ## recalc lib factors after dropping low-expressed genes
   dge <- dge[keep,,keep.lib.sizes=FALSE]
   dge <- calcNormFactors(dge,method = "TMM")
   
   return(dge)
 }
})
  
clus <- makeCluster(10,type = "PSOCK")
fulldgelres <- clusterApply(x=dgels,cl = clus,fun=function(X){
 
 library(data.table)
 setDTthreads(1,restore_after_fork = FALSE)
 library(SpatialExperiment)
 library(limma)
 library(edgeR)
 
 # prevent resource competition, though we aren't explicitly using DT at all here:
 setDTthreads(1) 
 
 dge <- X
 des <- model.matrix(~0+Sex+batch, data = dge$samples)
 # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
 sexdif <- makeContrasts(SexMale-SexFemale,levels=des)
 
 # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
 v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$BrNum),adaptive.span = T,sample.weights = T)
 
 v.swt.fit <- contrasts.fit(v.swt,sexdif)
 
 v.swt.fit.e <- eBayes(v.swt.fit)
 
 v.swt.e.tt <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
 
 return(v.swt.e.tt)
})

stopCluster(clus)
names(fulldgelres) <- names(ag)

### concatenate results for all cell types
fulldgelres <- lapply(fulldgelres,as.data.table,keep.rownames=T)
fulldgelres.arc <- rbindlist(fulldgelres,idcol="cluster")
setnames(fulldgelres.arc,"rn","gene_name")

## append cell type annotations
fulldgelres.arc <- merge.data.table(fulldgelres.arc,bkanno,by.x="cluster",by.y="clus",all.x=T)
setnames(fulldgelres.arc,c("bjm_annot","cluster"),c("cluster_annotated","cluster_id"))
fulldgelres.arc <- fulldgelres.arc[,.(cluster_annotated,gene_name,logFC,adj.P.Val,t,B,P.Value,AveExpr,cluster_id)]

### clean up
rm(ag,agc,agcd,agsc,clus,dgels,fulldgelres,hyp.adom,hyp.adom.ag)
gc(full=T)
```

### repeat as above for VMH
```{r}
hyp.vdom.ag <- dreamlet::aggregateToPseudoBulk(hyp.vdom,assay="counts",sample_id="sample_id",cluster_id = "banksy_clus",BPPARAM = MulticoreParam(8))

# extract colData to feed into DGEList as the "samples" data
agcd <- as.data.table(colData(hyp.vdom.ag),keep.rownames=T)
agcd[Date_Experiment_Began=="2/21/24",batch:="second"]
agcd[Date_Experiment_Began=="10/17/23",batch:="first"]
agcd <- as.data.frame(agcd)
rownames(agcd) <- agcd$rn
agcd$rn <- NULL

## only retain clusters for which at least 6 samples (size of the smaller sex group in total samps (females))) are available after pseudobulking+filtering through dreamlet and have at least 10,000 pseudobulked counts for the cluster of interest (or normalization and low-expression filtering gets to be impossible):
ag <- as.list(assays(hyp.vdom.ag))
agc <- lapply(ag,FUN=function(a){ncol(a[,colSums(a)>10000])})
# no samples pass for X37, which breaks this bit of code. substitute in 0 for that slot
agc["X37"] <- 0
ag <- ag[!(agc<6)]

## drop clusters where there are 6+ samples, but only 0 or 1 samples for one of the sexes:
agsc <- lapply(ag,FUN=function(a){
 cn <- colnames(a[,colSums(a)>10000])
 tmp <- as.data.table(agcd[cn,])
 if(nrow(tmp[,.N,by="Sex"][N<2])>0){
   return(0)
 } else {return(1)}
})

ag <- ag[agsc==1]

## make the DGELists per remaining cluster:
dgels <- lapply(ag,FUN=function(a){
   # filtering for 10,000 counts minimum as above
 X <- a[,colSums(a)>10000]
 s <- copy(agcd)[colnames(X),]
 if(length(unique(s$Sex))<2){
   return()
 } else {
 
   dge <- DGEList(counts=as.matrix(X),samples=s)
   dge <- normLibSizes(dge,method = "TMM")
   
   ## find the smaller group for this cluster
   smolgrp <- table(dge$samples$Sex)[which.min(table(dge$samples$Sex))]
   
   ## for the smallest group, determine low expression cutoff
   rowmn.m <- as.numeric(rowMeans(cpm(dge[,(dge$samples$Sex==names(smolgrp)[1])],log=TRUE)))
   
    ## get histogram distribution with 10 bins at the low end of the logcpm space, skipping to the second minimum in the distribution (to avoid including 0-count genes etc). this is tricky because with the small number of genes, the distribution is rather peaky. however, some spot-checking to tune this looks like using 8 bins for the lower norm-sized-logcpm range (second minima + 1.5 logcpm) retrieves an accurate last-minimum-before-the-normalish-part-of-the-norm-count-distribution.
   # additionally, most genes fall in a small normalized library size window (6-10ish) -- part of the 10x panel design process when providing seq data from a similar sample (as we did, providing visium), which they use to tweak the number of probes per gene for detection based on the tissue's expression of the targets.
   h <- hist(rowmn.m[which(rowmn.m>min(rowmn.m)&rowmn.m<(1.5+min(rowmn.m)))],breaks=8,plot=F)
   ## get the minimum in this low logcpm range and its corresponding bin (x axis value, i.e. log cpm). in case of repeat values, grab the first (lowest)
   minc <- h$breaks[which(h$counts==min(h$counts))][1]
   
   ## get the corresponding 33rd %ile-library count value for this minimum^
   smollib <- min(dge$samples[(dge$samples$Sex==names(smolgrp)[1]),"lib.size"])
   countthresh <- (minc^2)*(as.numeric(smollib)/10^6)
   
   ## and convert that smallest count to the logcpm value in said smallest library to use as a low-expression threshold. for conservativeness, round up, since we're dealing with relatively small numbers of counts here
   cpmthresh <- log2(ceiling(countthresh)/(as.numeric(smollib)*10^-6))
   
   ## filter for genes with at least thresh counts in at least (smaller sex group size) samples
   keep <- rowSums(cpm(dge,log=T)>cpmthresh) >= as.numeric(smolgrp)
   
   ## recalc lib factors after dropping low-expressed genes
   dge <- dge[keep,,keep.lib.sizes=FALSE]
   dge <- calcNormFactors(dge,method = "TMM")
   
   return(dge)
 }
})
  
clus <- makeCluster(10,type = "PSOCK")
fulldgelres.vmh <- clusterApply(x=dgels,cl = clus,fun=function(X){
 
 library(data.table)
 setDTthreads(1,restore_after_fork = FALSE)
 library(SpatialExperiment)
 library(limma)
 library(edgeR)
 
 # prevent resource competition, though we aren't explicitly using DT at all here:
 setDTthreads(1) 
 
 dge <- X
 des <- model.matrix(~0+Sex+batch, data = dge$samples)
 # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
 sexdif <- makeContrasts(SexMale-SexFemale,levels=des)
 
 # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
 v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$BrNum),adaptive.span = T,sample.weights = T)
 
 v.swt.fit <- contrasts.fit(v.swt,sexdif)
 
 v.swt.fit.e <- eBayes(v.swt.fit)
 
 v.swt.e.tt <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
 
 return(v.swt.e.tt)
})

stopCluster(clus)
names(fulldgelres.vmh) <- names(ag)

### concatenate results for all cell types
fulldgelres.vmh <- lapply(fulldgelres.vmh,as.data.table,keep.rownames=T)
fulldgelres.vmh <- rbindlist(fulldgelres.vmh,idcol="cluster")
setnames(fulldgelres.vmh,"rn","gene_name")

## append cell type annotations
fulldgelres.vmh <- merge.data.table(fulldgelres.vmh,bkanno,by.x="cluster",by.y="clus",all.x=T)
setnames(fulldgelres.vmh,c("bjm_annot","cluster"),c("cluster_annotated","cluster_id"))
fulldgelres.vmh <- fulldgelres.vmh[,.(cluster_annotated,gene_name,logFC,adj.P.Val,t,B,P.Value,AveExpr,cluster_id)]
```

merge and save
```{r}
setnames(fulldgelres.arc,c("logFC","adj.P.Val","t"),c("logFC_MvF","FDR","t_stat"))
setnames(fulldgelres.vmh,c("logFC","adj.P.Val","t"),c("logFC_MvF","FDR","t_stat"))

## save
fwrite(fulldgelres.arc,"processed-data/08_VMH-ARC cell type sex DE within domains/01a-withinARC_celltypeSexDE.txt")
fwrite(fulldgelres.arc,"processed-data/08_VMH-ARC cell type sex DE within domains/01b-withinVMH_celltypeSexDE.txt")

```


## reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     methods   base     

other attached packages:
 [1] edgeR_4.2.0                    dreamlet_1.2.1                
 [3] variancePartition_1.34.0       limma_3.60.3                  
 [5] BiocParallel_1.38.0            parallelly_1.38.0             
 [7] spatialLIBD_1.16.2             SpatialFeatureExperiment_1.6.1
 [9] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[11] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[13] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
[15] IRanges_2.38.1                 S4Vectors_0.42.1              
[17] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[19] matrixStats_1.3.0              gridExtra_2.3                 
[21] colorout_1.3-0.2               ggrastr_1.0.2                 
[23] ggplot2_3.5.1                  sf_1.0-16                     
[25] data.table_1.15.4              rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] bitops_1.0-7              EBImage_4.46.0            httr_1.4.7               
  [4] RColorBrewer_1.1-3        doParallel_1.0.17         Rgraphviz_2.48.0         
  [7] numDeriv_2016.8-1.1       backports_1.5.0           tools_4.4.1              
 [10] metafor_4.6-0             utf8_1.2.4                R6_2.5.1                 
 [13] DT_0.33                   HDF5Array_1.32.0          lazyeval_0.2.2           
 [16] rhdf5filters_1.16.0       withr_3.0.0               sp_2.1-4                 
 [19] prettyunits_1.2.0         cli_3.6.3                 labeling_0.4.3           
 [22] sass_0.4.9                KEGGgraph_1.64.0          SQUAREM_2021.1           
 [25] mvtnorm_1.2-5             proxy_0.4-27              mixsqp_0.3-54            
 [28] Rsamtools_2.20.0          R.utils_2.12.3            zenith_1.6.0             
 [31] scater_1.32.0             sessioninfo_1.2.2         attempt_0.3.1            
 [34] invgamma_1.1              maps_3.4.2                rstudioapi_0.16.0        
 [37] RSQLite_2.3.7             generics_0.1.3            BiocIO_1.14.0            
 [40] gtools_3.9.5              spdep_1.3-5               dplyr_1.1.4              
 [43] metadat_1.2-0             Matrix_1.7-0              ggbeeswarm_0.7.2         
 [46] fansi_1.0.6               abind_1.4-5               R.methodsS3_1.8.2        
 [49] terra_1.7-78              lifecycle_1.0.4           yaml_2.3.9               
 [52] mathjaxr_1.6-0            gplots_3.1.3.1            rhdf5_2.48.0             
 [55] SparseArray_1.4.8         BiocFileCache_2.12.0      paletteer_1.6.0          
 [58] grid_4.4.1                blob_1.2.4                promises_1.3.0           
 [61] dqrng_0.4.1               ExperimentHub_2.12.0      crayon_1.5.3             
 [64] lattice_0.22-6            msigdbr_7.5.1             beachmat_2.20.0          
 [67] cowplot_1.1.3             annotate_1.82.0           KEGGREST_1.44.1          
 [70] magick_2.8.3              zeallot_0.1.0             pillar_1.9.0             
 [73] knitr_1.48                rjson_0.2.21              boot_1.3-30              
 [76] corpcor_1.6.10            codetools_0.2-20          wk_0.9.2                 
 [79] glue_1.7.0                vctrs_0.6.5               png_0.1-8                
 [82] spam_2.10-0               Rdpack_2.6                gtable_0.3.5             
 [85] assertthat_0.2.1          rematch2_2.1.2            cachem_1.1.0             
 [88] xfun_0.45                 rbibutils_2.2.16          S4Arrays_1.4.1           
 [91] mime_0.12                 DropletUtils_1.24.0       Rfast_2.1.0              
 [94] sfheaders_0.4.4           iterators_1.0.14          units_0.8-5              
 [97] fields_16.2               statmod_1.5.0             nlme_3.1-165             
[100] pbkrtest_0.5.3            bit64_4.0.5               progress_1.2.3           
[103] EnvStats_2.8.1            filelock_1.0.3            rprojroot_2.0.4          
[106] bslib_0.7.0               irlba_2.3.5.1             vipor_0.4.7              
[109] KernSmooth_2.23-24        rmeta_3.0                 colorspace_2.1-0         
[112] spData_2.3.1              DBI_1.2.3                 tidyselect_1.2.1         
[115] bit_4.0.5                 compiler_4.4.1            curl_5.2.1               
[118] graph_1.82.0              BiocNeighbors_1.22.0      DelayedArray_0.30.1      
[121] plotly_4.10.4             rtracklayer_1.64.0        scales_1.3.0             
[124] caTools_1.18.2            remaCor_0.0.18            classInt_0.4-10          
[127] rappdirs_0.3.3            stringr_1.5.1             tiff_0.1-12              
[130] digest_0.6.36             minqa_1.2.7               fftwtools_0.9-11         
[133] rmarkdown_2.27            benchmarkmeData_1.0.4     aod_1.3.3                
[136] XVector_0.44.0            RhpcBLASctl_0.23-42       htmltools_0.5.8.1        
[139] pkgconfig_2.0.3           jpeg_0.1-10               lme4_1.1-35.5            
[142] sparseMatrixStats_1.16.0  mashr_0.2.79              dbplyr_2.5.0             
[145] fastmap_1.2.0             htmlwidgets_1.6.4         UCSC.utils_1.0.0         
[148] shiny_1.8.1.1             DelayedMatrixStats_1.26.0 farver_2.1.2             
[151] jquerylib_0.1.4           jsonlite_1.8.8            config_0.3.2             
[154] R.oo_1.26.0               BiocSingular_1.20.0       RCurl_1.98-1.14          
[157] magrittr_2.0.3            scuttle_1.14.0            GenomeInfoDbData_1.2.12  
[160] s2_1.1.6                  dotCall64_1.1-1           Rhdf5lib_1.26.0          
[163] munsell_0.5.1             Rcpp_1.0.13               babelgene_22.9           
[166] viridis_0.6.5             RcppZiggurat_0.1.6        EnrichmentBrowser_2.34.1 
[169] stringi_1.8.4             zlibbioc_1.50.0           MASS_7.3-61              
[172] plyr_1.8.9                AnnotationHub_3.12.0      ggrepel_0.9.5            
[175] deldir_2.0-4              splines_4.4.1             Biostrings_2.72.1        
[178] hms_1.1.3                 locfit_1.5-9.10           reshape2_1.4.4           
[181] ScaledMatrix_1.12.0       BiocVersion_3.19.1        XML_3.99-0.17            
[184] evaluate_0.24.0           golem_0.4.1               RcppParallel_5.1.8       
[187] BiocManager_1.30.23       nloptr_2.1.1              foreach_1.5.2            
[190] httpuv_1.6.15             tidyr_1.3.1               purrr_1.0.2              
[193] scattermore_1.2           benchmarkme_1.0.8         ashr_2.2-63              
[196] rsvd_1.0.5                broom_1.0.6               xtable_1.8-4             
[199] restfulr_0.0.15           fANCOVA_0.6-1             e1071_1.7-14             
[202] later_1.3.2               viridisLite_0.4.2         class_7.3-22             
[205] truncnorm_1.0-9           tibble_3.2.1              lmerTest_3.1-3           
[208] memoise_2.0.1             beeswarm_0.4.0            AnnotationDbi_1.66.0     
[211] GenomicAlignments_1.40.0  shinyWidgets_0.8.6        GSEABase_1.66.0          
[214] here_1.0.1               
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package                  * version    date (UTC) lib source
   abind                      1.4-5      2016-07-21 [1] CRAN (R 4.4.0)
   annotate                   1.82.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationDbi              1.66.0     2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationHub              3.12.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   aod                        1.3.3      2023-12-13 [1] CRAN (R 4.4.0)
   ashr                       2.2-63     2023-08-21 [1] CRAN (R 4.4.0)
   assertthat                 0.2.1      2019-03-21 [1] CRAN (R 4.4.0)
   attempt                    0.3.1      2020-05-03 [1] CRAN (R 4.4.0)
   babelgene                  22.9       2022-09-29 [1] CRAN (R 4.4.0)
   backports                  1.5.0      2024-05-23 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0      2021-06-01 [1] CRAN (R 4.4.0)
   benchmarkme                1.0.8      2022-06-12 [1] CRAN (R 4.4.0)
   benchmarkmeData            1.0.4      2020-04-23 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocFileCache              2.12.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocIO                     1.14.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocManager                1.30.23    2024-05-04 [1] CRAN (R 4.4.0)
   BiocNeighbors              1.22.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocSingular               1.20.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocVersion                3.19.1     2024-04-22 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1     2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5      2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5      2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7      2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4      2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30     2024-02-26 [1] CRAN (R 4.4.1)
   broom                      1.0.6      2024-05-17 [1] CRAN (R 4.4.0)
   bslib                      0.7.0      2024-03-29 [1] CRAN (R 4.4.0)
   cachem                     1.1.0      2024-05-16 [1] CRAN (R 4.4.0)
   caTools                    1.18.2     2021-03-28 [1] CRAN (R 4.4.0)
   class                      7.3-22     2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10     2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3      2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20     2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2    2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0      2023-01-23 [1] CRAN (R 4.4.0)
   config                     0.3.2      2023-08-30 [1] CRAN (R 4.4.0)
   corpcor                    1.6.10     2021-09-16 [1] CRAN (R 4.4.0)
   cowplot                    1.1.3      2024-01-22 [1] CRAN (R 4.4.0)
   crayon                     1.5.3      2024-06-20 [1] CRAN (R 4.4.0)
   curl                       5.2.1      2024-03-01 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4     2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3      2024-06-02 [1] CRAN (R 4.4.0)
   dbplyr                     2.5.0      2024-03-19 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1     2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4      2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36     2024-06-23 [1] CRAN (R 4.4.0)
   doParallel                 1.0.17     2022-02-07 [1] CRAN (R 4.4.1)
   dotCall64                  1.1-1      2023-11-28 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4      2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1      2024-05-28 [1] CRAN (R 4.4.0)
   dreamlet                 * 1.2.1      2024-06-09 [1] Bioconductor 3.19 (R 4.4.0)
   DropletUtils               1.24.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   DT                         0.33       2024-04-04 [1] CRAN (R 4.4.0)
   e1071                      1.7-14     2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                    * 4.2.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   EnrichmentBrowser          2.34.1     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   EnvStats                   2.8.1      2023-08-22 [1] CRAN (R 4.4.0)
   evaluate                   0.24.0     2024-06-10 [1] CRAN (R 4.4.0)
   ExperimentHub              2.12.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   fANCOVA                    0.6-1      2020-11-13 [1] CRAN (R 4.4.0)
   fansi                      1.0.6      2023-12-08 [1] CRAN (R 4.4.0)
   farver                     2.1.2      2024-05-13 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0      2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11     2021-03-01 [1] CRAN (R 4.4.0)
   fields                     16.2       2024-06-27 [1] CRAN (R 4.4.0)
   filelock                   1.0.3      2023-12-11 [1] CRAN (R 4.4.0)
   foreach                    1.5.2      2022-02-02 [1] CRAN (R 4.4.0)
   generics                   0.1.3      2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1     2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12     2024-05-01 [1] Bioconductor
   GenomicAlignments          1.40.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GenomicRanges            * 1.56.1     2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2      2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1      2024-04-23 [1] CRAN (R 4.4.0)
   ggrastr                  * 1.0.2      2023-06-01 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5      2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0      2024-01-09 [1] CRAN (R 4.4.0)
   golem                      0.4.1      2023-06-05 [1] CRAN (R 4.4.0)
   gplots                     3.1.3.1    2024-02-02 [1] CRAN (R 4.4.0)
   graph                      1.82.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   gridExtra                * 2.3        2017-09-09 [1] CRAN (R 4.4.0)
   GSEABase                   1.66.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   gtable                     0.3.5      2024-04-22 [1] CRAN (R 4.4.0)
   gtools                     3.9.5      2023-11-20 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1      2020-12-13 [1] CRAN (R 4.4.0)
   hms                        1.1.3      2023-03-21 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1    2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4      2023-12-06 [1] CRAN (R 4.4.0)
   httpuv                     1.6.15     2024-03-26 [1] CRAN (R 4.4.0)
   httr                       1.4.7      2023-08-15 [1] CRAN (R 4.4.0)
   invgamma                   1.1        2017-05-07 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1     2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                      2.3.5.1    2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14     2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10     2022-11-29 [1] CRAN (R 4.4.0)
   jquerylib                  0.1.4      2021-04-26 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8      2023-12-04 [1] CRAN (R 4.4.0)
   KEGGgraph                  1.64.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   KEGGREST                   1.44.1     2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24    2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48       2024-07-07 [1] CRAN (R 4.4.1)
   labeling                   0.4.3      2023-08-29 [1] CRAN (R 4.4.0)
   later                      1.3.2      2023-12-06 [1] CRAN (R 4.4.0)
   lattice                    0.22-6     2024-03-20 [1] CRAN (R 4.4.1)
   lazyeval                   0.2.2      2019-03-15 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4      2023-11-07 [1] CRAN (R 4.4.0)
   limma                    * 3.60.3     2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   lme4                       1.1-35.5   2024-07-03 [1] CRAN (R 4.4.1)
   lmerTest                   3.1-3      2020-10-23 [1] CRAN (R 4.4.0)
   locfit                     1.5-9.10   2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3      2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3      2022-03-30 [1] CRAN (R 4.4.0)
   maps                       3.4.2      2023-12-15 [1] CRAN (R 4.4.0)
   mashr                      0.2.79     2023-10-18 [1] CRAN (R 4.4.0)
   MASS                       7.3-61     2024-06-13 [1] CRAN (R 4.4.0)
   mathjaxr                   1.6-0      2022-02-28 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0      2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0      2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1      2021-11-26 [1] CRAN (R 4.4.0)
   metadat                    1.2-0      2022-04-06 [1] CRAN (R 4.4.0)
   metafor                    4.6-0      2024-03-28 [1] CRAN (R 4.4.0)
   mime                       0.12       2021-09-28 [1] CRAN (R 4.4.0)
   minqa                      1.2.7      2024-05-20 [1] CRAN (R 4.4.0)
   mixsqp                     0.3-54     2023-12-20 [1] CRAN (R 4.4.0)
   msigdbr                    7.5.1      2022-03-30 [1] CRAN (R 4.4.0)
   munsell                    0.5.1      2024-04-01 [1] CRAN (R 4.4.0)
   mvtnorm                    1.2-5      2024-05-21 [1] CRAN (R 4.4.0)
   nlme                       3.1-165    2024-06-06 [1] CRAN (R 4.4.0)
   nloptr                     2.1.1      2024-06-25 [1] CRAN (R 4.4.0)
   numDeriv                   2016.8-1.1 2019-06-06 [1] CRAN (R 4.4.0)
   paletteer                  1.6.0      2024-01-21 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0     2024-07-27 [1] CRAN (R 4.4.0)
   pbkrtest                   0.5.3      2024-06-26 [1] CRAN (R 4.4.0)
   pillar                     1.9.0      2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3      2019-09-22 [1] CRAN (R 4.4.0)
   plotly                     4.10.4     2024-01-13 [1] CRAN (R 4.4.0)
   plyr                       1.8.9      2023-10-02 [1] CRAN (R 4.4.0)
   png                        0.1-8      2022-11-29 [1] CRAN (R 4.4.0)
   prettyunits                1.2.0      2023-09-24 [1] CRAN (R 4.4.0)
   progress                   1.2.3      2023-12-06 [1] CRAN (R 4.4.0)
   promises                   1.3.0      2024-04-05 [1] CRAN (R 4.4.0)
   proxy                      0.4-27     2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2      2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2      2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0     2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3     2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1      2021-08-19 [1] CRAN (R 4.4.0)
   rappdirs                   0.3.3      2021-01-31 [1] CRAN (R 4.4.0)
   rbibutils                  2.2.16     2023-10-25 [1] CRAN (R 4.4.0)
   RColorBrewer               1.1-3      2022-04-03 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13     2024-07-17 [1] CRAN (R 4.4.0)
   RcppParallel               5.1.8      2024-07-06 [1] CRAN (R 4.4.0)
   RcppZiggurat               0.1.6      2020-10-20 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14  2024-01-09 [1] CRAN (R 4.4.0)
   Rdpack                     2.6        2023-11-08 [1] CRAN (R 4.4.0)
   remaCor                    0.0.18     2024-02-08 [1] CRAN (R 4.4.0)
   rematch2                   2.1.2      2020-05-01 [1] CRAN (R 4.4.0)
   reshape2                   1.4.4      2020-04-09 [1] CRAN (R 4.4.0)
   restfulr                   0.0.15     2022-06-16 [1] CRAN (R 4.4.0)
   Rfast                      2.1.0      2023-11-09 [1] CRAN (R 4.4.0)
   Rgraphviz                  2.48.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5                      2.48.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RhpcBLASctl                0.23-42    2023-02-11 [1] CRAN (R 4.4.0)
   rjson                      0.2.21     2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4      2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27       2024-05-17 [1] CRAN (R 4.4.0)
   rmeta                      3.0        2018-03-20 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4      2023-11-05 [1] CRAN (R 4.4.0)
   Rsamtools                  2.20.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RSQLite                    2.3.7      2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0     2024-03-24 [1] CRAN (R 4.4.0)
   rsvd                       1.0.5      2021-04-16 [1] CRAN (R 4.4.0)
   rtracklayer                1.64.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   s2                         1.1.6      2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1      2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1     2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   sass                       0.4.9      2024-03-15 [1] CRAN (R 4.4.0)
   ScaledMatrix               1.12.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0      2023-11-28 [1] CRAN (R 4.4.0)
   scater                     1.32.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scattermore                1.2        2023-06-12 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2      2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16     2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4      2024-01-17 [1] CRAN (R 4.4.0)
   shiny                      1.8.1.1    2024-04-02 [1] CRAN (R 4.4.0)
   shinyWidgets               0.8.6      2024-04-24 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4      2024-04-30 [1] CRAN (R 4.4.0)
   spam                       2.10-0     2023-10-23 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8      2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1      2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spatialLIBD              * 1.16.2     2024-05-28 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1      2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5      2024-06-10 [1] CRAN (R 4.4.0)
   SQUAREM                    2021.1     2021-01-13 [1] CRAN (R 4.4.0)
   statmod                    1.5.0      2023-01-06 [1] CRAN (R 4.4.0)
   stringi                    1.8.4      2024-05-06 [1] CRAN (R 4.4.0)
   stringr                    1.5.1      2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78     2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1      2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1      2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1      2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12     2023-11-28 [1] CRAN (R 4.4.0)
   truncnorm                  1.0-9      2023-03-20 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5      2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4      2023-10-22 [1] CRAN (R 4.4.0)
   variancePartition        * 1.34.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   vctrs                      0.6.5      2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7      2023-12-18 [1] CRAN (R 4.4.0)
   viridis                    0.6.5      2024-01-29 [1] CRAN (R 4.4.0)
   viridisLite                0.4.2      2023-05-02 [1] CRAN (R 4.4.0)
   withr                      3.0.0      2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2      2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45       2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17  2024-06-25 [1] CRAN (R 4.3.3)
   xtable                     1.8-4      2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9      2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0      2018-01-28 [1] CRAN (R 4.4.0)
   zenith                     1.6.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   zlibbioc                   1.50.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
