---
title: "05-Cell Type Sex DE within interior hi-conf 4-type-ARC and VMH Domains"
output: html_document
date: "2024-08-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(sf)
library(dreamlet)
library(limma)
library(edgeR)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallel)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(parallelly)
options(bphost="localhost")
library(BiocParallel)

```

### As one of two confirmatory analyses, we will restrict the analyses to the ~96% interior boundary of the ARC and VMH domains to avoid potentially overlapping cell types in their boundary areas. We can get this by using the buffer function of sf, which with a negative setting, will decrease the size of the polygonal domain.

### From here forward, we will only perform sex DE analyses using the more constrained ARC definition that only considered cell types 15/19/31/33 in the smoothing, and VMH as defined in this scheme.

```{r}
## read SFE and domain labels
hypb <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")


finesmoove <- fread("processed-data/07_Domains-subdomains_by_knnSmoothing_Banksytypes/03b-ARCVMHdomains_2stepsmooth_VMH-k10-0.2-VMH-k200-0.2_ARC-k50-0.1_ARC-k500-0.5.txt")

## since there are clusters wwe are discarding, we should drop those before proceeding lest hull boundary determinations are affected by such cells

## load banksy cell typing results
bkcl <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")

## load cell type annots, append to banksy clustering results; discard clusters labeled "DISCARD"
bkanno <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/02-M0l0kg6_topClusMarkers_celltypes_annotated.txt")
bkanno[,clus:=paste0("X",clus)]
bkanno <- unique(bkanno[,.(clus,bjm_annot)])

bkcl <- merge(bkcl,bkanno,by.x="M0lam0_leiden_multisamp_res2",by.y="clus",all.x=TRUE)

## drop the DISCARDs from the cell types and the SFE, as well as the cell-coord-domain table (so that we smooth only the non-discarded cells). we'll finish setup, including appending the domain labels and cell types, after we calculate the interior domain polygons.
bkcl <- bkcl[bjm_annot!="DISCARD",]
hypb <- hypb[,hypb$key %in% bkcl$rn]
stopifnot(ncol(hypb)==nrow(bkcl))

finesmoove <- finesmoove[rn %in% bkcl$rn]
```

### now, split the smoothing results by sample to define the "high-confidence" domain areas. the VMH is shaped such that convex hull determination works fine, but the ARC has concave portions in many samples, so use the concave hull function with ratio (lower --> concavity favoring) set to 0.1. (We used this previously in determining boundary domains from domain hulls in code / 07 / 02b, 02c, and 03c)
```{r}
samplist <- split(finesmoove,finesmoove$sample_id)

samplist <- bplapply(samplist,BPPARAM=MulticoreParam(10),FUN=function(s){
   vmh <- st_as_sf(s[dualVMHARC4=="VMH"],coords=c("xcoord","ycoord"))
   arc <- st_as_sf(s[dualVMHARC4=="ARC"],coords=c("xcoord","ycoord"))
   
   ## get the points defining the convex hull using sf by running st_union to make a single geometry of all component points, then st_convex_hull to get the hull
    vmh.sf <- st_as_sf(vmh,coords=c("xcoord","ycoord"))
    vmh.hull <- st_convex_hull(st_union(vmh.sf))
    
    ## arc is not necessarily a CONVEX hull, however. e.g, it's concave where it meets the VMH (in some cases). again, using X86_reg1, calling st_concave_hull with ratio 0.1 (concavity vs convexity parameter, lower = more concavity; 0 just gives a polygon tracing all points sequentially and is not useful) works beautifully, and this is probably the scoop-iest shaped ARC in the dataset. so let's try that for all samples
    arc.sf <- st_as_sf(arc,coords=c("xcoord","ycoord"))
    arc.hull <- st_concave_hull(st_union(arc.sf),ratio=0.1)
    
    ## add a NEGATIVE buffer zone the domain hull polygons, ie shrinking them inward. we previously benchmarked using a positive buffer for boundary detection by adding about 4% to the hull approximated by calculating the median of the pairwise distances between points defining the domain hull. so here, we will use the negative of that to shrink inwards by such a value.
    
    vmh.buffdist <- median(dist(st_coordinates(vmh.hull)))
    vmh.buffhull <- st_buffer(vmh.hull,dist=-0.04*vmh.buffdist)

    arc.buffdist <- median(dist(st_coordinates(arc.hull)))
    arc.buffhull <- st_buffer(arc.hull,dist=-0.04*arc.buffdist)
    
    # get the polygon defining where these negative-buffer hull overlaps the original, and the cell ids within it. these are what we want to subset to for "high-confidence domain" sex DE. 
    vmh.buffintsec <- st_intersection(vmh.buffhull,vmh.hull)
    arc.buffintsec <- st_intersection(arc.buffhull,arc.hull)

    ## really, truly finally, get the cells that (still) fall within the shrunken domain. st_within returns a list of 0/1 for within/not within per row, which is easier to use/spot check with sparse=F as all rows are accounted for in its output
    vmh.inboundary <- st_within(vmh.sf,vmh.buffintsec,sparse = F)
    vmh.innercells <- as.data.frame(vmh)[vmh.inboundary,"rn"]
    
    arc.inboundary <- st_within(arc.sf,arc.buffintsec,sparse=F)
    arc.innercells <- as.data.frame(arc)[arc.inboundary,"rn"]
    return(list(vmhinnercells=vmh.innercells,arcinnercells=arc.innercells))
})

### concatenate back to single vectors of VMH/ARC inner cells
vmhin <- unlist(lapply(samplist,function(x){x$vmhinnercells}))
arcin <- unlist(lapply(samplist,function(x){x$arcinnercells}))

### make sure there aren't conflicting cells after this (there shouldn't be--but if we're getting back erroneously convex hulls for ARC from the code above, then its boundary is going to cut at a weird diagonal straight across VMH, and cells would overlap then)
stopifnot(length(intersect(vmhin,arcin))==0)

### annotate which cells are in the hiconf domains in a new column of the smoothing result table, which we will use to subset the SFE into solely the cells of each domain
finesmoove[rn %in% vmhin,hiconfdom:="VMH"]
finesmoove[rn %in% arcin,hiconfdom:="ARC"]
finesmoove[is.na(hiconfdom),hiconfdom:="other"]
```


### finish sex DE setup by appending the cell type and domain labels to the SFE, and then subset to the high-confidence domains
```{r}
## create separate SFEs for 3-core-cell-type ARC domain and the VMH domain
finesmoove <- DataFrame(finesmoove,row.names=finesmoove$rn)
finesmoove <- finesmoove[colnames(hypb),]
hypb$hiconfdom <- finesmoove$hiconfdom

hyp.fineadom <- hypb[,hypb$hiconfdom=="ARC"]
stopifnot(ncol(hyp.fineadom)==sum(finesmoove$hiconfdom=="ARC"))

hyp.finevdom <- hypb[,hypb$hiconfdom=="VMH"]
stopifnot(ncol(hyp.finevdom)==sum(finesmoove$hiconfdom=="VMH"))

## label cell clusters from banksy for both, order levels 1-41, then drop nonapplicable cell types. (so the cell types are in numerical order)
bkcl <- DataFrame(bkcl,row.names=bkcl$rn)
bkcl.fineadom <- bkcl[bkcl$rn %in% colnames(hyp.fineadom),]
bkcl.fineadom <- bkcl.fineadom[colnames(hyp.fineadom),]
hyp.fineadom$banksy_clus <- factor(bkcl.fineadom$M0lam0_leiden_multisamp_res2,levels=c(paste0("X",c(1:41))))
hyp.fineadom$banksy_clus <- droplevels(hyp.fineadom$banksy_clus)

bkcl.finevdom <- bkcl[bkcl$rn %in% colnames(hyp.finevdom),]
bkcl.finevdom <- bkcl.finevdom[colnames(hyp.finevdom),]
hyp.finevdom$banksy_clus <- factor(bkcl.finevdom$M0lam0_leiden_multisamp_res2,levels=c(paste0("X",c(1:41))))
hyp.finevdom$banksy_clus <- droplevels(hyp.finevdom$banksy_clus)

## remove misc.
rm(bkcl,hypb,bkcl.fineadom,bkcl.finevdom,finesmoove,samplist,vmhin,arcin)
gc()
```

### use dreamlet to pseudobulk, then edgeR's voomLmFit to run sex DE (mirroring the procedures used for visium) for ARC. control for slide run (fall or winter).
```{r}
hyp.adom.ag <- dreamlet::aggregateToPseudoBulk(hyp.fineadom,assay="counts",sample_id="sample_id",cluster_id = "banksy_clus",BPPARAM = MulticoreParam(8))

# extract colData to feed into DGEList as the "samples" data
agcd <- as.data.table(colData(hyp.adom.ag),keep.rownames=T)
agcd[Date_Experiment_Began=="2/21/24",batch:="second"]
agcd[Date_Experiment_Began=="10/17/23",batch:="first"]
agcd <- as.data.frame(agcd)
rownames(agcd) <- agcd$rn
agcd$rn <- NULL

## only retain clusters for which at least 6 samples (size of the smaller sex group in total samps (females))) are available after pseudobulking+filtering through dreamlet and have at least 10,000 pseudobulked counts for the cluster of interest (or normalization and low-expression filtering gets to be impossible):
ag <- as.list(assays(hyp.adom.ag))
agc <- lapply(ag,FUN=function(a){ncol(a[,colSums(a)>10000])})
# no samples pass for X35, which breaks this bit of code. substitute in 0 for that slot
agc["X35"] <- 0
ag <- ag[!(agc<6)]

## drop clusters where there are 6+ samples, but only 0 or 1 samples for one of the sexes:
agsc <- lapply(ag,FUN=function(a){
 cn <- colnames(a[,colSums(a)>10000])
 tmp <- as.data.table(agcd[cn,])
 if(nrow(tmp[,.N,by="Sex"][N<2])>0){
   return(0)
 } else {return(1)}
})

ag <- ag[agsc==1]

## make the DGELists per remaining cluster:
dgels <- lapply(ag,FUN=function(a){
   # filtering for 10,000 counts minimum as above
 X <- a[,colSums(a)>10000]
 s <- copy(agcd)[colnames(X),]
 if(length(unique(s$Sex))<2){
   return()
 } else {
 
   dge <- DGEList(counts=as.matrix(X),samples=s)
   dge <- normLibSizes(dge,method = "TMM")
   
   ## find the smaller group for this cluster
   smolgrp <- table(dge$samples$Sex)[which.min(table(dge$samples$Sex))]
   
   ## for the smallest group, determine low expression cutoff
   rowmn.m <- as.numeric(rowMeans(cpm(dge[,(dge$samples$Sex==names(smolgrp)[1])],log=TRUE)))
   
    ## get histogram distribution with 10 bins at the low end of the logcpm space, skipping to the second minimum in the distribution (to avoid including 0-count genes etc). this is tricky because with the small number of genes, the distribution is rather peaky. however, some spot-checking to tune this looks like using 8 bins for the lower norm-sized-logcpm range (second minima + 1.5 logcpm) retrieves an accurate last-minimum-before-the-normalish-part-of-the-norm-count-distribution.
   # additionally, most genes fall in a small normalized library size window (6-10ish) -- part of the 10x panel design process when providing seq data from a similar sample (as we did, providing visium), which they use to tweak the number of probes per gene for detection based on the tissue's expression of the targets.
   h <- hist(rowmn.m[which(rowmn.m>min(rowmn.m)&rowmn.m<(1.5+min(rowmn.m)))],breaks=8,plot=F)
   ## get the minimum in this low logcpm range and its corresponding bin (x axis value, i.e. log cpm). in case of repeat values, grab the first (lowest)
   minc <- h$breaks[which(h$counts==min(h$counts))][1]
   
   ## get the corresponding 33rd %ile-library count value for this minimum^
   smollib <- min(dge$samples[(dge$samples$Sex==names(smolgrp)[1]),"lib.size"])
   countthresh <- (minc^2)*(as.numeric(smollib)/10^6)
   
   ## and convert that smallest count to the logcpm value in said smallest library to use as a low-expression threshold. for conservativeness, round up, since we're dealing with relatively small numbers of counts here
   cpmthresh <- log2(ceiling(countthresh)/(as.numeric(smollib)*10^-6))
   
   ## filter for genes with at least thresh counts in at least (smaller sex group size) samples
   keep <- rowSums(cpm(dge,log=T)>cpmthresh) >= as.numeric(smolgrp)
   
   ## recalc lib factors after dropping low-expressed genes
   dge <- dge[keep,,keep.lib.sizes=FALSE]
   dge <- calcNormFactors(dge,method = "TMM")
   
   return(dge)
 }
})
  
clus <- makeCluster(10,type = "PSOCK")
fulldgelres <- clusterApply(x=dgels,cl = clus,fun=function(X){
 
 library(data.table)
 setDTthreads(1,restore_after_fork = FALSE)
 library(SpatialExperiment)
 library(limma)
 library(edgeR)
 
 # prevent resource competition, though we aren't explicitly using DT at all here:
 setDTthreads(1) 
 
 dge <- X
 des <- model.matrix(~0+Sex+batch, data = dge$samples)
 # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
 sexdif <- makeContrasts(SexMale-SexFemale,levels=des)
 
 # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
 v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$BrNum),adaptive.span = T,sample.weights = T)
 
 v.swt.fit <- contrasts.fit(v.swt,sexdif)
 
 v.swt.fit.e <- eBayes(v.swt.fit)
 
 v.swt.e.tt <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
 
 return(v.swt.e.tt)
})

stopCluster(clus)
names(fulldgelres) <- names(ag)

### concatenate results for all cell types
fulldgelres <- lapply(fulldgelres,as.data.table,keep.rownames=T)
fulldgelres.arc <- rbindlist(fulldgelres,idcol="cluster")
setnames(fulldgelres.arc,"rn","gene_name")

## append cell type annotations
fulldgelres.arc <- merge.data.table(fulldgelres.arc,bkanno,by.x="cluster",by.y="clus",all.x=T)
setnames(fulldgelres.arc,c("bjm_annot","cluster"),c("cluster_annotated","cluster_id"))
fulldgelres.arc <- fulldgelres.arc[,.(cluster_annotated,gene_name,logFC,adj.P.Val,t,B,P.Value,AveExpr,cluster_id)]

### clean up
rm(ag,agc,agcd,agsc,clus,dgels,fulldgelres,hyp.fineadom,hyp.adom.ag)
gc(full=T)
```

### repeat as above for VMH
```{r}
hyp.vdom.ag <- dreamlet::aggregateToPseudoBulk(hyp.finevdom,assay="counts",sample_id="sample_id",cluster_id = "banksy_clus",BPPARAM = MulticoreParam(8))

# extract colData to feed into DGEList as the "samples" data
agcd <- as.data.table(colData(hyp.vdom.ag),keep.rownames=T)
agcd[Date_Experiment_Began=="2/21/24",batch:="second"]
agcd[Date_Experiment_Began=="10/17/23",batch:="first"]
agcd <- as.data.frame(agcd)
rownames(agcd) <- agcd$rn
agcd$rn <- NULL

## only retain clusters for which at least 6 samples (size of the smaller sex group in total samps (females))) are available after pseudobulking+filtering through dreamlet and have at least 10,000 pseudobulked counts for the cluster of interest (or normalization and low-expression filtering gets to be impossible):
ag <- as.list(assays(hyp.vdom.ag))
agc <- lapply(ag,FUN=function(a){ncol(a[,colSums(a)>10000])})
# no samples pass for X37, which breaks this bit of code. substitute in 0 for that slot
agc["X37"] <- 0
ag <- ag[!(agc<6)]

## drop clusters where there are 6+ samples, but only 0 or 1 samples for one of the sexes:
agsc <- lapply(ag,FUN=function(a){
 cn <- colnames(a[,colSums(a)>10000])
 tmp <- as.data.table(agcd[cn,])
 if(nrow(tmp[,.N,by="Sex"][N<2])>0){
   return(0)
 } else {return(1)}
})

ag <- ag[agsc==1]

## make the DGELists per remaining cluster:
dgels <- lapply(ag,FUN=function(a){
   # filtering for 10,000 counts minimum as above
 X <- a[,colSums(a)>10000]
 s <- copy(agcd)[colnames(X),]
 if(length(unique(s$Sex))<2){
   return()
 } else {
 
   dge <- DGEList(counts=as.matrix(X),samples=s)
   dge <- normLibSizes(dge,method = "TMM")
   
   ## find the smaller group for this cluster
   smolgrp <- table(dge$samples$Sex)[which.min(table(dge$samples$Sex))]
   
   ## for the smallest group, determine low expression cutoff
   rowmn.m <- as.numeric(rowMeans(cpm(dge[,(dge$samples$Sex==names(smolgrp)[1])],log=TRUE)))
   
    ## get histogram distribution with 10 bins at the low end of the logcpm space, skipping to the second minimum in the distribution (to avoid including 0-count genes etc). this is tricky because with the small number of genes, the distribution is rather peaky. however, some spot-checking to tune this looks like using 8 bins for the lower norm-sized-logcpm range (second minima + 1.5 logcpm) retrieves an accurate last-minimum-before-the-normalish-part-of-the-norm-count-distribution.
   # additionally, most genes fall in a small normalized library size window (6-10ish) -- part of the 10x panel design process when providing seq data from a similar sample (as we did, providing visium), which they use to tweak the number of probes per gene for detection based on the tissue's expression of the targets.
   h <- hist(rowmn.m[which(rowmn.m>min(rowmn.m)&rowmn.m<(1.5+min(rowmn.m)))],breaks=8,plot=F)
   ## get the minimum in this low logcpm range and its corresponding bin (x axis value, i.e. log cpm). in case of repeat values, grab the first (lowest)
   minc <- h$breaks[which(h$counts==min(h$counts))][1]
   
   ## get the corresponding 33rd %ile-library count value for this minimum^
   smollib <- min(dge$samples[(dge$samples$Sex==names(smolgrp)[1]),"lib.size"])
   countthresh <- (minc^2)*(as.numeric(smollib)/10^6)
   
   ## and convert that smallest count to the logcpm value in said smallest library to use as a low-expression threshold. for conservativeness, round up, since we're dealing with relatively small numbers of counts here
   cpmthresh <- log2(ceiling(countthresh)/(as.numeric(smollib)*10^-6))
   
   ## filter for genes with at least thresh counts in at least (smaller sex group size) samples
   keep <- rowSums(cpm(dge,log=T)>cpmthresh) >= as.numeric(smolgrp)
   
   ## recalc lib factors after dropping low-expressed genes
   dge <- dge[keep,,keep.lib.sizes=FALSE]
   dge <- calcNormFactors(dge,method = "TMM")
   
   return(dge)
 }
})
  
clus <- makeCluster(10,type = "PSOCK")
fulldgelres.vmh <- clusterApply(x=dgels,cl = clus,fun=function(X){
 
 library(data.table)
 setDTthreads(1,restore_after_fork = FALSE)
 library(SpatialExperiment)
 library(limma)
 library(edgeR)
 
 # prevent resource competition, though we aren't explicitly using DT at all here:
 setDTthreads(1) 
 
 dge <- X
 des <- model.matrix(~0+Sex+batch, data = dge$samples)
 # INITIALIZE the voomwithqualityweights model, equivalent to pseudobulkDGE(method="voom",qualities=T)
 sexdif <- makeContrasts(SexMale-SexFemale,levels=des)
 
 # run voomLmFit for the pseudobulked data, referring donor to duplicateCorrelation; using an adaptive span (number of genes, based on the number of genes in the dge) for smoothing the mean-variance trend
 v.swt <- voomLmFit(dge,design = des,block = as.factor(dge$samples$BrNum),adaptive.span = T,sample.weights = T)
 
 v.swt.fit <- contrasts.fit(v.swt,sexdif)
 
 v.swt.fit.e <- eBayes(v.swt.fit)
 
 v.swt.e.tt <- topTable(v.swt.fit.e,coef = 1,number=Inf,adjust.method = "BH")
 
 return(v.swt.e.tt)
})

stopCluster(clus)
names(fulldgelres.vmh) <- names(ag)

### concatenate results for all cell types
fulldgelres.vmh <- lapply(fulldgelres.vmh,as.data.table,keep.rownames=T)
fulldgelres.vmh <- rbindlist(fulldgelres.vmh,idcol="cluster")
setnames(fulldgelres.vmh,"rn","gene_name")

## append cell type annotations
fulldgelres.vmh <- merge.data.table(fulldgelres.vmh,bkanno,by.x="cluster",by.y="clus",all.x=T)
setnames(fulldgelres.vmh,c("bjm_annot","cluster"),c("cluster_annotated","cluster_id"))
fulldgelres.vmh <- fulldgelres.vmh[,.(cluster_annotated,gene_name,logFC,adj.P.Val,t,B,P.Value,AveExpr,cluster_id)]
```

### concat and save
```{r}
setnames(fulldgelres.arc,"logFC","logFC_MvF")
setnames(fulldgelres.vmh,"logFC","logFC_MvF")

## save
fwrite(fulldgelres.arc,"processed-data/08_VMH-ARC cell type sex DE within domains/05a-4celltypeARCsmoothing_hiconf-interior96pctOfARC_celltypeSexDE.txt",sep='\t',quote=F)
fwrite(fulldgelres.vmh,"processed-data/08_VMH-ARC cell type sex DE within domains/05b-4celltypeARCsmoothing_hiconf-interior96pctOfVMH_celltypeSexDE.txt",sep='\t',quote=F)

```


## reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] parallel  stats4    stats     graphics  grDevices utils     datasets  methods  
[9] base     

other attached packages:
 [1] sf_1.0-16                      parallelly_1.38.0             
 [3] colorout_1.3-0.2               edgeR_4.2.0                   
 [5] dreamlet_1.2.1                 variancePartition_1.34.0      
 [7] BiocParallel_1.38.0            limma_3.60.3                  
 [9] ggplot2_3.5.1                  SpatialFeatureExperiment_1.6.1
[11] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[13] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[15] GenomicRanges_1.56.1           GenomeInfoDb_1.40.1           
[17] IRanges_2.38.1                 S4Vectors_0.42.1              
[19] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[21] matrixStats_1.3.0              data.table_1.15.4             
[23] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] splines_4.4.1             bitops_1.0-7              tibble_3.2.1             
  [4] R.oo_1.26.0               graph_1.82.0              XML_3.99-0.17            
  [7] lifecycle_1.0.4           Rdpack_2.6                mixsqp_0.3-54            
 [10] rprojroot_2.0.4           lattice_0.22-6            MASS_7.3-61              
 [13] backports_1.5.0           magrittr_2.0.3            metafor_4.6-0            
 [16] rmarkdown_2.27            yaml_2.3.9                sessioninfo_1.2.2        
 [19] sp_2.1-4                  DBI_1.2.3                 minqa_1.2.7              
 [22] abind_1.4-5               zlibbioc_1.50.0           EnvStats_2.8.1           
 [25] purrr_1.0.2               R.utils_2.12.3            msigdbr_7.5.1            
 [28] rmeta_3.0                 RCurl_1.98-1.14           GenomeInfoDbData_1.2.12  
 [31] ggrepel_0.9.5             pbkrtest_0.5.3            irlba_2.3.5.1            
 [34] terra_1.7-78              units_0.8-5               annotate_1.82.0          
 [37] dqrng_0.4.1               DelayedMatrixStats_1.26.0 codetools_0.2-20         
 [40] DropletUtils_1.24.0       DelayedArray_0.30.1       scuttle_1.14.0           
 [43] tidyselect_1.2.1          UCSC.utils_1.0.0          lme4_1.1-35.5            
 [46] mathjaxr_1.6-0            jsonlite_1.8.8            BiocNeighbors_1.22.0     
 [49] e1071_1.7-14              iterators_1.0.14          progress_1.2.3           
 [52] tools_4.4.1               zenith_1.6.0              Rcpp_1.0.13              
 [55] glue_1.7.0                SparseArray_1.4.8         xfun_0.45                
 [58] here_1.0.1                EBImage_4.46.0            dplyr_1.1.4              
 [61] HDF5Array_1.32.0          withr_3.0.0               numDeriv_2016.8-1.1      
 [64] fastmap_1.2.0             boot_1.3-30               rhdf5filters_1.16.0      
 [67] fansi_1.0.6               spData_2.3.1              caTools_1.18.2           
 [70] digest_0.6.36             truncnorm_1.0-9           R6_2.5.1                 
 [73] colorspace_2.1-0          wk_0.9.2                  scattermore_1.2          
 [76] gtools_3.9.5              jpeg_0.1-10               RSQLite_2.3.7            
 [79] R.methodsS3_1.8.2         RhpcBLASctl_0.23-42       utf8_1.2.4               
 [82] tidyr_1.3.1               generics_0.1.3            corpcor_1.6.10           
 [85] class_7.3-22              prettyunits_1.2.0         httr_1.4.7               
 [88] htmlwidgets_1.6.4         S4Arrays_1.4.1            spdep_1.3-5              
 [91] pkgconfig_2.0.3           gtable_0.3.5              blob_1.2.4               
 [94] XVector_0.44.0            remaCor_0.0.18            htmltools_0.5.8.1        
 [97] fftwtools_0.9-11          GSEABase_1.66.0           scales_1.3.0             
[100] png_0.1-8                 fANCOVA_0.6-1             ashr_2.2-63              
[103] knitr_1.48                rstudioapi_0.16.0         reshape2_1.4.4           
[106] rjson_0.2.21              nlme_3.1-165              nloptr_2.1.1             
[109] proxy_0.4-27              cachem_1.1.0              rhdf5_2.48.0             
[112] stringr_1.5.1             KernSmooth_2.23-24        metadat_1.2-0            
[115] vipor_0.4.7               RcppZiggurat_0.1.6        AnnotationDbi_1.66.0     
[118] s2_1.1.6                  pillar_1.9.0              grid_4.4.1               
[121] vctrs_0.6.5               gplots_3.1.3.1            mashr_0.2.79             
[124] beachmat_2.20.0           xtable_1.8-4              sfheaders_0.4.4          
[127] beeswarm_0.4.0            Rgraphviz_2.48.0          KEGGgraph_1.64.0         
[130] evaluate_0.24.0           zeallot_0.1.0             invgamma_1.1             
[133] magick_2.8.3              mvtnorm_1.2-5             cli_3.6.3                
[136] locfit_1.5-9.10           compiler_4.4.1            crayon_1.5.3             
[139] SQUAREM_2021.1            classInt_0.4-10           ggbeeswarm_0.7.2         
[142] plyr_1.8.9                stringi_1.8.4             deldir_2.0-4             
[145] babelgene_22.9            assertthat_0.2.1          lmerTest_3.1-3           
[148] munsell_0.5.1             Biostrings_2.72.1         tiff_0.1-12              
[151] aod_1.3.3                 Matrix_1.7-0              hms_1.1.3                
[154] sparseMatrixStats_1.16.0  bit64_4.0.5               Rhdf5lib_1.26.0          
[157] KEGGREST_1.44.1           statmod_1.5.0             rbibutils_2.2.16         
[160] Rfast_2.1.0               broom_1.0.6               memoise_2.0.1            
[163] RcppParallel_5.1.8        bit_4.0.5                 EnrichmentBrowser_2.34.1 
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-01
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package                  * version    date (UTC) lib source
   abind                      1.4-5      2016-07-21 [1] CRAN (R 4.4.0)
   annotate                   1.82.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   AnnotationDbi              1.66.0     2024-05-01 [1] Bioconductor 3.19 (R 4.4.0)
   aod                        1.3.3      2023-12-13 [1] CRAN (R 4.4.0)
   ashr                       2.2-63     2023-08-21 [1] CRAN (R 4.4.0)
   assertthat                 0.2.1      2019-03-21 [1] CRAN (R 4.4.0)
   babelgene                  22.9       2022-09-29 [1] CRAN (R 4.4.0)
   backports                  1.5.0      2024-05-23 [1] CRAN (R 4.4.0)
   beachmat                   2.20.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   beeswarm                   0.4.0      2021-06-01 [1] CRAN (R 4.4.0)
   Biobase                  * 2.64.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   BiocGenerics             * 0.50.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Biostrings                 2.72.1     2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bit                        4.0.5      2022-11-15 [1] CRAN (R 4.4.0)
   bit64                      4.0.5      2020-08-30 [1] CRAN (R 4.4.0)
   bitops                     1.0-7      2021-04-24 [1] CRAN (R 4.4.0)
   blob                       1.2.4      2023-03-17 [1] CRAN (R 4.4.0)
   boot                       1.3-30     2024-02-26 [1] CRAN (R 4.4.1)
   broom                      1.0.6      2024-05-17 [1] CRAN (R 4.4.0)
   cachem                     1.1.0      2024-05-16 [1] CRAN (R 4.4.0)
   caTools                    1.18.2     2021-03-28 [1] CRAN (R 4.4.0)
   class                      7.3-22     2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10     2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.3      2024-06-21 [2] CRAN (R 4.4.0)
   codetools                  0.2-20     2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2    2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0      2023-01-23 [1] CRAN (R 4.4.0)
   corpcor                    1.6.10     2021-09-16 [1] CRAN (R 4.4.0)
   crayon                     1.5.3      2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4     2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3      2024-06-02 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1     2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4      2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.36     2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4      2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1      2024-05-28 [1] CRAN (R 4.4.0)
   dreamlet                 * 1.2.1      2024-06-09 [1] Bioconductor 3.19 (R 4.4.0)
   DropletUtils               1.24.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14     2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                    * 4.2.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   EnrichmentBrowser          2.34.1     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   EnvStats                   2.8.1      2023-08-22 [1] CRAN (R 4.4.0)
   evaluate                   0.24.0     2024-06-10 [1] CRAN (R 4.4.0)
   fANCOVA                    0.6-1      2020-11-13 [1] CRAN (R 4.4.0)
   fansi                      1.0.6      2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0      2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11     2021-03-01 [1] CRAN (R 4.4.0)
   generics                   0.1.3      2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1     2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12     2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1     2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggbeeswarm                 0.7.2      2023-04-29 [1] CRAN (R 4.4.0)
   ggplot2                  * 3.5.1      2024-04-23 [1] CRAN (R 4.4.0)
   ggrepel                    0.9.5      2024-01-10 [1] CRAN (R 4.4.0)
   glue                       1.7.0      2024-01-09 [1] CRAN (R 4.4.0)
   gplots                     3.1.3.1    2024-02-02 [1] CRAN (R 4.4.0)
   graph                      1.82.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   GSEABase                   1.66.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   gtable                     0.3.5      2024-04-22 [1] CRAN (R 4.4.0)
   gtools                     3.9.5      2023-11-20 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1      2020-12-13 [1] CRAN (R 4.4.0)
   hms                        1.1.3      2023-03-21 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1    2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4      2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7      2023-08-15 [1] CRAN (R 4.4.0)
   invgamma                   1.1        2017-05-07 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.1     2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   irlba                      2.3.5.1    2022-10-03 [1] CRAN (R 4.4.0)
   iterators                  1.0.14     2022-02-05 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10     2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8      2023-12-04 [1] CRAN (R 4.4.0)
   KEGGgraph                  1.64.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   KEGGREST                   1.44.1     2024-06-19 [1] Bioconductor 3.19 (R 4.4.0)
   KernSmooth                 2.23-24    2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.48       2024-07-07 [1] CRAN (R 4.4.1)
   lattice                    0.22-6     2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle                  1.0.4      2023-11-07 [1] CRAN (R 4.4.0)
   limma                    * 3.60.3     2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   lme4                       1.1-35.5   2024-07-03 [1] CRAN (R 4.4.1)
   lmerTest                   3.1-3      2020-10-23 [1] CRAN (R 4.4.0)
   locfit                     1.5-9.10   2024-06-24 [1] CRAN (R 4.4.0)
   magick                     2.8.3      2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                   2.0.3      2022-03-30 [1] CRAN (R 4.4.0)
   mashr                      0.2.79     2023-10-18 [1] CRAN (R 4.4.0)
   MASS                       7.3-61     2024-06-13 [1] CRAN (R 4.4.0)
   mathjaxr                   1.6-0      2022-02-28 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0      2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0      2024-04-11 [1] CRAN (R 4.4.0)
   memoise                    2.0.1      2021-11-26 [1] CRAN (R 4.4.0)
   metadat                    1.2-0      2022-04-06 [1] CRAN (R 4.4.0)
   metafor                    4.6-0      2024-03-28 [1] CRAN (R 4.4.0)
   minqa                      1.2.7      2024-05-20 [1] CRAN (R 4.4.0)
   mixsqp                     0.3-54     2023-12-20 [1] CRAN (R 4.4.0)
   msigdbr                    7.5.1      2022-03-30 [1] CRAN (R 4.4.0)
   munsell                    0.5.1      2024-04-01 [1] CRAN (R 4.4.0)
   mvtnorm                    1.2-5      2024-05-21 [1] CRAN (R 4.4.0)
   nlme                       3.1-165    2024-06-06 [1] CRAN (R 4.4.0)
   nloptr                     2.1.1      2024-06-25 [1] CRAN (R 4.4.0)
   numDeriv                   2016.8-1.1 2019-06-06 [1] CRAN (R 4.4.0)
   parallelly               * 1.38.0     2024-07-27 [1] CRAN (R 4.4.0)
   pbkrtest                   0.5.3      2024-06-26 [1] CRAN (R 4.4.0)
   pillar                     1.9.0      2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3      2019-09-22 [1] CRAN (R 4.4.0)
   plyr                       1.8.9      2023-10-02 [1] CRAN (R 4.4.0)
   png                        0.1-8      2022-11-29 [1] CRAN (R 4.4.0)
   prettyunits                1.2.0      2023-09-24 [1] CRAN (R 4.4.0)
   progress                   1.2.3      2023-12-06 [1] CRAN (R 4.4.0)
   proxy                      0.4-27     2022-06-09 [1] CRAN (R 4.4.0)
   purrr                      1.0.2      2023-08-10 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2      2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0     2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3     2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1      2021-08-19 [1] CRAN (R 4.4.0)
   rbibutils                  2.2.16     2023-10-25 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.13     2024-07-17 [1] CRAN (R 4.4.0)
   RcppParallel               5.1.8      2024-07-06 [1] CRAN (R 4.4.0)
   RcppZiggurat               0.1.6      2020-10-20 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14  2024-01-09 [1] CRAN (R 4.4.0)
   Rdpack                     2.6        2023-11-08 [1] CRAN (R 4.4.0)
   remaCor                    0.0.18     2024-02-08 [1] CRAN (R 4.4.0)
   reshape2                   1.4.4      2020-04-09 [1] CRAN (R 4.4.0)
   Rfast                      2.1.0      2023-11-09 [1] CRAN (R 4.4.0)
   Rgraphviz                  2.48.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5                      2.48.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   RhpcBLASctl                0.23-42    2023-02-11 [1] CRAN (R 4.4.0)
   rjson                      0.2.21     2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4      2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27       2024-05-17 [1] CRAN (R 4.4.0)
   rmeta                      3.0        2018-03-20 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4      2023-11-05 [1] CRAN (R 4.4.0)
   RSQLite                    2.3.7      2024-05-27 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0     2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6      2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1      2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.1     2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                     1.3.0      2023-11-28 [1] CRAN (R 4.4.0)
   scattermore                1.2        2023-06-12 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2      2021-12-06 [1] CRAN (R 4.4.0)
   sf                       * 1.0-16     2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4      2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4      2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8      2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1      2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1      2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5      2024-06-10 [1] CRAN (R 4.4.0)
   SQUAREM                    2021.1     2021-01-13 [1] CRAN (R 4.4.0)
   statmod                    1.5.0      2023-01-06 [1] CRAN (R 4.4.0)
   stringi                    1.8.4      2024-05-06 [1] CRAN (R 4.4.0)
   stringr                    1.5.1      2023-11-14 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78     2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1      2023-03-20 [1] CRAN (R 4.4.0)
   tidyr                      1.3.1      2024-01-24 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1      2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12     2023-11-28 [1] CRAN (R 4.4.0)
   truncnorm                  1.0-9      2023-03-20 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0      2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5      2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4      2023-10-22 [1] CRAN (R 4.4.0)
   variancePartition        * 1.34.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   vctrs                      0.6.5      2023-12-01 [1] CRAN (R 4.4.0)
   vipor                      0.4.7      2023-12-18 [1] CRAN (R 4.4.0)
   withr                      3.0.0      2024-01-16 [1] CRAN (R 4.4.0)
   wk                         0.9.2      2024-07-09 [1] CRAN (R 4.4.0)
   xfun                       0.45       2024-06-16 [1] CRAN (R 4.4.0)
   XML                        3.99-0.17  2024-06-25 [1] CRAN (R 4.3.3)
   xtable                     1.8-4      2019-04-21 [1] CRAN (R 4.4.0)
   XVector                    0.44.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.9      2024-07-05 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0      2018-01-28 [1] CRAN (R 4.4.0)
   zenith                     1.6.0      2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   zlibbioc                   1.50.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
