---
title: "07-Sex DE (02, 04-06) Xenium-Visium comparison"
output: html_document
date: "2024-08-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

```

### read in xenium cell type DE results, domain DE results, cell type DE in high-conf domain results, nuclear counts DE results; Visium k=15 bayesspace sex DE results with VMH and ARC domains individually and with them collapsed

```{r}
xencellarc <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/02a-4ARCtypesmoothed_withinARC_celltypeSexDE.txt")
xencellvmh <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/02b-4ARCtypesmoothed_withinVMH_celltypeSexDE.txt")

# domain-wide sex DE
xendomain <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/04-4typeARC-and-dualassignVMH_domainwise_sexDE.txt")

# cell type DE in high conf domains
xenhiconfarc <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/05a-4celltypeARCsmoothing_hiconf-interior96pctOfARC_celltypeSexDE.txt")
xenhiconfvmh <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/05b-4celltypeARCsmoothing_hiconf-interior96pctOfVMH_celltypeSexDE.txt")

# nuclear counts-based cell type DE
xennucarc <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/06a-4celltypeARCsmoothing_celltypeSexDE_nuclearcounts_inARC.txt")
xennucvmh <- fread("processed-data/08_VMH-ARC cell type sex DE within domains/06b-4celltypeARCsmoothing_celltypeSexDE_nuclearcounts_inVMH.txt")

## read in visium and reorganinze to match xenium de result table contents
## subset to VMH and ARC domains in these two
visdom <- fread("../spatial_HYP/processed-data/09-Sex DE/01b-voomLmFit_nnsvg10-HmnylmbNA-BS-15-singleClusts.txt")
visdom <- visdom[assay%in% c("X4","X6","X7","X12")]
visdom[assay=="X4",assay:="ARC.1"]
visdom[assay=="X6",assay:="ARC.2"]
visdom[assay=="X7",assay:="VMH.1"]
visdom[assay=="X12",assay:="VMH.2"]

visclps <- fread("../spatial_HYP/processed-data/09-Sex DE/01c-voomLmFit_nnsvg10-HmnylmbNA-BS-15-VMHARCclpsd.txt")
visclps <- visclps[assay %in% c("VMH","ARC")]

## get a list of these together for streamlined operations
des <- list(xencellarc,xencellvmh,xenhiconfarc,xenhiconfvmh,xennucarc,xennucvmh,xendomain,visdom,visclps)
names(des) <- c("xencellarc","xencellvmh","xenhiconfarc","xenhiconfvmh","xennucarc","xennucvmh","xendomain","visdom","visclps")
```

### this is clunky, but it'll work
```{r}
## subset to VMH/ARC clusters for xenium
des[1:6] <- lapply(des[1:6],FUN=function(x){
    x <- x[cluster_annotated %in% c(paste0("VMH_",c(1:3)),paste0("ARC_",c(1:5)))]
})

## subset visiums to single/joint arc/vmh domains
des[8:9] <- lapply(des[8:9],FUN=function(x){
    x <- x[assay %in% c("VMH","ARC","VMH.1","VMH.2","ARC.1","ARC.2")]
})
## make the cluster column uniformly "assay" across datasets
des[1:6] <- lapply(des[1:6],FUN=function(x){setnames(x,"cluster_annotated","assay")})
des[[7]] <- setnames(des[[7]],"domain","assay")

## get the unique combinations of these analyses
combs <- expand.grid(names(des),names(des))
combs <- as.data.table(combs)
setnames(combs,c("de1","de2"))
combs[,de1:=as.character(de1)]
combs[,de2:=as.character(de2)]

corstab <- as.data.frame(matrix(nrow=1,ncol=3))

i<-1
for (i in c(1:nrow(combs))){
    # get the unique "assay" values across these two analyses
    assays <- unique(c(des[[combs[i,de1]]]$assay,des[[combs[i,de2]]]$assay))
    assays <- expand.grid(assays,assays)
    assays <- as.data.table(assays)
    setnames(assays,c("a1","a2"))
    assays[,a1:=as.character(a1)]
    assays[,a2:=as.character(a2)]
    
    ## iterate through those combinations for the i'th pair of analyses
    j<-1
    for (j in c(1:nrow(assays))){
        assay1 <- des[[combs[i,de1]]][assay==assays[j,a1]]
        assay2 <- des[[combs[i,de2]]][assay==assays[j,a2]]
        
        t1 <- grep(names(assay1),pattern="^t.*$",value=T)
        setnames(assay1,t1,"t1")
        
        t2 <- grep(names(assay2),pattern="^t.*$",value=T)
        setnames(assay2,t2,"t2")
        
        ## line up the values for correlation
        curtest <- merge.data.table(assay1[,.(gene_name,t1)],assay2[,.(gene_name,t2)],by="gene_name")
        ## correlate
        if(nrow(curtest)==0){next(j)} else{
            spear <- cor(curtest$t1,curtest$t2,method="spearman",use="pairwise.complete.obs")
            corstab <- rbind(corstab, c(
                paste0(assays[j,a1],"_",combs[i,de1]),
                paste0(assays[j,a2],"_",combs[i,de2]),
                spear
            ))
            rm(spear)
        }
        rm(curtest,t2,t1,assay1,assay2)
    }
    rm(assays,j)
}

## remove self-correlations and NAs
corstab <- as.data.table(corstab)
setnames(corstab,c("DEcluster_domain1","DEcluster_domain2","spearman"))
corstab <- corstab[DEcluster_domain1!=DEcluster_domain2 & !is.na(spearman)]

```

### save
```{r}
fwrite(corstab,"processed-data/08_VMH-ARC cell type sex DE within domains/07-Sex DE correlations between pairs of cell type or domain wise analyses of Xenium, Visium.txt")
```

reprod info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
[1] colorout_1.3-0.2  data.table_1.15.4 rlang_1.1.4      

loaded via a namespace (and not attached):
 [1] compiler_4.4.1    here_1.0.1        fastmap_1.2.0     rprojroot_2.0.4  
 [5] cli_3.6.3         htmltools_0.5.8.1 tools_4.4.1       rstudioapi_0.16.0
 [9] yaml_2.3.9        rmarkdown_2.27    knitr_1.48        digest_0.6.36    
[13] xfun_0.45         evaluate_0.24.0  
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-08-23
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────────
 ! package     * version date (UTC) lib source
 P cli           3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   colorout    * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   data.table  * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   digest        0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   evaluate      0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fastmap       1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   here          1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools     0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   knitr         1.48    2024-07-07 [1] CRAN (R 4.4.1)
 P rlang       * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown     2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot     2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi    0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   sessioninfo   1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   xfun          0.45    2024-06-16 [1] CRAN (R 4.4.0)
   yaml          2.3.9   2024-07-05 [1] CRAN (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────────
