---
title: "01_1-QC Filters"
author: "Bernie Mulvey"
date: "2023-05-22"
output: html_document
---

```{r setup, include=FALSE}

#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(SpatialExperiment)
library(ggspavis)
library(scater) # addPerCellQC
library(nnSVG)
library(BiocParallel)
library(scran)
```

## load assembled, unfiltered data
```{r}
# from dcs04; in gitignore; new version from Ryan M 053123
load("data/spe_raw_053123.Rdata")
hyp <- spe
rm(spe)
```

```{r}
dim(hyp)
head(colData(hyp))
tail(colData(hyp))
```

## oo % mitochondrial already added in! Thanks Ryan!

### Ch9: QC / filtering
1. Is the tissue oriented correctly? -- REVISIT (not all anatomy/ish/slide images cross-mapped yet) (and it's impossible to tell just plotting spots since these are almost all perfect squares)
```{r}

# get rid of non-tissue spots
hyp2 <- hyp[,colData(hyp)$in_tissue]
ggspavis::plotSpots(hyp2)
sum(colData(hyp2)$in_tissue != 1, na.rm = T)
# qed
```

2. % mitochondrial reads--already tabulated it looks like, but since it says "expression" let's double check.
```{r}

## either of these approaches WILL work:
# 1. full-length vector of T/F for matching MT-.*
mitogenes <- rowData(hyp2)$gene_name %in% grep(rowData(hyp2)$gene_name, pattern = "MT-|mt-", value = T)
hyp2 <- addPerCellQC(hyp2, subsets = list(Mito = mitogenes))

# # 2. as coded in the OSTA book
# is_mito <- grepl("(^MT-)|(^mt-)", rowData(hyp2)$gene_name)
# rowData(hyp2)$gene_name[is_mito]
# hyp2 <- addPerCellQC(hyp2,subsets=list(Mito=is_mito))

## check equivalency to "expression ratio" column
tmp <- as.data.table(colData(hyp2))
tmp[,expr_chrM_ratio:=expr_chrM_ratio*100]

nrow(tmp[expr_chrM_ratio>=subsets_Mito_percent-0.1&expr_chrM_ratio<=subsets_Mito_percent+0.1])

nrow(tmp)
# ^ effectively, yes. but weird that there are 8 rows that are off. whys that?

tmp[!(expr_chrM_ratio>=subsets_Mito_percent-0.1)&!(expr_chrM_ratio<=subsets_Mito_percent+0.1)]

# k, not non-equivalency per se -- these are spots with zero genes, which will drop later
tmp[is.na(expr_chrM_ratio)]

# just that zero MT counts --> NA for the auto % calculation. so we're okay as long as we use the already-calculated column.
rm(tmp,mitogenes)
# 

# revert to spe without mito count columns added for sake of tidiness
hyp2 <- hyp[,colData(hyp)$in_tissue]

```

plot some qc stats
### mitochondrial read %age
```{r}
hist(colData(hyp2)$expr_chrM_ratio)
# nothing over ~50%, so that's good.
plotSpots(hyp2,annotate = "expr_chrM_ratio",palette = palette(value = "ggplot2"))
# and this is gradated over each slice rather than like distinct areas, so we're prob ok.
sum(colData(hyp2)$expr_chrM_ratio > 0.2,na.rm=T)
# 33 % of spots
sum(colData(hyp2)$expr_chrM_ratio > 0.3,na.rm=T)
# 9% of spots
sum(colData(hyp2)$expr_chrM_ratio > 0.35,na.rm=T)
# 5% of spots--lets start here so we know what we might be throwing away if raising later
```

### UMIs per spot
### standard threshold is 600 UMIs
```{r}
ggplot(as.data.frame(colData(hyp2)), aes(x = sum_umi)) +
    geom_histogram(binwidth = 100)+
  geom_vline(xintercept = 600,col="red")
dev.off()

sum(colData(hyp2)$sum_umi<600,na.rm=T)
# well 600 will do nicely! 4% of spots.
```

### also look at sum UMIs PER # cells per spot, if available (REVISIT-not available 052323). using ggspavis::plotQC. set threshold_y to the value we might plan to use--this function will draw on a horiz line to show what would be getting clipped off
# in the meantime, just use the cutoff of 600.
```{r}
# plotQC(hyp2,type = "scatter",metric_x = "cell_count",metric_y ="sum_umi",threshold_y = 600)
# splc@colData$cell_count <- sample(x=c(1:7),size=nrow(splc@colData),replace = T)
# hist(colData(splc)$cell_count,breaks = 20)
```

### genes per spot (sum_gene)
```{r}
ggplot(as.data.table(colData(hyp2)),aes(x=sum_gene)) +
    geom_histogram(binwidth = 100)+
    geom_vline(xintercept = 450,col="red")
dev.off()

### hm..there's this little jump in the ~1k-1.1k genes range. but otherwise ~450 genes looks like it cuts the very tail (and we certainly could have < 600 genes with 600 umis). so let's start with 450 genes minimum detected.
```

# we would look at # unique genes per spot as a function of cells per spot like this
```{r}
# plotQC(hyp2,type = "scatter",metric_x = "cell_count",metric_y ="sum",threshold_y = 600)
dev.off()
# as before, if we wanted to flag spots with more than 10 cells for exclusion (e.g., if that seems to be where the number of genes per spot dives off), then we add a new column to coldata as T/F:
```

### and eventually, let's visually check that spots with higher cell content are not all in one area of their tissue sections (which would indicate something biology-y happening)
```{r}
# plotSpots(hyp2,annotate = "cell_count")
# dev.off()
```

### double check UMI and gene count cutoffs against spots for spatial patterns
```{r}
plotSpots(hyp2,annotate = "sum_umi",palette = palette("ggplot2"))
### umis are definitely spatially differential to distinct areas (and not in a soft gradient at all like mitochondrial %age)
plotSpots(hyp2,annotate="sum_gene",palette=palette("ggplot2"))
dev.off()
### and as a rough estimator of gene diversity per RNA density andor seq depth:
colData(hyp2)$g_per_u <- colData(hyp2)$sum_gene/colData(hyp2)$sum_umi
plotSpots(hyp2,annotate="g_per_u",palette=palette("ggplot2"))
#nothing too wild going on. most spots pick up a new gene for every 1-2 additional UMIs. a few are very low xscriptional complexity, e.g. lower RH of 075-D1.

tmpdat <- as.data.table(colData(hyp2))
### now that we have cell auto-counts, i bet this g per u correlates decently to number of cells... 
ggplot(colData(hyp2),aes(x=g_per_u,y=Nmask_dark_blue))+geom_point()
## eh..correlates mediocrely

dev.off()
```

### apply mitochondria, UMI, unique gene cutoffs per above (35%, over 600, over 450)
```{r}
ncol(hyp2[,colData(hyp2)$sum_umi>600&colData(hyp2)$sum_gene>450&colData(hyp2)$expr_chrM_ratio<0.35])/ncol(hyp2)
# 91.7% spots retained for now!
hyp2 <- hyp2[,colData(hyp2)$sum_umi>600&colData(hyp2)$sum_gene>450&colData(hyp2)$expr_chrM_ratio<0.35]
```

### apply library-size normalized, LOG COUNT normalization AFTER filtering the data.
```{r}
library(scran)
hyp2 <- computeLibraryFactors(hyp2)
# peek
summary(sizeFactors(hyp2))
hist(sizeFactors(hyp2),breaks=100)
dev.off()
# append log counts using scater (don't use the SpatialExpt:: function here, that's for RETRIEVING an assay by the name of "logcounts" from spe objs)
hyp2 <- scater::logNormCounts(hyp2)
stopifnot(dim(counts(hyp2))==dim(logcounts(hyp2)))
# saveRDS(hyp2,"data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")

rm(hyp)
gc(full=T)
```
