---
title: "01-PCA, UMAP, Harmony Plots by Cohort Variables"
author: "Bernard Mulvey"
date: "2024-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(ggplot2)
library(gridExtra)
library(SpatialExperiment)
require(colorout)
library(BiocParallel)

ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(gtools)

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

```{r}
# load SPE--already contains the additional demographic/technical data at this point (age, sex, BMI, PMI, best RIN from cortex..), with samples rotated
hyp2 <- readRDS("processed-data/03-QC_filters/hypN10_umi210_gene126_chrm50_spotsweeped_lognorm_rotsNmirrors_071924.RDS")

# we only need the colData with spot identifiers here so
hypcd <- colData(hyp2)
rm(hyp2)
gc(full=T)
# load dimensionality reductions
dimredlist <- readRDS("processed-data/04-feature_selection/02-hypfiltered_4featureset-pca-umap-harmonydefault-harmonylambdanull-mnn30_reducedDim_slot.RDS")


# look thru  variables we want to plot
colnames(hypcd)

# for dimensionality reduction at the spot level, we want to include brnum, agedeath, race, sex, PMI, best_rin_pfc, bmi, sample_id, sum_umi, sum_gene, expr_chrM, slide, and to be sure nothing too strange is happening, array_row and array_col
labvars <- c("brnum","AgeDeath","Race","Sex","PrimaryDx","PMI","best_rin_pfc","bmi","sample_id","sum_umi","sum_gene","expr_chrM","slide","array_row","array_col")
```

```{r}
bpmapply(X=dimredlist,Y=names(dimredlist),SIMPLIFY=FALSE,BPPARAM=SnowParam(workers = 8,type="SOCK"),BPOPTIONS=bpoptions(exports=c("labvars","hypcd"),packages = c("ggplot2","SpatialExperiment","gridExtra","data.table")),FUN=function(X,Y){
    # dimred combinations to plot #
    if(length(grep(Y,pattern="UMAP",value=T))>0){
        zq <- rbind(c(1,2))
    }
    else{
        zq <- rbind(c(1,2),c(1,3),c(2,3))
    }
    
    plts <- list()
    
    v <- 1
    for (v in c(1:length(labvars))){
        m<-1
        for (m in c(1:nrow(zq))){
            z <- zq[m,1]
            q <- zq[m,2]
            pltdat <- cbind(X[,c(z,q)],hypcd[,labvars[v]])
            
            pltdat <- as.data.table(pltdat)
            setnames(pltdat,c(paste0(Y,paste0("_Dim",z)),paste0(Y,paste0("_Dim",q)),labvars[v]))
            xvar <- names(pltdat)[1]
            yvar <- names(pltdat)[2]
            
            # set dimensions to numeric, plotting variable to factor or numeric
            set(pltdat,j=xvar,value=as.numeric(pltdat[,get(xvar)]))
            set(pltdat,j=yvar,value=as.numeric(pltdat[,get(yvar)]))
            if(labvars[v] %in% c("AgeDeath","PMI","best_rin_pfc","bmi","sum_umi","sum_gene","expr_chrM","array_row","array_col")){
                set(pltdat,j=labvars[v],value=as.numeric(pltdat[,get(labvars[v])]))
            }
            else{
                set(pltdat,j=labvars[v],value=as.factor(pltdat[,get(labvars[v])]))
            }
     
            plts[[length(plts)+1]] <- ggplot(pltdat,aes(x=.data[[xvar]],y=.data[[yvar]],col=.data[[labvars[v]]]))+
                geom_point()
                #scale_color_discrete(DiscretePalette_scCustomize(num_colors = 2 + length(unique(colData(hyp2.tmp)$label)), palette = "varibow")[3:(length(unique(colData(hyp2.tmp)$label)) + 2)])
        }
        
    }
    
    ## determine appropriate subfolder for pllots -- github copilot came out swinging and filled most of this in near-perfectly!
    if (length(grep(Y,pattern="PCA",value=T))>0){subdir<-"02-PCA/"}
    else if (length(grep(Y,pattern="HARMONYdflt",value=T))>0){subdir<-"02-Harmonydefault/"}
    else if (length(grep(Y,pattern="HARMONYlmbna",value=T))>0){subdir<-"02-Harmonylambdanull/"}
    else if(length(grep(Y,pattern="mnn30",value=T))>0){subdir<-"02-mnn30/"}
    else{subdir<-"02-UMAP/"}
    
    ## render
    png(paste0("plots/04-Feature Selection, PCA, Harmony/",subdir,Y,"_by_cohortvariables.png"),width=48,height=40,units="in",res=300)
    do.call("grid.arrange",plts)
    dev.off()
    
    rm(plts,pltdat,xvar,yvar)
})

```

```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] parallelly_1.37.1           default_1.0.0              
 [3] gtools_3.9.5                BiocParallel_1.38.0        
 [5] colorout_1.3-0.2            SpatialExperiment_1.14.0   
 [7] SingleCellExperiment_1.26.0 SummarizedExperiment_1.34.0
 [9] Biobase_2.64.0              GenomicRanges_1.56.1       
[11] MatrixGenerics_1.16.0       matrixStats_1.3.0          
[13] gridExtra_2.3               ggplot2_3.5.1              
[15] Biostrings_2.72.1           GenomeInfoDb_1.40.1        
[17] XVector_0.44.0              IRanges_2.38.1             
[19] S4Vectors_0.42.1            BiocGenerics_0.50.0        
[21] data.table_1.15.4           rlang_1.1.4                

loaded via a namespace (and not attached):
 [1] gtable_0.3.5            rjson_0.2.21            xfun_0.45              
 [4] lattice_0.22-6          vctrs_0.6.5             tools_4.4.1            
 [7] generics_0.1.3          parallel_4.4.1          tibble_3.2.1           
[10] fansi_1.0.6             pkgconfig_2.0.3         Matrix_1.7-0           
[13] lifecycle_1.0.4         GenomeInfoDbData_1.2.12 compiler_4.4.1         
[16] munsell_0.5.1           codetools_0.2-20        htmltools_0.5.8.1      
[19] yaml_2.3.9              pillar_1.9.0            crayon_1.5.3           
[22] DelayedArray_0.30.1     magick_2.8.3            abind_1.4-5            
[25] tidyselect_1.2.1        digest_0.6.36           dplyr_1.1.4            
[28] rprojroot_2.0.4         fastmap_1.2.0           grid_4.4.1             
[31] here_1.0.1              colorspace_2.1-0        cli_3.6.3              
[34] SparseArray_1.4.8       magrittr_2.0.3          S4Arrays_1.4.1         
[37] utf8_1.2.4              withr_3.0.0             scales_1.3.0           
[40] UCSC.utils_1.0.0        rmarkdown_2.27          httr_1.4.7             
[43] evaluate_0.24.0         knitr_1.48              Rcpp_1.0.13            
[46] glue_1.7.0              rstudioapi_0.16.0       jsonlite_1.8.8         
[49] R6_2.5.1                zlibbioc_1.50.0        
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-07-20
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────
 ! package              * version date (UTC) lib source
   abind                  1.4-5   2016-07-21 [1] CRAN (R 4.4.0)
   Biobase              * 2.64.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics         * 0.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel         * 1.38.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings           * 2.72.1  2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
 P cli                    3.6.3   2024-06-21 [2] CRAN (R 4.4.0)
   codetools              0.2-20  2024-03-31 [1] CRAN (R 4.4.1)
   colorout             * 1.3-0.2 2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace             2.1-0   2023-01-23 [1] CRAN (R 4.4.0)
   crayon                 1.5.3   2024-06-20 [1] CRAN (R 4.4.0)
   data.table           * 1.15.4  2024-03-30 [2] CRAN (R 4.4.0)
   default              * 1.0.0   2017-08-07 [1] CRAN (R 4.4.0)
   DelayedArray           0.30.1  2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   digest                 0.6.36  2024-06-23 [1] CRAN (R 4.4.0)
   dplyr                  1.1.4   2023-11-17 [1] CRAN (R 4.4.0)
   evaluate               0.24.0  2024-06-10 [1] CRAN (R 4.4.0)
   fansi                  1.0.6   2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                1.2.0   2024-05-15 [1] CRAN (R 4.4.0)
   generics               0.1.3   2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb         * 1.40.1  2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData       1.2.12  2024-05-01 [1] Bioconductor
   GenomicRanges        * 1.56.1  2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggplot2              * 3.5.1   2024-04-23 [1] CRAN (R 4.4.0)
   glue                   1.7.0   2024-01-09 [1] CRAN (R 4.4.0)
   gridExtra            * 2.3     2017-09-09 [1] CRAN (R 4.4.0)
   gtable                 0.3.5   2024-04-22 [1] CRAN (R 4.4.0)
   gtools               * 3.9.5   2023-11-20 [1] CRAN (R 4.4.0)
   here                   1.0.1   2020-12-13 [1] CRAN (R 4.4.0)
   htmltools              0.5.8.1 2024-04-04 [1] CRAN (R 4.4.0)
   httr                   1.4.7   2023-08-15 [1] CRAN (R 4.4.0)
   IRanges              * 2.38.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   jsonlite               1.8.8   2023-12-04 [1] CRAN (R 4.4.0)
   knitr                  1.48    2024-07-07 [1] CRAN (R 4.4.1)
   lattice                0.22-6  2024-03-20 [1] CRAN (R 4.4.1)
   lifecycle              1.0.4   2023-11-07 [1] CRAN (R 4.4.0)
   magick                 2.8.3   2024-02-18 [1] CRAN (R 4.4.0)
   magrittr               2.0.3   2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                 1.7-0   2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics       * 1.16.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats          * 1.3.0   2024-04-11 [1] CRAN (R 4.4.0)
   munsell                0.5.1   2024-04-01 [1] CRAN (R 4.4.0)
   parallelly           * 1.37.1  2024-02-29 [1] CRAN (R 4.4.0)
   pillar                 1.9.0   2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig              2.0.3   2019-09-22 [1] CRAN (R 4.4.0)
   R6                     2.5.1   2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                   1.0.13  2024-07-17 [1] CRAN (R 4.4.1)
   rjson                  0.2.21  2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                * 1.1.4   2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown              2.27    2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot              2.0.4   2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi             0.16.0  2024-03-24 [1] CRAN (R 4.4.0)
   S4Arrays               1.4.1   2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors            * 0.42.1  2024-07-03 [1] Bioconductor 3.19 (R 4.4.1)
   scales                 1.3.0   2023-11-28 [1] CRAN (R 4.4.0)
   sessioninfo            1.2.2   2021-12-06 [1] CRAN (R 4.4.0)
   SingleCellExperiment * 1.26.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SparseArray            1.4.8   2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment    * 1.14.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SummarizedExperiment * 1.34.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   tibble                 3.2.1   2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect             1.2.1   2024-03-11 [1] CRAN (R 4.4.0)
   UCSC.utils             1.0.0   2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   utf8                   1.2.4   2023-10-22 [1] CRAN (R 4.4.0)
   vctrs                  0.6.5   2023-12-01 [1] CRAN (R 4.4.0)
   withr                  3.0.0   2024-01-16 [1] CRAN (R 4.4.0)
   xfun                   0.45    2024-06-16 [1] CRAN (R 4.4.0)
   XVector              * 0.44.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                   2.3.9   2024-07-05 [1] CRAN (R 4.4.0)
   zlibbioc               1.50.0  2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

─────────────────────────────────────────────────────────────────────────────────
