---
title: "04-DE of pseudobulk k2 VMH vs other"
author: "Bernard Mulvey"
date: "2023-06-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater)
library(scran)
library(edgeR)
library(jaffelab)
library(PCAtools) # findElbowPoint

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```


Load SPE with vmh already annotated
```{r}
hyp2.vmh <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_061323_HarmonyBSpc-q2VMH_rowname-key.RDS") 

# change/ensure colnames/colData's rownames == colData()$keys
colnames(hyp2.vmh) <- colData(hyp2.vmh)$key

## add demographics
demos <- fread("raw-data/demos.txt")
tmpcd <- as.data.table(colData(hyp2.vmh),keep.rownames=T)
tmpcd <- merge.data.table(tmpcd,demos,by.x="brnum",by.y="BrNum")
tmpcd <- as.data.frame(tmpcd)
rownames(tmpcd) <- tmpcd$rn
tmpcd <- tmpcd[colnames(hyp2.vmh),]

colData(hyp2.vmh) <- DataFrame(tmpcd) 
stopifnot(identical(colnames(hyp2.vmh),rownames(colData(hyp2.vmh))))
```


Going off https://github.com/LieberInstitute/spatial_DG_lifespan/blob/main/code/Pseudobulk/create_manual_annotated_pseudobulk_spe.R and https://github.com/LieberInstitute/spatial_DG_lifespan/blob/main/code/Pseudobulk/DE_pseudobulk.R
```{r}
hyp2_pbulk <- aggregateAcrossCells(hyp2.vmh,
                   DataFrame(is.vmh=hyp2.vmh$BSp_q2_consensus_vmh,
                             sample=hyp2.vmh$sample_id))

hyp2_pbulk$is.vmh <- factor(hyp2_pbulk$is.vmh)

# find a good expression cutoff using edgeR::filterByExpr
rowData(hyp2_pbulk)$high_expr <- filterByExpr(hyp2_pbulk)
rowData(hyp2_pbulk)$high_expr_cl <- filterByExpr(hyp2_pbulk, group = hyp2_pbulk$is.vmh)
rowData(hyp2_pbulk)$high_expr_sample <- filterByExpr(hyp2_pbulk, group = hyp2_pbulk$sample)

rowData(hyp2_pbulk)$high_expr_clorsam <- rowSums(cbind(rowData(hyp2_pbulk)$high_expr_sample,rowData(hyp2_pbulk)$high_expr_cl))

summary(rowData(hyp2_pbulk)$high_expr)
summary(rowData(hyp2_pbulk)$high_expr_sample)
summary(rowData(hyp2_pbulk)$high_expr_cl)
summary(rowData(hyp2_pbulk)$high_expr_clorsam>=1)

hyp2_pbulk2 <- hyp2_pbulk[rowData(hyp2_pbulk)$high_expr_clorsam>=1]
dim(hyp2_pbulk2)
```


### LEFT OFF HERE
```{r}
## Store the log normalized counts on the spe object
hyp2_pbulk3 <- edgeR::cpm(edgeR::calcNormFactors(hyp2_pbulk2), log = TRUE, prior.count = 0.1)

stopifnot(rownames(hyp2_pbulk3)==rownames(hyp2.vmh)[which(rownames(hyp2.vmh) %in% rownames(hyp2_pbulk3))])

#### fix dimnames, which right now got changed to "sample1, sample2...":
#### and aren't carried in by aggregate across cells above (colnames(hyp2_pbulk2) == null)

colnames(hyp2_pbulk3) <- paste0(hyp2_pbulk2$sample_id,"_",hyp2_pbulk2$is.vmh)

colnames(hyp2_pbulk2) <- paste0(hyp2_pbulk2$sample_id,"_",hyp2_pbulk2$is.vmh)

## and add the same as rownames of coldata for the full object hyp2_pbulk2
rownames(colData(hyp2_pbulk2)) <- colnames(hyp2_pbulk2)
## store norm counts on full spe
logcounts(hyp2_pbulk2) <- hyp2_pbulk3
```

### examine covariates vs PCs
```{r}
pca <- prcomp(t(logcounts(hyp2_pbulk2)))
pca_pbulk <- pca$x
reducedDims(hyp2_pbulk2) <- list(PCA=pca_pbulk)

pcavars <- jaffelab::getPcaVars(pca)
elbow <- PCAtools::findElbowPoint(pcavars)
# 6
# plot(pcavars)


```


### LEFT OFF AT LINe 104 of PSEUDOBULK SETUP SCRIPT @ https://github.com/LieberInstitute/spatial_DG_lifespan/blob/main/code/Pseudobulk/create_pseudobulk_spe.R
```{r}

### plot PCA x cohort vars
plotPCASCE(hyp2_pbulk2,color_by="Sex",ncomponents = 6)
plotPCASCE(hyp2_pbulk2,color_by="brnum",ncomponents = 6)
plotPCASCE(hyp2_pbulk2,color_by="AgeDeath",ncomponents = 6)
plotPCASCE(hyp2_pbulk2,color_by="PMI",ncomponents = 6)
plotPCASCE(hyp2_pbulk2,color_by="BMI..calculated.",ncomponents = 6)
plotPCASCE(hyp2_pbulk2,color_by="Best.RIN.PFC",ncomponents = 6)

varexpl <- getVarianceExplained(hyp2_pbulk2,c("Sex","brnum","AgeDeath","PMI","BMI..calculated.","sample_id"))

plotExplanatoryVariables(varexpl)
dev.off()
```


```{r}
colData(hyp2_pbulk2)$Sex <- factor(colData(hyp2_pbulk2)$Sex)
colData(hyp2_pbulk2)$is.vmh <- factor(colData(hyp2_pbulk2)$is.vmh)

# label $in_tissue --> all cells considered together, since we want to model by is.vmh (VMH vs other)
deres <- pseudoBulkDGE(
    hyp2_pbulk2,
    col.data = colData(hyp2_pbulk2),
    label = hyp2_pbulk2$in_tissue,
    design = ~is.vmh,
    coef="is.vmhVMH",
    row.data = rowData(hyp2_pbulk2),
    method = "voom",
    qualities = TRUE,
    robust = TRUE
)

deres.sexcov <- pseudoBulkDGE(
    hyp2_pbulk2,
    col.data = colData(hyp2_pbulk2),
    label = hyp2_pbulk2$in_tissue,
    design = ~is.vmh+Sex,
    coef="is.vmhVMH",
    row.data = rowData(hyp2_pbulk2),
    method = "voom",
    qualities = TRUE,
    robust = TRUE
)

deres.dt <- as.data.table(deres)
deres.sexcov <- as.data.table(deres.sexcov)

write.table(deres.dt,"analysis/data/H04-2cluster-consensus-VMH/pseudobulk-VMH-other-DE.txt",sep='\t',quote=F,row.names=F,col.names=T)
write.table(deres.sexcov,"analysis/data/H04-2cluster-consensus-VMH/pseudobulk-VMH-other-DE_sexcovaried.txt",sep='\t',quote=F,row.names=F,col.names=T)


```

#### 070823: alternate approach: crop the preprocessed, k=2 clustered sample V12-D075_B1 to the high-confidence VMH area, then proceed with DE. (clustering with k=2 with a pre-cropped tissue section and batch controlling w MNN does not isolate VMH, while the present k=2 approach w/ bayesspace had isolated cells negative for NR5A1 by ISH as VMH).

```{r}
hyp2.vmh <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_061323_HarmonyBSpc-q2VMH_rowname-key.RDS")

### crop V12D07-075_B1 first -- see code/00 blah blah .rmd
coordstmp <- as.data.frame(spatialCoords(hyp2.vmh))
rownames(coordstmp) <- colnames(hyp2.vmh)
coordstmp <- as.data.table(coordstmp,keep.rownames=T)
coordstmp <- coordstmp[,diag:=pxl_col_in_fullres+pxl_row_in_fullres]
coordstmp <- coordstmp[!(rn %in% grep(rn,pattern="V12D07-075_B1",value=T)&diag>30000)]

hyp2.vmh <- hyp2.vmh[,colnames(hyp2.vmh)%in%coordstmp$rn]
stopifnot(dim(hyp2.vmh)[2]==nrow(coordstmp))

### k now bulk
hyp2_pbulk <- aggregateAcrossCells(hyp2.vmh,
                   DataFrame(is.vmh=hyp2.vmh$BSp_q2_consensus_vmh,
                             sample=hyp2.vmh$sample_id))

hyp2_pbulk$is.vmh <- factor(hyp2_pbulk$is.vmh)

# find a good expression cutoff using edgeR::filterByExpr
rowData(hyp2_pbulk)$high_expr <- filterByExpr(hyp2_pbulk)
rowData(hyp2_pbulk)$high_expr_cl <- filterByExpr(hyp2_pbulk, group = hyp2_pbulk$is.vmh)
rowData(hyp2_pbulk)$high_expr_sample <- filterByExpr(hyp2_pbulk, group = hyp2_pbulk$sample)

rowData(hyp2_pbulk)$high_expr_clorsam <- rowSums(cbind(rowData(hyp2_pbulk)$high_expr_sample,rowData(hyp2_pbulk)$high_expr_cl))

hyp2_pbulk2 <- hyp2_pbulk[rowData(hyp2_pbulk)$high_expr_clorsam>=1]

## Store the log normalized counts on the spe object
hyp2_pbulk3 <- edgeR::cpm(edgeR::calcNormFactors(hyp2_pbulk2), log = TRUE, prior.count = 0.1)

stopifnot(rownames(hyp2_pbulk3)==rownames(hyp2.vmh)[which(rownames(hyp2.vmh) %in% rownames(hyp2_pbulk3))])

#### fix dimnames, which right now got changed to "sample1, sample2...":
#### and aren't carried in by aggregate across cells above (colnames(hyp2_pbulk2) == null)

colnames(hyp2_pbulk3) <- paste0(hyp2_pbulk2$sample_id,"_",hyp2_pbulk2$is.vmh)

colnames(hyp2_pbulk2) <- paste0(hyp2_pbulk2$sample_id,"_",hyp2_pbulk2$is.vmh)

## and add the same as rownames of coldata for the full object hyp2_pbulk2
rownames(colData(hyp2_pbulk2)) <- colnames(hyp2_pbulk2)
## store norm counts on full spe
logcounts(hyp2_pbulk2) <- hyp2_pbulk3

colData(hyp2_pbulk2)$is.vmh <- factor(colData(hyp2_pbulk2)$is.vmh)

# label $in_tissue --> all cells considered together, since we want to model by is.vmh (VMH vs other)
deres <- pseudoBulkDGE(
    hyp2_pbulk2,
    col.data = colData(hyp2_pbulk2),
    label = hyp2_pbulk2$in_tissue,
    design = ~is.vmh,
    coef="is.vmhVMH",
    row.data = rowData(hyp2_pbulk2),
    method = "voom",
    qualities = TRUE,
    robust = TRUE
)

deres.dt <- as.data.table(deres)
deres.dt[group_name==TRUE,group_name:="VMH"]

write.table(deres.dt,"analysis/data/spe_053123/H04-2cluster-consensus-VMH/pseudobulk_CROPPED-VMH-other-DE.txt",sep='\t',quote=F,row.names=F,col.names=T)

```
