---
title: "01-Plot putative VMH clusters and cluster markers from HVG10pctile-SNNk5"
author: "Bernie Mulvey"
date: "2023-06-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(scater)
library(SpatialExperiment)
library(Polychrome)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

hvg10, SNN k=5, cluster #5 looks promisingly like the VMH when plotting spots. Let's zoom in a little on this cluster and other possible VMH-accessory clusters.

```{r}
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")

walkclusts.hvg <- readRDS("analysis/data/H02-clustering/sampleBlock-hvg10-20_harmonyBrnumBlock-walktrap-SNNk5-10-15.RDS")

walktrap.markers <- readRDS("analysis/data/H03-marker_genes/sharmony_walktrapk5-15_markergenes.RDS")

colLabels(hyp2) <- factor(walkclusts.hvg[["hvg10_SNNk5"]])
# for plotting compatibility with marker output, change rownames to gene symbols
rownames(hyp2) <- rowData(hyp2)$gene_name
```

## Kim Yao Graybuck..Anderson Cell 19: "cells within the core are mostly glutamatergic and express Fezf1, Adcyap1, Gda, Nrgn, Cbln1, Lmo3, Nr5a1, C1ql2, and Rreb1[...]while those in the surrounding shell are mostly GABAergic "and express Gad1, Gad2, Tmem176a, Tmem176b, Six3, Dlk1, and Ecel"

```{r}
vmhcore <- toupper(c("Fezf1", "Adcyap1","Gda","Nrgn", "Cbln1", "Lmo3","Nr5a1","C1ql2"))
vmhshell <- toupper(c("Gad1", "Gad2", "Tmem176a", "Tmem176b", "Six3", "Dlk1"))
# ^ ECEL not in data
png("plots/H03-marker_genes/VMHcore-mousemarkers_hvg10_SNNk5.png",height=1500,width=2000)
plotExpression(hyp2,features=vmhcore,x="label",colour_by = "label")
dev.off()

png("plots/H03-marker_genes/VMHshell-mousemarkers_hvg10_SNNk5.png",height=1500,width=2000)
plotExpression(hyp2,features=vmhshell,x="label",colour_by = "label")
dev.off()

# plot spots only showing the two clusters of interest (5,14) vs all other

hvg10snn5.5.14.only <- walkclusts.hvg[["hvg10_SNNk5"]]
hvg10snn5.5.14.only[which(hvg10snn5.5.14.only!=5&hvg10snn5.5.14.only!=14)] <- 0 
colLabels(hyp2) <- factor(hvg10snn5.5.14.only)

png("plots/H03-marker_genes/hvg10snn5_clusters5and14only_plotspots.png")
plotSpots(hyp2,annotate="label")
dev.off()

rm(hvg10snn5.5.14.only)
```

Another gene set of interest: ESR1+ subtypes of VMH
•3 subpops of VMHvl marked by co-expression with tachikinin1 (Tac1), reprimo (Rprm), or prodynorphin (Pdyn) (cortes sturgeon forger horms and beh 2023)
•Additionally, Kim Yao..Graybuck '19 ID'ed sex-specific ESR1+ clusters based on coexpression of TSIX (female) or MOXD1 (male).
```{r}
png("plots/H03-marker_genes/VMH_ESR1-mouseSexSpecif-subcluster_markers_hvg10_SNNk5.png",height=750,width=800)
plotExpression(hyp2,features=c("ESR1","RPRM","PDYN","TSIX","MOXD1"),x="label",colour_by = "label")
dev.off()






```
