---
title: "02-plot Bayesspace q9-15-20-31 clusters from harmonyd hvg10-20 nnsvg10-20"
author: "Bernie Mulvey"
date: "2023-06-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
# library(scater) # addPerCellQC
# library(nnSVG)
library(BiocParallel)
library(scran)
library(parallel)
library(Polychrome)
library(spatialLIBD)
# library(fasthplus)

# theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load object with the harmonyPCA and harmonyUMAPs for these four feature sets for plotting purposes in those spaces. load bayesspace cluster assignments.
```{r}
pcas <- readRDS("analysis/data/H01-feature_selection/SPEs_sampleBlock-hvgs-10or20pctile_nnsvg10or20pctile_Harmony-PCA-UMAP.RDS")

### clone an spe out of this ^ for the spotplotting
hyp2.tmp <- pcas[[1]]

### set colnames(hyp2.tmp) to key for unique rownames etc; these are the rownames used with the bayesspace cluster results
colnames(hyp2.tmp) <- hyp2.tmp$key

bs_clusts <- list.files("analysis/data/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",pattern=".txt")
```

spotplots
```{r}
i <- 1
for (i in c(1:length(bs_clusts))) {
    curclusts <-
        fread(
            paste0(
                "analysis/data/H02-clustering/03-bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",
                bs_clusts[i]
            )
        )
    
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    
    if (sum(rownames(curclusts) == rownames(colData(hyp2.tmp))) != nrow(colData(hyp2.tmp))) {
        curclusts <- curclusts[rownames(colData(hyp2.tmp)),]
    }
    
    colLabels(hyp2.tmp) <- as.factor(curclusts[, 2])
    pal <- palette36.colors(n = length(unique(curclusts[, 2])))
    names(pal) <- unique(hyp2.tmp$label)
    
    plt <-
        spatialLIBD::vis_grid_clus(
            hyp2.tmp,
            clustervar = "label",
            pdf_file = NULL,
            colors = pal,
            spatial = F,
            auto_crop = T,
            return_plots = T,
            point_size = 2,
            sort_clust = F
        )
    
    ### adjust font sizes
    plt <- lapply(
        plt,
        FUN = function(x) {
            x +
                theme(
                    plot.title = element_text(size = 11, hjust = 0.5),
                    strip.text = element_text(size = 10),
                    legend.text = element_text(size = 10),
                    legend.title = element_text(size = 10, hjust = 0.5),
                    legend.spacing.y = unit(0, "cm"),
                    legend.spacing.x = unit(0.1, "cm")
                ) +
                guides(fill = guide_legend(
                    ncol = ceiling(length(unique(hyp2.tmp$label))/12),
                    byrow = T)
                )
        }
    )
    
    # save
    pdf(
        paste0(
            "plots/H02-clustering/02-bayesspace_q9-15-20-31_HVGorSVG_spotplots/",
            gsub(bs_clusts[i], pattern = "^(.*)\\.txt$", replacement = "\\1"),
            ".pdf"
        ),
        width = 20,
        height = 16
    )
    do.call("grid.arrange", plt)
    dev.off()
    rm(plt, curclusts)
}

rm(hyp2.tmp)
```

## for PCA and UMAP plotting below: set unadjusted data PCA, UMAP dimred names in the list of SPEs to uncor_PCA, uncor_UMAP so that we can trick plotDimRed into using the HARMONY results as in the plotting loop
```{r}
pcas <- lapply(
    pcas,
    FUN = function(x) {
        reducedDimNames(x)[which(reducedDimNames(x) == "PCA")] <-
            "uncor_PCA"
        reducedDimNames(x)[which(reducedDimNames(x) == "UMAP")] <-
            "uncor_UMAP"
        x
    }
)
```

Plot on feature-set-matched harmonyPCA

the BS_clusts file list conveniently is sorted in the same order as the PCA table (hvg10, hvg20, svg10, svg20)
```{r}
i <- 1
for (i in c(1:length(bs_clusts))) {
    k <- ceiling(i / 4)
    curclusts <-
        fread(
            paste0(
                "analysis/data/H02-clustering/bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",
                bs_clusts[i]
            )
        )
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    if (sum(rownames(curclusts) == rownames(colData(pcas[[k]]))) != nrow(colData(pcas[[k]]))) {
        curclusts <- curclusts[rownames(colData(pcas[[k]])), ]
    }
    
    colLabels(pcas[[k]]) <- factor(curclusts[, 2])
    reducedDimNames(pcas[[k]])[which(reducedDimNames(pcas[[k]]) == "HARMONY")] <-
        "PCA"
    colnames(pcas[[k]]@int_colData@listData$reducedDims@listData$PCA)[c(1,2)] <- c("PC1","PC2") 
    
    plt <-
        plotDimRed(
            pcas[[k]],
            type = "PCA",
            annotate = "label",
            palette = paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts[, 2])
            )))
        ) +
        ggtitle(paste0(
            "BSpace ",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_(.*)_.*\\.txt$", replacement =
                     "\\1"),
            " q=",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_.*_(.*)\\.txt$", replacement =
                     "\\1")
        ))
    pdf(
        paste0(
            "plots/H02-clustering/02-bayesspace_q9-15-20-31_HVGorSVG_harmonyPCA/",
            gsub(bs_clusts[i], pattern = "^(.*)\\.txt$", replacement = "\\1"),
            ".pdf"
        ),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    rm(plt, curclusts)
}

```


plot onto respective UMAPs
```{r}

i <- 1
for (i in c(1:length(bs_clusts))) {
    k <- ceiling(i / 4)
    
    curclusts <- fread(
        paste0("analysis/data/H02-clustering/bs_harmony_hvg1020_nnsvg1020_9-15-20-31_out/",
               bs_clusts[i]))
    
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    if (sum(rownames(curclusts) == rownames(colData(pcas[[k]]))) != nrow(colData(pcas[[k]]))) {
        curclusts <- curclusts[rownames(colData(pcas[[k]])), ]
    }
    
    colLabels(pcas[[k]]) <- factor(curclusts[, 2])
    reducedDimNames(pcas[[k]])[which(reducedDimNames(pcas[[k]]) == "HARMONY_UMAP")] <- "UMAP"
    colnames(pcas[[k]]@int_colData@listData$reducedDims@listData$UMAP) <- c("UMAP1","UMAP2") 
    
    plt <-
        plotDimRed(
            pcas[[k]],
            type = "UMAP",
            annotate = "label",
            palette = paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts[, 2]))))) +
        ggtitle(paste0("BSpace ",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_(.*)_.*\\.txt$", replacement =
                     "\\1"),
            " q=",
            gsub(bs_clusts[i], pattern = "^BSpace_Harmony_.*_(.*)\\.txt$", replacement =
                     "\\1")))
    
    pdf(paste0("plots/H02-clustering/02-bayesspace_q9-15-20-31_HVGorSVG_harmonyUMAP/",
            gsub(bs_clusts[i], pattern = "^(.*)\\.txt$", replacement = "\\1"),
            ".pdf"),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    
    rm(plt, curclusts)
}
```
