---
title: "03-plot PRECAST q9-15-20-31 clusters from harmonyd hvg10-20 nnsvg10-20"
author: "Bernie Mulvey"
date: "2023-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(ggplot2)
library(data.table)
library(Biostrings)
library(gridExtra)
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
library(SpatialExperiment)
library(ggspavis)
library(scater)
# library(nnSVG)
library(BiocParallel)
library(scran)
library(parallel)
library(Polychrome)
library(spatialLIBD)
# library(fasthplus)

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### load object with the PRECAST-generated dimension reductions for these each feature set-k combination for plotting purposes in those spaces. load precast cluster assignments.
```{r}
precast_dimreds <- readRDS("analysis/data/H02-clustering/04c-PRECAST_dimreductions_hvg1020_nnsvg1020_k9-15-20-31.RDS")

### clone an spe out of this ^ for the spotplotting
hyp2 <- readRDS("data/processing/hyp_umi600_gene450_chrm35_lognorm_053123.RDS")

# set colnames and rownames(coldata) to be coldata()$key for uniqueness
colnames(hyp2) <- colData(hyp2)$key

pcst_clusts <- readRDS("analysis/data/H02-clustering/04c-PRECAST_clusters_hvg1020_nnsvg1020_k9-15-20-31.RDS")
```

spotplots
```{r}
i <- 1
for (i in c(1:length(pcst_clusts))) {
    curclusts <- as.data.frame(pcst_clusts[[i]])
    rownames(curclusts) <- curclusts$rn
    
    if (sum(rownames(curclusts) == rownames(colData(hyp2.tmp))) != nrow(colData(hyp2.tmp))) {
        curclusts <- curclusts[rownames(colData(hyp2.tmp)), ]
    }
    
    colLabels(hyp2.tmp) <- as.factor(curclusts$label)
    pal <- palette36.colors(n=length(unique(curclusts[,2])))
    names(pal) <- levels(hyp2.tmp$label)
    
    plt <- spatialLIBD::vis_grid_clus(hyp2.tmp,clustervar = "label",pdf_file=NULL,colors = pal,spatial = F,auto_crop = T,return_plots = T,point_size = 2,sort_clust = F)
    
    ### adjust font sizes, pack legends tight (requires byrow arg in guides along with theme arg for spacing.y)
    plt <- lapply(
        plt,
        FUN = function(x) {
            x +
                theme(
                    plot.title = element_text(size = 11, hjust = 0.5),
                    strip.text = element_text(size = 10),
                    legend.text = element_text(size = 10),
                    legend.title = element_text(size = 10, hjust = 0.5),
                    legend.spacing.y = unit(0, "cm"),
                    legend.spacing.x = unit(0.1, "cm")
                ) +
                guides(fill = guide_legend(
                    ncol = ceiling(length(unique(hyp2.tmp$label))/12),
                    byrow = T)
                )
        }
    )
    #
    
    pdf(paste0(
            "plots/H02-clustering/03-PRECAST_q9-15-20-31_HVGorSVG_spotplots/", names(pcst_clusts)[i],".pdf"),
        width = 20,
        height = 16)
    do.call("grid.arrange",plt)
    dev.off()
    rm(plt, curclusts)
}

rm(hyp2.tmp)
```

## for PCA and UMAP plotting below: drop in PRECAST dimreduction to hyp2, generate UMAP therefrom per clustering run. plot colorized by cluster and sample id.

the pcst_clusts file list is sorted in the same order as the dimred table.
```{r}
i <- 1
for (i in c(1:length(pcst_clusts))) {
    curclusts <- pcst_clusts[[i]]
    curclusts <- as.data.frame(curclusts)
    rownames(curclusts) <- curclusts$rn
    if (sum(rownames(curclusts) == rownames(colData(hyp2))) != nrow(colData(hyp2))) {
        curclusts <- curclusts[rownames(colData(hyp2)), ]
    }
    
    curdimred <- precast_dimreds[[i]]
    curdimred <- as.data.frame(curdimred)
    rownames(curdimred) <- curdimred$rn
    if (sum(rownames(curdimred) == rownames(colData(hyp2))) != nrow(colData(hyp2))) {
        curdimred <- curdimred[rownames(colData(hyp2)), ]
    }
    curdimred$rn <- NULL
    colnames(curdimred)[c(1:2)] <- c("PC1","PC2")
    
    colLabels(hyp2) <- factor(curclusts$label)
    reducedDims(hyp2)$PCA <- curdimred
    
    # plot the PRECAST dimension reduction
    
    plt <-
        plotDimRed(
            hyp2,
            type = "PCA",
            annotate = "label",
            palette = paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts$label)
            )))
        ) +
        ggtitle(paste0("PRECAST",names(pcst_clusts)[i]))
    pdf(
        paste0("plots/H02-clustering/03-PRECAST_q9-15-20-31_HVGorSVG_PRECAST-PCA/",names(pcst_clusts)[i],".pdf"),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    rm(plt)
    
    # UMAP from temporarily placed PRECAST dimred
    set.seed(42)
    hyp2 <- runUMAP(hyp2,name = "UMAP",dimred="PCA")
    colnames(reducedDims(hyp2)$UMAP)[c(1,2)] <- c("UMAP1","UMAP2")
    
     plt <-
        plotDimRed(
            hyp2,
            type = "UMAP",
            annotate = "label",
            palette = paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts$label)
            )))
        ) +
        ggtitle(paste0("PRECAST",names(pcst_clusts)[i]))
    pdf(
        paste0("plots/H02-clustering/03-PRECAST_q9-15-20-31_HVGorSVG_PRECAST-UMAP/",names(pcst_clusts)[i],".pdf"),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()

    ## repeat the above two plots but by sample id
    # plot the PRECAST dimension reductions by sample id
    
    plt <-
        plotDimRed(
            hyp2,
            type = "PCA",
            annotate = "sample_id",
            size=0.6,
            palette = alpha(paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts$label)
            ))),0.4)
        ) +
        ggtitle(paste0("PRECAST",names(pcst_clusts)[i]))
    pdf(
        paste0("plots/H02-clustering/03-PRECAST_q9-15-20-31_HVGorSVG_dimreds_by_sample/",names(pcst_clusts)[i],"_bysample.pdf"),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    rm(plt)
    
     plt <-
        plotDimRed(
            hyp2,
            type = "UMAP",
            annotate = "sample_id",
            size=0.6,
            palette = alpha(paste0(Polychrome::palette36.colors(n = length(
                unique(curclusts$label)
            ))),0.4)
        ) +
        ggtitle(paste0("PRECAST",names(pcst_clusts)[i]))
    pdf(
        paste0("plots/H02-clustering/03-PRECAST_q9-15-20-31_HVGorSVG_dimreds_by_sample/",names(pcst_clusts)[i],"_bysample.pdf"),
        width = 10,
        height = 8
    )
    print(plt)
    dev.off()
    rm(plt,curclusts,curdimred)
}

```

