---
title: "02-ARC_subclustering_Banksy"
output: html_document
date: "2024-06-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Biostrings)
library(SpatialExperiment)
library(SpatialFeatureExperiment)
library(Banksy)

## rstudio GUI tweaks
require(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")
##

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc. seemed to need this a couple times, but otherwise havent so its here as a preventative measure.
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
```

## load SFE and annot clusters; subset to ARC clusters (15,19,23,31)
## importantly, we want the staggered-coordinates samples for subclustering so that samples across slides aren't misinterpreted as overlapping
```{r}
hypb <- readRDS("processed-data/05_Banksy_M0lam0_res2_multisamp/01-sfe-in_staggered-coords_genetarg-only_NONlog-norm.RDS")

m0l0 <- fread("processed-data/05_Banksy_M0lam0_res2_multisamp/01-BanksyClusts_M0lam0_leiden_multisamp.txt")

m0l0 <- DataFrame(m0l0,row.names=m0l0$rn)
m0l0 <- m0l0[colnames(hypb),]
colLabels(hypb) <- m0l0$M0lam0_leiden_multisamp_res2

hyparc <- hypb[,hypb$label %in% c("X15","X19","X23","X31")]

## clear workspace
rm(hypb,m0l0)
gc()
```

```{r}
## set arguments, which we'll try a variety of since the subsetted data is much smaller (71k cells)
lam <- c(0,0.2,0.5,0.8,1)
res <- seq(0.6,2.4,by=0.4)
gabor <- c(TRUE,FALSE)
# note that k geom of 6 is suggested in Banksy docs for VISIUM and was used for preceding cell typing.
# Banksy documentation uses 15 for non-AGF / 30 for AGF imaging-based spatial xscripto though, so this is what we'll use for granularity
kgeom <- c(15,30)
monocot <- 42
useassay <- "normcounts"
```


## Banksy
```{r}
hyparc <- computeBanksy(hyparc,
                      assay_name=useassay,
                      compute_agf = TRUE,
                      k_geom = kgeom,
                      seed=monocot)

hyparc <- runBanksyPCA(hyparc,
                         assay_name=useassay,
                         use_agf=gabor,
                         lambda=lam,
                         group="sample_id",
                         seed=monocot)

## since clustering runs on one CPU core per iteration and since the underlying dataset is rel. small, we can speed this up somewhat by running in parallel

cluspar <- expand.grid(lam=lam,res=res,gabor=gabor)
cluspar <- as.data.table(cluspar)
parm <- seq(1,nrow(cluspar),1)

arcclus <- bplapply(parm,BPPARAM=MulticoreParam(4),function(x){
  clust <- as.data.table(colData(
      clusterBanksy(hyparc,
                         assay_name=useassay,
                         use_agf=cluspar[x,gabor],
                         lambda=cluspar[x,lam],
                         resolution=cluspar[x,res],
                         group="sample_id",
                         seed=monocot)
  ),keep.rownames=T)
  keepnames <- c(paste0("clust_M",ifelse(cluspar[x,"gabor"],yes=1,no=0),"_lam",cluspar[x,lam],"_k50_res",cluspar[x,"res"]),"rn")
  clust <- clust[,..keepnames]
  stopifnot(ncol(clust)==2)
  return(clust)

})

hyparc <- clusterBanksy(hyparc,
                      assay_name=useassay,
                      use_agf = gabor,
                      lambda = lam,
                      resolution = res,
                      group="sample_id",
                      seed = monocot)


```

extract all of the clustering results into a table to be saved
```{r}
i<-1
for (i in c(1:length(arcclus))){
    if(i==1){arcclus2 <- copy(arcclus[[1]])}
    else{
        arcclus2 <- merge.data.table(arcclus2,arcclus[[i]],by="rn")
    }
}

fwrite(arcclus2,"processed-data/06_Banksy_VMH-ARC_subclustering/02_BanksyARCsubclusts_kg15-AGFk30_lam0-.2-.5-.8-1_res.6-1-1.4-1.8-2.2.txt")
```


reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```
R version 4.4.1 RC (2024-06-06 r86719)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.5

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods  
[8] base     

other attached packages:
 [1] magrittr_2.0.3                 BiocParallel_1.38.0           
 [3] parallelly_1.37.1              colorout_1.3-0.2              
 [5] Banksy_1.0.0                   SpatialFeatureExperiment_1.6.1
 [7] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
 [9] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[11] GenomicRanges_1.56.1           MatrixGenerics_1.16.0         
[13] matrixStats_1.3.0              Biostrings_2.72.1             
[15] GenomeInfoDb_1.40.1            XVector_0.44.0                
[17] IRanges_2.38.0                 S4Vectors_0.42.0              
[19] BiocGenerics_0.50.0            data.table_1.15.4             
[21] rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] DBI_1.2.3                 bitops_1.0-7             
  [3] deldir_2.0-4              s2_1.1.6                 
  [5] e1071_1.7-14              compiler_4.4.1           
  [7] sccore_1.0.5              DelayedMatrixStats_1.26.0
  [9] png_0.1-8                 sfheaders_0.4.4          
 [11] fftwtools_0.9-11          vctrs_0.6.5              
 [13] pkgconfig_2.0.3           wk_0.9.1                 
 [15] crayon_1.5.3              fastmap_1.2.0            
 [17] magick_2.8.3              scuttle_1.14.0           
 [19] utf8_1.2.4                rmarkdown_2.27           
 [21] UCSC.utils_1.0.0          xfun_0.45                
 [23] zlibbioc_1.50.0           beachmat_2.20.0          
 [25] jsonlite_1.8.8            rhdf5filters_1.16.0      
 [27] DelayedArray_0.30.1       Rhdf5lib_1.26.0          
 [29] jpeg_0.1-10               tiff_0.1-12              
 [31] terra_1.7-78              irlba_2.3.5.1            
 [33] parallel_4.4.1            aricode_1.0.3            
 [35] R6_2.5.1                  limma_3.60.3             
 [37] boot_1.3-30               leidenAlg_1.1.3          
 [39] Rcpp_1.0.12               knitr_1.47               
 [41] R.utils_2.12.3            tidyselect_1.2.1         
 [43] Matrix_1.7-0              igraph_2.0.3             
 [45] yaml_2.3.8                rstudioapi_0.16.0        
 [47] abind_1.4-5               EBImage_4.46.0           
 [49] codetools_0.2-20          lattice_0.22-6           
 [51] tibble_3.2.1              evaluate_0.24.0          
 [53] sf_1.0-16                 units_0.8-5              
 [55] spData_2.3.1              proxy_0.4-27             
 [57] mclust_6.1.1              pillar_1.9.0             
 [59] KernSmooth_2.23-24        generics_0.1.3           
 [61] dbscan_1.1-12             rprojroot_2.0.4          
 [63] sp_2.1-4                  RCurl_1.98-1.14          
 [65] ggplot2_3.5.1             sparseMatrixStats_1.16.0 
 [67] munsell_0.5.1             scales_1.3.0             
 [69] class_7.3-22              glue_1.7.0               
 [71] tools_4.4.1               BiocNeighbors_1.22.0     
 [73] locfit_1.5-9.9            rhdf5_2.48.0             
 [75] grid_4.4.1                spdep_1.3-5              
 [77] DropletUtils_1.24.0       edgeR_4.2.0              
 [79] colorspace_2.1-0          GenomeInfoDbData_1.2.12  
 [81] RcppHungarian_0.3         HDF5Array_1.32.0         
 [83] cli_3.6.2                 fansi_1.0.6              
 [85] S4Arrays_1.4.1            dplyr_1.1.4              
 [87] uwot_0.2.2                gtable_0.3.5             
 [89] R.methodsS3_1.8.2         zeallot_0.1.0            
 [91] digest_0.6.35             classInt_0.4-10          
 [93] SparseArray_1.4.8         dqrng_0.4.1              
 [95] rjson_0.2.21              htmlwidgets_1.6.4        
 [97] htmltools_0.5.8.1         R.oo_1.26.0              
 [99] lifecycle_1.0.4           httr_1.4.7               
[101] here_1.0.1                statmod_1.5.0            
> sessioninfo::session_info()
─ Session info ────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.1 RC (2024-06-06 r86719)
 os       macOS Sonoma 14.5
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-24
 rstudio  2024.07.0-daily+219 Cranberry Hibiscus (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ────────────────────────────────────────────────────────────────────
 ! package                  * version   date (UTC) lib source
   abind                      1.4-5     2016-07-21 [1] CRAN (R 4.4.0)
   aricode                    1.0.3     2023-10-20 [1] CRAN (R 4.4.0)
   Banksy                   * 1.0.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   beachmat                   2.20.0    2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   Biobase                  * 2.64.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocGenerics             * 0.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocNeighbors              1.22.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   BiocParallel             * 1.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.1)
   Biostrings               * 2.72.1    2024-06-02 [1] Bioconductor 3.19 (R 4.4.0)
   bitops                     1.0-7     2021-04-24 [1] CRAN (R 4.4.0)
   boot                       1.3-30    2024-02-26 [1] CRAN (R 4.4.1)
   class                      7.3-22    2023-05-03 [1] CRAN (R 4.4.1)
   classInt                   0.4-10    2023-09-05 [1] CRAN (R 4.4.0)
 P cli                        3.6.2     2023-12-11 [2] CRAN (R 4.4.0)
   codetools                  0.2-20    2024-03-31 [1] CRAN (R 4.4.1)
   colorout                 * 1.3-0.2   2024-05-01 [1] Github (jalvesaq/colorout@c6113a2)
   colorspace                 2.1-0     2023-01-23 [1] CRAN (R 4.4.0)
   crayon                     1.5.3     2024-06-20 [1] CRAN (R 4.4.0)
   data.table               * 1.15.4    2024-03-30 [2] CRAN (R 4.4.0)
   DBI                        1.2.3     2024-06-02 [1] CRAN (R 4.4.0)
   dbscan                     1.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
   DelayedMatrixStats         1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   deldir                     2.0-4     2024-02-28 [1] CRAN (R 4.4.0)
   digest                     0.6.35    2024-03-11 [1] CRAN (R 4.4.0)
   dplyr                      1.1.4     2023-11-17 [1] CRAN (R 4.4.0)
   dqrng                      0.4.1     2024-05-28 [1] CRAN (R 4.4.0)
   DropletUtils               1.24.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   e1071                      1.7-14    2023-12-06 [1] CRAN (R 4.4.0)
   EBImage                    4.46.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   edgeR                      4.2.0     2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   evaluate                   0.24.0    2024-06-10 [1] CRAN (R 4.4.0)
   fansi                      1.0.6     2023-12-08 [1] CRAN (R 4.4.0)
   fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
   fftwtools                  0.9-11    2021-03-01 [1] CRAN (R 4.4.0)
   generics                   0.1.3     2022-07-05 [1] CRAN (R 4.4.0)
   GenomeInfoDb             * 1.40.1    2024-05-24 [1] Bioconductor 3.19 (R 4.4.0)
   GenomeInfoDbData           1.2.12    2024-05-01 [1] Bioconductor
   GenomicRanges            * 1.56.1    2024-06-12 [1] Bioconductor 3.19 (R 4.4.1)
   ggplot2                    3.5.1     2024-04-23 [1] CRAN (R 4.4.0)
   glue                       1.7.0     2024-01-09 [1] CRAN (R 4.4.0)
   gtable                     0.3.5     2024-04-22 [1] CRAN (R 4.4.0)
   HDF5Array                  1.32.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   here                       1.0.1     2020-12-13 [1] CRAN (R 4.4.0)
   htmltools                  0.5.8.1   2024-04-04 [1] CRAN (R 4.4.0)
   htmlwidgets                1.6.4     2023-12-06 [1] CRAN (R 4.4.0)
   httr                       1.4.7     2023-08-15 [1] CRAN (R 4.4.0)
   igraph                     2.0.3     2024-03-13 [1] CRAN (R 4.4.0)
   IRanges                  * 2.38.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   irlba                      2.3.5.1   2022-10-03 [1] CRAN (R 4.4.0)
   jpeg                       0.1-10    2022-11-29 [1] CRAN (R 4.4.0)
   jsonlite                   1.8.8     2023-12-04 [1] CRAN (R 4.4.0)
   KernSmooth                 2.23-24   2024-05-17 [1] CRAN (R 4.4.1)
   knitr                      1.47      2024-05-29 [1] CRAN (R 4.4.0)
   lattice                    0.22-6    2024-03-20 [1] CRAN (R 4.4.1)
   leidenAlg                  1.1.3     2024-02-27 [1] CRAN (R 4.4.0)
   lifecycle                  1.0.4     2023-11-07 [1] CRAN (R 4.4.0)
   limma                      3.60.3    2024-06-16 [1] Bioconductor 3.19 (R 4.4.0)
   locfit                     1.5-9.9   2024-03-01 [1] CRAN (R 4.4.0)
   magick                     2.8.3     2024-02-18 [1] CRAN (R 4.4.0)
   magrittr                 * 2.0.3     2022-03-30 [1] CRAN (R 4.4.0)
   Matrix                     1.7-0     2024-04-26 [1] CRAN (R 4.4.1)
   MatrixGenerics           * 1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   matrixStats              * 1.3.0     2024-04-11 [1] CRAN (R 4.4.0)
   mclust                     6.1.1     2024-04-29 [1] CRAN (R 4.4.0)
   munsell                    0.5.1     2024-04-01 [1] CRAN (R 4.4.0)
   parallelly               * 1.37.1    2024-02-29 [1] CRAN (R 4.4.0)
   pillar                     1.9.0     2023-03-22 [1] CRAN (R 4.4.0)
   pkgconfig                  2.0.3     2019-09-22 [1] CRAN (R 4.4.0)
   png                        0.1-8     2022-11-29 [1] CRAN (R 4.4.0)
   proxy                      0.4-27    2022-06-09 [1] CRAN (R 4.4.0)
   R.methodsS3                1.8.2     2022-06-13 [1] CRAN (R 4.4.0)
   R.oo                       1.26.0    2024-01-24 [1] CRAN (R 4.4.0)
   R.utils                    2.12.3    2023-11-18 [1] CRAN (R 4.4.0)
   R6                         2.5.1     2021-08-19 [1] CRAN (R 4.4.0)
   Rcpp                       1.0.12    2024-01-09 [1] CRAN (R 4.4.0)
   RcppHungarian              0.3       2023-09-05 [1] CRAN (R 4.4.0)
   RCurl                      1.98-1.14 2024-01-09 [1] CRAN (R 4.4.0)
   rhdf5                      2.48.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rhdf5filters               1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   Rhdf5lib                   1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   rjson                      0.2.21    2022-01-09 [1] CRAN (R 4.4.0)
 P rlang                    * 1.1.4     2024-06-04 [2] CRAN (R 4.4.1)
   rmarkdown                  2.27      2024-05-17 [1] CRAN (R 4.4.0)
   rprojroot                  2.0.4     2023-11-05 [1] CRAN (R 4.4.0)
   rstudioapi                 0.16.0    2024-03-24 [1] CRAN (R 4.4.0)
   s2                         1.1.6     2023-12-19 [1] CRAN (R 4.4.0)
   S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
   S4Vectors                * 0.42.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   scales                     1.3.0     2023-11-28 [1] CRAN (R 4.4.0)
   sccore                     1.0.5     2024-02-23 [1] CRAN (R 4.4.0)
   scuttle                    1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sessioninfo                1.2.2     2021-12-06 [1] CRAN (R 4.4.0)
   sf                         1.0-16    2024-03-24 [1] CRAN (R 4.4.0)
   sfheaders                  0.4.4     2024-01-17 [1] CRAN (R 4.4.0)
   SingleCellExperiment     * 1.26.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   sp                         2.1-4     2024-04-30 [1] CRAN (R 4.4.0)
   SparseArray                1.4.8     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
   sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialExperiment        * 1.14.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   SpatialFeatureExperiment * 1.6.1     2024-05-15 [1] Bioconductor 3.19 (R 4.4.0)
   spData                     2.3.1     2024-05-31 [1] CRAN (R 4.4.0)
   spdep                      1.3-5     2024-06-10 [1] CRAN (R 4.4.0)
   statmod                    1.5.0     2023-01-06 [1] CRAN (R 4.4.0)
   SummarizedExperiment     * 1.34.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   terra                      1.7-78    2024-05-22 [1] CRAN (R 4.4.0)
   tibble                     3.2.1     2023-03-20 [1] CRAN (R 4.4.0)
   tidyselect                 1.2.1     2024-03-11 [1] CRAN (R 4.4.0)
   tiff                       0.1-12    2023-11-28 [1] CRAN (R 4.4.0)
   UCSC.utils                 1.0.0     2024-05-06 [1] Bioconductor 3.19 (R 4.4.0)
   units                      0.8-5     2023-11-28 [1] CRAN (R 4.4.0)
   utf8                       1.2.4     2023-10-22 [1] CRAN (R 4.4.0)
   uwot                       0.2.2     2024-04-21 [1] CRAN (R 4.4.0)
   vctrs                      0.6.5     2023-12-01 [1] CRAN (R 4.4.0)
   wk                         0.9.1     2023-11-29 [1] CRAN (R 4.4.0)
   xfun                       0.45      2024-06-16 [1] CRAN (R 4.4.0)
   XVector                  * 0.44.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
   yaml                       2.3.8     2023-12-11 [1] CRAN (R 4.4.0)
   zeallot                    0.1.0     2018-01-28 [1] CRAN (R 4.4.0)
   zlibbioc                   1.50.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)

 [1] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library
 [2] /Users/bmulvey/Library/R/arm64/4.4/library

 P ── Loaded and on-disk path mismatch.

───────────────────────────────────────────────────────────────────────────────
