---
title: "01-compare_per-sample-per-gene_logcounts_cellsegmentations_xenranger1.5vs1.7"
author: "Bernard Mulvey"
date: "2024-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 10,fig.width = 7,include = FALSE)
knitr::opts_chunk$set(fig.width=7,fig.height=10)
#### sets tab autocompletion for directories to begin from the directory containing the .Rproj session, instead of the script's directory if different
knitr::opts_knit$set(root.dir = here::here())

library(data.table)
library(Voyager)
library(patchwork)
library(SpatialFeatureExperiment)
library(SingleCellExperiment)
library(SpatialExperiment)
library(ggplot2)
library(gridExtra)
library(stringr)
library(scuttle)
library(BiocSingular)
library(scater)
library(Matrix)
library(DropletUtils)
library(vroom)
library(sf)
library(here)

library(colorout)
ColorOut()
options("styler.addins_style_transformer" = "biocthis::bioc_style()")

## change write.table to output a TSV without quotes or rownames by default
library(default)
default(write.table) <- list(row.names=F,col.names=T,sep='\t',quote=F)

## github copilot chat in the rstudio Viewer pane
library(chattr) # development branch @copilot installed 031324
# call chattr internal to disable debug mode so that this works
chattr:::ch_debug_set_false()
chattr_use("copilot")
chattr_defaults(model = "copilot",
                type = "chat",
                max_data_files = 20,
                max_data_frames = 20,
                include_doc_contents = TRUE,
                include_history = TRUE,
                prompt = "Use the 'R for Data Science' (https://r4ds.had.co.nz/) book as main reference
                        You can provide information about statistical methods and appropriateness of different statistical tests using material from https://compgenomr.github.io/book/, http://www.cookbook-r.com/, https://bookdown.org/rdpeng/advstatcomp/
                            Provide answers using data.table syntax as documented in https://cran.r-project.org/web/packages/data.table/vignettes/datatable-intro.html and  http://brooksandrew.github.io/simpleblog/articles/advanced-data-table/
                            Use packages: data.table, ggplot2, magrittr
                        Provide explanations with code
                        For code output, use RMarkdown code chunks
                        Avoid all code chunk options")
chattr::chattr_app("viewer",as_job=T,as_job_port=sample(c(4101:4199),size=1,replace=F))

## enable forked parallel processing with BiocParallel::multicoreParam, future::, etc on M processor macs
library(parallelly)
options(parallelly.supportsMulticore.disableOn="")
options(parallelly.fork.enable=TRUE)
library(BiocParallel)
options(bphost="localhost")
Sys.setenv("OBJC_DISABLE_INITIALIZE_FORK_SAFETY"="YES")

theme_set(theme_bw()+theme(axis.text.x = element_text(size = 14), axis.title.x = element_text(size = 16), axis.text.y = element_text(size = 14), axis.title.y = element_text(size =16), plot.title = element_text(size = 20,hjust=0.5), strip.text = element_text(size=18), legend.text = element_text(size=10), legend.title = element_text(size=11,hjust=0.5)))
```

### Define Xenium data importer to spatialfeatureexperiment format (courtesy Cindy Fang)
### edit this function to calculate log counts while in sce/spe formats, because trying to do so once in spatialfeatureexperiment won't work
```{r}
# Reference: https://github.com/pachterlab/SFEData/blob/main/inst/scripts/make-data.R

# Read in Xenium data and create an SFE object
# give this function a different name than readXenium, which is imported 
# from SpatialFeatureExperiment
readXenium_manual <- function(dir_name){

    counts_path <- here(dir_name, "cell_feature_matrix.h5")
    cell_info_path <- here(dir_name, "cells.csv.gz")
    cell_poly_path <- here(dir_name, "cell_boundaries.csv.gz")
    nuc_poly_path <- here(dir_name, "nucleus_boundaries.csv.gz")
    # 
    # Read in the data
    sce <- read10xCounts(counts_path)
    counts(sce) <- as(realize(counts(sce)), "dgCMatrix")
    cell_info <- vroom(cell_info_path)
    
    print("Read in files")
    
    cell_poly <- vroom(cell_poly_path)
    nuc_poly <- vroom(nuc_poly_path)
    
    names(cell_poly)[1] <- "ID"
    names(nuc_poly)[1] <- "ID"
    
    # change df to sf object for cell/nuc images
    cells_sf <- df2sf(cell_poly, c("vertex_x", "vertex_y"), geometryType = "POLYGON")
    nuc_sf <- df2sf(nuc_poly, c("vertex_x", "vertex_y"), geometryType = "POLYGON")
    
    # QC check
    all(st_is_valid(cells_sf))
    all(st_is_valid(nuc_sf))
    
    # get rid if invalid cells/nucs
    ind_invalid <- !st_is_valid(nuc_sf)
    nuc_sf[ind_invalid,] <- nngeo::st_remove_holes(st_buffer(nuc_sf[ind_invalid,], 0))
    
    # add cell info
    colData(sce) <- cbind(colData(sce), cell_info)
    print(cell_info)
    
    # make spatial objects
    spe <- toSpatialExperiment(sce, spatialCoordsNames = c("x_centroid", "y_centroid"))
    
    ###  log normalize the counts data, mostly just so that the plot doesn't look like its comprised of just a handful of points
    ### and to do that, we'll need to drop zero-count cells
    spe <- spe[,colSums(counts(spe))>0]
    spe <- scater::computeLibraryFactors(spe)
    spe <- scater::logNormCounts(spe)
    sfe <- toSpatialFeatureExperiment(spe)
    
    # add segmented cells/nuc to spatial object
    cellSeg(sfe, withDimnames = FALSE) <- cells_sf
    nucSeg(sfe, withDimnames = FALSE) <- nuc_sf
    
    
    # Add some QC Metrics
    colData(sfe)$nCounts <- colSums(counts(sfe))
    colData(sfe)$nGenes <- colSums(counts(sfe) > 0)
    
    is_blank <- str_detect(rownames(sfe), "^BLANK_")
    is_neg <- str_detect(rownames(sfe), "^NegControlProbe")
    is_neg2 <- str_detect(rownames(sfe), "^NegControlCodeword")
    is_anti <- str_detect(rownames(sfe), "^antisense")
    is_depr <- str_detect(rownames(sfe), "^DeprecatedCodeword")
    is_unassigned <- str_detect(rownames(sfe), "^Unassigned")
    
    is_any_neg <- is_blank | is_neg | is_neg2 | is_anti | is_depr | is_unassigned
    rowData(sfe)$is_neg <- is_any_neg
    
    n_panel <- nrow(sfe) - sum(is_any_neg)
    #print(n_panel)
    
    # normalize counts after QC
    colData(sfe)$nCounts_normed <- sfe$nCounts/n_panel
    colData(sfe)$nGenes_normed <- sfe$nGenes/n_panel
    colData(sfe)$prop_nuc <- sfe$nucleus_area / sfe$cell_area
    
    # add QC columns
    sfe <- addPerCellQCMetrics(sfe, subsets = list(blank = is_blank,
    negProbe = is_neg,
    negCodeword = is_neg2,
    anti = is_anti,
    depr = is_depr,
    unassigned = is_unassigned,
    any_neg = is_any_neg))

    # add features
    rowData(sfe)$means <- rowMeans(counts(sfe))
    rowData(sfe)$vars <- rowVars(counts(sfe))
    rowData(sfe)$cv2 <- rowData(sfe)$vars/rowData(sfe)$means^2

    
    # Add cell ids and make gene names unique
    #colnames(sfe) <- seq_len(ncol(sfe))
    rownames(sfe) <- uniquifyFeatureNames(ID=rownames(sfe),  names=rowData(sfe)$Symbol)
    
    return(sfe)
}
```

### load resegmentations and original segmentations as separate SpatialFeatureExperiment objects to start with
```{r}
resegs <- list.files("processed-data/01_xeniumranger-resegment1.7_20231019_slides",full.names = T)
resegsampname <- gsub(resegs,pattern="^.*slides/(.*)$",replacement="\\1")
resegs <- paste0(resegs,"/outs")

# outs directory files were previously placed directly inside the original samplewise directories
origsegs <- list.files("raw-data/xenium/20231019__161052__101923_KMon101923/",full.names=T)
origsegsampname <- gsub(origsegs,pattern="^.*output-(.*)$",replacement="\\1")
origsegsampname <- gsub(origsegsampname,pattern="^XETG.*__000[[:digit:]]{3}(..)__Region_(.).*$",replacement=paste0("\\1","_reg","\\2","_","origseg"))
i<-1
for (i in c(1:length(resegs))){
    if(i==1){
        resegsamps <- readXenium_manual(resegs[i])
        origsamps <- readXenium_manual(origsegs[i])
        
        resegsamps$Sample <- resegsampname[i]
        resegsamps$sample_id <- resegsampname[i]
        
        origsamps$Sample <- origsegsampname[i]
        origsamps$sample_id <- origsegsampname[i]
        # make unique colnames
        resegsamps$key <- paste0(resegsamps$Sample,"_",resegsamps$cell_id)
        colnames(resegsamps) <- resegsamps$key
        
        origsamps$key <- paste0(origsamps$Sample,"_",origsamps$cell_id)
        colnames(origsamps) <- origsamps$key
        
        # make unique colnames in rowdata
        colnames(rowData(origsamps))[c(5:7)] <- paste0(origsegsampname[i],"_",colnames(rowData(origsamps))[5:7]) 
        
        colnames(rowData(resegsamps))[c(5:7)] <- paste0(resegsampname[i],"_",colnames(rowData(resegsamps))[5:7]) 

    }
    else{
        tmpresegsamp <- readXenium_manual(resegs[i])
        tmpresegsamp$Sample <- resegsampname[i]
        tmpresegsamp$sample_id <- resegsampname[i]
        tmpresegsamp$key <- paste0(tmpresegsamp$Sample,"_",tmpresegsamp$cell_id)
        colnames(tmpresegsamp) <- tmpresegsamp$key
        
        ## remove the gene identifier/etc information here since we got it when i==1; make remaining colnames sample-specific 
        rowData(tmpresegsamp) <- rowData(tmpresegsamp)[,c(5:7)]
        colnames(rowData(tmpresegsamp)) <- paste0(resegsampname[i],"_",colnames(rowData(tmpresegsamp)))
        
        resegsamps <- cbind(resegsamps,tmpresegsamp)
        
        tmporigsamp <- readXenium_manual(origsegs[i])
        tmporigsamp$Sample <- origsegsampname[i]
        tmporigsamp$sample_id <- origsegsampname[i]
        tmporigsamp$key <- paste0(tmporigsamp$Sample,"_",tmporigsamp$cell_id)
        colnames(tmporigsamp) <- tmporigsamp$key
        
        ## remove the gene identifier/etc information here since we got it when i==1; make remaining colnames sample-specific 
        rowData(tmporigsamp) <- rowData(tmporigsamp)[,c(5:7)]
        colnames(rowData(tmporigsamp)) <- paste0(origsegsampname[i],"_",colnames(rowData(tmporigsamp)))
        
        origsamps <- cbind(origsamps,tmporigsamp)
        
        rm(tmpresegsamp,tmporigsamp)
    }
}
```


### getting the data paired back up for plotting will be very simple, because every nucleus is segmented to a cell and we did not alter the nuclear segmentations in the re-run, PLUS a nucleus and its cell have a shared unique identifier. so we can just match each cell identifier for each sample(1.5) to that of sample(1.7).

```{r}
## get logcounts matrices into d.t format for melting, merging
origcounts <- as.data.table(as.data.frame(as.matrix(logcounts(origsamps))),keep.rownames=T)

resegcounts <- as.data.table(as.data.frame(as.matrix(logcounts(resegsamps))),keep.rownames=T)

## melt so that each cell-gene-count value from a dataset is a row, i.e. so can have an x or y coord value for plotting
origcounts.m <- melt(origcounts,id.vars=c("rn"))
resegcounts.m <- melt(resegcounts,id.vars=c("rn"))

## make column names informative
setnames(origcounts.m,c("gene","cell","orig_value"))
setnames(resegcounts.m,c("gene","cell","reseg_value"))

## extract constant sample, cell identifiers for merging
origcounts.m[,sample:=gsub(cell,pattern="^(.*_reg.)_.*$",replacement="\\1")]
origcounts.m[,cell:=gsub(cell,pattern="^.*_reg._.*seg_(.*)$",replacement="\\1")]

resegcounts.m[,sample:=gsub(cell,pattern="^(.*_reg.)_.*$",replacement="\\1")]
resegcounts.m[,cell:=gsub(cell,pattern="^.*_reg._.*seg_(.*)$",replacement="\\1")]

## merge to one table for plotting
comparecounts <- merge.data.table(origcounts.m,resegcounts.m,by=c("sample","cell","gene"))

## easy to sanity check, since every nucleus is assigned a cell in both xeniumranger 1.5 and 1.7, and we didn't change nuclear segmentations here, so number of cells, genes, and samples (i.e. total rows) is consistent acorss original seg and reseg. doesn't appear there are different numbers of zero-count cells between the two segmentation runs (we filtered out zero-count in the readXenium_manual function); in rstudio the environment tab shows both tables as  the exact same number of rows (220,889,759)
stopifnot((nrow(comparecounts)==nrow(resegcounts.m)) & (nrow(comparecounts))==nrow(origcounts.m))

# spot check a row to make sure we got the correct table structure (e.g., cell aaaaaaa-1 counts of AABC1 in sample 86_reg1 are in a row)

## cleanup workspace/conserve memory
rm(origcounts,resegcounts,origcounts.m,resegcounts.m,i,origsegs,origsegsampname,resegs,resegsampname)
gc(full=T)
```

## finally, add some gene info (base panel vs. custom genes)
```{r}
genepanel <- fread("raw-data/humanbrain_basepanel_genes.txt")
comparecounts <- merge.data.table(comparecounts,genepanel,by.x="gene",by.y="Gene",all.x=T)
setnames(comparecounts,"Ensembl ID","ensembl")
comparecounts[!is.na(ensembl),gene_set:="Brain Base Panel"]
comparecounts[gene %in% grep(unique(comparecounts$gene),pattern="BLANK",value=T),gene_set:="BLANK"]
comparecounts[gene %in% grep(unique(comparecounts$gene),pattern="NegControl",value=T),gene_set:="Neg Ctrl"]
comparecounts[gene %in% grep(unique(comparecounts$gene),pattern="Deprecated",value=T),gene_set:="Deprecated Codeword"]
comparecounts[is.na(gene_set),gene_set:="HYP Custom Panel"]
```

### plot from here instead of saving this 220 million row table to load into a separate plotting script
### one pdf = one sample, one page = one gene
```{r}
samps <- as.list(unique(comparecounts$sample))

### divy up data before bplapply so each worker only loads relevant data
singsampdat <- lapply(samps,FUN=function(x){comparecounts[sample==x]})
names(singsampdat) <- unlist(samps)
rm(samps)

### Make a 5 color palette so points are colored by gene type
pals <- Polychrome::createPalette(5,seedcolors = "#43AE30")
pals <- as.data.table(cbind(unique(comparecounts$gene_set),pals))
setnames(pals,c("gene_set","pal"))
```

make a legend (and a useless paired plot) thatll give the color code for the rest of the plots
```{r}
## find a gene per gene_set to use for making a plot solely to extract a color legend from
# tmp <- unique(comparecounts[,.(gene,gene_set)])
leggenes <- c("ABCC9","BLANK_0054","CDH13","DeprecatedCodeword_0160","NegControlProbe_00041")
legdat <- comparecounts[sample=="86_reg1" & gene %in% leggenes]

tmppal <- pals$pal
names(tmppal) <- pals$gene_set
legplt <- ggplot(legdat,aes(x=orig_value,y=reseg_value,col=gene_set))+
  geom_point()+
  scale_color_manual(values=tmppal)

pdf("plots/02_compareCellSegmentations_1.5vs1.7/01a-colorlegend_for_01b.pdf",width=8.5,height=7.5)
legplt
dev.off()


# remove intermediate stuff from this block, other big memory stuff taking up RAM
# including the all-sample table, which is redundant with the per sample list of tables
rm(tmppal,leggenes,legdat,legplt,resegsamps,genepanel,origsamps,comparecounts)
gc(full=T)
```

### output plots in parallel
```{r}
bpp <- SnowParam(3)
register(bpp)
lapply(singsampdat,FUN=function(d){
    genes <- unique(d$gene)
    plts <- bplapply(genes,BPPARAM=bpp,BPOPTIONS = bpoptions(exportglobals =TRUE,packages=c("data.table","ggplot2","gridExtra"),exports = c("pals")),FUN=function(g){
        curcolor <- pals[gene_set==unique(d[gene==g,gene_set]),pal]
        ggplot(d[gene==g],aes(x=orig_value,y=reseg_value,col=curcolor))+
            geom_point(size = 0.5)+
            ggtitle(paste0(unique(d$sample),": ",g, " logcounts"))+
            xlim(0,max(c(d$orig_value,d$reseg_value)))+
            ylim(0,max(c(d$orig_value,d$reseg_value)))+
            geom_abline(intercept = 0,slope = 1,line.type="dashed",col="black")
    })
    ## add legend to front of list so it's on the first page
    plotsout <- marrangeGrob(plts,nrow=1,ncol=1)
    ggsave(path="plots/02_compareCellSegmentations_1.5vs1.7",filename = paste0("01b-",unique(d$sample),".pdf"),plot = plotsout,width = 8.5,height=7.5)
})

## stop parallel workers
register(SerialParam())

gc(full=T)
```

Reproducibility info
```{r}
sessionInfo()
sessioninfo::session_info()
```

R version 4.4.0 Patched (2024-04-30 r86503)
Platform: aarch64-apple-darwin20
Running under: macOS Sonoma 14.4.1

Matrix products: default
BLAS:   /System/Library/Frameworks/Accelerate.framework/Versions/A/Frameworks/vecLib.framework/Versions/A/libBLAS.dylib 
LAPACK: /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/lib/libRlapack.dylib;  LAPACK version 3.12.0

locale:
[1] en_US.UTF-8/en_US.UTF-8/en_US.UTF-8/C/en_US.UTF-8/en_US.UTF-8

time zone: America/Chicago
tzcode source: internal

attached base packages:
[1] stats4    stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] BiocParallel_1.38.0            parallelly_1.37.1             
 [3] chattr_0.1.0                   default_1.0.0                 
 [5] colorout_1.3-0.2               here_1.0.1                    
 [7] sf_1.0-16                      vroom_1.6.5                   
 [9] DropletUtils_1.24.0            Matrix_1.7-0                  
[11] scater_1.32.0                  BiocSingular_1.20.0           
[13] scuttle_1.14.0                 stringr_1.5.1                 
[15] gridExtra_2.3                  ggplot2_3.5.1                 
[17] SpatialExperiment_1.14.0       SingleCellExperiment_1.26.0   
[19] SummarizedExperiment_1.34.0    Biobase_2.64.0                
[21] GenomicRanges_1.56.0           GenomeInfoDb_1.40.1           
[23] IRanges_2.38.0                 S4Vectors_0.42.0              
[25] BiocGenerics_0.50.0            MatrixGenerics_1.16.0         
[27] matrixStats_1.3.0              patchwork_1.2.0               
[29] Voyager_1.6.0                  SpatialFeatureExperiment_1.6.1
[31] data.table_1.15.99             rlang_1.1.4                   

loaded via a namespace (and not attached):
  [1] later_1.3.2               bitops_1.0-7              tibble_3.2.1             
  [4] R.oo_1.26.0               lifecycle_1.0.4           httr2_1.0.1              
  [7] edgeR_4.2.0               rprojroot_2.0.4           processx_3.8.4           
 [10] lattice_0.22-6            magrittr_2.0.3            limma_3.60.2             
 [13] sass_0.4.9                rmarkdown_2.27            jquerylib_0.1.4          
 [16] yaml_2.3.8                httpuv_1.6.15             sp_2.1-4                 
 [19] sessioninfo_1.2.2         DBI_1.2.3                 abind_1.4-5              
 [22] zlibbioc_1.50.0           purrr_1.0.2               R.utils_2.12.3           
 [25] nngeo_0.4.8               RCurl_1.98-1.14           rappdirs_0.3.3           
 [28] GenomeInfoDbData_1.2.12   ggrepel_0.9.5             irlba_2.3.5.1            
 [31] terra_1.7-78              units_0.8-5               RSpectra_0.16-1          
 [34] dqrng_0.4.1               DelayedMatrixStats_1.26.0 codetools_0.2-20         
 [37] DelayedArray_0.30.1       tidyselect_1.2.1          farver_2.1.2             
 [40] UCSC.utils_1.0.0          memuse_4.2-3              ScaledMatrix_1.12.0      
 [43] viridis_0.6.5             jsonlite_1.8.8            BiocNeighbors_1.22.0     
 [46] e1071_1.7-14              systemfonts_1.1.0         tools_4.4.0              
 [49] ggnewscale_0.4.10         ragg_1.3.2                snow_0.4-4               
 [52] Rcpp_1.0.12               glue_1.7.0                SparseArray_1.4.8        
 [55] xfun_0.44                 EBImage_4.46.0            dplyr_1.1.4              
 [58] HDF5Array_1.32.0          withr_3.0.0               fastmap_1.2.0            
 [61] boot_1.3-30               rhdf5filters_1.16.0       bluster_1.14.0           
 [64] fansi_1.0.6               spData_2.3.1              callr_3.7.6              
 [67] digest_0.6.35             rsvd_1.0.5                R6_2.5.1                 
 [70] mime_0.12                 textshaping_0.4.0         colorspace_2.1-0         
 [73] wk_0.9.1                  jpeg_0.1-10               R.methodsS3_1.8.2        
 [76] utf8_1.2.4                generics_0.1.3            class_7.3-22             
 [79] httr_1.4.7                htmlwidgets_1.6.4         S4Arrays_1.4.1           
 [82] scatterplot3d_0.3-44      spdep_1.3-4               pkgconfig_2.0.3          
 [85] scico_1.5.0               gtable_0.3.5              XVector_0.44.0           
 [88] htmltools_0.5.8.1         fftwtools_0.9-11          scales_1.3.0             
 [91] png_0.1-8                 knitr_1.47                rstudioapi_0.16.0        
 [94] tzdb_0.4.0                rjson_0.2.21              proxy_0.4-27             
 [97] cachem_1.1.0              rhdf5_2.48.0              Polychrome_1.5.1         
[100] KernSmooth_2.23-24        parallel_4.4.0            vipor_0.4.7              
[103] s2_1.1.6                  pillar_1.9.0              grid_4.4.0               
[106] vctrs_0.6.5               promises_1.3.0            beachmat_2.20.0          
[109] xtable_1.8-4              sfheaders_0.4.4           cluster_2.1.6            
[112] beeswarm_0.4.0            evaluate_0.23             zeallot_0.1.0            
[115] magick_2.8.3              cli_3.6.2                 locfit_1.5-9.9           
[118] compiler_4.4.0            crayon_1.5.2              labeling_0.4.3           
[121] classInt_0.4-10           ps_1.7.6                  fs_1.6.4                 
[124] ggbeeswarm_0.7.2          stringi_1.8.4             viridisLite_0.4.2        
[127] deldir_2.0-4              munsell_0.5.1             tiff_0.1-12              
[130] sparseMatrixStats_1.16.0  bit64_4.0.5               Rhdf5lib_1.26.0          
[133] statmod_1.5.0             shiny_1.8.1.1             clipr_0.8.0              
[136] igraph_2.0.3              bslib_0.7.0               bit_4.0.5                
> sessioninfo::session_info()
─ Session info ──────────────────────────────────────────────────────────────────────
 setting  value
 version  R version 4.4.0 Patched (2024-04-30 r86503)
 os       macOS Sonoma 14.4.1
 system   aarch64, darwin20
 ui       RStudio
 language (EN)
 collate  en_US.UTF-8
 ctype    en_US.UTF-8
 tz       America/Chicago
 date     2024-06-10
 rstudio  2024.04.1+748 Chocolate Cosmos (desktop)
 pandoc   3.1.11 @ /Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools/aarch64/ (via rmarkdown)

─ Packages ──────────────────────────────────────────────────────────────────────────
 package                  * version   date (UTC) lib source
 abind                      1.4-5     2016-07-21 [2] CRAN (R 4.4.0)
 beachmat                   2.20.0    2024-05-06 [2] Bioconductor 3.19 (R 4.4.0)
 beeswarm                   0.4.0     2021-06-01 [2] CRAN (R 4.4.0)
 Biobase                  * 2.64.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocGenerics             * 0.50.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocNeighbors              1.22.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocParallel             * 1.38.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 BiocSingular             * 1.20.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 bit                        4.0.5     2022-11-15 [2] CRAN (R 4.4.0)
 bit64                      4.0.5     2020-08-30 [2] CRAN (R 4.4.0)
 bitops                     1.0-7     2021-04-24 [2] CRAN (R 4.4.0)
 bluster                    1.14.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 boot                       1.3-30    2024-02-26 [2] CRAN (R 4.4.0)
 bslib                      0.7.0     2024-03-29 [2] CRAN (R 4.4.0)
 cachem                     1.1.0     2024-05-16 [1] CRAN (R 4.4.0)
 callr                      3.7.6     2024-03-25 [2] CRAN (R 4.4.0)
 chattr                   * 0.1.0     2024-04-27 [2] CRAN (R 4.4.0)
 class                      7.3-22    2023-05-03 [2] CRAN (R 4.4.0)
 classInt                   0.4-10    2023-09-05 [2] CRAN (R 4.4.0)
 cli                        3.6.2     2023-12-11 [2] CRAN (R 4.4.0)
 clipr                      0.8.0     2022-02-22 [2] CRAN (R 4.4.0)
 cluster                    2.1.6     2023-12-01 [2] CRAN (R 4.4.0)
 codetools                  0.2-20    2024-03-31 [2] CRAN (R 4.4.0)
 colorout                 * 1.3-0.2   2024-05-01 [2] Github (jalvesaq/colorout@c6113a2)
 colorspace                 2.1-0     2023-01-23 [2] CRAN (R 4.4.0)
 crayon                     1.5.2     2022-09-29 [2] CRAN (R 4.4.0)
 data.table               * 1.15.99   2024-05-01 [2] local
 DBI                        1.2.3     2024-06-02 [2] CRAN (R 4.4.0)
 default                  * 1.0.0     2017-08-07 [2] CRAN (R 4.4.0)
 DelayedArray               0.30.1    2024-05-07 [1] Bioconductor 3.19 (R 4.4.0)
 DelayedMatrixStats         1.26.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 deldir                     2.0-4     2024-02-28 [2] CRAN (R 4.4.0)
 digest                     0.6.35    2024-03-11 [2] CRAN (R 4.4.0)
 dplyr                      1.1.4     2023-11-17 [2] CRAN (R 4.4.0)
 dqrng                      0.4.1     2024-05-28 [2] CRAN (R 4.4.0)
 DropletUtils             * 1.24.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 e1071                      1.7-14    2023-12-06 [2] CRAN (R 4.4.0)
 EBImage                    4.46.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 edgeR                      4.2.0     2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 evaluate                   0.23      2023-11-01 [2] CRAN (R 4.4.0)
 fansi                      1.0.6     2023-12-08 [2] CRAN (R 4.4.0)
 farver                     2.1.2     2024-05-13 [2] CRAN (R 4.4.0)
 fastmap                    1.2.0     2024-05-15 [1] CRAN (R 4.4.0)
 fftwtools                  0.9-11    2021-03-01 [2] CRAN (R 4.4.0)
 fs                         1.6.4     2024-04-25 [2] CRAN (R 4.4.0)
 generics                   0.1.3     2022-07-05 [2] CRAN (R 4.4.0)
 GenomeInfoDb             * 1.40.1    2024-05-24 [2] Bioconductor 3.19 (R 4.4.0)
 GenomeInfoDbData           1.2.12    2024-05-01 [2] Bioconductor
 GenomicRanges            * 1.56.0    2024-05-06 [2] Bioconductor 3.19 (R 4.4.0)
 ggbeeswarm                 0.7.2     2023-04-29 [2] CRAN (R 4.4.0)
 ggnewscale                 0.4.10    2024-02-08 [2] CRAN (R 4.4.0)
 ggplot2                  * 3.5.1     2024-04-23 [2] CRAN (R 4.4.0)
 ggrepel                    0.9.5     2024-01-10 [2] CRAN (R 4.4.0)
 glue                       1.7.0     2024-01-09 [2] CRAN (R 4.4.0)
 gridExtra                * 2.3       2017-09-09 [2] CRAN (R 4.4.0)
 gtable                     0.3.5     2024-04-22 [2] CRAN (R 4.4.0)
 HDF5Array                  1.32.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 here                     * 1.0.1     2020-12-13 [2] CRAN (R 4.4.0)
 htmltools                  0.5.8.1   2024-04-04 [2] CRAN (R 4.4.0)
 htmlwidgets                1.6.4     2023-12-06 [2] CRAN (R 4.4.0)
 httpuv                     1.6.15    2024-03-26 [2] CRAN (R 4.4.0)
 httr                       1.4.7     2023-08-15 [2] CRAN (R 4.4.0)
 httr2                      1.0.1     2024-04-01 [2] CRAN (R 4.4.0)
 igraph                     2.0.3     2024-03-13 [2] CRAN (R 4.4.0)
 IRanges                  * 2.38.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 irlba                      2.3.5.1   2022-10-03 [2] CRAN (R 4.4.0)
 jpeg                       0.1-10    2022-11-29 [2] CRAN (R 4.4.0)
 jquerylib                  0.1.4     2021-04-26 [2] CRAN (R 4.4.0)
 jsonlite                   1.8.8     2023-12-04 [2] CRAN (R 4.4.0)
 KernSmooth                 2.23-24   2024-05-17 [2] CRAN (R 4.4.0)
 knitr                      1.47      2024-05-29 [2] CRAN (R 4.4.0)
 labeling                   0.4.3     2023-08-29 [2] CRAN (R 4.4.0)
 later                      1.3.2     2023-12-06 [2] CRAN (R 4.4.0)
 lattice                    0.22-6    2024-03-20 [2] CRAN (R 4.4.0)
 lifecycle                  1.0.4     2023-11-07 [2] CRAN (R 4.4.0)
 limma                      3.60.2    2024-05-19 [1] Bioconductor 3.19 (R 4.4.0)
 locfit                     1.5-9.9   2024-03-01 [2] CRAN (R 4.4.0)
 magick                     2.8.3     2024-02-18 [2] CRAN (R 4.4.0)
 magrittr                   2.0.3     2022-03-30 [2] CRAN (R 4.4.0)
 Matrix                   * 1.7-0     2024-03-22 [2] CRAN (R 4.4.0)
 MatrixGenerics           * 1.16.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 matrixStats              * 1.3.0     2024-04-11 [2] CRAN (R 4.4.0)
 memuse                     4.2-3     2023-01-24 [2] CRAN (R 4.4.0)
 mime                       0.12      2021-09-28 [2] CRAN (R 4.4.0)
 munsell                    0.5.1     2024-04-01 [2] CRAN (R 4.4.0)
 nngeo                      0.4.8     2024-04-17 [2] CRAN (R 4.4.0)
 parallelly               * 1.37.1    2024-02-29 [2] CRAN (R 4.4.0)
 patchwork                * 1.2.0     2024-01-08 [2] CRAN (R 4.4.0)
 pillar                     1.9.0     2023-03-22 [2] CRAN (R 4.4.0)
 pkgconfig                  2.0.3     2019-09-22 [2] CRAN (R 4.4.0)
 png                        0.1-8     2022-11-29 [2] CRAN (R 4.4.0)
 Polychrome                 1.5.1     2022-05-03 [2] CRAN (R 4.4.0)
 processx                   3.8.4     2024-03-16 [2] CRAN (R 4.4.0)
 promises                   1.3.0     2024-04-05 [2] CRAN (R 4.4.0)
 proxy                      0.4-27    2022-06-09 [2] CRAN (R 4.4.0)
 ps                         1.7.6     2024-01-18 [2] CRAN (R 4.4.0)
 purrr                      1.0.2     2023-08-10 [2] CRAN (R 4.4.0)
 R.methodsS3                1.8.2     2022-06-13 [2] CRAN (R 4.4.0)
 R.oo                       1.26.0    2024-01-24 [2] CRAN (R 4.4.0)
 R.utils                    2.12.3    2023-11-18 [2] CRAN (R 4.4.0)
 R6                         2.5.1     2021-08-19 [2] CRAN (R 4.4.0)
 ragg                       1.3.2     2024-05-15 [2] CRAN (R 4.4.0)
 rappdirs                   0.3.3     2021-01-31 [2] CRAN (R 4.4.0)
 Rcpp                       1.0.12    2024-01-09 [2] CRAN (R 4.4.0)
 RCurl                      1.98-1.14 2024-01-09 [2] CRAN (R 4.4.0)
 rhdf5                      2.48.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 rhdf5filters               1.16.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 Rhdf5lib                   1.26.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 rjson                      0.2.21    2022-01-09 [2] CRAN (R 4.4.0)
 rlang                    * 1.1.4     2024-06-04 [1] CRAN (R 4.4.0)
 rmarkdown                  2.27      2024-05-17 [2] CRAN (R 4.4.0)
 rprojroot                  2.0.4     2023-11-05 [2] CRAN (R 4.4.0)
 RSpectra                   0.16-1    2022-04-24 [2] CRAN (R 4.4.0)
 rstudioapi                 0.16.0    2024-03-24 [2] CRAN (R 4.4.0)
 rsvd                       1.0.5     2021-04-16 [2] CRAN (R 4.4.0)
 s2                         1.1.6     2023-12-19 [2] CRAN (R 4.4.0)
 S4Arrays                   1.4.1     2024-05-20 [1] Bioconductor 3.19 (R 4.4.0)
 S4Vectors                * 0.42.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 sass                       0.4.9     2024-03-15 [2] CRAN (R 4.4.0)
 ScaledMatrix               1.12.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 scales                     1.3.0     2023-11-28 [2] CRAN (R 4.4.0)
 scater                   * 1.32.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 scatterplot3d              0.3-44    2023-05-05 [2] CRAN (R 4.4.0)
 scico                      1.5.0     2023-08-14 [2] CRAN (R 4.4.0)
 scuttle                  * 1.14.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 sessioninfo                1.2.2     2021-12-06 [2] CRAN (R 4.4.0)
 sf                       * 1.0-16    2024-03-24 [2] CRAN (R 4.4.0)
 sfheaders                  0.4.4     2024-01-17 [2] CRAN (R 4.4.0)
 shiny                      1.8.1.1   2024-04-02 [2] CRAN (R 4.4.0)
 SingleCellExperiment     * 1.26.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 snow                       0.4-4     2021-10-27 [2] CRAN (R 4.4.0)
 sp                         2.1-4     2024-04-30 [2] CRAN (R 4.4.0)
 SparseArray                1.4.8     2024-05-30 [2] Bioconductor 3.19 (R 4.4.0)
 sparseMatrixStats          1.16.0    2024-04-30 [1] Bioconductor 3.19 (R 4.4.0)
 SpatialExperiment        * 1.14.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 SpatialFeatureExperiment * 1.6.1     2024-05-15 [2] Bioconductor 3.19 (R 4.4.0)
 spData                     2.3.1     2024-05-31 [2] CRAN (R 4.4.0)
 spdep                      1.3-4     2024-05-31 [2] CRAN (R 4.4.0)
 statmod                    1.5.0     2023-01-06 [2] CRAN (R 4.4.0)
 stringi                    1.8.4     2024-05-06 [2] CRAN (R 4.4.0)
 stringr                  * 1.5.1     2023-11-14 [2] CRAN (R 4.4.0)
 SummarizedExperiment     * 1.34.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 systemfonts                1.1.0     2024-05-15 [2] CRAN (R 4.4.0)
 terra                      1.7-78    2024-05-22 [2] CRAN (R 4.4.0)
 textshaping                0.4.0     2024-05-24 [2] CRAN (R 4.4.0)
 tibble                     3.2.1     2023-03-20 [2] CRAN (R 4.4.0)
 tidyselect                 1.2.1     2024-03-11 [2] CRAN (R 4.4.0)
 tiff                       0.1-12    2023-11-28 [2] CRAN (R 4.4.0)
 tzdb                       0.4.0     2023-05-12 [2] CRAN (R 4.4.0)
 UCSC.utils                 1.0.0     2024-05-06 [2] Bioconductor 3.19 (R 4.4.0)
 units                      0.8-5     2023-11-28 [2] CRAN (R 4.4.0)
 utf8                       1.2.4     2023-10-22 [2] CRAN (R 4.4.0)
 vctrs                      0.6.5     2023-12-01 [2] CRAN (R 4.4.0)
 vipor                      0.4.7     2023-12-18 [2] CRAN (R 4.4.0)
 viridis                    0.6.5     2024-01-29 [2] CRAN (R 4.4.0)
 viridisLite                0.4.2     2023-05-02 [2] CRAN (R 4.4.0)
 Voyager                  * 1.6.0     2024-05-30 [1] Bioconductor 3.19 (R 4.4.0)
 vroom                    * 1.6.5     2023-12-05 [2] CRAN (R 4.4.0)
 withr                      3.0.0     2024-01-16 [2] CRAN (R 4.4.0)
 wk                         0.9.1     2023-11-29 [2] CRAN (R 4.4.0)
 xfun                       0.44      2024-05-15 [2] CRAN (R 4.4.0)
 xtable                     1.8-4     2019-04-21 [2] CRAN (R 4.4.0)
 XVector                    0.44.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)
 yaml                       2.3.8     2023-12-11 [2] CRAN (R 4.4.0)
 zeallot                    0.1.0     2018-01-28 [2] CRAN (R 4.4.0)
 zlibbioc                   1.50.0    2024-04-30 [2] Bioconductor 3.19 (R 4.4.0)

 [1] /Users/bmulvey/Library/R/arm64/4.4/library
 [2] /Library/Frameworks/R.framework/Versions/4.4-arm64/Resources/library

─────────────────────────────────────────────────────────────────────────────────────